### 2010-05-24 Documents/Manuals/flow-details.tex
initial commit

\documentclass[12pt]{article}
\usepackage{geometry}
\geometry{verbose,a4paper,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
\usepackage[pdftex]{graphicx}
%\usepackage{fancyvrb}
\usepackage{epstopdf}
%\usepackage{subfigure}
%\usepackage{alltt}
%\usepackage{amssymb}

%\newbox\subfigbox
%\makeatletter
%\newenvironment{subfloat}
%{\def\caption##1{\gdef\subcapsave{\relax##1}}%
%\let\subcapsave\@empty
%\setbox\subfigbox\hbox
%\bgroup}
%{\egroup
%\subfigure[\subcapsave]{\box\subfigbox}}
%\makeatother

\title{Compiler flow for Program to Hardware using AHIR}
\author{Sameer D. Sahasrabuddhe}

\begin{document}

\maketitle

\section{Overview}

The AHIR compiler flow involves a series of passes that result in the
generation of an AHIR specification from a program initially written
in C. The flow is divided into three main stages:

\begin{enumerate}
\item software compiler (LLVM)
\item AHIR generator
\item AHIR linker
\end{enumerate}

The implementation uses a number of freely available C++ software
libraries:

\begin{itemize}
\item LLVM Compiler Framework
\item Boost C++ libraries
\begin{itemize}
\item {\tt program\_options}
\item {\tt string\_algo}
\item {\tt lexical\_cast}
\end{itemize}
\item {\tt libxml++}, the C++ wrapper for the {\tt libxml} XML parser
\end{itemize}

\section{Calling conventions for C functions in AHIR}
\label{sect:calling-conventions}

AHIR, by definition, is independent of the input programming language. 
A specification in AHIR describes a system only in terms of modules
each made of a control and a data path, interacting through a link
layer. In addition, the modules interact with each other through a
separate link layer for control transfer, and pre-assigned memory
locations for data transfer.

A C program is a call-graph, where the vertices are functions and edges
represent function calls. Functions pass data through a function call
in the form of arguments and return values. When translating such a
call-graph to an AHIR specification, each function is represented by a
module, with its own control and data path. The function calls are
translated to control and data interaction between these modules. For
this inter-module communication, we need a mechanism that provides two
services:

\begin{itemize}
\item passing arguments and return values through postboxes
\item implementing a call by passing symbols through the inter-module
      link layer
\end{itemize}

\subsection{A call-site in LLVM}
\label{sect:call-site}

In LLVM, a function call is represented by a {\tt call} instruction,
that has the following properties:

\begin{itemize}
\item a pointer to the called function, or the {\it callee}
\item a list of actual arguments, if present, and their types
\item a return type if present, along with a register in
      the calling function or {\it caller}, to which the return value
      is assigned after the callee returns
\end{itemize}

The instance of a {\tt call} instruction in the body of a function is
called a {\it call-site}. When a call is executed the following
sequence of events happen, where $f$ is the caller and $g$ is the
callee:

\begin{enumerate}
\item prepare actual arguments
\item enter $g$
\item read arguments
\item execute body of $g$
\item prepare return value
\item exit $g$
\item return to the call instruction in $f$
\item read the return value
\end{enumerate}

The actual mechanism through which arguments and return values are
transferred between the caller and the callee, is called the {\it
calling convention} of the target platform. Typical examples are:

\begin{itemize}
\item register based, where the caller stores the arguments in a set
      of registers known to the callee, and expects the return value
      in some particular register.
\item stack based, where the caller pushes arguments onto the stack,
      and expects the return value on the stack after the callee
      returns.
\end{itemize}

There are a number of finer details in these conventions - the order
and locations where arguments occur on stack, for example. Or whether
the callee makes any guarantees about the contents of registers after
it returns.

\subsection{The postbox model}

We specify a similar convention meant for translating C call-graphs to
AHIR modules, that uses {\it postboxes}, or memory locations assigned
to each function, that are valid for the entire duration of execution.
Each function is characterised by a set of input locations, output
locations and internal static allocations:

\begin{table}[htb]
\small
\begin{eqnarray*}
S_F & : & (I_F, O_F, A_F) \mbox{ ... space for function F}\\
I_F & : & \mbox{input locations or arguments}\\
O_F & : & \mbox{output locations or return values (at most one in C)}\\
A_F & : & \mbox{static allocations}
\end{eqnarray*}
\caption{Memory locations assigned to function F}
\label{table:locations-f}
\end{table}

Note that the locations mentioned in Table~\ref{table:locations-f} are
simply addresses that are {\it owned} by a function. They do not
represent the actual data-space addressable by this function. This is
because most of the heavy lifting in C is done by passing pointers as
arguments. Predicting the address space accessed by a function during
the execution of the program requires a detailed memory reference
analysis to track down these argument pointers.

\subsection{A function call in AHIR}
\label{sect:ahir-function-call}

A function call in a C program implies transfer of control and data
across modules in the AHIR specification. To achieve this, a {\tt
call} instruction in the input function is translated into a series of
actions in the corresponding AHIR module. These correspond to the list
of events described in Section~\ref{sect:call-site}, that occur in a
function call. The equivalent AHIR actions are listed below:

\begin{enumerate}
\item a store operation for each actual argument, in the data-path for
      function $f$, using a symbolic address label
\item a transition in the control-path that sends a request symbol
      from alphabet $\Lambda$ to the link layer
\item a load operation for each formal argument in the data-path for
      function $g$
\item a transition in the control-path for function $g$ that waits for
      an input symbol that represents this particular call-site
\item a store operation for the return value in the data-path for
      function $g$
\item a transition that sends an acknowledge symbol for this
      call-site, to the link layer
\item a transition in the control-path for function $f$ that waits for
      an acknowledge symbol from the function $g$
\item a load operation for the return value in the data-path for
      function $f$
\end{enumerate}

\section{Software Compiler}

The LLVM compiler framework is used to provide the first stage of the
flow - a C front-end and an optimising software compiler. LLVM uses
the language front-end from the GCC project, modified to generate
bytecode in the LLVM IR. The GCC front-end is capable of parsing a
number of languages including C and C++. It produces {\tt .bc} files
that contain native LLVM bytecode which can be processed further by
various LLVM utilities.

\begin{figure}[htb]
\centering
\includegraphics[width=4.5in,keepaspectratio]{images/llvm-gcc.eps}
%\caption{LLVM-GCC front-end to translate C to native LLVM bytecode}
%\label{fig:llvm-gcc}
\end{figure}

The LLVM tool ``{\tt opt}'' provides the main optimisation framework. 
{\tt opt} is capable of scheduling various analyses and optmising
passes, based on parameters provided by the user. These passes can
work at the level of function, or the level of a compilation unit,
called ``module'' in LLVM terminology.

\subsection{XML Generator}

An AHIR specification is generated as XML by a pass called {\tt
ModuleGen} implemented in LLVM. This pass is invoked by passing the
``{\tt -module-gen}'' argument to {\tt opt}.

The {\tt ModuleGen} pass lowers a module in the LLVM bytecode to
AHIR-XML in stages. The LLVM bytecode uses an instruction as the
smallest unit of code, grouped into a CFG made of basic blocks, for
each function. {\tt ModuleGen} computes control and data dependences
for these instructions, and constructs a CDFG. This CDFG is translated
to an AHIR specification. The AHIR specification is then printed to a
file in XML format.

\begin{figure}[htb]
\centering
\includegraphics[width=4.5in,keepaspectratio]{images/xml-gen.eps}
%\caption{Generating AHIR-XML from LLVM bytecode}
%\label{fig:xml-gen}
\end{figure}

The {\tt ModuleGen} pass treats each function in the program
separately, generating a control-path, data-path and link layer for
it. The XML consists of a callgraph of functions, and a collection of
control-paths, data-paths and link layers. Each entry in the callgraph
describes the following attributes of a function.

\begin{itemize}
\item number and types of arguments and return value
\item init and fin symbols associated with the function
\item list of callsites that describe other functions called by this
      function
\item labels of associated control path, data path and link layer
\end{itemize}

\subsection{XML Linker}

The specification generated by {\tt ModuleGen} is not complete in
itself. The inter-module link layer (Inter-LL) that translates symbols
in the alphabet $\Omega$ is not present. Also, AHIR requires
predetermined memory locations assigned to each function, to be used
as post-boxes for argument passing, and for static allocations. The
addresses used for these locations do not have actual numeric values
assigned to them when the XML is first generated by {\tt ModuleGen}.

These two details in the XML are filled out by a separate linker, that
parses the generated XML and processes it further. The operations
performed by the linker are described in further detail in
Section~\ref{sect:calling-conventions}. The linked XML specification
is complete in all respects, and represents an AHIR version of the
original C program.

\begin{figure}[htb]
\centering
\includegraphics[width=4.5in,keepaspectratio]{images/ir-link.eps}
%\caption{Linker for functions in AHIR form}
%\label{fig:ir-link}
\end{figure}

The linker produces two files - a linked version of the input AHIR-XML
specification and a map of all the memory locations that have been
assigned addresses while linking. This map is useful for the user,
since it provides information about the location of ``postboxes'' used
by the functions for passing arguments. It also serves as a
description of the address space, that will be useful when designing a
memory subsystem that attempts to optimize parallel accesses to memory
locations.

\section{Two-pass flow: Generate and Link}

The input to the overall flow is a file containing all the C functions
that make up a program. The use of precompiled C libraries may be
enabled using suitable mechanisms in the C front-end. But when
producing the final AHIR specification, the bodies for all the
functions must be available.

\subsection{AHIR-XML Generator}

At the generator stage, each function is processed separately. The
LLVM infrastructure presents the program as an LLVM {\tt module}, that
contains all the functions. The generator iterates over the set of
functions in no particular order, and generates a separate (CP,DP,LN)
triplet from the body of each function.

For each function, the generator produces a preamble in the AHIR-XML
file that lists the following details, to be explained in later
subsections:
\begin{itemize}
\item name of the CP, DP and LN
\item values of the {\tt init} and {\tt fin} symbols
\item static allocations - name, size and address-label
\item formal arguments - name, type and address-label
\item return value - name, type and address-label
\item callsites - called function, req/ack symbol, actual arguments,
      return value
\end{itemize}

A callsite is the occurence of a function call in the body of a
function. The preamble of a function in AHIR-XML provides information
about the callsites occuring in that function.

\subsubsection{Passing data - labels for postboxes}

A number of load-store operations are created at the generator stage,
for passing arguments and return values, as described in
Section~\ref{sect:ahir-function-call}. But the actual addresses for the
corresponding postboxes are not known at this stage. Similarly, the
static allocations within the function bodies do not have valid
addresses as well. The operations involving all these memory locations
use labels instead of numerical addresses.

The address-labels mentioned in these preambles are used in the
various data-paths, as inputs for the load/store operations. With such
labels, each function essentially remains {\it relocatable}. The
functions desribed in such a manner can be used in multiple programs
provided they are eventually assigned valid address spaces.

When a function calls other functions, corresponding callsites are
included in the preamble of the caller. The callsite contains a list
of actual arguments and return value, in the form of address labels,
similar to the formal arguments and return value of a function. These
labels are used within the body of the calling function to generate
corresponding load/store operations. Thus, when a function $f$ calls a
function $g$, store operations are included in the data-path for
function $f$, that refer to the address-labels of the actual arguments
for this callsite. Eventually, the address-labels for the actual
arguments in a callsite and the address-labels for the formal
arguments in the called function must match, so that data is
transferred correctly.

\subsubsection{Passing control}

The control mechanisms of a function call are decoupled in a manner
similar to the data transfer. Each function declares a pair of symbols
that represent the start ({\tt init}) and completion ({\tt fin}) of
execution. Similarly, every callsite declares a pair of symbols for
the request and completion of a function call. The task of mapping a
call request to a function-init and mapping a function-fin to a call
completion symbol is done by the inter-module link layer.

In order to keep function definitions reusable, this assignment of
symbols is also deferred until link time. As a result, the
inter-module link layer that uses the alphabet $\Lambda$ is left
undefined in the AHIR-XML at this stage.

\subsection{AHIR-XML Linker}

The linker is responsible for linking together the functions in the
call-graph of a program. It accepts an AHIR-XML description of
functions, along with a call-graph.

\subsubsection{Addresses for postboxes}

The linker traverses the set of functions in no particular order, and
gathers information about the address-space associated with each
function - formal arguments, return values and static allocations. It
then assigns numerical values to these addresses, keeping in mind the
size of each data-type.

When all the addresses have been assigned values, the linker traverses
the list of callsites in each function. For each callsite, the actual
arguments are assigned the same address as the formal arguments in the
called function. This completes the allocation of memory space to the
address spaces of the functions, which had remained undefined prior to
linking.

\subsubsection{Inter Module Link Layer}

\begin{table}[htb]
\small
\begin{eqnarray*}
inter\;module\;link\;layer & : & \{(function,\{callsite\})\}\\
callsite & : & (req,ack)\\
function & : & (init,fin)
\end{eqnarray*}
\caption{Inter Module Link Layer}
\label{table:inter-ll}
\end{table}

The linker also generates the inter-module link layer, in order to
allow control-transfer across functions. For each function, the linker
generates a list of callsites and the req/ack symbols used by them.
These are placed together to form the specification of the
inter-module link layer.

As seen from Table~\ref{table:inter-ll}, the inter-module link layer
provides a mapping between pairs of symbols, or {\tt handshakes}. A
function can be called from multiple callsites, but this information
is not available when the function is compiled. Since this information
becomes available only at link time, the task of returning to the
correct caller must be performed by a module mechanism the function. 
Only the inter-module link layer is in a position to ``remember'' the
calling function and return to it when the called function completes
execution. Hence it has to be specified in terms of pairs of symbols,
in order to retain the association within request-acknowledge
handshakes.

\section{Limitations}

The C source code given as input to the frontend must conform to a
number of restrictions. Some of these restrictions are inherent in
AHIR, while others are a result of limitations in the current
implementation.

\subsubsection*{Constraints inherent in AHIR}
\begin{itemize}
\item pointers to functions are not allowed
\item variable number of arguments to functions are not allowed
\item recursion is not allowed; the call-graph must be a DAG
\end{itemize}

\subsubsection*{Limitations of the current implementation}
\begin{itemize}
\item structures are not supported
\item only 32-bit data is supported - smaller widths are treated as
      32-bit, and large widths will give undefined results, possibly a
      compilation error
\item the call-graph is further restricted to be a tree - one function
      can have only one callsite in the entire program
\item the {\tt underscore} ``{\tt \_}'' is not allowed in identifiers;
      it is reserved for generating internal identifiers
\item all the functions must be available in the same module when
      generating the XML - separate compilation for libraries is not
      supported yet
\end{itemize}

\section{SystemC Simulator}

The simulator provides an event-driven model for AHIR, using the
SystemC framework. The simulator is a single SystemC executable that
includes an XML parser. When invoked, the simulator reads the
specification and a memory map in XML format, and creates a SystemC
model of the description.

Typically, a SystemC model consists of a collection of classes
describing various entities. Instances of these classes are created in
the {\tt sc\_main} function, during the elaboration phase. After that,
the simulation is started by the {\tt sc\_start} function.

\begin{verbatim}
int sc_main(int argc, char* argv[])
{
  IRParser parser;
  parser.parse(ir.xml);
  parser.parse(map.xml);

  IrModel *ir = parser.get_ir();

  sc_clock clk();

  connect_signals(ir, clk);

  sc_start(-1) // run indefinitely until some module calls sc_end()
}
\end{verbatim}

The SystemC simulator is program-independent - no SystemC code is
generated that is specific to the current C program. Instead, the
simulator reads an AHIR-XML specification from a file and directly
creates instance of predefined components. It then proceeds to
simulate the components created, producing a trace of events during
execution. The simulator expects two files at invocation:

\begin{itemize}
\item $\langle$program.xml$\rangle$ - an XML file containing the AHIR
      specification
\item $\langle$map.xml$\rangle$ - an XML file describing contents at
      various memory locations
\end{itemize}

The memory map has to be created by hand, based on the list of
locations generated by the linker. It contains test data that is used
as input during simulation, and final data that is to be compared with
the contents of the simulated memory after execution.

%\section{VHDL Generation}

\end{document}

### 2010-05-24 Documents/Manuals/getting-started.tex
initial commit

\documentclass[12pt]{article}
\usepackage{geometry}
\geometry{verbose,a4paper,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
\usepackage{graphicx}

\title{Getting started with the AHIR tools}
\author{Sameer D. Sahasrabuddhe}
\date{February 2008}

\begin{document}

\maketitle

\section{Overview}

The compilation flow that generates an AHIR circuit specification from
an input C program is a three-stage process:
\begin{enumerate}
  \item parsing the C program using a standard compiler front-end
        (\texttt{llvm-gcc}, provided by LLVM)
  \item generating an unlinked AHIR specification from the compiler's
        standard IR (\texttt{irgen}, implemented using LLVM)
  \item linking the AHIR specification and generating a memory map that
        is assumed by the linked circuit (\texttt{irlink})
\end{enumerate}

\begin{figure}[htb]
  \centering
  \includegraphics[keepaspectratio,width=4in]{images/source-to-ahir.pdf}
  \caption{Compiling input source to AHIR}
  \label{figure:source-to-ahir}
\end{figure}

The linker stage is meant to allow separate compilation of ``source
libraries''. For now the tools only process a single standalone C
file, that must include all the functions required to create a complete
working application.

%The tools that process the linked AHIR specification are all
%standalone AHIR tools, that do not depend on any other project
%framework. Primary operations to be performed on the specification are
%analysis, optimisation, simulation and synthesis.

\begin{figure}[htb]
  \centering
  \includegraphics[keepaspectratio,width=4in]{images/simulation-and-synthesis.pdf}
  \caption{Tools for processing an AHIR specification}
  \label{figure:simulation-and-synthesis}
\end{figure}

\section{Expected input}

Currently, the toolchain works with a single C file that contains all
the functions of a program. There are a number of limitations on the C
code accepted.

\subsection{Restrictions imposed by the AHIR platform}

\begin{enumerate}
  \item The program's call-graph should be a DAG. This condition is
        currently not checked, so the result will be undefined if a
        cyclic call-graph is provided as input.
  \item Function pointers are not allowed.
  \item Functions with variable number of arguments are also not
        allowed.
  \item Dynamic memory allocation and other system calls are not
        possible. They may be implemented by an application
        internally, but they are not part of the AHIR platform.
\end{enumerate}

\subsection{Limitations in the current implementation}

\begin{enumerate}
  \item The complete program must be present in a single file.
  \item The complete DAG must contain a function called {\tt start},
        which denotes the root of the call-graph. The function {\tt
        main} should not be used, since it will trigger internal
        modifications by the C front-end, that are incompatible with
        the ahir tools.
  \item Bit selection operators have not been implemented.
\end{enumerate}

\subsection{Fully supported C constructs}

\begin{enumerate}
  \item 32-bit signed and unsigned integers, 32-bit floats and
        pointers (all other sizes are automatically typecast to 32
        bits)
  \item Arithmetic and bit-wise operators
  \item Pointers to data
  \item Structures and arrays
  \item Static memory allocation
  \item Control structures --- \texttt{if-then-else},
        \texttt{while-do}, \texttt{do-while}, \texttt{for},
        \texttt{break} and \texttt{continue}
\end{enumerate}

\subsection{Running example}

The rest of the document assumes an example input file named {\tt
sample.c}. This file contains a C-program with three functions, and a
call-graph as follows:

\[\texttt{start}\rightarrow\texttt{foo}\rightarrow\texttt{bar}\]

\section{Source compiler}

The input C source is parsed by the LLVM front-end, \texttt{llvm-gcc}. 
This produces a bytecode file that contains the input program in
LLVM's native IR. The command is invoked as follows:

\begin{quote}
\tt llvm-gcc sample.c -o sample.bc
\end{quote}

Note that \texttt{llvm-gcc} is part of the LLVM toolkit. It has a
number of command-line options to control its behaviour. The
installation script for the AHIR tools creates an alias in the users
environment that includes all the relevant switches.

\section{AHIR-XML generator}

The LLVM bytecode generated by the front-end is used as the actual
input that is converted into an AHIR specification. The command
\texttt{irgen} is an LLVM-based tool that reads the bytecode, converts
it to AHIR, and generates an XML file describing the AHIR virtual
circuit.

\begin{quote}
  irgen sample.bc
\end{quote}

This produces an XML file, which contains unlinked AHIR-XML
descriptions of the functions \texttt{start}, \texttt{foo} and
\texttt{bar}. It also contains meta-information about the call-graph
that constitutes \texttt{sample.c}

Since \texttt{irgen} is an LLVM tool, it has access to all the
transformations and optimisations that are available as part of the
LLVM framework. These can be specified by command-line arguments to
the \texttt{irgen} command. The entire list of available operations
can be viewed by invoking \texttt{irgen} as follows:

\begin{quote}
\tt  irgen -help
\end{quote}

%When generating the AHIR specification, \texttt{irgen} imposes a few
%structural constraints on the internal representation of the program.
%To enforce these constraints, some LLVM transformations are always
%invoked by \texttt{irgen} just before generating AHIR, but after
%applying the transformations specified on the command line. These
%hardcoded invocations cannot be controlled from the command-line. They
%can only be changed by modifying the utility itself.

\section{AHIR-XML linker}

The unlinked XML specification also includes the call-graph of the
original program. This information is required to provide
predetermined memory locations called \textit{postboxes} used to pass
arguments during function calls. The modules in the unlinked
specification refer to these locations only through names, since their
addresses are not yet known.

The linker assigns numerical addresses to all the postboxes required
for various function calls. In addition, it also creates locations for
global variables as well as statically allocated variables. It is
invoked on the unlinked XML as follows:

\begin{quote}
\tt  irlink sample.xml
\end{quote}

This produces a linked version of the input XML, in a file called
\texttt{sample\_linked.xml}. Additionally, a memory map is created with the
name \texttt{sample\_map.xml}.

The memory map is an XML file that describes known memory locations
ordered by their addresses. The format includes tags used to describe
structured values as well as scalar values. The map itself is divided
in two sections:

\begin{description}
  \item [\texttt{init}:] the initial state of the memory at invocation
        of the circuit
  \item [\texttt{fin}:] the expected state of the memory after
        execution
\end{description}

As a built-in precaution, \texttt{irlink} never over-writes an
existing memory map. When \texttt{irlink} is run, if the file
\texttt{sample\_map.xml} already exists, it instead creates a second
file called \texttt{sample\_new\_map.xml}. This second file is not
protected, and will be over-written by subsequent runs.

\section{SystemC simulator}

The linked AHIR specification can be simulated using SystemC. The
command \texttt{irsim} reads an AHIR description and a memory map, to
generate a set of SystemC modules in-memory. The modules are run until
the control-path of the module for the \texttt{start} function
triggers its \texttt{fin} transition, marking the end of execution.

\begin{quote}
\tt  irsim --ir sample\_linked.xml --map sample\_map.xml
\end{quote}

The simulator uses the \texttt{init} section of the memory map to
initialise the simulated memory, but ignores the contents of the
\texttt{fin} section.

The simulation results in a trace of the memory accesses seen by the
memory subsystem during execution. For each access to the memory
subsystem, the trace records the time of the access along with the
port and the address used.

Note that the simulator does not generate any actual SystemC
description for now. It simply instantiates a SystemC objects that
represent AHIR structures, and directly runs them.

%This has the
%following limitations:

%\begin{enumerate}
%  \item The SystemC ``library'' that implements AHIR is not pluggable.
%  \item Any implementation of arbitrary-width data-types requires
%        the widths to be constant at compile time. For this, the AHIR
%        specification should be translated to a set of SystemC source
%        files, that are then compiled to generate a simulation. As a
%        result, the current implementation can only support a known
%        data-width.
%\end{enumerate}

\section{VHDL generator}

The VHDL generator is a direct mapping of the AHIR specification to an
implementation. The generator creates a VHDL entity for each module in
the specification. These entities use standard names and interfaces to
instantiate and connect building blocks that are assumed to be
available in an independent library.

\begin{quote}
\tt irsyn --ir sample\_linked.xml --map sample\_map.xml
\end{quote}

This will generate the following set of VHDL files:

\begin{itemize}
  \item \texttt{start\_cp.vhdl}, etc --- one control-path for each function
  \item \texttt{start\_dp.vhdl}, etc --- data-path
  \item \texttt{start\_ln.vhdl}, etc --- intra-module link layer
  \item \texttt{omega.vhdl} --- a single inter-module link layer
  \item \texttt{memory.vhdl} --- the memory subsystem
  \item \texttt{system.vhdl} --- built from all of the above
  \item \texttt{testbench.vhdl} --- used to interact with the system
\end{itemize}

The testbench contains an array of values parsed from the
\texttt{init} section of the memory map. These are used to initialise
the memory subsystem. Similarly the testbench also contains values
specified in the \texttt{fin} section, that are used to verify the
contents of the memory at the end of execution.

\end{document}

### 2010-05-24 Documents/Manuals/memory_map.tex
initial commit

\documentclass[12pt]{article}
\usepackage{geometry}
\geometry{verbose,a4paper,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}

\usepackage[pdftex]{graphicx}
\usepackage{subfigure}

\usepackage{alltt}
%\usepackage{amssymb}

\title{The Memory Map in XML Format}
\author{Sameer D. Sahasrabuddhe}
\date{November 2006}

\newcommand{\irgen}{{\tt irgen}}
\newcommand{\irlink}{{\tt irlink}}
\newcommand{\irsim}{{\tt irsim}}
\newcommand{\irsyn}{{\tt irsyn}}

\begin{document}

  \maketitle

  \section{Introduction}

  The AHIR tools use an XML file to describe the contents of memory as
  seen by the generated AHIR specification. A template map is first
  created by \irlink{} using the locations assigned to various functions
  during the linking process. The map provides a descriptive label for
  each location, along with the type and numerical address. The user
  is expected to fill-in relevant values, using the label.

  \begin{alltt}
  <scalar addr="8" descr="start_formal_0" type="int"></scalar>\end{alltt}

  This instance of the tag ``{\tt <scalar>}'' declares a scalar
  variable described as the first formal argument of the function {\tt
  start}, of type {\tt int}. The address assigned to this variable is
  {\tt 8}. The user may initialise this variable by inserting the
  initial value (say, 10) as contents of the tag:

  \begin{alltt}
  <scalar addr="8" descr="start_formal_0" type="int">10</scalar>\end{alltt}

  \noindent{}{\bf NOTE:} The XML file for the memory map that
  \irlink{} generates, is given a name derived from the name of the
  module. For a module {\tt foo}, for example, the memory map is
  called {\tt foo\_map.xml}. This file is created only when \irgen{}
  runs for the first time. If a file of this name already exists,
  \irgen{} will instead creae a file with the name {\tt
  foo\_new\_map.xml}. This file is over-written on subsequent runs,
  with no warnings.

  \section{Structure of the Memory Map}

  The memory map is an XML document, which must start with the tag
  ``{\tt <map>}'', that has two compulsary attributes - ``{\tt id}''
  and ``{\tt size}''. The memory map is divided into two regions - the
  initial values and the final values present in memory.

  \begin{alltt}
  <map id="func" size="42">
  <init>
  <!-- Initial contents of memory -->

  </init>

  <!-- first free address: 42
   addresses from this value onwards can be used for external data. -->

  <fin>
  <!-- Final contents of memory -->
  </fin>
  </map>\end{alltt}

  \subsection{Populating the memory map}

  The memory map has to be populated only when simulating the
  generated AHIR description of a complete system. For example,
  \irsim{} uses the values to create an {\tt std::map<unsigned,
  bit\_vector>} look-up table which is used to model the external
  memory. In \irsyn{}, the generated test-bench contains an array of
  values, that are written to the system memory through suitable
  access ports before the test bench begins execution.

  In case of synthesisable VHDL code, the test bench is entirely
  ingored, so that the values serve no purpose either. Only the size
  of the memory map is important in order to generate a suitable
  memory component.

  \subsubsection*{The {\tt size} attribute}

  The initial memory map is generated by \irlink{}, and contains all
  the locations that are declared in the input C program. These
  include locations for arguments and return value of each function,
  as well as statically declared variables within functions. These
  locations are assigned addresses linearly, starting from {\tt 0}.

  The generated XML file also contains a comment that mentions the
  first free location in the map. The user is free to use addresses
  from there onwards for additional data. When such additional data is
  introduced, the {\tt size} attribute of the map must be set to a
  number that accomodates all the locations declared. This attribute
  is used as a hint by further tools, when creating a memory component
  in the system.

  \subsubsection*{The {\tt init} and {\tt fin} sections}

  The {\tt init} and {\tt fin} sections of the memory map serve two
  purposes:

  \begin{enumerate}
    \item as data entry points for a user during simulation
    \item as information for driver modules, that need to know the
      location of postboxes when sending data to the system during
      actual runs
  \end{enumerate}

\end{document}

### 2010-05-24 Documents/Notes/Compatibility/LRG.tex
initial commit

% Created 2010-02-14 Sun 14:27
\documentclass[12pt,a4paper]{article}
\usepackage{graphicx}
\usepackage{times}
\usepackage[margin=0.75in]{geometry}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}

\title{Testing for compatibility in the LRG}
%\author{Sameer D. Sahasrabuddhe}
\date{[2010-02-14 Sun]}

\begin{document}

\maketitle

\section{Algorithm}
\label{sec-1}

Given two vertices $u$ and $v$ in an LRG $G$, we test whether they are
compatible using the following algorithm.

\begin{enumerate}
\item Traverse depth-first starting from the vertex $u$, towards the
    root. Mark every vertex $x$ encountered in the traversal and
    record the fork indicated by the edges along which we reached the
    vertex $x$. Every such edge either indicates the same fork, or
    there is only one unlabelled edge along which $x$ is reachable as
    shown in Lemma \ref{all-or-none}.
\item Now traverse the LRG depth-first starting from $v$. When a marked
    vertex $x$ is reached then:

\begin{enumerate}
\item Return true (the two nodes are compatible) in each of the
       following cases:

\begin{enumerate}
\item The edge along which we reached $x$ is not labelled.
\item No fork is recorded at $x$.
\item The edge indicates a fork different from the one recorded at
       $x$.
\end{enumerate}

\item If none of the above three cases holds, then retrace the edge
       along which we reached $x$, and continue the depth-first
       traversal. The edges reachable beyond $x$ are not examined
       since they do not provide any conclusive information, as shown
       in Lemma\ref{look-no-further}.
\end{enumerate}

\end{enumerate}
\section{Details}
\label{sec-2}

Consider a Type-2 petri-net $P$ and its associated LRG $G$. Let $F$ be
a canonical fork region in $P$, such that all the child Class $C$
regions are replaced by simple places. $G'$ is the subgraph induced on
LRG $G$ by the labels assigned to the forks, joins and simple places
in the region $G$.

\begin{definition}

A transition $t_1\in F$ is said to be connected to a transition
$t_2\in F$ {\bf through a simple place} $p \in F$ when there exist
edges $(t_1,p),(p,t_2) \in F$.

\end{definition}

\begin{lemma}

Two transitions $t_1, t_2 \in F$ have the same label $v \in G'$ iff
one element is a join (say $t_1$) and the other element ($t_2$) such
that $t_1$ is connected to $t_2$ through a simple place $p$.
Consequently, the place $p$ is the only place in $F$ with label $v$.

\end{lemma}

\paragraph{Proof:}

\begin{description}
\item [Statement 1] Join $t_1$ is connected to fork $t_2$ through a
   simple place $p$.

   From the way labels are constructed in a Type-2 petri-net, the
   label assigned to the place $p$ and the fork $t_2$ is the same as
   the label assigned to join $t_1$.
\item [Statement 2] Transitions $t_1$ and $t_2$ have the same label.

   Consider transition $t_1\in F$ assigned a label $v_1\in G$. $t_1$
   is either the entry $f'$ of the fork region, or its label is
   derived from the label $v'$ assigned to $f'$. Every transition on
   the path from $f'$ to $t_1$ is either a fork that extends the
   label, or a join that creates a union with some other label. Let
   $S=[t_1,t_2,\ldots]$ be the sequence of forks and joins that occur
   along the path from $f$ to $t_1$.

   If transition $t_2$ has the same label $v_1$, then the path from
   $f$ to $t_2$ must pass through the same sequence $S$. At the end of
   this sequence, let $t_1$ occur first on the path. Since no
   modification occurs to the label when $t_2$ is encountered, the
   following statements must be true:

\begin{enumerate}
\item $t_1$ is not a fork. If it were a fork, it would extend the
      label on every outgoing edge. Hence, $t_1$ is a join.
\item There is no intervening fork or join on the path from $t_1$ to
      $t_2$, since this would also change the label that reaches
      $t_2$. Hence $t_1$ is connected to $t_2$ only through a simple
      place.
\item $t_2$ is not a join. It it were a join, it would generate a new
      label that is a union of the label at $p$ with other labels
      assigned to its other predecessors. Hence $t_2$ is a fork.
\end{enumerate}

\end{description}
\begin{lemma}

If $y_2$ is a vertex in G' with
unlabelled out-edges to distinct vertices $z_1,z_2,\ldots$, then at
most one vertex $z_i$ has out-egdges in the subgraph G'.

\end{lemma}

For vertices $u$ and $v$ in the LRG, let
\mbox{$E(u,v)=\{e_1,e_2,\ldots\}$} be the set of distinct out-edges at
$u$ along which $v$ is reachable from $u$. The following properties
hold for the set $E(u,v)$:

\begin{lemma}
\label{all-or-none}

If an edge $e_{i_0} \in E(u,v)$ is unlabelled, then $E(u,v) =
\{e_{i_0}\}$, else every edge $e_i \in E$ indicates the same fork $f$.

\end{lemma}

\end{document}

### 2010-05-24 Documents/Optimisations/register-elimination.tex
initial commit

\documentclass[a4paper,12pt]{article}
\usepackage{geometry}
\geometry{margin=1in}

\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{times}
\usepackage{tikz}

\newtheorem{statement}{Statement}[section]
\newtheorem{condition}{Condition}

\newcommand{\SYM}[1]{$\operatorname{#1}$}

\title{Eliminating registers at the outputs of data-path operators}
\date{}

\begin{document}

\maketitle

\section{Introduction}

Every operator in the data-path has a register at its output, that
takes at least one clock cycle before signalling completion. In some
cases it may be possible to eliminate this register. The result is a
combinational path from the input of the operator to the output
register of the next operator. For example,
Figure~\ref{fig:operator-sequence} shows a sequence of operations
($A\longrightarrow B\longrightarrow C$) where ovals represent
combinational logic and boxes represent registers. If the register
\SYM{reg\_B} is removed, the logic for \SYM{op\_B} and \SYM{op\_C} is
combined into one large combinational path as shown in
Figure~\ref{fig:combinational-short-circuit}.

\begin{figure}[htb]
  \centering
  \subfigure[Operator logic separated by registers.]{
    \label{fig:operator-sequence}
    \includegraphics[scale=0.75]{images/operator-sequence.pdf}}\\
  \vspace{0.25in}
  \subfigure[Combinational path lengthened by eliminating register
  reg\_B.]{
    \label{fig:combinational-short-circuit}
    \includegraphics[scale=0.75]{images/combinational-short-circuit.pdf}}
  \caption{Eliminating registers at operator outputs.}
  \label{fig:register-elimination}
\end{figure}

For this transformation to be correct, we must guarantee that the
absence of the register does not result in invalid data at the input
of other nodes in the data-path. When  \SYM{reg\_A} is updated,
the new value at the output of \SYM{op\_B} is immediately available at
the input of \SYM{op\_C}, in the absence of  \SYM{reg\_B}. This
may result in an incorrect computation at \SYM{C}, if the actual
operation was meant to use the old value stored in \SYM{reg\_B}. In
the following sections, we propose a set of necessary and/or
sufficient conditions that guarantee that the output of all
computations remains valid after eliminating a register.

\section{Terminology}
\label{sec:terminology}

\begin{description}
\item [Register definition:] An ``operation''(?) in the control-path
  that results in updating the value of a register in the data-path is
  called a \emph{definition} of that register. The set of such
  definitions of a register $A$ is written as $D_A =
  \{d^0_A,d^1_A,\ldots\}$.
\item [Register use:] A definition $d^i_B$ is said to \emph{use}
  register $A$ when the value assigned to $B$ is computed in a
  ``non-trivial''(?) manner from the value of $A$. This is written as
  $A\mapsto d^i_B$.
\item[Predecessor:] A register $A$ is termed as the \emph{predecessor}
  of a register $B$ if the value of $A$ is used in some definition of
  $B$, denoted by the operator ``$\longrightarrow$''. Also, $B$ is
  termed as a \textbf{successor} of $A$.

  \[A \longrightarrow B \iff \exists d^i_B \mbox{ such that } A\mapsto
  d^i_B\]
\end{description}

\section{Values latched from external sources}
\label{sec:external-data}

Some operators in the data-path acquire values from the environment,
such as memory access operators and input/output ports. The
environment is free to change the value after that particular
invocation of the operator is done --- the register at the output of
the operator is required to maintain it until it has been used by
other operations in the data-path. Hence that register cannot be
eliminated.

\begin{condition}
  \label{condition:external-data}
  Every definition of the candidate register must assign a value
  computed from an expression that consists of only constants and
  values stored in other registers in the data-path.
\end{condition}

\section{Registers with conditional definitions}
\label{sec:conditional-definitions}

In general, a data-path operator may have multiple control inputs and
multiple data outputs (registers). When the operator is invoked by a
particular control input, the state machine of the operator determines
whether a particular register is defined. Each such definition may
assign a different value to the same register. This behaviour results
in a multiplexer that selects the result of multiple combinational
circuits. If such a register is eliminated, the internal multiplexing
(which is dynamic) will drive invalid data at the output.

\begin{condition}
  \label{condition:conditional-definitions}
  Every definition of the candidate register must assign the same
  value to the register.
\end{condition}

\begin{figure}[htb]
  \centering
  \includegraphics[scale=0.3]{images/phi-function.pdf}
  \caption{Multiple definitions in a $\phi$-function.}
  \label{fig:phi-function}
\end{figure}

For example, the $\phi$-function in the AHIR data-path has two control
inputs, two data inputs and a single output register, as shown in
Figure~\ref{fig:phi-function}. When definition $d^0_M$ (or
respectively $d^1_M$) is invoked, the corresponding data-input $D_0$
(respectively $D_1$) is assigned to the output register. Eliminating
the output register will make it impossible to forward only one of the
two data inputs to the output. Hence, the register at the output of a
$\phi$-function cannot be eliminated.

\section{Definition-Use sequences}
\label{sec:def-use-chains}

A register $R$ that satisfies
Condition~\ref{condition:conditional-definitions} is assigned a value
generated by a single combinational circuit. The output of the circuit
changes when any of its input registers is updated. In the absence of
the register $R$, this value is immediately visible to the uses of
$R$. Hence $R$ can be eliminated only if the control-path guarantees
that when any of the inputs change, $R$ is always updated before it is
used.

\begin{condition}
  \label{condition:def-use-chains}
  For every predecessor $P$ and successor $S$ of the candidate
  register $B$, if there is a path from any definition of $P$ to any
  definition of $S$ that uses the value of $B$, then that path must
  pass through some definition of $B$.
\end{condition}

\begin{figure}[htb]
  \centering
  \includegraphics[scale=0.3]{images/branch-conditional-definition.pdf}
  \caption{Conditional definitions in the presence of branches.}
  \label{fig:branch-conditional-definition}
\end{figure}

For example, consider the output of operator $B$ shown in
Figure~\ref{fig:branch-conditional-definition}. Definition $d^0_B$ is
executed only if the appropriate branch is taken in the control-path
after definition $d^0_P$. The value of $B$ is used by definition
$d^1_M$, which occurs along the same path in the control-path. Thus,
every path from a definition of $P$ to a definition of $M$ that uses
$B$ (there is only one such path) passes through a definition of $B$.
Hence $B$ satisfies this condition for elimination.

\begin{figure}[htb]
  \centering
  \includegraphics[scale=0.3]{images/elimination-in-a-pipeline.pdf}
  \caption{Registers in a pipeline.}
  \label{fig:pipeline-registers}
\end{figure}

Figure~\ref{fig:pipeline-registers} shows a pipeline that invokes
$d_A$, $d_B$ and $d_C$ in parallel, where $A\mapsto d_B$ and $B\mapsto
d_C$. Register $B$ does not satisfy
Condition~\ref{condition:def-use-chains} since there is a path from
$d_A$ to $d_C$ which does not pass through $d_B$. Hence register $B$
cannot be eliminated.

\section{Cycles in the data-path}
\label{sec:data-path-cycles}

Every cycle in the data-path must have at least one register in order
to prevent a combinational loop.

\begin{condition}
  \label{condition:data-path-cycles}
  If the candidate register occurs in a cycle in the data-path, then
  it must not be the only register in that cycle.
\end{condition}
\end{document}

### 2010-05-24 Documents/Presentations/BARC/proposal.tex
initial commit

\documentclass[12pt]{article}

\usepackage{geometry}
\geometry{verbose,a4paper,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}

\usepackage{graphicx}
\usepackage{multirow}

\newenvironment{packed_enum}{
\begin{enumerate}
  \setlength{\itemsep}{1pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
  \setlength{\partopsep}{0pt}
  \setlength{\topsep}{0pt}
}{\end{enumerate}}

\begin{document}

\newcounter{srno}
\newcommand{\setsr}[1]{\setcounter{srno}{#1}}

\newcounter{srinc}
\newcommand{\incten}[0]{\setcounter{srinc}{10}}
\newcommand{\incone}[0]{\setcounter{srinc}{1}}

\newcommand{\srno}[2][\arabic{srno}]{\noindent\paragraph{#1. #2}\addtocounter{srno}{\value{srinc}}}

\incone
\setsr{100}

\newcommand{\mytitle}{Development of a high-level hardware synthesis
  tool-chain}

\begin{titlepage}
\begin{large}
\begin{center}
\vspace{0.5cm}
Project Proposal\\
\vspace{0.5cm}
for\\
\vspace{0.5cm}
{\Large \bf \mytitle}\\
\vspace{2.5cm}
\begin{tabular}{l}
\hline
{\bf Principal Investigator}\\
\hline
Prof. Kavi Arya\\
Centre for Formal Design and Verification of Software, and\\
Department of Computer Science and Engineering,\\
Indian Institute of Technology - Bombay,\\
Powai, Mumbai - 400076\\
\\
kavi@cse.iitb.ac.in\\
\hline
\end{tabular}
\end{center}
\end{large}
\end{titlepage}

\begin{center}
  \bf \Large Section-A\\
  \large Part I\\
  \vspace{2ex}
  Project Overview\\
\end{center}

\srno{Advisory Committee Code Number:} 34 (ATC)

\srno{Title:}
\begin{center}
  \large\mytitle
\end{center}

\srno{Keywords: } high-level synthesis, hardware compilation, EDA

\srno{Project Summary:}

\paragraph{} High-level synthesis is the process of compiling
high-level programs to hardware circuits. This reduces the cost of
hardware design by providing two advantages. Firstly, it lowers the
level of expertise required so that even programmers can generate
hardware designs, and secondly, it significantly reduces the time and
effort required to create efficient implementations. High-level
synthesis is also relevant to safety-critical applications since it
guarantees provably correct implementations of the input
specification. The resulting hardware does not have to be verified;
only the behaviour of the input program has to be functionally
verified.

Current work at IIT Bombay has established a scalable automated
high-level synthesis flow that guarantees correctness, called AHIR. It
has been tested on diverse examples such as cryptography, linear
algebra, data structures and signal processing. The flow takes
advantage of established compiler techniques to infer what the user
meant when they wrote the program, and produce the most suitable
hardware for it.

The proposed project aims to make the ongoing research at IIT Bombay
available to BARC in the form of a synthesis tool-chain based on AHIR.
The tool-chain will also be demonstrated on a system currently being
developed at BARC --- a hardware implementation of the TCP/IP network
protocol stack. Such a system is being designed at BARC using the
commercial high-level synthesis product called Handel-C. A second
implementation will be created under the proposed project, using AHIR
to synthesise a C implementation of TCP/IP to hardware. This will
provide the basis for a practical evaluation of the tool-chain.

The proposed project is a starting point for long-term development
work on the use of high-level synthesis at BARC. Access to such an
indigenous technology is of strategic importance to BARC in terms of
self-reliance. The users at BARC will have direct access to the
research group working on AHIR at IIT Bombay. This will allow BARC to
keep pace with further developments in high-level synthesis including
optimisations and the ability to target emerging hardware platforms.

\newpage

\srno{Principal Investigator}
\paragraph{}
\begin{tabular}{l}
  Prof. Kavi Arya\\
  \\
  Centre for Formal Design \& Verification\\
  of Software, and\\
  Dept. of Computer Science and Engg.,\\
  Indian Institute of Technology - Bombay,\\
  Powai, Mumbai - 400076\\
  \\
  kavi@cse.iitb.ac.in\\
\end{tabular}

\srno{Co-Investigator}
\paragraph{}
\begin{tabular}{l}
  Prof. Madhav P. Desai\\
  \\
  Dept. of Electrical Engineering,\\
  Indian Institute of Technology - Bombay,\\
  Powai, Mumbai - 400076\\
  \\
  madhav@ee.iitb.ac.in\\
\end{tabular}

\srno{Principal Collaborator}
\paragraph{}

\setsr{109}

\newpage

\srno{Budget Amounts in Rupees (Lakhs): }

\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
\multicolumn{2}{|c|}{Particulars} & Year I & Year II & Year III & Total \\
\hline
\hline
Equipment &  & 7.25 & -- & -- & 7.25 \\
Salaries & JRF & 1.38 & 1.38 & 1.56 & 4.32 \\
 & JRF & -- & 1.38 & 1.56 & 2.94 \\
 & RA-III & 4.20 & 4.20 & 4.80 & 13.20 \\
Consumables &  & 0.25 & 0.25 & 0.25 & 0.75 \\
Travel Allowance & PI & 1.05 & 1.05 & 1.05 & 3.15 \\
 & PC & 0.20 & 0.20 & 0.20 & 0.60 \\
Technical Assistance &  & 0.15 & 0.15 & 0.15 & 0.45 \\
Contingencies &  & 0.15 & 0.15 & 0.15 & 0.45 \\
\hline
Total &  & 14.63 & 8.76 & 9.72 & 33.11 \\
\hline 
Overheads &  & 2.19 & 1.31 & 1.46 & 4.97\\
\hline
Total &  & 16.82 & 10.07 & 11.18 & 38.08\\
\hline
\end{tabular}
\end{center}

\srno{Equipment Cost:}

\begin{center}
\begin{tabular}{|c|c|c|c|}
\multicolumn{4}{l}{\bf Amounts in Rupees (Lakhs)}\\
\hline
Item & Quantity & Rate & Amount \\
\hline
Desktop PC & 3 & 0.50 & 1.50 \\
External hard drives & 2 & 0.05 & 0.10 \\
Laptop & 1 & 0.65 & 0.65 \\
FPGA board & 2 & 1.00 & 2.00 \\
\hline
Total & & & 4.25 \\
\hline
\end{tabular}
\end{center}

\newpage

\begin{center} {\bf \large Part-II\\
\vspace{2ex}
Project Objectives, Research Plan and Deliverables}
\end{center}

\setsr{200}
\incten

\srno{List of Objectives}

The overall purpose of the project is to develop a high-level
synthesis software suite to be used at BARC, based on the research
being done at IIT Bombay. In particular there are two objectives:

\begin{enumerate}
\item Deliver a software tool-chain based on AHIR that provides the
  following facilities:
  \begin{enumerate}
  \item Compile C programs to synthesisable RTL descriptions.
  \item Run and evaluate the generated hardware on an FPGA board.
  \item Test and debug the behaviour of the FPGA implementation with
    respect to the input C programs.
  \end{enumerate}

\item Evaluate the effectiveness of the high-level synthesis flow by
  automatically generating a hardware implementation of the TCP/IP
  network protocol stack, starting from a behavioural description
  written in C.
\end{enumerate}

\srno{Yearly Research Plan and Deliverables}

\paragraph{} A high-level synthesis tool-chain based on AHIR was
implemented as part of of ongoing research at IIT Bombay on high-level
synthesis. The tool-chain consists of a set of tools that can
translate a subset of the C programming language to hardware modules.
It has been tried on a number of examples that represent various
applications such as cryptography, digital signal processing,
high-performance computing and high-level data structures. Work done
under this project will result in a complete software bundle being
delivered to BARC for regular use. This includes the implementation of
on-going research work at IIT Bombay aimed at enabling the compilation
of very large systems specified as C programs.

We propose to use a C implementation of the TCP/IP network protocol
stack as a working example during development. This will serve to
demonstrate the effectiveness of the synthesis tool-chain. Insights
gained during this exercise will also provide data for the on-going
research at IIT Bombay.

\subsubsection*{Developing the tool-chain}

The existing implementation at IIT Bombay has been used successfully
to demonstrate the potential gains of a high-level synthesis flow
based on AHIR. The work under the proposed project will be aimed at
realising these potential benefits and also packaging the prototype as
a complete end-to-end solution. In particular, the work will be
concentrated on following areas of development:

\begin{enumerate}
\item Support for a minimum set of features including the following:
  \begin{enumerate}
  \item Complete support for the ISO C language standard.
  \item Input/Output ports.
  \item Separate compilation and pre-compiled libraries.
  \end{enumerate}
\item Improvements in the hardware produced, including the following:
  \begin{enumerate}
  \item Improvements in the library of building blocks --- efficient
    operators, multicycle operators, pipelined operators etc.
  \item Efficient utilisation of hardware resources.
  \item Improvements in throughput by automatically recognising and
    eliminating unnecessary sequential delays when generating
    hardware.
  \item Support for sharing of pipelined multicycle operators.
  \end{enumerate}
\item Support for the design and implementation of large systems by
  separately designing various modules and automatically composing
  them into a single coherent system.
\end{enumerate}

The development work on these areas will involve a synergy between the
proposed project and the ongoing research work on AHIR at IIT Bombay.
Of the three areas of thrust highlighted here, the first two are
largely concerned with improving the existing implementation of the
AHIR tool-chain. The third area deals with extending the current
implementation to include new features being developed at IIT Bombay.
The current tool-chain takes a single C program and generates a
hardware implementation for it. But in order to support a large
system, the tool-chain must address the following issues:

\begin{enumerate}
\item Concurrent execution of modules in a system.
\item Efficient communication channels between the modules.
\item A memory subsystem for external storage of data, that can keep
  up with the demands of the system modules.
\item Facility to use a high-level description of the system
  architecture. The entire system will be automatically generated from
  this high-level description by combining the individual modules.
\end{enumerate}

Research work in this area being done at IIT Bombay will greatly
benefit from experience in using the existing AHIR tool-chain in
implementing a non-trivial system in hardware. The proposed project
will assist in this respect through the development of a TCP/IP
network protocol stack in hardware for use at BARC.

\subsubsection*{TCP/IP implementation}

The development of the TCP/IP implementation will begin in parallel
with the work on the AHIR tool-chain. Initial work will largely focus
on characterising the major components of such a system, and building
an early version using the existing AHIR implementation. Each
component will be implemented as a C program and compiled to hardware
using AHIR. The components will be put together manually to produce
the first prototype of the entire system.

The first manual effort is important in terms of highlighting the
automatic functionality expected from the AHIR tool-chain, that allows
a user to describe a large system as a collection of C programs, and
expect to automatically generated an efficient hardware
implementation. The result will be a complete high-level specification
of the TCP stack, that can be automatically compiled to competitive
hardware using AHIR.

\subsubsection*{Yearly Deliverables}

  \renewcommand\arraystretch{1.2}
\begin{tabular}{|c|p{0.45\textwidth}|p{0.45\textwidth}|}
  \hline
  Year & Activity & Deliverable\\
  \hline
  \hline
  I & Identifying and implementing ISO C features
  that are not  supported in the current implementation. &
  \multirow{4}{0.45\textwidth}{Complete    support for the C language
    in the AHIR tool-chain.}\\
  \cline{2-2}
  & Implementing I/O ports. & \\
  \cline{2-3}
  & Implementing multicycle and pipelined operators. & A library of
  efficient hardware operators to be used by the generated circuit.\\
  \cline{2-3}
  & Improving parallelisation of operators and memory accesses. &
  Improved throughput in the generated circuits. \\
  \cline{2-3}
  & \multirow{3}{0.45\textwidth}{Designing the overall architecture of
    the TCP/IP hardware    implementation.} & A preliminary overview
  of the system architecture,  along with expected performance and
  hardware costs. \\
  \hline
  \hline
  II & Implementing hardware optimisations including sharing of
  pipelined operators, sharing of registers and system-level resource
  reuse. & \multirow{4}{0.45\textwidth}{Hardware optimisation module
    in the AHIR tool-chain.} \\
  \cline{2-3}
  & Implementing separate compilation and    library linking. &
  Support  for pre-compiled libraries in the  AHIR  tool-chain.\\ 
  \cline{2-3}
  & \multirow{5}{0.45\textwidth}{Implementing the components in the
    TCP/IP system as separate     programs and manually combining them
    for a complete implementation.}   & A preliminary implementation
  of the TCP/IP system.\\
  \cline{3-3}
  & & A characterisation of the meta-level information required for
  automatic generation of entire systems. \\
  \hline
  \hline
  III & Implementing a system-level synthesis tool
  that combines  independent components using a high-level
  architectural description  to produce a complete system. & A
  complete high-level synthesis  tool-chain for generating complex
  systems from large programs. \\ 
  \cline{2-3}
  & Automatically generating a hardware of the TCP/IP network protocol
  stack using the AHIR tool-chain. & A complete high-level description
  of the TCP/IP stack along with a  working implementation on an FPGA.\\
  \hline
\end{tabular}

\newpage

\begin{center} {\bf \large Part-III\\
\vspace{2ex}
Budget Estimates}
\end{center}

\setsr{290}
\incten

\srno{Details of Budget Requirements --- Amounts in Rupees (Lakhs)}

\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
\multicolumn{2}{|c|}{Particulars} & Year I & Year II & Year III & Total \\
\hline
\hline
Equipment &  & 7.25 & -- & -- & 7.25 \\
Salaries & JRF & 1.38 & 1.38 & 1.56 & 4.32 \\
 & JRF & -- & 1.38 & 1.56 & 2.94 \\
 & RA-III & 4.20 & 4.20 & 4.80 & 13.20 \\
Consumables &  & 0.25 & 0.25 & 0.25 & 0.75 \\
Travel Allowance & PI & 1.05 & 1.05 & 1.05 & 3.15 \\
 & PC & 0.20 & 0.20 & 0.20 & 0.60 \\
Technical Assistance &  & 0.15 & 0.15 & 0.15 & 0.45 \\
Contingencies &  & 0.15 & 0.15 & 0.15 & 0.45 \\
\hline
Total &  & 14.63 & 8.76 & 9.72 & 33.11 \\
\hline 
Overheads &  & 2.19 & 1.31 & 1.46 & 4.97\\
\hline
Total &  & 16.82 & 10.07 & 11.18 & 38.08\\
\hline
\end{tabular}
\end{center}

\srno{Equipment}

\paragraph{} The PI proposes to procure three desktop PCs along with
two FPGA boards that will be used as the accelerator platform. The PI
also proposes to provide a laptop to research staff at RA-III level
for purposes of presentation of the technical work, progress reports
and documentation during the duration of the project. In addition,
this laptop will also carry the actual code of the tool for
development, debugging and demonstration purposes. This will be
important to ensure that there are no impediments to progress of the
work even when the key persons are travelling and may not be present
at the same physical location. Since the development work involves a
large amount of data, the PI proposes to purchase external hard drives
to be used with the desktop PC and the laptop. The budget requirements
for this equipment are listed below:

% \begin{enumerate}
% \item \textbf{Desktop PC:} Intel Core 2 Duo processor, 4GB DDR2 RAM,
%   2$\times$250GB SATA HDD, 19" TFT LCD monitor, DVD writer, keyboard,
%   mouse.
% \item \textbf{Laptop:} Intel Core 2 Duo processor, 2GB DDR2 RAM, 250GB
%   SATA HDD, 14" WXGA LCD, 10/100 Ethernet, 802.11g, Bluetooth.
% \item \textbf{FPGA board:} 
% \end{enumerate}

\begin{center}
\begin{tabular}{|c|c|c|c|}
\multicolumn{4}{l}{\bf Amounts in Rupees (Lakhs)}\\
\hline
Item & Quantity & Rate & Amount \\
\hline
Desktop PC & 3 & 0.50 & 1.50 \\
External hard drives & 2 & 0.05 & 0.10 \\
Laptop & 1 & 0.65 & 0.65 \\
FPGA board & 1 & 5.00 & 5.00 \\
\hline
Total & & & 7.25 \\
\hline
\end{tabular}
\end{center}

\newpage
\srno{Staff Salary --- Amounts in Rupees}

\paragraph{}The PI proposes to recruit two research staff for
undertaking the development work for the first year, and three
research staff for the second and third year. One position will be at
the level of RA-III for all three years, for a person with a PhD in
CSE/EE. One position at the level of JRF will be created for the first
year, with an additional position at the level of JRF for the second
and third year. The JRF positions will be open for persons with a
BTech in CSE/EE. The detailed budget estimate for staff salaries are
given below.

\begin{center}
\begin{tabular}{|c|c|c|c|c|}
\multicolumn{5}{l}{\bf Amounts in Rupees (Lakhs)}\\
\hline
Staff & Year I & Year II & Year III & Total \\
\hline
JRF & 1.38 & 1.38 & 1.56 & 4.32 \\
 & @Rs.11,500 p.m. & @Rs.11,500 p.m. & @Rs.13,000 p.m. & \\
\hline
JRF & -- & 1.38 & 1.56 & 2.94 \\
 & & @Rs.11,500 p.m. & @Rs.13,000 p.m. & \\
\hline
RA-III & 4.20 & 4.20 & 4.80 & 13.20 \\
 & @Rs.35,000 p.m. & @Rs.35,000 p.m. & @Rs.40,000 p.m. & \\
\hline
Total & 5.58 & 6.96 & 7.92 & 20.46 \\
\hline
\multicolumn{5}{l}{{\bf Note:} The HRA paid is 30\% of the consolidated salary.}
\end{tabular}
\end{center}

\setsr{330}

\srno{Consumables}

\paragraph{} The budgetary requirement of consumables like printer
cartridges, paper, backup CDs and other stationaries, is given below.

\begin{center}
\begin{tabular}{|c|c|c|c|c|}
\multicolumn{5}{l}{\bf Amounts in Rupees (Lakhs)}\\
\hline
Item & Year I & Year II & Year III & Total \\
\hline
Papers, Cartridges, CDs & 0.25 & 0.25 & 0.25 & 0.75 \\
and other consumables & & & & \\
\hline
Total & 0.25 & 0.25 & 0.25 & 0.75 \\
\hline
\end{tabular}
\end{center}

\newpage

\srno{Travel}

\paragraph{} It is expected that the research and development work
undertaken as part of this project will lead to at least two to three
international publications. This will help in getting the work peer
reviewed by the research community working in related areas of
verification. This may require travelling within India and also abroad
to present the research results in peer reviewed conferences. The
travel budget shown below also includes the budget required for travel
by the PC to PI's institute and by the PI to PC's institute for
regular meetings.

\begin{center}
\begin{tabular}{|c|c|c|c|c|}
\multicolumn{5}{l}{\bf Amounts in Rupees (Lakhs)}\\
\hline
Item & \multicolumn{4}{c|}{Funds required} \\
\cline{2-5}
 & Year I & Year II & Year III & Total \\
\hline
Weekly visits by PC & 0.20 & 0.20 & 0.20 & 0.60 \\
to PI's premises & & & & \\
Monthly visits by PI & 0.05 & 0.05 & 0.05 & 0.15 \\
to PC's premises & & & & \\
Attending national \& & 1.00 & 1.00 & 1.00 & 3.00 \\
international conferences & & & & \\
\hline
Total & 1.25 & 1.25 & 1.25 & 3.75 \\
\hline
\end{tabular}
\end{center}

% \srno{Contingency}

% \srno{Overheads}

% \srno{Grand Total}

% \begin{center}
%   \large
%   \vspace{2ex}
%   BUDGET DETAILS
%   \vspace{2ex}
% \end{center}

% \srno[310]{Details of the budget for equipment to be procured by the
%   PI (Amount in Rupees):}

% \srno[320]{Details of staff salary (Amount in Rupees):}

% \srno[340]{Details of budget for consumables to be procured by the PI
%   (Amount in Rupees):}

% \setsr{350}
% \incone

% \srno{Details of travel:}
% \srno{Proposed number of visits of PC/DC to PI's institute}
% \srno{Duration of stay during each visit (No. of days)}
% \srno{Total funds required}
% \srno{Proposed number of visits of PI to PC/DC's institute}
% \srno{Duration of stay during each visit (No. of days)}
% \srno{Total funds required}
% \srno{Funds required by PI for travel to attend conferences within India.}
% \srno{Funds for Other visits}

% \begin{center}
%   \large
%   \vspace{2ex}
%   BUDGET JUSTIFICATIONS
%   \vspace{2ex}
% \end{center}

\newpage

\begin{center}
  \bf \Large Section-B\\
  \vspace{2ex}
\end{center}

\incone
\setsr{499}
\srno{Technical Information}

\section{Introduction}

Given an algorithm, one can either implement it on a microprocessor,
or design hardware with an equivalent behaviour. Although compiling a
program to a microprocessor is convenient, a hardware implementation
has a number of advantages:

\begin{enumerate}
  \item Higher performance.
  \item Decreased power dissipation.
  \item Reduced size.
\end{enumerate}

But hardware platforms continue to evolve, making it possible to
integrate ever larger systems onto a single device. As the size
increases, it becomes more and more difficulty to cope with the
complexity of the design. In addition, it is important that the design
be independent of the underlying hardware, so that the target platform
may be changed without affecting the design itself. This can be only
be addressed by making effective use of {\em high-level synthesis},
the process of translating high-level programs into hardware.

The use of high-level programs as the starting point makes hardware
design accessible to a very large set of users. If the compiler is
guaranteed to produce a correct implementation of the input
specification, the resulting hardware does not have to be verified. 
Only the behaviour of the input specification has to be verified,
which can be done using existing software verification practices.

A competitive high-level synthesis process must preserve common
practices in software programming while providing a verifiable and
optimisable path to a hardware implementation. This requires a
compiler flow that can generate efficient circuits from complex
high-level programs. Such a compiler flow must have the following
features:

\begin{enumerate}
  \item It should be independent of the programming language
        used.
  \item It should guarantee a correct implementation of the
        specified behaviour.
  \item It should support optimisations that can scale to very
        large systems.
  \item It should be able to target new hardware platforms as
        they emerge.
\end{enumerate}

AHIR is an attempt at implementing such a high-level synthesis flow. 
AHIR uses an intermediate representation that makes it possible to
decouple high-level issues encountered when writing programs, from
low-level issues in the hardware implementation. An AHIR specification
is factorised into three components: control-flow, data-flow and
storage. The three components can be analysed and transformed
separately without affecting each other, enabling optimisations that
can scale with the size of the circuit. At the same time, AHIR
captures sufficient information from the input program, so that it can
be routinely mapped to a variety of hardware platforms.

\section{AHIR}

AHIR is a compact graph-based hardware representation that provides a
delay-independent specification of the circuit. A specification in
AHIR can be routinely translated to hardware. Features such as
parallelism and resource sharing may be available in the input
language, or introduced by the compiler. In either case, the
intermediate representation is rich enough to describe them.

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.25]{omega.pdf}
  \caption{An AHIR system.}
  \label{fig:system}
\end{figure}

A system in AHIR consists of a number of modules, each consisting of a
control-path and a data-path. Typically, each function in an input
program is translated to a module. Function calls are implemented by
passing the call request along with the relevant arguments through an
inter-module link layer. The modules can also share data through an
external memory subsystem.

AHIR does not include any details about the memory architecture; only
a simple linear address space is assumed. Each module can specify an
arbitrary number of memory access ports. The only requirement from the
memory subsystem is that a request made on an access port must be
served eventually.

\begin{figure}[h]
\centering
\includegraphics[scale=0.3]{decouple.pdf}
\caption{An AHIR module.}
\end{figure}

A module in AHIR consists of two closely interacting components: a
control-path and a data-path. The data-path in a module is a pool of
operators connected by wires. The control-path is a petri-net that
specifies the ordering of events in the module. AHIR uses a subclass
of petri-nets called {\em Type-2 petri-nets}, which has been defined
to guarantee liveness and safety, while supporting scalable analyses.

%\begin{figure}[h]
%\centering
%\includegraphics[scale=0.35]{alpha-schema.pdf}
%\caption{Handshakes between control and data paths.}
%\end{figure}

The control-path and data-path interact through a link layer by
exchanging request-acknowledge handshakes that encapsulate delays
occurring in the implementation. AHIR specifies a set of constraints
on the delays involved --- an implementation that satisfies these
constraints is guaranteed to be correct.

\begin{figure}[h]
\centering
\includegraphics[scale=0.3]{isochron.pdf}
\caption{Delay constraints in AHIR.}
\label{fig:isochron}
\end{figure}

Figure~\ref{fig:isochron} shows a hypothetical example with associated
delays. The numbered delays $d_0$ to $d_5$ in this figure are not
individual values, but representatives of their respective class of
delays. The delays in the implementation must satisfy the following
two inequalities in order to guarantee correctness:

\[d_5 \le d_0 + d_1 + d_3\]
\[d_2 \le d_3 + d_4 + d_0\]

\subsection{Key innovations}

\begin{enumerate}
\item An orthogonal representation of the control, data and storage
      aspects of the input program, that can be routinely translated
      to a hardware implementation.
\item Support for static analysis and optimisation procedures that can
      scale with the size of the program.
\item A verifiable compilation path from algorithms to digital
      hardware, based on the intermediate representation.
\end{enumerate}

\subsection{High-level Synthesis using AHIR}

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.3]{flow.pdf}
  \caption{High-level Synthesis using AHIR.}
\end{figure}

The high-level synthesis flow based on AHIR consists of two phases: a
software phase that optimises and translates a high-level program to
AHIR, and a hardware phase that optimises and translates an AHIR
specification to a hardware implementation. The result is an
end-to-end compiler flow that translates C programs to synthesisable
VHDL descriptions.

The software phase consists of existing software compilation
techniques along with a backend that generates AHIR specifications. 
The translation to AHIR uses a well-known intermediate form called a
CDFG (Control Data Flow Graph). The CDFG is first derived from the
optimised program and then translated to an AHIR specification using a
simple piece-wise construction method. The method used guarantees
correctness {\it by construction} --- the generated AHIR specification
correctly implements the behaviour specified by the CDFG.

The control, data and storage components of the AHIR specification  can
now be implemented independently. The factorisation allows us to
efficiently analyse and transform these components in order to
optimise the circuit. We demonstrate this with an optimisation
that improves resource utilisation by sharing resources in the
data-path.

The resource sharing is based on a static analysis of the control-path
that identifies pairs of operations that cannot be active at the same
time. The algorithms and supplementary data structures used by the
analysis are very simple and close to linear in complexity. 
Experimental results show that even this simple approach is quite
effective in improving the resource utilisation of the circuit.

The optimised AHIR specification is translated to a circuit
implementation by the hardware phase. The compiler is free to explore
various architectures in this phase, as long as the timing constraints
in AHIR are satisfied. The constraints are easy to satisfy in practice
by sufficiently padding the relevant circuit delays. The AHIR
specification is then translated to a synchronous circuit
implementation by directly replacing each component with an equivalent
VHDL implementation.

\subsection{Results}

\begin{figure}[h]
\centering
\includegraphics[scale=0.8]{tpg.pdf}
\caption{Comparison of an FPGA implementation with the P-4 and RTL}
\label{figure:tput-area}
\end{figure}

The current implementation has been tested on a number of examples
chosen from diverse application domains: LINPACK, Red-Black Trees,
FFT, AES block cipher and A5/1 stream cipher. One set of experiments
compares the performance of an FPGA implementation with two extremes:
equivalent programs running on a microprocessor (Intel Pentium~IV),
and hand-crafted circuits that implement the same behaviour (results
obtained from third-party literature). The metric used for comparing
performance is the ratio of the throughput in terms of jobs per second
completed by the circuit to the area occupied by it in terms of
equivalent gates. Figure~\ref{figure:tput-area} shows the performance
of the different implementations on a logarithmic scale.

\begin{figure}[h]
\centering
\includegraphics[scale=0.8]{ppw.pdf}
\caption{Comparison of an ASIC implementation with the P-4
(normalised to ``1'')}
\label{figure:ppw}
\end{figure}

A second set of experiments uses an automated VHDL-to-ASIC flow based
on commercially available tools. ASIC layouts were automatically
generated for TSMC 180nm technology using this flow. Since actual SRAM
cells were not available, the experiment uses SRAM models provided by
the CACTI project. The estimated power dissipation and area were
scaled to the 130nm technology used for the Pentium~IV. 
Figure~\ref{figure:ppw} shows the ``performance~per~watt'' delivered
by the ASIC implementation, normalised by the performance of the
Pentium~IV.

The synthesis results show that it is possible to translate complex
programs to hardware circuits using AHIR. The generated circuits are
competitive with equivalent programs running on a general-purpose
microprocessor. But the performance is less than that of hand-crafted
circuits by two orders of magnitude. This gap can be bridged by
further work in three key areas: expressing parallelism in the
high-level specification, introducing optimisations on the AHIR
specification, and using customised memory subsystems that support
parallel accesses.

\subsection{Future work}

AHIR can potentially be used as a universal design platform that
unifies {\it software compilation} (translating programs to
executables) with {\it hardware compilation} (synthesising circuits
from programs). The expressive power of AHIR is evident at two levels:
as a framework for exploring hardware architectures, and as a target
for the implementation of high-level languages. A combination of these
two aspects will result in a compiler flow that can map various
classes of high-level languages to different low-level architectures.

\subsubsection{Programming languages}

Current work in translating C programs to hardware demonstrate that
AHIR can support imperative languages, with suitable implementations
for specific features. Further research should explore the possibility
of supporting system-level languages such as SystemC and Esterel.

\subsubsection{Hardware Platforms}

AHIR has already been used for mapping C programs to FPGAs and ASICs.
But AHIR itself is not limited to only these platforms. A circuit
described in AHIR can result in a custom microprocessor, for example,
by suitably modifying the data path. AHIR can also be used to
target emerging platforms such as SoCs, arrays of functional units,
ASIC-FPGA hybrids, etc.

\subsubsection{Hardware optimisations}

The synthesis of large systems starting from complex high-level
programs requires further work towards improving performance at the
system level. The compiler must be able to generate implementations
that deliver good throughput while occupying acceptable amounts of
resources. These two parameters are usually traded-off with each other
since faster computations come at the cost of more hardware, and {\it
vice versa}. AHIR provides many opportunities for creating
transformations that improve the final circuit, such as pipelining for
improved throughput and sharing resources across modules for reduction
in area.

\subsubsection{Memory subsystem}

In addition to optimised hardware, the performance delivered by the
memory subsystem is critical to the overall performance of the system. 
High-level synthesis has to be coupled with an automated memory design
flow to generate application-specific memory subsystems that can keep
up with the hardware system.

Some work has been done by related projects at IIT Bombay towards
developing simple performance models for memory subsystems. Such
models will support an integrated memory sub-system design procedure
that can explore large areas of the design space in a feasible manner. 
For a given application, the design process may be parameterised by
the properties of the expected memory access trace.

\subsection{Publications}

\begin{enumerate}
\item ``AHIR: A Hardware Intermediate Representation for Hardware
      Generation from High-level Languages'' by Sameer D. 
      Sahasrabuddhe, Hakim Raja, Kavi Arya and Madhav P. Desai,
      presented at the 20th International Conference on VLSI Design,
      January 2007.
\item PhD Thesis entitled ``A competitive pathway from high-level
      programs to hardware specifications'' by Sameer D. Sahasrabuddhe
      (Dept. of CSE, IIT Bombay), submitted for review in October
      2008.
\end{enumerate}

% \newpage

% \begin{center} {\bf \large Part-V\\
% \vspace{5ex}
% Appendix}
% \end{center}

% \section{CV of Principal Investigator (PI)}

% \section{CV of Principal Collaborator (PC)}

\end{document}

### 2010-05-24 Documents/Publications/dsd2010/paper.tex
initial commit

\documentclass[conference]{IEEEtran} 
\usepackage{graphicx}
\usepackage[tight,footnotesize]{subfigure}
\usepackage{amsmath}
\usepackage{alltt}
\usepackage{multirow}
\usepackage{times}

\title{A C-to-RTL flow as an energy efficient alternative to embedded
  processors in digital systems}

\author{
  Sameer D. Sahasrabuddhe, Kunal P. Ghosh, Sreenivas
  Subramanian, Kavi Arya and Madhav P. Desai\\
  Indian Institute of Technology -- Bombay, Powai, Mumbai -- 400076, INDIA\\
  Email: \{sameerds@cse, ssreenivas@ee, kunalghosh@ee, kavi@cse, madhav@ee\}.iitb.ac.in
}
\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]

\newcommand{\sym}[1]{$\operatorname{#1}$}

\newbox\subfigbox             % Create a box to hold the subfigure.
\makeatletter
  \newenvironment{subfloat}% % Create the new environment.
    {\def\caption##1{\gdef\subcapsave{\relax##1}}%
     \let\subcapsave=\@empty % Save the subcaption text.
     \let\sf@oldlabel=\label
     \def\label##1{\xdef\sublabsave{\noexpand\label{##1}}}%
     \let\sublabsave\relax    % Save the label key.
     \setbox\subfigbox\hbox
       \bgroup}%              % Open the box...
      {\egroup                % ... close the box and call \subfigure.
     \let\label=\sf@oldlabel
     \subfigure[\subcapsave]{\box\subfigbox\sublabsave}}%
\makeatother

\pagestyle{empty}

\begin{document}

\maketitle
\thispagestyle{empty}

\begin{abstract}

  We present a high-level synthesis flow for mapping an algorithm
  description (in C) to a provably equivalent register-transfer level
  (RTL) description of hardware. This flow uses an intermediate
  representation which is an orthogonal factorization of the program
  behavior into control, data and memory aspects, and is suitable for
  the description of large systems. We show that optimizations such as
  arbiter-less resource sharing can be efficiently computed on this
  representation. We apply the flow to a wide range of examples
  ranging from stream ciphers to database and linear algebra
  applications. The resulting RTL is then put through a standard ASIC
  tool chain (synthesis followed by automatic place-and-route), and
  the performance and power dissipation of the resulting layout is
  computed. We observe that the energy consumption (per completed
  task) of each resulting circuit is considerably lower than that of
  an equivalent executable running on a low-power processor,
  indicating that this C-to-RTL flow offers an energy efficient
  alternative to the use of embedded processors in mapping algorithms
  to digital VLSI systems.

\end{abstract}

\section{Introduction}

The design of complex digital systems is expensive due to two reasons:
the need for trained manpower, and the difficulty of verification at
every step of the design process. Thus, it is often the case that
low-power embedded microprocessors are used to implement complex
algorithms in digital systems, so that the design and verification
problems are moved to the software domain. A high-level synthesis flow
which maps an algorithm directly to a hardware description can be a
potential alternative if it provides a verifiable and optimisable path
from executable specifications\cite{gajski-executable-specification}
to hardware.

In addition to the ease of use, performance measures and energy
considerations are standards by which a high-level synthesis flow
should be judged relative to the competing methodologies --- custom
designed RTL and embedded processors.  It is very critical
that the results of the high-level synthesis flow are characterized
with respect to their area/power/delay/energy, both in absolute terms
and relative to competing methodologies.  As far as comparisons
with  custom designed hardware are concerned, a high-level synthesis flow
which starts from a C/C++ description of the algorithm is unlikely to
be competitive, mainly because the amount of parallelism expressible in 
a C/C++ program is not sufficient.   It should be possible to address
this shortcoming by starting with algorithms described in parallel
programming languages, but such discussions are beyond the scope of
this paper.

We provide comparisons between
the hardware generated by our high level synthesis flow and  
processor implementations of an algorithm described in C.  
Note that it is not entirely obvious that algorithm-specific hardware 
generated by a high-level synthesis flow will be superior
to embedded processor implementations of the same algorithm.  
One can argue that the algorithm-specific hardware
can exploit the characteristics of the algorithm to a greater
degree than a general purpose processor.  
On the other hand, when a C/C++ program is compiled to
a processor, the platform (that is, the processor) is usually a
heavily optimized one in which each component has been tweaked to a
high level, whereas the algorithm-specific
hardware is generated by automated tools whose optimization abilities
will most probably not match those of a human designer.   
Concrete data which throws light on these aspects is valuable.

Our high-level synthesis flow is based on a factored intermediate
representation which guarantees the correctness of the implementation,
and supports optimisations that can scale with the size of the
program. The synthesis flow introduces an intermediate step in the
form of a representation called AHIR (A Hardware Intermediate
Representation). The representation is independent of the programming
language used, and can be routinely translated to a hardware
implementation, while also supporting scalable
optimisations\cite{AHIR_sameerds_hakim_2007}\cite{ahir_thesis}.

In order to characterize the energy efficiency of the resulting
circuits, we select a wide range of applications (from stream ciphers
to databases and linear algebra) and run them through our C-to-RTL
flow. The resulting RTL is then mapped to an ASIC using industry
standard automatic synthesis tools. This ASIC is characterized for
power, delay, area and energy, and the results are compared with those
obtained when the same program is mapped to a low power processor.
These results indicate that even in its current state, our high-level
synthesis flow is an energy competitive alternative to the use of
embedded processors in the design of complex systems.

\section{Related work}

There has been a great deal of activity in the field of 
high-level synthesis in the context of converting a high
level program to RTL.   Attempts at creating a path from 
high-level programs to hardware descriptions can be loosely 
categorised as follows:

\subsection{Improvements over RTL}

Efforts such as Bluespec\cite{bluespec-memo} raise the abstraction in an
RTL description in order to support higher-level constructs. Such a
language can be very powerful in expressing the architecture of the
hardware, but the target user is a hardware designer who can
effectively utilize this expressive power.

\subsection{Modified high-level languages}

Some efforts reinterpret programming languages as hardware
descriptions, and also extend them with special features. The language
SA-C\cite{SA-C}, for example, is a purely
functional subset of C that disallows pointers. On the other hand,
Handel-C\cite{handel-c} is a language that guarantees complete ISO-C
compatibility and also provides additional primitives.

In both examples, the designer must use specific features to generate
efficient hardware, instead of the compiler inferring a hardware
implementation. This forces the programmer to reevaluate standard
programming practices.

\subsection{High-level programs as hardware specifications}

Some efforts simply interpret a program as a behavioral
specification, which is mapped to a hardware implementation using an
intermediate representation.

For example, the Phoenix project uses an intermediate representation
called Pegasus\cite{pegasus-ir} for a compiler flow from C to
hardware\cite{pegasus-cash}. A description in Pegasus can be
implemented as a micro-pipeline. The representation allows the compiler
to natively implement a number of high-level transformations.

The SPARK\cite{spark-vlsi-paper} project uses an internal
representation based on hierarchical task graphs (HTG). The compiler
uses a heuristic to combine high-level transformations on the HTG ---
such as code motion and speculation\cite{spark-code-motion} --- with
scheduling and resource binding to produce efficient hardware.

Our work is similar, since the goal is to transparently compile
programs into hardware. AHIR differs from both Pegasus and SPARK since
it factorises the system into three separate components: control flow,
data flow and memory.  The components can be optimised and
implemented separately as long as specified constraints are satisfied.

Several commercial tools which translate a C program to hardware are
also available.  For example, Catapult-C \cite{catapult-c} from
Mentor Graphics, C-to-Silicon \cite{c-to-silicon} from Cadence,
and Synopsys HLS \cite{synopsys-hls} from Synopsys.  Currently,
we have no access to the internal details of these proprietary tools.

\section{AHIR}

A system in AHIR is a collection of modules connected to a memory
subsystem as shown in Fig.~\ref{fig:system}. Each module represents
one function from the input program. Function calls are implemented
through an \emph{inter-module link layer}. The architecture of the
memory subsystem is not defined in AHIR. It is only required to
service every request \emph{eventually}.

\begin{figure}[!t]
  \centering
  \includegraphics[scale=0.2]{images/omega.eps}
  \caption{An AHIR system.}
  \label{fig:system}
\end{figure}

\subsection{A module in AHIR}

\begin{figure}[!t]
  \centering
  \includegraphics[scale=0.25]{images/decouple.eps}
  \caption{An AHIR module.}
  \label{fig:decouple}
\end{figure}

Two flows are described in an AHIR module --- control-path and
data-path --- that communicate through an \emph{intra-module link
  layer} as shown in Fig.~\ref{fig:decouple}. The control-path is a
petri-net that specifies the ordering of events in the module. The
data-path is a pool of hardware resources connected by wires.

Communication through the link layers is specified as an exchange of
symbols. The set of symbols associated with a component is called its
alphabet. The data-path uses alphabet $\Sigma$, while the control-path
uses alphabet $\Lambda$. The interaction with the inter-module link
layer is represented by the alphabet $\Omega$.

\subsection{Data-path}

The data-path is a directed hyper-graph, where the edges represent
values, and the nodes represent operations on these values. Each edge
is a hyper-edge with a single tail and one or more heads. The tail
drives a value on the edge, which reaches all the heads
instantaneously.

A data-path node is described by a state machine with an \sym{idle}
state, and one or more \sym{busy} states. When a request
$\operatorname{req}_i$ arrives at an idle node, the node samples all
its incoming data-edges and changes state to $\operatorname{busy}_i$.
The node then operates on the sampled values and updates the outgoing
data-edges. On completion, the node emits an acknowledge symbol
$\operatorname{ack}_j$, and then returns to the \sym{idle} state.

The data-path uses load and store operators to communicate with
external memory. Additionally, there are input and output ports that
are used to exchange arguments and return values with the inter-module
link layer during a function call.

\subsection{Control-path}
\index{control-path}

%\begin{figure}
%  \centering
%  \includegraphics[scale=0.25]{images/control-path-example.eps}
%  \caption{An AHIR Control-path.}
%  \label{fig:control-path-example}
%\end{figure}

The control-path is a petri-net that expresses the sequence in which
events occur in a module. It is required to be in a class called
``Type-2 Petri-nets'', as defined in Section~\ref{section:type-2}.
The circuit-level implementation paradigm to be used for an AHIR module is
not specified (the implementation may be synchronous or asynchronous).
Thus, petri-nets provide a neutral, powerful and compact mechanism to 
describe parallelism in the control flow of an algorithm.  
As we shall see in Section~\ref{section:type-2}, a ``Type-2'' petri-net
provides an easily verifiable, yet powerful enough mechanism to
describe control flow.

The transitions in the control path are associated with input and
output symbols.  An {\em input transition} is associated with
an input symbol (that is, the transition is assumed to have fired
when the input symbol is received by the control path), and an {\em output transition}
is associated with an output symbol (the output symbol is emitted
when the transition fires).

The control-path has a single marked place in the initial marking.
This enables a single transition called \sym{init}, which responds to
a request symbol in alphabet $\Omega$ and begins execution of the
module. The end of execution is represented by the \sym{fin}
transition that emits an acknowledgment symbol in $\Omega$ and marks
the initially marked place.

\subsection{The Intra-module Link Layer}
\index{link layer!intra-module}

The intra-module link layer translates symbols generated by the
control-path (in $\Lambda$) to symbols for the data-path (in $\Sigma$)
or the inter-module link layer (in $\Omega$), and \emph{vice versa}.
It is defined as a set of functions that instantaneously consume
symbols presented to them and generate new symbols.

\subsection{Handshakes and delay constraints}
\label{sect:handshakes}

\begin{figure}[!t]
\centering
\includegraphics[scale=0.25]{images/isochron.eps}
\caption{Delays in an AHIR specification.}
\label{fig:isochron}
\end{figure}

Operations in AHIR are managed by symbolic handshakes. The
control-path emits a \sym{request} in $\Lambda$, which triggers an
operator in the data-path. When done, the operator emits an
\sym{acknowledge} in $\Sigma$, which causes further events in the
control-path. This \emph{request-acknowledge handshake} encapsulates
any delays in the implementation.

An implementation must satisfy a pair of one-sided delay constraints
for the handshake in order to ensure correctness.
Fig.~\ref{fig:isochron} shows a hypothetical example with associated
delays. When the control-path emits the request symbol \sym{Req}, it
must update its state before the arrival of the acknowledgment symbol
\sym{Ack}. Hence we have:
\[d_5 \le d_0 + d_1 + d_3\]

Similarly, when the data-path emits an \sym{Ack}, it eventually
receives a \sym{Req}. The data inputs must have stabilized before
this request arrives. Hence we have:
\[d_2 \le d_3 + d_4 + d_0\]

Note that the term $d_0 + d_3$ is common to both expressions. An
implementation can always guarantee timing correctness by sufficiently
padding these delays to satisfy the constraints.

\subsection{Execution model}
\index{execution!AHIR module}

AHIR uses the following execution model. The control-path responds
instantaneously to the arrival of symbols, while data-path elements
take some finite non-zero amount of time to execute. Values in the data-path are also
propagated instantaneously. Clearly, this satisfies the delay
constraints, since $d_1$ is finite, while other delays are zero.  
In principle, RTL implementations of AHIR can be synchronous or asynchronous.  In 
our current flow however, we generate synchronous RTL implementations from
AHIR.

\subsection{The Inter-module Link Layer}
\index{link layer!inter-module}

The inter-module link layer is used to route function calls between
modules. It has one arbiter for each module, that manages the flow of
input arguments and return values between the calling module and the
called module.

\section{Type-2 petri-nets}
\label{section:type-2}

\begin{figure}[!t]
  \centering
  \includegraphics[scale=0.2]{images/TPR.eps}
  \caption{A TPR and a Type-1 petri-net}
  \label{figure:TPR}
\end{figure}

\begin{figure}[!t]
  \centering
  \subfigure[Primitive]{
    \label{figure:primitive-TPRs}
    \includegraphics[scale=0.2]{images/Type-2-terminals.eps}}
  \subfigure[Series]{
    \label{figure:series-TPR}
    \hspace{0.125in}
    \includegraphics[scale=0.2]{images/Type-2-linear.eps}
    \hspace{0.125in}}
  \subfigure[Fork]{
    \label{figure:fork-TPR}
    \includegraphics[scale=0.15]{images/Type-2-forks.eps}}
  \hspace{0.125in}
  \subfigure[Branch]{
    \label{figure:branch-TPR}
    \includegraphics[scale=0.15]{images/Type-2-branches.eps}}
  \subfigure[Parallel merges]{
    \label{figure:merge-region}
    \includegraphics[scale=0.2]{images/merge-region.eps}}
  \caption{Type-2 construction rules.}
  \label{figure:Type-2-TPRs}
\end{figure}

Control paths in AHIR are instances of Type-2 petri-nets (introduced
in \cite{ahir_thesis}), which
are defined using a set of standard construction rules. 
The structure of a Type-2 petri-net is
designed to enable analyses and transformations that are scalable with
the size of the petri-net.

\begin{definition} A {\bf simple place (transition)} is a place
(transition) with one incoming edge and one outgoing edge.
\end{definition}

\begin{definition} A {\bf token-preserving region (TPR)} is a
petri-net $P$ that can be augmented with one simple place $\hat{p}$
and a sufficient number of simple transitions and edges, to produce a
live and safe petri-net $P'$ such that $\hat{p}$ is the only marked
place in the initial marking (as shown in Fig.~\ref{figure:TPR}).
\label{definition:TPR}
\end{definition}

\begin{definition} A {\bf Type-1 Petri-net} is a live and safe
petri-net that marks one simple place in the initial marking. Clearly,
a Type-1 petri-net is constructed by augmenting a TPR.
\end{definition}

\begin{definition}
A {\bf Standard TPR (STPR)} is a TPR constructed using the standard
set of rules (defined below).
\end{definition}

\begin{definition} A {\bf Type-2 petri-net} is a Type-1 petri-net
created by augmenting a Standard TPR.
\end{definition}

\subsection*{Type-2 construction rules:}

\begin{enumerate}
  \item A simple place or transition is a {\em primitive} STPR.

  \item A {\em series region} is an STPR formed by joining two
        standard STPRs in series.

  \item A connected {\it acyclic} sub-graph made of STPRs, forks and
        joins is itself an STPR called a {\em fork region}.

  \item A connected (possibly cyclic) sub-graph made of STPRs, branches
        and merges is an STPR called a {\em branch region}.

  \item Replacing a merge place in a branch region with parallel
        merges as shown in Fig.~\ref{figure:merge-region} also
        results in a standard TPR. The set of parallel merges
        introduced by this replacement is called a {\em parallel-merge
        region}.
\end{enumerate}

The Type-2 construction rules are illustrated in
Fig.~\ref{figure:Type-2-TPRs}. The branch region allows cycles in
order to express arbitrary branch and loop structures. The fork region
does not allow cycles since that introduces further conditions for
liveness. But this does not affect the expressive power of Type-2
petri-nets. Parallel-merge regions implement the run-time selection of
values at the exit of a branch, such as variable {\tt d} in
Fig.~\ref{figure:C-and-SSA}.

\section{Arbiter-less sharing of data-path operators}

We describe an optimization that reuses a data-path operator for
multiple operations that are never active at the same time. This
avoids arbitration overheads since there is no contention. The
optimisation uses an almost linear static analysis of the control-path
to identify sharing opportunities.

\begin{definition} An operator is said to be {\bf active} at a given
instant of time if and only if it has received a request, but not
yet emitted an acknowledgment.
\end{definition}

\begin{definition} Two operators $M_1$ and $M_2$ are said to be {\bf
compatible} if and only if $M_1$ does not receive a request while
$M_2$ is active, and {\it vice versa}.
\end{definition}

\subsection{Compatibility in Type-2 petri-nets}

In a Type-2 petri-net, compatibility of two transitions is determined
by the nature of the smallest STPR that contains them, which we term
as their {\em nearest common ancestor} (NCA). Two operations can
potentially be incompatible only if the NCA is a fork region;
operations in a branch or series NCA region are always compatible.

\begin{figure}[!t]
  \centering
  \subfigure[A fork region]{
    \label{figure:compatibility-fork-paths}
    \includegraphics[scale=0.225]{images/compatibility-fork-paths.eps}}
  \subfigure[Compatibility]{
    \begin{tabular}[b]{|c|ccccc|}
      \hline
      & 1 & 2 & 3 & 4 & 5\\
      \hline
      1 & Y & - & Y & Y & Y\\
      2 & - & Y & - & Y & -\\
      3 & Y & - & Y & Y & -\\
      4 & Y & Y & Y & Y & -\\
      5 & Y & - & - & - & Y\\
      \hline
    \end{tabular}
    \label{figure:compatibility-fork-table}}
  \caption{Compatibility in a Type-2 petri-net.}
  \label{figure:compatibility-fork}
\end{figure}

Fig.~\ref{figure:compatibility-fork} shows a fork region in a Type-2
petri-net with numbered segments and their compatibility with each
other. Segments $1$ and $2$ are \emph{incompatible} since they can
execute concurrently. But segments $1$ and $4$ are compatible, since a
sequence is enforced by the path through segment $3$.

\begin{definition} Two elements $e_1$ and $e_2$ in a Type-2 petri-net
are compatible if and only if one of the following is true:
\begin{enumerate}
\item their NCA is {\em not} a fork region
\item there is a path within the NCA fork region, joining the two elements
\end{enumerate}
\end{definition}

\subsection{Compatibility labeling}

We use a labeling scheme to record the paths reaching a petri-net
element from the \sym{init} transition. The labeling is a symbolic
execution of the Type-2 petri-net. Two elements can be compared for
compatibility using their labels instead.

A label is a set $L = \{l_0,l_1,\ldots\}$ where each $l \in L$ is a
sequence of $n$ label elements $l = [a_0,a_1,\ldots,a_{n-1}]$. A label
element is a 3-tuple $(f,k,i)$ made of a fork identifier $f$, the
fan-out $k$ of the fork, and an index $i$ into the fan-out. A label
element $a = (f,k,i)$ is said to {\it indicate} the fork $f$.

The product of two labels $(A * B)$ is
defined as the concatenation of pairs of sequences from the two
labels: $A*B = \{a.b |\forall a \in A, \forall b \in B \}$. For
convenience, the product operator is overloaded to represent the
product of a label with a single element: $A*b = A*\{[b]\}$.

%% Common operators such as equality $(A = B)$, concatenation $(a.b)$ and
%% prefix $(a \leq b)$ are defined to have sensible meanings. The prefix
%% operator defines a partial order on label sequences. The longest
%% common prefix (LCP) of two sequences is the longest sequence that is a
%% prefix of both the sequences.

\subsection{Labeling scheme}

Parallel-merge regions are first reduced to simple merges, which
simplifies the labeling without affecting compatibility. The $init$
transition is assigned an empty label. The label of every element is
computed from its predecessors. Only forks and joins result in a new
label; other elements are assigned the same label as their
predecessors.

\subsubsection{Labeling at a fork}

If $L$ is the label assigned to a fork $f$ with $k(f)$ successors,
then each successor $s_i$ is assigned the label $L *
\bigl(f,k(f),i\bigr)$, as shown in Fig.~\ref{figure:fork-labels}.

\begin{figure}[!t]
  \centering
  \subfigure[Fork.]{
    \label{figure:fork-labels}
    \includegraphics[scale=0.25]{images/fork-label-products.eps}}
  \subfigure[Union at a join.]{
    \label{figure:join-union}
    \hspace{0.125in}
    \includegraphics[scale=0.25]{images/join-union-labels.eps}
    \hspace{0.125in}}
  \subfigure[Reduction at a join.]{
    \label{figure:join-reduce}
    \includegraphics[scale=0.25]{images/join-reduce-labels.eps}}
  \caption{Labeling scheme.}
\end{figure}

\subsubsection{Labeling at a join}

In the general case --- such as transition $m'$ in
Fig.\ref{figure:compatibility-fork-paths} --- a join is assigned the
union of the labels assigned to all its predecessors, as shown in
Fig.~\ref{figure:join-union}. But when the join receives all the
tokens starting from a particular fork --- such as transition $m$ in
Fig.\ref{figure:compatibility-fork-paths} --- the union is reduced to
remove the label elements indicating that fork, as shown in
Fig.~\ref{figure:join-reduce}. Usually, only a subset of the union is
reduced, since paths from unrelated forks may reach a join.

The reduction at the join ensures that the labeling scheme is
``closed'' --- all the extensions created within a fork region
disappear at the exit of a fork region\cite{ahir_thesis}. Finally, the
\sym{fin} transition is assigned an empty label.

\subsection{Label Representation Graph (LRG)}

The compatibility label is a record of every path reaching that
element from the $init$ transition, which results in an exponential
size. Comparing two labels for compatibility also has exponential
complexity, since every sequence in one label has to be compared with
every sequence in the other label.

We eliminate the complexity by using a {\em label representation
  graph} (LRG) as shown in Fig.~\ref{figure:LRG}. The LRG represents
labels as nodes, where edges represent the manner in which a label is
constructed from other labels. The LRG is a directed acyclic graph
with a single root node that represents the empty label.

\begin{figure}[!t]
  \centering
  \includegraphics[scale=0.225]{images/LRG-example-cp.eps}
  \hspace{0.125in}
  \includegraphics[scale=0.225]{images/LRG-example.eps}
  \caption{Label Representation Graph.}
  \label{figure:LRG}
\end{figure}

Let $l(u)$ be the label represented by a node $u$. An edge $(u,v)$ in
the LRG may itself be labeled with a label element $e$, in which case
it represents the product operation $l(v) = l(u) * e$. If multiple
incoming edges are incident at a node $v$, then it represents a label
that is the union of all its predecessors. In a well-formed LRG,
multiple incoming edges are incident on a node if and only if they are
all unlabeled.

The LRG is a compact representation of compatibility labels in a
Type-2 petri-net. Each path reaching a node $u$ from the root of the
LRG represents one label sequence in the label $l(u)$. The following
result is proved in \cite{ahir_thesis}.

\begin{theorem}

  Two operations with labels represented by nodes $u$ and $v$ in the
  LRG are said to be compatible, if and only if one of the following
  is true:

\begin{enumerate}
  \item There is a path from $u$ to $v$ or \emph{vice versa}.
  \item There exists a node $a$ in the LRG, from which $u$ and $v$ are
        reachable along non-intersecting paths such that one of the
        following is true:
    \begin{enumerate}
      \item One or both paths begin with an unlabeled edge.
      \item The labels on the first edges in the paths indicate
            different forks.
    \end{enumerate}
\end{enumerate}
\label{theorem:compatibility-LRG}
\end{theorem}

\subsection{Shared operators}

We have used the LRG to implement arbiterless sharing in the AHIR
synthesis flow. We use a simple greedy algorithm to generate cliques
of pair-wise compatible operations that are mapped to a single
operator in the data-path. The incoming data-edges are routed through
multiplexers; the registers for the outgoing data-edges are not
shared.

This scheme for arbiterless sharing is quite effective in reducing
hardware costs, demonstrating support for scalable optimisations in
AHIR. Synthesis results show improvements in the throughput-area ratio
(measured in $\operatorname{Hz}/\operatorname{slice}$) in the range of
15--190\% depending on the application\cite{ahir_thesis}.

\section{The synthesis flow}

The synthesis flow uses the LLVM \cite{ref:LLVM}  framework to parse and optimise the
%
% http://www.llvm.org
%
input C program. This is translated to an AHIR specification using a
CDFG as an intermediate step.

\subsection{C to CDFG}

\begin{figure}[!t]
\centering
\begin{subfloat}%
\hspace{0.125in}
  \begin{minipage}[b]{0.75in}%
    \small%
\begin{alltt}
d = m + n
b = m - n
if (b > 0):
  a = b + c
  d = e + a
x = d + 2
\end{alltt}%
  \end{minipage}%
  \caption{Pseudo-code.}%
  \label{figure:C-snippet}%
\hspace{0.125in}
\end{subfloat}%
\hspace{0.5in}%
\begin{subfloat}%
  \begin{minipage}[b]{0.85in}%
    \small%
\begin{alltt}
d1 = m + n
b = m - n
if (b > 0):
  a = b + c
  d2 = e + a
d3 = \(\phi\)(d1,d2)
x = d3 + 2
\end{alltt}%
  \end{minipage}%
  \caption{SSA form.}%
  \label{figure:ssa-snippet}%
\end{subfloat}%
\caption{A code fragment and its SSA form.}%
\label{figure:C-and-SSA}%
\end{figure}

The starting point is a C program.  In the current implementation,
the only restrictions on the C program are that the call graph
should not contain any cycles, and that function pointers should
not be used (pointers to data can be used in the usual way)\footnote{These restrictions
are not fundamental to the approach, but we have not gotten around to eliminating
them so far.}.
The C program is first converted to LLVM byte-code, which is based on
the SSA form\cite{muchnick}. This is a purely functional form that
removes the notion of individual {\it variables} from a program. Every
assignment to a variable is a distinct value; multiple assignments
that occur in branches are handled by a special instruction called the
$\phi$-function, as shown in Fig.~\ref{figure:C-and-SSA}.

\nocite{conditional_branches_rim_jain_1992}
\nocite{DFG_jong_1991}
\nocite{DFG_standard_eijndhoven_1992}
\nocite{flow_graph_orailoglu_1986}
\nocite{functional_synthesis_TASS_amellal_kaminska_1994}

The LLVM bytecode is then translated to a control data flow graph
(CDFG)\cite{conditional_branches_rim_jain_1992}-%
\cite{functional_synthesis_TASS_amellal_kaminska_1994}, as shown in
Fig.~\ref{fig:snippet-cdfg}. The CDFG represents instructions as
nodes connected by control and data flow edges. The edges in the CDFG
arise from three kinds of dependence in the original program:

\begin{itemize}
\item Data dependence that creates control and data flow.
\item Control structures that create control flow.
\item External dependence (in storage) that creates control flow.
\end{itemize}

\subsection{Generating AHIR from a CDFG}

\begin{figure*}
\centering
\subfigure[CDFG.]{%
  \includegraphics[scale=0.2]{images/cdfg.eps}%
  \label{fig:snippet-cdfg}}%
\hspace{0.25in}%
\subfigure[Control Path.]{%
  \includegraphics[scale=0.3]{images/controlpath.eps}%
  \label{fig:snippet-cp}}%
\hspace{0.25in}%
\subfigure[Data Path.]{%
  \includegraphics[scale=0.22]{images/datapath.eps}%
  \label{fig:snippet-dp}}%
\caption{Translating a CDFG to AHIR.}%
\label{figure:cdfg-to-ahir}%
\end{figure*}

The AHIR specification is generated by piece-wise translation. Each
node or edge in the CDFG is replaced by an equivalent \emph{AHIR
  fragment}, and the fragments are connected to obtain the control and
data paths. Fig.~\ref{figure:cdfg-to-ahir} shows the input CDFG and
the resulting AHIR specification for our example. Elements that are
obvious from the context have been hidden. Two structures are shown in
detail --- a decoder element (\texttt{D1}) that examines the condition
for a branch, and a multiplexer element (\texttt{P1}) that implements
a $\phi$-function.

\subsection{Correctness}

The method used by our synthesis flow guarantees that the generated
circuit specification is correct {\em by construction}. The first step
of translating a C program to a CDFG is a routine one that does not
need to be verified separately.

The correctness of the second step (CDFG to AHIR) is shown
as follows \cite{ahir_thesis}: suppose that
$A$ was the AHIR specification generated from the
reference CDFG $G$.  Then every execution sequence
in $A$ is a member of the set of
possible execution sequences of $G$.  Now,
from $A$, we construct a new CDFG $G'$ such
that every execution sequence of  $G'$  is
an element of the set of execution sequences
of $A$.  Now, $G'$ and $G$ are shown to be isomorphic,
so that their sets of execution sequences are in 
a one-to-one correspondence with each other.  This
shows that the sets of execution sequences in $A$ and
$G$ are in a one-to-one correspondence.

\subsection{AHIR to VHDL}

The conversion of an AHIR network to VHDL is straightforward.
Currently, we produce a system which is synchronous and uses
a single clock (positive edges only).  Each symbol that is
exchanged between the control path and the data path is coded 
by a pulse which is sampled high by exactly one clock
edge.  Symbols are associated with transitions as indicated
earlier.

Places in the AHIR control path are handled as follows:
\begin{enumerate}
\item If a place has a single predecessor transition which is
an output transition and a single successor which is an input
transition, then it is optimized away.
\item If a place has a single successor transition which is
an output transition, and if this successor transition is
not a join, then the place is modeled by an OR gate whose inputs
are its predecessor transitions.
\item In the rest of the cases, the place is modeled as 
a finite state machine with two states, with the state being
set by any of its predecessors, and reset by any of its successors.
\end{enumerate}
The number of places that need to be modeled by state machines is
usually a small fraction of the total number of places.
Output transitions are modeled by AND gates, with each predecessor place
to the transition providing an input to the AND gate.  The cost
of implementing the control path is thus minimal in relation to
the implementation of the data path and memory subsystem.

Operators in the data path compute their results in a single
clock cycle, and the outputs are registered.  Shared operators
are constructed using multiplexers in a standard manner.
The instantiation of operators in the data path, 
and the interconnections within
the data path and between the data path and the control path
mirror those in the AHIR representation, except for the fact
that several operators in the AHIR data path can be mapped
to the same physical operator (as identified by arbiterless
operating sharing).
The inter module link layer is completed using the call graph
corresponding to the source code.

The memory sub-system is implemented as a multiple-bank (each
bank is a single-cycle synchronous SRAM)
pipe-lined memory which offers as many load/store ports as
there are load/store operators across the data paths.  The memory
sub-system is constructed using a request-complete protocol and
ensures that read/write operations complete in the order
that they were requested. The number of memory banks in the memory
subsystem,
and the degree of pipe-lining are parameters/generics that
can be selected based on the memory reference characteristics
of the application.

\section{Experimental Evaluation of the end-to-end synthesis flow}

We have implemented a complete C-to-ASIC tool-flow based on commercial
synthesis tools from Cadence and Synopsys.  We selected the
following set of applications for the experiments (the task used
to characterize the energy dissipation is also indicated).
\begin{enumerate}
\item A5/1, a simple stream cipher (basic task: produce one key bit).
\item AES, the standard AES-128 block cipher (basic task: encrypt a 128-bit block).
\item LPK, the standard Linpack benchmark (basic task: LU factorization of $100\times100$ matrix with
floating point entries).
\item FFT (basic task: 64 point floating point FFT).
\item RBT, the red-black tree data structure (basic task: insertion of $1000$ elements into
the tree).
\end{enumerate}

In each case, the VHDL generated by
the C-to-RTL flow was synthesized using Synposys Design Compiler,
using the OSU standard cell library \cite{OSU} for the TSMC $0.18\mu$ 
%
% vcag.ecen.okstate.edu/projects/scells/
%
technology.  
Memories were modeled using CACTI SRAM models \cite{CACTI}. 
The resulting net-list was mapped to layout using
the Cadence SOC Encounter tool.  No human intervention was
used in the entire process from C to the final layout (other
than specifying the operating frequency to the automatic place-and-route
tools).   The total time needed for the entire process (from C to layout)
was of the order of a few minutes for A5/1 to a few hours for the LPK
case.
%
% Premkishore Shivakumar, Norman P. Jouppi. Cacti 3.0 : An Integreted Cache Timing, Power, and Area model, Western Research
% Laboratory, 250 University Avenue Palo Alto, California 94301 USA, August 2001
%
A simulation based toggle file was then used with the extracted
net-list to estimate power dissipation in each of the generated
circuits.
The reciprocal of the energy used for completing a task is equivalent to
the throughput achieved for each watt of power supplied, commonly
termed as {\em performance per watt} 
%
% TODO: add a comment that the critical path was through 
% adders/multipliers which were unoptimized!

\begin{table}[htb]
  \centering
  \caption{Area/Delay/Power/Energy data for AHIR circuits in the TSMC $0.18\mu$ technology}
  \label{table:tsmc-180nm-data}
  \renewcommand\arraystretch{1.2}
  \setlength{\tabcolsep}{1ex}
  \begin{tabular}{c|c|c|c|c|c}
    \hline
    & Area & Freq & Delay & Power & Energy \\
    & (mm$^2$) & (MHz) & (ms) & (mW) & ($\mu$J) \\
    \hline
    \hline
    A5/1 & 1.45 & 71.4 & 0.28$\mu$s & 73 & 20.44\,nJ \\
    \hline
    AES & 6.5 & 71.4 & 0.428 & 338 & 144.7 \\
    \hline
    FFT & 5.1 & 41.7 & 0.155  & 95 & 14.7  \\
    \hline
    LPK & 27 & 41.7 & 37.7 & 273 & 10300  \\
    \hline
    RBT & 18 & 41.7 & 9.9 & 153 & 1511 \\
    \hline
  \end{tabular}

\end{table}

As a reference to judge the energy efficiency of
the AHIR generated circuits, we consider a 
popular low power processor (henceforth denoted
by $P$), namely the Intel Atom N270 processor \cite{ref:Atom},
%
% www.intel.com/products/processor/atom/index.htm
% 
which is built in Intel's $45nm$ technology.
This processor is readily available, combines high performance
with low power dissipation, has full floating point
support, and is a part of many low power
computing devices.   
The sample set of C programs was run on the processor $P$ (exactly
the same program used in the hardware generation was run on the processor).
In each case, the number of cycles needed
to finish one task (as described above) specific to that program was measured. 
The data
sets for the programs were small enough to fit into the processor
cache, and the cycle count was averaged across multiple runs.  These
cycle counts were then used to find the delay and energy consumption
for the processor.  All values for area, frequency and
power dissipation for the processor were obtained from the
corresponding data-sheet.

\begin{table}[htb]
  \centering
  \caption{Area/Delay/Power/Energy results for the processor $P$ built in a $45nm$ process}
  \label{table:atom-power-delay}
  \renewcommand\arraystretch{1.2}
  \setlength{\tabcolsep}{1ex}
  \begin{tabular}{c|c|c|c|c|c}
    \hline
    & Area & Freq & Delay & Power & Energy \\
    & (mm$^2$) & (MHz) & (ms) & (mW) & ($\mu$J) \\
    \hline
    \hline
    A5/1 & 25 & 1600 & 0.12$\mu$s & 2500 & 298.44\,nJ  \\
    \hline
    AES & 25 & 1600 & 0.036 & 2500 & 89.362 \\
    \hline
    FFT & 25 & 1600 & 0.022 & 2500 & 55.64  \\
    \hline
    LPK & 25 & 1600 & 7.90 & 2500 & 19740  \\
    \hline
    RBT & 25 & 1600 & 0.36 & 2500 & 891.89 \\
    \hline
  \end{tabular}

\end{table}

If we compare the results for the $180nm$ AHIR circuits with
the $45nm$ processor $P$, we find that except for the AES and
RBT cases,
the energy consumed per task by the AHIR circuit is appreciably
lower than that for the processor $P$.  This is significant
because the processor $P$ is implemented in a technology which
is four generations ahead of the AHIR circuits.

Now, suppose
that  we scale the 
energy numbers assuming migration of the AHIR circuits
from $180nm$ to the $45nm$ technology node, using 
the following  scaling rules (assume that $S = 45/180 = 0.25$):
\begin{itemize}
\item The operating voltage is scaled from $1.8V$ to $0.9V$.
\item The internal capacitances are assumed to be scaled by $S$.
\item The operating frequency of the generated circuit is scaled
by $1/S$.
\item Half of the total power dissipation of the $45nm$ circuits
is due to leakage \cite{ref:Frank}, and the other half is due to switching power. 
Note that this is a conservative estimate, and the actual leakage is
likely to be lower at the power levels corresponding to the
AHIR circuits.
%
% 
The leakage power is a negligible fraction of the
total power dissipation in the $180nm$ technology.
\item The area is scaled by $S^2$.
\end{itemize}
Thus, the per task energy consumption due to switching
activity alone will scale by $1/16$, and since
we are assuming that leakage power dissipation is half the total,
the total per task energy consumption of the circuit will 
scale by $1/8$.
The area of the circuits will scale by $1/16$.  
The power dissipation will scale by $1/2$ (this would have been
$1/4$ without leakage).
This scaling, though approximate, can be justified since the 
area of the resulting circuits (in each case) is small, so 
that interconnect effects are not likely to be pronounced.

\begin{table}[htb]
  \centering
  \caption{Area/Delay/Power/Energy RATIOS (processor $P$ values relative to the scaled AHIR values)}
  \label{table:Ratios}
  \renewcommand\arraystretch{1.2}
  \setlength{\tabcolsep}{1ex}
  \begin{tabular}{c|c|c|c|c|c}
    \hline
    & Area & Freq & Delay & Power & Energy \\
    \hline
    \hline
    A5/1 & 275.8 & 5.6 & 1.7 & 68.5 & 116.6  \\
    \hline
    AES & 61.5 & 5.6 & 0.34 & 14.8 & 4.9 \\
    \hline
    FFT & 78.4 & 9.6 & 0.57 & 52.6 & 30.3  \\
    \hline
    LPK & 14.8 & 9.6 & 0.84 & 18.3 & 15.3  \\
    \hline
    RBT & 22.2 & 9.6 & 0.15 & 32.7 & 4.7 \\
    \hline
  \end{tabular}

\end{table}

In Table \ref{table:Ratios}, we tabulate the ratios of
the performance parameters area, frequency, delay, power
and energy for the processor $P$ relative to the scaled values
for the AHIR circuits assuming the scaling rules shown above.
From the table, we see that across all applications, the energy consumed
by AHIR circuits is considerably lower than the
processor (the energy per task for the processor is
between $4.7X$ and $116.6X$ that of the corresponding AHIR circuit).  
The improvement is highest in the case of
bit-level manipulation applications such as A5/1, and
lowest in the case of random ``pointer-chasing'' code.
The delays in completing the tasks are lower 
for the processor $P$ in almost all the cases.  This
is mainly due to the fact that currently, the operators used in 
the AHIR generated RTL are not pipelined or optimized in any
manner during the logic synthesis and physical
design. This limits the operating clock frequency of the circuits
synthesized from the AHIR generated RTL.  
%
% Need more elaboration here
%

\section{Conclusion}

We have presented a high-level synthesis flow that converts
an algorithm described in C into a VHDL description.  The
synthesis flow uses a factored control-data-storage internal representation
and is correct by construction.  On this internal
representation, optimizations such as arbiter-less operator sharing 
can be carried out in an efficient manner.  The flow can thus
be applied to large programs describing a wide variety of algorithms.
Our experiments demonstrate that across this variety of
algorithms, an ASIC
implemented using the RTL generated by this high-level
synthesis flow uses a significantly lower energy per task (or equivalently,
has higher performance per watt) than industry-standard low-power microprocessors. 
Thus, the range of algorithms
which can be handled, and the ease of obtaining an implementation
which is likely to be more energy efficient than a processor 
indicate that this high-level synthesis flow provides an 
alternative to embedded processors for implementing complex algorithms in hardware.

%
% Need to revisit this
%

\bibliography{ref}
\bibliographystyle{IEEEtran}

\end{document}

% LocalWords:  Req Ack LL DP CP cmpgt br cdfg cp dp ir intra req init NCA SA mW
% LocalWords:  TPR STPR STPRs ccccc RTL ASIC petri ack versa liveness LRG LLVM
% LocalWords:  VHDL SRAM Synopsys AES LPK Linpack FFT RBT Synposys OSU TSMC nJ
% LocalWords:  nm

### 2010-05-24 Documents/Publications/type-2/type-2.tex
initial commit

\documentclass[12pt,a4paper]{article}
\usepackage{graphicx}
\usepackage{times}
\usepackage[margin=0.75in]{geometry}

\title{A class of live and safe petri-nets that supports scalable
  concurrency analysis}

\author{Sameer D. Sahasrabuddhe \\ Kavi Arya \\ Madhav P. Desai}

\date{}

\begin{document}

\maketitle

\begin{abstract}
  \large

  Petri-nets are a compact way of representing concurrent sequences of
  events in a system, where events are represented as transitions in
  the corresponding petri-net. Given two events in a system, it is
  useful to determine whether they can occur simultaneously. But it is
  difficult to make such a determination in an arbitrary petri-net. We
  present a class of petri-nets called ``Type-2'' which restricts the
  structure of the petri-net in order to enable a scalable analysis,
  without severely affecting the expressive power of the petri-net.
  Given a Type-2 petri-net, it is possible to determine in linear time
  whether two transitions can fire simultaneously. In addition, it is
  also possible to generate an exhaustive listing of sets of
  transitions such that no two transitions in each set can fire
  simultaneously.

  We first define the concept of a ``token-preserving region'' which
  produces a live and safe petri-net when augmented by a marked place.
  We further define a set of production rules to combine such regions
  into larger token-preserving regions. The petri-net created by
  augmenting such a region is said to be of the class ``Type-2''. The
  initial marking of the petri-net is restricted to a single marked
  place.

  The restriction on the structure of the Type-2 petri-net allow a
  simple labelling scheme to effectively simulates an execution of the
  petri-net. These labels can be compared in linear time to determine
  the concurrency of the labelled transitions. Despite the
  restrictions imposed by the production rules, the Type-2 class of
  petri-nets is powerful enough to represent synchronous systems. For
  example, this class has been used as a transition step in
  translating imperative programs to hardware descriptions.

\end{abstract}

\end{document}

### 2010-05-24 Documents/Reports/Kunal-Sreenivas/a5.tex
initial commit

\subsection{Reports}
The energy, delay and area estimates for each example are as follows:

\subsubsection{A5/1}
\\
The energy, delay and area estimates for each architecture / frequency combination is given in TABLE III.
In Fig. 2,  Fig. 3 we show the Power-Delay and Area-Delay plot for A5/1. 
\begin{table}[h!]
\caption{Energy / Delay / Area for each architecture/frequency combination}
\begin{center}
{\begin{tabular}{c | c   c   c   c    c   c}
\hline
A5/1 &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
\hline
Frequency (MHz) & 71.42& 71.42& 83.33& 71.42& 83.33& 71.42 \\ [1ex]
Energy (nJ) &25.26 &20.44 &29.484 &22.078 &34.56 &23.436 \\ [1ex]

Delay (ns)& 308& 280& 252& 266& 240& 252\\[1ex] 
Area (mm^2)& 1.45& 1.3& 1.77& 1.52& 2.3& 1.9\\[1ex]
\hline

\end{tabular}}
\label{diffstruc}
\end{center}
\end{table}

\begin{figure}[h!]
{\centering \resizebox*{6in}{4in}{\includegraphics{a5p.ps}} \par}
\caption{Power-Delay plot for A5/1}
\end{figure}

\begin{figure}[h!]
{\centering \resizebox*{6in}{4in}{\includegraphics{a5a.ps}} \par}
\caption{Area-Delay plot for A5/1}
\end{figure}

### 2010-05-24 Documents/Reports/Kunal-Sreenivas/abstract.tex
initial commit

\begin{abstract}
{
The goal of the research project was to characterize RTL generated by the AHIR flow. 
We have taken a set of examples, used AHIR to generate RTL, mapped the RTL to layout and finally performed power, area and delay estimation for the above examples.
For all our circuit examples we have used the OSU Standard Cell library for TSMC 0.18 technology and Synopsis  Design Compiler to perform synthesis.
The netlist thus obtained is used as an input to Cadence SOC Encounter to generate the layout.
The layout being closest to the actual hardware is most suiatble for estimation of power consumed in the design for a particular input sequence.
Five examples viz. A5, AES, Red Black trees, Linpack and FFT were selected, six memory architectures were explored for each and results are summarized in final section.
}
\end{abstract}

### 2010-05-24 Documents/Reports/Kunal-Sreenivas/arch_choice.tex
initial commit

\begin{table}[t]

\caption{Architecture choices for each example}
\begin{center}
{\begin{tabular}{c  c | c  c  c}
\hline
ARCHITECTURE && \multicolumn{3}{c}{MEMORY BANK CHOICE} \\  
CHOICES && 1& 2& 4 \\ [3ex] 
\hline
Example& Memory Size &&Base Bank Address Width \\ [1ex]
\hline 
A5& 16& 4& 4& 4\\[1ex]
LINPACK& 16K& 12& 12& 12\\[1ex]
R-B TREES& 16k& 12& 12& 12\\[1ex]
FFT& 512& 8& 8& 8\\[1ex]
AES& 1024& 8& 8& 8\\[1ex]
\hline

\end{tabular}}
\label{diffstruc}
\end{center}	
\end{table}

For each of memory architecture choices, deeply pipelined and non-pipelined was selected.
The range of frequencies tried for each architecture is shown in TABLE II:

\begin{table}[t]
\caption{Range of Frequencies for Each Architecture}
\begin{center}
{\begin{tabular}{c | c  c  c  c  c  c}
\hline
 & \multicolumn{6}{c}{ARCHITECTURE CHOICES} \\[1ex]
 &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
\hline
Example& \multicolumn{6}{c}{Frequency of Operation (in MHz)} \\ [1ex]
\hline
A5& 71.42& 71.42& 83.33& 71.42& 83.33& 71.43\\[1ex]
LINPACK& 45.45& 41.67& 41.67& 41.67& 41.67& 38.4\\[1ex]
R-B TREES& 41.67 and 62.5 & 41.67 and 55.56& 71.42& 50& 71.42& 38.36\\[1ex]
FFT& 41.66& 41.66& 41.66& 45.45& 45.45& 45.45\\[1ex]
AES& 45.45 and 71.42& 45.45 and 71.42& 71.42& 71.42& 71.42& 45.45\\[1ex]
\hline

\end{tabular}}
\\[1ex]
Note : Architecture choices are in the form [memory bank x deeply pipelined/non-pipelined degree]. For eg. 4 x 2 indicates an architecture with number of memory banks as 4. 2 and 0 indicates pipelined and non-pipelined memory, repectively.
\label{diffstruc}
\end{center}
\end{table}

### 2010-05-24 Documents/Reports/Kunal-Sreenivas/artex.tex
initial commit

\documentclass[a4paper, twoside]{article}      % Comments after  % are ignored
\usepackage{amsmath,amssymb,amsfonts} % Typical maths resource packages
\usepackage{graphics}                 % Packages to allow inclusion of graphics
\usepackage{color}                    % For creating coloured text and background
\usepackage{hyperref}                 % For creating hyperlinks in cross references

\oddsidemargin 0cm
\evensidemargin 0cm

\pagestyle{myheadings}         % Option to put page headers
                               % Needed \documentclass[a4paper,twoside]{article}

%\markboth{{\small\it PDF\ \LaTeX \ coloured text and graphics}}

%{{\small\it C.T.J. Dodson} }
\textwidth 15.5cm
\topmargin -1cm
\parindent 0cm
\textheight 24cm
\parskip 1mm

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{definition}[theorem]{Definition}

\def\R{\mathbb{ R}}

\def\S{\mathbb{ S}}

\date{\small\it April 20, 2009}

\title{\fcolorbox{white}{white}{\color{black}Power / Delay Estimation of Auto Generated Circuits}}

%\footnote{A demonstration example including colored text and graphics}}}

\author{ K. P. Ghosh, S.Subramanian \\
{\small Department of Electrical Engineering, IIT, Bombay, India}
 }

\begin{document}
\maketitle
\begin{abstract}
{
The goal of the research project was to characterize RTL generated by AHIR flow. We had taken a set of examples, used AHIR to generate RTL, mapped the RTL tolayout, and finally performed power, area and delay estimation for the above examples.  For all our circuit examples we have used  OSU Standard Cell library for TSMC 0.18 technology and Synopsis  Design Compiler to perform synthesis. The netlist thus obtained is used as an input to Cadence SOC Encounter  to generate the layout. The layout being closest to the actual hardware is most suiatble for estimation of power consumed in the design for a particular input sequence.Hence, we present an alternate approach of implementing memory in our circuits as Blackboxes. Cacti 3.2[1] was used to model the power and delay of the memory blackboxes. Five examples viz. A5, AES, Red Black trees, Linpack and FFT were selected, six memory architectures were explored for each and results are summarized in final section.
}
\end{abstract}

%{\color{black}{\sc keywords: AHIR, Memory Modelling, Cacti}}

\section{Introduction}
{AHIR provides a platform to describe hardware in a high level programming language like C and translates it to its equivalent hardware description. When we use this description of hardware, followed by the RTL synthesis and then its layout, we would like to characterise the hardware thus generated to meet certain physical and performance criteria. Area , power and delay are  important criteria that mark the quality of the hardware. The current approach followed by us to obtain layout from a VHDL description is as shown in the figure 1. Memories in the hardware have been treated as blackboxes with suitable power and area models obtained from Cacti 3.2.

\begin{figure}[ht]
{\centering \resizebox*{6in}{4in}{\includegraphics{list_of_figures1.ps}} \par}
\caption{Flow describing VHDL Description to Layout}
\end{figure}

At each level, a different tool is used for a specific purpose. At first,  the VHDL code  is simulated and checked for  its functionality using ModelSim 6.3a. The next step is  to perform synthesis and obtain a gate level netlist. For all our circuit examples we have used  OSU Standard Cell library for TSMC 0.18 technology and Synopsis  Design Compiler (DC) to perform synthesis. The netlist thus obtained is used as an input to Cadence SOC Encounter  to generate the layout. The layout being closest to the actual hardware is most suitable for estimation of powert consumed in the design for a particular input sequence.  Power has been calculated using Cadence SOC Encounter.

The memory could either be built up as a single bank or as a combination of multiple banks. The later would aid parallelism in the circuit. Also, the memory could be pipelined. We have tried to explore 6 different memory architectures for different examples. This would help us compare the performance of the circuits with multiple banks versus single banks, pipelined as opposed to non-pipelined memory. 
}
\section{Memory Modelling}
For encounter to estimate power, delay and area of the entire hardware, the memory black boxes must have the respective information. We use Cacti 3.2 to estimate power and delay in the memory black box. The area is modelled using  [2].

area(in mm2) = (0.001){ * }(tech)^{2.07} { * } (bits)^{0.9} { * } (port)^{0.7} + {0.0048}

The values of power, area and delay obtained from the above models are used to generate the necessary information for the blackboxes in timing library file (TLF) and library exchange file (LEF) formats.The vital timing information is also obtained from delay models. 
To describe a memory element as a black box, we instruct Synopsys DC to turn off its synthesis when it encounters the architecture for the memory.
Thus for synopsys, the memory is an entity definition with well defined input & output ports but a blank architecture.
Similarly, Cadence SOC encounter is made to infer black boxes whenever a memory component is instantiated from the TLF and LEF files.

\section{Power Estimation}
Using the final gate level verilog netlist provided by encounter we perform a post -layout simulation and dump all the signal information to a vcd file. This vcd file which now contains all the switching activity information is used by SOC Encounter to generate a power report. One is also required to mention the top level entity name for which power is to be estimated.
For examples like linpack, Redblack which required huge size of memories and also which took a long time to simulate, a sampling technique was followed, in order to reduce the size of VCD files (which was around 25Gb). In this technique, the final post layout netlist is simulated and the switching activities are captured at 10 random intervals of time. It was observed that, when power is calculated using these samples, the variance in power estimates for all the samples is in the orders of 1e-5.

\section{Results}
Five examples viz. A5, AES, Red Black trees, Linpack and FFT were selected. 
The memory architecture choices are shown table 1:

\begin{table}[t]

\caption{Architecture choices for each example}
\begin{center}
{\begin{tabular}{c  c | c  c  c}
\hline
ARCHITECTURE && \multicolumn{3}{c}{MEMORY BANK CHOICE} \\  
CHOICES && 1& 2& 4 \\ [3ex] 
\hline
Example& Memory Size &&Base Bank Address Width \\ [1ex]
\hline 
A5& 16& 4& 4& 4\\[1ex]
LINPACK& 16K& 12& 12& 12\\[1ex]
R-B TREES& 16k& 12& 12& 12\\[1ex]
FFT& 512& 8& 8& 8\\[1ex]
AES& 1024& 8& 8& 8\\[1ex]
\hline

\end{tabular}}
\label{diffstruc}
\end{center}	
\end{table}

For each of memory architecture choices, mux and demux degree of 2 and infinity was selected.
The range of frequencies tried for each architecture is shown in table 2:

\begin{table}[t]
\caption{Range of Frequencies for Each Architecture}
\begin{center}
{\begin{tabular}{c | c  c  c  c  c  c}
\hline
 & \multicolumn{6}{c}{ARCHITECTURE CHOICES} \\[1ex]
 &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
\hline
Example& \multicolumn{6}{c}{Frequency of Operation (in MHz)} \\ [1ex]
\hline
A5& 71.42& 71.42& 83.33& 71.42& 83.33& 71.43\\[1ex]
LINPACK& 45.45& 41.67& 41.67& 41.67& 41.67& 38.4\\[1ex]
R-B TREES& 41.67 and 62.5 & 41.67 and 55.56& 71.42& 50& 71.42& 38.36\\[1ex]
FFT& 41.66& 41.66& 41.66& 45.45& 45.45& 45.45\\[1ex]
AES& 45.45 and 71.42& 45.45 and 71.42& 71.42& 71.42& 71.42& 45.45\\[1ex]
\hline

\end{tabular}}
\\[1ex]
Note : Architecture choices are in the form [memory bank x mux/demux degree]. For eg. 4 x 2 indicates an architecture with number of memory banks as 4 and mux/demux degree of 2. 0 indicates a mux/demux degree of infinity.
\label{diffstruc}
\end{center}
\end{table}

\subsection{Reports}
The energy, delay and area estimates for each example with their respective 6 architectures are as follows:

\subsubsection{A5/1}
\\
The energy, delay and area estimates for each architecture / frequency combination is given in table 3.
Figure 2 and 3 shows the Power-Delay and Area-Delay plot for A5/1. 
\begin{table}[h!]
\caption{Energy / Delay / Area for each architecture/frequency combination}
\begin{center}
{\begin{tabular}{|c | c |  c |  c  | c  |  c |  c|}
\hline
A5/1 &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
\hline
Energy &&&&&& \\ [1ex]
\hline
Delay (ns)& 308& 280& 252& 266& 240& 252\\[1ex] \hline
Area (mm^2)& 1.45& 1.3& 1.77& 1.52& 2.3& 1.9\\[1ex]
\hline

\end{tabular}}
\label{diffstruc}
\end{center}
\end{table}

\begin{figure}[h!]
{\centering \resizebox*{6in}{4in}{\includegraphics{a5p.ps}} \par}
\caption{Power-Delay plot for A5/1}
\end{figure}

\begin{figure}[h!]
{\centering \resizebox*{6in}{4in}{\includegraphics{a5a.ps}} \par}
\caption{Area-Delay plot for A5/1}
\end{figure}

\subsubsection{LINPACK}
\\
The energy, delay and area estimates for each architecture / frequency combination is given in table 4.
Figure 4 and 5 shows the Power-Delay and Area-Delay plot for Linpack.

\begin{table}[h!]
\caption{Energy / Delay / Area for each architecture/frequency combination}
\begin{center}
{\begin{tabular}{|c | c |  c |  c  | c  |  c |  c|}
\hline
LINPACK &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
\hline
Energy &&&&&& \\ [1ex]
\hline
Delay (ms)& 56.48& 43.65& 55.63& 37.68& 51.64& 38.71\\[1ex] \hline
Area (mm^2)& 27.5& 25.5& 30.8& 27& 36.2& 28.9\\[1ex]
\hline

\end{tabular}}
\label{diffstruc}
\end{center}
\end{table}

\begin{figure}[h!]
{\centering \resizebox*{6in}{4in}{\includegraphics{lpp.ps}} \par}
\caption{Power-Delay plot for LINPACK}
\end{figure}

\begin{figure}[h!]
{\centering \resizebox*{6in}{4in}{\includegraphics{lpa.ps}} \par}
\caption{Area-Delay plot for LINPACK}
\end{figure}

\subsubsection{R-B TREES}
\\
The energy, delay and area estimates for each architecture / frequency combination is given in table 5.
Figure 6 and 7 shows the Power-Delay and Area-Delay plot for Red-Black Trees.

\begin{table}[h!]
\caption{Energy / Delay / Area for each architecture/frequency combination}
\begin{center}
{\begin{tabular}{|c | c |  c |  c  | c  |  c |  c|}
\hline
R-B TREES &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
\hline
Energy &&&&&& \\ [1ex]
\hline
Delay (ms)& 18.74 and 12.5& 9.88 and 7.41& 10.9& 8.17& 10.87& 10.57\\[1ex] \hline
Area (mm^2)& 19 and 23& 18 and 21& 26& 22& 32& 25\\[1ex]
\hline

\end{tabular}}
\label{diffstruc}
\end{center}
\end{table}

\begin{figure}[h!]
{\centering \resizebox*{6in}{4in}{\includegraphics{rbp.ps}} \par}
\caption{Power-Delay plot for RED-BLACK TREES}
\end{figure}

\begin{figure}[h!]
{\centering \resizebox*{6in}{4in}{\includegraphics{rba.ps}} \par}
\caption{Area-Delay plot for RED-BLACK TREES}
\end{figure}

\subsubsection{FFT}
\\
The energy, delay and area estimates for each architecture / frequency combination is given in table 6.
Figure 8 and 9 shows the Power-Delay and Area-Delay plot for FFT.

\begin{table}[h!]
\caption{Energy / Delay / Area for each architecture/frequency combination}
\begin{center}
{\begin{tabular}{|c | c |  c |  c  | c  |  c |  c|}
\hline
FFT &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
\hline
Energy &&&&&& \\ [1ex]
\hline
Delay (us)& 203.792& 154.88& 189.27& 141.606& 172.453& 140.55\\[1ex] \hline
Area (mm^2)& 5.4& 5.07 & 6.6& 5.7& 9.2& 6.7\\[1ex]
\hline

\end{tabular}}
\label{diffstruc}
\end{center}
\end{table}

\begin{figure}[h!]
{\centering \resizebox*{6in}{4in}{\includegraphics{fftp.ps}} \par}
\caption{Power-Delay plot for FFT}
\end{figure}

\begin{figure}[h!]
{\centering \resizebox*{6in}{4in}{\includegraphics{ffta.ps}} \par}
\caption{Area-Delay plot for FFT}
\end{figure}

\subsubsection{AES}
\\
The energy, delay and area estimates for each architecture / frequency combination is given in table 7.
Figure 10 and 11 shows the Power-Delay and Area-Delay plot for AES.

\begin{table}[h!]
\caption{Energy / Delay / Area for each architecture/frequency combination}
\begin{center}
{\begin{tabular}{|c | c |  c |  c  | c  |  c |  c|}
\hline
AES &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
\hline
Energy &&&&&& \\ [1ex]
\hline
Delay (ms)& 1.21 and 0.769& 0.67 and 0.43& 0.762& 0.418& 0.759& 0.655\\[1ex] \hline
Area (mm^2)& 7.67 and 9.03& 6.63 and 6.5 & 13& 7.5& 20.25& 9.4\\[1ex]
\hline

\end{tabular}}
\label{diffstruc}
\end{center}
\end{table}

\begin{figure}[h!]
{\centering \resizebox*{6in}{4in}{\includegraphics{aesp.ps}} \par}
\caption{Power-Delay plot for AES}
\end{figure}

\begin{figure}[h!]
{\centering \resizebox*{6in}{4in}{\includegraphics{aesa.ps}} \par}
\caption{Area-Delay plot for AES}
\end{figure}

\section{Conclusion}
From the results, tabulated earlier, we can conclude the following:
The circuits generated showed very limited parallelism in memory access patterns. Also, there is no performance boost in pipelined memory as opposed to the non-pipelined memory. We can observe that there's no significant improvement in delay as we increase the number of banks.
Also another trend that we observe is that circuits are more energy efficient at higher frequencies.

\newpage
\addcontentsline{toc}{chapter}{References}
\begin{thebibliography}{99}

\bibitem{} Premkishore Shivakumar, Norman P. Jouppi. Cacti 3.0 : An Integreted Cache Timing, Power, and Area model,
{\bf Western Research Laboratory}, 250 University Avenue Palo Alto, California 94301 USA.

\bibitem{} Banit Agrawal, Timothy Sherwood. Guiding Architectural SRAM models, 
{\bf Published in International Conference of Computer Design (ICCD) 2006},Department of Computer Science, University of California, Santa Barbara.

\end{thebibliography}

\end{document}

### 2010-05-24 Documents/Reports/Kunal-Sreenivas/artex_old.tex
initial commit

\documentclass[a4paper,twoside]{article}      % Comments after  % are ignored
\usepackage{amsmath,amssymb,amsfonts} % Typical maths resource packages
\usepackage{graphics}                 % Packages to allow inclusion of graphics
\usepackage{color}                    % For creating coloured text and background
\usepackage{hyperref}                 % For creating hyperlinks in cross references

\oddsidemargin 0cm
\evensidemargin 0cm

\pagestyle{myheadings}         % Option to put page headers
                               % Needed \documentclass[a4paper,twoside]{article}
\markboth{{\small\it PDF\ \LaTeX \ coloured text and graphics}}
{{\small\it C.T.J. Dodson} }

\textwidth 15.5cm
\topmargin -1cm
\parindent 0cm
\textheight 24cm
\parskip 1mm
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{definition}[theorem]{Definition}

\def\R{\mathbb{ R}}
\def\S{\mathbb{ S}}

\date{\small\it May 16, 2000}
\title{\fcolorbox{red}{blue}{\color{white}Including colour, pdf graphics and hyperlinks
\footnote{A demonstration example including colored text and graphics}}}

\author{ C.T.J. Dodson \\
{\small Department of Mathematics, UMIST, Manchester M60 1QD, UK}
 }

\begin{document}
\maketitle
\begin{abstract}
{\color{red}
You are unlikely to want your research articles to use coloured text
but this illustrates how it can be done}
\end{abstract}

{\color{green}{\sc keywords: dpa, zero-knowledge, information theory,
distributions, metric}}

\section{Introduction}
Zero-Knowledge proofs allow verification of
secret-based actions without revealing the secrets. Goldreich et
al.~\cite{goldreich} discussed the class of promise problems in
which interaction may give additional information in the context
of Statistical Zero-Knowlege. They invoked two types of difference
between distributions: the `statistical difference' and the
`entropy difference' of two random variables. In this context,
typically, one of the distributions is the uniform distribution.

Thus, in the contexts of DPA and SZK tests, it is necessary to
compare two nearby distributions on bounded domains.
We describe the following result and discuss applications.
{\color{cyan}
\begin{proposition}\label{colour}
The family of probability density functions for random variable
$N\in [0,1]$ given by
\begin{equation}
g(N,\mu,\beta) = {\frac{{{{\frac{1}{N}}}^
       {1 - {\frac{\beta }{\mu }}}}\,
     ({{{\frac{\beta }{\mu }}})^{\beta }}\,
     {{(\log {\frac{1}{N}})}^
       { \beta -1}}}{
      \Gamma(\beta )}} \ \ \ {\rm for} \  \mu \  > 0 \ {\rm and} \ \beta \geq 1
  \label{loggamma}    \end{equation}
determines a metric space of distributions with the following
properties\\ $\bullet$ it contains the uniform distribution\\
$\bullet$ it contains approximations to truncated Gaussian
distributions\\ $\bullet$ the difference structure is given by the
information-theoretic metric\\ $\bullet$ as a Riemannian
2-manifold it is an isometric isomorph of the manifold of gamma
distributions.
\end{proposition} }

\section{In this example}
Examples are provided of
\begin{enumerate}
\item Coloured text~\pageref{colour}
\item Graphics~\pageref{graphics}
\item Tables~\pageref{table}
\end{enumerate}

\section{Proof of Proposition 1.1}
\subsection{Log-gamma PDFs}
By integration, it is easily checked that the family given by
equation (\ref{loggamma}) consists of probability density
functions for the random variable $N\in [0,1];$ some with central
mean are shown in Figure~\ref{pdf}. The limiting densities are
given by
\begin{eqnarray}
\lim_{\beta\rightarrow 1^+}g(N,\mu,\beta) =
g(N,\mu,1)&=&\frac{1}{\mu}\left(\frac{1}{N}\right)^
      {1 - \frac{1}{\mu }} \\
     \lim_{\mu\rightarrow 1}g(N,\mu,1) = g(N,1,1) &=& 1 \ .
\end{eqnarray}

\subsection{Information metric structure} For the log-gamma
densities, the Fisher information matrix determines a Riemannian
information metric~\cite{amari} on the parameter space ${\cal
S}=\{(\mu,\beta)\in (0,\infty)\times[1,\infty)\}.$ Its arc length
function is given by
\begin{equation}
ds_{\cal{S}}^2=\sum_{ij} g_{ij} \, dx^i dx^j =\frac{\beta}{\mu^2}
\, d\mu^2 +
        \left(\psi'(\beta) -\frac{1}{\beta}\right)\, d\beta^2 , \label{gammametric}
\end{equation}
where $\psi(\beta)=\frac{\Gamma'(\beta)}{\Gamma(\beta)}$ is the
logarithmic derivative of the gamma function, evaluated at
$\beta.$

In fact, (\ref{loggamma}) arises from the gamma family
\begin{equation}
f(x,\mu,\beta) = \frac{ x^{\beta -1} \,
     (\frac{\beta }{\mu})^\beta    }
     {\Gamma(\beta )} \,  e^{-\frac{V\,\beta }{\mu }}
     \label{gamma}
\end{equation}
for the non-negative random variable $x=\log\frac{1}{N}.$ It is
known that the gamma family (\ref{gamma}) has also the information
metric (\ref{gammametric}) (cf \cite{lauritzen}) so the identity
map on the space of coordinates $(\mu,\beta)$ is an isometry of
Riemannian manifolds. \hfill $\Box$

\section{Tables}\label{table}
Table~\ref{diffstruc} lists the number of differentiable
structures on spheres.
\begin{table}
\begin{center}
\framebox[1.5in]{\begin{tabular}{c || c | c }
$n$ & $\S^n$ & $\R^n$ \\ \hline
1 & 1 & 1 \\
2 & 1 & 1 \\
3 & 1 & 1 \\
4 & 1 & $\infty$ \\
5 & 1 & 1 \\
6 & 1 & 1 \\
7 & 28 & 1 \\
8 & 2 & 1 \\
9 & 8 & 1 \\
10 & 6 & 1 \\
11 & 992 & 1 \\
12 & 1 & 1 \\
13 & 3 & 1 \\
14 & 2 & 1 \\
15 & 16256 & 1 \\
\end{tabular}}
\caption{Numbers of distinct differentiable structures on real $n$-space
and $n$-spheres}
\label{diffstruc}
\end{center}
\end{table}

Here is how to set out a table in \LaTeX:
\begin{verbatim}
\begin{table}
\begin{center}
\framebox[1.5in]{\begin{tabular}{c | c | c }
$n$ & $\S^n$ & $\R^n$ \\ \hline
1 & 1 & 1 \\
2 & 1 & 1 \\
3 & 1 & 1 \\
4 & 1 & $\infty$ \\
5 & 1 & 1 \\
6 & 1 & 1 \\
7 & 28 & 1 \\
8 & 2 & 1 \\
9 & 8 & 1 \\
10 & 6 & 1 \\
11 & 992 & 1 \\
12 & 1 & 1 \\
13 & 3 & 1 \\
14 & 2 & 1 \\
15 & 16256 & 1 \\
\end{tabular}}
\caption{Numbers of distinct differentiable structures on real $n$-space
and $n$-spheres}
\label{diffstruc}
\end{center}
\end{table}
\end{verbatim}
This allowed us to cross-reference the Table via:
\begin{verbatim}
~\ref{diffstruc}
\end{verbatim}

\section{Graphics}\label{graphics}
\begin{figure}
\begin{picture}(300,220)(0,0)
\put(-20,20){\resizebox{20 cm}{!}{\includegraphics{3dpdf}}}
\put(260,30){\resizebox{15 cm}{!}{\includegraphics{contpdf}}}
\put(210,100){$\beta$}
\put(400,25){$N$}
\put(260,200){$\beta$}
\put(90,40){$N$}
\end{picture}
\caption{{\em The log-gamma family of densities with central mean
$<N> \, = \frac{1}{2}$ as a surface and as a contour plot. }}
\label{pdf}
\end{figure}

Figure~\ref{pdf} was created by calling in the two pdf graphics files
{\tt 3dpdf.pdf} , {\tt contpdf.pdf}
placed together in the following {\tt picture} environment
inside a {\tt figure} environment with a caption and label:
\begin{verbatim}
\begin{figure}
\begin{picture}(300,220)(0,0)
\put(-20,20){\resizebox{20 cm}{!}{\includegraphics{3dpdf}}}
\put(260,30){\resizebox{15 cm}{!}{\includegraphics{contpdf}}}
\put(210,100){$\beta$}
\put(400,25){$N$}
\put(260,200){$\beta$}
\put(90,40){$N$}
\end{picture}
\caption{{\em The log-gamma family of densities with central mean
$<N> \, = \frac{1}{2}$ as a surface and as a contour plot. }}
\label{pdf}
\end{figure}
\end{verbatim}

\begin{thebibliography}{99}
\bibitem{amari} S-I. Amari. {\bf Differential Geometrical Methods in
Statistics}, Springer Lecture Notes in Statistics 28,
Springer-Verlag, Berlin 1985.

\bibitem{chari} S. Chari, C.S. Jutla, J.R. Rao and P. Rohatgi.
Towards sound approaches to counteract power-analysis attacks. In
{\bf Advances in Cryptology-CRYPTO '99}, Ed. M. Wiener, Lecture
Notes in Computer Science 1666, Springer, Berlin 1999 pp 398-412.

\bibitem{crescenzo} G. Di Crescenzo and R. Ostrovsky.
On concurrent zero-knowledge with pre-processing. In {\bf Advances
in Cryptology-CRYPTO '99}  Ed. M. Wiener, Lecture Notes in
Computer Science 1666, Springer, Berlin 1999 pp 485-502.

\bibitem{tg} C.T.J. Dodson and T. Poston. {\bf Tensor Geometry}, Graduate
Texts in Mathematics 130, Second edition, Springer-Verlag, New
York, 1991. \htmladdnormallink{http://www.ma.umist.ac.uk/kd/tg.html}
{http://www.ma.umist.ac.uk/kd/tg.html}

\bibitem{goldreich} O. Goldreich, A. Sahai and S. Vadham.
Can Statistical Zero-Knowledge be made non-interactive? Or, on the
relationship of SZK and NISZK. In {\bf Advances in
Cryptology-CRYPTO '99},  Ed. M. Wiener, Lecture Notes in Computer
Science 1666, Springer, Berlin 1999 pp 467-484.

\bibitem{kocher} P. Kocher, J. Jaffe and B.Jun.
Differential Power Analysis. In {\bf Advances in Cryptology-CRYPTO
'99},  Ed. M. Wiener, Lecture Notes in Computer Science 1666,
Springer, Berlin 1999 pp 388-397.

\bibitem{lauritzen} S.L. Lauritzen. Statistical Manifolds. In {\bf
Differential Geometry in Statistical Inference}, Institute of
Mathematical Statistics Lecture Notes, Volume 10, Berkeley 1987,
pp 163-218.

\end{thebibliography}
\end{document}

### 2010-05-24 Documents/Reports/Kunal-Sreenivas/conclusion.tex
initial commit

\clearpage

\section{Conclusion}

From the results, tabulated and plotted earlier, we can conclude the following:
\begin{enumerate}
\item The nonpipelined memory looks more efficient in power-delay and area-delay, when compared to the deeply pipelined case. 
There's no significant improvement in delays as we increase the number of banks except A5/1.
This indicates that there is limited parallelism in the memory accesses.
\item The circuits are usually more energy efficient at higher frequency of operation.
\item The floating point operators limit the higher operating frequency in the case of FFT and Linpack.
\item Energy per operation for FFT and Linpack is in the order of 10 nJ and 12 nJ respectively. 
\end{enumerate}

### 2010-05-24 Documents/Reports/Kunal-Sreenivas/introduction.tex
initial commit

--- 2010-05-24 Documents/Reports/Kunal-Sreenivas/artex.tex
@@
-\documentclass[a4paper, twoside]{article}      % Comments after  % are ignored
-\usepackage{amsmath,amssymb,amsfonts} % Typical maths resource packages
-\usepackage{graphics}                 % Packages to allow inclusion of graphics
-\usepackage{color}                    % For creating coloured text and background
-\usepackage{hyperref}                 % For creating hyperlinks in cross references
-
-\oddsidemargin 0cm
-\evensidemargin 0cm
-
-\pagestyle{myheadings}         % Option to put page headers
-                               % Needed \documentclass[a4paper,twoside]{article}
-
-%\markboth{{\small\it PDF\ \LaTeX \ coloured text and graphics}}
-
-%{{\small\it C.T.J. Dodson} }
-\textwidth 15.5cm
-\topmargin -1cm
-\parindent 0cm
-\textheight 24cm
-\parskip 1mm
-
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{proposition}[theorem]{Proposition}
-\newtheorem{corollary}[theorem]{Corollary}
-\newtheorem{lemma}[theorem]{Lemma}
-\newtheorem{remark}[theorem]{Remark}
-\newtheorem{definition}[theorem]{Definition}
-
-\def\R{\mathbb{ R}}
-
-\def\S{\mathbb{ S}}
-
-\date{\small\it April 20, 2009}
-
-\title{\fcolorbox{white}{white}{\color{black}Power / Delay Estimation of Auto Generated Circuits}}
-
-%\footnote{A demonstration example including colored text and graphics}}}
-
-\author{ K. P. Ghosh, S.Subramanian \\
-{\small Department of Electrical Engineering, IIT, Bombay, India}
- }
-
-\begin{document}
-\maketitle
-\begin{abstract}
-{
-The goal of the research project was to characterize RTL generated by AHIR flow.
-We had taken a set of examples, used AHIR to generate RTL, mapped the RTL tolayout, and finally performed power, area and delay estimation for the above examples.
-For all our circuit examples we have used  OSU Standard Cell library for TSMC 0.18 technology and Synopsis  Design Compiler to perform synthesis.
-The netlist thus obtained is used as an input to Cadence SOC Encounter  to generate the layout.
-The layout being closest to the actual hardware is most suiatble for estimation of power consumed in the design for a particular input sequence.Hence, we present an alternate approach of implementing memory in our circuits as Blackboxes.
-Cacti 3.2[1] was used to model the power and delay of the memory blackboxes.
-Five examples viz.
-A5, AES, Red Black trees, Linpack and FFT were selected, six memory architectures were explored for each and results are summarized in final section.
-}
-\end{abstract}
-
-%{\color{black}{\sc keywords: AHIR, Memory Modelling, Cacti}}
-
 \section{Introduction}
-{AHIR provides a platform to describe hardware in a high level programming language like C and translates it to its equivalent hardware description.
-When we use this description of hardware, followed by the RTL synthesis and then its layout, we would like to characterise the hardware thus generated to meet certain physical and performance criteria.
-Area , power and delay are  important criteria that mark the quality of the hardware.
-The current approach followed by us to obtain layout from a VHDL description is as shown in the figure 1.
-Memories in the hardware have been treated as blackboxes with suitable power and area models obtained from Cacti 3.2.
+AHIR provides a platform to translate an algorithm described in a high level programming language like C to a VHDL circuit description.
+In this work, we characterise the resulting hardware in terms of physical and performance parameters such as area, power and delay.
+Our approach followed for obtaining layout from a VHDL description is as shown in the Fig.
+1.
 
@@
 The next step is  to perform synthesis and obtain a gate level netlist.
-For all our circuit examples we have used  OSU Standard Cell library for TSMC 0.18 technology and Synopsis  Design Compiler (DC) to perform synthesis.
+For all our circuit examples we have used the OSU Standard Cell library for TSMC 0.18 technology and Synopsis  Design Compiler (DC) to perform synthesis.
 The netlist thus obtained is used as an input to Cadence SOC Encounter to generate the layout.
-The layout being closest to the actual hardware is most suitable for estimation of powert consumed in the design for a particular input sequence.
-Power has been calculated using Cadence SOC Encounter.
 
-The memory could either be built up as a single bank or as a combination of multiple banks.
+The memory is designed as single bank or as a combination of multiple banks.
 The later would aid parallelism in the circuit.
 Also, the memory could be pipelined.
-We have tried to explore 6 different memory architectures for different examples.
-This would help us compare the performance of the circuits with multiple banks versus single banks, pipelined as opposed to non-pipelined memory.
-}
-\section{Memory Modelling}
-For encounter to estimate power, delay and area of the entire hardware, the memory black boxes must have the respective information.
-We use Cacti 3.2 to estimate power and delay in the memory black box.
-The area is modelled using  [2].
-
-area(in mm2) = (0.001){ * }(tech)^{2.07} { * } (bits)^{0.9} { * } (port)^{0.7} + {0.0048}
-
-The values of power, area and delay obtained from the above models are used to generate the necessary information for the blackboxes in timing library file (TLF) and library exchange file (LEF) formats.The vital timing information is also obtained from delay models.
-To describe a memory element as a black box, we instruct Synopsys DC to turn off its synthesis when it encounters the architecture for the memory.
-Thus for synopsys, the memory is an entity definition with well defined input & output ports but a blank architecture.
-Similarly, Cadence SOC encounter is made to infer black boxes whenever a memory component is instantiated from the TLF and LEF files.
-
-\section{Power Estimation}
-Using the final gate level verilog netlist provided by encounter we perform a post -layout simulation and dump all the signal information to a vcd file.
-This vcd file which now contains all the switching activity information is used by SOC Encounter to generate a power report.
-One is also required to mention the top level entity name for which power is to be estimated.
-For examples like linpack, Redblack which required huge size of memories and also which took a long time to simulate, a sampling technique was followed, in order to reduce the size of VCD files (which was around 25Gb).
-In this technique, the final post layout netlist is simulated and the switching activities are captured at 10 random intervals of time.
-It was observed that, when power is calculated using these samples, the variance in power estimates for all the samples is in the orders of 1e-5.
-
-\section{Results}
-Five examples viz. A5, AES, Red Black trees, Linpack and FFT were selected. 
-The memory architecture choices are shown table 1:
-
-\begin{table}[t]
-
-\caption{Architecture choices for each example}
-\begin{center}
-{\begin{tabular}{c  c | c  c  c}
-\hline
-ARCHITECTURE && \multicolumn{3}{c}{MEMORY BANK CHOICE} \\  
-CHOICES && 1& 2& 4 \\ [3ex] 
-\hline
-Example& Memory Size &&Base Bank Address Width \\ [1ex]
-\hline 
-A5& 16& 4& 4& 4\\[1ex]
-LINPACK& 16K& 12& 12& 12\\[1ex]
-R-B TREES& 16k& 12& 12& 12\\[1ex]
-FFT& 512& 8& 8& 8\\[1ex]
-AES& 1024& 8& 8& 8\\[1ex]
-\hline
-
-\end{tabular}}
-\label{diffstruc}
-\end{center}	
-\end{table}
-
-For each of memory architecture choices, mux and demux degree of 2 and infinity was selected.
-The range of frequencies tried for each architecture is shown in table 2:
-
-\begin{table}[t]
-\caption{Range of Frequencies for Each Architecture}
-\begin{center}
-{\begin{tabular}{c | c  c  c  c  c  c}
-\hline
- & \multicolumn{6}{c}{ARCHITECTURE CHOICES} \\[1ex]
- &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
-\hline
-Example& \multicolumn{6}{c}{Frequency of Operation (in MHz)} \\ [1ex]
-\hline
-A5& 71.42& 71.42& 83.33& 71.42& 83.33& 71.43\\[1ex]
-LINPACK& 45.45& 41.67& 41.67& 41.67& 41.67& 38.4\\[1ex]
-R-B TREES& 41.67 and 62.5 & 41.67 and 55.56& 71.42& 50& 71.42& 38.36\\[1ex]
-FFT& 41.66& 41.66& 41.66& 45.45& 45.45& 45.45\\[1ex]
-AES& 45.45 and 71.42& 45.45 and 71.42& 71.42& 71.42& 71.42& 45.45\\[1ex]
-\hline
-
-\end{tabular}}
-\\[1ex]
-Note : Architecture choices are in the form [memory bank x mux/demux degree].
-For eg.
-4 x 2 indicates an architecture with number of memory banks as 4 and mux/demux degree of 2.
-0 indicates a mux/demux degree of infinity.
-\label{diffstruc}
-\end{center}
-\end{table}
-
-\subsection{Reports}
-The energy, delay and area estimates for each example with their respective 6 architectures are as follows:
-
-\subsubsection{A5/1}
-\\
-The energy, delay and area estimates for each architecture / frequency combination is given in table 3.
-Figure 2 and 3 shows the Power-Delay and Area-Delay plot for A5/1. 
-\begin{table}[h!]
-\caption{Energy / Delay / Area for each architecture/frequency combination}
-\begin{center}
-{\begin{tabular}{|c | c |  c |  c  | c  |  c |  c|}
-\hline
-A5/1 &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
-\hline
-Energy &&&&&& \\ [1ex]
-\hline
-Delay (ns)& 308& 280& 252& 266& 240& 252\\[1ex] \hline
-Area (mm^2)& 1.45& 1.3& 1.77& 1.52& 2.3& 1.9\\[1ex]
-\hline
-
-\end{tabular}}
-\label{diffstruc}
-\end{center}
-\end{table}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{a5p.ps}} \par}
-\caption{Power-Delay plot for A5/1}
-\end{figure}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{a5a.ps}} \par}
-\caption{Area-Delay plot for A5/1}
-\end{figure}
-
-\subsubsection{LINPACK}
-\\
-The energy, delay and area estimates for each architecture / frequency combination is given in table 4.
-Figure 4 and 5 shows the Power-Delay and Area-Delay plot for Linpack.
-
-\begin{table}[h!]
-\caption{Energy / Delay / Area for each architecture/frequency combination}
-\begin{center}
-{\begin{tabular}{|c | c |  c |  c  | c  |  c |  c|}
-\hline
-LINPACK &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
-\hline
-Energy &&&&&& \\ [1ex]
-\hline
-Delay (ms)& 56.48& 43.65& 55.63& 37.68& 51.64& 38.71\\[1ex] \hline
-Area (mm^2)& 27.5& 25.5& 30.8& 27& 36.2& 28.9\\[1ex]
-\hline
-
-\end{tabular}}
-\label{diffstruc}
-\end{center}
-\end{table}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{lpp.ps}} \par}
-\caption{Power-Delay plot for LINPACK}
-\end{figure}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{lpa.ps}} \par}
-\caption{Area-Delay plot for LINPACK}
-\end{figure}
-
-\subsubsection{R-B TREES}
-\\
-The energy, delay and area estimates for each architecture / frequency combination is given in table 5.
-Figure 6 and 7 shows the Power-Delay and Area-Delay plot for Red-Black Trees.
-
-\begin{table}[h!]
-\caption{Energy / Delay / Area for each architecture/frequency combination}
-\begin{center}
-{\begin{tabular}{|c | c |  c |  c  | c  |  c |  c|}
-\hline
-R-B TREES &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
-\hline
-Energy &&&&&& \\ [1ex]
-\hline
-Delay (ms)& 18.74 and 12.5& 9.88 and 7.41& 10.9& 8.17& 10.87& 10.57\\[1ex] \hline
-Area (mm^2)& 19 and 23& 18 and 21& 26& 22& 32& 25\\[1ex]
-\hline
-
-\end{tabular}}
-\label{diffstruc}
-\end{center}
-\end{table}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{rbp.ps}} \par}
-\caption{Power-Delay plot for RED-BLACK TREES}
-\end{figure}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{rba.ps}} \par}
-\caption{Area-Delay plot for RED-BLACK TREES}
-\end{figure}
-
-\subsubsection{FFT}
-\\
-The energy, delay and area estimates for each architecture / frequency combination is given in table 6.
-Figure 8 and 9 shows the Power-Delay and Area-Delay plot for FFT.
-
-\begin{table}[h!]
-\caption{Energy / Delay / Area for each architecture/frequency combination}
-\begin{center}
-{\begin{tabular}{|c | c |  c |  c  | c  |  c |  c|}
-\hline
-FFT &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
-\hline
-Energy &&&&&& \\ [1ex]
-\hline
-Delay (us)& 203.792& 154.88& 189.27& 141.606& 172.453& 140.55\\[1ex] \hline
-Area (mm^2)& 5.4& 5.07 & 6.6& 5.7& 9.2& 6.7\\[1ex]
-\hline
-
-\end{tabular}}
-\label{diffstruc}
-\end{center}
-\end{table}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{fftp.ps}} \par}
-\caption{Power-Delay plot for FFT}
-\end{figure}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{ffta.ps}} \par}
-\caption{Area-Delay plot for FFT}
-\end{figure}
-
-\subsubsection{AES}
-\\
-The energy, delay and area estimates for each architecture / frequency combination is given in table 7.
-Figure 10 and 11 shows the Power-Delay and Area-Delay plot for AES.
-
-\begin{table}[h!]
-\caption{Energy / Delay / Area for each architecture/frequency combination}
-\begin{center}
-{\begin{tabular}{|c | c |  c |  c  | c  |  c |  c|}
-\hline
-AES &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
-\hline
-Energy &&&&&& \\ [1ex]
-\hline
-Delay (ms)& 1.21 and 0.769& 0.67 and 0.43& 0.762& 0.418& 0.759& 0.655\\[1ex] \hline
-Area (mm^2)& 7.67 and 9.03& 6.63 and 6.5 & 13& 7.5& 20.25& 9.4\\[1ex]
-\hline
-
-\end{tabular}}
-\label{diffstruc}
-\end{center}
-\end{table}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{aesp.ps}} \par}
-\caption{Power-Delay plot for AES}
-\end{figure}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{aesa.ps}} \par}
-\caption{Area-Delay plot for AES}
-\end{figure}
-
-\section{Conclusion}
-From the results, tabulated earlier, we can conclude the following:
-The circuits generated showed very limited parallelism in memory access patterns.
-Also, there is no performance boost in pipelined memory as opposed to the non-pipelined memory.
-We can observe that there's no significant improvement in delay as we increase the number of banks.
-Also another trend that we observe is that circuits are more energy efficient at higher frequencies.
-
-\newpage
-\addcontentsline{toc}{chapter}{References}
-\begin{thebibliography}{99}
-
-\bibitem{} Premkishore Shivakumar, Norman P. Jouppi. Cacti 3.0 : An Integreted Cache Timing, Power, and Area model,
-{\bf Western Research Laboratory}, 250 University Avenue Palo Alto, California 94301 USA.
-
-\bibitem{} Banit Agrawal, Timothy Sherwood. Guiding Architectural SRAM models, 
-{\bf Published in International Conference of Computer Design (ICCD) 2006},Department of Computer Science, University of California, Santa Barbara.
-
-\end{thebibliography}
-
-\end{document}
+We have explored 6 different memory architectures for different examples and compared the performance of the circuits with multiple banks versus single banks, pipelined as opposed to non-pipelined memory.

### 2010-05-24 Documents/Reports/Kunal-Sreenivas/memory.tex
initial commit

\section{Memory Modelling}
Memories has been treated as blackboxes due to the absence of SRAM cells in the OSU standard cell library.
For SOC Encounter to estimate power, delay and area of the entire hardware, the memory black boxes must have the required information. We use Cacti 3.2[1] to estimate power and delay in the memory blackboxes. The area is modelled using the following expression[2].

area(in mm2) = (0.001){ * }(tech)^{2.07} { * } (bits)^{0.9} { * } (port)^{0.7} + {0.0048}

The values of power, area and delay obtained from the above models are used to generate the necessary information for the blackboxes in timing library file (TLF) and library exchange file (LEF) formats. The vital timing information is also obtained from delay models.

To describe a memory element as a black box, we instruct Synopsys DC to turn off its synthesis when it encounters the architecture for the memory.
Thus for Synopsys DC, the memory is an entity definition with well defined input and output ports but a blank architecture.
Similarly, Cadence SOC Encounter is made to infer black boxes whenever a memory component is instantiated from the TLF and LEF files.

\section{Power Estimation}
The layout being closest to the actual hardware is most suitable for estimation of power consumed in the design for a particular input sequence.
Power has been calculated using SOC Encounter.

Using the final gate level verilog netlist provided by SOC Encounter we perform a post -layout simulation and dump all the signal information to a vcd file. This vcd file is used by SOC Encounter to generate a power report. One is also required to mention the top level entity name for which power is to be estimated.

Circuits like Linpack, Redblack have larger memory requirement and also take long time to simulate. For such examples, a sampling technique was followed, in order to reduce the size of VCD files (which was around 25Gb). In this technique, the final post layout netlist is simulated and the switching activities are captured at 10 random intervals of time. It was observed that, when power is calculated using these samples, the variance in power estimates for the samples is negligible.

\section{Results}
Five examples viz. A5, AES, Red Black trees, Linpack and FFT were selected. 
The memory architecture choices are shown TABLE I:

### 2010-05-24 Documents/Reports/Kunal-Sreenivas/old_tite.tex
initial commit

--- 2010-05-24 Documents/Reports/Kunal-Sreenivas/artex.tex
@@
-\documentclass[a4paper, twoside]{article}      % Comments after  % are ignored
+\documentclass[12pt]{report}      % Comments after  % are ignored
+
 \usepackage{amsmath,amssymb,amsfonts} % Typical maths resource packages
+
 \usepackage{graphics}                 % Packages to allow inclusion of graphics
+
 \usepackage{color}                    % For creating coloured text and background
+
 \usepackage{hyperref}                 % For creating hyperlinks in cross references
@@
 \oddsidemargin 0cm
+
 \evensidemargin 0cm
@@
 \pagestyle{myheadings}         % Option to put page headers
+
                                % Needed \documentclass[a4paper,twoside]{article}
 
-%\markboth{{\small\it PDF\ \LaTeX \ coloured text and graphics}}
+\markboth{{\small\it PDF\ \LaTeX \ coloured text and graphics}}
 
-%{{\small\it C.T.J. Dodson} }
+{{\small\it C.T.J. Dodson} }
+
 \textwidth 15.5cm
+
 \topmargin -1cm
+
 \parindent 0cm
+
 \textheight 24cm
+
 \parskip 1mm
@@
 \newtheorem{theorem}{Theorem}[section]
+
 \newtheorem{proposition}[theorem]{Proposition}
+
 \newtheorem{corollary}[theorem]{Corollary}
+
 \newtheorem{lemma}[theorem]{Lemma}
+
 \newtheorem{remark}[theorem]{Remark}
+
 \newtheorem{definition}[theorem]{Definition}
@@
 
-\date{\small\it April 20, 2009}
+\date{\small\it May 16, 2000}
 
@@
 \author{ K. P. Ghosh, S.Subramanian \\
+
 {\small Department of Electrical Engineering, IIT, Bombay, India}
+
  }
-
-\begin{document}
-\maketitle
-\begin{abstract}
-{
-The goal of the research project was to characterize RTL generated by AHIR flow.
-We had taken a set of examples, used AHIR to generate RTL, mapped the RTL tolayout, and finally performed power, area and delay estimation for the above examples.
-For all our circuit examples we have used  OSU Standard Cell library for TSMC 0.18 technology and Synopsis  Design Compiler to perform synthesis.
-The netlist thus obtained is used as an input to Cadence SOC Encounter  to generate the layout.
-The layout being closest to the actual hardware is most suiatble for estimation of power consumed in the design for a particular input sequence.Hence, we present an alternate approach of implementing memory in our circuits as Blackboxes.
-Cacti 3.2[1] was used to model the power and delay of the memory blackboxes.
-Five examples viz.
-A5, AES, Red Black trees, Linpack and FFT were selected, six memory architectures were explored for each and results are summarized in final section.
-}
-\end{abstract}
-
-%{\color{black}{\sc keywords: AHIR, Memory Modelling, Cacti}}
-
-\section{Introduction}
-{AHIR provides a platform to describe hardware in a high level programming language like C and translates it to its equivalent hardware description.
-When we use this description of hardware, followed by the RTL synthesis and then its layout, we would like to characterise the hardware thus generated to meet certain physical and performance criteria.
-Area , power and delay are  important criteria that mark the quality of the hardware.
-The current approach followed by us to obtain layout from a VHDL description is as shown in the figure 1.
-Memories in the hardware have been treated as blackboxes with suitable power and area models obtained from Cacti 3.2.
-
-\begin{figure}[ht]
-{\centering \resizebox*{6in}{4in}{\includegraphics{list_of_figures1.ps}} \par}
-\caption{Flow describing VHDL Description to Layout}
-\end{figure}
-
-At each level, a different tool is used for a specific purpose.
-At first,  the VHDL code  is simulated and checked for  its functionality using ModelSim 6.3a.
-The next step is  to perform synthesis and obtain a gate level netlist.
-For all our circuit examples we have used  OSU Standard Cell library for TSMC 0.18 technology and Synopsis  Design Compiler (DC) to perform synthesis.
-The netlist thus obtained is used as an input to Cadence SOC Encounter  to generate the layout.
-The layout being closest to the actual hardware is most suitable for estimation of powert consumed in the design for a particular input sequence.
-Power has been calculated using Cadence SOC Encounter.
-
-The memory could either be built up as a single bank or as a combination of multiple banks.
-The later would aid parallelism in the circuit.
-Also, the memory could be pipelined.
-We have tried to explore 6 different memory architectures for different examples.
-This would help us compare the performance of the circuits with multiple banks versus single banks, pipelined as opposed to non-pipelined memory.
-}
-\section{Memory Modelling}
-For encounter to estimate power, delay and area of the entire hardware, the memory black boxes must have the respective information.
-We use Cacti 3.2 to estimate power and delay in the memory black box.
-The area is modelled using  [2].
-
-area(in mm2) = (0.001){ * }(tech)^{2.07} { * } (bits)^{0.9} { * } (port)^{0.7} + {0.0048}
-
-The values of power, area and delay obtained from the above models are used to generate the necessary information for the blackboxes in timing library file (TLF) and library exchange file (LEF) formats.The vital timing information is also obtained from delay models.
-To describe a memory element as a black box, we instruct Synopsys DC to turn off its synthesis when it encounters the architecture for the memory.
-Thus for synopsys, the memory is an entity definition with well defined input & output ports but a blank architecture.
-Similarly, Cadence SOC encounter is made to infer black boxes whenever a memory component is instantiated from the TLF and LEF files.
-
-\section{Power Estimation}
-Using the final gate level verilog netlist provided by encounter we perform a post -layout simulation and dump all the signal information to a vcd file.
-This vcd file which now contains all the switching activity information is used by SOC Encounter to generate a power report.
-One is also required to mention the top level entity name for which power is to be estimated.
-For examples like linpack, Redblack which required huge size of memories and also which took a long time to simulate, a sampling technique was followed, in order to reduce the size of VCD files (which was around 25Gb).
-In this technique, the final post layout netlist is simulated and the switching activities are captured at 10 random intervals of time.
-It was observed that, when power is calculated using these samples, the variance in power estimates for all the samples is in the orders of 1e-5.
-
-\section{Results}
-Five examples viz. A5, AES, Red Black trees, Linpack and FFT were selected. 
-The memory architecture choices are shown table 1:
-
-\begin{table}[t]
-
-\caption{Architecture choices for each example}
-\begin{center}
-{\begin{tabular}{c  c | c  c  c}
-\hline
-ARCHITECTURE && \multicolumn{3}{c}{MEMORY BANK CHOICE} \\  
-CHOICES && 1& 2& 4 \\ [3ex] 
-\hline
-Example& Memory Size &&Base Bank Address Width \\ [1ex]
-\hline 
-A5& 16& 4& 4& 4\\[1ex]
-LINPACK& 16K& 12& 12& 12\\[1ex]
-R-B TREES& 16k& 12& 12& 12\\[1ex]
-FFT& 512& 8& 8& 8\\[1ex]
-AES& 1024& 8& 8& 8\\[1ex]
-\hline
-
-\end{tabular}}
-\label{diffstruc}
-\end{center}	
-\end{table}
-
-For each of memory architecture choices, mux and demux degree of 2 and infinity was selected.
-The range of frequencies tried for each architecture is shown in table 2:
-
-\begin{table}[t]
-\caption{Range of Frequencies for Each Architecture}
-\begin{center}
-{\begin{tabular}{c | c  c  c  c  c  c}
-\hline
- & \multicolumn{6}{c}{ARCHITECTURE CHOICES} \\[1ex]
- &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
-\hline
-Example& \multicolumn{6}{c}{Frequency of Operation (in MHz)} \\ [1ex]
-\hline
-A5& 71.42& 71.42& 83.33& 71.42& 83.33& 71.43\\[1ex]
-LINPACK& 45.45& 41.67& 41.67& 41.67& 41.67& 38.4\\[1ex]
-R-B TREES& 41.67 and 62.5 & 41.67 and 55.56& 71.42& 50& 71.42& 38.36\\[1ex]
-FFT& 41.66& 41.66& 41.66& 45.45& 45.45& 45.45\\[1ex]
-AES& 45.45 and 71.42& 45.45 and 71.42& 71.42& 71.42& 71.42& 45.45\\[1ex]
-\hline
-
-\end{tabular}}
-\\[1ex]
-Note : Architecture choices are in the form [memory bank x mux/demux degree].
-For eg.
-4 x 2 indicates an architecture with number of memory banks as 4 and mux/demux degree of 2.
-0 indicates a mux/demux degree of infinity.
-\label{diffstruc}
-\end{center}
-\end{table}
-
-\subsection{Reports}
-The energy, delay and area estimates for each example with their respective 6 architectures are as follows:
-
-\subsubsection{A5/1}
-\\
-The energy, delay and area estimates for each architecture / frequency combination is given in table 3.
-Figure 2 and 3 shows the Power-Delay and Area-Delay plot for A5/1. 
-\begin{table}[h!]
-\caption{Energy / Delay / Area for each architecture/frequency combination}
-\begin{center}
-{\begin{tabular}{|c | c |  c |  c  | c  |  c |  c|}
-\hline
-A5/1 &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
-\hline
-Energy &&&&&& \\ [1ex]
-\hline
-Delay (ns)& 308& 280& 252& 266& 240& 252\\[1ex] \hline
-Area (mm^2)& 1.45& 1.3& 1.77& 1.52& 2.3& 1.9\\[1ex]
-\hline
-
-\end{tabular}}
-\label{diffstruc}
-\end{center}
-\end{table}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{a5p.ps}} \par}
-\caption{Power-Delay plot for A5/1}
-\end{figure}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{a5a.ps}} \par}
-\caption{Area-Delay plot for A5/1}
-\end{figure}
-
-\subsubsection{LINPACK}
-\\
-The energy, delay and area estimates for each architecture / frequency combination is given in table 4.
-Figure 4 and 5 shows the Power-Delay and Area-Delay plot for Linpack.
-
-\begin{table}[h!]
-\caption{Energy / Delay / Area for each architecture/frequency combination}
-\begin{center}
-{\begin{tabular}{|c | c |  c |  c  | c  |  c |  c|}
-\hline
-LINPACK &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
-\hline
-Energy &&&&&& \\ [1ex]
-\hline
-Delay (ms)& 56.48& 43.65& 55.63& 37.68& 51.64& 38.71\\[1ex] \hline
-Area (mm^2)& 27.5& 25.5& 30.8& 27& 36.2& 28.9\\[1ex]
-\hline
-
-\end{tabular}}
-\label{diffstruc}
-\end{center}
-\end{table}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{lpp.ps}} \par}
-\caption{Power-Delay plot for LINPACK}
-\end{figure}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{lpa.ps}} \par}
-\caption{Area-Delay plot for LINPACK}
-\end{figure}
-
-\subsubsection{R-B TREES}
-\\
-The energy, delay and area estimates for each architecture / frequency combination is given in table 5.
-Figure 6 and 7 shows the Power-Delay and Area-Delay plot for Red-Black Trees.
-
-\begin{table}[h!]
-\caption{Energy / Delay / Area for each architecture/frequency combination}
-\begin{center}
-{\begin{tabular}{|c | c |  c |  c  | c  |  c |  c|}
-\hline
-R-B TREES &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
-\hline
-Energy &&&&&& \\ [1ex]
-\hline
-Delay (ms)& 18.74 and 12.5& 9.88 and 7.41& 10.9& 8.17& 10.87& 10.57\\[1ex] \hline
-Area (mm^2)& 19 and 23& 18 and 21& 26& 22& 32& 25\\[1ex]
-\hline
-
-\end{tabular}}
-\label{diffstruc}
-\end{center}
-\end{table}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{rbp.ps}} \par}
-\caption{Power-Delay plot for RED-BLACK TREES}
-\end{figure}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{rba.ps}} \par}
-\caption{Area-Delay plot for RED-BLACK TREES}
-\end{figure}
-
-\subsubsection{FFT}
-\\
-The energy, delay and area estimates for each architecture / frequency combination is given in table 6.
-Figure 8 and 9 shows the Power-Delay and Area-Delay plot for FFT.
-
-\begin{table}[h!]
-\caption{Energy / Delay / Area for each architecture/frequency combination}
-\begin{center}
-{\begin{tabular}{|c | c |  c |  c  | c  |  c |  c|}
-\hline
-FFT &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
-\hline
-Energy &&&&&& \\ [1ex]
-\hline
-Delay (us)& 203.792& 154.88& 189.27& 141.606& 172.453& 140.55\\[1ex] \hline
-Area (mm^2)& 5.4& 5.07 & 6.6& 5.7& 9.2& 6.7\\[1ex]
-\hline
-
-\end{tabular}}
-\label{diffstruc}
-\end{center}
-\end{table}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{fftp.ps}} \par}
-\caption{Power-Delay plot for FFT}
-\end{figure}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{ffta.ps}} \par}
-\caption{Area-Delay plot for FFT}
-\end{figure}
-
-\subsubsection{AES}
-\\
-The energy, delay and area estimates for each architecture / frequency combination is given in table 7.
-Figure 10 and 11 shows the Power-Delay and Area-Delay plot for AES.
-
-\begin{table}[h!]
-\caption{Energy / Delay / Area for each architecture/frequency combination}
-\begin{center}
-{\begin{tabular}{|c | c |  c |  c  | c  |  c |  c|}
-\hline
-AES &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
-\hline
-Energy &&&&&& \\ [1ex]
-\hline
-Delay (ms)& 1.21 and 0.769& 0.67 and 0.43& 0.762& 0.418& 0.759& 0.655\\[1ex] \hline
-Area (mm^2)& 7.67 and 9.03& 6.63 and 6.5 & 13& 7.5& 20.25& 9.4\\[1ex]
-\hline
-
-\end{tabular}}
-\label{diffstruc}
-\end{center}
-\end{table}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{aesp.ps}} \par}
-\caption{Power-Delay plot for AES}
-\end{figure}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{aesa.ps}} \par}
-\caption{Area-Delay plot for AES}
-\end{figure}
-
-\section{Conclusion}
-From the results, tabulated earlier, we can conclude the following:
-The circuits generated showed very limited parallelism in memory access patterns.
-Also, there is no performance boost in pipelined memory as opposed to the non-pipelined memory.
-We can observe that there's no significant improvement in delay as we increase the number of banks.
-Also another trend that we observe is that circuits are more energy efficient at higher frequencies.
-
-\newpage
-\addcontentsline{toc}{chapter}{References}
-\begin{thebibliography}{99}
-
-\bibitem{} Premkishore Shivakumar, Norman P. Jouppi. Cacti 3.0 : An Integreted Cache Timing, Power, and Area model,
-{\bf Western Research Laboratory}, 250 University Avenue Palo Alto, California 94301 USA.
-
-\bibitem{} Banit Agrawal, Timothy Sherwood. Guiding Architectural SRAM models, 
-{\bf Published in International Conference of Computer Design (ICCD) 2006},Department of Computer Science, University of California, Santa Barbara.
-
-\end{thebibliography}
-
-\end{document}

### 2010-05-24 Documents/Reports/Kunal-Sreenivas/rb_trees.tex
initial commit

--- 2010-05-24 Documents/Reports/Kunal-Sreenivas/artex.tex
@@
-\documentclass[a4paper, twoside]{article}      % Comments after  % are ignored
-\usepackage{amsmath,amssymb,amsfonts} % Typical maths resource packages
-\usepackage{graphics}                 % Packages to allow inclusion of graphics
-\usepackage{color}                    % For creating coloured text and background
-\usepackage{hyperref}                 % For creating hyperlinks in cross references
-
-\oddsidemargin 0cm
-\evensidemargin 0cm
-
-\pagestyle{myheadings}         % Option to put page headers
-                               % Needed \documentclass[a4paper,twoside]{article}
-
-%\markboth{{\small\it PDF\ \LaTeX \ coloured text and graphics}}
-
-%{{\small\it C.T.J. Dodson} }
-\textwidth 15.5cm
-\topmargin -1cm
-\parindent 0cm
-\textheight 24cm
-\parskip 1mm
-
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{proposition}[theorem]{Proposition}
-\newtheorem{corollary}[theorem]{Corollary}
-\newtheorem{lemma}[theorem]{Lemma}
-\newtheorem{remark}[theorem]{Remark}
-\newtheorem{definition}[theorem]{Definition}
-
-\def\R{\mathbb{ R}}
-
-\def\S{\mathbb{ S}}
-
-\date{\small\it April 20, 2009}
-
-\title{\fcolorbox{white}{white}{\color{black}Power / Delay Estimation of Auto Generated Circuits}}
-
-%\footnote{A demonstration example including colored text and graphics}}}
-
-\author{ K. P. Ghosh, S.Subramanian \\
-{\small Department of Electrical Engineering, IIT, Bombay, India}
- }
-
-\begin{document}
-\maketitle
-\begin{abstract}
-{
-The goal of the research project was to characterize RTL generated by AHIR flow.
-We had taken a set of examples, used AHIR to generate RTL, mapped the RTL tolayout, and finally performed power, area and delay estimation for the above examples.
-For all our circuit examples we have used  OSU Standard Cell library for TSMC 0.18 technology and Synopsis  Design Compiler to perform synthesis.
-The netlist thus obtained is used as an input to Cadence SOC Encounter  to generate the layout.
-The layout being closest to the actual hardware is most suiatble for estimation of power consumed in the design for a particular input sequence.Hence, we present an alternate approach of implementing memory in our circuits as Blackboxes.
-Cacti 3.2[1] was used to model the power and delay of the memory blackboxes.
-Five examples viz.
-A5, AES, Red Black trees, Linpack and FFT were selected, six memory architectures were explored for each and results are summarized in final section.
-}
-\end{abstract}
-
-%{\color{black}{\sc keywords: AHIR, Memory Modelling, Cacti}}
-
-\section{Introduction}
-{AHIR provides a platform to describe hardware in a high level programming language like C and translates it to its equivalent hardware description.
-When we use this description of hardware, followed by the RTL synthesis and then its layout, we would like to characterise the hardware thus generated to meet certain physical and performance criteria.
-Area , power and delay are  important criteria that mark the quality of the hardware.
-The current approach followed by us to obtain layout from a VHDL description is as shown in the figure 1.
-Memories in the hardware have been treated as blackboxes with suitable power and area models obtained from Cacti 3.2.
-
-\begin{figure}[ht]
-{\centering \resizebox*{6in}{4in}{\includegraphics{list_of_figures1.ps}} \par}
-\caption{Flow describing VHDL Description to Layout}
-\end{figure}
-
-At each level, a different tool is used for a specific purpose.
-At first,  the VHDL code  is simulated and checked for  its functionality using ModelSim 6.3a.
-The next step is  to perform synthesis and obtain a gate level netlist.
-For all our circuit examples we have used  OSU Standard Cell library for TSMC 0.18 technology and Synopsis  Design Compiler (DC) to perform synthesis.
-The netlist thus obtained is used as an input to Cadence SOC Encounter  to generate the layout.
-The layout being closest to the actual hardware is most suitable for estimation of powert consumed in the design for a particular input sequence.
-Power has been calculated using Cadence SOC Encounter.
-
-The memory could either be built up as a single bank or as a combination of multiple banks.
-The later would aid parallelism in the circuit.
-Also, the memory could be pipelined.
-We have tried to explore 6 different memory architectures for different examples.
-This would help us compare the performance of the circuits with multiple banks versus single banks, pipelined as opposed to non-pipelined memory.
-}
-\section{Memory Modelling}
-For encounter to estimate power, delay and area of the entire hardware, the memory black boxes must have the respective information.
-We use Cacti 3.2 to estimate power and delay in the memory black box.
-The area is modelled using  [2].
-
-area(in mm2) = (0.001){ * }(tech)^{2.07} { * } (bits)^{0.9} { * } (port)^{0.7} + {0.0048}
-
-The values of power, area and delay obtained from the above models are used to generate the necessary information for the blackboxes in timing library file (TLF) and library exchange file (LEF) formats.The vital timing information is also obtained from delay models.
-To describe a memory element as a black box, we instruct Synopsys DC to turn off its synthesis when it encounters the architecture for the memory.
-Thus for synopsys, the memory is an entity definition with well defined input & output ports but a blank architecture.
-Similarly, Cadence SOC encounter is made to infer black boxes whenever a memory component is instantiated from the TLF and LEF files.
-
-\section{Power Estimation}
-Using the final gate level verilog netlist provided by encounter we perform a post -layout simulation and dump all the signal information to a vcd file.
-This vcd file which now contains all the switching activity information is used by SOC Encounter to generate a power report.
-One is also required to mention the top level entity name for which power is to be estimated.
-For examples like linpack, Redblack which required huge size of memories and also which took a long time to simulate, a sampling technique was followed, in order to reduce the size of VCD files (which was around 25Gb).
-In this technique, the final post layout netlist is simulated and the switching activities are captured at 10 random intervals of time.
-It was observed that, when power is calculated using these samples, the variance in power estimates for all the samples is in the orders of 1e-5.
-
-\section{Results}
-Five examples viz. A5, AES, Red Black trees, Linpack and FFT were selected. 
-The memory architecture choices are shown table 1:
-
-\begin{table}[t]
-
-\caption{Architecture choices for each example}
-\begin{center}
-{\begin{tabular}{c  c | c  c  c}
-\hline
-ARCHITECTURE && \multicolumn{3}{c}{MEMORY BANK CHOICE} \\  
-CHOICES && 1& 2& 4 \\ [3ex] 
-\hline
-Example& Memory Size &&Base Bank Address Width \\ [1ex]
-\hline 
-A5& 16& 4& 4& 4\\[1ex]
-LINPACK& 16K& 12& 12& 12\\[1ex]
-R-B TREES& 16k& 12& 12& 12\\[1ex]
-FFT& 512& 8& 8& 8\\[1ex]
-AES& 1024& 8& 8& 8\\[1ex]
-\hline
-
-\end{tabular}}
-\label{diffstruc}
-\end{center}	
-\end{table}
-
-For each of memory architecture choices, mux and demux degree of 2 and infinity was selected.
-The range of frequencies tried for each architecture is shown in table 2:
-
-\begin{table}[t]
-\caption{Range of Frequencies for Each Architecture}
-\begin{center}
-{\begin{tabular}{c | c  c  c  c  c  c}
-\hline
- & \multicolumn{6}{c}{ARCHITECTURE CHOICES} \\[1ex]
- &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
-\hline
-Example& \multicolumn{6}{c}{Frequency of Operation (in MHz)} \\ [1ex]
-\hline
-A5& 71.42& 71.42& 83.33& 71.42& 83.33& 71.43\\[1ex]
-LINPACK& 45.45& 41.67& 41.67& 41.67& 41.67& 38.4\\[1ex]
-R-B TREES& 41.67 and 62.5 & 41.67 and 55.56& 71.42& 50& 71.42& 38.36\\[1ex]
-FFT& 41.66& 41.66& 41.66& 45.45& 45.45& 45.45\\[1ex]
-AES& 45.45 and 71.42& 45.45 and 71.42& 71.42& 71.42& 71.42& 45.45\\[1ex]
-\hline
-
-\end{tabular}}
-\\[1ex]
-Note : Architecture choices are in the form [memory bank x mux/demux degree].
-For eg.
-4 x 2 indicates an architecture with number of memory banks as 4 and mux/demux degree of 2.
-0 indicates a mux/demux degree of infinity.
-\label{diffstruc}
-\end{center}
-\end{table}
-
-\subsection{Reports}
-The energy, delay and area estimates for each example with their respective 6 architectures are as follows:
-
-\subsubsection{A5/1}
+\subsubsection{R-B TREES}
 \\
-The energy, delay and area estimates for each architecture / frequency combination is given in table 3.
-Figure 2 and 3 shows the Power-Delay and Area-Delay plot for A5/1. 
-\begin{table}[h!]
-\caption{Energy / Delay / Area for each architecture/frequency combination}
-\begin{center}
-{\begin{tabular}{|c | c |  c |  c  | c  |  c |  c|}
-\hline
-A5/1 &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
-\hline
-Energy &&&&&& \\ [1ex]
-\hline
-Delay (ns)& 308& 280& 252& 266& 240& 252\\[1ex] \hline
-Area (mm^2)& 1.45& 1.3& 1.77& 1.52& 2.3& 1.9\\[1ex]
-\hline
-
-\end{tabular}}
-\label{diffstruc}
-\end{center}
-\end{table}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{a5p.ps}} \par}
-\caption{Power-Delay plot for A5/1}
-\end{figure}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{a5a.ps}} \par}
-\caption{Area-Delay plot for A5/1}
-\end{figure}
-
-\subsubsection{LINPACK}
-\\
-The energy, delay and area estimates for each architecture / frequency combination is given in table 4.
-Figure 4 and 5 shows the Power-Delay and Area-Delay plot for Linpack.
+The energy, delay and area estimates for each architecture / frequency combination is given in TABLE V.
+In Fig. 6, Fig. 7 shows the Power-Delay and Area-Delay plot for Red-Black Trees.
 
@@
 \begin{center}
-{\begin{tabular}{|c | c |  c |  c  | c  |  c |  c|}
+{\begin{tabular}{ c |c  c  c  c   c   c    c   c}
 \hline
-LINPACK &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
+R-B TREES &\multicolumn{2}{c}{1x2} &\multicolumn{2}{c}{1x0} &2x0 &2x2 &4x2 &4x0 \\ [1ex]
 \hline
-Energy &&&&&& \\ [1ex]
-\hline
-Delay (ms)& 56.48& 43.65& 55.63& 37.68& 51.64& 38.71\\[1ex] \hline
-Area (mm^2)& 27.5& 25.5& 30.8& 27& 36.2& 28.9\\[1ex]
-\hline
+Frequency (MHz)& 41.67& 62.5& 41.67& 55.56& 71.42& 50 & 71.42&38.36\\[1ex]
+Energy (mJ)&3.935 & 7.062 &1.511 &2.54 &8.07 &2.851 &11.47  &3.23 \\ [1ex]
 
-\end{tabular}}
-\label{diffstruc}
-\end{center}
-\end{table}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{lpp.ps}} \par}
-\caption{Power-Delay plot for LINPACK}
-\end{figure}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{lpa.ps}} \par}
-\caption{Area-Delay plot for LINPACK}
-\end{figure}
-
-\subsubsection{R-B TREES}
-\\
-The energy, delay and area estimates for each architecture / frequency combination is given in table 5.
-Figure 6 and 7 shows the Power-Delay and Area-Delay plot for Red-Black Trees.
-
-\begin{table}[h!]
-\caption{Energy / Delay / Area for each architecture/frequency combination}
-\begin{center}
-{\begin{tabular}{|c | c |  c |  c  | c  |  c |  c|}
-\hline
-R-B TREES &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
-\hline
-Energy &&&&&& \\ [1ex]
-\hline
-Delay (ms)& 18.74 and 12.5& 9.88 and 7.41& 10.9& 8.17& 10.87& 10.57\\[1ex] \hline
-Area (mm^2)& 19 and 23& 18 and 21& 26& 22& 32& 25\\[1ex]
+Delay (ms)& 18.74 & 12.5& 9.88 & 7.41& 10.9& 8.17& 10.87& 10.57\\[1ex] 
+Area (mm^2)& 19 & 23& 18 & 21& 26& 22& 32& 25\\[1ex]
 \hline
@@
 \end{figure}
-
-\subsubsection{FFT}
-\\
-The energy, delay and area estimates for each architecture / frequency combination is given in table 6.
-Figure 8 and 9 shows the Power-Delay and Area-Delay plot for FFT.
-
-\begin{table}[h!]
-\caption{Energy / Delay / Area for each architecture/frequency combination}
-\begin{center}
-{\begin{tabular}{|c | c |  c |  c  | c  |  c |  c|}
-\hline
-FFT &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
-\hline
-Energy &&&&&& \\ [1ex]
-\hline
-Delay (us)& 203.792& 154.88& 189.27& 141.606& 172.453& 140.55\\[1ex] \hline
-Area (mm^2)& 5.4& 5.07 & 6.6& 5.7& 9.2& 6.7\\[1ex]
-\hline
-
-\end{tabular}}
-\label{diffstruc}
-\end{center}
-\end{table}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{fftp.ps}} \par}
-\caption{Power-Delay plot for FFT}
-\end{figure}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{ffta.ps}} \par}
-\caption{Area-Delay plot for FFT}
-\end{figure}
-
-\subsubsection{AES}
-\\
-The energy, delay and area estimates for each architecture / frequency combination is given in table 7.
-Figure 10 and 11 shows the Power-Delay and Area-Delay plot for AES.
-
-\begin{table}[h!]
-\caption{Energy / Delay / Area for each architecture/frequency combination}
-\begin{center}
-{\begin{tabular}{|c | c |  c |  c  | c  |  c |  c|}
-\hline
-AES &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
-\hline
-Energy &&&&&& \\ [1ex]
-\hline
-Delay (ms)& 1.21 and 0.769& 0.67 and 0.43& 0.762& 0.418& 0.759& 0.655\\[1ex] \hline
-Area (mm^2)& 7.67 and 9.03& 6.63 and 6.5 & 13& 7.5& 20.25& 9.4\\[1ex]
-\hline
-
-\end{tabular}}
-\label{diffstruc}
-\end{center}
-\end{table}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{aesp.ps}} \par}
-\caption{Power-Delay plot for AES}
-\end{figure}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{aesa.ps}} \par}
-\caption{Area-Delay plot for AES}
-\end{figure}
-
-\section{Conclusion}
-From the results, tabulated earlier, we can conclude the following:
-The circuits generated showed very limited parallelism in memory access patterns.
-Also, there is no performance boost in pipelined memory as opposed to the non-pipelined memory.
-We can observe that there's no significant improvement in delay as we increase the number of banks.
-Also another trend that we observe is that circuits are more energy efficient at higher frequencies.
-
-\newpage
-\addcontentsline{toc}{chapter}{References}
-\begin{thebibliography}{99}
-
-\bibitem{} Premkishore Shivakumar, Norman P. Jouppi. Cacti 3.0 : An Integreted Cache Timing, Power, and Area model,
-{\bf Western Research Laboratory}, 250 University Avenue Palo Alto, California 94301 USA.
-
-\bibitem{} Banit Agrawal, Timothy Sherwood. Guiding Architectural SRAM models, 
-{\bf Published in International Conference of Computer Design (ICCD) 2006},Department of Computer Science, University of California, Santa Barbara.
-
-\end{thebibliography}
-
-\end{document}

### 2010-05-24 Documents/Reports/Kunal-Sreenivas/ref.tex
initial commit

--- 2010-05-24 Documents/Reports/Kunal-Sreenivas/artex.tex
@@
-\documentclass[a4paper, twoside]{article}      % Comments after  % are ignored
-\usepackage{amsmath,amssymb,amsfonts} % Typical maths resource packages
-\usepackage{graphics}                 % Packages to allow inclusion of graphics
-\usepackage{color}                    % For creating coloured text and background
-\usepackage{hyperref}                 % For creating hyperlinks in cross references
-
-\oddsidemargin 0cm
-\evensidemargin 0cm
-
-\pagestyle{myheadings}         % Option to put page headers
-                               % Needed \documentclass[a4paper,twoside]{article}
-
-%\markboth{{\small\it PDF\ \LaTeX \ coloured text and graphics}}
-
-%{{\small\it C.T.J. Dodson} }
-\textwidth 15.5cm
-\topmargin -1cm
-\parindent 0cm
-\textheight 24cm
-\parskip 1mm
-
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{proposition}[theorem]{Proposition}
-\newtheorem{corollary}[theorem]{Corollary}
-\newtheorem{lemma}[theorem]{Lemma}
-\newtheorem{remark}[theorem]{Remark}
-\newtheorem{definition}[theorem]{Definition}
-
-\def\R{\mathbb{ R}}
-
-\def\S{\mathbb{ S}}
-
-\date{\small\it April 20, 2009}
-
-\title{\fcolorbox{white}{white}{\color{black}Power / Delay Estimation of Auto Generated Circuits}}
-
-%\footnote{A demonstration example including colored text and graphics}}}
-
-\author{ K. P. Ghosh, S.Subramanian \\
-{\small Department of Electrical Engineering, IIT, Bombay, India}
- }
-
-\begin{document}
-\maketitle
-\begin{abstract}
-{
-The goal of the research project was to characterize RTL generated by AHIR flow.
-We had taken a set of examples, used AHIR to generate RTL, mapped the RTL tolayout, and finally performed power, area and delay estimation for the above examples.
-For all our circuit examples we have used  OSU Standard Cell library for TSMC 0.18 technology and Synopsis  Design Compiler to perform synthesis.
-The netlist thus obtained is used as an input to Cadence SOC Encounter  to generate the layout.
-The layout being closest to the actual hardware is most suiatble for estimation of power consumed in the design for a particular input sequence.Hence, we present an alternate approach of implementing memory in our circuits as Blackboxes.
-Cacti 3.2[1] was used to model the power and delay of the memory blackboxes.
-Five examples viz.
-A5, AES, Red Black trees, Linpack and FFT were selected, six memory architectures were explored for each and results are summarized in final section.
-}
-\end{abstract}
-
-%{\color{black}{\sc keywords: AHIR, Memory Modelling, Cacti}}
-
-\section{Introduction}
-{AHIR provides a platform to describe hardware in a high level programming language like C and translates it to its equivalent hardware description.
-When we use this description of hardware, followed by the RTL synthesis and then its layout, we would like to characterise the hardware thus generated to meet certain physical and performance criteria.
-Area , power and delay are  important criteria that mark the quality of the hardware.
-The current approach followed by us to obtain layout from a VHDL description is as shown in the figure 1.
-Memories in the hardware have been treated as blackboxes with suitable power and area models obtained from Cacti 3.2.
-
-\begin{figure}[ht]
-{\centering \resizebox*{6in}{4in}{\includegraphics{list_of_figures1.ps}} \par}
-\caption{Flow describing VHDL Description to Layout}
-\end{figure}
-
-At each level, a different tool is used for a specific purpose.
-At first,  the VHDL code  is simulated and checked for  its functionality using ModelSim 6.3a.
-The next step is  to perform synthesis and obtain a gate level netlist.
-For all our circuit examples we have used  OSU Standard Cell library for TSMC 0.18 technology and Synopsis  Design Compiler (DC) to perform synthesis.
-The netlist thus obtained is used as an input to Cadence SOC Encounter  to generate the layout.
-The layout being closest to the actual hardware is most suitable for estimation of powert consumed in the design for a particular input sequence.
-Power has been calculated using Cadence SOC Encounter.
-
-The memory could either be built up as a single bank or as a combination of multiple banks.
-The later would aid parallelism in the circuit.
-Also, the memory could be pipelined.
-We have tried to explore 6 different memory architectures for different examples.
-This would help us compare the performance of the circuits with multiple banks versus single banks, pipelined as opposed to non-pipelined memory.
-}
-\section{Memory Modelling}
-For encounter to estimate power, delay and area of the entire hardware, the memory black boxes must have the respective information.
-We use Cacti 3.2 to estimate power and delay in the memory black box.
-The area is modelled using  [2].
-
-area(in mm2) = (0.001){ * }(tech)^{2.07} { * } (bits)^{0.9} { * } (port)^{0.7} + {0.0048}
-
-The values of power, area and delay obtained from the above models are used to generate the necessary information for the blackboxes in timing library file (TLF) and library exchange file (LEF) formats.The vital timing information is also obtained from delay models.
-To describe a memory element as a black box, we instruct Synopsys DC to turn off its synthesis when it encounters the architecture for the memory.
-Thus for synopsys, the memory is an entity definition with well defined input & output ports but a blank architecture.
-Similarly, Cadence SOC encounter is made to infer black boxes whenever a memory component is instantiated from the TLF and LEF files.
-
-\section{Power Estimation}
-Using the final gate level verilog netlist provided by encounter we perform a post -layout simulation and dump all the signal information to a vcd file.
-This vcd file which now contains all the switching activity information is used by SOC Encounter to generate a power report.
-One is also required to mention the top level entity name for which power is to be estimated.
-For examples like linpack, Redblack which required huge size of memories and also which took a long time to simulate, a sampling technique was followed, in order to reduce the size of VCD files (which was around 25Gb).
-In this technique, the final post layout netlist is simulated and the switching activities are captured at 10 random intervals of time.
-It was observed that, when power is calculated using these samples, the variance in power estimates for all the samples is in the orders of 1e-5.
-
-\section{Results}
-Five examples viz. A5, AES, Red Black trees, Linpack and FFT were selected. 
-The memory architecture choices are shown table 1:
-
-\begin{table}[t]
-
-\caption{Architecture choices for each example}
-\begin{center}
-{\begin{tabular}{c  c | c  c  c}
-\hline
-ARCHITECTURE && \multicolumn{3}{c}{MEMORY BANK CHOICE} \\  
-CHOICES && 1& 2& 4 \\ [3ex] 
-\hline
-Example& Memory Size &&Base Bank Address Width \\ [1ex]
-\hline 
-A5& 16& 4& 4& 4\\[1ex]
-LINPACK& 16K& 12& 12& 12\\[1ex]
-R-B TREES& 16k& 12& 12& 12\\[1ex]
-FFT& 512& 8& 8& 8\\[1ex]
-AES& 1024& 8& 8& 8\\[1ex]
-\hline
-
-\end{tabular}}
-\label{diffstruc}
-\end{center}	
-\end{table}
-
-For each of memory architecture choices, mux and demux degree of 2 and infinity was selected.
-The range of frequencies tried for each architecture is shown in table 2:
-
-\begin{table}[t]
-\caption{Range of Frequencies for Each Architecture}
-\begin{center}
-{\begin{tabular}{c | c  c  c  c  c  c}
-\hline
- & \multicolumn{6}{c}{ARCHITECTURE CHOICES} \\[1ex]
- &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
-\hline
-Example& \multicolumn{6}{c}{Frequency of Operation (in MHz)} \\ [1ex]
-\hline
-A5& 71.42& 71.42& 83.33& 71.42& 83.33& 71.43\\[1ex]
-LINPACK& 45.45& 41.67& 41.67& 41.67& 41.67& 38.4\\[1ex]
-R-B TREES& 41.67 and 62.5 & 41.67 and 55.56& 71.42& 50& 71.42& 38.36\\[1ex]
-FFT& 41.66& 41.66& 41.66& 45.45& 45.45& 45.45\\[1ex]
-AES& 45.45 and 71.42& 45.45 and 71.42& 71.42& 71.42& 71.42& 45.45\\[1ex]
-\hline
-
-\end{tabular}}
-\\[1ex]
-Note : Architecture choices are in the form [memory bank x mux/demux degree].
-For eg.
-4 x 2 indicates an architecture with number of memory banks as 4 and mux/demux degree of 2.
-0 indicates a mux/demux degree of infinity.
-\label{diffstruc}
-\end{center}
-\end{table}
-
-\subsection{Reports}
-The energy, delay and area estimates for each example with their respective 6 architectures are as follows:
-
-\subsubsection{A5/1}
-\\
-The energy, delay and area estimates for each architecture / frequency combination is given in table 3.
-Figure 2 and 3 shows the Power-Delay and Area-Delay plot for A5/1. 
-\begin{table}[h!]
-\caption{Energy / Delay / Area for each architecture/frequency combination}
-\begin{center}
-{\begin{tabular}{|c | c |  c |  c  | c  |  c |  c|}
-\hline
-A5/1 &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
-\hline
-Energy &&&&&& \\ [1ex]
-\hline
-Delay (ns)& 308& 280& 252& 266& 240& 252\\[1ex] \hline
-Area (mm^2)& 1.45& 1.3& 1.77& 1.52& 2.3& 1.9\\[1ex]
-\hline
-
-\end{tabular}}
-\label{diffstruc}
-\end{center}
-\end{table}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{a5p.ps}} \par}
-\caption{Power-Delay plot for A5/1}
-\end{figure}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{a5a.ps}} \par}
-\caption{Area-Delay plot for A5/1}
-\end{figure}
-
-\subsubsection{LINPACK}
-\\
-The energy, delay and area estimates for each architecture / frequency combination is given in table 4.
-Figure 4 and 5 shows the Power-Delay and Area-Delay plot for Linpack.
-
-\begin{table}[h!]
-\caption{Energy / Delay / Area for each architecture/frequency combination}
-\begin{center}
-{\begin{tabular}{|c | c |  c |  c  | c  |  c |  c|}
-\hline
-LINPACK &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
-\hline
-Energy &&&&&& \\ [1ex]
-\hline
-Delay (ms)& 56.48& 43.65& 55.63& 37.68& 51.64& 38.71\\[1ex] \hline
-Area (mm^2)& 27.5& 25.5& 30.8& 27& 36.2& 28.9\\[1ex]
-\hline
-
-\end{tabular}}
-\label{diffstruc}
-\end{center}
-\end{table}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{lpp.ps}} \par}
-\caption{Power-Delay plot for LINPACK}
-\end{figure}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{lpa.ps}} \par}
-\caption{Area-Delay plot for LINPACK}
-\end{figure}
-
-\subsubsection{R-B TREES}
-\\
-The energy, delay and area estimates for each architecture / frequency combination is given in table 5.
-Figure 6 and 7 shows the Power-Delay and Area-Delay plot for Red-Black Trees.
-
-\begin{table}[h!]
-\caption{Energy / Delay / Area for each architecture/frequency combination}
-\begin{center}
-{\begin{tabular}{|c | c |  c |  c  | c  |  c |  c|}
-\hline
-R-B TREES &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
-\hline
-Energy &&&&&& \\ [1ex]
-\hline
-Delay (ms)& 18.74 and 12.5& 9.88 and 7.41& 10.9& 8.17& 10.87& 10.57\\[1ex] \hline
-Area (mm^2)& 19 and 23& 18 and 21& 26& 22& 32& 25\\[1ex]
-\hline
-
-\end{tabular}}
-\label{diffstruc}
-\end{center}
-\end{table}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{rbp.ps}} \par}
-\caption{Power-Delay plot for RED-BLACK TREES}
-\end{figure}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{rba.ps}} \par}
-\caption{Area-Delay plot for RED-BLACK TREES}
-\end{figure}
-
-\subsubsection{FFT}
-\\
-The energy, delay and area estimates for each architecture / frequency combination is given in table 6.
-Figure 8 and 9 shows the Power-Delay and Area-Delay plot for FFT.
-
-\begin{table}[h!]
-\caption{Energy / Delay / Area for each architecture/frequency combination}
-\begin{center}
-{\begin{tabular}{|c | c |  c |  c  | c  |  c |  c|}
-\hline
-FFT &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
-\hline
-Energy &&&&&& \\ [1ex]
-\hline
-Delay (us)& 203.792& 154.88& 189.27& 141.606& 172.453& 140.55\\[1ex] \hline
-Area (mm^2)& 5.4& 5.07 & 6.6& 5.7& 9.2& 6.7\\[1ex]
-\hline
-
-\end{tabular}}
-\label{diffstruc}
-\end{center}
-\end{table}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{fftp.ps}} \par}
-\caption{Power-Delay plot for FFT}
-\end{figure}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{ffta.ps}} \par}
-\caption{Area-Delay plot for FFT}
-\end{figure}
-
-\subsubsection{AES}
-\\
-The energy, delay and area estimates for each architecture / frequency combination is given in table 7.
-Figure 10 and 11 shows the Power-Delay and Area-Delay plot for AES.
-
-\begin{table}[h!]
-\caption{Energy / Delay / Area for each architecture/frequency combination}
-\begin{center}
-{\begin{tabular}{|c | c |  c |  c  | c  |  c |  c|}
-\hline
-AES &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
-\hline
-Energy &&&&&& \\ [1ex]
-\hline
-Delay (ms)& 1.21 and 0.769& 0.67 and 0.43& 0.762& 0.418& 0.759& 0.655\\[1ex] \hline
-Area (mm^2)& 7.67 and 9.03& 6.63 and 6.5 & 13& 7.5& 20.25& 9.4\\[1ex]
-\hline
-
-\end{tabular}}
-\label{diffstruc}
-\end{center}
-\end{table}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{aesp.ps}} \par}
-\caption{Power-Delay plot for AES}
-\end{figure}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{aesa.ps}} \par}
-\caption{Area-Delay plot for AES}
-\end{figure}
-
-\section{Conclusion}
-From the results, tabulated earlier, we can conclude the following:
-The circuits generated showed very limited parallelism in memory access patterns.
-Also, there is no performance boost in pipelined memory as opposed to the non-pipelined memory.
-We can observe that there's no significant improvement in delay as we increase the number of banks.
-Also another trend that we observe is that circuits are more energy efficient at higher frequencies.
-
-\newpage
 \addcontentsline{toc}{chapter}{References}
@@
 \end{thebibliography}
-
-\end{document}

### 2010-05-24 Documents/Reports/Kunal-Sreenivas/report.tex
initial commit

--- 2010-05-24 Documents/Reports/Kunal-Sreenivas/artex.tex
@@
-\documentclass[a4paper, twoside]{article}      % Comments after  % are ignored
+\documentclass[conference, onecolumn]{IEEEtran}     % Comments after  % are ignored
 \usepackage{amsmath,amssymb,amsfonts} % Typical maths resource packages
@@
 
-\pagestyle{myheadings}         % Option to put page headers
+\pagestyle{myheadings}         % Option to put headers
                                % Needed \documentclass[a4paper,twoside]{article}
@@
 
-\title{\fcolorbox{white}{white}{\color{black}Power / Delay Estimation of Auto Generated Circuits}}
+\title{{Power / Delay Estimation of \\ Auto Generated Circuits}}
 
@@
 \begin{document}
+
+\pagenumbering{arabic}
 \maketitle
 \begin{abstract}
-{
-The goal of the research project was to characterize RTL generated by AHIR flow.
-We had taken a set of examples, used AHIR to generate RTL, mapped the RTL tolayout, and finally performed power, area and delay estimation for the above examples.
-For all our circuit examples we have used  OSU Standard Cell library for TSMC 0.18 technology and Synopsis  Design Compiler to perform synthesis.
+
+The goal of the research project was to characterize RTL generated by the AHIR flow.
+We have taken a set of examples, used AHIR to generate RTL, mapped the RTL to layout and finally performed power, area and delay estimation for the above examples.
+For all our circuit examples we have used the OSU Standard Cell library for TSMC 0.18 technology and Synopsis  Design Compiler to perform synthesis.
 The netlist thus obtained is used as an input to Cadence SOC Encounter to generate the layout.
-The layout being closest to the actual hardware is most suiatble for estimation of power consumed in the design for a particular input sequence.Hence, we present an alternate approach of implementing memory in our circuits as Blackboxes.
-Cacti 3.2[1] was used to model the power and delay of the memory blackboxes.
-Five examples viz.
-A5, AES, Red Black trees, Linpack and FFT were selected, six memory architectures were explored for each and results are summarized in final section.
-}
+The layout being closest to the actual hardware is most suiatble for estimation of power consumed in the design for a particular input sequence.
+Five examples viz. A5, AES, Red Black trees, Linpack and FFT were selected, six memory architectures were explored for each and results are summarized in final section.
+
 \end{abstract}
-
-%{\color{black}{\sc keywords: AHIR, Memory Modelling, Cacti}}
-
-\section{Introduction}
-{AHIR provides a platform to describe hardware in a high level programming language like C and translates it to its equivalent hardware description.
-When we use this description of hardware, followed by the RTL synthesis and then its layout, we would like to characterise the hardware thus generated to meet certain physical and performance criteria.
-Area , power and delay are  important criteria that mark the quality of the hardware.
-The current approach followed by us to obtain layout from a VHDL description is as shown in the figure 1.
-Memories in the hardware have been treated as blackboxes with suitable power and area models obtained from Cacti 3.2.
-
-\begin{figure}[ht]
-{\centering \resizebox*{6in}{4in}{\includegraphics{list_of_figures1.ps}} \par}
-\caption{Flow describing VHDL Description to Layout}
-\end{figure}
-
-At each level, a different tool is used for a specific purpose.
-At first,  the VHDL code  is simulated and checked for  its functionality using ModelSim 6.3a.
-The next step is  to perform synthesis and obtain a gate level netlist.
-For all our circuit examples we have used  OSU Standard Cell library for TSMC 0.18 technology and Synopsis  Design Compiler (DC) to perform synthesis.
-The netlist thus obtained is used as an input to Cadence SOC Encounter  to generate the layout.
-The layout being closest to the actual hardware is most suitable for estimation of powert consumed in the design for a particular input sequence.
-Power has been calculated using Cadence SOC Encounter.
-
-The memory could either be built up as a single bank or as a combination of multiple banks.
-The later would aid parallelism in the circuit.
-Also, the memory could be pipelined.
-We have tried to explore 6 different memory architectures for different examples.
-This would help us compare the performance of the circuits with multiple banks versus single banks, pipelined as opposed to non-pipelined memory.
-}
-\section{Memory Modelling}
-For encounter to estimate power, delay and area of the entire hardware, the memory black boxes must have the respective information.
-We use Cacti 3.2 to estimate power and delay in the memory black box.
-The area is modelled using  [2].
-
-area(in mm2) = (0.001){ * }(tech)^{2.07} { * } (bits)^{0.9} { * } (port)^{0.7} + {0.0048}
-
-The values of power, area and delay obtained from the above models are used to generate the necessary information for the blackboxes in timing library file (TLF) and library exchange file (LEF) formats.The vital timing information is also obtained from delay models.
-To describe a memory element as a black box, we instruct Synopsys DC to turn off its synthesis when it encounters the architecture for the memory.
-Thus for synopsys, the memory is an entity definition with well defined input & output ports but a blank architecture.
-Similarly, Cadence SOC encounter is made to infer black boxes whenever a memory component is instantiated from the TLF and LEF files.
-
-\section{Power Estimation}
-Using the final gate level verilog netlist provided by encounter we perform a post -layout simulation and dump all the signal information to a vcd file.
-This vcd file which now contains all the switching activity information is used by SOC Encounter to generate a power report.
-One is also required to mention the top level entity name for which power is to be estimated.
-For examples like linpack, Redblack which required huge size of memories and also which took a long time to simulate, a sampling technique was followed, in order to reduce the size of VCD files (which was around 25Gb).
-In this technique, the final post layout netlist is simulated and the switching activities are captured at 10 random intervals of time.
-It was observed that, when power is calculated using these samples, the variance in power estimates for all the samples is in the orders of 1e-5.
-
-\section{Results}
-Five examples viz. A5, AES, Red Black trees, Linpack and FFT were selected. 
-The memory architecture choices are shown table 1:
-
-\begin{table}[t]
-
-\caption{Architecture choices for each example}
-\begin{center}
-{\begin{tabular}{c  c | c  c  c}
-\hline
-ARCHITECTURE && \multicolumn{3}{c}{MEMORY BANK CHOICE} \\  
-CHOICES && 1& 2& 4 \\ [3ex] 
-\hline
-Example& Memory Size &&Base Bank Address Width \\ [1ex]
-\hline 
-A5& 16& 4& 4& 4\\[1ex]
-LINPACK& 16K& 12& 12& 12\\[1ex]
-R-B TREES& 16k& 12& 12& 12\\[1ex]
-FFT& 512& 8& 8& 8\\[1ex]
-AES& 1024& 8& 8& 8\\[1ex]
-\hline
-
-\end{tabular}}
-\label{diffstruc}
-\end{center}	
-\end{table}
-
-For each of memory architecture choices, mux and demux degree of 2 and infinity was selected.
-The range of frequencies tried for each architecture is shown in table 2:
-
-\begin{table}[t]
-\caption{Range of Frequencies for Each Architecture}
-\begin{center}
-{\begin{tabular}{c | c  c  c  c  c  c}
-\hline
- & \multicolumn{6}{c}{ARCHITECTURE CHOICES} \\[1ex]
- &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
-\hline
-Example& \multicolumn{6}{c}{Frequency of Operation (in MHz)} \\ [1ex]
-\hline
-A5& 71.42& 71.42& 83.33& 71.42& 83.33& 71.43\\[1ex]
-LINPACK& 45.45& 41.67& 41.67& 41.67& 41.67& 38.4\\[1ex]
-R-B TREES& 41.67 and 62.5 & 41.67 and 55.56& 71.42& 50& 71.42& 38.36\\[1ex]
-FFT& 41.66& 41.66& 41.66& 45.45& 45.45& 45.45\\[1ex]
-AES& 45.45 and 71.42& 45.45 and 71.42& 71.42& 71.42& 71.42& 45.45\\[1ex]
-\hline
-
-\end{tabular}}
-\\[1ex]
-Note : Architecture choices are in the form [memory bank x mux/demux degree].
-For eg.
-4 x 2 indicates an architecture with number of memory banks as 4 and mux/demux degree of 2.
-0 indicates a mux/demux degree of infinity.
-\label{diffstruc}
-\end{center}
-\end{table}
-
-\subsection{Reports}
-The energy, delay and area estimates for each example with their respective 6 architectures are as follows:
-
-\subsubsection{A5/1}
-\\
-The energy, delay and area estimates for each architecture / frequency combination is given in table 3.
-Figure 2 and 3 shows the Power-Delay and Area-Delay plot for A5/1. 
-\begin{table}[h!]
-\caption{Energy / Delay / Area for each architecture/frequency combination}
-\begin{center}
-{\begin{tabular}{|c | c |  c |  c  | c  |  c |  c|}
-\hline
-A5/1 &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
-\hline
-Energy &&&&&& \\ [1ex]
-\hline
-Delay (ns)& 308& 280& 252& 266& 240& 252\\[1ex] \hline
-Area (mm^2)& 1.45& 1.3& 1.77& 1.52& 2.3& 1.9\\[1ex]
-\hline
-
-\end{tabular}}
-\label{diffstruc}
-\end{center}
-\end{table}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{a5p.ps}} \par}
-\caption{Power-Delay plot for A5/1}
-\end{figure}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{a5a.ps}} \par}
-\caption{Area-Delay plot for A5/1}
-\end{figure}
-
-\subsubsection{LINPACK}
-\\
-The energy, delay and area estimates for each architecture / frequency combination is given in table 4.
-Figure 4 and 5 shows the Power-Delay and Area-Delay plot for Linpack.
-
-\begin{table}[h!]
-\caption{Energy / Delay / Area for each architecture/frequency combination}
-\begin{center}
-{\begin{tabular}{|c | c |  c |  c  | c  |  c |  c|}
-\hline
-LINPACK &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
-\hline
-Energy &&&&&& \\ [1ex]
-\hline
-Delay (ms)& 56.48& 43.65& 55.63& 37.68& 51.64& 38.71\\[1ex] \hline
-Area (mm^2)& 27.5& 25.5& 30.8& 27& 36.2& 28.9\\[1ex]
-\hline
-
-\end{tabular}}
-\label{diffstruc}
-\end{center}
-\end{table}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{lpp.ps}} \par}
-\caption{Power-Delay plot for LINPACK}
-\end{figure}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{lpa.ps}} \par}
-\caption{Area-Delay plot for LINPACK}
-\end{figure}
-
-\subsubsection{R-B TREES}
-\\
-The energy, delay and area estimates for each architecture / frequency combination is given in table 5.
-Figure 6 and 7 shows the Power-Delay and Area-Delay plot for Red-Black Trees.
-
-\begin{table}[h!]
-\caption{Energy / Delay / Area for each architecture/frequency combination}
-\begin{center}
-{\begin{tabular}{|c | c |  c |  c  | c  |  c |  c|}
-\hline
-R-B TREES &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
-\hline
-Energy &&&&&& \\ [1ex]
-\hline
-Delay (ms)& 18.74 and 12.5& 9.88 and 7.41& 10.9& 8.17& 10.87& 10.57\\[1ex] \hline
-Area (mm^2)& 19 and 23& 18 and 21& 26& 22& 32& 25\\[1ex]
-\hline
-
-\end{tabular}}
-\label{diffstruc}
-\end{center}
-\end{table}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{rbp.ps}} \par}
-\caption{Power-Delay plot for RED-BLACK TREES}
-\end{figure}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{rba.ps}} \par}
-\caption{Area-Delay plot for RED-BLACK TREES}
-\end{figure}
-
-\subsubsection{FFT}
-\\
-The energy, delay and area estimates for each architecture / frequency combination is given in table 6.
-Figure 8 and 9 shows the Power-Delay and Area-Delay plot for FFT.
-
-\begin{table}[h!]
-\caption{Energy / Delay / Area for each architecture/frequency combination}
-\begin{center}
-{\begin{tabular}{|c | c |  c |  c  | c  |  c |  c|}
-\hline
-FFT &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
-\hline
-Energy &&&&&& \\ [1ex]
-\hline
-Delay (us)& 203.792& 154.88& 189.27& 141.606& 172.453& 140.55\\[1ex] \hline
-Area (mm^2)& 5.4& 5.07 & 6.6& 5.7& 9.2& 6.7\\[1ex]
-\hline
-
-\end{tabular}}
-\label{diffstruc}
-\end{center}
-\end{table}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{fftp.ps}} \par}
-\caption{Power-Delay plot for FFT}
-\end{figure}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{ffta.ps}} \par}
-\caption{Area-Delay plot for FFT}
-\end{figure}
-
-\subsubsection{AES}
-\\
-The energy, delay and area estimates for each architecture / frequency combination is given in table 7.
-Figure 10 and 11 shows the Power-Delay and Area-Delay plot for AES.
-
-\begin{table}[h!]
-\caption{Energy / Delay / Area for each architecture/frequency combination}
-\begin{center}
-{\begin{tabular}{|c | c |  c |  c  | c  |  c |  c|}
-\hline
-AES &1x2 &1x0 &2x0 &2x2 &4x2 &4x0 \\ [1ex]
-\hline
-Energy &&&&&& \\ [1ex]
-\hline
-Delay (ms)& 1.21 and 0.769& 0.67 and 0.43& 0.762& 0.418& 0.759& 0.655\\[1ex] \hline
-Area (mm^2)& 7.67 and 9.03& 6.63 and 6.5 & 13& 7.5& 20.25& 9.4\\[1ex]
-\hline
-
-\end{tabular}}
-\label{diffstruc}
-\end{center}
-\end{table}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{aesp.ps}} \par}
-\caption{Power-Delay plot for AES}
-\end{figure}
-
-\begin{figure}[h!]
-{\centering \resizebox*{6in}{4in}{\includegraphics{aesa.ps}} \par}
-\caption{Area-Delay plot for AES}
-\end{figure}
-
-\section{Conclusion}
-From the results, tabulated earlier, we can conclude the following:
-The circuits generated showed very limited parallelism in memory access patterns.
-Also, there is no performance boost in pipelined memory as opposed to the non-pipelined memory.
-We can observe that there's no significant improvement in delay as we increase the number of banks.
-Also another trend that we observe is that circuits are more energy efficient at higher frequencies.
-
-\newpage
-\addcontentsline{toc}{chapter}{References}
-\begin{thebibliography}{99}
-
-\bibitem{} Premkishore Shivakumar, Norman P. Jouppi. Cacti 3.0 : An Integreted Cache Timing, Power, and Area model,
-{\bf Western Research Laboratory}, 250 University Avenue Palo Alto, California 94301 USA.
-
-\bibitem{} Banit Agrawal, Timothy Sherwood. Guiding Architectural SRAM models, 
-{\bf Published in International Conference of Computer Design (ICCD) 2006},Department of Computer Science, University of California, Santa Barbara.
-
-\end{thebibliography}
-
+\input{introduction.tex}
+\input{memory.tex}
+\input{arch_choice.tex}
+\input{a5.tex}
+\input{linpack.tex}
+\input{rb_trees.tex}
+\input{fft.tex}
+\input{aes.tex}
+\input{conclusion.tex}
+\input{ref.tex}
 \end{document}

### 2010-05-24 examples/exhaustive-dp-test/test.c

@@ -1,0 +1,180 @@
+#ifdef RUN
+#include <stdio.h>
+#endif
+
+int int_x[6];
+int int_y[6];
+
+int int_arith[3];
+unsigned int_cmp[6];
+int int_shift[2];
+
+unsigned short short_x[6];
+unsigned short short_y[6];
+
+unsigned short short_arith[3];
+unsigned short short_cmp[6];
+unsigned short short_shift[2];
+
+float float_x[6];
+float float_y[6];
+
+float float_arith[3];
+unsigned float_cmp[6];
+
+int test_ints()
+{
+  unsigned int cmp_vector = 0;
+  int result = 0;
+  int i;
+  
+  int_arith[0] = int_x[0] + int_y[0];
+  int_arith[1] = int_x[1] - int_y[1];
+  int_arith[2] = int_x[2] * int_y[2];
+
+  int_cmp[0] = int_x[0] > int_y[0];
+  int_cmp[1] = int_x[1] >= int_y[1];
+  int_cmp[2] = int_x[2] < int_y[2];
+  int_cmp[3] = int_x[3] <= int_y[3];
+  int_cmp[4] = int_x[4] == int_y[4];
+  int_cmp[5] = int_x[5] != int_y[5];
+
+  for (i = 0; i < 6; i++)
+    cmp_vector = cmp_vector & (int_cmp[i] << (i << 2));
+
+  int_shift[0] = int_x[0] >> 4;
+  int_shift[1] = int_y[0] << 4;
+
+  result = int_arith[0];
+  result &= int_arith[1];
+  result |= int_arith[2];
+  
+  result ^= int_shift[0];
+  result ^= int_shift[1];
+
+  result += cmp_vector;
+
+#ifdef RUN
+  printf("%x\n", result);
+#endif
+  return result;
+}
+
+short test_shorts()
+{
+  short cmp_vector = 0;
+  short result = 0;
+  int i;
+  
+  short_arith[0] = short_x[0] + short_y[0];
+  short_arith[1] = short_x[1] - short_y[1];
+  short_arith[2] = short_x[2] * short_y[2];
+
+  short_cmp[0] = short_x[0] > short_y[0];
+  short_cmp[1] = short_x[1] >= short_y[1];
+  short_cmp[2] = short_x[2] < short_y[2];
+  short_cmp[3] = short_x[3] <= short_y[3];
+  short_cmp[4] = short_x[4] == short_y[4];
[…]

### 2010-05-24 examples/globals/globals.c

@@ -1,0 +1,11 @@
+int a[3] = { 2, 3, 4 };
+int b = 10;
+int select = 1;
+
+int main (void)
+{
+  if (select == 0)
+    return (a[0] + a[1] + a[2]);
+  else
+    return (a[0] + a[1] - a[2]);
+}

### 2010-05-24 examples/lu/lu.c

@@ -1,0 +1,144 @@
+float divide( float dividend, float divisor)
+{
+        float eps = 0.001;
+        float diplus1,di,niplus1,nicurr,ni,riplus1,ri;
+        int i;
+
+        di = divisor;
+        ni = dividend;
+
+        for (i=0; di<=0.5; i++)
+        {
+                di = di * 2;
+                ni = ni * 2;
+        }
+
+        for ( i=0; di>1; i++)
+        {
+                di = di * 0.5;
+                ni = ni * 0.5;
+        }
+
+        nicurr = ni -1;
+        ri = 2 - di;
+
+        for ( i =0; (ni - nicurr > eps) ; i++ )
+        {
+		nicurr = ni;
+                di = di * ri;
+                ni = ni * ri;
+                ri = 2 - di;
+        }
+
+        return ni;
+}
+
+
+void start(int noofelem,float *A, int *rcoeff,int *ccoeff,float *l_array,float *u_array)
+{	
+	int n=3;
+	float B[3][3];
+	float L[3][3],U[3][3];
+	int maxcoeff=0;
+	float temp=0;
+	int i,j,k,p;
+
+  n = 3;
+
+  for (i = 0; i < n; i++)
+    {
+      for (j = 0; j < n; j++)
+	{
+	  B[i][j] = 0;
+	}
+    }
+
+  for (i = 0; i < noofelem; i++)
+    {
+      B[rcoeff[i]][ccoeff[i]] = A[i];
+    }
+
+
+  for (i = 0; i < n; i++)
+    {
+      maxcoeff = i;
+      for (j = i + 1; j < n; j++)
+	{
+	  if (B[j][i] > B[i][i])
+	    maxcoeff = j;
+	}
+      for (j = 0; j < n; j++)
+	{
+	  temp = B[i][j];
+	  B[i][j] = B[maxcoeff][j];
+	  B[maxcoeff][j] = temp;
+	}
+    }
+
+  for (p = 0; p < n - 1; p++)
+    {
+      for (k = p + 1; k < n; k++)
+	{
+	  B[k][p] = B[k][p] / B[p][p];
+	  for (j = p + 1; j < n; j++)
+	    {
+	      B[k][j] = B[k][j] - B[k][p] * B[p][j];
+	    }
+	}
+<<<<<<< .mine
+    }
+
[…]

### 2010-05-24 examples/redblack/redblack.c

@@ -1,0 +1,1683 @@
+/*
+ * RCS $Id: redblack.h,v 1.9 2003/10/24 01:31:21 damo Exp $
+ */
+
+/*
+   Redblack balanced tree algorithm
+   Copyright (C) Damian Ivereigh 2000
+
+   This program is free software; you can redistribute it and/or modify
+   it under the terms of the GNU Lesser General Public License as published by
+   the Free Software Foundation; either version 2.1 of the License, or
+   (at your option) any later version. See the file COPYING for details.
+
+   This program is distributed in the hope that it will be useful,
+   but WITHOUT ANY WARRANTY; without even the implied warranty of
+   MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+   GNU General Public License for more details.
+
+   You should have received a copy of the GNU Lesser General Public License
+   along with this program; if not, write to the Free Software
+   Foundation, Inc., 675 Mass Ave, Cambridge, MA 02139, USA.
+ */
+
+/* Header file for redblack.c, should be included by any code that 
+** uses redblack.c since it defines the functions 
+*/ 
+ 
+#ifndef RB_STATIC
+#define RB_STATIC
+#endif
+
+#ifdef RUN
+#include <stdio.h>
+#include <stdlib.h>
+#endif
+
+typedef struct {
+    int key;
+    int value;
+} my_data;
+
+#define RB_CUSTOMIZE
+#define RB_ENTRY(name)	rb##name
+#define rbdata_t my_data
+
+#define RB_CMP(s,t) 0
+#define RB_LT(s,t) (s)->key < (t)->key
+#define RB_GT(s,t) (s)->key > (t)->key
+
+enum nodecolour { BLACK = 0, RED = 1};
+
+#define RB_GET(x,y)		&x->y
+#define RB_SET(x,y,v)		x->y = *(v)
+
+struct RB_ENTRY(node)
+{
+	struct RB_ENTRY(node) *left;		/* Left down */
+	struct RB_ENTRY(node) *right;		/* Right down */
+	struct RB_ENTRY(node) *up;		/* Up */
+	enum nodecolour colour;		/* Node colour */
+	RB_ENTRY(data_t) key;		/* User's key (and data) */
+};
+
+struct RB_ENTRY(tree) {
[…]

### 2010-05-24 examples/rw/rw.c

@@ -1,0 +1,14 @@
+
+int x;
+
+int rw()
+{
+  return x;
+}
+
+int start(void)
+{
+  x = 42;
+
+  return rw();
+}

### 2010-05-24 examples/shared_load/shared_load.c

@@ -1,0 +1,19 @@
+
+float a;
+float b;
+float c;
+
+float foo(float *x, float *y, float *z)
+{
+  *x = *y;
+  *y = *z;
+}
+
+float start (void)
+{
+  a = 6.789;
+  b = 9.432;
+  c = 8.0;
+
+  return foo(&a, &b, &c);
+}

### 2010-05-24 iolib/portmap.c

@@ -1,0 +1,60 @@
+#include "portmap.h"
+#include "libds/ds.h"
+
+#include <sys/stat.h>
+#include <stdlib.h>
+#include <unistd.h>
+
+#define TABLE_SIZE 7
+
+static HASHTABLE portmap = NULL;
+
+typedef struct _FileEntry 
+{
+  FILE *fd;
+  int mode;
+} FileEntry;
+
+void init_portmap()
+{
+  /* assert(NULL == portmap); */
+
+  portmap = htMake(TABLE_SIZE);
+}
+
+FILE* port2fifo(char *id, int rwbar)
+{
+  void *data;
+  FILE *F = NULL;
+  FileEntry *entry;
+  
+  if (NULL == portmap)
+    init_portmap();
+
+  data = htFind(portmap, id);
+
+  if (NULL != data) {
+    entry = (FileEntry*)data;
+    /* assert(!((rwbar == 0) ^ (entry->mode == 0))); */
+    return entry->fd;
+  }
+
+  if (rwbar != 0) {
+    F = fopen(id, "r");
+    while (!F) {
+      sleep(5);
+      F = fopen(id, "r");
+    }
+  } else {
+    unlink(id);
+    mkfifo(id, S_IRUSR | S_IWUSR | S_IRGRP | S_IROTH);
+    F = fopen(id, "w");
+  }
+
+  entry = (FileEntry*)malloc(sizeof(FileEntry));
+  entry->fd = F;
+  entry->mode = rwbar;
+  
+  htAdd(portmap, id, (void *)entry);
+  return F;
+}

### 2010-05-24 toolset/lib/AHIR/AHIRFactory.cpp

@@ -1,0 +1,291 @@
+#include "AHIRFactory.hpp"
+
+#include "AHIRModule.hpp"
+#include "DataPath.hpp"
+#include "ControlPath.hpp"
+#include "LinkLayer.hpp"
+#include "Arbiter.hpp"
+#include <Base/Program.hpp>
+#include "CPUtils.hpp"
+
+#include <boost/lexical_cast.hpp>
+
+#include <iostream>
+
+using namespace ahir;
+using namespace hls;
+
+AHIRFactory::AHIRFactory()
+{
+  dp = NULL;
+  port = NULL;
+  dpe = NULL;
+  cp = NULL;
+  place = NULL;
+  trans = NULL;
+  ln = NULL;
+  omega = NULL;
+}
+
+void AHIRFactory::create_module_hook(const std::string &id
+				     , const std::string &type)
+{
+  assert(type == "ahir");
+  ahir_module = new ahir::Module(id);
+  module = ahir_module;
+}
+
+// /* ----- data-path ----- */
+
+void AHIRFactory::create_dp(const std::string& id)
+{
+  dp = new DataPath(id);
+  ahir_module->register_dp(dp);
+}
+
+void AHIRFactory::create_dpe(const std::string& _id
+			     , const std::string& ntype
+			     , const std::string &description)
+{
+  unsigned id = boost::lexical_cast<unsigned>(_id);
+  dpe = new DPElement(id, hls::ntype(ntype), description);
+  dp->register_dpe(dpe);
+}
+
+void AHIRFactory::dpe_set_req(const std::string &sym)
+{
+  Symbol s = Symbol(boost::lexical_cast<unsigned>(sym));
+  dp->register_req(dpe, dpe_req, s);
+}
+
+void AHIRFactory::dpe_set_ack(const std::string &sym)
+{
+  Symbol s = Symbol(boost::lexical_cast<unsigned>(sym));
+  dp->register_ack(dpe, dpe_ack, s);
+}
+
+void AHIRFactory::dpe_set_value(const std::string &characters)
+{
+  assert(dpe->ntype == Constant);
+  dpe->value = characters;
+}
+
+void AHIRFactory::dpe_set_addressable(const std::string &characters)
+{
+  assert(dpe->ntype == Address);
+  dpe->addressable = program->find_addressable(characters);
+  assert(dpe->addressable);
+}
+
+void AHIRFactory::dpe_set_counterpart(const std::string &characters)
+{
[…]

### 2010-05-24 toolset/lib/AHIR/MuxRegions.cpp

@@ -1,0 +1,419 @@
+#include "MuxRegions.hpp"
+
+#include "../AHIR/ControlPath.hpp"
+#include "../AHIR/CPUtils.hpp"
+
+#include <iostream>
+#include <assert.h>
+#include <set>
+
+using namespace ahir;
+
+namespace ahir {
+  
+  // The parallel-merge region. This structure records the entry forks
+  // and the exit join that demarcate the parallel-merge. These are
+  // replaced by proxies connected to a single merge place.
+  struct MuxRegion {
+    unsigned id;
+
+    // elements within the region, not including the end points
+    std::set<CPElement*> elements;
+
+    typedef std::map<Transition*, Transition*> ProxyMap;
+    ProxyMap proxies, reverse_proxies;
+    void register_proxy(Transition* orig, Transition *proxy);
+    Transition* find_proxy(Transition *orig);
+    Transition* find_reverse_proxy(Transition *proxy);
+
+    // this merge replaces the entire region
+    Place *merge;
+
+    MuxRegion(unsigned _id)
+      : id(_id), merge(NULL)
+    {}
+
+    ~MuxRegion();
+  };
+
+  // List of MuxRegion identified by the id of their exit joins (the
+  // proxy exit has the same id when the MuxRegion is replaced).
+  typedef std::map<unsigned, MuxRegion*> MuxRegionMap;
+  
+}
+
+MuxRegion::~MuxRegion()
+{
+  // member of the elements vector are not destroyed since they are
+  // owned by the control-path.
+
+  // destroy all the proxy forks
+  for (ProxyMap::iterator fi = proxies.begin(), fe = proxies.end();
+       fi != fe; ++fi) {
+    delete (*fi).second;
+  }
+
+  delete merge;
+}
+
+void MuxRegion::register_proxy(Transition *orig, Transition *proxy)
+{
+  assert(!find_proxy(orig));
+  proxies[orig] = proxy;
+  reverse_proxies[proxy] = orig;
+}
+
+Transition* MuxRegion::find_proxy(Transition *orig)
+{
+  if (proxies.find(orig) != proxies.end())
+    return proxies[orig];
+  return NULL;
+}
+
+Transition* MuxRegion::find_reverse_proxy(Transition *proxy)
+{
+  if (reverse_proxies.find(proxy) != reverse_proxies.end())
+    return reverse_proxies[proxy];
+  return NULL;
+}
+
+namespace {
+
[…]

### 2010-05-24 toolset/lib/AHIR/Symbol.hpp

@@ -1,0 +1,15 @@
+#ifndef __SYMBOL_HPP__
+#define __SYMBOL_HPP__
+
+/*! \file Symbol.hpp
+  Header file that declares types for Symbol and SymbolList.
+  */
+
+#include <vector>
+
+//! A Symbol used in AHIR handshakes
+typedef unsigned Symbol;
+//! A list of Symbols
+typedef std::vector<Symbol> SymbolList;
+
+#endif

### 2010-05-24 toolset/lib/Base/Module.cpp

@@ -1,0 +1,11 @@
+#include "Module.hpp"
+
+#include <assert.h>
+
+using namespace hls;
+
+// void Module::register_callsite(unsigned id)
+// {
+  // assert(callsites.count(id) == 0);
+  // callsites.insert(id);
+// }

### 2010-05-24 toolset/lib/Base/Program.cpp

@@ -1,0 +1,53 @@
+#include "Program.hpp"
+#include "Module.hpp"
+#include "Addressable.hpp"
+
+#include <assert.h>
+
+using namespace hls;
+
+Program::Program(const std::string& _id)
+  : id(_id), start(NULL), first_free_address(0)
+{}
+
+const Type* Program::find_type(const std::string &id)
+{
+  if (types.find(id) != types.end())
+    return types[id];
+  return NULL;
+}
+
+void Program::register_type(const std::string &id, const Type *type)
+{
+  assert(!find_type(id));
+  types[id] = type;
+}
+
+Module* Program::find_module(const std::string &id)
+{
+  if (modules.find(id) != modules.end())
+    return modules[id];
+  else
+    return NULL;
+}
+
+void Program::register_module(Module *func)
+{
+  assert(!find_module(func->id));
+  modules[func->id] = func;
+}
+
+void Program::register_addressable(Addressable *addr)
+{
+  assert(addrs.find(addr->id) == addrs.end()
+	 && "addressable already exists");
+  addrs[addr->id] = addr;
+}
+
+Addressable* Program::find_addressable(const std::string &id)
+{
+  if (addrs.find(id) != addrs.end())
+    return addrs[id];
+  return NULL;
+}
+

### 2010-05-24 toolset/lib/Base/Type.cpp

@@ -1,0 +1,51 @@
+#include "Type.hpp"
+
+#include <sstream>
+
+using namespace hls;
+
+std::string hls::str(TypeID t) 
+{
+  switch (t) {
+    case APIntID: return "APInt"; break;
+    case APFloatID: return "APFloat"; break;
+    default: break;
+  }
+
+  assert(false);
+  return "xyzzy";
+}
+
+std::string hls::str(const Type *t)
+{
+  switch (t->type_id) {
+    case APIntID:
+      return get_apint_description(t->int_width);
+      break;
+
+    case APFloatID:
+      return get_apfloat_description(t->exp_width, t->frc_width);
+      break;
+
+    default:
+      assert(false);
+      return "xyzzy";
+      break;
+  }
+}
+
+std::string hls::get_apint_description(unsigned w)
+{
+  std::ostringstream str;
+  str << hls::str(APIntID) << "(" << w << ")";
+
+  return str.str();
+}
+
+std::string hls::get_apfloat_description(unsigned e, unsigned f)
+{
+  std::ostringstream str;
+  str << hls::str(APFloatID) << "(" << e << ", " << f << ")";
+
+  return str.str();
+}

### 2010-05-24 toolset/lib/CDFG/Utils.cpp

@@ -1,0 +1,198 @@
+#include "Utils.hpp"
+
+#include <deque>
+#include <iostream>
+
+#include <assert.h>
+
+using namespace cdfg;
+using namespace hls;
+
+void cdfg::check_wellformedness(CDFG *cdfg)
+{
+  if (!cdfg)
+    return;
+
+  for (CDFG::NodeList::iterator ni = cdfg->nodes.begin(), ne = cdfg->nodes.end();
+       ni != ne; ++ni) {
+    CDFGNode *node = (*ni).second;
+
+    node->colour = WHITE;
+  }
+
+  std::deque<CDFGNode*> queue;
+  queue.push_back(cdfg->start);
+  cdfg->start->colour = BLACK;
+
+  while (!queue.empty()) {
+    CDFGNode *node = queue.front();
+    queue.pop_front();
+    
+    assert(node->ports.size()
+	   == node->input_data_ports.size()
+	   + node->output_data_ports.size()
+	   + node->output_control_ports.size()
+	   + node->input_control_ports.size());
+
+    if (is_control(node)) {
+      if (node->ntype != Stop)
+	assert(node->output_control_ports.size() > 0);
+      if (node->ntype != Start)
+	assert(node->input_control_ports.size() > 0);
+    }
+
+    for (CDFGNode::PortList::iterator pi = node->ports.begin(), pe = node->ports.end();
+	 pi != pe; ++pi) {
+      Port *port = (*pi).second;
+      
+      assert(port->edge);
+
+      assert(port->edge->driver);
+      // if (port->edge->users.size() == 0)
+	// assert(cdfg->retval == port->parent);
+
+      if (is_output(port))
+	assert(port->edge->driver == port);
+      else
+	assert(port->edge->users.find(port) != port->edge->users.end());
+      
+      if (is_control(port))
+	assert(port->edge->users.size() == 1);
+    }
+
+    for (CDFGNode::PortList::iterator ci = node->output_control_ports.begin()
+	   , ce = node->output_control_ports.end();
+	 ci != ce; ++ci) {
+      
+      Port *port = (*ci).second;
+      CDFGEdge *edge = port->edge;
+      Port *user = (*(edge->users.begin()));
+      assert(user);
+      assert(user->parent);
+      
+      if (user->parent->colour != BLACK) {
+	user->parent->colour = BLACK;
+	queue.push_back(user->parent);
+      }
+    }
+  }
+
[…]

### 2010-05-24 toolset/tools/ahir2vhdl/ArbiterBuilder.hpp

@@ -1,0 +1,14 @@
+#ifndef VHDL_ARBITERBUILDER_HPP
+#define VHDL_ARBITERBUILDER_HPP
+
+namespace ahir {
+  class Arbiter;
+}
+
+namespace vhdl {
+  class Arbiter;
+  class DataPath;
+  Arbiter* create_arbiter(ahir::Arbiter *arb, DataPath *dp);
+}
+
+#endif

### 2010-05-24 toolset/tools/ahir2vhdl/LinkLayer.hpp

@@ -1,0 +1,26 @@
+#ifndef __VHDL_LINKLAYER_HPP__
+#define __VHDL_LINKLAYER_HPP__
+
+#include "Entity.hpp"
+
+namespace ahir {
+  class LinkLayer;
+}
+
+namespace vhdl {
+  
+  struct LinkLayer : public Entity
+  {
+    std::string instance_name();
+    std::string component_name();
+    ahir::LinkLayer *aln;
+    
+    LinkLayer(const std::string _id, ahir::LinkLayer *_aln);
+  };
+
+  LinkLayer* create_ln(ahir::LinkLayer *aln);
+
+  void print_ln(LinkLayer *ln);
+}
+
+#endif

### 2010-05-24 toolset/tools/ahir2vhdl/SimpleEntity.hpp

@@ -1,0 +1,20 @@
+#ifndef SIMPLEENTITY_HPP
+#define SIMPLEENTITY_HPP
+
+#include "Entity.hpp"
+
+namespace vhdl {
+
+  struct SimpleEntity : Entity 
+  {
+    SimpleEntity(const std::string instance, const std::string component)
+      : Entity(instance, SIMPLE, component)
+    {}
+
+    std::string instance_name() { return id; }
+    std::string component_name() { return description; } 
+  };
+  
+}
+
+#endif

### 2010-05-24 toolset/tools/ahir2vhdl/System.cpp

@@ -1,0 +1,624 @@
+#include "Module.hpp"
+#include "EntityBuilder.hpp"
+#include "EntityPrinter.hpp"
+#include "ControlPath.hpp"
+#include "DataPath.hpp"
+#include "Arbiter.hpp"
+#include "LinkLayer.hpp"
+#include "Utils.hpp"
+
+#include <Base/ostream.hpp>
+#include <Base/Program.hpp>
+
+#include <fstream>
+#include <sstream>
+
+using namespace vhdl;
+using namespace hls;
+
+namespace vhdl {
+
+  void generate_system(Program *program);
+  
+}
+
+namespace {
+
+  struct Memory : public Entity
+  {
+    std::string component_name() { return id; }
+    std::string instance_name() { return id + "_inst"; }
+
+    Memory(const std::string &id, const std::string &d = "")
+      : Entity(id, MEMORY, d)
+    {}
+  };
+
+  struct System : public Entity 
+  {
+    std::string component_name() { return id; }
+    std::string instance_name() { return id + "_inst"; }
+
+    Memory memory;
+
+    System(const std::string &id, const std::string &d = "")
+      : Entity(id, SYSTEM, d), memory("memory_subsystem", "system memory")
+    {} 
+  };
+  
+  void system_create_memory_wires(System &system
+                                  , unsigned load_lines, unsigned store_lines)
+  {
+    system.register_wire(new Wire("lr_req"
+                                  , vhdl::Type("std_logic_vector"
+                                               , Range(DOWNTO, load_lines))));
+    system.register_wire(new Wire("lr_ack"
+                                  , vhdl::Type("std_logic_vector"
+                                               , Range(DOWNTO, load_lines))));
+    system.register_wire(new Wire("lr_addr"
+                                  , vhdl::Type("std_logic_vector"
+                                               , Range(DOWNTO
+                                                       , load_lines
+                                                       * memory::address_width))));
+    system.register_wire(new Wire("lr_tag"
+                                  , vhdl::Type("std_logic_vector"
+                                               , Range(DOWNTO
+                                                       , load_lines
+                                                       * memory::tag_width))));
+    system.register_wire(new Wire("lc_req"
+                                  , vhdl::Type("std_logic_vector"
+                                               , Range(DOWNTO, load_lines))));
+    system.register_wire(new Wire("lc_ack"
[…]

### 2010-05-24 toolset/tools/ahir2vhdl/VHDLBuilder.hpp

@@ -1,0 +1,16 @@
+#ifndef __VHDL_BUILDER_HPP__
+#define __VHDL_BUILDER_HPP__
+
+#include "typedefs.hpp"
+
+namespace hls {
+  class Program;
+}
+
+namespace vhdl {
+
+  void ahir2vhdl(hls::Program *program);
+  
+}
+
+#endif

### 2010-05-24 toolset/tools/llvm2cdfg/LowerConstantExpr.cpp

@@ -1,0 +1,106 @@
+#include <llvm/Pass.h>
+#include <llvm/Support/InstIterator.h>
+#include <llvm/Instructions.h>
+#include <llvm/DerivedTypes.h>
+#include <llvm/Constants.h>
+#include <llvm/GlobalVariable.h>
+#include <llvm/Transforms/Utils/BasicBlockUtils.h>
+
+#include "Utils.hpp"
+
+#include <iostream>
+#include <fstream>
+#include <sstream>
+#include <numeric>
+#include <functional>
+
+using namespace llvm;
+
+namespace cdfg {
+
+  struct LowerConstantExpr : public FunctionPass {
+    static char ID;
+    virtual bool runOnFunction(Function &F);
+    virtual void getAnalysisUsage(AnalysisUsage &AU) const;
+
+    void process(Instruction *inst);
+    void lower(ConstantExpr *ce, Instruction *inst);
+
+    LowerConstantExpr() : FunctionPass((intptr_t)&ID) {}
+  };
+
+  FunctionPass* createLowerConstantExprPass()
+  {
+    return new LowerConstantExpr();
+  }
+
+  char LowerConstantExpr::ID = 0;
+}
+
+namespace {
+  RegisterPass<cdfg::LowerConstantExpr>
+  X("lower-const-expr", "Lower ConstantExprs to Insts to be eliminated");
+}
+
+void cdfg::LowerConstantExpr::getAnalysisUsage(AnalysisUsage &AU) const {}
+
+bool cdfg::LowerConstantExpr::runOnFunction(Function &F)
+{
+  for (Function::iterator bi = F.begin(), be = F.end(); bi != be; ++bi) {
+    BasicBlock *bb = bi;
+    
+    BasicBlock::iterator ii = bb->begin();
+    while (ii != bb->end()) {
+      Instruction *inst = ii;
+      BasicBlock::iterator II = ii++;
+      
+      process(inst);
+    }
+  }
+
+  return true;
+}
+
+void cdfg::LowerConstantExpr::process(Instruction *inst)
+{
+  for (unsigned oi = 0, oe = inst->getNumOperands();
+       oi != oe; oi++) {
+    llvm::Value *opnd = inst->getOperand(oi);
+    
+    if (ConstantExpr *ce = dyn_cast<ConstantExpr>(opnd)) {
+      lower(ce, inst);
+    }
+  }
+}
+
+
[…]

### 2010-05-24 toolset/tools/llvm2cdfg/LowerGep.cpp

@@ -1,0 +1,150 @@
+#include <llvm/Pass.h>
+#include <llvm/Support/InstIterator.h>
+#include <llvm/Instructions.h>
+#include <llvm/DerivedTypes.h>
+#include <llvm/Constants.h>
+#include <llvm/GlobalVariable.h>
+#include <llvm/Transforms/Utils/BasicBlockUtils.h>
+#include <llvm/Target/TargetData.h>
+#include <llvm/Module.h>
+
+#include "Utils.hpp"
+
+#include <iostream>
+#include <fstream>
+#include <sstream>
+#include <numeric>
+#include <functional>
+
+using namespace llvm;
+using namespace cdfg;
+
+namespace cdfg {
+
+  struct LowerGepPass : public ModulePass {
+    static char ID;
+    virtual bool runOnModule(Module &M);
+    bool runOnFunction(Function &F);
+    virtual void getAnalysisUsage(AnalysisUsage &AU) const;
+
+    TargetData *TD;
+
+    LowerGepPass();
+  };
+
+  ModulePass* createLowerGepPass()
+  {
+    return new LowerGepPass();
+  }
+
+  char LowerGepPass::ID = 0;
+}
+
+namespace {
+  RegisterPass<cdfg::LowerGepPass>
+  X("lower-gep", "Convert GEP into mul-add trees");
+}
+
+LowerGepPass::LowerGepPass() : ModulePass((intptr_t)&ID)
+{
+  TD = NULL;
+};
+
+bool LowerGepPass::runOnModule(Module &M)
+{
+  TD = new TargetData(&M);
+  
+  for (Module::iterator fi = M.begin(), fe = M.end();
+       fi != fe; ++fi) {
+    runOnFunction(*fi);
+  }
+  
+  return false; // we didn't touch anything
+}
+
+void cdfg::LowerGepPass::getAnalysisUsage(AnalysisUsage &AU) const {}
+
+bool cdfg::LowerGepPass::runOnFunction(Function &F)
+{
+  for (Function::iterator bi = F.begin(), be = F.end(); bi != be; ++bi) {
+    BasicBlock *bb = bi;
+
+    BasicBlock::iterator ii = bb->begin();
+    while (ii != bb->end()) {
+      GetElementPtrInst *gep = dyn_cast<GetElementPtrInst>(ii);
+      BasicBlock::iterator gi = ii++;
+      if (!gep) {
+	continue;
+      }
+
+      assert(gep->hasIndices() && "GEP without indices??");
[…]

### 2010-05-24 Documents/Publications/dsd2010/paper.tex
dsd2010: fixed references for AHIR, also put all URLs in footnotes

@@
 \usepackage{times}
+\usepackage{url}
 
@@
 
-\author{
-  Sameer D. Sahasrabuddhe, Kunal P. Ghosh, Sreenivas
-  Subramanian, Kavi Arya and Madhav P. Desai\\
+\author{Sameer D. Sahasrabuddhe, Sreenivas
+  Subramanian, Kunal P. Ghosh, Kavi Arya and Madhav P. Desai\\
   Indian Institute of Technology -- Bombay, Powai, Mumbai -- 400076, INDIA\\
-  Email: \{sameerds@cse, ssreenivas@ee, kunalghosh@ee, kavi@cse, madhav@ee\}.iitb.ac.in
-}
+  Email: \{sameerds@cse, ssreenivas@ee, kunalghosh@ee, kavi@cse,
+  madhav@ee\}.iitb.ac.in }
 \newtheorem{definition}{Definition}[section]
@@
 Several commercial tools which translate a C program to hardware are
-also available.  For example, Catapult-C \cite{catapult-c} from
-Mentor Graphics, C-to-Silicon \cite{c-to-silicon} from Cadence,
-and Synopsys HLS \cite{synopsys-hls} from Synopsys.  Currently,
-we have no access to the internal details of these proprietary tools.
+also available. For example,
+Catapult-C\footnote{\url{http://www.mentor.com/products/esl/high_level_synthesis/}}
+from Mentor Graphics,
+C-to-Silicon\footnote{\url{http://www.cadence.com/products/sd/silicon_compiler/}}
+from Cadence, and Synopsys
+HLS\footnote{\url{http://www.synopsys.com/Tools/SLD/HLS/}} from
+Synopsys. Currently, we have no access to the internal details of
+these proprietary tools.
 
@@
 
-%% Common operators such as equality $(A = B)$, concatenation $(a.b)$ and
-%% prefix $(a \leq b)$ are defined to have sensible meanings. The prefix
-%% operator defines a partial order on label sequences. The longest
-%% common prefix (LCP) of two sequences is the longest sequence that is a
-%% prefix of both the sequences.
-
 \subsection{Labeling scheme}
@@
 
-The synthesis flow uses the LLVM \cite{ref:LLVM}  framework to parse and optimise the
-%
-% http://www.llvm.org
-%
-input C program. This is translated to an AHIR specification using a
-CDFG as an intermediate step.
+The synthesis flow uses the LLVM Compiler
+Infrastructure\footnote{\url{http://llvm.org/}} to parse and optimise
+the input C program. This is translated to an AHIR specification using
+a CDFG as an intermediate step.
 
@@
 The correctness of the second step (CDFG to AHIR) is shown
-as follows \cite{ahir_thesis}: suppose that
+as follows\cite{ahir_thesis}: suppose that
 $A$ was the AHIR specification generated from the
@@
 
-In each case, the VHDL generated by
-the C-to-RTL flow was synthesized using Synposys Design Compiler,
-using the OSU standard cell library \cite{OSU} for the TSMC $0.18\mu$ 
-%
-% vcag.ecen.okstate.edu/projects/scells/
-%
-technology.  
-Memories were modeled using CACTI SRAM models \cite{CACTI}. 
-The resulting net-list was mapped to layout using
-the Cadence SOC Encounter tool.  No human intervention was
-used in the entire process from C to the final layout (other
-than specifying the operating frequency to the automatic place-and-route
-tools).   The total time needed for the entire process (from C to layout)
-was of the order of a few minutes for A5/1 to a few hours for the LPK
-case.
-%
-% Premkishore Shivakumar, Norman P. Jouppi. Cacti 3.0 : An Integreted Cache Timing, Power, and Area model, Western Research
-% Laboratory, 250 University Avenue Palo Alto, California 94301 USA, August 2001
-%
-A simulation based toggle file was then used with the extracted
-net-list to estimate power dissipation in each of the generated
-circuits.
-The reciprocal of the energy used for completing a task is equivalent to
+In each case, the VHDL generated by the C-to-RTL flow was synthesized
+using Synposys Design Compiler, using the OSU Standard Cell
+Library\footnote{\url{http://vcag.ecen.okstate.edu/projects/scells/}}
+for the TSMC $0.18\mu$ technology. Memories were modeled using CACTI
+SRAM
+models\cite{CACTI}\footnote{\url{http://www.hpl.hp.com/research/cacti/}}.
+The resulting net-list was mapped to layout using the Cadence SOC
+Encounter tool. No human intervention was used in the entire process
+from C to the final layout (other than specifying the operating
+frequency to the automatic place-and-route tools). The total time
+needed for the entire process (from C to layout) was of the order of a
+few minutes for A5/1 to a few hours for the LPK case. A simulation
+based toggle file was then used with the extracted net-list to
+estimate power dissipation in each of the generated circuits. The
+reciprocal of the energy used for completing a task is equivalent to
 the throughput achieved for each watt of power supplied, commonly
 termed as {\em performance per watt}
-%
-% TODO: add a comment that the critical path was through 
-% adders/multipliers which were unoptimized!
 
@@
 
-As a reference to judge the energy efficiency of
-the AHIR generated circuits, we consider a 
-popular low power processor (henceforth denoted
-by $P$), namely the Intel Atom N270 processor \cite{ref:Atom},
-%
-% www.intel.com/products/processor/atom/index.htm
-% 
-which is built in Intel's $45nm$ technology.
-This processor is readily available, combines high performance
-with low power dissipation, has full floating point
-support, and is a part of many low power
-computing devices.   
-The sample set of C programs was run on the processor $P$ (exactly
-the same program used in the hardware generation was run on the processor).
-In each case, the number of cycles needed
-to finish one task (as described above) specific to that program was measured. 
-The data
-sets for the programs were small enough to fit into the processor
-cache, and the cycle count was averaged across multiple runs.  These
-cycle counts were then used to find the delay and energy consumption
-for the processor.  All values for area, frequency and
+As a reference to judge the energy efficiency of the AHIR generated
+circuits, we consider a popular low power processor (henceforth
+denoted by $P$), namely the Intel Atom N270
+processor\footnote{\url{http://www.intel.com/products/processor/atom/index.htm}},
+which is built in Intel's $45nm$ technology. This processor is readily
+available, combines high performance with low power dissipation, has
+full floating point support, and is a part of many low power computing
+devices. The sample set of C programs was run on the processor $P$
+(exactly the same program used in the hardware generation was run on
+the processor). In each case, the number of cycles needed to finish
+one task (as described above) specific to that program was measured.
+The data sets for the programs were small enough to fit into the
+processor cache, and the cycle count was averaged across multiple
+runs. These cycle counts were then used to find the delay and energy
+consumption for the processor. All values for area, frequency and
 power dissipation for the processor were obtained from the
@@
 \item Half of the total power dissipation of the $45nm$ circuits
-is due to leakage \cite{ref:Frank}, and the other half is due to switching power. 
+is due to leakage\cite{ref:Frank}, and the other half is due to switching power. 
 Note that this is a conservative estimate, and the actual leakage is
@@
 synthesized from the AHIR generated RTL.
-%
-% Need more elaboration here
-%
 
@@
 
-%
-% Need to revisit this
-%
-
 \bibliography{ref}

### 2010-05-27 Documents/Publications/dsd2010/paper.tex
Modified paper.tex to add a line in the introduction. No other changes
Note: have also prepared paper.ps and sent to IEEE
pdfExpress to convert ot pdf.  If this tex file is
final, we can submit the pdf file generated by
pdfExpress.

@@
 a processor, the platform (that is, the processor) is usually a
-heavily optimized one in which each component has been tweaked to a
+heavily optimized circuit in which each functional
+element has been tweaked to a
 high level, whereas the algorithm-specific
-hardware is generated by automated tools whose optimization abilities
+circuit is generated by automated tools whose optimization abilities
 will most probably not match those of a human designer.   
-Concrete data which throws light on these aspects is valuable.
+Finally, since it is not possible to express parallelism
+directly in a C program (although a compiler can infer it),
+it is not too clear whether a high-level synthesis system can
+identify enough parallelism to make the resulting hardware competitive
+with a heavily optimized processor.  Concrete data which throws light on these aspects is valuable.
 
@@
 obtained when the same program is mapped to a low power processor.
-These results indicate that even in its current state, our high-level
+The results indicate that even in its current state, our high-level
 synthesis flow is an energy competitive alternative to the use of
@@
 bit-level manipulation applications such as A5/1, and
-lowest in the case of random ``pointer-chasing'' code.
+lowest in the case of random ``pointer-chasing'' code.  Note
+that the improvement in the
+case of  floating point programs (FFT and LPK) is also
+substantial.
 The delays in completing the tasks are lower 

### 2010-05-31 Documents/Publications/dsd2010/paper.tex
final version of tex file.

@@
 which maps an algorithm directly to a hardware description can be a
-potential alternative if it provides a verifiable and optimisable path
+potential alternative if it provides a verifiable and optimizable path
 from executable specifications\cite{gajski-executable-specification}
@@
 representation which guarantees the correctness of the implementation,
-and supports optimisations that can scale with the size of the
+and supports optimizations that can scale with the size of the
 program. The synthesis flow introduces an intermediate step in the
@@
 implementation, while also supporting scalable
-optimisations\cite{AHIR_sameerds_hakim_2007}\cite{ahir_thesis}.
+optimizations\cite{AHIR_sameerds_hakim_2007}\cite{ahir_thesis}.
 
@@
 
-There has been a great deal of activity in the field of 
-high-level synthesis in the context of converting a high
-level program to RTL.   Attempts at creating a path from 
-high-level programs to hardware descriptions can be loosely 
-categorised as follows:
+There have been several approaches to the creation of a path from 
+high-level programs to hardware descriptions, which
+can be loosely categorized as follows:
 
@@
 
-Some efforts simply interpret a program as a behavioral
+Some efforts interpret a program as a behavioral
 specification, which is mapped to a hardware implementation using an
@@
 programs into hardware. AHIR differs from both Pegasus and SPARK since
-it factorises the system into three separate components: control flow,
-data flow and memory.  The components can be optimised and
+it factorizes the system into three separate components: control flow,
+data flow and memory.  The components can be optimized and
 implemented separately as long as specified constraints are satisfied.
@@
 Several commercial tools which translate a C program to hardware are
-also available. For example,
+also available, such as 
 Catapult-C\footnote{\url{http://www.mentor.com/products/esl/high_level_synthesis/}}
@@
 The control-path has a single marked place in the initial marking.
-This enables a single transition called \sym{init}, which responds to
+This enables a single transition called {\em init}, which responds to
 a request symbol in alphabet $\Omega$ and begins execution of the
-module. The end of execution is represented by the \sym{fin}
+module. The end of execution is represented by the {\em fin}
 transition that emits an acknowledgment symbol in $\Omega$ and marks
@@
 in \cite{ahir_thesis}), which
-are defined using a set of standard construction rules. 
-The structure of a Type-2 petri-net is
-designed to enable analyses and transformations that are scalable with
+are defined using a set of standard construction rules, which
+ensure that the resulting petri-net is live and safe. 
+Further, the structure of a Type-2 petri-net  
+enables analyses and transformations that are scalable with
 the size of the petri-net.
@@
 avoids arbitration overheads since there is no contention. The
-optimisation uses an almost linear static analysis of the control-path
+optimization uses an almost linear static analysis of the control-path
 to identify sharing opportunities.
@@
 \begin{definition} Two elements $e_1$ and $e_2$ in a Type-2 petri-net
-are compatible if and only if one of the following is true:
-\begin{enumerate}
-\item their NCA is {\em not} a fork region
-\item there is a path within the NCA fork region, joining the two elements
-\end{enumerate}
+are compatible if their NCA is {\em not} a fork region or if
+there is a directed path within the NCA region which joins the two elements.
 \end{definition}
@@
 The compatibility label is a record of every path reaching that
-element from the $init$ transition, which results in an exponential
-size. Comparing two labels for compatibility also has exponential
+element from the $init$ transition, and can thus have an exponential
+size (the number of elements in the label set can be
+exponential in the size of the petri-net). 
+Comparing two labels for compatibility can also have exponential
 complexity, since every sequence in one label has to be compared with
@@
 
-We eliminate the complexity by using a {\em label representation
+We eliminate the exponential complexity by using a {\em label representation
   graph} (LRG) as shown in Fig.~\ref{figure:LRG}. The LRG represents
@@
 
-We have used the LRG to implement arbiterless sharing in the AHIR
+We have used the LRG to implement arbiter-less sharing in the AHIR
 synthesis flow. We use a simple greedy algorithm to generate cliques
@@
 
-This scheme for arbiterless sharing is quite effective in reducing
-hardware costs, demonstrating support for scalable optimisations in
-AHIR. Synthesis results show improvements in the throughput-area ratio
+This scheme for arbiter-less sharing is quite effective in reducing
+hardware costs, demonstrating support for scalable optimizations in
+AHIR. Synthesis results (targeted at field programmable gate arrays)
+show improvements in the throughput-area ratio
 (measured in $\operatorname{Hz}/\operatorname{slice}$) in the range of
@@
 The synthesis flow uses the LLVM Compiler
-Infrastructure\footnote{\url{http://llvm.org/}} to parse and optimise
-the input C program. This is translated to an AHIR specification using
-a CDFG as an intermediate step.
+Framework\footnote{\url{http://llvm.org/}} to parse and optimize the
+input C program. The resulting optimized LLVM bytecode is then
+translated to an AHIR specification using a CDFG as an intermediate
+step.
 
@@
 
-The starting point is a C program.  In the current implementation,
-the only restrictions on the C program are that the call graph
-should not contain any cycles, and that function pointers should
-not be used (pointers to data can be used in the usual way)\footnote{These restrictions
-are not fundamental to the approach, but we have not gotten around to eliminating
-them so far.}.
-The C program is first converted to LLVM byte-code, which is based on
-the SSA form\cite{muchnick}. This is a purely functional form that
-removes the notion of individual {\it variables} from a program. Every
-assignment to a variable is a distinct value; multiple assignments
-that occur in branches are handled by a special instruction called the
+The starting point is a C program. In the current implementation, the
+only restrictions on the C program are that the call graph should not
+contain any cycles, and that function pointers should not be used
+(pointers to data can be used in the usual way)\footnote{These
+  restrictions are not fundamental to the approach, but we have not
+  gotten around to eliminating them so far.}. The C program is first
+converted to LLVM byte-code, which is based on the SSA
+form\cite{muchnick}. This is a purely functional form that removes the
+notion of individual {\it variables} from a program. Every assignment
+to a variable is a distinct value; multiple assignments that occur in
+branches are handled by a special instruction called the
 $\phi$-function, as shown in Fig.~\ref{figure:C-and-SSA}.
@@
 \cite{functional_synthesis_TASS_amellal_kaminska_1994}, as shown in
-Fig.~\ref{fig:snippet-cdfg}. The CDFG represents instructions as
-nodes connected by control and data flow edges. The edges in the CDFG
-arise from three kinds of dependence in the original program:
-
-\begin{itemize}
-\item Data dependence that creates control and data flow.
-\item Control structures that create control flow.
-\item External dependence (in storage) that creates control flow.
-\end{itemize}
+Fig.~\ref{fig:snippet-cdfg}. The CDFG is a hypergraph where the nodes
+represent instructions and edges represent control and data flow. A
+data edge is a hyperedge with a single tail and one or more heads,
+that represents a value in the program, defined by the tail and used
+by the heads. The control edge has a single tail and a single head,
+and represents the passing of control from the tail to the head.
+
+Each value in the program that is defined by some instruction $u$ and
+used by a set of instructions $s_v$ is mapped a data edge in the CDFG
+where the tail $n_u$ corresponds to the instruction $u$ and the set of
+heads has a one to one correspondence with the set $s_v$. Control
+edges are created in a manner that exposes instruction-level
+parallelism in the program. Sequence is enforced only where the
+execution of a node may affect execution of some other node. For
+example, when a node $v$ uses the output of a node $u$, a control edge
+is introduced to ensure that $v$ is executed only when $u$ is
+finished. Similarly, control edges are used to eliminate hazards
+between pairs of load instructions, store instructions and function
+calls that may access the same memory location, as determined by a
+memory reference analysis.
 
@@
 
-The conversion of an AHIR network to VHDL is straightforward.
-Currently, we produce a system which is synchronous and uses
-a single clock (positive edges only).  Each symbol that is
-exchanged between the control path and the data path is coded 
-by a pulse which is sampled high by exactly one clock
-edge.  Symbols are associated with transitions as indicated
-earlier.
+The conversion of an AHIR circuit to VHDL is straightforward.
+Currently, we produce a system which is synchronous and uses a single
+clock (positive edges only). Each symbol exchanged between the control
+path and the data path is coded by a pulse which is sampled high by
+exactly one clock edge. Symbols are associated with transitions as
+indicated earlier.
 
@@
 are its predecessor transitions.
-\item In the rest of the cases, the place is modeled as 
-a finite state machine with two states, with the state being
-set by any of its predecessors, and reset by any of its successors.
+\item In all other cases, the place is modeled as a finite state
+  machine with two states, with the state being set by any of its
+  predecessors, and reset by any of its successors.
 \end{enumerate}
@@
 
-Operators in the data path compute their results in a single
-clock cycle, and the outputs are registered.  Shared operators
-are constructed using multiplexers in a standard manner.
-The instantiation of operators in the data path, 
-and the interconnections within
-the data path and between the data path and the control path
-mirror those in the AHIR representation, except for the fact
-that several operators in the AHIR data path can be mapped
-to the same physical operator (as identified by arbiterless
-operating sharing).
-The inter module link layer is completed using the call graph
-corresponding to the source code.
+Operators in the data path compute their results in a single clock
+cycle, and the outputs are registered. Shared operators are
+constructed using multiplexers in a standard manner. The instantiation
+of operators in the data path, the interconnections within the data
+path and those between the data path and the control path mirror those
+in the AHIR representation, except for the fact that several operators
+in the AHIR data path can be mapped to the same physical operator
+(through arbiterless sharing). The inter module link layer is
+completed using the call graph corresponding to the source code.
 
@@
 the processor). In each case, the number of cycles needed to finish
-one task (as described above) specific to that program was measured.
-The data sets for the programs were small enough to fit into the
-processor cache, and the cycle count was averaged across multiple
-runs. These cycle counts were then used to find the delay and energy
-consumption for the processor. All values for area, frequency and
-power dissipation for the processor were obtained from the
-corresponding data-sheet.
+one task (as described above) was measured. The data sets for the
+programs were small enough to fit into the processor cache, and the
+cycle count was averaged across multiple runs. These cycle counts were
+then used to find the delay and energy consumption for the processor.
+All values for area, frequency and power dissipation for the processor
+were obtained from the corresponding data-sheet.
 
@@
 
-If we compare the results for the $180nm$ AHIR circuits with
-the $45nm$ processor $P$, we find that except for the AES and
-RBT cases,
-the energy consumed per task by the AHIR circuit is appreciably
-lower than that for the processor $P$.  This is significant
-because the processor $P$ is implemented in a technology which
-is four generations ahead of the AHIR circuits.
+We find that the energy consumed per task by the $180nm$ AHIR circuits
+is appreciably lower than that for the $45nm$ processor $P$ (except
+for AES and RBT). This is significant because the processor $P$ is
+implemented in a technology which is four generations ahead of the
+AHIR circuits.
 

### 2010-06-29 Documents/Publications/type-2/type-2.tex
added the missing update to type-2.tex

@@
 
-\title{A class of live and safe petri-nets that supports scalable
-  concurrency analysis}
+\title{A class of untimed live and safe petri-nets that supports
+  scalable static analysis}
 
@@
   Petri-nets are a compact way of representing concurrent sequences of
-  events in a system, where events are represented as transitions in
-  the corresponding petri-net. Given two events in a system, it is
-  useful to determine whether they can occur simultaneously. But it is
-  difficult to make such a determination in an arbitrary petri-net. We
-  present a class of petri-nets called ``Type-2'' which restricts the
-  structure of the petri-net in order to enable a scalable analysis,
-  without severely affecting the expressive power of the petri-net.
-  Given a Type-2 petri-net, it is possible to determine in linear time
-  whether two transitions can fire simultaneously. In addition, it is
-  also possible to generate an exhaustive listing of sets of
+  events in a system, where events are represented as transitions
+  connected to places that encode the state of the system. We present
+  a class of untimed live and safe petri-nets called ``Type-2'' that
+  allows a scalable static analysis, without severely affecting the
+  expressive power of the petri-net.
+
+  The simplest Type-2 petri-net is a cycle consisting of a single
+  marked place and a single transition. Larger petri-nets are produced
+  by progressive application of production rules that replace an
+  unmarked place or a transition with a specific petri-net subgraph
+  with no marked places. The initial marking of a Type-2 petri-net is
+  thus restricted to the single marked place in the initial cycle. The
+  production rules also ensure that every Type-2 petri-net is live and
+  safe.
+
+  The restriction on the structure of the Type-2 petri-net
+  allows the definition of a simple labelling scheme that effectively
+  simulates an execution of the petri-net. The labels of any two
+  transitions can be compared in linear time to statically determine
+  their relation in the execution of the petri-net. For example, the
+  labels can be used determine whether two transitions can fire
+  concurrently, thus identifying concurrent events in the system. It
+  is also possible to generate an exhaustive listing of sets of
   transitions such that no two transitions in each set can fire
@@
 
-  We first define the concept of a ``token-preserving region'' which
-  produces a live and safe petri-net when augmented by a marked place.
-  We further define a set of production rules to combine such regions
-  into larger token-preserving regions. The petri-net created by
-  augmenting such a region is said to be of the class ``Type-2''. The
-  initial marking of the petri-net is restricted to a single marked
-  place.
-
-  The restriction on the structure of the Type-2 petri-net allow a
-  simple labelling scheme to effectively simulates an execution of the
-  petri-net. These labels can be compared in linear time to determine
-  the concurrency of the labelled transitions. Despite the
-  restrictions imposed by the production rules, the Type-2 class of
-  petri-nets is powerful enough to represent synchronous systems. For
-  example, this class has been used as a transition step in
-  translating imperative programs to hardware descriptions.
+  Despite the restrictions imposed by the production rules, the Type-2
+  class of petri-nets is powerful enough to represent synchronous
+  systems. It has been used as a transition step in translating
+  imperative programs to hardware descriptions\cite{ahir-thesis}. The
+  resulting high-level synthesis flow uses the labelling scheme to
+  optimise the generated hardware by identifying opportunities for
+  sharing hardware resources in the absence of timing information.
 

### 2010-07-03 Ahir-asm/antlr2/docs/AaLRM.tex
more progress.  a cleaner version, with more powerful branching support (also hopefully easier to program)

\documentclass{article}

\title{Aa Language Reference Manual}
\author{Madhav Desai \\ Department of Electrical Engineering \\ Indian Institute of Technology \\
	Mumbai 400076 India}

\newcommand{\Aa}{{\bf Aa}~}

\begin{document}
\maketitle

\section{Introduction}

The {\bf Aa} language is intended to be used to
describe an algorithm so that the control flow
in the description can be modeled by a petri net
of a specific class (the type-2 petri-net as
introduced in \cite{ref:SameerPhD}).  A program in
\Aa is also a description of a system which reacts with 
its environment through input and output ports and 
through message queues.

In the rest of this document, we outline the
structure and the syntax of the \Aa language,
and also describe the semantics of an \Aa
program, especially the execution model and
the behaviour of the equivalent system.

\section{Program structure}

A program in \Aa consists of a sequence of
declarations and module definitions.  
\begin{verbatim}
program := ( module-definition | program-object-declaration )*
\end{verbatim}
Declarations at the program level 
can belong to one of three classes:
\begin{itemize}
\item {\bf Storage} objects declared in the program
are globally visible variables that can be read from
anywhere in the program and can be written from anywhere
in the program.
\item {\bf Pipe} objects declared in the program are
globally visible first-in-first-out queues that can 
be read from anywhere in the program. 
written into by at most one module in the program.
\begin{itemize}
\item A pipe can be written into by at most one module
in the program, and can be read from by at most one module
in the program.
\item A pipe that is only written into in the program, but
is not read from, is assumed to have a destination outside
the program.
\item A pipe that is only read from in a program but not
written into, is assumed to have a source outside the program.
\end{itemize}
\item {\bf Constant} objects are globally visible and have
initial values which can never be altered in the program.
Thus, an assignment to a constant is an error.
\end{itemize}

A module in \Aa is the basic unit of compilation, and
has the following structure
\begin{verbatim}
module-definition :=  
                 module-specifier module-name  
                 module-input-arguments
                 module-output-arguments
                 module-object-declarations
                 { <module-body> }
\end{verbatim}
Thus, a module has a name, has input and output arguments,
declares objects and consists of a sequence of statements.
Objects (storage/pipe/constant) declared in the module are visible only in the
module body.

The module body itself is a sequence of statements.
\begin{verbatim}
module-body := (statement)+
\end{verbatim}
The control flow implied is serial in nature.  The
module starts executing when it gets a token.
This token is passed to the first statement in the
list which executes and releases the token when it
finishes execution, and the token  is passed serially down the
list of statements.  When the last statement finishes 
executing, the token passes out of the module.

The region between \{ and \} is termed a {\bf scope}.  Thus, 
the module description defines a scope.  In this case, 
the scope has a label which is the same as the module name.
As we shall see later, the statements in the statement sequence
may in turn define scopes, which may be labeled or anonymous.
Each scope defines a name-space.  The visibility
rules and access mechanisms between scopes will be described
in Section \ref{sec:Scopes}

The control flow in an \Aa module can be visualized
in terms of the flow of tokens.
A module is invoked when it receives a token, and 
this token flows down the sequence of statements.
Whenever a statement receives a token, it executes,
and when it has finished executing, the token is
released to the next statement.

\section{Statements}

Statements in the module-body can be simple statements
or block statements.

\subsection{Simple Statements}

Simple statements can be one of the following:
\begin{itemize}
\item An {\bf assignment} statement is of the form
\begin{verbatim}
target-ref := expression
\end{verbatim}
where {\em target-ref} either specifies a declared object
or an undeclared name.  If it is an undeclared name, then
it is an implicit variable that is attached to the
assignment statement.  No other statement can write to this
variable, but any statement can read from this variable
subject to scoping rules.  Thus, every implicit variable
is such that exactly one statement defines its value.
When control flow reaches an assignment statement, it
evaluates the expression, updates the value of the target
and passes the control flow forward (we will see what this
means in the sequel).
Some examples of simple statements are 
\begin{verbatim}
a  :=  ($not b)
b  :=   (c + (d + e))
\end{verbatim}
etc.  
More details of the syntax are provided in 
Section \ref{sec:Syntax}.
\item A {\bf call} statement is of the form
\begin{verbatim}
call-spec module-name input-argument-values output-argument-targets
\end{verbatim}
The call statement thus specifies a module to which control flow
is to be passed.  The call statement may itself define new implicit
variables through its specified targets (and thus is the only
statement that can modify these newly defined implicit variables).
When control flow reaches the call statement, it forwards the
token to the called module, and finishes when the token exits
the called module.  Recursive calls and cyclical dependencies between
modules through call statements are not permitted in the \Aa language.
Here is an example of a call statement
\begin{verbatim}
// pass control to a module named foo
// foo has two inputs and three outputs
$call foo (p,q)  (r,s,t)
\end{verbatim}
More details of the syntax are provided in 
Section \ref{sec:Syntax}.
\item The {\bf null} statement does nothing, and just passes the
token onwards.  This is how it looks like
\begin{verbatim}
$null
\end{verbatim}
\end{itemize}

\subsection{Block Statements}

Block statements can be used to describe complex
control flows, and consist of a sequence of statements.
Block statements are of
the following types:
\begin{itemize}
\item {\bf Series} block statements are of the form
\begin{verbatim}
series-block-statement :=
             series-block-specifier block-name
             { 
                declarations
                sequence-of-statements
             }
\end{verbatim}
The behaviour is similar to the module, in that the control
token flows serially down the sequence of statements.
Here is an example:
\begin{verbatim}
$seriesblock [s1] 
{
   $declare $storage b $uint<32>
   b := a
   b := ( $mux (a > c)  (b+c) (b-c))
   q := ( b * 2)
}
\end{verbatim}
In this example, the variable $b$ was declared as
a storage variable in scope s1, and hence was legally
able to be the target of multiple assignment statements.
\item {\bf Parallel} block statements are of the form
\begin{verbatim}
parallel-block-statement :=
             parallel-block-specifier block-name
             { 
                declarations
                sequence-of-statements
             }
\end{verbatim}
When a token enters the parallel block statement, it
is replicated into as many tokens as there are statements in
the sequence, and all statements are started in parallel.
When all statements have finished and released their token,
the parallel block statement
ends and a single token exits the parallel block statement (this
is essentially a fork followed by a join).
Here is an example
\begin{verbatim}
$parallelblock [p1] 
{
   b := a + c
   d := a - c
}
\end{verbatim}
The two statements will start in parallel, and the block 
will finish when both have finished.  The
order in which the two statements are executed
is {\bf not} specified.
\item {\bf Fork} block statements are generalizations of the
parallel block region in the sense that they
allow the programmer to express complex fork-join interactions.
They have the following structure:
\begin{verbatim}
fork-block-statement :=
             fork-block-specifier block-name
             {
                declarations
                sequence-of-statements
             }
\end{verbatim}
The control flow is similar to the parallel block region,
in the sense that all statements in the sequence are started
in parallel, and the fork block statement ends when all statements
in the sequence have ended.  The difference is that
fork blocks can have an additional statement which
allows the programmer to provide additional synchronization
points.  These statements are termed {\bf join} statements
and have the form
\begin{verbatim}
join-statement :=
         join-specifier list-of-labels 
          fork-specifier  list-of-statements
\end{verbatim}
The meaning of this statement is that it waits until all statements
in the list-of-labels have finished, then starts the list-of-statements
(in parallel) and finishes.  This essentially defines a node in a directed graph,
with the arcs corresponding to statements in the statement-lists.  
The join indicates the arcs (or statements) 
from whom tokens have to arrive.  When all tokens have arrived
on the incoming arcs, tokens are sent along the outgoing arcs specified
in the list of statements following the optional fork specifier.  There is an implicit fork
at the entry point to the fork block statement and an implicit join at the
end of the fork block statement.  It is thus easy to describe
an arbitrary directed graph with arbitrary forks and joins.  The only
restriction is that this directed graph must be acyclic!
Here is an example:
\begin{verbatim}
$forkblock [f1] {
  a := (b+c)
  $seriesblock [s1] { ... }
  $seriesblock [s2] { ... }
  $seriesblock [s3] { ... }
  $join s1 s2 $and $fork 
      $seriesblock [s4] { ... }
  $join s3 s4
}
\end{verbatim}
In this example, the first assignment statement as well as s1, s2, s3 are 
started in parallel. When both s1 and s2 have finished, s4 is started and
when s3, s4 have joined and when the first assignment statement has
finished, the forkblock f1 finishes execution.
\item {\bf Branch} block statements are constructs which allow the programmer
to describe arbitrary sequential branching behaviour.  They 
describe a control flow in which a single token 
is active within the block.  The movement of this token is
controlled by special flow control statements.
They are structured 
as follows
\begin{verbatim}
branch-block-statement :=
             branch-block-specifier block-name
             { sequence-of-statements}
\end{verbatim}
The sequence of statements appearing in a branch block
consist of
\begin{itemize}
\item Simple statements or Block statements.
\item Switch statements: A switch statement has
the form
\begin{verbatim}
switch-statement :=
         switch-spec switch-expression
           ( expr-value  statement-to-be-executed )*
           ( default statement-to-be-executed )?

\end{verbatim}
The switch-expression is checked, and depending on its
value, one of the alternative statements is selected.
Thus, the incoming token to the switch statement is 
passed to one of the alternatives.  For example,
\begin{verbatim}
$switch a $when 0 $then a1 := (b + c)
          $when 1 $then a2 := (b - c)
          $default $null
$endswitch
\end{verbatim}
The control token is routed to the appropriate
choice and if the token is not routed out 
of the block (by the {\bf takeroute} statement 
described below), then the token passes to the
statement following the switch. 
\item If statements: An if statement has the form
\begin{verbatim}
if-statement :=
       if-spec test-expression 
           true-statement
       (else
           false-statement)?

\end{verbatim}
For example, 
\begin{verbatim}
$if (a != 0) $then
   q := (r + s)
   t := 0
$else
   qdash :  (r - s)
$endif
\end{verbatim}
\item Routing statement: The {\bf takeroute} statement
defines a labeled path along which it sends the
incoming control token.  For example,
\begin{verbatim}
$takeroute [fastpath]
\end{verbatim}
means that the incoming token is not sent out to the
next statement in the sequence, but is sent out along
a path labeled ``fastpath''.  The token will be captured
either by a {\bf merge} statement, or will exit the
containing branch block.
\item Merge statements: A merge statement is
specified as 
\begin{verbatim}
merge-statement :=
          merge-spec label-list merge-assignments
\end{verbatim}
which is to be interpreted as follows.  The labels in the label list
refer to paths defined by ``takeroute'' statements within the branch block.
Whenever a token arrives along any of these paths,
the merge statement starts and executes a series of special assignments
which multiplex values into new variables based on which arc the
token arrived from.  The merge statement then releases its token to
the next statement. 
\end{itemize}

Here is an example of a branch block constructed in this manner
\begin{verbatim}
$branchblock [b1] 
{
  $merge loop $entry // $entry is the path 
                     // into the block

       $phi q := 0 on $entry r on loop 
                    // q is defined by where 
                    // the token came from
  r := q + 1
  $if r < 10 $then 
       $takeroute [loop]
  $endif
}
\end{verbatim}

\end{itemize}

Thus, the program is constructed as a collection of modules,
each of which is a sequence of statements.  The use of series,
parallel, fork and branch block
statements enables the programmer to describe a highly concurrent
structured system with complex branching behaviour.  
The resulting control flow structure is a
type-2 petri-net with provable liveness and safety properties \cite{ref:SameerPhD}.

\section{Scoping Rules}

An \Aa program is made of modules which in turn contain statements
and so on.  The program thus has a hierarchy of scopes with each
scope being contained in another (except for the program itself,
which is not contained in any scope).

The rules for scoping are as follows:
\begin{enumerate}
\item A variable (declared or implicit) defined in a scope is visible
in all descendant scopes.
\item A reference to a variable $b$ in a scope X is resolved by checking
whether the variable is defined in that scope, and if not found there,
by checking in the scope that contains the scope X, and so on.
\item A scope can read from variables that are defined in descendant
scopes.  
\item A scope can read from variables that are defined in ancestor scopes.
\end{enumerate}

A variable which is to be read in a statement may be specified 
as follows
\begin{verbatim}
a            look for variable a in current scope; 
             if not found look in the parent.

:a           same as the previous case

^:a          look for variable a defined in the parent of
             the current scope.

%p:a         look for variable a defined in the child scope 
	     with label p (child scope of the current scope).

^^:a         look for variable a defined in the parent of
             the parent of the current scope.

%p%q:a       look for variable a defined in the child scope 
             with label q of the child scope with label p of 
             the current scope

a[10]        look for a storage variable a defined in the 
             current module.  If not found, look for it in
             the program scope.  If found in either of the places,
             access the corresponding element of the array.
\end{verbatim}
Thus, a generic variable reference has the form 
\begin{verbatim} 
    scope-id : variable-reference 
\end{verbatim} 
and the scoping rules forbid a scope from accessing variables
which are not defined in either an ancestor or a descendant
of the scope.

\section{Syntax}

The syntax for \Aa is described by the following lexical analysis and
parsing rules (written as rules to be parsed by antlr2 \cite{ref:antlr2},
a LL(k) parser).

This is the set of tokens recognized by the lexical analyzer (or lexer).
\input{AaLexer}
As one can see, all keywords in the \Aa language begin with \$.

The grammar for the parser is (using the EBNF notation) given below. 
\input{AaParser}

\section{Examples}

Here is a very simple program
\begin{verbatim}
// an array of 32-bit unsigned integers.
$storage mem $array<1024> of $uint<32> 
$module [sel_mod]
   $in (a $uint<32> b $uint<10>)
   $out (c $uint<32>)
$is
{
   c := (mem[b] + a)
   mem[b] := c
}
\end{verbatim}
This consists of a single module, which accumulates
a value into an array position.

A little bit more complicated:
\begin{verbatim}
// an array of 32-bit unsigned integers.
$storage mem $array<1024> of $uint<32>

// module returns the sum of mem[I] from
// I=low to I=high
$module [sum_mod]
    $in (low $uint<10> high $uint<10>)
    $out (sum $uint<32>)
$is
    $storage I $uint<10>
    $storage J $uint<10>
    $storage sum1 $uint<32>
    $storage sum2 $uint<32>
{

   d := (high-low)  
   mp := ((high-low)/2)

   I := low
   J := (mp + 1)

   \branchblock[trivcheck]
    {
       $merge $entry
       $if (d > 0) $then
           // do two summations in parallel
           // parallel summations
           \parallelblock[parsum]
           {
               \branchblock[sb1]{
                    merge $entry loopback
                       $phi s := 0 on $entry %forward:s on loopback
                    $if (I < mp) $then
                        \seriesblock[loopback]{
                            I := (I+1)
                            s := $mux (I == 0)  0 (s + mem[I])
                         }
               } 
               \branchblock[sb2]{
                    merge $entry loopback
                       $phi s := 0 on $entry %forward:s on loopback
                    $if (J < high) $then
                        \seriesblock[loopback]{
                            J := (J+1)
                            s := $mux (J == (mp+1))  0 (s + mem[J])
                         }
               } 
           }

           // combine results from parallel statement above
           snontriv := (%parsum%sb1:s  + %parsum%sb2:s)
           $takeroute [nontrivsum]
      $else

           // summation is trivial 
           striv := mem[low]
           $takeroute [trivsum]
      $merge nontrivsum trivsum
         // which sum do you pick? depends on which path was taken
         $phi sum := snontriv on nontrivsum strivsum on trivsum
      $exit
    }

    // bring sum up to the surface.
    sum := %trivcheck:sum
}
\end{verbatim}

\end{document}

### 2010-07-06 Ahir-asm/antlr2/docs/AaLRM.tex
Some more changes to the grammar to rationalize the syntax.  Introduced a $place statement to enable branches.  Also started adding link code in the parsing phase.

@@
 
-The {\bf Aa} language is intended to be used to
-describe an algorithm so that the control flow
-in the description can be modeled by a petri net
+{\bf Aa} is intended to be a low-level 
+programming language for the description
+of algorithms in which the control flow
+can be modeled by a petri net
 of a specific class (the type-2 petri-net as
 introduced in \cite{ref:SameerPhD}).  A program in
-\Aa is also a description of a system which reacts with 
+\Aa can also be viewed as a 
+description of a system which reacts with 
 its environment through input and output ports and 
-through message queues.
+through message queues.  Thus, an \Aa program
+can be directly executed on a computer, and
+can also be mapped to a logic circuit.
 
@@
 \end{verbatim}
-The control flow implied is serial in nature.  The
-module starts executing when it gets a token.
-This token is passed to the first statement in the
-list which executes and releases the token when it
-finishes execution, and the token  is passed serially down the
-list of statements.  When the last statement finishes 
-executing, the token passes out of the module.
+Each statement can be viewed as a transition
+in a petri-net, and has a 
+set of input (or source) places  and a set of output
+(or sink)  places.  
+The execution of the statement is triggered when
+there is a token present in each of its source 
+places and when the statement finishes execution,
+a token is placed in each of its output places.
+The control flow in the sequence of statements
+in a module is serial in nature.  Thus each
+statement has a single source place and a single
+sink place, sink place of a statement being 
+the source place of its successor.
+The module as a whole can be viewed as having
+a single entry place and a single exit place.
+When a token appears in the entry place, the first
+statement can trigger, and when the last statement
+has finished, a token is placed in the exit
+place.
 
@@
 
-The control flow in an \Aa module can be visualized
-in terms of the flow of tokens.
-A module is invoked when it receives a token, and 
-this token flows down the sequence of statements.
-Whenever a statement receives a token, it executes,
-and when it has finished executing, the token is
-released to the next statement.
-
 \section{Statements}
 
-Statements in the module-body can be simple statements
-or block statements.
+Statements in the module-body are can be of two kinds:
+\begin{itemize}
+\item Atomic statements: those that have a single
+entry place and a single exit place.  These are
+the only statements that can occur in a module body.
+These statements are further divided into 
+{\em simple} statements and {\em block} statements.
+\item Sub-atomic statements:  these statements
+can appear only inside block statements,  and they
+can have multiple sources and multiple sinks.
+Examples of these statements are statements for
+combining tokens (merge and join statements) and
+statements for redirecting tokens (the toss statement).
+\end{itemize}
 
@@
                 declarations
-                sequence-of-statements
+                sequence-of-atomic-statements
              }
@@
 an arbitrary directed graph with arbitrary forks and joins.  The only
-restriction is that this directed graph must be acyclic!
+restriction is that this directed graph must be acyclic!  This is enforced
+by requiring that a join statement refers only to labels of statements
+that appear before the join.
 Here is an example:
@@
          switch-spec switch-expression
-           ( expr-value  statement-to-be-executed )*
-           ( default statement-to-be-executed )?
+           ( expr-value  serial-statement-sequence )*
+           ( default serial-statement-sequence )?
 
@@
 The control token is routed to the appropriate
-choice and if the token is not routed out 
-of the block (by the {\bf takeroute} statement 
+choice sequence and if the token is not routed out 
+of the block (by the {\bf place} statement 
 described below), then the token passes to the
@@
        if-spec test-expression 
-           true-statement
+           serial-statement-sequence
        (else
-           false-statement)?
+           serial-statement-sequence)?
 
 \end{verbatim}
-For example, 
+For example:
 \begin{verbatim}
@@
 \end{verbatim}
-\item Routing statement: The {\bf takeroute} statement
-defines a labeled path along which it sends the
-incoming control token.  For example,
+If the control token reaches the end of a selected
+segment in the if statement (that is, without being
+rerouted by a place statement), then the control
+token is passed on to the statement immediately
+following the if statement.
+\item The place statement: The {\bf place} statement
+defines a place with a label  into which it places
+a token after it has executed.  For example,
 \begin{verbatim}
-$takeroute [fastpath]
+$place [fastpath]
 \end{verbatim}
-means that the incoming token is not sent out to the
-next statement in the sequence, but is sent out along
-a path labeled ``fastpath''.  The token will be captured
-either by a {\bf merge} statement, or will exit the
-containing branch block.
+means that the incoming token is placed in a labeled
+place ``fastpath''.  The token will either
+trigger a {\bf merge} statement which is required
+to depend on this labeled place in the same branch
+block, or if no such merge statement
+exists, then ``fastpath'' is identified
+with the exit place of the  containing branch block.
 \item Merge statements: A merge statement is
@@
 which is to be interpreted as follows.  The labels in the label list
-refer to paths defined by ``takeroute'' statements within the branch block.
-Whenever a token arrives along any of these paths,
+refer to token labels defined by ``toss'' statements within the branch block.
+Whenever a token labeled by any of the items in label-list arrives
+at the input of the merge,
 the merge statement starts and executes a series of special assignments
@@
 {
-  $merge loop $entry // $entry is the path 
+  $merge loop $entry // $entry is the place
                      // into the block
+                     // merge triggered by
+                     // token at entry or
+                     // by token in place "loop"
 
@@
   $if r < 10 $then 
-       $takeroute [loop]
+       $place [loop] // put token in place "loop"
   $endif
@@
 \end{verbatim}
-
+This is to be interpreted as follows: the merge executes
+whenever a token is present  either in the entry place
+of the branch-block or in the place labeled ``loop''.
+The merge defines variable ``q'' with a value which is
+either 0 (if a token was present in the entry place)
+or r (if a token was present in the ``loop'' place).
+By the construction rules in an \Aa program, it is
+impossible for there to be a token present in more
+than one input place to a merge.
 \end{itemize}
 
-Thus, the program is constructed as a collection of modules,
+Thus, an \Aa program is constructed as a collection of modules,
 each of which is a sequence of statements.  The use of series,
@@
 
-\section{Scoping Rules}
+\section{Scoping Rules} \label{sec:Scopes}
 
@@
 
-\section{Syntax}
-
-The syntax for \Aa is described by the following lexical analysis and
-parsing rules (written as rules to be parsed by antlr2 \cite{ref:antlr2},
-a LL(k) parser).
-
-This is the set of tokens recognized by the lexical analyzer (or lexer).
-\input{AaLexer}
-As one can see, all keywords in the \Aa language begin with \$.
-
-The grammar for the parser is (using the EBNF notation) given below. 
-\input{AaParser}
-
 \section{Examples}
@@
 
-A little bit more complicated:
+An example which is a little bit more complicated:
 \begin{verbatim}
 // an array of 32-bit unsigned integers.
-$storage mem $array<1024> of $uint<32>
+$declare $storage mem $array<1024> $of $uint<32>
 
@@
 $is
-    $storage I $uint<10>
-    $storage J $uint<10>
-    $storage sum1 $uint<32>
-    $storage sum2 $uint<32>
 {
@@
 
-   I := low
-   J := (mp + 1)
-
-   \branchblock[trivcheck]
+   $branchblock[trivcheck]
     {
-       $merge $entry
+       // d from parent scope
        $if (d > 0) $then  
@@
            // parallel summations
-           \parallelblock[parsum]
+           $parallelblock[parsum]
            {
-               \branchblock[sb1]{
-                    merge $entry loopback
-                       $phi s := 0 on $entry %forward:s on loopback
+               $branchblock[sb1]{
+                    $declare $storage  I $uint<10> := 0
+                    $merge $entry loopback
+                       $phi s := 0 $on $entry  s1 $on loopback
+                    $endmerge
                     $if (I < mp) $then
-                        \seriesblock[loopback]{
                        I := (I+1)
-                            s := $mux (I == 0)  0 (s + mem[I])
+                       s1 := ($mux (I == 0)  0 (s + mem[I]))
+                       $place [loopback]
+                    $endif
                } 
-               } 
-               \branchblock[sb2]{
-                    merge $entry loopback
-                       $phi s := 0 on $entry %forward:s on loopback
+               $branchblock[sb2]{
+                    $declare $storage  J $uint<10> := 0
+                    J := (mp + 1)
+                    $merge $entry loopback
+                       $phi s := 0 $on $entry s1 $on loopback
+                    $endmerge
                     $if (J < high) $then
-                        \seriesblock[loopback]{
                          J := (J+1)
-                            s := $mux (J == (mp+1))  0 (s + mem[J])
-                         }
+                         s1 := ($mux (J == (mp+1))  0 (s + mem[J]))
+                         $place [loopback]
+                    $endif
                } 
@@
            snontriv := (%parsum%sb1:s  + %parsum%sb2:s)
-           $takeroute [nontrivsum]
+           $place [nontrivsum]
       $else
@@
            striv := mem[low]
-           $takeroute [trivsum]
+           $place [trivsum]
+      $endif
       $merge nontrivsum trivsum
          // which sum do you pick? depends on which path was taken
-         $phi sum := snontriv on nontrivsum strivsum on trivsum
-      $exit
+         $phi sum := snontriv $on nontrivsum strivsum $on trivsum
+      $endmerge
     }
@@
 \end{verbatim}
+This example describes an algorithm which computes the sum of
+a section of an array by dividing the problem into two partial
+summations.
+
+\section{Syntax} \label{sec:Syntax}
+
+The syntax for \Aa follows the following
+principles
+\begin{itemize}
+\item All keywords begin with the \$ sign.
+\item The region between \{ and \} defines a new scope.
+\item Statements are space separated (no semicolons at all).
+\item Expressions are fully parenthitized.  Thus
+$(a + b)$ is a legal expression, but $a+b$ is not.
+\end{itemize}
+
+The parser is implemented using 
+an LL(k) parser (written as rules to be parsed by antlr2 \cite{ref:antlr2}).
+The grammar for the parser is (using the EBNF notation) given below. 
+\input{AaParser}
+
+This is the set of tokens recognized by the lexical analyzer (or lexer).
+\input{AaLexer}
+
+\begin{thebibliography}{99}
+\bibitem{ref:SameerPhD}
+Sameer D. Sahasrabuddhe,
+``A competitive pathway from high-level programs to hardware,''
+Ph.D. thesis, IIT Bombay, 2009.
+
+\bibitem{ref:antlr2}
+http://www.antlr2.org.
+\end{thebibliography}
 

### 2010-07-06 Ahir-asm/antlr2/include/AaObject.h

@@ -1,0 +1,93 @@
+#ifndef _Aa_Object__
+#define _Aa_Object__
+
+#include <AaIncludes.h>
+#include <AaUtil.h>
+#include <AaRoot.h>
+#include <AaScope.h>
+#include <AaType.h>
+#include <AaValue.h>
+#include <AaExpression.h>
+
+
+/*****************************************  OBJECT  ****************************/
+// base object
+// Each object has a type, a name, a value and a parent
+
+class AaObject: public AaRoot
+{
+  string _name;
+  AaConstantLiteralReference* _value;
+  AaType* _object_type;
+  AaScope* _scope;
+
+ public:
+
+  AaType* Get_Object_Type() {return(this->_object_type);}
+  AaConstantLiteralReference* Get_Value() {return(this->_value);}
+  void Set_Value(AaConstantLiteralReference* v) {this->_value = v;}
+  virtual string Get_Name() {return(this->_name);}
+  virtual AaScope* Get_Scope() {return(this->_scope);}
+  virtual string Tab();
+
+  AaObject(AaScope* scope_tpr, string oname, AaType* object_type);
+  ~AaObject();
+
+  virtual void Print(ostream& ofile);
+  virtual string Kind() {return("AaObject");}
+};
+
+// interface object: function arguments
+class AaInterfaceObject: public AaObject
+{
+  // arguments can be input or output
+  string _mode; // IN or OUT?
+
+ public:
+  string Get_Mode() {return(this->_mode);}
+
+  AaInterfaceObject(AaScope* scope_tpr, string oname, AaType* otype, string mode);
+  ~AaInterfaceObject();
+  virtual string Kind() {return("AaInterfaceObject");}
+  // uses AaObject::Print method
+};
+
+class AaConstantObject: public AaObject
+{
+  // 
+ // constants are visible only in pure ancestors
+  // or pure descendants.
+  //
+ public:
+  AaConstantObject(AaScope* scope_tpr, string oname, AaType* otype, AaConstantLiteralReference* value);
+  ~AaConstantObject();
+
+  virtual void Print(ostream& ofile); 
+  virtual string Kind() {return("AaConstantObject");}
+};
+
+class AaStorageObject: public AaObject
+{
+  // objects will be stored in memory
+  //
+ public:
+
+  AaStorageObject(AaScope* scope_tpr, string oname, AaType* otype, AaConstantLiteralReference* initial_value);
[…]

### 2010-07-06 Ahir-asm/antlr2/include/AaStatement.h

@@ -1,0 +1,313 @@
+#ifndef _Aa_Statement__
+#define _Aa_Statement__
+
+#include <AaIncludes.h>
+#include <AaUtil.h>
+#include <AaRoot.h>
+#include <AaScope.h>
+#include <AaType.h>
+#include <AaValue.h>
+#include <AaExpression.h>
+#include <AaObject.h>
+
+
+// *******************************************  STATEMENT etc. ************************************
+// base class for statements
+// A statement may specify a set of targets which act as
+// implicit variable declarations.  These need to be 
+// mapped in the containing scope
+class AaStatement: public AaScope
+{
+ public:
+  AaStatement(AaScope* scope);
+  
+  // add map entries in parent´s lookup map for easy access
+  virtual void Map_Targets() 
+  {
+    // derived statements will map their targets
+    // in their containing scopes.
+    assert(0); 
+  }
+
+  virtual void Map_Target(AaObjectReference* obj_ref);
+  virtual string Tab();
+
+  ~AaStatement();
+
+  virtual string Kind() {return("AaStatement");}
+};
+
+// statement sequence (is used in block statements which lead to programs)
+class AaStatementSequence: public AaScope
+{
+  vector<AaStatement*> _statement_sequence;
+ public:
+
+  AaStatementSequence(AaScope* scope, vector<AaStatement*>& statement_sequence);
+  ~AaStatementSequence();
+
+  unsigned int Get_Statement_Count() { return this->_statement_sequence.size(); }
+  virtual void Print(ostream& ofile);
+  virtual string Kind() {return("AaStatementSequence");}
+};
+
+// null statement
+class AaNullStatement: public AaStatement
+{
+ public:
+  AaNullStatement(AaScope* parent_tpr);
+  ~AaNullStatement();
+
+  virtual void Print(ostream& ofile) { ofile << this->Tab() << "$null" << endl; }
+  virtual string Kind() {return("AaNullStatement");}
+};
+
+
+// assignment statement
+class AaAssignmentStatement: public AaStatement
+{
+  AaObjectReference* _target;
+  AaExpression* _source;
+ public:
+  AaObjectReference* Get_Target() {return(this->_target);}
+  AaExpression* Get_Source() {return(this->_source);}
+
+  virtual void Map_Targets() 
+  {
[…]

### 2010-07-06 Ahir-asm/antlr2/include/AaType.h

@@ -1,0 +1,108 @@
+#ifndef _Aa_Type__
+#define _Aa_Type__
+
+#include <AaIncludes.h>
+#include <AaUtil.h>
+#include <AaRoot.h>
+#include <AaScope.h>
+
+
+/*****************************************  TYPE ****************************/
+
+class AaType: public AaRoot
+{
+  AaScope* _scope;
+  // types dont have names, only declarations.
+ public:
+  virtual AaScope* Get_Scope() {return(this->_scope);}
+
+  AaType(AaScope* parent);
+  ~AaType();
+  virtual string Kind() {return("AaType");}
+};
+
+class AaScalarType: public AaType
+{
+
+ public:
+  AaScalarType(AaScope* parent);
+  ~AaScalarType();
+  virtual string Kind() {return("AaScalarType");}
+};
+
+
+class AaUintType: public AaScalarType
+{
+  // width > 0
+  unsigned int _width;
+
+ public:
+  virtual unsigned int Get_Width() {return(this->_width);}
+
+  AaUintType(AaScope* scope, unsigned int width);
+  ~AaUintType();
+  void Print(ostream& ofile);
+  virtual string Kind() {return("AaUintType");}
+};
+
+class AaIntType: public AaUintType
+{
+  // gets width from Uint
+
+ public:
+  AaIntType(AaScope* scope,unsigned int width);
+  ~AaIntType();
+  void Print(ostream& ofile);
+  virtual string Kind() {return("AaIntType");}
+};
+
+class AaPointerType: public AaUintType
+{
+ public :
+  AaPointerType(AaScope* scope, unsigned int _object_width);
+  ~AaPointerType();
+
+  virtual void Print(ostream& ofile);
+  virtual string Kind() {return("AaPointerType");}
+};
+
+
+class AaFloatType : public AaScalarType
+{
+  // both > 0
+  unsigned int _characteristic;
+  unsigned int _mantissa;
+
+ public:
+  unsigned int Get_Characteristic() {return(this->_characteristic);}
+  unsigned int Get_Mantissa() {return(this->_mantissa);}
+
+  AaFloatType(AaScope* scope, unsigned int characteristic, unsigned int mantissa);
+  ~AaFloatType();
+  void Print(ostream& ofile);
+  virtual string Kind() {return("AaFloatType");}
+};
+
+class AaArrayType: public AaType
+{
+  // multi-dimensional array types are possible
[…]

### 2010-07-06 Documents/Publications/type-2/type-2.tex
modified the abstract..

@@
 
-\title{A class of untimed live and safe petri-nets that supports
-  scalable static analysis}
+\title{On the use of petri-nets for modeling concurrent programs}
 
@@
 
-  Petri-nets are a compact way of representing concurrent sequences of
-  events in a system, where events are represented as transitions
-  connected to places that encode the state of the system. We present
-  a class of untimed live and safe petri-nets called ``Type-2'' that
-  allows a scalable static analysis, without severely affecting the
-  expressive power of the petri-net.
-
-  The simplest Type-2 petri-net is a cycle consisting of a single
-  marked place and a single transition. Larger petri-nets are produced
-  by progressive application of production rules that replace an
-  unmarked place or a transition with a specific petri-net subgraph
-  with no marked places. The initial marking of a Type-2 petri-net is
-  thus restricted to the single marked place in the initial cycle. The
-  production rules also ensure that every Type-2 petri-net is live and
-  safe.
-
-  The restriction on the structure of the Type-2 petri-net
-  allows the definition of a simple labelling scheme that effectively
-  simulates an execution of the petri-net. The labels of any two
-  transitions can be compared in linear time to statically determine
-  their relation in the execution of the petri-net. For example, the
-  labels can be used determine whether two transitions can fire
-  concurrently, thus identifying concurrent events in the system. It
-  is also possible to generate an exhaustive listing of sets of
-  transitions such that no two transitions in each set can fire
-  simultaneously.
-
+  Petri-nets provide a powerful mechanism for the modeling of
+  concurrent systems.   However, because of their generality, 
+  it is often difficulty to prove properties about petri-net
+  models unless some restrictions are placed on their construction.
+  We consider the use of petri-nets to model algorithms described
+  in high level programming languages.  In particular, we introduce
+  a class of live and safe petri-nets (which we term as {\em Type-2}
+  petri-nets) for this purpose.   This class is defined by a
+  set of production rules which ensure liveness and safety of
+  the resulting petri-net.  
+  Further, we show that concurrency
+  analysis of Type-2 petri-nets can be carried out very efficiently
+  using a labelling scheme which is analogous to a symbolic simulation
+  of the petri-net.
+  Thus, it is easy to determine the sets of events (or transitions)
+  which satisfy ordering relationships (two transitions are said to
+  be ordered if it is impossible for both of them to fire simultaneously).  
+  Such information is useful in identifying possibilities for conflict 
+  free resource sharing in the implementations of the programs being
+  modeled.
   Despite the restrictions imposed by the production rules, the Type-2
-  class of petri-nets is powerful enough to represent synchronous
-  systems. It has been used as a transition step in translating
+  class of petri-nets is powerful enough to represent programs
+  written in traditional imperative languages such as C/C++ as well in synchronous
+  languages such as Esterel. It has been used as a transition step in translating
   imperative programs to hardware descriptions\cite{ahir-thesis}. The

### 2010-07-12 Ahir-asm/antlr2/docs/AaLRM.tex
some modifications to the program structure to prepare for the following 1. cycle check in call graph 2. type propagation Added a BGL wrapper class to get dfs functionality...

@@
 \begin{enumerate}
-\item A variable (declared or implicit) defined in a scope is visible
-in all descendant scopes.
+\item A declared variable defined in a scope is visible
+in all descendant scopes. In addition
+\begin{itemize}
+\item If $x$ is a storage or pipe variable defined in a scope, then
+there can be no other variable named $x$ in any of the scopes
+contained in this scope.
+\item If $x$ is an input or output argument of a module, then
+there can be no other variable named $x$ declared in any of
+the descendant scopes of this module.
+\end{itemize}
 \item A reference to a variable $b$ in a scope X is resolved by checking
@@
 \item A scope can read from variables that are defined in ancestor scopes.
+\item A scope can only write to one of the following:
+\begin{itemize}
+\item a storage or pipe variable defined in the scope or an ancestor
+of the scope.
+\item an output argument of a module of which the scope is a descendant.
+In this case, there can be at most one statement in the entire module
+which writes to this output argument.
+\end{itemize}
 \end{enumerate}
 
-A variable which is to be read in a statement may be specified 
+A variable reference in a statement may be specified 
 as follows
@@
 
-^:a          look for variable a defined in the parent of
+^:a          look for variable a starting from the parent of
              the current scope.
 
-%p:a         look for variable a defined in the child scope 
+%p:a         look for variable a starting from the child scope 
 	     with label p (child scope of the current scope).
 
-^^:a         look for variable a defined in the parent of
+^^:a         look for variable starting from the parent of
              the parent of the current scope.
 
-%p%q:a       look for variable a defined in the child scope 
+%p%q:a       look for variable starting from the child scope 
              with label q of the child scope with label p of 
@@
 a[10]        look for a storage variable a defined in the 
-             current module.  If not found, look for it in
-             the program scope.  If found in either of the places,
+             current scope.  If not found, look for it in
+             the parent scope, and so on.  If eventually found, 
              access the corresponding element of the array.
@@
            snontriv := (%parsum%sb1:s  + %parsum%sb2:s)
+           // put token in place nontrivsum.  It will be picked
+           // up by succeeding merge
            $place [nontrivsum]
@@
            striv := mem[low]
+           // put token in place trivsum.  It will be picked
+           // up by succeeding merge
            $place [trivsum]
@@
          // which sum do you pick? depends on which path was taken
-         $phi sum := snontriv $on nontrivsum strivsum $on trivsum
+         // note that sum here refers to the output argument of 
+         // the parent module which is globally visible, but can
+         // be written into exactly once!
+         $phi sum := snontriv $on nontrivsum striv $on trivsum
       $endmerge
@@
 
-    // bring sum up to the surface.
-    sum := %trivcheck:sum
 }

### 2010-07-12 Ahir-asm/antlr2/include/AaExpression.h

@@ -14,6 +14,9 @@ class AaExpression: public AaRoot
 {
   // the containing scope of this expression
   AaScope* _scope;
+  
+  // type of the expression
+  AaType* _type;
 
  public:
   virtual AaScope* Get_Scope() { return(this->_scope);}
@@ -21,6 +24,11 @@ class AaExpression: public AaRoot
   AaExpression(AaScope* scope_tpr);
   ~AaExpression();
   virtual string Kind() {return("AaExpression");}
+
+  virtual void Set_Type(AaType* t) { this->_type = t;}
+  virtual AaType* Get_Type() {return(this->_type);}
+
+  virtual void Map_Source_References() { assert(0); }
 };
 
 
@@ -71,6 +79,7 @@ class AaObjectReference: public AaExpression
     return((this->_hier_ids.size() > 0) || (this->_search_ancestor_level > 0));
   }
   virtual string Kind() {return("AaObjectReference");}
+  virtual void Map_Source_References(); // important
 };
 
 // simple reference to a constant string (must be integer or real scalar or array)
@@ -80,6 +89,7 @@ class AaConstantLiteralReference: public AaObjectReference
   AaConstantLiteralReference(AaScope* scope_tpr, string literal_string);
   ~AaConstantLiteralReference();
   virtual string Kind() {return("AaConstantLiteralReference");}
+  virtual void Map_Source_References() {} // do nothing
 };
 
 // simple reference (no array indices)
@@ -125,6 +135,7 @@ class AaTypeCastExpression: public AaExpression
   ~AaTypeCastExpression();
   void Print(ostream& ofile);
   virtual string Kind() {return("AaTypeCastExpression");}
+  virtual void Map_Source_References() {if(this->_rest) this->_rest->Map_Source_References();}
 };
 
 
@@ -144,6 +155,7 @@ class AaUnaryExpression: public AaExpression
   AaStringValue* Get_Operation() {return(this->_operation);}
   AaExpression* Get_Rest() {return(this->_rest);}
   virtual string Kind() {return("AaUnaryExpression");}
+  virtual void Map_Source_References() {if(this->_rest) this->_rest->Map_Source_References();}
 };
 
 // 
@@ -164,6 +176,11 @@ class AaBinaryExpression: public AaExpression
   AaExpression* Get_First() {return(this->_first);}
   AaExpression* Get_Second() {return(this->_second);}
   virtual string Kind() {return("AaBinaryExpression");}
[…]

### 2010-07-13 Documents/Publications/type-2/type-2.tex
type-2: Introduction and early description of Type-1 petri nets

@@
 \usepackage{times}
+\usepackage{amsfonts}
 \usepackage[margin=0.75in]{geometry}
 
-\title{On the use of petri-nets for modeling concurrent programs}
+\title{On the use of petri nets for modelling high-level programs}
 
@@
 \date{}
+
+\newtheorem{definition}{Definition}
+
+\newtheorem{srule}{Rule}
 
@@
 \begin{abstract}
-  \large
 
-  Petri-nets provide a powerful mechanism for the modeling of
-  concurrent systems.   However, because of their generality, 
-  it is often difficulty to prove properties about petri-net
-  models unless some restrictions are placed on their construction.
-  We consider the use of petri-nets to model algorithms described
-  in high level programming languages.  In particular, we introduce
-  a class of live and safe petri-nets (which we term as {\em Type-2}
-  petri-nets) for this purpose.   This class is defined by a
-  set of production rules which ensure liveness and safety of
-  the resulting petri-net.  
-  Further, we show that concurrency
-  analysis of Type-2 petri-nets can be carried out very efficiently
-  using a labelling scheme which is analogous to a symbolic simulation
-  of the petri-net.
-  Thus, it is easy to determine the sets of events (or transitions)
-  which satisfy ordering relationships (two transitions are said to
-  be ordered if it is impossible for both of them to fire simultaneously).  
-  Such information is useful in identifying possibilities for conflict 
-  free resource sharing in the implementations of the programs being
-  modeled.
+  Petri nets provide a powerful mechanism for the modelling of
+  concurrent systems. However, because of their generality, it is
+  often difficult to prove properties about petri net models unless
+  some restrictions are placed on their construction. We consider the
+  use of petri nets to model algorithms described in high level
+  programming languages. In particular, we introduce a class of live
+  and safe petri nets (which we term as {\em Type-2} petri nets) for
+  this purpose. This class is defined by a set of production rules
+  which ensure liveness and safety of the resulting petri net.
+
+  Further, we show that Type-2 petri nets support an efficient method
+  for concurrency analysis that uses a labelling scheme which is
+  analogous to a symbolic simulation of the petri net. Thus, it is
+  easy to determine the sets of events (or transitions) which satisfy
+  ordering relationships (two transitions are said to be ordered if it
+  is impossible for both of them to fire simultaneously). This is
+  useful in identifying possibilities for conflict free resource
+  sharing when implementing the programs being modelled.
+
   Despite the restrictions imposed by the production rules, the Type-2
-  class of petri-nets is powerful enough to represent programs
-  written in traditional imperative languages such as C/C++ as well in synchronous
-  languages such as Esterel. It has been used as a transition step in translating
-  imperative programs to hardware descriptions\cite{ahir-thesis}. The
-  resulting high-level synthesis flow uses the labelling scheme to
-  optimise the generated hardware by identifying opportunities for
-  sharing hardware resources in the absence of timing information.
+  class of petri nets is powerful enough to represent programs written
+  in traditional imperative languages such as C/C++ and also in
+  synchronous languages such as Esterel. It has been used as a
+  transition step in translating imperative programs to hardware
+  descriptions\cite{ahir-thesis}. The resulting high-level synthesis
+  flow uses the labelling scheme to optimise the generated hardware by
+  identifying opportunities for sharing hardware resources in the
+  absence of timing information.
 
@@
 
+\section{Introduction}
+\label{sec:introduction}
+
+Petri nets provide a powerful formalism that has been used to model
+and analyse the behaviour of a large class of systems. We consider the
+use of petri nets to model the behaviour of algorithms described in
+high-level programming languages in order to derive efficient
+implementations. We use the petri net to capture the control flow in
+the program which is coupled with a data-path (a set of computational
+resources) in order to implement the intended behaviour. The petri net
+model can be analysed to unveil properties that are important when
+deriving an implementation, such as safety and boundedness. In
+addition, the analysis can also provide information for efficient
+utilisation of resources in the data-path.
+
+However, the complexity of analysing a petri net in general quickly
+increases with its size. We adopt the strategy of reducing this
+complexity by restricting the structure of the petri net itself. We
+propose a class of petri nets (that we call ``Type-2'') that is live
+and safe, and can be easily analysed to determine concurrency
+relationships between different computations in the program.
+
+The structure of the Type-2 petri net is well suited for representing
+the control flow observed in high-level programs --- in particular,
+the interaction between concurrency and choice (branches). Threads
+that are started within a branch must synchronise before exiting the
+branch, and conversely, the targets of a branch encountered in a
+thread must be restricted to that thread. A Type-2 petri net is built
+using a set of construction rules that restrict the structure of the
+petri net to ensure this behaviour. The same restrictions enable an
+efficient labelling scheme that can be used to analyse the concurrency
+of operations in the program. Note that the behaviour allowed by a
+Type-2 petri net is a generalised version of a series-parallel
+net\cite{something}, as can be seen in Section~\cite{sec:somewhere}.
+
+\section{Definitions}
+\label{sec:defintions}
+
+A petri net is a directed bipartite graph with two kinds of nodes ---
+\textit{places} and \textit{transitions}. A \textit{marking} is a
+function that assigns a non-negative number to each place, indicating
+the number of tokens in that place.
+
+\begin{eqnarray*}
+  N & = & (P, T, E, M)\\
+  \textrm{where } P & = & \textrm{finite set of places}\\
+  T & = & \textrm{finite set of transitions}\\
+  E & \subseteq & (P \times T) \cup (T \times P)\\
+  M & : & P \rightarrow \mathbb{N}
+\end{eqnarray*}
+
+The above defintion allows at most one edge from a place (transition)
+to a transition (place). A \textit{simple place (transition)} is a
+place (transition) with one incoming edge and one outgoing place. A
+\textit{branch (fork)} is a place (transition) with one incoming edge
+and multiple outgoing edges, while a \textit{merge (join)} is a place
+(transition) with multiple incoming edges and one outgoing edge.
+
+\section{Type-1 petri nets}
+\label{sec:type-1}
+
+We first define the class of Type-1 petri nets, which are constructed
+by a set of construction rules combined with the most basic Type-1
+petri net. The class of Type-2 petri net is a subclass that is defined
+by an additional construction rule introduced in
+Section~\ref{sec:type-2}.
+
+\begin{definition}
+  A \emph{simple cycle} is the petri net $N = (P, T, E, M_0)$ where $P
+  = \{p\}$, $T =\{t\}$, $E = \{(p,t), (t,p)\}$ and $M_0(p) = 1$.
+\end{definition}
+
+Clearly, a simple cycle is live and safe; and it is defined as the
+simplest type-1 petri net. A type-1 petri net is defined as the petri
+net obtained by repeatedly applying one more construction rules to the
+simple cycle. Each rule replaces a simple place or transition in a
+type-1 petri net with a petri net subgraph called a ``standard
+region''. There are three types of standard regions, defined below.
+The places occuring in a standard region are \textbf{not marked} in
+the initial marking.
+
+\begin{description}
+\item [Series region:]
+  A \emph{p-type series region} (t-type series region) is a petri net
+  consiting of a sequence of simple transitions and places, where the
+  first and the last elements are both places (transitions) with only
+  one edge within the region. The first and the last elements
+  are of the series region are called the \emph{entry} and the
+  \emph{exit} of the region respectively.
+
+\item [Branch region:]
+  A \emph{branch region} is a conected petri net where every
+  transition is simple, while every place is either a branch or a
+  merge. No place in the branch region has an edge incident on an
+  element outside the region, except the \emph{entry} which has one
+  incoming edge from outside the region and the \emph{exit} which has
+  one outgoing edge outside the region.
+
+\item [Fork region:]
+  A \emph{fork region} is a connected petri net where every place is
+  simple, while every transition is either a fork or a join. No
+  transition in a fork region has an edge incident on an element
+  outside the region, execep the \emph{entry} and \emph{exit}
+  transitions, analogous to the entry and exit of a branch region. In
+  addition, a fork region is required to be a DAG.
+\end{description}
+
+The construction rules of a Type-1 petri net can now be stated as the
+following replacements:
+
+\begin{srule}
+  The petri net obtained by replacing a simple place (transition) in a
+  Type-1 petri net by a p-type series region (t-type series region) is
+  a also a Type-1 petri net.
+\end{srule}
+
+\begin{srule}
+  The petri net obtained by replacing a simple place in a Type-1 petri
+  net by a branch region is also a Type-1 petri net.
+\end{srule}
+
+\begin{srule}
+  The petri net obtained by replacing a simple transition in a Type-1
+  petri net by a fork region is also a Type-1 petri net.
+\end{srule}
+
+\section{Liveness and Safety in a Type-1 petri net}
+\label{sec:live-safe}
+
+We prove that a Type-1 petri net is live and safe by induction over
+the structure of the petri net. In particular, we show that the
+replacement of a simple transition or place in a Type-1 petri net by
+applying any of the construction rules does not affect the liveness
+and safety of the rest of the petri net.
+
+\subsection{Series Region}
+\label{sec:series-live-and-safe}
+
+\subsection{Branch Region}
+\label{sec:branch-live-and-safe}
+
+A petri net formed by a branch region connected to simple marked place
+is a \emph{state machine} which is strongly connected and has exactly
+one token. Hence it is live and safe\ref{murata-tutorial}.
+
+\subsection{Fork Region}
+\label{sec:fork-live-and-safe}
+
+All cycles in a petri net formed by a fork region connected to a
+simple marked place $p_0$ pass through the place $p_0$. Thus every
+cycle in the petri net has a marked place (hence it is live) and every
+place is on a cycle with exactly one marked place (hence it is safe).
+
 \end{document}

### 2010-07-19 Documents/Publications/type-2/type-2.tex
type-2: Towards a proof for Type-1 liveness and safety

@@
   descriptions\cite{ahir-thesis}. The resulting high-level synthesis
-  flow uses the labelling scheme to optimise the generated hardware by
-  identifying opportunities for sharing hardware resources in the
-  absence of timing information.
+  flow uses the labelling scheme to identify opportunities for sharing
+  hardware resources in the absence of timing information.
 
@@
 that are started within a branch must synchronise before exiting the
-branch, and conversely, the targets of a branch encountered in a
+branch, and conversely, the targets of a branch that occurs in a
 thread must be restricted to that thread. A Type-2 petri net is built
@@
   A \emph{simple cycle} is the petri net $N = (P, T, E, M_0)$ where $P
-  = \{p\}$, $T =\{t\}$, $E = \{(p,t), (t,p)\}$ and $M_0(p) = 1$.
+  = \{p\}$, $T =\{t\}$, $E = \{(p,t), (t,p)\}$ and the
+  initial marking $M_0(p) = 1$.
 \end{definition}
 
-Clearly, a simple cycle is live and safe; and it is defined as the
-simplest type-1 petri net. A type-1 petri net is defined as the petri
-net obtained by repeatedly applying one more construction rules to the
-simple cycle. Each rule replaces a simple place or transition in a
-type-1 petri net with a petri net subgraph called a ``standard
-region''. There are three types of standard regions, defined below.
-The places occuring in a standard region are \textbf{not marked} in
-the initial marking.
+Clearly, a simple cycle is live and safe. It is defined as the
+simplest type-1 petri net, and any petri net obtained by repeatedly
+applying one or more construction rules to the simple cycle is also a
+Type-1 petri net. Each rule replaces a simple place or transition in
+the original petri net with a subgraph called a ``standard region''.
+There are three types of standard regions, defined below. Places
+occuring in the standard regions are \textbf{not marked} in the
+initial marking.
 
 \begin{description}
-\item [Series region:]
-  A \emph{p-type series region} (t-type series region) is a petri net
-  consiting of a sequence of simple transitions and places, where the
-  first and the last elements are both places (transitions) with only
-  one edge within the region. The first and the last elements
-  are of the series region are called the \emph{entry} and the
-  \emph{exit} of the region respectively.
+\item [Series region:] A \emph{p-type series region} (t-type series
+  region) is a petri net consiting of a sequence of simple transitions
+  and places, where the first and the last elements are both places
+  (transitions) with only one edge within the region. The first and
+  the last elements of the series region are called the \emph{entry}
+  and the \emph{exit} of the region respectively.
 
-\item [Branch region:]
-  A \emph{branch region} is a conected petri net where every
-  transition is simple, while every place is either a branch or a
-  merge. No place in the branch region has an edge incident on an
-  element outside the region, except the \emph{entry} which has one
-  incoming edge from outside the region and the \emph{exit} which has
-  one outgoing edge outside the region.
+\item [Branch region:] A \emph{branch region} is a conected petri net
+  where every transition is simple, while every place is either a
+  branch or a merge. No place in the branch region has an edge
+  incident on an element outside the region, except the \emph{entry}
+  which has one incoming edge from outside the region and the
+  \emph{exit} which has one outgoing edge outside the region.
 
-\item [Fork region:]
-  A \emph{fork region} is a connected petri net where every place is
-  simple, while every transition is either a fork or a join. No
-  transition in a fork region has an edge incident on an element
-  outside the region, execep the \emph{entry} and \emph{exit}
+\item [Fork region:] A \emph{fork region} is a connected petri net
+  where every place is simple, while every transition is either a fork
+  or a join. No transition in a fork region has an edge incident on an
+  element outside the region, execep the \emph{entry} and \emph{exit}
   transitions, analogous to the entry and exit of a branch region. In
@@
 
-The construction rules of a Type-1 petri net can now be stated as the
-following replacements:
+The construction rules for obtaining a  Type-1 petri net can now be
+stated as follows:
+
+\setcounter{srule}{-1}
 
 \begin{srule}
-  The petri net obtained by replacing a simple place (transition) in a
-  Type-1 petri net by a p-type series region (t-type series region) is
-  a also a Type-1 petri net.
+  The simple cycle is a Type-1 petri net.
 \end{srule}
@@
 \begin{srule}
-  The petri net obtained by replacing a simple place in a Type-1 petri
-  net by a branch region is also a Type-1 petri net.
+  The petri net obtained by replacing an unmarked simple place
+  (transition) in a Type-1 petri net by a p-type series region (t-type
+  series region) is a also a Type-1 petri net.
+\end{srule}
+
+\begin{srule}
+  The petri net obtained by replacing an unmarked simple place in a
+  Type-1 petri net by a branch region is also a Type-1 petri net.
 \end{srule}
@@
 \end{srule}
+
+The above construction rules ensure that the place occuring in the
+original simple cycle is the only marked place and it is never
+replaced during the construction of the Type-1 petri net.
 
@@
 
-\subsection{Series Region}
+\subsection{Simplification}
+\label{sec:simplification-live-safe}
+
+Consider a simple place $p$ in a live and safe petri net $N =
+(P,T,E,M)$. Let $u,v$ be the predecessor and successor transitions of
+$p$ in $N$, i.e, $u,v \in T$ and $(u,p), (p,v)\in E$. Since $N$ is
+live and safe, it follows that when $u$ fires, $v$ must fire before
+$u$ may fire again (safety), and $v$ must eventually fire when $u$
+fires (liveness). Thus, any firing sequence of the petri net $N$ that
+includes the firing of $u$ (and consequently also $v$) effectively
+puts a token in $p$ and then eventually removes it. Thus, the petri
+net $N' = N - \{p,u,v\}$ can be modelled as an initially marked simple
+place $p_0$ that forms a cycle with $u$, $v$ and $p$. Similarly,
+relative to a simple transition $t$, the rest of the petri net can be
+modelled as an initially marked simple place $p_0$ that forms a simple
+cycle with $t$. Both replacements are valid if and only if the
+original petri net is live and safe; the resulting simple cycle is
+also live and safe. We use this simplification to examine the effect
+of applying a construction rule on a Type-1 petri net:
+
+\begin{enumerate}
+\item When a rule replaces a simple place $p$, the rest of the petri
+  net is reduced to a simple marked place $p_0$ connected to $p$ by
+  simple transitions.
+\item When a rule replaces a simple transition $t$, the rest of the
+  petri net is first reduced to a single marked place $p_0$.
+\end{enumerate}
+
+Thus, in each of the following subsections, we model an arbitrary
+Type-1 petri-net as a cycle consisting of a marked place $p_0$
+connected to the candidate simple place or transition through a
+suitable number of additional simple transitions. We refer to this
+cycle as the \emph{base petri net} in the rest of this section.
+
+\subsection{p-type Series Region}
 \label{sec:series-live-and-safe}
+
+When a simple place (transition) in the base petri net is replaced by
+p-type series region (t-type series region), the result is simply a
+larger cycle. The resulting petri net is live and safe, since it is a
+strongly connected \emph{state machine} with exactly one marked place.
 
@@
 
-A petri net formed by a branch region connected to simple marked place
-is a \emph{state machine} which is strongly connected and has exactly
-one token. Hence it is live and safe\ref{murata-tutorial}.
+When a place in the base petri net is replaced by a branch region, the
+resulting petri net is a strongly connected \emph{state machine} which
+is that has exactly one marked place. Hence it is live and
+safe\ref{murata-tutorial}.
 
@@
 
-All cycles in a petri net formed by a fork region connected to a
-simple marked place $p_0$ pass through the place $p_0$. Thus every
-cycle in the petri net has a marked place (hence it is live) and every
-place is on a cycle with exactly one marked place (hence it is safe).
+When a transition in the base petri net is replaced by a fork region,
+the resulting petri net is a \emph{marked graph} such that every cycle
+in the marked graph pass through the single marked place $p_0$. Thus
+every cycle in the petri net has a marked place (hence it is live) and
+every place is on a cycle with exactly one marked place (hence it is
+safe).
 

### 2010-07-20 Documents/Publications/type-2/type-2.tex
type-2: using firing sequences when proving liveness and safety

@@
   descriptions\cite{ahir-thesis}. The resulting high-level synthesis
-  flow uses the labelling scheme to identify opportunities for sharing
-  hardware resources in the absence of timing information.
+  flow uses the labelling scheme to reuse hardware resources in the
+  absence of timing information.
 
@@
 (transition) with multiple incoming edges and one outgoing edge.
+
+A firing sequence is a sequence of transitions $\sigma = [t_{i_0},
+t_{i_1}, \ldots]$ in $T$ that may fire in the petri net $N$ starting
+at some marking $M$. $L(N)$ is the set of firing sequences possible in
+the petri net $N$ starting from the initial marking $M_0$.
 
@@
 (P,T,E,M)$. Let $u,v$ be the predecessor and successor transitions of
-$p$ in $N$, i.e, $u,v \in T$ and $(u,p), (p,v)\in E$. Since $N$ is
-live and safe, it follows that when $u$ fires, $v$ must fire before
-$u$ may fire again (safety), and $v$ must eventually fire when $u$
-fires (liveness). Thus, any firing sequence of the petri net $N$ that
-includes the firing of $u$ (and consequently also $v$) effectively
-puts a token in $p$ and then eventually removes it. Thus, the petri
-net $N' = N - \{p,u,v\}$ can be modelled as an initially marked simple
-place $p_0$ that forms a cycle with $u$, $v$ and $p$. Similarly,
-relative to a simple transition $t$, the rest of the petri net can be
-modelled as an initially marked simple place $p_0$ that forms a simple
-cycle with $t$. Both replacements are valid if and only if the
-original petri net is live and safe; the resulting simple cycle is
-also live and safe. We use this simplification to examine the effect
-of applying a construction rule on a Type-1 petri net:
+$p$ in $N$, i.e, $u,v \in T$ and $(u,p), (p,v)\in E$. For any firing
+sequence $\sigma \in L(N)$, $u$ must occur before any occurence of
+$v$, and occurences of $u$ and $v$ must strictly alternate in the
+entire sequence.
+
+The petri net $N' = N - \{p,u,v\}$ can be modelled as an initially
+marked simple place $p_0$ that forms a cycle with $u$, $v$ and $p$.
+The cycle $N'$ is live and safe. Every sequence in $L(N)$ can be
+mapped to a sequence in $L(N')$ by simply deleting all transitions
+other than $u$ and $v$, if and only if the petri net $N$ is live and
+safe.
+
+Similarly, relative to a simple transition $t$, the rest of the petri
+net can be modelled as an initially marked simple place $p_0$ that
+forms a simple cycle with $t$. Both replacements are valid if and only
+if the original petri net is live and safe; the resulting simple cycle
+is also live and safe. We use this simplification to examine the
+effect of applying a construction rule on a Type-1 petri net:
 

### 2010-07-27 Documents/Publications/type-2/type-2.tex
type-2: liveness and safety for Type-1 proven in a straight-forward         observation of structure.         Also introducing Type-2 petri nets

--- 2010-07-13 Documents/Publications/type-2/type-2.tex
@@
   and safe petri nets (which we term as {\em Type-2} petri nets) for
-  this purpose. This class is defined by a set of production rules
+  this purpose. This class is defined by a set of refinement rules
   which ensure liveness and safety of the resulting petri net.
@@
   descriptions\cite{ahir-thesis}. The resulting high-level synthesis
-  flow uses the labelling scheme to optimise the generated hardware by
-  identifying opportunities for sharing hardware resources in the
+  flow uses the labelling scheme to reuse hardware resources in the
   absence of timing information.
@@
 branch, and conversely, the targets of a branch encountered in a
-thread must be restricted to that thread. A Type-2 petri net is built
-using a set of construction rules that restrict the structure of the
-petri net to ensure this behaviour. The same restrictions enable an
-efficient labelling scheme that can be used to analyse the concurrency
-of operations in the program. Note that the behaviour allowed by a
-Type-2 petri net is a generalised version of a series-parallel
-net\cite{something}, as can be seen in Section~\cite{sec:somewhere}.
+thread must be restricted to that thread. The structure of a Type-2
+petri net is restricted to guarantee this priority. The same
+restrictions enable an efficient labelling scheme that can be used to
+analyse the concurrency of operations in the program. Note that the
+behaviour allowed by a Type-2 petri net is a generalised version of a
+series-parallel net\cite{something}, as can be seen in
+Section~\cite{sec:somewhere}.
 
@@
 
-The above defintion allows at most one edge from a place (transition)
+The above definition allows at most one edge from a place (transition)
 to a transition (place). A \textit{simple place (transition)} is a
@@
 
+A firing sequence is a sequence of transitions $\sigma = [t_{i_0},
+t_{i_1}, \ldots]$ in $T$ that may fire in the petri net $N$ starting
+at some marking $M$. $L(N)$ is the set of firing sequences possible in
+the petri net $N$ starting from the initial marking $M_0$.
+
 \section{Type-1 petri nets}
@@
 
-We first define the class of Type-1 petri nets, which are constructed
-by a set of construction rules combined with the most basic Type-1
-petri net. The class of Type-2 petri net is a subclass that is defined
-by an additional construction rule introduced in
-Section~\ref{sec:type-2}.
+We first define the class of Type-1 petri nets, in terms of a set of
+refinement rules (which we call \emph{standard refinements} combined
+with a pre-defined basic Type-1 petri net. The class of Type-2 petri
+net is a subclass that is defined by an additional refinement rule
+introduced in Section~\ref{sec:type-2}.
 
@@
   A \emph{simple cycle} is the petri net $N = (P, T, E, M_0)$ where $P
-  = \{p\}$, $T =\{t\}$, $E = \{(p,t), (t,p)\}$ and $M_0(p) = 1$.
+  = \{p\}$, $T =\{t\}$, $E = \{(p,t), (t,p)\}$ and the
+  initial marking $M_0(p) = 1$.
 \end{definition}
 
-Clearly, a simple cycle is live and safe; and it is defined as the
-simplest type-1 petri net. A type-1 petri net is defined as the petri
-net obtained by repeatedly applying one more construction rules to the
-simple cycle. Each rule replaces a simple place or transition in a
-type-1 petri net with a petri net subgraph called a ``standard
-region''. There are three types of standard regions, defined below.
-The places occuring in a standard region are \textbf{not marked} in
-the initial marking.
+Clearly, a simple cycle is live and safe. It is defined as the
+simplest Type-1 petri net --- any petri net obtained by a sequence of
+standard refinements starting with the simple cycle is also a Type-1
+petri net. Each standard refinement replaces a simple place or
+transition in the original petri net with a subgraph called a
+\emph{standard region}. There are two kinds of standard regions ---
+\emph{p-type} and \emph{t-type} --- with the following common
+properties:
 
-\begin{description}
-\item [Series region:]
-  A \emph{p-type series region} (t-type series region) is a petri net
-  consiting of a sequence of simple transitions and places, where the
-  first and the last elements are both places (transitions) with only
-  one edge within the region. The first and the last elements
-  are of the series region are called the \emph{entry} and the
-  \emph{exit} of the region respectively.
+  \begin{enumerate}
+  \item Every standard region has a unique \emph{entry} and a unique
+    \emph{exit} of the same kind (place or transition) such that:
+    \begin{enumerate}
+    \item All elements in the region are reachable from the entry.
+    \item The exit is reachable from all elements in the region.
+    \end{enumerate}
+  \item No element in a standard region has an edge incident outside
+    the region, with the following exceptions:
+    \begin{enumerate}
+    \item The entry has exactly one incoming edge incident outside.
+    \item The exit has exactly one outgoing edge incident outside.
+    \end{enumerate}
+  \item Places occuring in a standard region are \textbf{not
+      marked} in the initial marking.
+\end{enumerate}
 
-\item [Branch region:]
-  A \emph{branch region} is a conected petri net where every
-  transition is simple, while every place is either a branch or a
-  merge. No place in the branch region has an edge incident on an
-  element outside the region, except the \emph{entry} which has one
-  incoming edge from outside the region and the \emph{exit} which has
-  one outgoing edge outside the region.
+\begin{definition} The two kinds of standard regions are as follows:
+  \begin{enumerate}
+  \item A \emph{p-type standard region} is a standard region where
+    every transition is simple, while the entry and exit are both
+    places.
+  \item A \emph{t-type standard region} is an \emph{acyclic} standard
+    region where every place is simple, while the entry and the exit
+    are both transitions.
+  \end{enumerate}
+\end{definition}
 
-\item [Fork region:]
-  A \emph{fork region} is a connected petri net where every place is
-  simple, while every transition is either a fork or a join. No
-  transition in a fork region has an edge incident on an element
-  outside the region, execep the \emph{entry} and \emph{exit}
-  transitions, analogous to the entry and exit of a branch region. In
-  addition, a fork region is required to be a DAG.
-\end{description}
+A p-type standard region has a structure that allows branching and
+loops but not concurrency, while a t-type standard region allows
+concurrency but not branching or loops.
 
-The construction rules of a Type-1 petri net can now be stated as the
-following replacements:
+\begin{definition}
+  A Type-1 petri net is a petri net generated by the following
+  production rules:
 
+\setcounter{srule}{-1}
 \begin{srule}
-  The petri net obtained by replacing a simple place (transition) in a
-  Type-1 petri net by a p-type series region (t-type series region) is
-  a also a Type-1 petri net.
+  The simple cycle is a Type-1 petri net.
 \end{srule}
@@
 \begin{srule}
-  The petri net obtained by replacing a simple place in a Type-1 petri
-  net by a branch region is also a Type-1 petri net.
+  The petri net obtained from a Type-1 petri net by refining an
+  unmarked simple place as a p-type region is also a Type-1 petri net.
 \end{srule}
@@
 \begin{srule}
-  The petri net obtained by replacing a simple transition in a Type-1
-  petri net by a fork region is also a Type-1 petri net.
+  The petri net obtained from a Type-1 petri net by refining a simple
+  transition as a t-type region is also a Type-1 petri net.
 \end{srule}
+
+\end{definition}
+
+These production rules imply that the place occuring in the original
+simple cycle is the only marked place in a Type-1 petri net.
 
@@
 the structure of the petri net. In particular, we show that the
-replacement of a simple transition or place in a Type-1 petri net by
-applying any of the construction rules does not affect the liveness
-and safety of the rest of the petri net.
+application of a standard refinement to a Type-1 petri net preserves
+liveness and safety in the resulting petri net.
 
-\subsection{Series Region}
-\label{sec:series-live-and-safe}
+\subsection{p-type standard region}
+\label{sec:refinement-place}
 
-\subsection{Branch Region}
-\label{sec:branch-live-and-safe}
+Consider a simple place $p \in P$ in a live and safe petri net $N =
+(P,T,E,M)$. Let $u,v$ be the predecessor and successor transitions of
+$p$ in $N$, i.e, $u,v \in T$ and $(u,p), (p,v)\in E$. Safety implies
+that when $u$ fires, it may not fire again until $v$ has fired.
+Liveness implies that $u$ can eventually fire again after $v$ has
+fired (besides the first firing reachable from the initial marking
+$M_0$).
 
-A petri net formed by a branch region connected to simple marked place
-is a \emph{state machine} which is strongly connected and has exactly
-one token. Hence it is live and safe\ref{murata-tutorial}.
+Let $N'$ be the petri net obtained by refining place $p$ as a p-type
+region $R$. Then $N'$ is safe if every place in $R$ has at most one
+token for any firing sequence starting from $u$. Also, $N'$ is live if
+a token that enters region $R$ via transition $u$ can eventually exit
+via transition $v$.
 
-\subsection{Fork Region}
-\label{sec:fork-live-and-safe}
+\paragraph{Proof:} When $u$ fires, a token is placed in the entry of
+the region $R$. Since there are no forks or joins in the region, the
+number of tokens within the region is exactly one for any sequence of
+firings within the region. Also, no token exits the region until $v$
+has fired, and hence $u$ cannot fire again before $v$ does. Hence the
+petri net $N'$ is safe. The single token circulates within the region
+and can eventually reach the exit, since the exit is reachable from
+every element in a standard region. Thus, transition $v$ can
+eventually fire after $u$ has fired and hence the petri net $N'$ is
+live.
 
-All cycles in a petri net formed by a fork region connected to a
-simple marked place $p_0$ pass through the place $p_0$. Thus every
-cycle in the petri net has a marked place (hence it is live) and every
-place is on a cycle with exactly one marked place (hence it is safe).
+\subsection{t-type standard region}
+\label{sec:refinement-transition}
+
+Consider a simple transition $t \in T$ in a live and safe petri net $N
+= (P,T,E,M)$. Let $u,v$ be the predecessor and successor places of $t$
+in $N$, i.e, $u,v \in P$ and $(u,p), (p,v)\in E$. Safety implies that
+when $t$ fires, it will not be enabled again (no token will be placed
+$u$) until the token in $v$ has been removed. Liveness implies that
+$t$ can eventually be enabled again after the token has been removed
+from $v$ (besides the first time when $t$ is enabled in a marking
+reachable from $M_0$).
+
+Let $N'$ be the petri net obtained by refining a transition $t$ as a
+t-type region $R$. Then $N'$ is safe if every place in $R$ has at most
+one token for any firing sequence starting from the entry. Also, $N'$
+is live if a token that enters region $R$ via the entry transition
+eventually exit via transition the exit transition.
+
+\paragraph{Proof:} The entry of the t-type region fires when a token
+is placed in place $u$. Since there are no branches in a t-type
+region, at least one token flows along every path starting from the
+entry. Thus every transition in the t-type region will fire at least
+once, including the exit and hence the petri net $N'$ is live. Since
+no new token will be placed in $u$ until a token is removed from $v$,
+at most one token (in fact, exactly one token) travels along every
+path in the region $R$ for any firing sequence starting with the entry
+transition. Hence the petri net $N'$ is safe.
+
+\section{Type-2 petri nets}
+\label{sec:type-2}
+
+The Type-1 petri net provides the basic framework for expressing
+concurrency and choice encountered in a high-level program. In
+addition to these features, a special pattern is required to
+represent the effect of branching on values in the program. This
+pattern is the \emph{parallel-merge} region that we describe below.
+
+\begin{definition}
+  A \emph{Type-2 petri net} is a Type-1 petri net where some of the
+  merge places are refined as parallel-merge regions.
+\end{definition}
+
+\subsection{Parallel-merge region}
+\label{sec:parallel-merge}
+
+When a branch occurs in a program, a variable may be assigned
+different values along each path in the branch. The actual value of
+the variable at the exit of the branch is decided by the path through
+the branch taken at run-time. This is represented by a multiplexer in
+the data-path with as many inputs as there are paths merging at a
+given merge. Thus, every merge in the control-path may be associated
+with one or more such multiplexers to be triggered in parallel. We
+represent the control-logic for these multiplexers as a petri net
+subgraph called the \emph{parallel-merge region} defined as follows.
+
+\begin{definition}
+  Parallel-merge region:
+  \begin{enumerate}
+  \item A {\em merge fragment} is an acyclic subgraph made of a single
+    place with two input edges and one output edge, with a simple
+    transition connected to each edge.
+  \item An {\em $m$-way merge tree} is a connected acyclic subgraph
+    made of $m-1$ merge fragments. The output edge of every merge
+    fragment except one is connected to the input edge of some other
+    merge fragment through a simple place. The subgraph has $m$ input
+    edges and a single output edge. A $2$-way merge tree is a merge
+    fragment.
+  \item A {\em $k \times m$ parallel-merge region} is a connected
+    acyclic subgraph made of the following components:
+    \begin{enumerate}
+    \item $k$ $m$-way merge trees
+    \item $m$ forks with fan-out $k$, such that each output edge of a
+      fork is connected to an input edge of a distinct merge tree
+      through a simple place
+    \item a single join with fan-in $k$, such that each input edge is
+      connected to the single output edge of a distinct merge tree
+      through a simple place
+    \end{enumerate}
+  \end{enumerate}
+\end{definition}
+
+\subsection{Refining a merge as a parallel-merge region}
+\label{sec:refine-merge}
 

### 2010-07-29 Ahir-asm/antlr2/docs/AaLRM.tex
Completed first cut of Aa2C.  Updated reference manual, added more examples

@@
 
-{\bf Aa} is intended to be a low-level 
+{\bf Aa} is a low-level 
 programming language for the description
-of algorithms in which the control flow
+of algorithms.  The control flow in
+an \Aa program 
 can be modeled by a petri net
 of a specific class (the type-2 petri-net as
-introduced in \cite{ref:SameerPhD}).  A program in
-\Aa can also be viewed as a 
+introduced in \cite{ref:SameerPhD}).  
+A program in \Aa can also be viewed as a 
 description of a system which reacts with 
@@
 through message queues.  Thus, an \Aa program
-can be directly executed on a computer, and
-can also be mapped to a logic circuit.
+can either be executed on a computer, or
+be mapped to a logic circuit.
 
@@
 \item {\bf Pipe} objects declared in the program are
-globally visible first-in-first-out queues that can 
+globally visible first-in-first-out buffers that can 
 be read from anywhere in the program. 
@@
 \begin{verbatim}
-module-definition :=  
-                 module-specifier module-name  
-                 module-input-arguments
-                 module-output-arguments
-                 module-object-declarations
-                 { <module-body> }
+$module [module-name] 
+   $in(<input-arguments>)
+   $out(<output-arguments>)
+$is
+{
+   <object-declarations>
+   <sequence-of-statements>
+}
 \end{verbatim}
@@
 
-The module body itself is a sequence of statements.
-\begin{verbatim}
-module-body := (statement)+
-\end{verbatim}
-Each statement can be viewed as a transition
+Each statement in the sequence
+of statements can be viewed as a transition
 in a petri-net, and has a 
@@
 a token is placed in each of its output places.
+
 The control flow in the sequence of statements
@@
 The region between \{ and \} is termed a {\bf scope}.  Thus, 
-the module description defines a scope.  In this case, 
+a module description defines a scope.  In this case, 
 the scope has a label which is the same as the module name.
 As we shall see later, the statements in the statement sequence
-may in turn define scopes, which may be labeled or anonymous.
-Each scope defines a name-space.  The visibility
+may in turn define scopes.  Each scope defines a name-space.  
+The visibility
 rules and access mechanisms between scopes will be described
-in Section \ref{sec:Scopes}
+in Section \ref{sec:Scopes}.
 
@@
 \begin{itemize}
-\item Atomic statements: those that have a single
+\item Simple statements: those that have a single
 entry place and a single exit place.  These are
@@
 can appear only inside block statements,  and they
-can have multiple sources and multiple sinks.
+can have multiple source and sink places.
 Examples of these statements are statements for
 combining tokens (merge and join statements) and
-statements for redirecting tokens (the toss statement).
+statements for redirecting tokens (the place statement).
 \end{itemize}
@@
 \begin{verbatim}
-a  :=  ($not b)
+a  :=  (~b)
 b  :=   (c + (d + e))
@@
 {
-   $declare $storage b $uint<32>
+   $storage b $uint<32>
    b := a
@@
   $seriesblock [s3] { ... }
-  $join s1 s2 $and $fork 
+  $join s1 s2 $fork 
       $seriesblock [s4] { ... }
@@
 merge-statement :=
-          merge-spec label-list merge-assignments
+          merge-spec label-list merge-assignments end-merge-spec
 \end{verbatim}
@@
 the next statement. 
+The merge assignments inside a merge block are all of a specific
+form, called the phi-statement.  A phi-statement has the following
+form
+\begin{verbatim}
+$phi a := b $on $entry c $on loopback 
+\end{verbatim}
+This says that the target $a$ is to take the value of $b$ if
+the merge statement was started from its entry place (ie from a
+token passed from its predecessor) and is to take the value of $c$
+if the merge statement was started from the place ``loopback".
 \end{itemize}
@@
 {
-  $merge loop $entry // $entry is the place
+  $merge loopback $entry // $entry is the place
                      // into the block
@@
                     // the token came from
+  $endmerge
   r := q + 1
   $if r < 10 $then 
-       $place [loop] // put token in place "loop"
+       $place [loopback] // put token in place "loopback"
   $endif
@@
 This is to be interpreted as follows: the merge executes
-whenever a token is present  either in the entry place
-of the branch-block or in the place labeled ``loop''.
+whenever a token is present in its entry place  or in 
+the place labeled ``loopback''.
 The merge defines variable ``q'' with a value which is
@@
 \begin{itemize}
+\item an implicit variable defined in the scope.
 \item a storage or pipe variable defined in the scope or an ancestor
@@
 
-^:a          look for variable a starting from the parent of
+#:a          look for variable a starting from the parent of
              the current scope.
@@
 
-^^:a         look for variable starting from the parent of
+##:a         look for variable starting from the parent of
              the parent of the current scope.
@@
 \begin{verbatim} 
-    scope-id : variable-reference 
+    scope-reference : variable-reference 
 \end{verbatim} 
@@
 of the scope.
+
+\section{Types}
+
+Types in \Aa can be either scalars or arrays.
+
+Scalar types can be one of 
+\begin{itemize}
+\item Unsigned integers:  An unsigned integer type
+has a width parameter and is specified as 
+\begin{verbatim}
+$uint<width>
+\end{verbatim}
+The width parameter can be any positive number.
+Values corresponding to this type are to be
+viewed as unsigned integers represented by 
+a binary sequence of the specified width.
+\item Signed integers:  A signed integer type
+has a width parameter and is specified as 
+\begin{verbatim}
+$int<width>
+\end{verbatim}
+The width parameter can be any positive number.
+Values corresponding to this type are to be
+viewed as integers maintained in the
+two-s complement form  by 
+a binary sequence of the specified width.
+\item Floats: A float is parametrized by
+two integers, the width of the exponent,
+and the width of the mantissa.  
+The specification is
+\begin{verbatim}
+$float<exponent,mantissa>
+\end{verbatim}
+where the exponent and mantissa must be positive
+integers.  The float is represented by 
+a word with $exponent+mantissa+1$ bits
+(with the additional bit needed for the sign).
+The standard IEEE 754 float and double
+precision representations correspond
+to
+\begin{verbatim}
+$float<8><23> 
+\end{verbatim}
+and
+\begin{verbatim}
+$float<11><52> 
+\end{verbatim}
+respectively.
+\item Pointers:  A pointer type has the form
+\begin{verbatim}
+$pointer<ref-width>
+\end{verbatim}
+A value of this type is considered to be
+a handle to an item whose width is specified
+by ref-width (ref-width must be positive).  
+A pointer value can only be
+referred to and {\bf can never} be used in a
+unary, binary or ternary expression.
+\end{itemize}
+
+Array types in \Aa have the form
+\begin{verbatim}
+$array[d1][d2]..[dk] $of <scalar-type-spec>
+\end{verbatim}
+The number of dimensions $k$ must be positive (there is
+no limit on this number), and the scalar type spec
+must specify one of the scalar types described above.
+
+\section{Expressions}
+
+Expressions in \Aa fall into the following classes
+\begin{itemize}
+\item {\bf Object References}: references to an object
+which have the form
+\begin{verbatim}
+<scope-specifier>:<object-specifier>
+\end{verbatim}
+The scope specifier can either specify a
+parent scope of the scope in which the expression appears,
+or a child scope of the scope in which the 
+expression appears.   The scope-specifier can be omitted
+if the reference is to be resolved in the same scope
+as the expression.
+\item {\bf Unary} Expressions:  There are two
+possible unary expressions.  The {\bf type-cast}
+expression is of the form
+\begin{verbatim}
+($cast (<type-spec>)  <expression>)
+\end{verbatim}
+For example
+\begin{verbatim}
+($cast ($uint<10>) ##:a)
+\end{verbatim}
+The type-cast expression takes the value of the
+specified expression and converts it to a value
+of the specified type.
+The {\bf bit-wise complement} expression is of the form
+\begin{verbatim}
+( ~ <expression> )
+\end{verbatim}
+The symbol for the complement is the same
+as the one used in the C programming language,
+and the evaluation of the expression proceeds in the
+usual way.
+\item {\bf Binary} Expressions: These are of the
+form
+\begin{verbatim}
+(<expression> <operation-id>  <expression>)
+\end{verbatim}
+The following binary operations are supported
+\begin{verbatim}
+// arithmetic operators
+PLUS              + 
+MINUS             - 
+MUL               * 
+DIV               / 
+SHL               <<
+SHR               >>
+
+// bit-wise logical operators
+NOT               ~     
+OR                |     
+AND               &    
+XOR               ^   
+NOR               ~|     
+NAND              ~&    
+XNOR              ~^
+
+// comparison operators
+EQUAL             ==
+NOTEQUAL          !=
+LESS              < 
+LESSEQUAL         <=
+GREATER           > 
+GREATEREQUAL      >=
+\end{verbatim}
+The evaluation of a binary expression proceeds in
+the usual way.
+\item {\bf Ternary} Expressions:  There is only one
+form for the ternary expression.
+\begin{verbatim}
+( $mux <test-expression>  <true-expression> <false-expression> )
+\end{verbatim}
+The test-expression is evaluated, and if true, the true-expression
+is evaluated, and if false, the false-expression is evaluated.
+\end{itemize}
+
+\section{Foreign Modules}
+
+A module can be marked as foreign by using the
+\$foreign keyword.
+\begin{verbatim}
+$foreign $ module [GetValue]
+	$in (ptr $pointer<32>)
+	$out (val $uint<32>)
+\end{verbatim}
+The \Aa compiler then considers that the module GetValue
+is defined ``elsewhere'' and does not try to link to it
+directly.  This linking is done outside by other tools which
+use the results of the \Aa compilation process.
 
@@
 // an array of 32-bit unsigned integers.
-$declare $storage mem $array<1024> $of $uint<32>
+$storage mem $array<1024> $of $uint<32>
 
@@
                $branchblock[sb1]{
-                    $declare $storage  I $uint<10> := 0
+                    $storage  I $uint<10> := 0
                     $merge $entry loopback
@@
                $branchblock[sb2]{
-                    $declare $storage  J $uint<10> := 0
+                    $storage  J $uint<10> := 0
                     J := (mp + 1)
@@
            snontriv := (%parsum%sb1:s  + %parsum%sb2:s)
-           // put token in place nontrivsum.  It will be picked
-           // up by succeeding merge
            $place [nontrivsum]
@@
            striv := mem[low]
-           // put token in place trivsum.  It will be picked
-           // up by succeeding merge
            $place [trivsum]
@@
          // which sum do you pick? depends on which path was taken
-         // note that sum here refers to the output argument of 
-         // the parent module which is globally visible, but can
-         // be written into exactly once!
          $phi sum := snontriv $on nontrivsum striv $on trivsum

### 2010-08-18 Ahir-asm/devel/docs/AaLRM.tex
renamed antlr2 folder to devel folder

--- 2010-07-29 Ahir-asm/antlr2/docs/AaLRM.tex
@@
 The execution of the statement is triggered when
-there is a token present in each of its source 
+there is a token present in some subset of its source 
 places and when the statement finishes execution,
@@
 means that the incoming token is placed in a labeled
-place ``fastpath''.  The token will either
-trigger a {\bf merge} statement which is required
+place ``fastpath'' (the place statement never
+puts a token into its default exit place).  
+The token placed in ``fastpath" must be used to
+trigger a unique {\bf merge} statement which is required
 to depend on this labeled place in the same branch
-block, or if no such merge statement
-exists, then ``fastpath'' is identified
-with the exit place of the  containing branch block.
+block. If no such merge exists, or if multiple such merges exist, 
+then this is an error.
 \item Merge statements: A merge statement is
@@
 which is to be interpreted as follows.  The labels in the label list
-refer to token labels defined by ``toss'' statements within the branch block.
-Whenever a token labeled by any of the items in label-list arrives
-at the input of the merge,
+refer to token labels defined by {\bf place}  statements within the branch block.
+Whenever a token is present in any of the places in
+the label-list for a merge,
 the merge statement starts and executes a series of special assignments
@@
 if the merge statement was started from the place ``loopback".
+Note that if the merge statement is to be triggered by a
+token in its default entry place, one of the labels in the
+label-list for the merge must contain the identifier \$entry.
 \end{itemize}

### 2010-08-20 Documents/Publications/dsd2010/presentation/dsd2010.tex
added presentation for DSD

\documentclass{beamer}
\usepackage{epsfig}
\title{A C-to-RTL flow as an Alternative to the Use of Embedded Processors in Digital Systems}
\author{Sameer D. Sahasrabuddhe, Sreenivas
  Subramanian, Kunal P. Ghosh, Kavi Arya \\ Madhav P. Desai\\
  Indian Institute of Technology -- Bombay, Powai, Mumbai -- 400076, INDIA}
\begin{document}

\maketitle

\frame[containsverbatim]{\frametitle{Overview}

\begin{itemize}
\item We present a correct-by-construction and scalable 
C-to-RTL design flow. 
	\begin{itemize}
	\item with a factored intermediate representation (AHIR), in which
		control-flow, data-flow and storage are represented as concurrent
		interacting subsystems.
	\end{itemize}
\item Experimental data to characterize the C-to-RTL flow
	\begin{itemize}
	\item comparison in terms of area, delay and {\bf energy}.
	\item for a set of programs, use standard commercial synthesis tools to obtain
		 an ASIC starting from the RTL generated by the C-to-RTL flow.
	\item compare with processor implementations of the same set of programs.
	\end{itemize}
\end{itemize}

}

\frame[containsverbatim]{\frametitle{A system description in AHIR}

\begin{figure}
\begin{centering}
\centerline{\psfig{figure=AhirSystem.eps,width=2.5in,height=2.0in}}
\end{centering}
\end{figure}

A system description consists of a collection of modules (a module is
similar to a function in a C program).  The environment is responsible
for memory initialization and for providing input arguments to
the system.
}

\frame[containsverbatim]{\frametitle{An AHIR module}

An AHIR module description is of the form:

\begin{displaymath}
{\bf CP} \ \times \ {\bf DP} \ \times \ {\bf S}
\end{displaymath}

where ${\bf CP}$ represents control flow, ${\bf DP}$ represents 
data flow, and {\bf S} represents storage.
}

\frame[containsverbatim]{\frametitle{The control path ${\bf CP}$}

\begin{itemize}
\item The control path is modeled by a petri-net.  Each transition 
is associated with a symbol, which can either be an input symbol
or an output symbol.
\item When a transition labeled by an input symbol is enabled, 
it fires when the input symbol is received.
\item When a transition labeled by an output symbol is enabled,
it fires eventually and emits the output symbol.
\item The petri-net has a special structure, which guarantees that
{\bf CP} is live and safe.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{The data path ${\bf DP}$}

\begin{itemize}
\item A directed graph of operators.
\item Each operator has a set of request symbols and a set of acknowledge symbols.
The operator responds to the arrival of request symbols by eventually generating
appropriate acknowledge symbols.
\item Operator library: all the standard arithmetic and logical operators, as
well as a multiplexor and load/store operators.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{The storage subsystem ${\bf S}$}
\begin{itemize}
\item The storage subsystem connects to the load/store operators in the datapath,
and is required to eventually service the load/store requests.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{Illustration of an AHIR module} 
\begin{figure}
\begin{centering}
\centerline{\psfig{figure=Illustration.eps,width=2.5in,height=2.0in}}
\end{centering}
The computation being performed is
\begin{verbatim}
               x = y + 1 || w = mem(y)
               y = x + w 
\end{verbatim}
\end{figure}
}

\frame[containsverbatim]{\frametitle{Type-2 petri-nets} 
Type-2 petri-nets are live and safe petri-nets constructed using a
set of simple rules.  The simplest type-2 petri-net has the form
shown in the figure.
\begin{figure}
\begin{centering}
\centerline{\psfig{figure=Type2Base.eps,width=2.5in,height=2.0in}}
\end{centering}
\end{figure}
}

\frame[containsverbatim]{\frametitle{Production rules} 
Replace a transition by a transition-place-transition (any region produced
from a transition is called a t-region):
\begin{figure}
\begin{centering}
\centerline{\psfig{figure=Type2T1.eps,width=2.5in,height=1.5in}}
\end{centering}
\end{figure}
}

\frame[containsverbatim]{\frametitle{Production rules} 
Replace a transition by an acyclic graph constructed using
transitions and p-regions: the acyclic graph must have a
single source and a single sink.
\begin{figure}
\begin{centering}
\centerline{\psfig{figure=Type2T2.eps,width=2.5in,height=2.0in}}
\end{centering}
\end{figure}
}

\frame[containsverbatim]{\frametitle{Production rules} 
Replace a place by a possibly cyclic graph consisting of
places and t-regions.
\begin{figure}
\begin{centering}
\centerline{\psfig{figure=Type2P1.eps,width=2.3in,height=2.0in}}
\end{centering}
\end{figure}
}

\frame[containsverbatim]{\frametitle{Production rules} 
A place with more than one incoming arc may be replaced
by a parallel merge region as shown.  The region obtained by
this replacement is not allowed to be refined any further..
\begin{figure}
\begin{centering}
\centerline{\psfig{figure=Type2P2.eps,width=2.5in,height=2.0in}}
\end{centering}
\end{figure}
}

\frame[containsverbatim]{\frametitle{Liveness and Safety}

\begin{itemize}
\item   Assume that an enable input transition eventually fires.  Then the type-2
petri-net is live and safe.
\item The type-2 petri-net is also powerful enough to describe the control
flow in imperative languages such as C/C++ as well as in synchronous
languages such as Esterel.
\end{itemize}

}

\frame[containsverbatim]{\frametitle{Optimizations}

\begin{itemize}
\item Define two transitions to be compatible if it is not possible
for both of them to fire simultaneously.
\item The compatibility of two transitions in a type-2 petri-net can 
be determined in linear time.
\item Thus, in an AHIR module, it is easy to determine which operators
can be shared without the need for arbitration. This leads to 
considerable hardware reduction without compromising performance.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{Our C-to-VHDL flow} 
\begin{itemize}
\item The starting point is a program written in C
\begin{itemize}
\item restrictions: no cycles in call graph, and no {\bf function} pointers.
\end{itemize}
\item Use the LLVM front end to convert C program into LLVM byte code (CDFG).
\item Map the CDFG to AHIR (the resulting AHIR description can be proved
to be equivalent to the CDFG).
\item Map AHIR to VHDL
\begin{itemize}
\item identify arbiter-less resource sharing opportunities to reduce hardware
cost.
\item instantiate the system: modules with their control and data-paths, the
inter-module link layer to handle calls, the memory subsystem.
\end{itemize}
\end{itemize}
}

\frame[containsverbatim]{\frametitle{Mapping AHIR to VHDL} 
\begin{itemize}
\item A synchronous, single clock, positive edge-triggered paradigm 
is used.
\item Transitions are coded as pulses which are sampled high by one
clock edge.  An output transition is implemented by an AND gate whose
inputs are the input places to the transition.
\item A place is modeled by a flip-flop which is set when any of its
input transitions fires, and is reset when any of its successor transitions
fire.
\begin{itemize}
\item A place with only a single input transition and a single output transition
is optimized away.
\end{itemize}
\item The datapath is mapped to an equivalent VHDL netlist constructed
using a library of operators.  A set of compatible operators is replaced
by a single operator with input multiplexors and output demultiplexors.
\item The memory subsystem is implemented using multiple banks, and offers
as many ports as there are shared load/store operators which use it.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{Experimental Evaluation}
\begin{itemize}
\item Select a range of programs
\begin{itemize}
\item A5 (stream cipher), AES encryption, Red-black Trees (data-structure),
      Linpack (LU factorization), Fast-Fourier Transform (FFT).
\end{itemize}
\item Using the C-to-RTL flow, map each program to RTL.  The run-time for
mapping is neglible in all cases (less than a minute).
\item Use standard synthesis tools (Synopsys Design Compiler, Cadence SOC encounter)
to implement ASIC from RTL (we use the 180nm TSMC CMOS process, with OSU standard
cell libraries).  Extract area, delay and energy numbers for the implemented ASIC.
\begin{itemize}
\item Operators are not pipelined or optimized in any fashion.
\end{itemize}
\item Run each program on a  processor based platform (we use the Intel Atom N270 as
a reference).  Extract area, delay, energy numbers using the processor data sheet.
\item Compare the ASIC numbers with the processor numbers.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{Data for 180nm implementations of C-to-RTL results }

\begin{table}[htb]
  \centering
  \caption{Area/Delay/Power/Energy data for AHIR circuits in the TSMC $0.18\mu$ technology}
  \label{table:tsmc-180nm-data}
  \renewcommand\arraystretch{1.2}
  \setlength{\tabcolsep}{1ex}
  \begin{tabular}{c|c|c|c|c|c}
    \hline
    & Area & Freq & Delay & Power & Energy \\
    & (mm$^2$) & (MHz) & (ms) & (mW) & ($\mu$J) \\
    \hline
    \hline
    A5/1 & 1.45 & 71.4 & 0.28$\mu$s & 73 & 20.44\,nJ \\
    \hline
    AES & 6.5 & 71.4 & 0.428 & 338 & 144.7 \\
    \hline
    FFT & 5.1 & 41.7 & 0.155  & 95 & 14.7  \\
    \hline
    LPK & 27 & 41.7 & 37.7 & 273 & 10300  \\
    \hline
    RBT & 18 & 41.7 & 9.9 & 153 & 1511 \\
    \hline
  \end{tabular}
\end{table}

}

\frame[containsverbatim]{\frametitle{Data for 45nm Processor }
  \begin{table}[htb]
  \centering
  \caption{Area/Delay/Power/Energy results for the processor $P$ built in a $45nm$ process}
  \label{table:atom-power-delay}
  \renewcommand\arraystretch{1.2}
  \setlength{\tabcolsep}{1ex}
  \begin{tabular}{c|c|c|c|c|c}
    \hline
    & Area & Freq & Delay & Power & Energy \\
    & (mm$^2$) & (MHz) & (ms) & (mW) & ($\mu$J) \\
    \hline
    \hline
    A5/1 & 25 & 1600 & 0.12$\mu$s & 2500 & 298.44\,nJ  \\
    \hline
    AES & 25 & 1600 & 0.036 & 2500 & 89.362 \\
    \hline
    FFT & 25 & 1600 & 0.022 & 2500 & 55.64  \\
    \hline
    LPK & 25 & 1600 & 7.90 & 2500 & 19740  \\
    \hline
    RBT & 25 & 1600 & 0.36 & 2500 & 891.89 \\
    \hline
  \end{tabular}
  \end{table}
}

\frame[containsverbatim]{\frametitle{Observations}

\begin{itemize}
\item In terms of energy, the 180nm C-to-RTL circuits are in the
same range as the 45nm processor numbers: for A5, the C-to-RTL circuit
is better by 10X, but for RBT, it is worse by 2X.
\item For a more fair comparison, we should compare 45nm circuits
with the 45nm processor.
\begin{itemize}
\item Since we did not have access to a 45nm library, we scaled the
numbers for the 180nm circuits.
\item Area scaled by 1/16.
\item Delay scaled by 1/4.
\item Power dissipation scaled by 1/2 (we assume that in the 45nm 
circuit, half the power dissipation is due to leakage), so that
energy consumption per completed task will scale by 1/8.
\end{itemize}
\item We compared the scaled numbers with the processor. 
\end{itemize}
}

\frame[containsverbatim]{\frametitle{Comparison of projected 45nm C-to-RTL circuits with 45nm processor}
\begin{table}[htb]
  \centering
  \caption{Area/Delay/Power/Energy RATIOS (processor values relative to the scaled C-to-RTL circuit values)}
  \label{table:Ratios}
  \renewcommand\arraystretch{1.2}
  \setlength{\tabcolsep}{1ex}
  \begin{tabular}{c|c|c|c|c|c}
    \hline
    & Area & Freq & Delay & Power & Energy \\
    \hline
    \hline
    A5/1 & 275.8 & 5.6 & 1.7 & 68.5 & 116.6  \\
    \hline
    AES & 61.5 & 5.6 & 0.34 & 14.8 & 4.9 \\
    \hline
    FFT & 78.4 & 9.6 & 0.57 & 52.6 & 30.3  \\
    \hline
    LPK & 14.8 & 9.6 & 0.84 & 18.3 & 15.3  \\
    \hline
    RBT & 22.2 & 9.6 & 0.15 & 32.7 & 4.7 \\
    \hline
  \end{tabular}
\end{table}
}

\frame[containsverbatim]{\frametitle{Analysis}

\begin{itemize}
\item The C-to-RTL circuits are between 4.7X and 116X more efficient than the processor.
\item The processor delays are lower than those of the C-to-RTL circuit in most
cases (except for A5) but the difference is less than one order of magnitude in
all cases.
\begin{itemize}
\item pipelining of operators and retiming of the C-to-RTL circuits should reduce
this gap.
\end{itemize}
\item To summarize, the C-to-RTL circuits have much better energy efficiency than 
the processor, but have marginally lower performance than the processor.
\end{itemize}

}

\frame[containsverbatim]{\frametitle{Conclusion}
\begin{itemize}
\item We have described a C-to-RTL flow which can be applied to complex
C programs across a wide range of applications.
\item The flow is provably correct by construction (the generated RTL 
is equivalent to the CDFG corresponding to the program from which it
was created).
\item Even in its current state, the flow produces circuits which are
considerably more energy efficient than a state-of-the-art processor.
\item Correctness by construction, energy efficiency and fast turnaround
time imply that this C-to-RTL flow can be an effective alternative to 
the use of embedded processors in implementing complex algorithms in
systems-on-chip.
\item If you wish to try out the flow, please contact madhav@ee.iitb.ac.in
or sameerds@it.iitb.ac.in.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{Thank you!}
}

\end{document}

### 2011-01-07 source/Ahir-asm/devel/docs/AaParser.tex
commmitting some recent changes.. lots of flux in libAhirV2 and Ahir-asm but will converge soon

\begin{verbatim}

Definition of parser AaParser, which is a subclass of LLkParser.

aA_Program
	:	(	( aA_Module ) 
		|	( aA_Object_Declaration ) 
		)* 
	;

aA_Module
	:	(	FOREIGN 
		|	
		) 
		MODULE aA_Label aA_In_Args aA_Out_Args 
		(	IS LBRACE ( aA_Object_Declaration )* aA_Atomic_Statement_Sequence RBRACE 
		|	
		) 
	;

aA_Object_Declaration
	:	( aA_Storage_Object_Declaration ) 
	|	( aA_Constant_Object_Declaration ) 
	|	( aA_Pipe_Object_Declaration ) 
	;

aA_Label
	:	( LBRACKET ) ( SIMPLE_IDENTIFIER ) ( RBRACKET ) 
	;

aA_In_Args
	:	IN LPAREN ( aA_Interface_Object_Declaration )* RPAREN 
	;

aA_Out_Args
	:	OUT LPAREN ( aA_Interface_Object_Declaration )* RPAREN 
	;

aA_Atomic_Statement_Sequence
	:	( aA_Atomic_Statement )+ 
	;

aA_Interface_Object_Declaration
	:	( aA_Object_Declaration_Base ) 
	;

aA_Atomic_Statement
	:	aA_Assignment_Statement 
	|	aA_Call_Statement 
	|	aA_Null_Statement 
	|	aA_Block_Statement 
	;

aA_Assignment_Statement
	:	( aA_Object_Reference ) ASSIGNEQUAL ( aA_Expression ) 
	;

aA_Call_Statement
	:	CALL SIMPLE_IDENTIFIER aA_Argv_In aA_Argv_Out 
	;

aA_Null_Statement
	:	NuLL 
	;

aA_Block_Statement
	:	aA_Series_Block_Statement 
	|	aA_Parallel_Block_Statement 
	|	aA_Fork_Block_Statement 
	|	aA_Branch_Block_Statement 
	;

aA_Object_Reference
	:	(	(	( PERCENT SIMPLE_IDENTIFIER )* 
			|	( HASH )+ 
			) 
			COLON 
		|	
		) 
		( SIMPLE_IDENTIFIER ) ( LBRACKET aA_Expression RBRACKET )* 
	;

aA_Expression
	:	(	( aA_Constant_Literal_Reference ) 
		|	( aA_Object_Reference ) 
		|	( aA_Unary_Expression ) 
		|	( aA_Binary_Expression ) 
		|	( aA_Ternary_Expression ) 
		) 
	;

aA_Argv_In
	:	LPAREN ( aA_Object_Reference )* RPAREN 
	;

aA_Argv_Out
	:	LPAREN ( SIMPLE_IDENTIFIER )* RPAREN 
	;

aA_Series_Block_Statement
	:	SERIESBLOCK 
		(	aA_Label 
		|	
		) 
		LBRACE ( aA_Object_Declaration )* aA_Atomic_Statement_Sequence RBRACE 
	;

aA_Parallel_Block_Statement
	:	PARALLELBLOCK 
		(	aA_Label 
		|	
		) 
		LBRACE ( aA_Object_Declaration )* aA_Atomic_Statement_Sequence RBRACE 
	;

aA_Fork_Block_Statement
	:	FORKBLOCK 
		(	aA_Label 
		|	
		) 
		LBRACE ( aA_Object_Declaration )* aA_Fork_Block_Statement_Sequence RBRACE 
	;

aA_Branch_Block_Statement
	:	BRANCHBLOCK 
		(	aA_Label 
		|	
		) 
		LBRACE ( aA_Object_Declaration )* aA_Branch_Block_Statement_Sequence RBRACE 
	;

aA_Fork_Block_Statement_Sequence
	:	(	( aA_Join_Fork_Statement ) 
		|	( aA_Atomic_Statement ) 
		)+

	;

aA_Join_Fork_Statement
	:	JOIN ( SIMPLE_IDENTIFIER )* 
		(	FORK aA_Atomic_Statement_Sequence 
		|	
		) 
		ENDJOIN 
	;

aA_Branch_Block_Statement_Sequence
	:	( (	( aA_Merge_Statement ) 
			|	( (	aA_Switch_Statement 
					|	aA_If_Statement 
					) ) 
			|	( aA_Atomic_Statement ) 
			|	( aA_Place_Statement ) 
			) 
			)+ 
	;

aA_Merge_Statement
	:	MERGE 
		(	( SIMPLE_IDENTIFIER ) 
		|	( ENTRY ) 
		)+ 
		(	( aA_Phi_Statement )+ 
		|	
		) 
		ENDMERGE 
	;

aA_Switch_Statement
	:	SWITCH aA_Expression ( WHEN aA_Constant_Literal_Reference THEN aA_Branch_Block_Statement_Sequence )* 
		(	DEFAULT aA_Branch_Block_Statement_Sequence 
		|	
		) 
		ENDSWITCH 
	;

aA_If_Statement
	:	IF ( aA_Expression ) THEN aA_Branch_Block_Statement_Sequence 
		(	ELSE aA_Branch_Block_Statement_Sequence 
		|	
		) 
		ENDIF 
	;

aA_Place_Statement
	:	PLACE aA_Label 
	;

aA_Block_Statement_Sequence
	:	( aA_Block_Statement )+ 
	;

aA_Phi_Statement
	:	PHI SIMPLE_IDENTIFIER ASSIGNEQUAL ( aA_Expression ON 
			(	( SIMPLE_IDENTIFIER ) 
			|	( ENTRY ) 
			) 
			)+ 
	;

aA_Constant_Literal_Reference
	:	(	PLUS 
		|	MINUS 
		|	
		) 
		(	(	( UINTEGER ) 
			|	( UFLOAT ) 
			) 
		|	( LESS 
				(	( UINTEGER )+ 
				|	( UFLOAT )+ 
				) 
				GREATER ) 
		)

	;

aA_Unary_Expression
	:	LPAREN 
		(	( ( NOT ) ( aA_Expression ) ) 
		|	( CAST LPAREN ( aA_Type_Reference ) RPAREN ( aA_Expression ) ) 
		) 
		RPAREN 
	;

aA_Binary_Expression
	:	LPAREN aA_Expression aA_Binary_Op aA_Expression RPAREN 
	;

aA_Ternary_Expression
	:	LPAREN MUX aA_Expression ( aA_Expression ) ( aA_Expression ) RPAREN 
	;

aA_Type_Reference
	:	( aA_Scalar_Type_Reference ) 
	|	( aA_Array_Type_Reference ) 
	;

aA_Binary_Op
	:	( OR ) 
	|	( AND ) 
	|	( NOR ) 
	|	( NAND ) 
	|	( XOR ) 
	|	( XNOR ) 
	|	( SHL ) 
	|	( SHR ) 
	|	( PLUS ) 
	|	( MINUS ) 
	|	( DIV ) 
	|	( MUL ) 
	|	( EQUAL ) 
	|	( NOTEQUAL ) 
	|	( LESS ) 
	|	( LESSEQUAL ) 
	|	( GREATER ) 
	|	( GREATEREQUAL ) 
	;

aA_Storage_Object_Declaration
	:	( STORAGE aA_Object_Declaration_Base ) 
	;

aA_Constant_Object_Declaration
	:	( CONSTANT aA_Object_Declaration_Base ) 
	;

aA_Pipe_Object_Declaration
	:	( PIPE aA_Object_Declaration_Base ) 
	;

aA_Object_Declaration_Base
	:	( SIMPLE_IDENTIFIER ) ( aA_Type_Reference ) 
		(	ASSIGNEQUAL aA_Constant_Literal_Reference 
		|	
		) 
	;

aA_Scalar_Type_Reference
	:	( aA_Uint_Type_Reference ) 
	|	( aA_Int_Type_Reference ) 
	|	( aA_Float_Type_Reference ) 
	|	( aA_Pointer_Type_Reference ) 
	;

aA_Uint_Type_Reference
	:	UINT LESS UINTEGER GREATER 
	;

aA_Int_Type_Reference
	:	INT LESS UINTEGER GREATER 
	;

aA_Float_Type_Reference
	:	FLOAT LESS UINTEGER COMMA UINTEGER GREATER 
	;

aA_Pointer_Type_Reference
	:	POINTER LESS UINTEGER GREATER 
	;

aA_Array_Type_Reference
	:	ARRAY ( LESS UINTEGER GREATER )+ OF ( aA_Scalar_Type_Reference ) 
	;

\end{verbatim}

### 2011-01-07 libAhirV2/include/vcSystem.hpp

@@ -1,62 +1,0 @@
-
-#ifndef _VC_System_H_
-#define _VC_System_H_
-
-#include <vcIncludes.hpp>
-#include <vcRoot.hpp>
-class vcType;
-class vcIntType;
-class vcFloatType;
-class vcArrayType;
-class vcRecordType;
-class vcPointerType;
-class vcMemorySpace;
-class vcDatapathElementLibrary;
-class vcDatapathElementTemplate;
-class vcModule;
-
-class vcSystem: public vcRoot
-{
-
-  map<string, vcType*> _type_map;
-  vector<vcType*> _type_vector;
-
-  map<string, vcMemorySpace*> _memory_space_map;
-  vector<vcModule*> _top_modules;
-  map<string, vcModule*> _modules;
-  map<string, vcDatapathElementLibrary*> _dpe_libraries;
-
-  static bool _error_flag;
- public:
-  vcSystem(string id);
-  virtual void Print(ostream& ofile);
-  void Add_Module(vcModule* module);
-  void Add_Library(vcDatapathElementLibrary* lib);
-  void Set_As_Top_Module(vcModule* module);
-  void Add_Memory_Space(vcMemorySpace* ms);
-  vcMemorySpace* Find_Memory_Space(string ms_name);
-  vcModule* Find_Module(string m_name);
-  vcDatapathElementLibrary* Find_Library(string lib_name);
-
-  void Elaborate(); // elaborate the system...
-  void Print_VHDL(ostream& ofile);
-  void Print_VHDL(ofstream& ofile);
-
-
-
-  static void Error(string err_msg); // {_error_flag = true;}
-  static void Warning(string err_msg); // {_warning_flag = true;}
-  static bool Get_Error_Flag(); // { return _error_flag; }
-
-
-  vcDatapathElementTemplate* Get_DPE_Template(string library_id, string template_id);
-  void Add_Type(string tid, vcType* t);
-  vcIntType* Make_Integer_Type(unsigned int w);
-  vcFloatType* Make_Float_Type(unsigned int c, unsigned int m);
-  vcArrayType* Make_Array_Type(vcType* etype, unsigned int dim);
-  vcRecordType* Make_Record_Type(vector<vcType*>& etypes);
-  vcPointerType* Make_Pointer_Type(string scope_id, string space_id);
-
-};
-
-#endif

### 2011-01-07 libAhirV2/src/vcLexer.cpp

@@ -1,2118 +1,0 @@
-/* $ANTLR 2.7.7 (2006-11-01): "vc.g" -> "vcLexer.cpp"$ */
-#include "vcLexer.hpp"
-#include <antlr/CharBuffer.hpp>
-#include <antlr/TokenStreamException.hpp>
-#include <antlr/TokenStreamIOException.hpp>
-#include <antlr/TokenStreamRecognitionException.hpp>
-#include <antlr/CharStreamException.hpp>
-#include <antlr/CharStreamIOException.hpp>
-#include <antlr/NoViableAltForCharException.hpp>
-
-#line 10 "vc.g"
-
-
-#line 15 "vcLexer.cpp"
-#line 1 "vc.g"
-#line 17 "vcLexer.cpp"
-vcLexer::vcLexer(ANTLR_USE_NAMESPACE(std)istream& in)
-	: ANTLR_USE_NAMESPACE(antlr)CharScanner(new ANTLR_USE_NAMESPACE(antlr)CharBuffer(in),true)
-{
-	initLiterals();
-}
-
-vcLexer::vcLexer(ANTLR_USE_NAMESPACE(antlr)InputBuffer& ib)
-	: ANTLR_USE_NAMESPACE(antlr)CharScanner(ib,true)
-{
-	initLiterals();
-}
-
-vcLexer::vcLexer(const ANTLR_USE_NAMESPACE(antlr)LexerSharedInputState& state)
-	: ANTLR_USE_NAMESPACE(antlr)CharScanner(state,true)
-{
-	initLiterals();
-}
-
-void vcLexer::initLiterals()
-{
-	literals["M"] = 61;
-	literals["C"] = 60;
-}
-
-ANTLR_USE_NAMESPACE(antlr)RefToken vcLexer::nextToken()
-{
-	ANTLR_USE_NAMESPACE(antlr)RefToken theRetToken;
-	for (;;) {
-		ANTLR_USE_NAMESPACE(antlr)RefToken theRetToken;
-		int _ttype = ANTLR_USE_NAMESPACE(antlr)Token::INVALID_TYPE;
-		resetText();
-		try {   // for lexical and char stream error handling
-			switch ( LA(1)) {
-			case 0x2c /* ',' */ :
-			{
-				mCOMMA(true);
-				theRetToken=_returnToken;
-				break;
-			}
[…]

### 2011-01-07 source/libAhirV2/include/vcType.hpp

@@ -1,0 +1,154 @@
+
+#ifndef _VC_TYPE_H
+#define _VC_TYPE_H
+
+#include <vcIncludes.hpp>
+#include <vcRoot.hpp>
+
+class vcType: public vcRoot
+{
+  unsigned int _index;
+
+public:
+
+  vcType();
+
+  virtual string Kind()
+  {
+    return("vcType");
+  }
+
+  void Set_Index(unsigned int w) {this->_index = w;}
+  unsigned int Get_Index() {return(this->_index);}
+
+  virtual void Print(ostream& ofile) {assert(0);}
+  virtual int Size() { assert(0);}
+};
+
+class vcScalarTypeTemplate: public vcType
+{
+  bool _int_flag;
+  bool _float_flag;
+  string _width;
+  string _characteristic;
+  string _mantissa;
+public:
+  vcScalarTypeTemplate(string width); // int type
+  vcScalarTypeTemplate(string characteristic, string mantissa); // float type
+
+  virtual void Print(ostream& ofile);
+  virtual int Size() {return(-1);}
+};
+
+class vcScalarType: public vcType
+{
+
+public:
+  vcScalarType();
+  virtual string Kind() {return("vcScalarType");}
+  virtual int Size() { assert(0);}
+};
+
+
+class vcIntType: public vcScalarType
+{
+
+protected:
+  // width > 0
+  unsigned int _width;
+
+public:
+  virtual unsigned int Get_Width() {return(this->_width);}
+  vcIntType(unsigned int width);
+
+  void Print(ostream& ofile);
+  virtual string Kind() {return("vcIntType");}
+  virtual int Size() { return(this->_width);}
+};
+
+class vcMemorySpace;
+class vcPointerType: public vcIntType
+{
+  vcMemorySpace* _memory_space;
+
+public :
+  vcPointerType(vcMemorySpace* mem_space);
+  vcMemorySpace* Get_Memory_Space();
+
+  virtual void Print(ostream& ofile);
+  virtual string Kind() {return("vcPointerType");}
+  virtual int Size();
+};
+
+
+class vcFloatType : public vcScalarType
+{
+protected:
+  // both > 0
+  vcIntType* _characteristic_type;
+  vcIntType* _mantissa_type;
+
+public:
+  unsigned int Get_Characteristic_Width() {return(this->_characteristic_type->Get_Width());}
+  unsigned int Get_Mantissa_Width() {return(this->_mantissa_type->Get_Width());}
+
[…]

### 2011-01-07 source/libAhirV2/src/vcLexer.hpp

@@ -1,0 +1,142 @@
+#ifndef INC_vcLexer_hpp_
+#define INC_vcLexer_hpp_
+
+#include <antlr/config.hpp>
+/* $ANTLR 2.7.7 (2006-11-01): "vc.g" -> "vcLexer.hpp"$ */
+#include <antlr/CommonToken.hpp>
+#include <antlr/InputBuffer.hpp>
+#include <antlr/BitSet.hpp>
+#include "vcParserTokenTypes.hpp"
+#include <antlr/CharScanner.hpp>
+#line 13 "vc.g"
+
+#include <vcHeader.hpp>
+#include <antlr/RecognitionException.hpp>
+	ANTLR_USING_NAMESPACE(antlr)
+
+#line 18 "vcLexer.hpp"
+class CUSTOM_API vcLexer : public ANTLR_USE_NAMESPACE(antlr)CharScanner, public vcParserTokenTypes
+{
+#line 1 "vc.g"
+#line 22 "vcLexer.hpp"
+private:
+	void initLiterals();
+public:
+	bool getCaseSensitiveLiterals() const
+	{
+		return true;
+	}
+public:
+	vcLexer(ANTLR_USE_NAMESPACE(std)istream& in);
+	vcLexer(ANTLR_USE_NAMESPACE(antlr)InputBuffer& ib);
+	vcLexer(const ANTLR_USE_NAMESPACE(antlr)LexerSharedInputState& state);
+	ANTLR_USE_NAMESPACE(antlr)RefToken nextToken();
+	public: void mATTRIBUTE(bool _createToken);
+	public: void mDPE(bool _createToken);
+	public: void mLIBRARY(bool _createToken);
+	public: void mMEMORYSPACE(bool _createToken);
+	public: void mOBJECT(bool _createToken);
+	public: void mCAPACITY(bool _createToken);
+	public: void mDATAWIDTH(bool _createToken);
+	public: void mADDRWIDTH(bool _createToken);
+	public: void mMODULE(bool _createToken);
+	public: void mSERIESBLOCK(bool _createToken);
+	public: void mPARALLELBLOCK(bool _createToken);
+	public: void mFORKBLOCK(bool _createToken);
+	public: void mBRANCHBLOCK(bool _createToken);
+	public: void mOF(bool _createToken);
+	public: void mFORK(bool _createToken);
+	public: void mJOIN(bool _createToken);
+	public: void mBRANCH(bool _createToken);
[…]

### 2011-01-07 source/libAhirV2/src/vcOperator.cpp

@@ -1,0 +1,173 @@
+#include <vcIncludes.hpp>
+#include <vcRoot.hpp>
+#include <vcType.hpp>
+#include <vcValue.hpp>
+#include <vcControlPath.hpp>
+#include <vcDataPath.hpp>
+#include <vcOperator.hpp>
+
+
+vcLoadStore::vcLoadStore(string id, vcMemorySpace* ms, vcWire* addr, vcWire* data):vcSplitOperator(id)
+{
+  _memory_space = ms;
+  _address = addr;
+  _data = data;
+}
+
+vcLoad::vcLoad(string id, vcMemorySpace* ms, vcWire* addr, vcWire* data):vcLoadStore(id,ms,addr,data) {}
+void vcLoad::Print(ostream& ofile)
+{
+
+}
+vcStore::vcStore(string id, vcMemorySpace* ms, vcWire* addr, vcWire* data):vcLoadStore(id,ms,addr,data) {}
+void vcStore::Print(ostream& ofile)
+{
+  //todo
+}
+
+vcPhi::vcPhi(string id, vector<vcWire*>& inwires, vcWire* outwire):vcRoot(id)
+{
+  assert(inwires.size() > 0 && outwire != NULL);
+
+  // all wires must have the same type
+  vcType* t = outwire->Get_Type();
+  for(int idx = 0; idx < inwires.size(); idx++)
+    assert(t == inwires[idx]->Get_Type());
+
+  this->_inwires = inwires;
+  this->_outwire = outwire;
+  this->_ack = NULL;
+}
+
+void vcPhi::Set_Inreqs(vector<vcTransition*>& reqs)
+{
+  assert(reqs.size() == this->_inwires.size());
+  this->_inreqs = reqs;
+}
+
+void vcPhi::Set_Ack(vcTransition* t)
+{
+  assert(this->_ack == NULL);
+  assert(t->Get_Transition_Type() == _IN_TRANSITION);
+  this->_ack = t;
+}
+
+void vcPhi::Print(ostream& ofile)
+{
+  ofile << vcLexerKeywords[__PHI] << " " << this->Get_Label() << " ";
+  for(int idx = 0; idx < this->_inwires.size(); idx++)
+    {
+      ofile << this->_inwires[idx]->Get_Id() << " ";
+    }
[…]

### 2011-01-07 source/libAhirV2/src/vcParse.cpp

@@ -1,0 +1,42 @@
+#include <signal.h>
+#include <vcSystem.hpp>
+#include <vcParser.hpp>
+#include <vcLexer.hpp>
+
+using namespace std;
+using namespace antlr;
+
+void vcParse(string filename, vcSystem* sys)
+{
+  ifstream infile;
+  
+  infile.open(filename.c_str());
+  if(!infile.is_open())
+    {
+      cerr << "Error: Could not read file " << filename << endl;
+      exit(1);
+    }
+  
+  vcLexer* lexer = new vcLexer(infile);
+  vcParser* parser = new vcParser(*lexer);
+  
+  lexer->setFilename(filename);
+  parser->setFilename(filename);
+  
+  try
+    {
+      parser->vc_System(sys);
+    }
+  catch(ANTLR_USE_NAMESPACE(antlr)RecognitionException& re)
+    {
+      cerr << "Error: Parsing Exception: " << re.toString() << endl;
+      vcSystem::Error("");
+    }
+  
+  infile.close();
+  delete parser;
+  delete lexer;
+}
+
+
+

### 2011-01-07 source/libAhirV2/src/vcParserTokenTypes.hpp

@@ -1,0 +1,125 @@
+#ifndef INC_vcParserTokenTypes_hpp_
+#define INC_vcParserTokenTypes_hpp_
+
+/* $ANTLR 2.7.7 (2006-11-01): "vc.g" -> "vcParserTokenTypes.hpp"$ */
+
+#ifndef CUSTOM_API
+# define CUSTOM_API
+#endif
+
+#ifdef __cplusplus
+struct CUSTOM_API vcParserTokenTypes {
+#endif
+	enum {
+		EOF_ = 1,
+		LIBRARY = 4,
+		LBRACE = 5,
+		RBRACE = 6,
+		DPE = 7,
+		PARAMETER = 8,
+		MIN = 9,
+		UINTEGER = 10,
+		MAX = 11,
+		IN = 12,
+		OUT = 13,
+		COLON = 14,
+		FLOAT = 15,
+		LESS = 16,
+		SIMPLE_IDENTIFIER = 17,
+		COMMA = 18,
+		GREATER = 19,
+		INT = 20,
+		REQS = 21,
+		ACKS = 22,
+		MEMORYSPACE = 23,
+		CAPACITY = 24,
+		DATAWIDTH = 25,
+		ADDRWIDTH = 26,
+		OBJECT = 27,
+		ASSIGNEQUAL = 28,
+		MODULE = 29,
+		LINK = 30,
+		EQUIVALENT = 31,
+		SLASH = 32,
+		PHI = 33,
+		CONTROLPATH = 34,
+		PLACE = 35,
+		TRANSITION = 36,
+		HIDDEN = 37,
+		SERIESBLOCK = 38,
+		PARALLELBLOCK = 39,
+		BRANCHBLOCK = 40,
+		AT = 41,
+		MERGE = 42,
+		ENTRY = 43,
+		FROM = 44,
+		BRANCH = 45,
+		EXIT = 46,
+		FORKBLOCK = 47,
+		JOIN = 48,
+		FORK = 49,
+		DATAPATH = 50,
+		PLUS_OP = 51,
+		MINUS_OP = 52,
+		MUL_OP = 53,
+		DIV_OP = 54,
+		SHL_OP = 55,
+		SHR_OP = 56,
+		GT_OP = 57,
+		GE_OP = 58,
+		EQ_OP = 59,
+		LT_OP = 60,
+		LE_OP = 61,
+		NEQ_OP = 62,
+		BITSEL_OP = 63,
[…]

### 2011-02-26 source/Ahir-asm/devel/docs/AaLRM.tex
finished the first round of testing.   all basic structures seem to work fine.    - parallel,series,fork, branch blocks    - storage object access and memory space management.    - pipe access.    - function calls. I will move on to bc->Aa stuff now.

--- 2010-08-18 Ahir-asm/devel/docs/AaLRM.tex
@@
 \end{verbatim}
-Declarations at the program level 
+Program object declarations
 can belong to one of three classes:
@@
 \item {\bf Storage} objects declared in the program
-are globally visible variables that can be read from
-anywhere in the program and can be written from anywhere
-in the program.
-\item {\bf Pipe} objects declared in the program are
-globally visible first-in-first-out buffers that can 
-be read from anywhere in the program. 
-written into by at most one module in the program.
+are variables that can be written into from multiple
+points in the program.  When declared at the program
+scope, storage objects are visible throughout
+the program.
+\item {\bf Pipe} objects declared at the program scope
+are globally visible first-in-first-out buffers that can 
+be read from anywhere in the program and can be
+written into from anywhere in th program.  A pipe
+is a single-place buffer.  A write to a pipe succeeds
+only if the single place buffer is empty.  A read
+from the pipe succeeds only if the single place
+buffer is not empty (the read empties the buffer).
+Thus, pipes can be used for synchronization in 
+an \Aa program.
 \begin{itemize}
-\item A pipe can be written into by at most one module
-in the program, and can be read from by at most one module
-in the program.
 \item A pipe that is only written into in the program, but
 is not read from, is assumed to have a destination outside
-the program.
+the program.  Such a pipe is an output port of the
+program.
 \item A pipe that is only read from in a program but not
 written into, is assumed to have a source outside the program.
+Such a pipe is an input port of the program.
 \end{itemize}
-\item {\bf Constant} objects are globally visible and have
+\item {\bf Constant} objects declared at
+the program scope are globally visible and have
 initial values which can never be altered in the program.
@@
 Each statement in the sequence
-of statements can be viewed as a transition
+of statements can be abstractly 
+viewed as a transition
 in a petri-net, and has a 
@@
 \begin{itemize}
-\item Simple statements: those that have a single
+\item Complete or atomic statements: these that have a single
 entry place and a single exit place.  These are
@@
 {\em simple} statements and {\em block} statements.
-\item Sub-atomic statements:  these statements
+\item Incomplete or sub-atomic statements:  these statements
 can appear only inside block statements,  and they
@@
 
-\subsection{Simple Statements}
+\subsection{Complete Statements}
 
-Simple statements can be one of the following:
+Complete statements can be one of the following:
 \begin{itemize}
@@
 \end{verbatim}
+
 More details of the syntax are provided in 
@@
 
-\subsection{Block Statements}
+\subsection{Complete Block Statements}
 
@@
 }
-\end{verbatim}
+\end{verbatim} %$
 The two statements will start in parallel, and the block 
@@
 $place [fastpath]
-\end{verbatim}
+\end{verbatim}%$
 means that the incoming token is placed in a labeled
@@
 token passed from its predecessor) and is to take the value of $c$
-if the merge statement was started from the place ``loopback".
+if the merge statement was started from the place ``loopback''.
 Note that if the merge statement is to be triggered by a
@@
 }
-\end{verbatim}
+\end{verbatim}%$
 This is to be interpreted as follows: the merge executes
@@
 $float<exponent,mantissa>
-\end{verbatim}
+\end{verbatim} %$
 where the exponent and mantissa must be positive
@@
 \begin{verbatim}
-$float<8><23> 
+$float<8,23> 
 \end{verbatim}
@@
 \begin{verbatim}
-$float<11><52> 
+$float<11,52> 
 \end{verbatim}
@@
 \begin{verbatim}
-$array[d1][d2]..[dk] $of <scalar-type-spec>
-\end{verbatim}
-The number of dimensions $k$ must be positive (there is
-no limit on this number), and the scalar type spec
+$array[d1] $of <scalar-type-spec>
+\end{verbatim} %$
+Only single dimensional arrays are allowed
+in an \Aa program, and the scalar type spec
 must specify one of the scalar types described above.
@@
 ($cast (<type-spec>)  <expression>)
-\end{verbatim}
+\end{verbatim} %$
 For example
@@
 The evaluation of a binary expression proceeds in
-the usual way.
+the usual way.  Note that when specifying an 
+expression, you {\bf must} use parentheses around
+each expression.
 \item {\bf Ternary} Expressions:  There is only one
@@
 ( $mux <test-expression>  <true-expression> <false-expression> )
-\end{verbatim}
+\end{verbatim}%$
 The test-expression is evaluated, and if true, the true-expression
 is evaluated, and if false, the false-expression is evaluated.
+Note the parentheses delimiting the expression.
 \end{itemize}
@@
 A module can be marked as foreign by using the
+
 \$foreign keyword.
@@
 \end{verbatim}
+
 The \Aa compiler then considers that the module GetValue
@@
 $module [sel_mod]
-   $in (a $uint<32> b $uint<10>)
-   $out (c $uint<32>)
+   $in (a:$uint<32> b:$uint<10>)
+   $out (c:$uint<32>)
 $is
 {
-   c := (mem[b] + a)
-   mem[b] := c
+   t := (mem[b] + a)
+   mem[b] := t
+   c := t
 }
@@
 // an array of 32-bit unsigned integers.
-$storage mem $array<1024> $of $uint<32>
+$storage mem:$array[1024] $of $uint<32>
 
@@
 $module [sum_mod]
-    $in (low $uint<10> high $uint<10>)
-    $out (sum $uint<32>)
+    $in (low:$uint<10> high:$uint<10>)
+    $out (sum:$uint<32>)
 $is
 {
-
    d := (high-low)  
@@
                $branchblock[sb1]{
-                    $storage  I $uint<10> := 0
+                    $storage  I:$uint<10>
                     $merge $entry loopback
@@
                $branchblock[sb2]{
-                    $storage  J $uint<10> := 0
+                    $storage  J:$uint<10>
                     J := (mp + 1)

### 2011-02-26 source/libAhirV2/src/vcControlPath.cpp

@@ -359,11 +359,11 @@ void vcTransition::Print_VHDL(ostream& ofile)
 
 string vcTransition::Get_DP_To_CP_Symbol()
 {
-  return(To_VHDL(this->Get_Hierarchical_Id()) + "_dp_to_cp");
+  return(To_VHDL(this->Get_Hierarchical_Id() + "_dp_to_cp"));
 }
 string vcTransition::Get_CP_To_DP_Symbol()
 {
-  return(To_VHDL(this->Get_Hierarchical_Id()) + "_cp_to_dp");
+  return(To_VHDL(this->Get_Hierarchical_Id() + "_cp_to_dp"));
 }
 vcPlace::vcPlace(vcCPElement* parent, string id, unsigned int init_marking):vcCPElement(parent, id)
 {
@@ -399,10 +399,13 @@ vcCPBlock::vcCPBlock(vcCPBlock* parent, string id): vcCPElement((vcCPElement*)pa
 
 void vcCPBlock::Add_CPElement(vcCPElement* cpe)
 {
-  assert(this->_element_map.find(cpe->Get_Id()) == this->_element_map.end());
-
-  this->_element_map[cpe->Get_Id()] = cpe;
-  this->_elements.push_back(cpe);
+  if(cpe != NULL)
+    {
+      assert(this->_element_map.find(cpe->Get_Id()) == this->_element_map.end());
+      
+      this->_element_map[cpe->Get_Id()] = cpe;
+      this->_elements.push_back(cpe);
+    }
 }
 
 vcCPElement* vcCPBlock::Find_CPElement(string cname)

### 2011-04-11 Documents/AhirV2/AhirV2.tex
Circular references!

\documentclass{article}

\title{AhirV2: from algorithms to hardware}
\author{Madhav Desai \\ Department of Electrical Engineering \\ Indian Institute of Technology \\
	Mumbai 400076 India}

\newcommand{\Aa}{{\bf Aa}~}
\newcommand{\vC}{{\bf vC}~}

\begin{document}
\maketitle

\section{Introduction}

We describe the essential features
of the second version of the AhirV2 toolset
developed at IIT-Bombay.

\section{What is AhirV2?}

Perhaps the simplest way to understand AhirV2 is
through an example.

Suppose we start with a simple program.
\begin{verbatim}
int add(int a, int b)
{
        int c = (a+b);
        return(c);
}
\end{verbatim}
This program describes an algorithm.  We normally
compile this program and convert it to a machine-level
program which is executed on a processor.

What if we wished to convert this program to
a circuit?  
Such a circuit would have inputs 
$a,b$ and an output $c$ (together with some
interface hand-shake signals).  When activated,
the circuit should read its inputs and compute 
a response which is the same as would be expected
if the program were executed on a computer.

AhirV2 consists of a set of tools which takes
a C/C++ program and converts this collection to
a hardware circuit (described in VHDL).   This
conversion is done in two steps:
\begin{itemize}
\item The high-level program is compiled
to an interemediate assembly form.  AhirV2
introduces an intermediate assembly language
\Aa  which can serve as a target for sequential
programming languages (such as C/C++) as
well as for parallel programming languages.
\item From the \Aa description, a virtual
circuit (described in a virtual circuit
description language \vC) is generated.
The chief optimization carried out at this
step is that decomposing the system memory
into disjoint segments based on usage
analysis (this considerably improves the
available memory bandwidth and reduces
system cost).
\item From the \vC description, a
VHDL description of the system is generated.
The chief optimization carried out at this
stage is resource sharing.  The \vC description
is analyzed to identify operations which
cannot be concurrently active and this information
is used to reduce the hardware required.
\item 
The VHDL description produced from 
\vC is in terms of a library
of VHDL design units which has been developed
as part of the AhirV2  effort.  This library
consists of control-flow elements, data-path
elements and memory elements.
\item 
For the C/C++ to \Aa translation, the
AhirV2 toolset uses the front-end C/C++ compiler
infrastructure developed by the LLVM project (www.llvm.org), which
compiles the C/C++ program to LLVM byte-code
(which is a processor independent representation
of the compiled program).  The AhirV2 toolset
has a utility which translates the LLVM byte-code
to an \Aa description.
\end{itemize}
In Section \ref{sec:Example}, 
we illustrate the use of these tools on the
simple example shown above.

The earlier version of the AHIR flow has been
used to convert non-trivial programs to hardware.
The resulting circuits are upto two orders of
magnitude more energy-efficient than
a processor \cite{ref:dsd2010}.  The AhirV2
toolset incorporates several optimizations which
make the resulting circuits more 
competitive.

\section{The tools}

We assume that you have access to either {\bf llvm-gcc}
of {\bf clang} as the front-end compiler which generates
LLVM byte-code from C/C++.  The current AhirV2 toolset
is consistent with llvm 2.8 and clang 2.8.

The other tools in the chain are described below.

\subsection{{\bf llvm2aa}}

This tool takes LLVM byte code and converts it into an
\Aa file.
\begin{verbatim}
llvm2aa bytecode.o > bytecode.aa
\end{verbatim}
The generated \Aa code is sent to {\bf stdout} and all informational
messages are sent to {\bf stderr}.  On success, the tool returns 0.

\subsection{{\bf Aa2VC}}

This tool takes a list of \Aa programs and converts them
to a \vC description. 
\begin{verbatim}
Aa2VC [-O] file1.aa file2.aa ...  > result.vc
\end{verbatim}
The generated \vC code is sent to {\bf stdout} and all informational
messages are sent to {\bf stderr}.  On success the tool returns 0.

The tool performs memory space decomposition and if the {\bf -O} option
is specified, it also attempts to parallelize sequential code using 
dependency analysis.

\subsection{\bf vc2vhdl}

Takes a collection of \vC descriptions and converts them to
VHDL.
\begin{verbatim}
vc2vhdl -t foo [-t bar -t bar2 ...]  -f file1.vc -f file2.vc ...  > system.vhdl
\end{verbatim}
Using the {\bf -t} option, one can specify the modules which are to be 
accessible from the ports of the generated VHDL system.
Using the {\bf -f} option, one specifies the \vC files to be analyzed.

The tool performs concurrency analysis to determine operations which
can be mapped to the same physical operator without the need for
arbitration.  It also instantiates separate memory subsystems for
the disjoint memory spaces (in practice many of the memory spaces
are small and are converted to register banks).

\subsection{Miscellaneous: {\bf vcFormat} and {\bf vhdlFormat}}

The outputs produces by {\bf Aa2VC} and {\bf vc2vhdl} are
not well formatted.  One can format \Aa and \vC files
using  {\bf vcFormat} as follows
\begin{verbatim}
vcFormat < unformatted-vc/aa-file  > formatted-vc/aa-file
\end{verbatim}
and similarly use {\bf vhdlFormat} to format generated
VHDL file.

\section{An example} \label{sec:Example}

Let us revisit the simple example considered in
the first section:
\begin{verbatim}
int add(int a, int b)
{
        int c = (a+b);
        return(c);
}
\end{verbatim}
We wish to generate a circuit which {\em implements}
the specification implied by this program.

We convert the program to LLVM byte code using
the {\bf clang} compiler (www.llvm.org)
\begin{verbatim}
      clang -std=gnu89 -emit-llvm -c add.c
\end{verbatim}
This produces a binary file {\bf add.o} which is
the LLVM byte-code.  To make the byte-code human
readable, we dis-assemble it using an LLVM utility
\begin{verbatim}
     llvm-dis add.o
\end{verbatim}
This is what the LLVM assembly code looks like
\begin{verbatim}
; ModuleID = 'add.o'
target datalayout = "e-p ...... "
target triple = "i386-pc-linux-gnu"

define i32 @add(i32 %a, i32 %b) nounwind {
  %1 = alloca i32, align 4
  %2 = alloca i32, align 4
  %c = alloca i32, align 4
  store i32 %a, i32* %1, align 4
  store i32 %b, i32* %2, align 4
  %3 = load i32* %1, align 4
  %4 = load i32* %2, align 4
  %5 = add nsw i32 %3, %4
  store i32 %5, i32* %c, align 4
  %6 = load i32* %c, align 4
  ret i32 %6
}
\end{verbatim}
To get to this point, we could have used several
optimizations which are available in the LLVM frame-work.
But we work with the unoptimized version to illustrate
the storage decomposition which is carried out by
the AhirV2 tools.

The LLVM byte-code is our starting point.  We first convert it
to \Aa.
\begin{verbatim}
llvm2aa add.o | vcFormat > add.o.aa
\end{verbatim}
This produces an \Aa program
\begin{verbatim}
// Aa code produced by llvm2aa (version 1.0)
$module [add]
// arguments
$in (a : $uint<32> b : $uint<32> )
$out (ret_val__ : $uint<32>)
$is
{
  $storage stored_ret_val__ : $uint<32>
  $branchblock [add]
  {
    //begin: basic-block bb_0
    $storage iNsTr_0 : $uint<32>
    $storage iNsTr_1 : $uint<32>
    $storage c : $uint<32>
    iNsTr_0 := a
    iNsTr_1 := b
    // load
    iNsTr_4 := iNsTr_0
    // load
    iNsTr_5 := iNsTr_1
    iNsTr_6 := (iNsTr_4 + iNsTr_5)
    c := iNsTr_6
    // load
    iNsTr_8 := c
    stored_ret_val__ := iNsTr_8
    $place [return__]
    $merge return__ $endmerge
    ret_val__ := stored_ret_val__
  }
}
\end{verbatim}

Now, this \Aa code is converted to a virtual circuit \vC representation.
\begin{verbatim}
     Aa2VC -O add.o.aa | vcFormat > add.o.aa.vc
\end{verbatim}
The virtual circuit representation is a bit too verbose to reproduce entirely
here, but we show some critical fragments
\begin{verbatim}
$module [add] 
{
  $in a:$int<32> b:$int<32>
  $out ret_val__:$int<32>
  $memoryspace [memory_space_0] 
  {
    $capacity 1
    $datawidth 32
    $addrwidth 1
    // ret-val is kept here
    $object [xxaddxxstored_ret_val__] : $int<32>
  }
  $memoryspace [memory_space_1] 
  {
    $capacity 1
    $datawidth 32
    $addrwidth 1
    // a is kept here.
    $object [xxaddxxaddxxiNsTr_0] : $int<32>
  }
  $memoryspace [memory_space_2] 
  {
    $capacity 1
    $datawidth 32
    $addrwidth 1
    // b is kept her
    $object [xxaddxxaddxxiNsTr_1] : $int<32>
  }
  $memoryspace [memory_space_3] 
  {
    $capacity 1
    $datawidth 32
    $addrwidth 1
    // c is kept here.
    $object [xxaddxxaddxxc] : $int<32>
  }
  $CP 
  {
     // a control-flow petri-net..  verbose..
  }
  // end control-path
  $DP 
  {
     // wires and operators.
  }

   // links between CP and DP
}
\end{verbatim}
The important points to note are that the stored objects a,b,c and
ret\_val\_\_ are mapped to different memory spaces.  Thus, the
chief difference between a \vC description and a processor is 
that the \vC program partitions storage into small units which
are accessed only by operators that need them.

Finally, we take the \vC description and convert it to
VHDL
\begin{verbatim}
vc2vhdl -t add -f add.o.aa.vc | vhdlFormat > system.vhdl
\end{verbatim}
This produces a VHDL implementation of the system with
{\bf add} marked as a top-level module.  The VHDL that
is produced is too voluminous to reproduce here, but
the top-level system entity is
\begin{verbatim}
entity test_system is  -- system
  port (--
    add_a : in  std_logic_vector(31 downto 0);
    add_b : in  std_logic_vector(31 downto 0);
    add_ret_val_x_x : out  std_logic_vector(31 downto 0);
    add_tag_in: in std_logic_vector(0 downto 0);
    add_tag_out: out std_logic_vector(0 downto 0);
    add_start : in std_logic;
    add_fin   : out std_logic;
    clk : in std_logic;
    reset : in std_logic); --
  --
end entity;
\end{verbatim}
There are ports corresponding to the arguments of the top-level module,
and the add\_start/add\_fin is a handshake pair.  One sets of the inputs,
starts the system and waits until the fin is asserted.  After the fin
is asserted, one has the return value of at the appropriate port.

\section{Handling complex programs}

This example is trivial.  Real, non-trivial programs can be
mapped in this manner.  Currently, there are only
two restrictions
\begin{itemize}
\item No recursion, no cycles in the call-graph of the original
program.
\item No function pointers.
\end{itemize}

The AhirV2 flow produces a modular system, and by default,
produces one VHDL entity for every function in the original
program.  Further, the concept of message pipes is native
to the \Aa and \vC descriptions.  Thus, it is easy to
map parallel programs or programs consisting of multiple
concurrent processes to hardware.

\section{To be continued}

This document is a work-in-progress.  More details and
related documentation will be added shortly.

The examples folder contains some explanations which might
make things easier to understand.

\begin{thebibliography}{99}
\bibitem{ref:dsd2010}
Sameer D. Sahasrabuddhe, Sreenivas Subramanian, Kunal P. Ghosh, Kavi Arya, Madhav P. Desai, 
"A C-to-RTL Flow as an Energy Efficient Alternative to Embedded Processors in Digital Systems," 
DSD, pp.147-154, 2010 13th Euromicro Conference on Digital 
System Design: Architectures, Methods and Tools, 2010
\end{thebibliography}
\end{document}

### 2011-04-11 source/Ahir-asm/devel/docs/AaLRM.tex
Circular references!

@@
 
-{\bf Aa} is a low-level 
+{\bf Aa} is a low-level {\em algorithm-assembly}
 programming language for the description
-of algorithms.  The control flow in
-an \Aa program 
+of algorithms.   An \Aa program 
 can be modeled by a petri net
@@
 Objects (storage/pipe/constant) declared in the module are visible only in the
-module body.
+module body.   For example:
+\begin{verbatim}
+$module [sum] 
+   $in (a: $uint<32> b: $uint<32>) 
+   $out (c : $uint<32>)
+$is
+{
+   c := (a+b)
+}
+\end{verbatim}
+In this example, {\em sum} is 
+a module which has two inputs and a single output.
+The type of the inputs and the output is \$uint<32>
+which is an unsigned 32-bit integer (More details
+on types are given in Section \ref{sec:Types}).
 
@@
 of statements can be abstractly 
-viewed as a transition
+viewed as a region
 in a petri-net,  and has a 
@@
 The execution of the statement is triggered when
-there is a token present in some subset of its source 
+there is a token present in every place in some specified subsets
+of its source 
 places and when the statement finishes execution,
-a token is placed in each of its output places.
+a token is placed in some specified subset of its sink places.
 
@@
 statement has a single source place and a single
-sink place, sink place of a statement being 
+sink place, with the sink place of a statement being 
 the source place of its successor.
@@
 \begin{itemize}
-\item Complete or atomic statements: these that have a single
+\item Complete or atomic statements: these are statements
+which have a single
 entry place and a single exit place.  These are
@@
 \end{verbatim}
-etc.  
-More details of the syntax are provided in 
+etc.   Expressions can be {\em unary},
+{\em binary} or {\em ternary}, and are
+described in detail in Section \ref{sec:Expressions}.
+Details of the syntax are provided in 
 Section \ref{sec:Syntax}.
@@
 \begin{verbatim}
-call-spec module-name input-argument-values output-argument-targets
+call-spec module-name input-arguments output-arguments
 \end{verbatim}
@@
 // foo has two inputs and three outputs
-$call foo (p,q)  (r,s,t)
+$call foo (p q)  (r s t)
 \end{verbatim}
+Here $p,q,r,s,t$ can be expressions.
 
@@
 The join indicates the arcs (or statements) 
-from whom tokens have to arrive.  When all tokens have arrived
+from whom tokens have to arrive.  After all tokens have arrived
 on the incoming arcs, tokens are sent along the outgoing arcs specified
@@
 controlled by special flow control statements.
-They are structured 
+A branch block is constructed 
 as follows
@@
 \item The place statement: The {\bf place} statement
-defines a place with a label  into which it places
+identifies a place into which it places
 a token after it has executed.  For example,
@@
 means that the incoming token is placed in a labeled
-place ``fastpath'' (the place statement never
+place {\bf fastpath} (the place statement never
 puts a token into its default exit place).  
-The token placed in ``fastpath" must be used to
+The token placed in {\bf fastpath} must be used to
 trigger a unique {\bf merge} statement which is required
@@
 the next statement. 
+\begin{itemize}
+\item 
 The merge assignments inside a merge block are all of a specific
@@
 \end{itemize}
+\end{itemize}
 
@@
 {
-  $merge loopback $entry // $entry is the place
-                     // into the block
-                     // merge triggered by
-                     // token at entry or
-                     // by token in place "loop"
+  $merge loopback $entry // $entry is the entry place
+                         // of the merge statement.
+                         // this merge is triggered by
+                         // a token entering it from
+                         // its predecessor or by  
+                         // a token in the place "loop"
 
@@
   $endmerge
-  r := q + 1
-  $if r < 10 $then 
+  r := (q + 1)
+  $if (r < 10) $then 
        $place [loopback] // put token in place "loopback"
@@
 whenever a token is present in its entry place  or in 
-the place labeled ``loopback''.
-The merge defines variable ``q'' with a value which is
+the place labeled {\bf loopback}
+The merge defines variable {\bf q} with a value which is
 either 0 (if a token was present in the entry place)
@@
 An \Aa program is made of modules which in turn contain statements
-and so on.  The program thus has a hierarchy of scopes with each
+and so on.  The program thus has a hierarchy of scopes 
+(except for the program, each scope is delimited by \{ and \}) with each
 scope being contained in another (except for the program itself,
@@
 \item A declared variable defined in a scope is visible
-in all descendant scopes. In addition
-\begin{itemize}
-\item If $x$ is a storage or pipe variable defined in a scope, then
-there can be no other variable named $x$ in any of the scopes
-contained in this scope.
-\item If $x$ is an input or output argument of a module, then
-there can be no other variable named $x$ declared in any of
-the descendant scopes of this module.
-\end{itemize}
+in all descendent scopes. 
 \item A reference to a variable $b$ in a scope X is resolved by checking
@@
 
-#:a          look for variable a starting from the parent of
+../:a          look for variable a starting from the parent of
              the current scope.
@@
 
-##:a         look for variable starting from the parent of
+../../:a     look for variable starting from the parent of
              the parent of the current scope.
@@
              the parent scope, and so on.  If eventually found, 
-             access the corresponding element of the array.
+             access the corresponding element of the composite
+             object.
 \end{verbatim}
@@
 
-\section{Types}
+\section{Types} \label{sec:Types}
 
-Types in \Aa can be either scalars or arrays.
+Types in \Aa can be either scalar types or composite types.
 
@@
 a binary sequence of the specified width.
+\item Pointers:  A pointer is an unsigned
+integer with a default pointer width
+(set to 32 for now), which specified the
+type of object to which it points.
+For example
+\begin{verbatim}
+$pointer< $uint<32> > 
+\end{verbatim}
+is a pointer which refers to a storage
+object of type \$uint<32>.
 \item Floats: A float is parametrized by
@@
 \end{verbatim}
-respectively.
-\item Pointers:  A pointer type has the form
-\begin{verbatim}
-$pointer<ref-width>
-\end{verbatim}
-A value of this type is considered to be
-a handle to an item whose width is specified
-by ref-width (ref-width must be positive).  
-A pointer value can only be
-referred to and {\bf can never} be used in a
-unary, binary or ternary expression.
+respectively.  Currently, \$float<8,23> 
+and \$float<11,42> are the
+only float types that are supported.
 \end{itemize}
 
+Composite types in \Aa can be either
+array types or record types.
+\begin{itemize}
+\item 
 Array types in \Aa have the form
 \begin{verbatim}
-$array[d1] $of <scalar-type-spec>
+$array [d1][d2]...[dn] $of <element-type-spec>
 \end{verbatim} %$
-Only single dimensional arrays are allowed
-in an \Aa program, and the scalar type spec
-must specify one of the scalar types described above.
+The values $ d1,d2, \ldots dn$ must
+be positive integers, and element-type-spec
+must refer to a type.
+For exampe, 
+\begin{verbatim}
+$array [10][10] of $array [10] $of $uint<32>
+\end{verbatim}
+is an two-dimensional array type whose
+elements are one dimensional arrays of 32-bit
+unsigned integers.
+\item
+Record types in \Aa have the form
+\begin{verbatim}
+$record <type-1> <type-2> ... <type-n>
+\end{verbatim}
+An element of such a record type is
+an aggregate whose first element is of type
+type-1, second element is of type type-2 and so on.
+For example:
+\begin{verbatim}
+$record <$uint<32> > <$array [10] $of $uint<32> > 
+\end{verbatim}
+\end{itemize}
 
-\section{Expressions}
+\section{Expressions} \label{sec:Expressions}
 
@@
 \begin{itemize}
-\item {\bf Object References}: references to an object
-which have the form
+\item Constant literal references.
+\item Simple object references.
+\item Indexed object references.
+\item Pointer de-reference expressions.
+\item Address-of expressions.
+\item Unary expressions.
+\item Binary expressions.
+\item Ternary expressions.
+\end{itemize}
+
+\subsection{Constant literal references}
+
+Constants can be specified in one of many
+ways, depending on their type.
+\begin{itemize}
+\item Integers can be specified in their
+decimal form, as for example 
+\begin{verbatim}
+23
+-8
+\end{verbatim}
+or in binary form, as for example
+\begin{verbatim}
+_b10111
+\end{verbatim}
+When specified in binary form, the twos-complement
+value of the integer being specified should be
+used.
+\item Floats are specified in the exponentiated
+form
+\begin{verbatim}
+_f2.3000e+10
+_f-1.354e+10
+\end{verbatim}
+where the mantissa has to have exactly one digit
+to the left of the decimal point.
+\item Composite constants are specified
+as a space separated list of values.
+\begin{verbatim}
+( 12 32 43 10)
+( (1 5) (2 8) (9 100) )
+\end{verbatim}
+etc.  The elements are listed in 
+row major form.  Thus a[2][2] is
+listed as
+\begin{verbatim}
+a[0][0] a[0][1] a[1][0] a[1][1]
+\end{verbatim}
+\end{itemize}
+
+\subsection{Simple object references}
+
+References to an object have the form
 \begin{verbatim}
@@
 or a child scope of the scope in which the 
-expression appears.   The scope-specifier can be omitted
-if the reference is to be resolved in the same scope
+expression appears (as described in Section \ref{sec:Scopes}).  
+The scope-specifier can be omitted
+if the reference is to be resolved starting from the same scope
 as the expression.
-\item {\bf Unary} Expressions:  There are two
-possible unary expressions.  The {\bf type-cast}
+For example:
+\begin{verbatim}
+../:a
+\end{verbatim}
+
+\subsection{Indexed object references}
+
+These have the form
+\begin{verbatim}
+<scope-specifier>:<root-object-specifier>[i1][i2]...[im]
+\end{verbatim}
+The scope and root-object-specifier are interpreted in the
+same manner as before.  The indices $i1,i2, \ldots im$
+must be non-negative integers.   The 
+object-specifier must be one of
+\begin{itemize}
+\item A storage object whose type is composite.  For
+example, if $a$ is a storage object whose
+type is a two-dimensional composite type, then 
+\begin{verbatim}
+a[0][1]
+\end{verbatim}
+is interpreted in the usual manner.  Thus, when
+the root-object-specifier is a declared storage
+object, the indexed expression evaluates to an
+element of the object.
+\item A scalar object of type pointer.  For example if $ptr$ is a pointer
+to an object of type $T$, then
+\begin{verbatim}
+ptr[1]
+\end{verbatim}
+is a pointer which will point to the adjacent object
+of type T if we assume that ptr points to an array
+of objects of type $T$.  This is similar to the following
+evaluation in {\bf C}
+\begin{verbatim}
+int* ptr;     // ptr points to some int a[i].
+ptr = ptr[1]  // ptr now points to a[i+1] 
+\end{verbatim}
+Now, if type T is a one-dimensional array type,
+say
+\begin{verbatim}
+$array [10] of $uint<32>
+\end{verbatim}
+then the following expression makes sense
+\begin{verbatim}
+ptr[1][2]
+\end{verbatim}
+and is similar to the following evaluation in {\bf C}
+\begin{verbatim}
+int a[10][10];
+ptr = &a[0]      // points to a[0] which is
+                 // array of 10 integers
+ptr = ptr[1][2]; // ptr points to a[1][2]
+\end{verbatim} 
+Thus, the evaluation of an indexed expression
+whose root is a pointer evaluates to a
+pointer of some element type.
+\end{itemize}
+
+\section{Pointer de-reference expressions}
+
+If {\bf ptr} is a pointer to a scalar object,
+then the expression 
+\begin{verbatim}
+->(ptr)
+\end{verbatim}
+refers to the value of the object pointed to
+by {\bf ptr}.
+Such an expression can occur {\bf only} as 
+the target or source of a simple assignment statement.
+For example
+\begin{verbatim}
+a = ->(ptr)
+->(ptr) = (a+1)
+\end{verbatim}
+
+\section{Address-of expressions}
+
+If {\bf a} is a declared {\bf storage} object,
+then
+\begin{verbatim}
+@(a)
+\end{verbatim}
+is a pointer which points to {\bf a}.  Such 
+an expression can occur {\bf only} as a source
+of a simple assignment statement.
+For example
+\begin{verbatim}
+ptr = @(a)
+\end{verbatim}
+
+\section{Unary expressions}
+
+These can be of three types:
+\begin{itemize}
+\item 
+The {\bf cast}
 expression is of the form
@@
 \begin{verbatim}
-($cast ($uint<10>) ##:a)
+($cast ($uint<10>) ../:a)
 \end{verbatim}
-The type-cast expression takes the value of the
+The cast expression takes the value of the
 specified expression and converts it to a value
-of the specified type.
+of the specified type.  This is equivalent
+to a type-cast in {\bf C}, as for example
+in
+\begin{verbatim}
+float a;
+int b = (int) a;
+\end{verbatim}
+
+\item 
+The {\bf bitcast}
+expression is of the form
+\begin{verbatim}
+($bitcast (<type-spec>)  <expression>)
+\end{verbatim} %$
+For example
+\begin{verbatim}
+($bitcast ($uint<10>) a)
+\end{verbatim}
+The bitcast expression takes takes the bits
+corresponding to the value of the specified expression
+and treats them as being of the specified cast type.
+This is similar to the following in {\bf C}
+\begin{verbatim}
+float a;
+int b = *((int*) &a);
+\end{verbatim}
+except that in \Aa, the destination type need not
+be the same size as the source type (higher
+order bits are truncated or padded with 0
+as necessary).
+\item 
 The {\bf bit-wise complement} expression is of the form
@@
 usual way.
-\item {\bf Binary} Expressions: These are of the
+\end{itemize}
+
+The parentheses around a unary expression are
+{\bf required}.
+
+\section{Binary expressions}
+
+These are of the
 form
@@
 GREATEREQUAL      >=
+
+// concatenation operator
+CONCATENATION     &&
+     c = (a && b) means that the bits of
+     a and b are concatenated to produce
+     the bits of c.  a,b must be of type
+     $uint<>
+
+// bit-select operator
+BITSELECT         []
+    c = (a [] i) means that c gets the
+    value of the bit of a with index i 
+    (index 0 is the least significant bit)
 \end{verbatim}
@@
 each expression.
-\item {\bf Ternary} Expressions:  There is only one
+
+\section{Ternary expressions}
+
+There is only one
 form for the ternary expression.
@@
 Note the parentheses delimiting the expression.
-\end{itemize}
+For example
+\begin{verbatim}
+($mux a (b+1) (c+d))
 
@@
 // an array of 32-bit unsigned integers.
-$storage mem $array<1024> of $uint<32> 
+$storage mem: $array<1024> of $uint<32> 
 $module [sel_mod]
@@
     }
-
 }
@@
 an LL(k) parser (written as rules to be parsed by antlr2 \cite{ref:antlr2}).
-The grammar for the parser is (using the EBNF notation) given below. 
-\input{AaParser}
-
-This is the set of tokens recognized by the lexical analyzer (or lexer).
-\input{AaLexer}
+The grammar for the parser is (using the EBNF notation) given in 
+the html file {\bf AaParser.html} which is part of this distribution. 
+The set of tokens recognized by the lexical analyzer (or lexer).
+is available in the html file {\bf AaLexer.html}.
 

### 2011-04-11 Release/examples/pipes/main.c

@@ -1,0 +1,62 @@
+#include <stdlib.h>
+#include <stdint.h>
+
+float read_float32(const char *id);
+uint32_t read_uint32(const char *id);
+
+void write_float32(const char *id, float data);
+void write_uint32(const char *id, uint32_t data);
+
+
+// To get a "software" model, you will need to
+// create two executables which will use the
+// named pipes to communicate.
+//
+// gcc -DFOO -o foo main.c 
+//    will produce foo.
+//
+// gcc -DBAR -o bar main.c 
+//    will produce bar
+//
+// run foo and bar as two distinct processes
+// they will exchange information and terminate.
+//
+// To get the hardware module compile without
+// defining FOO, BAR.  This will generate
+// two VHDL modules, which communicate through
+// the pipes.
+#ifdef FOO
+int main()
+{
+	return(foo(10));
+}
+#endif 
+
+#ifdef BAR
+int main()
+{
+	return(bar());
+}
+#endif
+
+int foo(int a)
+{
+	
+	int b;
+
+	write_uint32("pipe1",a);
+	b = read_uint32("pipe2");
+
+	return(b);
+}
+
+int bar()
+{
+	int b;
+
+	b = read_uint32("pipe1");
+	write_uint32("pipe2",b);
+
+	return(1);
+}
+

### 2011-04-11 source/Ahir-asm/devel/examples/ToVC/BranchBlock/compile.sh

@@ -1,2 +1,3 @@
-Aa2VC BranchBlock.aa | vcFormat > BranchBlock.vc
+Aa2VC -O BranchBlock.aa | vcFormat > BranchBlock.vc
 vc2vhdl -t sum_mod -f BranchBlock.vc | vhdlFormat > system.vhdl
+vc2vhdl -O -t sum_mod -f BranchBlock.vc | vhdlFormat > system_optimized.vhdl

### 2011-04-11 source/libAhirV2/util/vc2vhdl.cpp

@@ -20,9 +20,9 @@ void Handle_Segfault(int signal)
 void Usage_Vc2VHDL()
 {
   cerr << "Usage: " << endl;
-  cerr << "vc2vhdl -f <vc-file> [-f <vc-file>...] -t <top-module> [-t <top-module>...] " << endl << endl;
+  cerr << "vc2vhdl [-O] -f <vc-file> [-f <vc-file>...] -t <top-module> [-t <top-module>...] " << endl << endl;
   cerr << "specify vc-files using -f, top-modules in system using -t.. for example" << endl
-       << "    vc2vhdl -t foo -f file1.vc -f file2.vc -t bar" << endl;
+       << "    vc2vhdl -O -t foo -f file1.vc -f file2.vc -t bar" << endl;
   cerr << "file1.vc and file2.vc will be parsed in order, and the instantiated " << endl
        << "system will have modules foo and bar as top-modules. " << endl;
 }
@@ -30,8 +30,6 @@ void Usage_Vc2VHDL()
 
 int main(int argc, char* argv[])
 {
-
-
 
 
   string sys_name = "test_system";
@@ -61,10 +59,11 @@ int main(int argc, char* argv[])
   string mod_name;
   string opt_string;
 
+  vcSystem::_opt_flag = false;
   while ((opt = 
 	  getopt_long(argc, 
 		      argv, 
-		      "t:f:",
+		      "t:f:O",
 		      long_options, &option_index)) != -1)
     {
       switch (opt)
@@ -84,6 +83,9 @@ int main(int argc, char* argv[])
 	case 't':
 	  mod_name = string(optarg);	
 	  top_modules.insert(mod_name);
+	  break;
+	case 'O':
+	  vcSystem::_opt_flag = true;
 	  break;
 	case '?':		  // incorrect option
 	  opt_string = opt;
@@ -119,6 +121,7 @@ int main(int argc, char* argv[])
 	{
 	  test_system.Set_As_Top_Module(*iter);
 	}
+
       test_system.Elaborate();
       cout << "-- VHDL produced by vc2vhdl from virtual circuit (vc) description " << endl;
       test_system.Print_VHDL(cout);

### 2011-04-11 source/llvmbc2Aa/Utils.cpp

@@ -70,7 +70,7 @@ std::string Aa::get_aa_constant_string(llvm::Constant *konst)
 	  std::ostringstream id;
 
 	  //TODO: revisit.
-	  ret_val =  g->getNameStr();
+	  ret_val =  to_aa(g->getNameStr());
 	} 
       else 
 	{
@@ -186,6 +186,20 @@ std::string Aa::to_aa(std::string x)
   return(ret_string);
 }
 
+bool Aa::is_io_read(IOCode ioc)
+{
+  if(ioc == READ_UINT32 || ioc == READ_FLOAT32)
+    return(true);
+  else
+    return(false);
+}
+bool Aa::is_io_write(IOCode ioc)
+{
+  if(ioc == WRITE_UINT32 || ioc == WRITE_FLOAT32)
+    return(true);
+  else
+    return(false);
+}
 IOCode Aa::get_io_code(Use &u)
 {
   return get_io_code(u.getUser());
@@ -200,7 +214,11 @@ IOCode Aa::get_io_code(User *u)
 IOCode Aa::get_io_code(CallInst &C)
 {
   llvm::Function *f = C.getCalledFunction();
-  assert(f && "function pointers are not currently supported");
+
+  if(f == NULL)
+    {
+      return NOT_IO;
+    }
   
   if (!f->isDeclaration())
     return NOT_IO;
@@ -265,26 +283,29 @@ std::string Aa::get_aa_type_name(const llvm::Type* ptr)
 
       const llvm::SequentialType *ptr_seq = dyn_cast<llvm::SequentialType>(ptr);
       const llvm::Type* el_type = ptr_seq->getElementType();
-
-      if(isa<PointerType>(el_type) || !isa<CompositeType>(el_type))
-	{
-	  int dim = 0;
-	  const llvm::ArrayType* ptr_array = dyn_cast<llvm::ArrayType>(ptr);
-	  if(ptr_array != NULL)
-	    dim = ptr_array->getNumElements();
-	  else
-	    {
-	      const llvm::VectorType* ptr_vec = dyn_cast<llvm::VectorType>(ptr);
-	      dim = ptr_vec->getNumElements();
-	    }
-	  
-	  ret_string = "$array [" + int_to_str(dim) + "] $of ";
-	  ret_string += get_aa_type_name(el_type);
-	}
+      
+      int dim = 0;
+      const llvm::ArrayType* ptr_array = dyn_cast<llvm::ArrayType>(ptr);
[…]

### 2011-04-15 Documents/AhirV2/AhirV2.tex
some more optimizations.. flatten the vc control path and reduce it somewhat...

@@
 \begin{verbatim}
-llvm2aa bytecode.o > bytecode.aa
+llvm2aa [-modules=<listfile>] bytecode.o > bytecode.aa
 \end{verbatim}
@@
 messages are sent to {\bf stderr}.  On success, the tool returns 0.
+
+The option 
+\begin{verbatim}
+-modules=listfile
+\end{verbatim}
+can be used to specify the list of functions in the bytecode
+which should be converted to \Aa.  If absent, all functions
+are converted.
+
+\subsection{{\bf AaLinkExtMem}}
+
+This tool takes a list of \Aa files, elaborates the program,
+does memory space decomposition.  The externally visible memory space is
+linked in one of two ways: either it is assumed to be external
+and all accesses to it are routed out of the \Aa program,
+or it is assumed to be internal and assumed to consist of
+a memory object (an array of bytes).  External pointer dereferences
+are handled as if they are directed at this memory object.
+\begin{verbatim}
+AaLinkExtMem [-I n] [-E obj-name] bytecode.o > bytecode.aa
+\end{verbatim}
+The generated \Aa code is sent to {\bf stdout} and all informational
+messages are sent to {\bf stderr}.  On success, the tool returns 0.
+
+The options:
+\begin{itemize}
+\item {\bf -I n}: specifies that external references to memory
+are to be mapped as if they are to an internal object whose size
+is $n$ bytes.
+\item {\bf -E obj-name} : specifies that the object to which
+external references are mapped is to be named obj-name.
+\end{itemize}
+If the {\bf -I} option is not used, then all external memory
+references are routed out of the \Aa program through pipes.
 
@@
 \begin{verbatim}
-Aa2VC [-O] file1.aa file2.aa ...  > result.vc
+Aa2VC [-O] [-I obj-name] file1.aa file2.aa ...  > result.vc
 \end{verbatim}
@@
 
-The tool performs memory space decomposition and if the {\bf -O} option
-is specified, it also attempts to parallelize sequential code using 
-dependency analysis.
+The options:
+\begin{itemize}
+\item {\bf -O} : if used, sequential statement blocks are parallelized
+by doing dependency analysis.
+\item {\bf -I obj-name} : if specified, all external memory references
+are considered as being directed at the storage object named obj-name.
+\end{itemize}
 
@@
 and similarly use {\bf vhdlFormat} to format generated
-VHDL file.
+VHDL files.
 
@@
 
+\section{External and internal memory spaces}
+
+Consider the following C program:
+\begin{verbatim}
+int main(int* b)
+{
+   int q[2];
+   q[0] = *b;
+   q[1] = q[0];
+   return(q[1]);
+}
+\end{verbatim}
+When this program is mapped to a circuit, we identify 
+two distinct memory spaces, one which contains the
+array $q$ and the other corresponding to the 
+external world (the one referred to by the pointer $b$).
+Where is the external memory physically located?  In 
+the AhirV2 flow, we can either locate it outside
+the system or inside the system which is being 
+described by this program.
+
+If the external memory is to be placed outside, then
+accesses to it from within the system must be routed
+outside the system.  On the other hand if it is
+to be placed inside, a storage object corresponding 
+to it must be created and all accesses to the external
+memory must be directed at this storage object.
+Further, the external world must have a mechanism for
+accessing this storage object.
+
+Both options are supported in the AhirV2 flow through
+the utility {\bf AaLinkExtMem} described earlier.
+
 \section{Handling complex programs}

### 2011-04-15 source/Ahir-asm/devel/include/AaRoot.h

@@ -115,6 +115,22 @@ class AaRoot
 
   virtual void Err_Check() {}
 
+  virtual string Get_Source_Info() 
+  {
+     return(string("file ") 
+	   + this->Get_File_Name() 
+	   + ", line " 
+	   + 
+	   IntToStr(this->Get_Line_Number())); 
+  }
+
+  virtual string Debug_Info()
+  {
+    return(this->Get_Source_Info());
+  };
+
+
+
 };
 
 struct AaRootCompare:public binary_function

### 2011-04-15 source/Ahir-asm/devel/src/AaExpression.cpp

@@ -1530,6 +1530,7 @@ AaPointerDereferenceExpression::AaPointerDereferenceExpression(AaScope* scope,
   
   obj_ref->Add_Target(this);
   AaProgram::Add_Storage_Dependency_Graph_Vertex(this);
+  AaProgram::_pointer_dereferences.insert(this);
 }
 
 
@@ -1540,7 +1541,7 @@ void AaPointerDereferenceExpression::Print(ostream& ofile)
   this->_reference_to_object->Print(ofile);
   ofile << ")";
   
-  if(_addressed_object_representative == NULL)
+  if(_addressed_object_representative == NULL && !AaProgram::_keep_extmem_inside)
     {
       AaRoot::Error("illegal pointer-dereference expression... not associated with any memory space!", this);
     }
@@ -1613,11 +1614,6 @@ void AaPointerDereferenceExpression::Propagate_Addressed_Object_Representative(A
 	    AaProgram::Add_Storage_Dependency(this,obj);
 	}
       
-      if(obj->Is_Foreign_Storage_Object())
-	{
-	  AaProgram::Add_ExtMem_Access_Width(this->Get_Type()->Size());
-	}
-
       this->Set_Addressed_Object_Representative(obj);
       
 
@@ -1677,6 +1673,8 @@ void AaPointerDereferenceExpression::Write_VC_Control_Path( ostream& ofile)
 void AaPointerDereferenceExpression::Write_VC_Control_Path_As_Target( ostream& ofile)
 {
   ofile << "// " << this->To_String() << endl;
+
+
   if((this->Get_Addressed_Object_Representative() == NULL)
      || this->Get_Addressed_Object_Representative()->Is_Foreign_Storage_Object())
     {

### 2011-04-15 source/Ahir-asm/devel/src/AaModule.cpp

@@ -295,19 +295,35 @@ void AaModule::Set_Foreign_Object_Representatives()
 	    {
 	      AaRoot::Info("input argument " + inobj->Get_Name() + " of module " + this->Get_Label()
 			   + " points to foreign storage ");
-	      AaType* el_type = ((AaPointerType*)inobj->Get_Type())->Get_Ref_Type();
 
-	      AaForeignStorageObject* fobj = AaProgram::Make_Foreign_Storage_Object(el_type);
+	      AaStorageObject* fobj = NULL;
+	      if(!AaProgram::_keep_extmem_inside)
+		{
+		  AaType* el_type = ((AaPointerType*)inobj->Get_Type())->Get_Ref_Type();
+		  fobj = AaProgram::Make_Foreign_Storage_Object(el_type);
+		}
+	      else
+		fobj = AaProgram::Get_Extmem_Object();
+
 	      fobj->Add_Source_Reference(inobj);  // inobj uses fobj as a source
 	      inobj->Add_Target_Reference(fobj);  // fobj uses inobj as a target
 	      inobj->Propagate_Addressed_Object_Representative(fobj);
 	    }
 	  else
 	    {
+	      AaStorageObject* fobj = NULL;
+
 	      // not a pointer, so we cannot say anything about the
 	      // type of the object to which it points.
-	      AaForeignStorageObject* fobj = 
-		AaProgram::Make_Foreign_Storage_Object(AaProgram::Make_Void_Type());
+	      if(!AaProgram::_keep_extmem_inside)
+		{
+		  fobj = AaProgram::Make_Foreign_Storage_Object(AaProgram::Make_Void_Type());
+		}
+	      else
+		fobj = AaProgram::Get_Extmem_Object();
+
+
+
 	      fobj->Add_Source_Reference(inobj);  // inobj uses fobj as a source
 	      inobj->Add_Target_Reference(fobj);  // fobj uses inobj as a target
 	      inobj->Propagate_Addressed_Object_Representative(fobj);
@@ -324,8 +340,16 @@ void AaModule::Set_Foreign_Object_Representatives()
 		    
 	      AaRoot::Info("output argument " + outobj->Get_Name() + " of module " + this->Get_Label()
 			   + " points to foreign storage ");
-	      AaType* el_type = ((AaPointerType*)outobj->Get_Type())->Get_Ref_Type();
[…]

### 2011-04-15 source/Ahir-asm/devel/src/AaUtil.cpp

@@ -236,7 +236,7 @@ string To_Alphanumeric(string x)
   return(ret_string);
 }
 
-int LCM(set<int>::iterator iL, set<int>::iterator iR)
+int GCD(set<int>::iterator iL, set<int>::iterator iR)
 {
   int ret_val = 0;
   if(iL != iR)
@@ -246,7 +246,7 @@ int LCM(set<int>::iterator iL, set<int>::iterator iR)
       set<int>::iterator nextiL = iL;
       nextiL++;
 
-      int r_val = LCM(nextiL,iR);
+      int r_val = GCD(nextiL,iR);
 
       if(r_val > 0)
 	{
@@ -260,9 +260,9 @@ int LCM(set<int>::iterator iL, set<int>::iterator iR)
   return(ret_val);
 }
 
-int LCM(set<int>& s)
+int GCD(set<int>& s)
 {
-  return(LCM(s.begin(), s.end()));
+  return(GCD(s.begin(), s.end()));
 }
 
 string Augment_Hier_Id(string hid, string suffix)

### 2011-04-15 source/libAhirV2/include/vcControlPath.hpp

@@ -74,10 +74,8 @@ public:
   virtual int64_t Get_Index() {return(this->_index);}
 
 
-  virtual void Construct_CPElement_Group_Graph(vcControlPath* cp);
-  virtual void Construct_CPElement_Group_Graph(vcControlPath* cp,
-					       vector<vcCPElement*>& preds,
-					       vector<vcCPElement*>& succs);
+  virtual void Construct_CPElement_Group_Graph_Vertices(vcControlPath* cp) {}
+  virtual void Connect_CPElement_Group_Graph(vcControlPath* cp);
   
   virtual vcCPElement* Get_Exit_Element()
   {
@@ -130,6 +128,8 @@ public:
       }
     return(ret_string);
   }
+
+
 
 
   friend class vcControlPath;
@@ -193,6 +193,7 @@ public:
   string Get_DP_To_CP_Symbol();
   string Get_CP_To_DP_Symbol();
 
+  virtual void Construct_CPElement_Group_Graph_Vertices(vcControlPath* cp);
 
 };
 
@@ -206,7 +207,7 @@ public:
   virtual bool Is_Place() {return(true);}
 
   virtual void Print_VHDL(ostream& ofile);
-
+  virtual void Construct_CPElement_Group_Graph_Vertices(vcControlPath* cp);
 
   friend class ControlPath;
 
@@ -254,7 +255,8 @@ public:
   virtual void Print_VHDL_Exit_Symbol_Assignment(ostream& ofile);
 
   void Print_Missing_Elements(set<vcCPElement*>& visited_set); // print to cerr
-  virtual void Construct_CPElement_Group_Graph(vcControlPath* cp);
+  virtual void Construct_CPElement_Group_Graph_Vertices(vcControlPath* cp);
+  virtual void Connect_CPElement_Group_Graph(vcControlPath* cp);
 
   virtual vcCPElement* Get_Exit_Element()
   {
@@ -349,6 +351,7 @@ public:
 };
 
 
+class vcControlPath;
 class vcCPElementGroup: public vcRoot
 {
   int64_t _group_index;
@@ -357,6 +360,8 @@ class vcCPElementGroup: public vcRoot
   set<vcCPElementGroup*> _successors;
   set<vcCPElementGroup*> _predecessors;
 
+  bool _has_transition;
+  bool _has_place;
   bool _has_input_transition;
   bool _has_output_transition;
   bool _has_dead_transition;
[…]

### 2011-04-19 source/Ahir-asm/devel/docs/AaLRM.tex
Some more bugs. combinational cycles through register_bank due to over-aggressive timing. removed latch in InputMuxBase which should not have been present.

@@
 a module which has two inputs and a single output.
-The type of the inputs and the output is \$uint<32>
+The type of the inputs and the output is 
+\begin{verbatim}
+$uint<32>
+\end{verbatim}
 which is an unsigned 32-bit integer (More details
@@
 is a pointer which refers to a storage
-object of type \$uint<32>.
+object of type \begin{verbatim}$uint<32>\end{verbatim}
 \item Floats: A float is parametrized by
@@
 \end{verbatim}
-respectively.  Currently, \$float<8,23> 
-and \$float<11,42> are the
+respectively.  Currently, these are the
 only float types that are supported.
@@
 \end{verbatim}
+\item Named record types are used to avoid the circular
+reference problem (for example, when a record has a pointer to
+itself).
+\begin{verbatim}
+$record [myrec] <$uint<32> >  <$pointer<myrec> >
+\end{verbatim}
+defines a record type with name myrec, one of whose
+fields is a pointer to myrec.
 \end{itemize}
@@
 \begin{verbatim}
-a = ->(ptr)
-->(ptr) = (a+1)
+a := ->(ptr)
+->(ptr) := (a+1)
 \end{verbatim}
@@
 \begin{verbatim}
-ptr = @(a)
+ptr := @(a)
 \end{verbatim}

### 2011-04-19 source/Ahir-asm/devel/examples/ToC/Pointers/aa_c_model.h

@@ -7,7 +7,7 @@ typedef struct increment_State__
   unsigned int _assign_line_12_entry:1;
   unsigned int _assign_line_12_in_progress:1;
   unsigned int _assign_line_12_exit:1;
-  uint_32 *incr_t1;
+    uint_32 (*incr_t1);
   unsigned int _assign_line_13_entry:1;
   unsigned int _assign_line_13_in_progress:1;
   unsigned int _assign_line_13_exit:1;
@@ -36,23 +36,31 @@ typedef struct passpointer_State__
   unsigned int _assign_line_25_entry:1;
   unsigned int _assign_line_25_in_progress:1;
   unsigned int _assign_line_25_exit:1;
-  uint_32 *pu;
+    uint_32 (*pu);
   unsigned int _assign_line_26_entry:1;
   unsigned int _assign_line_26_in_progress:1;
   unsigned int _assign_line_26_exit:1;
-  uint_32 *pv;
+    uint_32 (*pv);
   unsigned int _assign_line_27_entry:1;
   unsigned int _assign_line_27_in_progress:1;
   unsigned int _assign_line_27_exit:1;
-  uint_32 *px;
+  uint_32 t;
   unsigned int _assign_line_28_entry:1;
   unsigned int _assign_line_28_in_progress:1;
   unsigned int _assign_line_28_exit:1;
+  uint_32 s;
+  unsigned int _assign_line_29_entry:1;
+  unsigned int _assign_line_29_in_progress:1;
+  unsigned int _assign_line_29_exit:1;
+  uint_32 t1;
+  unsigned int _assign_line_30_entry:1;
+  unsigned int _assign_line_30_in_progress:1;
+  unsigned int _assign_line_30_exit:1;
   uint_32 q;
-  unsigned int _call_line_29_entry:1;
-  unsigned int _call_line_29_in_progress:1;
-  unsigned int _call_line_29_exit:1;
-  increment_State *_call_line_29_called_fn_struct;
+  unsigned int _call_line_31_entry:1;
+  unsigned int _call_line_31_in_progress:1;
+  unsigned int _call_line_31_exit:1;
+  increment_State *_call_line_31_called_fn_struct;
   unsigned passpointer_entry:1;
   unsigned passpointer_in_progress:1;
   unsigned passpointer_exit:1;

### 2011-04-20 Documents/AhirV2/AhirV2.tex
Some minor bugs. also updating the documentation.

@@
 
-\title{AhirV2: from algorithms to hardware}
+\title{AhirV2: from algorithms to hardware \\ A tutorial introduction}
 \author{Madhav Desai \\ Department of Electrical Engineering \\ Indian Institute of Technology \\
@@
 We assume that you have access to either {\bf llvm-gcc}
-of {\bf clang} as the front-end compiler which generates
+or {\bf clang} as the front-end compiler which generates
 LLVM byte-code from C/C++.  The current AhirV2 toolset
@@
 \begin{verbatim}
-llvm2aa [-modules=<listfile>] bytecode.o > bytecode.aa
+llvm2aa [-modules=<listfile>] [-storageinit] bytecode.o > bytecode.aa
 \end{verbatim}
@@
 
-The option 
-\begin{verbatim}
--modules=listfile
-\end{verbatim}
-can be used to specify the list of functions in the bytecode
-which should be converted to \Aa.  If absent, all functions
+The options:
+\begin{itemize}
+\item {\bf -modules=listfile} : Specify the list of functions in the bytecode
+which should be converted to \Aa.   The names of these functions should be
+listed in the text-file listfile. If absent, all functions
 are converted.
+\item {\bf -storageinit} :  Storage objects in the llvm bytecode
+are explicitly initialized in the generated \Aa code.   An initializer
+routine {\bf global\_storage\_initializer\_} is instantiated in
+the \Aa code for this purpose.
+\end{itemize}
 
@@
 \begin{verbatim}
-vc2vhdl -t foo [-t bar -t bar2 ...]  -f file1.vc -f file2.vc ...  > system.vhdl
+vc2vhdl [-O] -t foo [-t bar -t bar2 ...]  -f file1.vc -f file2.vc ...  > system.vhdl
 \end{verbatim}
-Using the {\bf -t} option, one can specify the modules which are to be 
+
+The options:
+\begin{itemize}
+\item {\bf -t} : to specify the modules which are to be 
 accessible from the ports of the generated VHDL system.
-Using the {\bf -f} option, one specifies the \vC files to be analyzed.
+Multiple top-level modules can be specified in this way.
+\item {\bf -f} : specifies the \vC files to be analyzed. 
+Multiple \vC files may be specified.
+\item {\bf -O} : optimize the generated VHDL by compacting
+the control-path.  This does not change the resulting
+hardware, but makes the generated VHDL file smaller.
+\end{itemize}
 
@@
 
+\subsection{Keeping the external memory outside the system}
+
+For this example, if you want to keep the external memory
+outside, you will have to go through the following sequence
+\begin{verbatim}
+# first use clang (or llvm-gcc) to generate llvm-byte-code
+clang -std=gnu89 -emit-llvm -c foo.c
+#
+# disassemble so that you can make sense of the llvm bc.
+llvm-dis foo.o
+#
+# OK, now take the llvm byte code
+# and generate an Aa description.
+# use the storageinit option to initialize
+# global storage.
+# (the pipe to vcFormat is to prettify the output)
+ llvm2aa -storageinit foo.o | vcFormat > foo.o.aa
+#
+#
+# Do an Aa -> Aa transformation: map external
+# memory outside..
+AaLinkExtMem foo.o.aa | vcFormat > foo.o.memlinked.ExternalOutside.aa
+#
+# Now take the Aa code and generate a virtual
+# circuit..
+# the -O flag does dependency analysis in straight-line
+# code and parallelizes it.
+#
+Aa2VC -O foo.o.memlinked.ExternalOutside.aa | vcFormat\
+                 > foo.o.memlinked.ExternalOutside.aa.vc
+#
+# finally, generate vhdl from the vc description.  Note that
+# you will have to mark the module foo as well as the
+# extmem_store_32/load_32 modules as top-level modules
+# so that it is possible for the outside world to serve
+# requests made from inside.
+#
+vc2vhdl -O -t foo -t extmem_store_32 -t extmem_load_32\ 
+         -f foo.o.memlinked.ExternalOutside.aa.vc | vhdlFormat\
+              > foo_o_aa_memlinked_external_outside_vc.vhdl
+
+\end{verbatim}
+
+If you look at the generate top-level VHDL entity, its ports
+will be
+\begin{verbatim}
+entity test_system is  -- system
+  port (--
+    foo_b : in  std_logic_vector(31 downto 0);
+    foo_ret_val_x_x : out  std_logic_vector(31 downto 0);
+    foo_tag_in: in std_logic_vector(0 downto 0);
+    foo_tag_out: out std_logic_vector(0 downto 0);
+    foo_start : in std_logic;
+    foo_fin   : out std_logic;
+    clk : in std_logic;
+    reset : in std_logic;
+    extmem_read_address_32_pipe_read_data: out std_logic_vector(31 downto 0);
+    extmem_read_address_32_pipe_read_req : in std_logic_vector(0 downto 0);
+    extmem_read_address_32_pipe_read_ack : out std_logic_vector(0 downto 0);
+    extmem_read_data_32_pipe_write_data: in std_logic_vector(31 downto 0);
+    extmem_read_data_32_pipe_write_req : in std_logic_vector(0 downto 0);
+    extmem_read_data_32_pipe_write_ack : out std_logic_vector(0 downto 0);
+    extmem_write_address_32_pipe_read_data: out std_logic_vector(31 downto 0);
+    extmem_write_address_32_pipe_read_req : in std_logic_vector(0 downto 0);
+    extmem_write_address_32_pipe_read_ack : out std_logic_vector(0 downto 0);
+    extmem_write_data_32_pipe_read_data: out std_logic_vector(31 downto 0);
+    extmem_write_data_32_pipe_read_req : in std_logic_vector(0 downto 0);
+    extmem_write_data_32_pipe_read_ack : out std_logic_vector(0 downto 0)); --
+  --
+end entity;
+\end{verbatim}
+The external memory read and write address and data are clearly
+visible.   The outside world is responsible for serving the
+read/write requests made from the inside.
+
+\subsection{Keeping the external memory inside the system}
+
+For this example, if you want to keep the external memory
+outside, you will have to go through the following sequence
+\begin{verbatim}
+# use clang (or llvm-gcc) to generate llvm-byte-code
+clang -std=gnu89 -emit-llvm -c foo.c
+#
+# disassemble so that you can make sense of the llvm bc.
+llvm-dis foo.o
+#
+# OK, now take the llvm byte code
+# and generate an Aa description.
+# use the storageinit option to initialize
+# global storage.
+# (the pipe to vcFormat is to prettify the output)
+ llvm2aa -storageinit foo.o | vcFormat > foo.o.aa
+#
+#
+# Do an Aa -> Aa transformation: map external
+# memory to a storage area inside the system...
+# -I 1024 says that the amount of memory that will be
+# referred to is 1024 bytes.
+# -E mempool says that the storage object corresponding
+# to external memory is named mempool.
+AaLinkExtMem  -I 1024 -E mempool foo.o.aa | vcFormat\
+               > foo.o.memlinked.ExternalInside.aa
+#
+# Now take the Aa code and generate a virtual
+# circuit..
+# the -O flag does dependency analysis in straight-line
+# code and parallelizes it.
+# the -I mempool option says that external memory is
+# to be mapped inside the system to object mempool..
+#
+Aa2VC -O -I mempool foo.o.memlinked.ExternalInside.aa\
+           | vcFormat > foo.o.memlinked.ExternalInside.aa.vc
+#
+# finally, generate vhdl from the vc description.
+# note that you will have to mark mem_load__ and mem_store__
+# as top-level modules, so that the external world can
+# access its memory pool inside the system.
+#
+vc2vhdl -O -t foo -t mem_load__ -t mem_store__ \ 
+     -f foo.o.memlinked.ExternalInside.aa.vc\
+       | vhdlFormat > foo_o_aa_memlinked_external_inside_vc.vhdl
+\end{verbatim}
+
+The generated top-level VHDL entity has the following
+ports:
+\begin{verbatim}
+entity test_system is  -- system
+  port (--
+    foo_b : in  std_logic_vector(31 downto 0);
+    foo_ret_val_x_x : out  std_logic_vector(31 downto 0);
+    foo_tag_in: in std_logic_vector(0 downto 0);
+    foo_tag_out: out std_logic_vector(0 downto 0);
+    foo_start : in std_logic;
+    foo_fin   : out std_logic;
+    mem_load_x_x_address : in  std_logic_vector(31 downto 0);
+    mem_load_x_x_data : out  std_logic_vector(7 downto 0);
+    mem_load_x_x_tag_in: in std_logic_vector(0 downto 0);
+    mem_load_x_x_tag_out: out std_logic_vector(0 downto 0);
+    mem_load_x_x_start : in std_logic;
+    mem_load_x_x_fin   : out std_logic;
+    mem_store_x_x_address : in  std_logic_vector(31 downto 0);
+    mem_store_x_x_data : in  std_logic_vector(7 downto 0);
+    mem_store_x_x_tag_in: in std_logic_vector(0 downto 0);
+    mem_store_x_x_tag_out: out std_logic_vector(0 downto 0);
+    mem_store_x_x_start : in std_logic;
+    mem_store_x_x_fin   : out std_logic;
+    clk : in std_logic;
+    reset : in std_logic); --
+  --
+end entity;
+\end{verbatim}
+
+The system provides a memory load and memory store
+port to the external world (through the mem\_load.. and mem\_store..)
+ports.
+
 \section{Handling complex programs}

### 2011-04-21 source/CtestBench/vhpi/src/Vhpi.c
started adding the CtestBench stuff..

@@ -1,0 +1,632 @@
+// author: Madhav P. Desai
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include <assert.h>
+#include <signal.h>
+#include <sys/types.h>
+#include <sys/socket.h>
+#include <netinet/in.h>
+#include <netdb.h>
+#include "Vhpi.h"
+
+#define MAX_CLIENT_SOCKET_COUNT 65536
+#define z32__ "00000000000000000000000000000000"
+
+#define APPEND(lst,lnk) if(lst.tail)\
+      {\
+	lst.tail->next = lnk;\
+	lnk->prev = lst.tail;\
+	lnk->next = NULL;\
+	lst.tail = lnk;\
+      }\
+    else\
+      {\
+	lst.head = lnk;\
+	lst.tail = lnk;\
+	lnk->prev = NULL;\
+	lnk->next = NULL;\
+      }
+
+#define REMOVE(lst,lnk) if((lst.head == lnk) && (lst.tail == lnk))	\
+    {\
+      lst.head = NULL;\
+      lst.tail = NULL;\
+      lnk->next = lnk->prev = NULL;\
+    }\
+  else if(lst.head == lnk)\
+    {\
+      lst.head = lnk->next;\
+      lnk->next->prev = NULL;\
+      lnk->next = lnk->prev = NULL;\
+    }\
+  else if(lst.tail == lnk)\
+    {\
+      lst.tail = lnk->prev;\
+      lnk->prev->next = NULL;\
+      lnk->prev = lnk->next = NULL;\
+    }\
+  else\
+    {\
+      lnk->prev->next = lnk->next;\
+      lnk->next->prev = lnk->prev;\
+      lnk->prev = lnk->next = NULL;\
+    }
+
+
+// maximum id..
+int max_socket_id = 0;
+
+// server socket
+int server_socket_id = -1;
+
+// client socket array
+int client_sockets[MAX_CLIENT_SOCKET_COUNT];
+int client_socket_count = 0;
+
+// joblist.
+JobList new_jobs;
+JobList active_jobs;
+JobList finished_jobs;
+
+///////////////////////////////////////////////////////////////////////////////////////////////
+// some utility functions.
+///////////////////////////////////////////////////////////////////////////////////////////////
+
+// convert 32 character string pointed to by a to an 
+// unsigned.
+unsigned Bit_Vector_To_Unsigned(char* a)
+{
+  unsigned ret_val = 0;
+  int index;
+  if(a != NULL)
+    {
[…]

### 2011-04-26 Documents/AhirV2/AhirV2.tex
quicksort runs, C testbench runs.. alpha label can be applied..

@@
 \end{itemize}
+It is better if you use the {\bf -I} and {\bf -E} options.
+
 If the {\bf -I} option is not used, then all external memory
 references are routed out of the \Aa program through pipes.
+In this case, if the \Aa compiler determines that there is some pointer 
+in the program which can point
+to both internal and external memory, then this will be
+declared as an error!
 
@@
 \begin{verbatim}
-Aa2VC [-O] [-I obj-name] file1.aa file2.aa ...  > result.vc
+Aa2VC [-O] [-C] [-I obj-name] file1.aa file2.aa ...  > result.vc
 \end{verbatim}
@@
 by doing dependency analysis.
+\item {\bf -C} : if used, a C stub is created for every module that
+is not called from within the system.  These stubs can be used to
+interface to a VHDL simuator (or even drive hardware) to verify
+the VHDL code generated by downstream tools.
 \item {\bf -I obj-name} : if specified, all external memory references
@@
 \begin{verbatim}
-vc2vhdl [-O] -t foo [-t bar -t bar2 ...]  -f file1.vc -f file2.vc ...  > system.vhdl
+vc2vhdl [-O] [-C] -t foo [-t bar -t bar2 ...]\
+         -f file1.vc -f file2.vc ...  > system.vhdl
 \end{verbatim}
@@
 hardware, but makes the generated VHDL file smaller.
+\item {\bf -C} : the VHDL code has a system test bench which
+interfaces to foreign code using a VHPI/Modelsim-FLI interface.
+If this is not specified, the  generated test bench simply
+instantiates the system and starts all top-level modules
+off (you will need to fill in your own test bench here).
+The C testbench is usually easier to write (it probably
+already exists in the form of the original program).
 \end{itemize}

### 2011-04-26 Ahir-asm/antlr2/examples/ToC/CallModules/compile.sh

@@ -1,7 +1,0 @@
-../../../bin/Aa2C CallModules.aa
-indent aa_c_model.c
-indent aa_c_model.h
-gcc -g -c -I../../../include -I./ aa_c_model.c
-gcc -g -c -I../../../include -I./ driver.c
-gcc -g -o driver driver.o aa_c_model.o
-rm *.o *~

### 2011-04-26 source/Ahir-asm/devel/examples/ToC/ParallelBlock/aa_c_model.c

@@ -49,7 +49,7 @@ sum_mod_ (sum_mod_State * __top)
 	    {
 	      if (1)
 		{
-		  (__top->p.q).__val = __PLUS ((__top->a).__val, (__top->b).__val);	//  file ParallelBlock.aa, line 8
+		  (__top->p.q).__val = __PLUS ((__top->a).__val, (__top->b).__val);	// file ParallelBlock.aa, line 8
 		  __top->p._assign_line_8_entry = 0;
 		  __top->p._assign_line_8_exit = 1;
 		}
@@ -64,7 +64,7 @@ sum_mod_ (sum_mod_State * __top)
 	    {
 	      if (1)
 		{
-		  (__top->p.r).__val = __PLUS ((__top->c).__val, (__top->d).__val);	//  file ParallelBlock.aa, line 9
+		  (__top->p.r).__val = __PLUS ((__top->c).__val, (__top->d).__val);	// file ParallelBlock.aa, line 9
 		  __top->p._assign_line_9_entry = 0;
 		  __top->p._assign_line_9_exit = 1;
 		}
@@ -96,7 +96,7 @@ sum_mod_ (sum_mod_State * __top)
 	{
 	  if (1)
 	    {
-	      (__top->result).__val = __PLUS ((__top->p.q).__val, (__top->p.r).__val);	//  file ParallelBlock.aa, line 11
+	      (__top->result).__val = __PLUS ((__top->p.q).__val, (__top->p.r).__val);	// file ParallelBlock.aa, line 11
 	      __top->_assign_line_11_entry = 0;
 	      __top->_assign_line_11_exit = 1;
 	    }

### 2011-04-27 v2/Ahir-asm/devel/docs/AaLRM.tex
reorganization

--- 2011-04-19 source/Ahir-asm/devel/docs/AaLRM.tex
@@
 \newcommand{\Aa}{{\bf Aa}~}
+\newcommand{\vC}{{\bf vC}~}
 
@@
 $else
-   qdash :  (r - s)
+   qdash :=  (r - s)
 $endif
@@
 The resulting control flow structure is a
-type-2 petri-net with provable liveness and safety properties \cite{ref:SameerPhD}.
+petri-net with provable liveness and safety properties \cite{ref:SameerPhD}.
 
@@
 
-\section{Pointer de-reference expressions}
+\subsection{Pointer de-reference expressions}
 
@@
 
-\section{Address-of expressions}
+\subsection{Address-of expressions}
 
@@
 
-\section{Unary expressions}
+\subsection{Unary expressions}
 
@@
 
-\section{Binary expressions}
+\subsection{Binary expressions}
 
@@
 
-\section{Ternary expressions}
+\subsection{Ternary expressions}
 
@@
 ($mux a (b+1) (c+d))
+\end{verbatim}
+
+\section{Memory Spaces}
+
+An \Aa program can contain declarations to storage objects.  These
+storage objects need to be grouped into memory spaces using
+the following rule: two storage objects are put in the same
+memory space if there is a value in the program which
+is infected by (determined by) the values of pointers to the two objects.
+
+To understand this concept, consider the following
+example
+\begin{verbatim}
+$module [foo] 
+  $in (a: $uint<32>)
+  $out (b: $uint<32>) 
+$is
+{
+   $storage u: $uint<32>
+   $storage v: $uint<32>
+   $storage w: $uint<32>
+
+   w := a
+   ptr := @(u)
+   ->(ptr) := w
+   ptr := @(v)
+   ->ptr := u
+   b := u
+}
+\end{verbatim}
+Now consider the object {\bf ptr}, which can hold a value which
+is a pointer to either {\bf u} or a pointer to {\bf v}.  Thus,
+{\bf u} and {\bf v} must belong to the same memory space.
+However, {\bf w} can sit in a memory space by itself.  The
+store accesses implied by the $->(ptr)$ will point to the
+memory space which contains {\bf u} and {\bf v}.
+
+What should we do in the following case?
+\begin{verbatim}
+$module [foo]
+   $in (a : $uint<32>)
+   $out (b: $uint<32>)
+$is
+{
+   $storage u: $uint<32> 
+   $storage ptr: $uint<32>
+   ptr := ($bitcast ($pointer<$uint<32> >) a)
+   tmp := ->(ptr)
+   ptr := @(u)
+   ->(ptr) := tmp
+   b := u
+}
+\end{verbatim}
+In this case, {\bf ptr} takes the value of $a$,
+and as a pointer, it must point to 
+an {\bf external} object.  Since ptr can 
+point to this external object and also to {\bf u},
+the external object and {\bf u} must be kept in
+the same memory space.  In this case, the external
+object needs
+to be made internal, and moved {\bf inside} the system.  
+This is done by the {\bf AaLinkExtMem} utility
+which is described in a separate document.
 

### 2011-04-27 v2/llvmbc2Aa/docs/llvm2aa.tex
reorganization

--- 2011-04-26 Documents/AhirV2/AhirV2.tex
@@
 
-\title{AhirV2: from algorithms to hardware \\ A tutorial introduction}
+\title{About {\bf llvm2aa}}
 \author{Madhav Desai \\ Department of Electrical Engineering \\ Indian Institute of Technology \\
@@
 
-We describe the essential features
-of the second version of the AhirV2 toolset
-developed at IIT-Bombay.
+{\bf llvm2aa}  is a tool which reads in LLVM byte-code (see http://www.llvm.org for
+details about LLVM) and produces \Aa code which can then be further used
+to produce VHDL using the AhirV2 tool chain developed at IIT Bombay.
 
-\section{What is AhirV2?}
+\section{Synopsys}
 
-Perhaps the simplest way to understand AhirV2 is
-through an example.
-
-Suppose we start with a simple program.
-\begin{verbatim}
-int add(int a, int b)
-{
-        int c = (a+b);
-        return(c);
-}
-\end{verbatim}
-This program describes an algorithm.  We normally
-compile this program and convert it to a machine-level
-program which is executed on a processor.
-
-What if we wished to convert this program to
-a circuit?  
-Such a circuit would have inputs 
-$a,b$ and an output $c$ (together with some
-interface hand-shake signals).  When activated,
-the circuit should read its inputs and compute 
-a response which is the same as would be expected
-if the program were executed on a computer.
-
-AhirV2 consists of a set of tools which takes
-a C/C++ program and converts this collection to
-a hardware circuit (described in VHDL).   This
-conversion is done in two steps:
-\begin{itemize}
-\item The high-level program is compiled
-to an interemediate assembly form.  AhirV2
-introduces an intermediate assembly language
-\Aa  which can serve as a target for sequential
-programming languages (such as C/C++) as
-well as for parallel programming languages.
-\item From the \Aa description, a virtual
-circuit (described in a virtual circuit
-description language \vC) is generated.
-The chief optimization carried out at this
-step is that decomposing the system memory
-into disjoint segments based on usage
-analysis (this considerably improves the
-available memory bandwidth and reduces
-system cost).
-\item From the \vC description, a
-VHDL description of the system is generated.
-The chief optimization carried out at this
-stage is resource sharing.  The \vC description
-is analyzed to identify operations which
-cannot be concurrently active and this information
-is used to reduce the hardware required.
-\item 
-The VHDL description produced from 
-\vC is in terms of a library
-of VHDL design units which has been developed
-as part of the AhirV2  effort.  This library
-consists of control-flow elements, data-path
-elements and memory elements.
-\item 
-For the C/C++ to \Aa translation, the
-AhirV2 toolset uses the front-end C/C++ compiler
-infrastructure developed by the LLVM project (www.llvm.org), which
-compiles the C/C++ program to LLVM byte-code
-(which is a processor independent representation
-of the compiled program).  The AhirV2 toolset
-has a utility which translates the LLVM byte-code
-to an \Aa description.
-\end{itemize}
-In Section \ref{sec:Example}, 
-we illustrate the use of these tools on the
-simple example shown above.
-
-The earlier version of the AHIR flow has been
-used to convert non-trivial programs to hardware.
-The resulting circuits are upto two orders of
-magnitude more energy-efficient than
-a processor \cite{ref:dsd2010}.  The AhirV2
-toolset incorporates several optimizations which
-make the resulting circuits more 
-competitive.
-
-\section{The tools}
-
-We assume that you have access to either {\bf llvm-gcc}
-or {\bf clang} as the front-end compiler which generates
-LLVM byte-code from C/C++.  The current AhirV2 toolset
-is consistent with llvm 2.8 and clang 2.8.
-
-The other tools in the chain are described below.
-
-\subsection{{\bf llvm2aa}}
-
-This tool takes LLVM byte code and converts it into an
-\Aa file.
+The typical usage of the tool is 
 \begin{verbatim}

### 2011-04-28 v2/llvmbc2Aa/docs/llvm2aa.tex
committing some misc. changes and bug fixes.

@@
 \begin{verbatim}
-llvm2aa [-modules=<listfile>] [-storageinit] bytecode.o > bytecode.aa
+llvm2aa [-modules=<listfile>] [-storageinit] [llvm-passes] bytecode.o > bytecode.aa
 \end{verbatim}
@@
 which should be converted to \Aa.   The names of these functions should be
-listed in the text-file listfile. If absent, all functions
+listed in the text-file listfile. If this option is not specified, all functions
 are converted.
@@
 the \Aa code for this purpose.
+\item {\bf llvmpasses} : {\bf llvm2aa} uses the LLVM compiler
+infrastructure to perform LLVM byte-code optimizations.  A large list
+of these optimizations is available through the llvm2aa command-line.
+For more details, see LLVM documentation at http://www.llvm.org.
 \end{itemize}
 
-\subsection{{\bf AaLinkExtMem}}
+\section{Limitations}
 
-This tool takes a list of \Aa files, elaborates the program,
-does memory space decomposition.  The externally visible memory space is
-linked in one of two ways: either it is assumed to be external
-and all accesses to it are routed out of the \Aa program,
-or it is assumed to be internal and assumed to consist of
-a memory object (an array of bytes).  External pointer dereferences
-are handled as if they are directed at this memory object.
-\begin{verbatim}
-AaLinkExtMem [-I n] [-E obj-name] bytecode.o > bytecode.aa
-\end{verbatim}
-The generated \Aa code is sent to {\bf stdout} and all informational
-messages are sent to {\bf stderr}.  On success, the tool returns 0.
-
-The options:
+Several LLVM byte-code constructs are not supported.  Most 
+importantly:
 \begin{itemize}
-\item {\bf -I n}: specifies that external references to memory
-are to be mapped as if they are to an internal object whose size
-is $n$ bytes.
-\item {\bf -E obj-name} : specifies that the object to which
-external references are mapped is to be named obj-name.
-\end{itemize}
-It is better if you use the {\bf -I} and {\bf -E} options.
-
-If the {\bf -I} option is not used, then all external memory
-references are routed out of the \Aa program through pipes.
-In this case, if the \Aa compiler determines that there is some pointer 
-in the program which can point
-to both internal and external memory, then this will be
-declared as an error!
-
-\subsection{{\bf Aa2VC}}
-
-This tool takes a list of \Aa programs and converts them
-to a \vC description. 
-\begin{verbatim}
-Aa2VC [-O] [-C] [-I obj-name] file1.aa file2.aa ...  > result.vc
-\end{verbatim}
-The generated \vC code is sent to {\bf stdout} and all informational
-messages are sent to {\bf stderr}.  On success the tool returns 0.
-
-The options:
-\begin{itemize}
-\item {\bf -O} : if used, sequential statement blocks are parallelized
-by doing dependency analysis.
-\item {\bf -C} : if used, a C stub is created for every module that
-is not called from within the system.  These stubs can be used to
-interface to a VHDL simuator (or even drive hardware) to verify
-the VHDL code generated by downstream tools.
-\item {\bf -I obj-name} : if specified, all external memory references
-are considered as being directed at the storage object named obj-name.
+\item Function pointers are not supported.
+\item Functions with a variable number of arguments are not 
+supported.
+\item Calls to LLVM intrinsics are just passed through to the
+output \Aa file.  The \Aa file will then contain calls
+to these intrinsics without there being a corresponding
+module declaration in the \Aa file.
+\item If the LLVM byte-code has cycles in its call graph,
+then the code is translated, but will create an error
+in downstream \Aa analysis and transformation tools.
+\item System calls made from the \Aa code are simply
+passed through and would need to be supplied as an
+\Aa library in order to perform downstream analysis and
+transformation.
+\item The LLVM integer, floating-point, array, structure,
+vector and void types are the only ones currently supported.
+\item The LLVM switch instruction is currently not supported,
+but support is planned in the near future.
+\item The LLVM indirect-branch, invoke, unwind and unreachable
+instructions will not be supported.
+\item The LLVM division and remainder instructions are currently
+not supported.  Support for these is planned.
+\item LLVM vector instructions are currently not supported.
+\item LLVM aggregate instructions (extractvalue, insertvalue)
+are currently not supported, but will be supported in the
+near future.
 \end{itemize}
 
-\subsection{\bf vc2vhdl}
+\section{Examples}
 
-Takes a collection of \vC descriptions and converts them to
-VHDL.
-\begin{verbatim}
-vc2vhdl [-O] [-C] -t foo [-t bar -t bar2 ...]\
-         -f file1.vc -f file2.vc ...  > system.vhdl
-\end{verbatim}
-
-The options:
-\begin{itemize}
-\item {\bf -t} : to specify the modules which are to be 
-accessible from the ports of the generated VHDL system.
-Multiple top-level modules can be specified in this way.
-\item {\bf -f} : specifies the \vC files to be analyzed. 
-Multiple \vC files may be specified.
-\item {\bf -O} : optimize the generated VHDL by compacting
-the control-path.  This does not change the resulting
-hardware, but makes the generated VHDL file smaller.
-\item {\bf -C} : the VHDL code has a system test bench which
-interfaces to foreign code using a VHPI/Modelsim-FLI interface.
-If this is not specified, the  generated test bench simply
-instantiates the system and starts all top-level modules
-off (you will need to fill in your own test bench here).
-The C testbench is usually easier to write (it probably
-already exists in the form of the original program).
-\end{itemize}
-
-The tool performs concurrency analysis to determine operations which
-can be mapped to the same physical operator without the need for
-arbitration.  It also instantiates separate memory subsystems for
-the disjoint memory spaces (in practice many of the memory spaces
-are small and are converted to register banks).
-
-\subsection{Miscellaneous: {\bf vcFormat} and {\bf vhdlFormat}}
-
-The outputs produces by {\bf Aa2VC} and {\bf vc2vhdl} are
-not well formatted.  One can format \Aa and \vC files
-using  {\bf vcFormat} as follows
-\begin{verbatim}
-vcFormat < unformatted-vc/aa-file  > formatted-vc/aa-file
-\end{verbatim}
-and similarly use {\bf vhdlFormat} to format generated
-VHDL files.
-
-\section{An example} \label{sec:Example}
-
-Let us revisit the simple example considered in
-the first section:
+Let us start with the following C program, kept
+in file ``add.c''.
 \begin{verbatim}
@@
 \end{verbatim}
-We wish to generate a circuit which {\em implements}
-the specification implied by this program.
 
-We convert the program to LLVM byte code using
-the {\bf clang} compiler (www.llvm.org)
+We will first need to compile this program down to 
+LLVM byte code.  For this, we use the {\bf clang}
+compiler (http://www.clang.org)
 \begin{verbatim}
@@
 \end{verbatim}
-This produces a binary file {\bf add.o} which is
-the LLVM byte-code.  To make the byte-code human
-readable, we dis-assemble it using an LLVM utility
+This produces an LLVM byte-code file {\bf add.o},
+which contains a compiled version of the function
+in the file shown above.
+This is our starting point.
+
+We use 
 \begin{verbatim}
-     llvm-dis add.o
+llvm2aa  -storageinit add.o > add.o.aa 
 \end{verbatim}
-This is what the LLVM assembly code looks like
-\begin{verbatim}
-; ModuleID = 'add.o'
-target datalayout = "e-p ...... "
-target triple = "i386-pc-linux-gnu"
-
-define i32 @add(i32 %a, i32 %b) nounwind {
-  %1 = alloca i32, align 4
-  %2 = alloca i32, align 4
-  %c = alloca i32, align 4
-  store i32 %a, i32* %1, align 4
-  store i32 %b, i32* %2, align 4
-  %3 = load i32* %1, align 4
-  %4 = load i32* %2, align 4
-  %5 = add nsw i32 %3, %4
-  store i32 %5, i32* %c, align 4
-  %6 = load i32* %c, align 4
-  ret i32 %6
-}
-\end{verbatim}
-To get to this point, we could have used several
-optimizations which are available in the LLVM frame-work.
-But we work with the unoptimized version to illustrate
-the storage decomposition which is carried out by
-the AhirV2 tools.
-
-The LLVM byte-code is our starting point.  We first convert it
-to \Aa.
-\begin{verbatim}
-llvm2aa add.o | vcFormat > add.o.aa
-\end{verbatim}
-This produces an \Aa program
+to generate an \Aa version of the LLVM bytecode.
+All functions in the LLVM bytecode will be translated
+and initial values of globally declared objects will
+be ignored.
+The \Aa file that is produced is
 \begin{verbatim}
@@
 $module [add] 
-// arguments
 $in (a : $uint<32> b : $uint<32> )
@@
   $storage stored_ret_val__ : $uint<32>
-  $branchblock [add]
+  $branchblock [body] 
   {
     //begin: basic-block bb_0
-    $storage iNsTr_0 : $uint<32>
-    $storage iNsTr_1 : $uint<32>
-    $storage c : $uint<32>
-    iNsTr_0 := a
-    iNsTr_1 := b
+    $storage iNsTr_0_alloc : $uint<32>
+    $storage iNsTr_1_alloc : $uint<32>
+    $storage c_alloc : $uint<32>
+    iNsTr_0 := @(iNsTr_0_alloc)
+    iNsTr_1 := @(iNsTr_1_alloc)
+    c := @(c_alloc)
+    ->(iNsTr_0) := a
+    ->(iNsTr_1) := b
     // load 
-    iNsTr_4 := iNsTr_0
+    iNsTr_4 := ->(iNsTr_0) 
     // load 
-    iNsTr_5 := iNsTr_1
+    iNsTr_5 := ->(iNsTr_1) 
     iNsTr_6 := (iNsTr_4 + iNsTr_5)
-    c := iNsTr_6
+    ->(c) := iNsTr_6
     // load 
-    iNsTr_8 := c
+    iNsTr_8 := ->(c) 
     stored_ret_val__ := iNsTr_8
@@
 
-Now, this \Aa code is converted to a virtual circuit \vC representation.
-\begin{verbatim}
-     Aa2VC -O add.o.aa | vcFormat > add.o.aa.vc
-\end{verbatim}
-The virtual circuit representation is a bit too verbose to reproduce entirely
-here, but we show some critical fragments
-\begin{verbatim}
-$module [add] 
-{
-  $in a:$int<32> b:$int<32>
-  $out ret_val__:$int<32>
-  $memoryspace [memory_space_0] 
-  {
-    $capacity 1
-    $datawidth 32
-    $addrwidth 1
-    // ret-val is kept here
-    $object [xxaddxxstored_ret_val__] : $int<32>
-  }
-  $memoryspace [memory_space_1] 
-  {
-    $capacity 1
-    $datawidth 32
-    $addrwidth 1
-    // a is kept here.
-    $object [xxaddxxaddxxiNsTr_0] : $int<32>
-  }
-  $memoryspace [memory_space_2] 
-  {
-    $capacity 1
-    $datawidth 32
-    $addrwidth 1
-    // b is kept her
-    $object [xxaddxxaddxxiNsTr_1] : $int<32>
-  }
-  $memoryspace [memory_space_3] 
-  {
-    $capacity 1
-    $datawidth 32
-    $addrwidth 1
-    // c is kept here.
-    $object [xxaddxxaddxxc] : $int<32>
-  }
-  $CP 
-  {
-     // a control-flow petri-net..  verbose..
-  }
-  // end control-path
-  $DP 
-  {
-     // wires and operators.
-  }
-
-   // links between CP and DP
-}
-\end{verbatim}
-The important points to note are that the stored objects a,b,c and
-ret\_val\_\_ are mapped to different memory spaces.  Thus, the
-chief difference between a \vC description and a processor is 
-that the \vC program partitions storage into small units which
-are accessed only by operators that need them.
-
-Finally, we take the \vC description and convert it to
-VHDL
-\begin{verbatim}
-vc2vhdl -t add -f add.o.aa.vc | vhdlFormat > system.vhdl
-\end{verbatim}
-This produces a VHDL implementation of the system with
-{\bf add} marked as a top-level module.  The VHDL that
-is produced is too voluminous to reproduce here, but
-the top-level system entity is
-\begin{verbatim}
-entity test_system is  -- system
-  port (--
-    add_a : in  std_logic_vector(31 downto 0);
-    add_b : in  std_logic_vector(31 downto 0);
-    add_ret_val_x_x : out  std_logic_vector(31 downto 0);
-    add_tag_in: in std_logic_vector(0 downto 0);
-    add_tag_out: out std_logic_vector(0 downto 0);
-    add_start : in std_logic;
-    add_fin   : out std_logic;
-    clk : in std_logic;
-    reset : in std_logic); --
-  --
-end entity;
-\end{verbatim}
-There are ports corresponding to the arguments of the top-level module,
-and the add\_start/add\_fin is a handshake pair.  One sets of the inputs,
-starts the system and waits until the fin is asserted.  After the fin
-is asserted, one has the return value of at the appropriate port.
-
-\section{External and internal memory spaces}
-
-Consider the following C program:
-\begin{verbatim}
-int main(int* b)
-{
-   int q[2];
-   q[0] = *b;
-   q[1] = q[0];
-   return(q[1]);
-}
-\end{verbatim}
-When this program is mapped to a circuit, we identify 
-two distinct memory spaces, one which contains the
-array $q$ and the other corresponding to the 
-external world (the one referred to by the pointer $b$).
-Where is the external memory physically located?  In 
-the AhirV2 flow, we can either locate it outside
-the system or inside the system which is being 
-described by this program.
-
-If the external memory is to be placed outside, then
-accesses to it from within the system must be routed
-outside the system.  On the other hand if it is
-to be placed inside, a storage object corresponding 
-to it must be created and all accesses to the external
-memory must be directed at this storage object.
-Further, the external world must have a mechanism for
-accessing this storage object.
-
-Both options are supported in the AhirV2 flow through
-the utility {\bf AaLinkExtMem} described earlier.
-
-\subsection{Keeping the external memory outside the system}
-
-For this example, if you want to keep the external memory
-outside, you will have to go through the following sequence
-\begin{verbatim}
-# first use clang (or llvm-gcc) to generate llvm-byte-code
-clang -std=gnu89 -emit-llvm -c foo.c
-#
-# disassemble so that you can make sense of the llvm bc.
-llvm-dis foo.o
-#
-# OK, now take the llvm byte code
-# and generate an Aa description.
-# use the storageinit option to initialize
-# global storage.
-# (the pipe to vcFormat is to prettify the output)
- llvm2aa -storageinit foo.o | vcFormat > foo.o.aa
-#
-#
-# Do an Aa -> Aa transformation: map external
-# memory outside..
-AaLinkExtMem foo.o.aa | vcFormat > foo.o.memlinked.ExternalOutside.aa
-#
-# Now take the Aa code and generate a virtual
-# circuit..
-# the -O flag does dependency analysis in straight-line
-# code and parallelizes it.
-#
-Aa2VC -O foo.o.memlinked.ExternalOutside.aa | vcFormat\
-                 > foo.o.memlinked.ExternalOutside.aa.vc
-#
-# finally, generate vhdl from the vc description.  Note that
-# you will have to mark the module foo as well as the
-# extmem_store_32/load_32 modules as top-level modules
-# so that it is possible for the outside world to serve
-# requests made from inside.
-#
-vc2vhdl -O -t foo -t extmem_store_32 -t extmem_load_32\ 
-         -f foo.o.memlinked.ExternalOutside.aa.vc | vhdlFormat\
-              > foo_o_aa_memlinked_external_outside_vc.vhdl
-
-\end{verbatim}
-
-If you look at the generate top-level VHDL entity, its ports
-will be
-\begin{verbatim}
-entity test_system is  -- system
-  port (--
-    foo_b : in  std_logic_vector(31 downto 0);
-    foo_ret_val_x_x : out  std_logic_vector(31 downto 0);
-    foo_tag_in: in std_logic_vector(0 downto 0);
-    foo_tag_out: out std_logic_vector(0 downto 0);
-    foo_start : in std_logic;
-    foo_fin   : out std_logic;
-    clk : in std_logic;
-    reset : in std_logic;
-    extmem_read_address_32_pipe_read_data: out std_logic_vector(31 downto 0);
-    extmem_read_address_32_pipe_read_req : in std_logic_vector(0 downto 0);
-    extmem_read_address_32_pipe_read_ack : out std_logic_vector(0 downto 0);
-    extmem_read_data_32_pipe_write_data: in std_logic_vector(31 downto 0);
-    extmem_read_data_32_pipe_write_req : in std_logic_vector(0 downto 0);
-    extmem_read_data_32_pipe_write_ack : out std_logic_vector(0 downto 0);
-    extmem_write_address_32_pipe_read_data: out std_logic_vector(31 downto 0);
-    extmem_write_address_32_pipe_read_req : in std_logic_vector(0 downto 0);
-    extmem_write_address_32_pipe_read_ack : out std_logic_vector(0 downto 0);
-    extmem_write_data_32_pipe_read_data: out std_logic_vector(31 downto 0);
-    extmem_write_data_32_pipe_read_req : in std_logic_vector(0 downto 0);
-    extmem_write_data_32_pipe_read_ack : out std_logic_vector(0 downto 0)); --
-  --
-end entity;
-\end{verbatim}
-The external memory read and write address and data are clearly
-visible.   The outside world is responsible for serving the
-read/write requests made from the inside.
-
-\subsection{Keeping the external memory inside the system}
-
-For this example, if you want to keep the external memory
-outside, you will have to go through the following sequence
-\begin{verbatim}
-# use clang (or llvm-gcc) to generate llvm-byte-code
-clang -std=gnu89 -emit-llvm -c foo.c
-#
-# disassemble so that you can make sense of the llvm bc.
-llvm-dis foo.o
-#
-# OK, now take the llvm byte code
-# and generate an Aa description.
-# use the storageinit option to initialize
-# global storage.
-# (the pipe to vcFormat is to prettify the output)
- llvm2aa -storageinit foo.o | vcFormat > foo.o.aa
-#
-#
-# Do an Aa -> Aa transformation: map external
-# memory to a storage area inside the system...
-# -I 1024 says that the amount of memory that will be
-# referred to is 1024 bytes.
-# -E mempool says that the storage object corresponding
-# to external memory is named mempool.
-AaLinkExtMem  -I 1024 -E mempool foo.o.aa | vcFormat\
-               > foo.o.memlinked.ExternalInside.aa
-#
-# Now take the Aa code and generate a virtual
-# circuit..
-# the -O flag does dependency analysis in straight-line
-# code and parallelizes it.
-# the -I mempool option says that external memory is
-# to be mapped inside the system to object mempool..
-#
-Aa2VC -O -I mempool foo.o.memlinked.ExternalInside.aa\
-           | vcFormat > foo.o.memlinked.ExternalInside.aa.vc
-#
-# finally, generate vhdl from the vc description.
-# note that you will have to mark mem_load__ and mem_store__
-# as top-level modules, so that the external world can
-# access its memory pool inside the system.
-#
-vc2vhdl -O -t foo -t mem_load__ -t mem_store__ \ 
-     -f foo.o.memlinked.ExternalInside.aa.vc\
-       | vhdlFormat > foo_o_aa_memlinked_external_inside_vc.vhdl
-\end{verbatim}
-
-The generated top-level VHDL entity has the following
-ports:
-\begin{verbatim}
-entity test_system is  -- system
-  port (--
-    foo_b : in  std_logic_vector(31 downto 0);
-    foo_ret_val_x_x : out  std_logic_vector(31 downto 0);
-    foo_tag_in: in std_logic_vector(0 downto 0);
-    foo_tag_out: out std_logic_vector(0 downto 0);
-    foo_start : in std_logic;
-    foo_fin   : out std_logic;
-    mem_load_x_x_address : in  std_logic_vector(31 downto 0);
-    mem_load_x_x_data : out  std_logic_vector(7 downto 0);
-    mem_load_x_x_tag_in: in std_logic_vector(0 downto 0);
-    mem_load_x_x_tag_out: out std_logic_vector(0 downto 0);
-    mem_load_x_x_start : in std_logic;
-    mem_load_x_x_fin   : out std_logic;
-    mem_store_x_x_address : in  std_logic_vector(31 downto 0);
-    mem_store_x_x_data : in  std_logic_vector(7 downto 0);
-    mem_store_x_x_tag_in: in std_logic_vector(0 downto 0);
-    mem_store_x_x_tag_out: out std_logic_vector(0 downto 0);
-    mem_store_x_x_start : in std_logic;
-    mem_store_x_x_fin   : out std_logic;
-    clk : in std_logic;
-    reset : in std_logic); --
-  --
-end entity;
-\end{verbatim}
-
-The system provides a memory load and memory store
-port to the external world (through the mem\_load.. and mem\_store..)
-ports.
-
-\section{Handling complex programs}
-
-This example is trivial.  Real, non-trivial programs can be
-mapped in this manner.  Currently, there are only
-two restrictions
-\begin{itemize}
-\item No recursion, no cycles in the call-graph of the original
-program.
-\item No function pointers.
-\end{itemize}
-
-The AhirV2 flow produces a modular system, and by default,
-produces one VHDL entity for every function in the original
-program.  Further, the concept of message pipes is native
-to the \Aa and \vC descriptions.  Thus, it is easy to
-map parallel programs or programs consisting of multiple
-concurrent processes to hardware.
-
-\section{To be continued}
-
-This document is a work-in-progress.  More details and
-related documentation will be added shortly.
-
-The examples folder contains some explanations which might
-make things easier to understand.
-
-\begin{thebibliography}{99}
-\bibitem{ref:dsd2010}
-Sameer D. Sahasrabuddhe, Sreenivas Subramanian, Kunal P. Ghosh, Kavi Arya, Madhav P. Desai, 
-"A C-to-RTL Flow as an Energy Efficient Alternative to Embedded Processors in Digital Systems," 
-DSD, pp.147-154, 2010 13th Euromicro Conference on Digital 
-System Design: Architectures, Methods and Tools, 2010
-\end{thebibliography}
 \end{document}

### 2011-04-29 v2/CtestBench/docs/CtestBench.tex
added the Ctestbench stuff to v2/

\documentclass{article}

\title{On verifying AhirV2 generated VHDL using software testbenches}
\author{Madhav Desai \\ Department of Electrical Engineering \\ Indian Institute of Technology \\
	Mumbai 400076 India}

\newcommand{\Aa}{{\bf Aa}~}
\newcommand{\vC}{{\bf vC}~}

\begin{document}
\maketitle

The AhirV2 tool chain can be used to convert parts of a C program to VHDL
(essentially, some of the functions in a program are mapped to VHDL).
To verify the resulting VHDL, one would like to simulate it in a
VHDL simulator (such as Modelsim from Mentor Graphics).  The most
natural way to do this is to use the original program itself
as a testbench for this purpose.

\begin{itemize}
\item Stubs are created for the set of functions which are mapped to  
VHDL by the AhirV2 flow.
\item The software testbench is compiled and linked with these stubs.
\item Whenever a stub function is called, it tries to connect with
a server created by the VHDL simulation process.
\item The VHDL simulation process listens for calls from the stubs
and exchanges data between the stubs and the actual VHDL being simulated.
\end{itemize}

We assume that you have a {\bf Release} directory of the AhirV2 flow
installed.   The Release/bin subdirectory contains the binaries for
the translation of LLVM-byte-code down to a VHDL description. 
The Release/vhdl directory contains VHDL library elements that 
are instantiated in this VHDL description.  The Release/CtestBench 
directory contains the libraries and headers that you will need.  
Please see Release/CtestBench/examples/helloworld for an 
illustration of the whole flow.
\end{document}

### 2011-04-29 v2/CtestBench/examples/quicksort/vhdl/compile_testbench.sh

@@ -1,0 +1,3 @@
+gcc -g -c vhdlCStubs.c -I./ -I../../../include 
+gcc -g -c testbench.c -I./ -I../../../include
+gcc -g -o testbench testbench.o vhdlCStubs.o  -L../../../lib -lSocketLib -lpthread

### 2011-05-21 ericsson/docs/plan.tex
added write_uint32_n/read_uint32_n methods to Ctestbench stuff. also allow branch regions to have looping which prevents exit (simple infinite loops), now a warning is generated. also started process of fitting to NetFPGA

% Created 2010-08-17 Tue 13:49
\documentclass{article}
\usepackage{graphicx}
\usepackage{times}
\usepackage[margin=0.75in]{geometry}

\title{Inserting an AHIR module in the NetFPGA framework}
\author{Madhav Desai}

\begin{document}

\maketitle

\section{Overview}

This document describes the mechanism by which an 
AHIR generated VHDL system can be included in the NetFPGA packet
processing implementation framework.

The AHIR generated VHDL system is required to fit into 
a NetFPGA user-module template (described below).  Thus,
the AHIR generated VHDL system must be wrapped into
a form expected by the NetFPGA user-module template.

\section{Mechanism}

The user module template has the form (the original
NetFPGA module is in Verilog; we use a VHDL equivalent).
\begin{verbatim}
entity module_template is
   generic (
      DATA_WIDTH: integer := 64;
      CTR_WIDTH: integer := 8;
      UDP_REG_SRC_WIDTH: integer := 2
   );
   port (
      in_data: in std_logic_vector(DATA_WIDTH-1 downto 0);
      in_ctrl: in std_logic_vector(CTRL_WIDTH-1 downto 0);
      in_wr: in std_logic;
      in_rdy: out std_logic;

      out_data: out std_logic_vector(DATA_WIDTH-1 downto 0);
      out_ctrl: out std_logic_vector(CTRL_WIDTH-1 downto 0);
      out_wr: out std_logic;
      out_rdy: in std_logic;

      clk: in std_logic;
      reset: in std_logic);
end entity;
\end{verbatim}
Thus, the NetFPGA user-module template can be viewed
as a system with a single input queue and a single output queue.

The AHIR system has form:
\begin{verbatim}
entity ahir_system is
   port(
          -- for each top-level function 
          --   (in this case int foo(int a))
          foo_a : in std_logic_vector(31 downto 0); -- argument a
          foo_ret_val_x_X: out std_logic_vector(31 downto 0); -- ret value
          foo_start: in std_logic;  -- pulse to start.
          foo_fin: out std_logic;   -- pulsed to indicate finish.

          -- for each input pipe
          --   (in this case inpipe of type int)
          inpipe_pipe_write_data: in std_logic_vector(31 downto 0); -- data
          inpipe_pipe_write_req: in std_logic_vector(0 downto 0); -- ready-in
          inpipe_pipe_write_ack: out std_logic_vector(0 downto 0); -- ready-ack.

          -- for each output pipe
          --   (in this case outpipe of type int)
          outpipe_pipe_read_data: out std_logic_vector(31 downto 0); -- data
          outpipe_pipe_read_req: in std_logic_vector(0 downto 0); -- ready-in
          outpipe_pipe_read_ack: out std_logic_vector(0 downto 0); -- ready-ack.

          clk: in std_logic;
          reset: in std_logic;
       );
end entity;
\end{verbatim}
An AHIR system can thus be viewed as having
\begin{itemize}
\item An input queue for each top-level function.
\item An output queue for each top-level function.
\item An input queue for each input pipe.
\item An output queue for each output pipe.
\end{itemize}

Thus, in order to insert an AHIR system into a NetFPGA user-module template,
all we need to do is
\begin{itemize}
\item Include an input-module which will collect data from the NetFPGA user-module
input data interface.  This input-module will feed the AHIR system input queues.
\begin{itemize}
\item This is especially simple if the AHIR system consists of a single top-level
function without pipes, since there is only one possible destination for the
incoming data.
\end{itemize}
\item Include an output-module which will collect data from the AHIR system
output queues
and forward it out of the output data interface of the NetFPGA user-module.
\begin{itemize}
\item Especially easy if there is only one top-level function in the AHIR
system without output pipes.
\end{itemize}
\item  In the general case (many top-level AHIR functions, with input and output
pipes), we will have to implement some kind of distributor/concentrator to
stream the data consistently between the NetFPGA user-module
input/output queues and the AHIR system input/output queues.
\end{itemize}

\section{Looking Ahead: a proposal for rapid deployment}

The process of including the AHIR system in a NetFPGA user template is
exactly the same as that of verifying an AHIR system using a C-testbench
with the assumption that the testbench only sends/receives packets to the AHIR 
system.  Most of the basic infrastructure for this is in place.  The only
necessary components are the input and output modules outlined in the 
previous section.

There could be three approaches:
\begin{itemize}
\item The input and output modules are written as VHDL modules with
adequate parametrization.  This should work if the AHIR system is
simple enough (consists of a system with a single top-level function
without any input/output pipes, for example).  However, writing
VHDL (or Verilog) implementations is a pain for more complex scenarios.
\item The input and output modules are written as C functions and
are included into the AHIR system.  This gives much more flexibility.
However, in a complex AHIR system, this approach will limit the
ability to exploit parallelism (note: we cannot use threads in these
C functions because AhirV2 tools cannot handle threads yet).
\item The input and output modules are written as Aa modules.  This
gives not only flexibility, but also the ability to use the
parallelism in the system.
\end{itemize}

Originally, Sameer had implemented the input and output module for
a specific scenario (as documented in the previous plan).  In this
scenario, the input module would receive the data from the NetFPGA
surroundings, would put the data packet into memory and would trigger
the AHIR module.  On completion of the AHIR module, the output module
would take the modified data from memory and would forward it to
the NetFPGA environment.   I have looked at the code and it
seems reasonably clean (will probably have to redo parts of it
to fit into the new AhirV2 scheme).

What would be easiest, in my opinion would be
\begin{itemize}
\item Use the parts of Sameer's code that do the protocol
translation (between the NetFPGA user-module and an AHIR module).
\item Write the input and output modules in Aa.  For the moment,
we would do so for a specific scenario (much faster to develop and
easier to maintain).  This would remove the need to do memory 
management from the VHDL interface code (as is done in Sameer's
current implementation, in which the VHDL input module directly
accesses AHIR memory).
\item Generate the final AHIR system by including the input/output modules
written in Aa.
\item Instantiate the final AHIR system in a NetFPGA wrapper which 
just does the protocol translation between NetFPGA and AHIR.
\end{itemize}
This is indicated in Figure \ref{fig:Wrap}.   The only
additional effort is to write the input and output modules
and to write the wrap functionality in vc2vhdl.  This should
take a week.

\begin{figure}[!t]
  \centering
  \includegraphics[scale=0.7]{NetFPGAWrap.eps}
  \caption{NetFPGA Wrap Flow}
  \label{fig:Wrap}
\end{figure}

\end{document}

### 2011-05-21 v2/CtestBench/include/SocketLib.h

@@ -98,7 +98,7 @@ int receive_string(int n, char* buffer);
 // the buffer is used for the sent as well
 // as the received data.
 //
-void  send_packet_and_wait_for_response(char* buffer, char* server_host_name, int server_port_number);
+void  send_packet_and_wait_for_response(char* buffer, int send_length, char* server_host_name, int server_port_number);
 
 
 

### 2011-05-21 v2/iolib/fifo.c

@@ -220,3 +220,84 @@ void write_uint8(char *id, uint8_t data)
   fclose(F);
   wait_for_ack(id);
 }
+
+
+
+// burst read/write..
+void read_uint64_n(const char *id, uint64_t* buf, int buf_len)
+{
+  int i;
+  for(i = 0; i < buf_len; i++)
+    {
+      buf[i] = read_uint64((char*) id);
+    }
+}
+void write_uint64_n(const char *id, uint64_t* buf, int buf_len)
+{
+  int i;
+  for(i = 0; i < buf_len; i++)
+    {
+      write_uint64((char*) id,buf[i]);
+    }
+}
+
+
+
+void read_uint32_n(const char *id, uint32_t* buf, int buf_len)
+{
+  int i;
+  for(i = 0; i < buf_len; i++)
+    {
+      buf[i] = read_uint32((char*) id);
+    }
+}
+void write_uint32_n(const char *id, uint32_t* buf, int buf_len)
+{
+  int i;
+  for(i = 0; i < buf_len; i++)
+    {
+      write_uint32((char*) id,buf[i]);
+    }
+}
+
+
+void read_uint16_n(const char *id, uint16_t* buf, int buf_len)
+{
+  int i;
+  for(i = 0; i < buf_len; i++)
+    {
+      buf[i] = read_uint16((char*) id);
+    }
+}
+void write_uint16_n(const char *id, uint16_t* buf, int buf_len)
+{
+  int i;
+  for(i = 0; i < buf_len; i++)
+    {
+      write_uint16((char*) id,buf[i]);
+    }
+}
+
+void read_uint8_n(const char *id, uint8_t* buf, int buf_len)
+{
+  int i;
+  for(i = 0; i < buf_len; i++)
+    {
+      buf[i] = read_uint8((char*) id);
+    }
+}
+
+void write_uint8_n(const char *id, uint8_t* buf, int buf_len)
+{
+  int i;
+  for(i = 0; i < buf_len; i++)
+    {
+      write_uint8((char*) id,buf[i]);
+    }
+}
+
+
+
+
+
+

### 2011-05-25 Documents/AhirV2/AhirV2.tex
updated the documentation..

@@
 \end{itemize}
-It is better if you use the {\bf -I} and {\bf -E} options.
+It is recommended that you use the {\bf -I} and {\bf -E} options.
 
@@
 are considered as being directed at the storage object named obj-name.
+If not specified, then the tool will throw an error if it finds
+a pointer dereference that cannot be resolved as pointing only to
+storage objects declared inside the \Aa program.
 \end{itemize}
@@
 Takes a collection of \vC descriptions and converts them to
-VHDL.
+an AHIR system described in VHDL.
 \begin{verbatim}
-vc2vhdl [-O] [-C] -t foo [-t bar -t bar2 ...]\
+vc2vhdl [-O] [-C] -t/-T foo [-t/-T bar -t/-T bar2 ...]\
          -f file1.vc -f file2.vc ...  > system.vhdl
@@
 accessible from the ports of the generated VHDL system.
+Such modules have to be top-level (that is, they cannot
+be called from within the program).
 Multiple top-level modules can be specified in this way.
+The control and argument ports for these modules are
+visible at the interface of the generated AHIR system.
+\item {\bf -T} : to specify top-level modules which are to be 
+free-running inside the AHIR system.
+Multiple top-level modules can be specified in this way.
+Such modules do not have any arguments and do not return
+any values.  Their only mechanism of communication with
+the world outside the AHIR system is through pipes.
+The control ports for these modules are
+{\bf not visible} at the interface of the generated AHIR system.
+In the AHIR system, these modules are started on reset
+and are run forever (restarted after they finish, forever).
 \item {\bf -f} : specifies the \vC files to be analyzed. 
-Multiple \vC files may be specified.
+Multiple \vC files may be specified.  An object must be
+defined before it is used, so the \vC files must be 
+specified in the correct order.
 \item {\bf -O} : optimize the generated VHDL by compacting
@@
 already exists in the form of the original program).
+\item {\bf -s ghdl/modelsim} :  If {\bf ghdl} is specified
+with the -s option, then the generated testbench (if -C is specified)
+uses the VHPI interface to link with foreign code.  Otherwise,
+the generated testbench (if -C is specified) uses the Modelsim FLI
+interface to link with foreign code.
 \end{itemize}
@@
 starts the system and waits until the fin is asserted.  After the fin
-is asserted, one has the return value of at the appropriate port.
+is asserted, one has the return value of {\bf add} at the appropriate port.
 
@@
 Both options are supported in the AhirV2 flow through
-the utility {\bf AaLinkExtMem} described earlier.
+{\bf AaLinkExtMem}.
 
@@
 For this example, if you want to keep the external memory
-outside, you will have to go through the following sequence
+inside, you will have to go through the following sequence
 \begin{verbatim}
@@
 
-The examples folder contains some explanations which might
-make things easier to understand.
-
 \begin{thebibliography}{99}

### 2011-05-26 v2/Ahir-asm/devel/docs/AaLRM.tex
getting the documentation in order..  also updated release area with latest binaries and documentation.

@@
              branch-block-specifier block-name
-             { sequence-of-statements}
+             { branch-block-statement-sequence }
 \end{verbatim}
@@
          switch-spec switch-expression
-           ( expr-value  serial-statement-sequence )*
-           ( default serial-statement-sequence )?
-
+           ( expr-value  branch-block-statement-sequence )*
+           ( default branch-block-statement-sequence )?
+         end-switch-spec
 \end{verbatim}
@@
        if-spec test-expression 
-           serial-statement-sequence
+           branch-block-statement-sequence
        (else
-           serial-statement-sequence)?
+           branch-block-statement-sequence)?
+       end-if-spec
 

### 2011-05-26 v2/libAhirV2/util/vc2vhdl.cpp

@@ -5,11 +5,8 @@
 #include <vcLexer.hpp>
 
 using namespace std;
-struct option long_options[] = {
-    {"relaxed-component-visibility", 0, 0, 0},
-    {"depend", required_argument, 0, 0},
-    {0, 0, 0, 0}
-};
+
+
 
 void Handle_Segfault(int signal)
 {
@@ -19,22 +16,43 @@ void Handle_Segfault(int signal)
 
 void Usage_Vc2VHDL()
 {
+
+  cerr << "brief description: produces unformatted VHDL from specified VC files,  " << endl;
   cerr << "Usage: " << endl;
-  cerr << "vc2vhdl [-O] [-C] [-s <ghdl/modelsim>] -f <vc-file> [-f <vc-file>...] -t/-T <top-module> [-t/-T <top-module>...] " << endl << endl;
-  cerr << "specify vc-files using -f, top-modules in system using -t.. for example" << endl
+  cerr << "vc2vhdl ... options ... " << endl;
+  cerr << " options " << endl;
+  cerr <<  " -t mod-name: will make mod-name a top-level module which must" << endl
+       <<  "              be started from outside the AHIR system.  Such a module can " << endl
+       <<  "              have input/output arguments from/to the outside world" << endl
+       <<  " Notes: " << endl
+       <<  "   1. alternate form: --top mod-name" << endl
+       <<  "   2. option can be repeated to specify multiple top-level modules" << endl;
+  cerr << endl;  
+  cerr <<  " -T mod-name: will make mod-name a free-running top-level module " << endl
+       <<  "              which need not be started from outside.  Such a module cannot " << endl
+       <<  "              have input/output arguments from/to the outside world" << endl
+       <<  " Notes: " << endl
+       <<  "   1. alternate form: --free_running_top mod-name" << endl
+       <<  "   2. option can be repeated to specify multiple top-level modules." << endl;
+  cerr << endl;
+  cerr <<  " -O: will lead to smaller VHDL files due to some compaction." << endl
+       <<  " -C: the generated system testbench connects to foreign language testbench (Ctestbench). " << endl;
+  cerr <<  " -s ghdl/modelsim: specify the simulator with which one is going to simulate the system." << endl
+       <<  "    -s ghdl: will produce a testbench which can be used in GHDL. " << endl
+       <<  "    -s modelsim: will produce a testbench which can be used in Modelsim. " << endl
+       <<  " Notes: " << endl
[…]

### 2011-05-29 ericsson/docs/plan.tex
documentation about putting AHIR system into NetFPGA.

@@
 The AHIR generated VHDL system is required to fit into 
-a NetFPGA user-module template (described below).  Thus,
-the AHIR generated VHDL system must be wrapped into
-a form expected by the NetFPGA user-module template.
+a NetFPGA user-module template. It is relatively easy to
+achieve this.
 
@@
 
-The user module template has the form (the original
-NetFPGA module is in Verilog; we use a VHDL equivalent).
-\begin{verbatim}
-entity module_template is
-   generic (
-      DATA_WIDTH: integer := 64;
-      CTR_WIDTH: integer := 8;
-      UDP_REG_SRC_WIDTH: integer := 2
-   );
-   port (
-      in_data: in std_logic_vector(DATA_WIDTH-1 downto 0);
-      in_ctrl: in std_logic_vector(CTRL_WIDTH-1 downto 0);
-      in_wr: in std_logic;
-      in_rdy: out std_logic;
+The user module template has the form shown in Figure \ref{fig:NetFPGAUserModule}.
+\begin{figure}
+  \centering
+  \includegraphics[scale=0.7]{NetFPGAUserModule.eps}
+  \caption{NetFPGA User Module}
+  \label{fig:NetFPGAUserModule}
+\end{figure}
+The interface is very simple: an input data and control bus,
+and an output data and control bus.
 
-      out_data: out std_logic_vector(DATA_WIDTH-1 downto 0);
-      out_ctrl: out std_logic_vector(CTRL_WIDTH-1 downto 0);
-      out_wr: out std_logic;
-      out_rdy: in std_logic;
+A generic AHIR system can have the form:
+\begin{figure}
+  \centering
+  \includegraphics[scale=0.7]{GenericAhirSystem.eps}
+  \caption{Generic AHIR system}
+  \label{fig:GenericAhirSystem}
+\end{figure}
 
-      clk: in std_logic;
-      reset: in std_logic);
-end entity;
-\end{verbatim}
-Thus, the NetFPGA user-module template can be viewed
-as a system with a single input queue and a single output queue.
-
-The AHIR system has form:
-\begin{verbatim}
-entity ahir_system is
-   port(
-          -- for each top-level function 
-          --   (in this case int foo(int a))
-          foo_a : in std_logic_vector(31 downto 0); -- argument a
-          foo_ret_val_x_X: out std_logic_vector(31 downto 0); -- ret value
-          foo_start: in std_logic;  -- pulse to start.
-          foo_fin: out std_logic;   -- pulsed to indicate finish.
-
-          -- for each input pipe
-          --   (in this case inpipe of type int)
-          inpipe_pipe_write_data: in std_logic_vector(31 downto 0); -- data
-          inpipe_pipe_write_req: in std_logic_vector(0 downto 0); -- ready-in
-          inpipe_pipe_write_ack: out std_logic_vector(0 downto 0); -- ready-ack.
-
-          -- for each output pipe
-          --   (in this case outpipe of type int)
-          outpipe_pipe_read_data: out std_logic_vector(31 downto 0); -- data
-          outpipe_pipe_read_req: in std_logic_vector(0 downto 0); -- ready-in
-          outpipe_pipe_read_ack: out std_logic_vector(0 downto 0); -- ready-ack.
-
-          clk: in std_logic;
-          reset: in std_logic;
-       );
-end entity;
-\end{verbatim}
-An AHIR system can thus be viewed as having
+Thus, in order to insert a generic AHIR system into a NetFPGA user-module template:
 \begin{itemize}
-\item An input queue for each top-level function.
-\item An output queue for each top-level function.
-\item An input queue for each input pipe.
-\item An output queue for each output pipe.
+\item we ensure that the only interaction of the AHIR system with the outside world is
+through four pipes: two input pipes for the input data and control, and
+two output pipes for the output data and control.
+\item The top-level modules in the AHIR system are free-running.  One solution
+would be to include two top-level free-running modules which are 
+responsible for interactions with the outside world (through the four pipes
+discussed above).
+\begin{itemize}
+\item an input-module which will collect data from the input pipes and
+forward to appropriate pipes/modules in the AHIR system.
+\item an output-module which will collect data from internal pipes/modules
+in the AHIR system and forward it to the the system output pipes.
+\end{itemize}
+\item The AHIR system would be instantiated in a {\em fixed} wrapper
+which would be responsible for the protocol matching between the AHIR
+system interface and the NetFPGA user module interface.
 \end{itemize}
 
-Thus, in order to insert an AHIR system into a NetFPGA user-module template,
-all we need to do is
+\section{Issues: memory management}
+
+One issue that needs to be resolved is that of memory management.
+This relates to two processes:
 \begin{itemize}
-\item Include an input-module which will collect data from the NetFPGA user-module
-input data interface.  This input-module will feed the AHIR system input queues.
-\begin{itemize}
-\item This is especially simple if the AHIR system consists of a single top-level
-function without pipes, since there is only one possible destination for the
-incoming data.
-\end{itemize}
-\item Include an output-module which will collect data from the AHIR system
-output queues
-and forward it out of the output data interface of the NetFPGA user-module.
-\begin{itemize}
-\item Especially easy if there is only one top-level function in the AHIR
-system without output pipes.
-\end{itemize}
-\item  In the general case (many top-level AHIR functions, with input and output
-pipes), we will have to implement some kind of distributor/concentrator to
-stream the data consistently between the NetFPGA user-module
-input/output queues and the AHIR system input/output queues.
+\item Initialization: if the memory subsystem in the AHIR module needs
+to be initialized, then this must be done explicitly by the outside
+world.  The llvm2aa tool can generate memory initialization modules
+that initialize the data structures in the system to their initial
+values as declared in the original C program.  These modules are
+auto-generated and need to be called once after the AHIR system
+is reset (this call-once feature can be incorporated into vc2vhdl).
+\item Exchange of information between the AHIR memory subsystem and
+the outside world:  If this is needed, it would need to be explicitly 
+built into the input/output modules.  Note that an exchange of
+information (in general) may be initiated from ``inside'' or
+from ``outside'' the AHIR system.
 \end{itemize}
@@
 previous section.
-
-There could be three approaches:
-\begin{itemize}
-\item The input and output modules are written as VHDL modules with
-adequate parametrization.  This should work if the AHIR system is
-simple enough (consists of a system with a single top-level function
-without any input/output pipes, for example).  However, writing
-VHDL (or Verilog) implementations is a pain for more complex scenarios.
-\item The input and output modules are written as C functions and
-are included into the AHIR system.  This gives much more flexibility.
-However, in a complex AHIR system, this approach will limit the
-ability to exploit parallelism (note: we cannot use threads in these
-C functions because AhirV2 tools cannot handle threads yet).
-\item The input and output modules are written as Aa modules.  This
-gives not only flexibility, but also the ability to use the
-parallelism in the system.
-\end{itemize}
 
@@
 \begin{itemize}
-\item Use the parts of Sameer's code that do the protocol
+\item use the parts of Sameer's code that do the protocol
 translation (between the NetFPGA user-module and an AHIR module).
-\item Write the input and output modules in Aa.  For the moment,
-we would do so for a specific scenario (much faster to develop and
-easier to maintain).  This would remove the need to do memory 
-management from the VHDL interface code (as is done in Sameer's
-current implementation, in which the VHDL input module directly
-accesses AHIR memory).
-\item Generate the final AHIR system by including the input/output modules
-written in Aa.
-\item Instantiate the final AHIR system in a NetFPGA wrapper which 
-just does the protocol translation between NetFPGA and AHIR.
+\begin{itemize}
+\item we implement a VHDL entity {\em NetFpgaAhirSystemWrap} which 
+would instantiate an AHIR system with four interface pipes (see Figure \ref{fig:NetFpgaAhirSystemWrap}).
+\begin{figure}
+  \centering
+  \includegraphics[scale=0.7]{NetFpgaAhirSystemWrap.eps}
+  \caption{NetFpgaAhirSystemWrap Schematic}
+  \label{fig:NetFpgaAhirSystemWrap}
+\end{figure}
 \end{itemize}
-This is indicated in Figure \ref{fig:Wrap}.   The only
-additional effort is to write the input and output modules
-and to write the wrap functionality in vc2vhdl.  This should
-take a week.
+\item write the input and output modules in C/Aa and add
+to the AHIR flow to generate the final system. 
+\begin{itemize}
+\item this can be user-written (and eventually auto-generated) for
+the specific system being generated.  Note that we will have
+to pay some attention to memory initialization and transfer of
+information between the AHIR memory subsystem and the outside
+world.
+\end{itemize}
+\item generate the final AHIR system by including the input/output modules
+written in C/Aa.
+\item Instantiate the final AHIR system in the NetFpgaAhirSystemWrap wrapper.
+\end{itemize}
+The process is indicated in Figure \ref{fig:Wrap}.   The only
+additional effort is to write the input and output modules (in C or Aa).
 
-\begin{figure}[!t]
+\begin{figure}
   \centering

### 2011-06-07 ericsson/docs/plan.tex
Bug fix: initialize variables in storage objects and also avoid needless printing of objects that were never used.
Committed some documentation about how to fit ahir system into netfpga pipeline.

@@
 
-A generic AHIR system can have the form:
+A generic AHIR system can have the form shown in Figure \ref{fig:GenericAhirSystem}.
 \begin{figure}
@@
 
-Thus, in order to insert a generic AHIR system into a NetFPGA user-module template:
+In order to insert a generic AHIR system into a NetFPGA user-module template:
 \begin{itemize}
@@
 two output pipes for the output data and control.
-\item The top-level modules in the AHIR system are free-running.  One solution
-would be to include two top-level free-running modules which are 
+\item The top-level modules in the AHIR system are free-running.  
+The set of modules (functions) are of two kinds.  The first
+kind consists of functions generated from the click to llvm to AHIR
+flow.  The second kind consists of three fixed free-running modules:
+two free-running modules which are 
 responsible for interactions with the outside world (through the four pipes
-discussed above).
+discussed above) and a third free-running module which manages a packet-queue.
 \begin{itemize}
@@
 in the AHIR system and forward it to the the system output pipes.
+\item a free-queue-manager which provides packet slots to the input-module
+and receives empty packet slots from the output-module.
 \end{itemize}
@@
 \begin{itemize}
-\item Initialization: if the memory subsystem in the AHIR module needs
-to be initialized, then this must be done explicitly by the outside
-world.  The llvm2aa tool can generate memory initialization modules
-that initialize the data structures in the system to their initial
-values as declared in the original C program.  These modules are
-auto-generated and need to be called once after the AHIR system
-is reset (this call-once feature can be incorporated into vc2vhdl).
+\item Initialization: must be explicitly done in 
+the source program.
 \item Exchange of information between the AHIR memory subsystem and
@@
 
-\section{Looking Ahead: a proposal for rapid deployment}
+\section{Implementation}
 
@@
 scenario, the input module would receive the data from the NetFPGA
-surroundings, would put the data packet into memory and would trigger
+surroundings, would put the data packet into a free memory slot and would trigger
 the AHIR module.  On completion of the AHIR module, the output module
 would take the modified data from memory and would forward it to
-the NetFPGA environment.   I have looked at the code and it
-seems reasonably clean (will probably have to redo parts of it
-to fit into the new AhirV2 scheme).
+the NetFPGA environment.   The memory slots would be organized into
+a queue emptied by the input module and filled by the output module.
 
@@
 \begin{itemize}
-\item we implement a VHDL entity {\em NetFpgaAhirSystemWrap} which 
+\item we re-implement a VHDL entity {\em netfpga\_module} which 
 would instantiate an AHIR system with four interface pipes (see Figure \ref{fig:NetFpgaAhirSystemWrap}).
@@
 \end{itemize}
-\item write the input and output modules in C/Aa and add
-to the AHIR flow to generate the final system. 
+\item write the input and output modules and
+a free queue manager in C and add them 
+to the AHIR flow to generate the final AHIR system. 
 \begin{itemize}
-\item this can be user-written (and eventually auto-generated) for
+\item This can be written for
 the specific system being generated.  Note that we will have
@@
 information between the AHIR memory subsystem and the outside
-world.
+world.  I have done a reference implementation based on
+Sameer's VHDL code.  This needs to be reviewed.
 \end{itemize}
 \item generate the final AHIR system by including the input/output modules
-written in C/Aa.
-\item Instantiate the final AHIR system in the NetFpgaAhirSystemWrap wrapper.
+written in C.  This is easiest to do by linking the llvm byte-code for
+the input/output modules with the click2llvm generated byte-code.
+\item Instantiate the final AHIR system in a netfpga pipeline module.
 \end{itemize}
-The process is indicated in Figure \ref{fig:Wrap}.   The only
-additional effort is to write the input and output modules (in C or Aa).
+The process is indicated in Figure \ref{fig:Wrap}.
 
@@
 
+In the folders AHIRGIT/ericsson/C and AHIRGIT/ericsson/vhdl,
+I have put an implementation of the C source (input/output module, free-queue
+manager) and the netfpga-wrapper module.  The C source implements
+Sameer's input module, output module and free queue manager (this should
+be linked with the click2llvm stuff during the AHIR compile chain).
+The vhdl is a very simple wrapper which just does protocol
+translation.
+
 \end{document}

### 2011-07-13 Documents/AhirV2/AhirV2.tex
Fixed a problem in memory coalescing.  The solution is more refined, but needs more memory (and time).  Two storage objects are merged only in the following cases   - a pointer calculation expression which can point     to either of the two objects.   - a pointer dereference expression in which the pointer     can point to either of the two objects. This seems to have a big impact on performance, because memories are getting split into smaller units, hence more parallelism.
Bug fix: if interface object is target/source of statements, then dependency analysis was being incorrectly handled.

@@
 
-\title{AhirV2: from algorithms to hardware \\ A tutorial introduction}
+\title{AhirV2: from algorithms to hardware \\ An overview }
 \author{Madhav Desai \\ Department of Electrical Engineering \\ Indian Institute of Technology \\
@@
 toolset incorporates several optimizations which
-make the resulting circuits more 
-competitive.
+make the resulting circuits smaller and 
+faster than before.
 
@@
 \begin{verbatim}
-vc2vhdl [-O] [-C] -t/-T foo [-t/-T bar -t/-T bar2 ...]\
+vc2vhdl [-O] [-C] [-a] [-e <entity-name] [-w]\
+         -t/-T foo [-t/-T bar -t/-T bar2 ...]\
          -f file1.vc -f file2.vc ...  > system.vhdl
@@
 already exists in the form of the original program).
+\item {\bf -a} : try to minimize the area of the resulting
+VHDL by sharing operators to the maximum extent possible
+(allowing potential contention for resources).  This will
+result in a slower (usually by 2X) system, but will
+also reduce the area (usually by 0.5X).  If not specified,
+two operations will be mapped to the same
+operator  only if it can be proved that they cannot be active simultaneously.
+\item {\bf -e top-entity-name} : The generated top-level VHDL entity
+corresponding to the AHIR system is named top-entity-name.  The default
+is test\_system.
+\item {\bf -w} :  If specified, the VHDL system and test-bench are
+generated as separate unformatted VHDL files.  You will need to
+format these using the vhdlFormat command.
 \item {\bf -s ghdl/modelsim} :  If {\bf ghdl} is specified
@@
 
-The outputs produces by {\bf Aa2VC} and {\bf vc2vhdl} are
+The outputs produced by {\bf Aa2VC} and {\bf vc2vhdl} are
 not well formatted.  One can format \Aa and \vC files
@@
 There are ports corresponding to the arguments of the top-level module,
-and the add\_start/add\_fin is a handshake pair.  One sets of the inputs,
+and the add\_start/add\_fin is a handshake pair.  One sets up the inputs,
 starts the system and waits until the fin is asserted.  After the fin
@@
 
-For this example, if you want to keep the external memory
+In this scenario, all memory accesses which are resolved
+to be to a storage object which is not declared in the
+\Aa program are redirected outside the system by using
+pipes.
+
+If you want to keep the external memory
 outside, you will have to go through the following sequence
@@
 
-If you look at the generate top-level VHDL entity, its ports
+If you look at the generated top-level VHDL entity, its ports
 will be
@@
 
-For this example, if you want to keep the external memory
-inside, you will have to go through the following sequence
+In this scenario, we will assume that all accesses to 
+storage variables not defined in the \Aa program are 
+to be directed to a storage variable which is to be
+a viewed as a shared memory pool that is visible to the
+\Aa program as well as to the outside world.  The visibility
+to the outside world is provided by access functions
+that the outside world can use to read/write from this
+shared memory pool.
+
+You will have to go through the following sequence:
 \begin{verbatim}
@@
 
+In this example, we are saying that the shared memory pool
+variable is named mempool, and it is an array of 1024 {\bf bytes}.
 The generated top-level VHDL entity has the following
@@
 port to the external world (through the mem\_load.. and mem\_store..)
-ports.
+ports.  The shared memory variable is kept in a memory
+space and is guaranteed to have a base address of $0$.  Thus,
+mempool[I] will be present at address $I$.
 
@@
 
-This example is trivial.  Real, non-trivial programs can be
-mapped in this manner.  Currently, there are only
-two restrictions
+Currently, there are only two restrictions in mapping
+a C program to VHDL using the AhirV2 flow:
 \begin{itemize}
@@
 
-\section{To be continued}
+\subsection{Pipelines}
 
-This document is a work-in-progress.  More details and
-related documentation will be added shortly.
+A pipeline is a powerful mechanism by which an application
+can be parallelized.
+
+Consider the following pipeline which has two stages
+{\bf foo} and {\bf bar}:  
+\begin{verbatim}
+Environment --> foo --> bar --> Environment
+\end{verbatim}
+The stage {\bf foo} takes a 32-bit
+integer from the external world, complements it and passes it on to stage
+bar.  The stage {\bf bar} takes the 32-bit number from {\bf foo}, complements
+it and sends it to the external world (in effect, nothing useful
+is done, this is only an illustration).
+
+If we are writing this pipe-line as a program, we could implement
+two independent processes (or threads) and use named pipes to
+perform the communication between these processes/threads (and the
+``outside´´ world).  For example:
+\begin{verbatim}
+#include <iolib.h>
+void foo()
+{
+   while(1)
+   {
+     uint32_t data = read_uint32("inpipe");
+     write_uint32("midpipe", ~data);
+   }
+}
+
+void bar()
+{
+   while(1)
+   {
+     uint32_t data = read_uint32("midpipe");
+     write_uint32("outpipe", ~data);
+   }
+}
+\end{verbatim}
+The two functions can be compiled into separate processes
+or threads in a multi-threaded program (using pthreads, for
+example), and one gets a software pipeline, in which
+the inter-process communication is done using the
+read/write function calls, which use named pipes.
+
+One can also map this pipeline to a VHDL system, using the
+following flow.  One uses the standard flow that has already
+been described:
+\begin{verbatim}
+clang -std=gnu89 -I../../../iolib/ -emit-llvm -c prog.c
+llvm2aa prog.o | vcFormat >  prog.o.aa
+Aa2VC -O -I mempool -C prog.o.aa | vcFormat > prog.o.aa.vc
+vc2vhdl -C -s ghdl -T foo -T bar -f prog.o.aa.vc\
+           | vhdlFormat > system.vhdl
+\end{verbatim}
+Note that in this case, we use the -T option in vc2vhdl to
+specify that foo and bar are free-running top-level modules.
+The resulting AHIR system in system.vhdl implements the pipeline:
+the system has interfaces corresponding to the pipes inpipe and
+outpipe through which data is exchanged.
 

### 2011-07-13 v2/Ahir-asm/devel/docs/AaLRM.tex
Bug fix: if interface object is target/source of statements, then dependency analysis was being incorrectly handled.

@@
 
-{\bf Aa} is a low-level {\em algorithm-assembly}
-programming language for the description
-of algorithms.   An \Aa program 
-can be modeled by a petri net
+{\bf Aa} is a programming language for the 
+description of algorithms.   The control-flow
+in an \Aa program corresponds to a petri net
 of a specific class (the type-2 petri-net as
 introduced in \cite{ref:SameerPhD}).  
+The data-flow is constructed using static
+single-assignment variables (which can be assigned to only once),
+storage objects, messaging pipes and constants.
+Pointers to storage objects can be created
+and manipulated (pointer arithmetic, dereferencing
+etc.) in the usual way.
+
 A program in \Aa can also be viewed as a 
 description of a system which reacts with 
-its environment through input and output ports and 
-through message queues.  Thus, an \Aa program
-can either be executed on a computer, or
-be mapped to a logic circuit.
+its environment through input and output ports.
+Thus, an \Aa program can either be executed on a 
+computer, or be mapped to a logic circuit.
 
@@
 \begin{verbatim}
-program := ( module-definition | program-object-declaration )*
-\end{verbatim}
-Program object declarations
-can belong to one of three classes:
-\begin{itemize}
-\item {\bf Storage} objects declared in the program
-are variables that can be written into from multiple
-points in the program.  When declared at the program
-scope, storage objects are visible throughout
-the program.
-\item {\bf Pipe} objects declared at the program scope
-are globally visible first-in-first-out buffers that can 
-be read from anywhere in the program and can be
-written into from anywhere in th program.  A pipe
-is a single-place buffer.  A write to a pipe succeeds
-only if the single place buffer is empty.  A read
-from the pipe succeeds only if the single place
-buffer is not empty (the read empties the buffer).
-Thus, pipes can be used for synchronization in 
-an \Aa program.
-\begin{itemize}
-\item A pipe that is only written into in the program, but
-is not read from, is assumed to have a destination outside
-the program.  Such a pipe is an output port of the
-program.
-\item A pipe that is only read from in a program but not
-written into, is assumed to have a source outside the program.
-Such a pipe is an input port of the program.
-\end{itemize}
-\item {\bf Constant} objects declared at
-the program scope are globally visible and have
-initial values which can never be altered in the program.
-Thus, an assignment to a constant is an error.
-\end{itemize}
+program := ( module-definition | program-variable-declaration )*
+\end{verbatim}
+Program variable declarations
+can belong to one of three classes as described in Section \ref{sec:Objects}.
 
@@
 
-\section{Statements}
+\section{Program Variables}
+
+Variables in  the program can belong to the following classes:
+\begin{itemize}
+\item {\bf Storage} variables declared in the program
+can be written into from multiple
+points in the program.  When declared at the program
+scope, storage variables are visible throughout
+the program.  Storage variables can be thought of 
+as memory locations which can be accessed from different
+points in the program.
+\item {\bf Pipe} variables declared at the program scope
+are globally visible first-in-first-out (FIFO) buffers that can 
+be read from anywhere in the program and can be
+written into from anywhere in th program.  A pipe
+is a FIFO buffer with a specified number of
+slots.  A write to a pipe succeeds
+only if there is an empty slot in the FIFO (a successful
+write uses up an empty slots.  A read
+from the pipe succeeds only if there is at
+least one occupied slot in the FIFO (a successful read
+creates an empty slot in the FIFO). 
+Thus, pipes provide a mechanism for synchronization and
+for message passing in an \Aa program.
+\begin{itemize}
+\item A pipe that is only written into in the program, but
+is not read from, is assumed to have a destination outside
+the program.  Such a pipe is an output port of the
+program.
+\item A pipe that is only read from in a program but not
+written into, is assumed to have a source outside the program.
+Such a pipe is an input port of the program.
+\end{itemize}
+\item {\bf Constant} variables declared at
+the program scope are globally visible and have
+initial values which can never be altered in the program.
+Thus, an assignment to a constant is an error.
+\item {\bf Implicit} or {\bf Static Single Assignment}
+variables are created by statements in the \Aa program
+(as described below in Section \ref{sec:Statements}).
+These are analogous to registers that can be written
+into only from one point,  but can be
+read from several points in the program.
+\end{itemize}
+
+\section{Statements} \label{sec:Statements}
 
@@
 means in the sequel).
-Some examples of simple statements are 
+Some examples of assignment statements are 
 \begin{verbatim}
@@
 Section \ref{sec:Syntax}.
+
 \item The {\bf null} statement does nothing, and just passes the
@@
 Block statements can be used to describe complex
-control flows, and consist of a sequence of statements.
+control flow,  and consist of a sequence of statements.
 Block statements are of
@@
                 sequence-of-atomic-statements
-             }
+             } [(variable-exports)]
 \end{verbatim}
@@
    q := ( b * 2)
-}
+} ( q => q )
 \end{verbatim}
@@
 able to be the target of multiple assignment statements.
+The optional export specification
+\begin{verbatim}
+( q => q )
+\end{verbatim}
+after the curly brace \} specifies that the variable
+$q$ defined inside the block $s1$ is visible outside
+the block as $q$ (so that other scopes can use this
+variable, see Section \ref{sec:Scopes} below).
 \item {\bf Parallel} block statements are of the form
@@
 \item {\bf Fork} block statements are generalizations of the
-parallel block region in the sense that they
-allow the programmer to express complex fork-join interactions.
+parallel block region, and they allow the programmer to express 
+complex fork-join interactions.
 They have the following structure:
@@
                 sequence-of-statements
-             }
+             } [(variable-exports)]
 \end{verbatim}
@@
              branch-block-specifier block-name
-             { branch-block-statement-sequence }
+             { 
+		 branch-block-statement-sequence 
+	     } [(variable-exports)]
 \end{verbatim}
@@
 \item The place statement: The {\bf place} statement
-identifies a place into which it places
-a token after it has executed.  For example,
+identifies a place which will contain 
+a token after the place statemet has executed.  For example,
 \begin{verbatim}
@@
 the merge statement starts and executes a series of special assignments
-which multiplex values into new variables based on which arc the
+which multiplex values into targets based on which arc the
 token arrived from.  The merge statement then releases its token to
@@
 \begin{verbatim}
+phi-spec target-expr := expr1 on-spec place-spec-1  [expr2 on-spec place-spec-2 ... ]
+\end{verbatim}
+For example:
+\begin{verbatim}
 $phi a := b $on $entry c $on loopback 
@@
 This says that the target $a$ is to take the value of $b$ if
-the merge statement was started from its entry place (ie from a
+the merge statement was started from its entry place (from a
 token passed from its predecessor) and is to take the value of $c$
@@
                          // its predecessor or by  
-                         // a token in the place "loop"
+                         // a token in the place "loopback"
 
@@
   $endif
-}
+} (r => r)
 \end{verbatim}%$
@@
 impossible for there to be a token present in more
-than one input place to a merge.
+than one input place to a merge.  The final export
+makes $r$ directly visible as $r$  in the parent scope of the
+branch-block.
 \end{itemize}
@@
 scopes.  
+\item A scope can export a variable defined within its scope to
+its parent scope.   The name  of the exported variable must not
+clash with the name of any other variable visible in the
+parent scope.
 \item A scope can read from variables that are defined in ancestor scopes.
@@
 which are not defined in either an ancestor or a descendant
-of the scope.
+of the scope (the export mechanism should be used if this
+is needed).
 
@@
 a binary sequence of the specified width.
+
 \item Pointers:  A pointer is an unsigned
 integer with a default pointer width
-(set to 32 for now), which specified the
+(set to 32 for now), which specifies the
 type of object to which it points.
@@
 must refer to a type.
-For exampe, 
+For example, 
 \begin{verbatim}
@@
 \end{verbatim}
-is an two-dimensional array type whose
+is a two-dimensional array type whose
 elements are one dimensional arrays of 32-bit
@@
 \end{verbatim}
-except that in \Aa, the destination type need not
+The destination type need not
 be the same size as the source type (higher
@@
 
-\section{Memory Spaces}
+\section{Storage variables and Memory Spaces}
 
 An \Aa program can contain declarations to storage objects.  These
-storage objects need to be grouped into memory spaces using
-the following rule: two storage objects are put in the same
-memory space if there is a value in the program which
-is infected by (determined by) the values of pointers to the two objects.
+storage objects are grouped into memory spaces using
+the following rule: two storage objects must be in the same
+memory space if there is a de-referenced pointer in the
+program which can point to either of these two storage objects.
 
@@
    $storage w: $uint<32>
-
+   $storage ptr: $pointer< $uint<32> >
    w := a
@@
 \end{verbatim}
-Now consider the object {\bf ptr}, which can hold a value which
+Now consider the storage variable {\bf ptr}, which can hold a value which
 is a pointer to either {\bf u} or a pointer to {\bf v}.  Thus,
@@
 memory space which contains {\bf u} and {\bf v}.
+
+\subsection{Pointers from ``outside´´? }
 
@@
 the external object and {\bf u} must be kept in
-the same memory space.  In this case, the external
+the same memory space.  In this case, there
+are two possible solutions: either the external
 object needs
-to be made internal, and moved {\bf inside} the system.  
+to be made internal, and moved {\bf inside} the system,
+or, the object {\bf u} must be moved outside the system
+and be made part of the outside world.
 This is done by the {\bf AaLinkExtMem} utility

### 2011-07-13 v2/Ahir-asm/devel/include/AaObject.h

@@ -94,18 +94,33 @@ class AaObject: public AaRoot
 };
 
 // interface object: function arguments
+class AaStatement;
 class AaInterfaceObject: public AaObject
 {
   // arguments can be input or output
   string _mode; // "in" or "out"
 
+  AaStatement* _unique_driver_statement;
  public:
   string Get_Mode() {return(this->_mode);}
 
   AaInterfaceObject(AaScope* scope_tpr, string oname, AaType* otype, string mode);
   ~AaInterfaceObject();
   virtual string Kind() {return("AaInterfaceObject");}
+  virtual bool Is_Interface_Object() {return(true);}
   // uses AaObject::Print method
+
+  void Set_Unique_Driver_Statement(AaStatement* stmt)
+  {
+    if(_unique_driver_statement == NULL)
+      _unique_driver_statement = stmt;
+
+  }
+  AaStatement* Get_Unique_Driver_Statement()
+  {
+    return(_unique_driver_statement);
+  }
+  
 };
 
 class AaConstantObject: public AaObject

### 2011-07-15 v2/CtestBench/docs/CtestBench.tex
fixed some bugs in offset calculation.

@@
 
-We assume that you have a {\bf Release} directory of the AhirV2 flow
-installed.   The Release/bin subdirectory contains the binaries for
-the translation of LLVM-byte-code down to a VHDL description. 
-The Release/vhdl directory contains VHDL library elements that 
-are instantiated in this VHDL description.  The Release/CtestBench 
-directory contains the libraries and headers that you will need.  
-Please see Release/CtestBench/examples/helloworld for an 
-illustration of the whole flow.
 \end{document}

### 2011-07-15 v2/Ahir-asm/devel/src/AaExpression.cpp

@@ -225,20 +225,30 @@ void AaObjectReference::Assign_Expression_Value(AaValue* expr_value)
 }
 
 // return -1 if not evaluatable.
-int AaObjectReference::Evaluate(vector<AaExpression*>* indices, vector<int>* scale_factors)
+int AaObjectReference::Evaluate(vector<AaExpression*>* indices, vector<int>* scale_factors, vector<int>* shift_factors)
 {
   bool address_is_const = true;
   int address_offset = 0;
+
+
   if(indices != NULL)
     {
+      
+      assert(scale_factors);
+      assert(shift_factors);
+      assert(scale_factors->size() == shift_factors->size());
+
       for(int idx = 0; idx < indices->size(); idx++)
 	{
 	  if(!(*indices)[idx]->Is_Constant())
 	    address_is_const = false;
 	  else
 	    {
-	      address_offset += (*indices)[idx]->Get_Expression_Value()->To_Integer() *
-		(*scale_factors)[idx];
+	      if((*shift_factors)[idx] < 0)
+		address_offset += (*indices)[idx]->Get_Expression_Value()->To_Integer() *
+		  (*scale_factors)[idx];
+	      else
+		address_offset += (*shift_factors)[idx];
 	    }
 	}
     }
@@ -631,7 +641,7 @@ void AaSimpleObjectReference::Write_VC_Control_Path( ostream& ofile)
       // as well.
       else if(this->_object->Is("AaStorageObject"))
 	{
-	  this->Write_VC_Load_Control_Path(NULL,NULL,ofile);
+	  this->Write_VC_Load_Control_Path(NULL,NULL,NULL,ofile);
 	}
       // else if the object being referred to is
       // a pipe, instantiate a series r-a
@@ -671,7 +681,7 @@ void AaSimpleObjectReference::Write_VC_Control_Path_As_Target( ostream& ofile)
 
       // followed by several parallel stores..
       // note that you will need a split operation here
-      this->Write_VC_Store_Control_Path(NULL,NULL,ofile);
+      this->Write_VC_Store_Control_Path(NULL,NULL,NULL,ofile);
     }
   // else if the object being referred to is
   // a pipe, instantiate a series r-a
@@ -763,7 +773,7 @@ void AaSimpleObjectReference::Write_VC_Constant_Wire_Declarations(ostream& ofile
[…]

### 2011-07-25 Documents/AhirV2/AhirV2.tex
Small bug-fix: race-through in non-pipelined module.  fin_ack will retrigger.

@@
-\documentclass{article}
+%%\begin{figure}
+%%\begin{centering}
+%%\centerline{\psfig{figure=OurApproach.eps,width=3.0in,height=3.0in}}
+%%\caption{Our Approach to Fault Simulation using a Hardware Emulator}
+%% \label{fig:OurApproach}
+%%\end{centering}
+%%\end{figure}
+
+\documentclass[12pt]{article}
+\usepackage{epsfig}
 
@@
 
-\section{Introduction}
-
-We describe the essential features
-of the second version of the AhirV2 toolset
-developed at IIT-Bombay.
-
 \section{What is AhirV2?}
 
-Perhaps the simplest way to understand AhirV2 is
-through an example.
+AhirV2 is a set of tools 
+\begin{itemize}
+\item which can convert a C/C++ progam to  an equivalent hardware
+implementation (described in VHDL).
+\item which provide an algorithmic approach to the design of 
+hardware.
+\end{itemize}
 
-Suppose we start with a simple program.
-\begin{verbatim}
-int add(int a, int b)
-{
-        int c = (a+b);
-        return(c);
-}
-\end{verbatim}
-This program describes an algorithm.  We normally
-compile this program and convert it to a machine-level
-program which is executed on a processor.
+The flow of transformations is illustrated in Figure \ref{fig:AhirFlow}.
+\begin{figure}
+\begin{centering}
+\centerline{\psfig{figure=AHIRv2Flow.eps,width=3.0in,height=3.0in}}
+\caption{AhirV2 flow}
+ \label{fig:AhirFlow}
+\end{centering}
+\end{figure}
 
-What if we wished to convert this program to
-a circuit?  
-Such a circuit would have inputs 
-$a,b$ and an output $c$ (together with some
-interface hand-shake signals).  When activated,
-the circuit should read its inputs and compute 
-a response which is the same as would be expected
-if the program were executed on a computer.
-
-AhirV2 consists of a set of tools which takes
-a C/C++ program and converts this collection to
-a hardware circuit (described in VHDL).   This
-conversion is done in two steps:
 \begin{itemize}
-\item The high-level program is compiled
-to an interemediate assembly form.  AhirV2
+\item Given a high-level C/C++ program, we rely on
+an LLVM (www.llvm.org) compatible compiler such as clang (www.clang.org)
+or dragon-egg (dragonegg.llvm.org) to produce LLVM byte code.
+Currently, the AhirV2 flow uses LLVM byte code as a starting point.
+\item The LLVM byte-code program is compiled
+to an intermediate assembly form.  AhirV2
 introduces an intermediate assembly language
-\Aa  which can serve as a target for sequential
+\Aa  which serves as a target for sequential
 programming languages (such as C/C++) as
 well as for parallel programming languages.
+An \Aa program consists of modules (analogous
+to sub-programs in C/C++) which can call each 
+other, and can communicate through storage objects
+as well as through pipes (first-in-first-out buffers).
 \item From the \Aa description, a virtual
@@
 The chief optimization carried out at this
-step is that decomposing the system memory
-into disjoint segments based on usage
+step is the decomposition of the system memory
+into disjoint spaces based on usage
 analysis (this considerably improves the
 available memory bandwidth and reduces
-system cost).
+system cost).  A \vC description also
+consists of modules: however, the modules
+are presented in a factored form (control X data X storage).
 \item From the \vC description, a
 VHDL description of the system is generated.
+The system consists of modules, memory spaces
+and FIFO buffers.  The modules are further
+broken down into a control-path (a live and safe
+Petri net), and a data-path (a graph of operators
+and wires).
 The chief optimization carried out at this
@@
 elements and memory elements.
-\item 
-For the C/C++ to \Aa translation, the
-AhirV2 toolset uses the front-end C/C++ compiler
-infrastructure developed by the LLVM project (www.llvm.org), which
-compiles the C/C++ program to LLVM byte-code
-(which is a processor independent representation
-of the compiled program).  The AhirV2 toolset
-has a utility which translates the LLVM byte-code
-to an \Aa description.
-\end{itemize}
-In Section \ref{sec:Example}, 
-we illustrate the use of these tools on the
-simple example shown above.
-
-The earlier version of the AHIR flow has been
-used to convert non-trivial programs to hardware.
-The resulting circuits are upto two orders of
-magnitude more energy-efficient than
-a processor \cite{ref:dsd2010}.  The AhirV2
-toolset incorporates several optimizations which
-make the resulting circuits smaller and 
-faster than before.
-
-\section{The tools}
-
-We assume that you have access to either {\bf llvm-gcc}
-or {\bf clang} as the front-end compiler which generates
-LLVM byte-code from C/C++.  The current AhirV2 toolset
-is consistent with llvm 2.8 and clang 2.8.
-
-The other tools in the chain are described below.
-
-\subsection{{\bf llvm2aa}}
-
-This tool takes LLVM byte code and converts it into an
-\Aa file.
-\begin{verbatim}
-llvm2aa [-modules=<listfile>] [-storageinit] bytecode.o > bytecode.aa
-\end{verbatim}
-The generated \Aa code is sent to {\bf stdout} and all informational
-messages are sent to {\bf stderr}.  On success, the tool returns 0.
-
-The options:
-\begin{itemize}
-\item {\bf -modules=listfile} : Specify the list of functions in the bytecode
-which should be converted to \Aa.   The names of these functions should be
-listed in the text-file listfile. If absent, all functions
-are converted.
-\item {\bf -storageinit} :  Storage objects in the llvm bytecode
-are explicitly initialized in the generated \Aa code.   An initializer
-routine {\bf global\_storage\_initializer\_} is instantiated in
-the \Aa code for this purpose.
 \end{itemize}
 
-\subsection{{\bf AaLinkExtMem}}
+Thus, to generate hardware using the AhirV2 flow,
+it is possible to start at C/C++, at the \Aa level
+at the \vC level or at the VHDL level (or a combination
+of all these levels).  Starting at a higher level is
+easier for the programmer, but using lower level representations
+will probably lead to more efficient hardware.
+Typically, if we start from C/C++, circuits
+produced by the AhirV2 flow are upto two orders of
+magnitude more energy-efficient than
+a processor \cite{ref:dsd2010}.
 
-This tool takes a list of \Aa files, elaborates the program,
-does memory space decomposition.  The externally visible memory space is
-linked in one of two ways: either it is assumed to be external
-and all accesses to it are routed out of the \Aa program,
-or it is assumed to be internal and assumed to consist of
-a memory object (an array of bytes).  External pointer dereferences
-are handled as if they are directed at this memory object.
-\begin{verbatim}
-AaLinkExtMem [-I n] [-E obj-name] bytecode.o > bytecode.aa
-\end{verbatim}
-The generated \Aa code is sent to {\bf stdout} and all informational
-messages are sent to {\bf stderr}.  On success, the tool returns 0.
-
-The options:
+Currently, there are only two restrictions in mapping
+a C/C++ program to VHDL using the AhirV2 flow:
 \begin{itemize}
-\item {\bf -I n}: specifies that external references to memory
-are to be mapped as if they are to an internal object whose size
-is $n$ bytes.
-\item {\bf -E obj-name} : specifies that the object to which
-external references are mapped is to be named obj-name.
-\end{itemize}
-It is recommended that you use the {\bf -I} and {\bf -E} options.
-
-If the {\bf -I} option is not used, then all external memory
-references are routed out of the \Aa program through pipes.
-In this case, if the \Aa compiler determines that there is some pointer 
-in the program which can point
-to both internal and external memory, then this will be
-declared as an error!
-
-\subsection{{\bf Aa2VC}}
-
-This tool takes a list of \Aa programs and converts them
-to a \vC description. 
-\begin{verbatim}
-Aa2VC [-O] [-C] [-I obj-name] file1.aa file2.aa ...  > result.vc
-\end{verbatim}
-The generated \vC code is sent to {\bf stdout} and all informational
-messages are sent to {\bf stderr}.  On success the tool returns 0.
-
-The options:
-\begin{itemize}
-\item {\bf -O} : if used, sequential statement blocks are parallelized
-by doing dependency analysis.
-\item {\bf -C} : if used, a C stub is created for every module that
-is not called from within the system.  These stubs can be used to
-interface to a VHDL simuator (or even drive hardware) to verify
-the VHDL code generated by downstream tools.
-\item {\bf -I obj-name} : if specified, all external memory references
-are considered as being directed at the storage object named obj-name.
-If not specified, then the tool will throw an error if it finds
-a pointer dereference that cannot be resolved as pointing only to
-storage objects declared inside the \Aa program.
+\item No recursion, no cycles in the call-graph of the original
+program.
+\item No function pointers.
 \end{itemize}
 
-\subsection{\bf vc2vhdl}
+\section{Use Scenario 1:  a transformational system}
 
-Takes a collection of \vC descriptions and converts them to
-an AHIR system described in VHDL.
-\begin{verbatim}
-vc2vhdl [-O] [-C] [-a] [-e <entity-name] [-w]\
-         -t/-T foo [-t/-T bar -t/-T bar2 ...]\
-         -f file1.vc -f file2.vc ...  > system.vhdl
-\end{verbatim}
+Many programs are transformational in nature.  There
+is some input data $X$, and the program can be modeled as
+a function $f$ which acts on $X$ to produce output data $Y$.
 
-The options:
-\begin{itemize}
-\item {\bf -t} : to specify the modules which are to be 
-accessible from the ports of the generated VHDL system.
-Such modules have to be top-level (that is, they cannot
-be called from within the program).
-Multiple top-level modules can be specified in this way.
-The control and argument ports for these modules are
-visible at the interface of the generated AHIR system.
-\item {\bf -T} : to specify top-level modules which are to be 
-free-running inside the AHIR system.
-Multiple top-level modules can be specified in this way.
-Such modules do not have any arguments and do not return
-any values.  Their only mechanism of communication with
-the world outside the AHIR system is through pipes.
-The control ports for these modules are
-{\bf not visible} at the interface of the generated AHIR system.
-In the AHIR system, these modules are started on reset
-and are run forever (restarted after they finish, forever).
-\item {\bf -f} : specifies the \vC files to be analyzed. 
-Multiple \vC files may be specified.  An object must be
-defined before it is used, so the \vC files must be 
-specified in the correct order.
-\item {\bf -O} : optimize the generated VHDL by compacting
-the control-path.  This does not change the resulting
-hardware, but makes the generated VHDL file smaller.
-\item {\bf -C} : the VHDL code has a system test bench which
-interfaces to foreign code using a VHPI/Modelsim-FLI interface.
-If this is not specified, the  generated test bench simply
-instantiates the system and starts all top-level modules
-off (you will need to fill in your own test bench here).
-The C testbench is usually easier to write (it probably
-already exists in the form of the original program).
-\item {\bf -a} : try to minimize the area of the resulting
-VHDL by sharing operators to the maximum extent possible
-(allowing potential contention for resources).  This will
-result in a slower (usually by 2X) system, but will
-also reduce the area (usually by 0.5X).  If not specified,
-two operations will be mapped to the same
-operator  only if it can be proved that they cannot be active simultaneously.
-\item {\bf -e top-entity-name} : The generated top-level VHDL entity
-corresponding to the AHIR system is named top-entity-name.  The default
-is test\_system.
-\item {\bf -w} :  If specified, the VHDL system and test-bench are
-generated as separate unformatted VHDL files.  You will need to
-format these using the vhdlFormat command.
-\item {\bf -s ghdl/modelsim} :  If {\bf ghdl} is specified
-with the -s option, then the generated testbench (if -C is specified)
-uses the VHPI interface to link with foreign code.  Otherwise,
-the generated testbench (if -C is specified) uses the Modelsim FLI
-interface to link with foreign code.
-\end{itemize}
+\subsection{A trivial example} \label{sec:Example}
 
-The tool performs concurrency analysis to determine operations which
-can be mapped to the same physical operator without the need for
-arbitration.  It also instantiates separate memory subsystems for
-the disjoint memory spaces (in practice many of the memory spaces
-are small and are converted to register banks).
-
-\subsection{Miscellaneous: {\bf vcFormat} and {\bf vhdlFormat}}
-
-The outputs produced by {\bf Aa2VC} and {\bf vc2vhdl} are
-not well formatted.  One can format \Aa and \vC files
-using  {\bf vcFormat} as follows
-\begin{verbatim}
-vcFormat < unformatted-vc/aa-file  > formatted-vc/aa-file
-\end{verbatim}
-and similarly use {\bf vhdlFormat} to format generated
-VHDL files.
-
-\section{An example} \label{sec:Example}
-
-Let us revisit the simple example considered in
-the first section:
+Consider the following trivial example:
 \begin{verbatim}
@@
 \end{verbatim}
-The important points to note are that the stored objects a,b,c and
+
+Note that the stored objects a,b,c and
 ret\_val\_\_ are mapped to different memory spaces.  Thus, the
@@
     add_tag_out: out std_logic_vector(0 downto 0);
-    add_start : in std_logic;
-    add_fin   : out std_logic;
+    add_start_req : in std_logic;
+    add_start_ack : out std_logic;
+    add_fin_req   : in std_logic;
+    add_fin_ack   : out std_logic;
     clk : in std_logic;
@@
 \end{verbatim}
-There are ports corresponding to the arguments of the top-level module,
-and the add\_start/add\_fin is a handshake pair.  One sets up the inputs,
-starts the system and waits until the fin is asserted.  After the fin
-is asserted, one has the return value of {\bf add} at the appropriate port.
 
-\section{External and internal memory spaces}
+The AHIR system is a purely synchronous implementation
+which uses only the rising edge of the {\bf clk} input.
+The {\bf reset} is synchronous, and is active high.
 
+The VHDL implementation of a module in the AHIR system
+corresponds to that of a pipeline stage (see Figure \ref{fig:ModulePipeline}).
+\begin{figure}
+\begin{centering}
+\centerline{\psfig{figure=ModulePipeline.eps,width=3.0in,height=3.0in}}
+\caption{System and Module Interfaces}
+ \label{fig:ModulePipeline}
+\end{centering}
+\end{figure}
+The ports 
+\begin{verbatim}
+add_a add_b
+\end{verbatim}
+correspond to the input arguments of the top-level 
+function {\bf add}.  The port 
+\begin{verbatim}
+ret_val_x_x
+\end{verbatim}
+corresponds to the value returned by {\bf add}.
+The pair of ports
+\begin{verbatim}
+add_start_req  add_start_ack
+\end{verbatim}
+implement a {\bf start} protocol; The environment
+asserts {add\_start\_req} when it wants to start
+{\bf add}, and the AHIR system asserts {\bf add\_start\_ack}
+whenever it is ready to start.  The environment is required
+to hold the input arguments steady until it observes
+the acknowledge from the AHIR system.
+The pair of ports
+\begin{verbatim}
+add_fin_req  add_fin_ack
+\end{verbatim}
+implement a {\bf finish} protocol; The environment
+asserts {add\_fin\_req} when it is in a position to accept
+the returned value from a previously started 
+{\bf add}, and the AHIR system asserts {\bf add\_fin\_ack}
+whenever it has a return-value avaiable.  The returned value
+is valid only when the acknowledge from the AHIR system is
+asserted. 
+The ports
+\begin{verbatim}
+    add_tag_in  add_tag_out
+\end{verbatim}
+provide a mechanism by which a tag can be presented
+by the environment to the AHIR system to identify
+a particular invocation of the {\bf add} function.  The
+width of the tag is chosen (by the AhirV2 tool-chain,
+specifically, the vc2vhdl tool) to be large enough
+that each active call to {\bf add} can be identified
+uniquely.
+
+\section{Use Scenario 2: pipelines}
+
+A pipeline is a collection of parallel processes which
+work together to accomplish a certain function (or to finish
+a job).   Pipelines are commonly used in software and
+in hardware systems.
+
+The concept of pipes or sockets provide a natural mechanism
+for communication between parallel processes in a software
+pipeline.  In \Aa (and \vC), pipes are first-in-first-out
+buffers which provide a corresponding communication construct
+in the generated hardware.
+
+\subsection{An example of a trivial pipeline}
+
+Consider the following pipeline which has two stages
+{\bf foo} and {\bf bar}:  
+\begin{verbatim}
+Environment --> foo --> bar --> Environment
+\end{verbatim}
+The stage {\bf foo} takes a 32-bit
+integer from the external world, complements it and passes it on to stage
+bar.  The stage {\bf bar} takes the 32-bit number from {\bf foo}, complements
+it and sends it to the external world (in effect, nothing useful
+is done, this is only an illustration).
+
+If we are writing this pipe-line as a program, we could implement
+two independent processes (or threads) and use named pipes to
+perform the communication between these processes/threads (and the
+``outside´´ world).  For example:
+\begin{verbatim}
+#include <iolib.h>
+void foo()
+{
+   while(1)
+   {
+     uint32_t data = read_uint32("inpipe");
+     write_uint32("midpipe", ~data);
+   }
+}
+
+void bar()
+{
+   while(1)
+   {
+     uint32_t data = read_uint32("midpipe");
+     write_uint32("outpipe", ~data);
+   }
+}
+\end{verbatim}
+In this example, the read/write functions are assumed
+to use named pipes for inter-process communication.
+
+The two functions can be compiled into separate processes
+or can be used in threads in a multi-threaded program (using pthreads, for
+example), and one gets a software pipeline, in which
+the inter-process communication is done using the
+read/write function calls, which use named pipes.
+
+One can also map this pipeline to a VHDL system, using the
+following flow.  One uses the standard flow that has already
+been described:
+\begin{verbatim}
+clang -std=gnu89 -I../../../iolib/ -emit-llvm -c prog.c
+llvm2aa prog.o | vcFormat >  prog.o.aa
+Aa2VC -O -I mempool -C prog.o.aa | vcFormat > prog.o.aa.vc
+vc2vhdl -C -s ghdl -T foo -T bar -f prog.o.aa.vc\
+           | vhdlFormat > system.vhdl
+\end{verbatim}
+Note that in this case, we use the -T option in vc2vhdl to
+specify that foo and bar are free-running top-level modules.
+\begin{figure}
+\begin{centering}
+\centerline{\psfig{figure=FooBarPipeline.eps,width=3.0in,height=3.0in}}
+\caption{Hardware implementation of Foo-Bar pipeline}
+ \label{fig:FooBarPipeline}
+\end{centering}
+\end{figure}
+The resulting AHIR system in system.vhdl implements the pipeline
+(the block diagram of the system is shown in Figure \ref{fig:FooBarPipeline})
+with the following interface:
+\begin{verbatim}
+entity test_system is  -- system 
+  port (-- 
+    clk : in std_logic;
+    reset : in std_logic;
+    inpipe_pipe_write_data: in std_logic_vector(31 downto 0);
+    inpipe_pipe_write_req : in std_logic_vector(0 downto 0);
+    inpipe_pipe_write_ack : out std_logic_vector(0 downto 0);
+    outpipe_pipe_read_data: out std_logic_vector(31 downto 0);
+    outpipe_pipe_read_req : in std_logic_vector(0 downto 0);
+    outpipe_pipe_read_ack : out std_logic_vector(0 downto 0)); -- 
+  -- 
+end entity;
+\end{verbatim}
+The system has interfaces corresponding to the pipes inpipe and
+outpipe through which data is exchanged.
+
+In practice, one can have any number of pipes and 
+interacting processes in a pipeline implementation.
+
+\section{Storage variables and memory spaces}
+
+In an \Aa program, variables can be of three kinds:
+storage variables, pipe variables, or single-static-assignment
+variables.  Storage variables are implemented in memory,
+pipe variables as FIFO buffers, and single-static-assignment
+variables as registers.
+
+While transforming an \Aa description down to a \vC description,
+the storage variables in the \Aa program need to be grouped
+into memory spaces.  Two storage variables are put in the 
+same memory space only if we
+determine that a pointer de-reference in the \Aa
+program can point into either of the two storage variables
+(this is determined by a conservative static analysis).
+
+A memory space in a \vC description is characterized by
+a word-length (the greatest common divisor of the widths
+of accesses to this memory space), an address-width (wide enough
+to allow access to all words in the memory space), and
+a capacity (the number of words in the memory space).  Typically,
+a program will have many small memory spaces corresponding to
+scratch storage and some large memory spaces which correspond
+to arrays etc.
+
+There is one small issue, however.
 Consider the following C program:
@@
   port (--
-    foo_b : in  std_logic_vector(31 downto 0);
-    foo_ret_val_x_x : out  std_logic_vector(31 downto 0);
-    foo_tag_in: in std_logic_vector(0 downto 0);
-    foo_tag_out: out std_logic_vector(0 downto 0);
-    foo_start : in std_logic;
-    foo_fin   : out std_logic;
     clk : in std_logic;
     reset : in std_logic;
+
+    --   some-lines-omitted --
+    --   foo-related ports --
+    --   some-lines-omitted --
+
     extmem_read_address_32_pipe_read_data: out std_logic_vector(31 downto 0);
@@
     mem_load_x_x_tag_out: out std_logic_vector(0 downto 0);
-    mem_load_x_x_start : in std_logic;
-    mem_load_x_x_fin   : out std_logic;
+    mem_load_x_x_start_req : in std_logic;
+    mem_load_x_x_start_ack : out std_logic;
+    mem_load_x_x_fin_req   : in std_logic;
+    mem_load_x_x_fin_ack   : out std_logic;
     mem_store_x_x_address : in  std_logic_vector(31 downto 0);
@@
     mem_store_x_x_tag_out: out std_logic_vector(0 downto 0);
-    mem_store_x_x_start : in std_logic;
-    mem_store_x_x_fin   : out std_logic;
+    mem_store_x_x_start_req : in std_logic;
+    mem_store_x_x_start_ack : out std_logic;
+    mem_store_x_x_fin_req   : in std_logic;
+    mem_store_x_x_fin_ack   : out std_logic;
     clk : in std_logic;
@@
 
-The system provides a memory load and memory store
-port to the external world (through the mem\_load.. and mem\_store..)
-ports.  The shared memory variable is kept in a memory
-space and is guaranteed to have a base address of $0$.  Thus,
-mempool[I] will be present at address $I$.
+The system provides  memory load and memory store
+function interfaces to the external world 
+(through  mem\_load.. and mem\_store..).
+The shared memory variable is guaranteed to have a base address of $0$.  Thus,
+byte mempool[I] will be present at address $I$.
 
-\section{Handling complex programs}
+\section{The tools}
 
-Currently, there are only two restrictions in mapping
-a C program to VHDL using the AhirV2 flow:
+We assume that you have access to either {\bf llvm-gcc}
+or {\bf clang} as the front-end compiler which generates
+LLVM byte-code from C/C++.  The current AhirV2 toolset
+is consistent with llvm 2.8 and clang 2.8.
+
+The other tools in the chain are described below.
+
+\subsection{{\bf llvm2aa}}
+
+This tool takes LLVM byte code and converts it into an
+\Aa file.
+\begin{verbatim}
+llvm2aa [-modules=<listfile>] [-storageinit] bytecode.o > bytecode.aa
+\end{verbatim}
+The generated \Aa code is sent to {\bf stdout} and all informational
+messages are sent to {\bf stderr}.  On success, the tool returns 0.
+
+The options:
 \begin{itemize}
-\item No recursion, no cycles in the call-graph of the original
-program.
-\item No function pointers.
+\item {\bf -modules=listfile} : Specify the list of functions in the bytecode
+which should be converted to \Aa.   The names of these functions should be
+listed in the text-file listfile. If absent, all functions
+are converted.
+\item {\bf -storageinit} :  Storage objects in the llvm bytecode
+are explicitly initialized in the generated \Aa code.   An initializer
+routine {\bf global\_storage\_initializer\_} is instantiated in
+the \Aa code for this purpose.
 \end{itemize}
 
-The AhirV2 flow produces a modular system, and by default,
-produces one VHDL entity for every function in the original
-program.  Further, the concept of message pipes is native
-to the \Aa and \vC descriptions.  Thus, it is easy to
-map parallel programs or programs consisting of multiple
-concurrent processes to hardware.
+\subsection{{\bf AaLinkExtMem}}
 
-\subsection{Pipelines}
+This tool takes a list of \Aa files, elaborates the program,
+does memory space decomposition.  The externally visible memory space is
+linked in one of two ways: either it is assumed to be external
+and all accesses to it are routed out of the \Aa program,
+or it is assumed to be internal and assumed to consist of
+a memory object (an array of bytes).  External pointer dereferences
+are handled as if they are directed at this memory object.
+\begin{verbatim}
+AaLinkExtMem [-I n] [-E obj-name] bytecode.o > bytecode.aa
+\end{verbatim}
+The generated \Aa code is sent to {\bf stdout} and all informational
+messages are sent to {\bf stderr}.  On success, the tool returns 0.
 
-A pipeline is a powerful mechanism by which an application
-can be parallelized.
+The options:
+\begin{itemize}
+\item {\bf -I n}: specifies that external references to memory
+are to be mapped as if they are to an internal object whose size
+is $n$ bytes.
+\item {\bf -E obj-name} : specifies that the object to which
+external references are mapped is to be named obj-name.
+\end{itemize}
+We recommend that you use the {\bf -I} and {\bf -E} options to
+locate externally visible memory into a specified object in the
+\Aa program.
 
-Consider the following pipeline which has two stages
-{\bf foo} and {\bf bar}:  
+If the {\bf -I} option is not used, then all external memory
+references are routed out of the \Aa program through pipes.
+In this case, if the \Aa compiler determines that there is some pointer 
+in the program which can point
+to both internal and external memory, then this will be
+declared as an error!
+
+\subsection{{\bf Aa2VC}}
+
+This tool takes a list of \Aa programs and converts them
+to a \vC description. 
 \begin{verbatim}
-Environment --> foo --> bar --> Environment
+Aa2VC [-O] [-C] [-I obj-name] file1.aa file2.aa ...  > result.vc
 \end{verbatim}
-The stage {\bf foo} takes a 32-bit
-integer from the external world, complements it and passes it on to stage
-bar.  The stage {\bf bar} takes the 32-bit number from {\bf foo}, complements
-it and sends it to the external world (in effect, nothing useful
-is done, this is only an illustration).
+The generated \vC code is sent to {\bf stdout} and all informational
+messages are sent to {\bf stderr}.  On success the tool returns 0.
 
-If we are writing this pipe-line as a program, we could implement
-two independent processes (or threads) and use named pipes to
-perform the communication between these processes/threads (and the
-``outside´´ world).  For example:
+The options:
+\begin{itemize}
+\item {\bf -O} : if used, sequential statement blocks are parallelized
+by doing dependency analysis.
+\item {\bf -C} : if used, a C stub is created for every module that
+is not called from within the system.  These stubs can be used to
+interface to a VHDL simulator (or even drive hardware) to verify
+the VHDL code generated by downstream tools.
+\item {\bf -I obj-name} : if specified, all external memory references
+are considered as being directed at the storage object named obj-name.
+If not specified, then the tool will throw an error if it finds
+a pointer dereference that cannot be resolved as pointing only to
+storage objects declared inside the \Aa program.
+\end{itemize}
+
+\subsection{\bf vc2vhdl}
+
+Takes a collection of \vC descriptions and converts them to
+an AHIR system described in VHDL.
 \begin{verbatim}
-#include <iolib.h>
-void foo()
-{
-   while(1)
-   {
-     uint32_t data = read_uint32("inpipe");
-     write_uint32("midpipe", ~data);
-   }
-}
+vc2vhdl [-O] [-C] [-a] [-e <entity-name] [-w]\
+         -t/-T foo [-t/-T bar -t/-T bar2 ...]\
+         -f file1.vc -f file2.vc ...  > system.vhdl
+\end{verbatim}
 
-void bar()
-{
-   while(1)
-   {
-     uint32_t data = read_uint32("midpipe");
-     write_uint32("outpipe", ~data);
-   }
-}
+The options:
+\begin{itemize}
+\item {\bf -t} : to specify the modules which are to be 
+accessible from the ports of the generated VHDL system.
+Such modules have to be top-level (that is, they cannot
+be called from within the program).
+Multiple top-level modules can be specified in this way.
+The control and argument ports for these modules are
+visible at the interface of the generated AHIR system.
+\item {\bf -T} : to specify top-level modules which are to be 
+free-running inside the AHIR system.
+Multiple top-level modules can be specified in this way.
+Such modules do not have any arguments and do not return
+any values.  Their only mechanism of communication with
+the world outside the AHIR system is through pipes.
+The control ports for these modules are
+{\bf not visible} at the interface of the generated AHIR system.
+In the AHIR system, these modules are started on reset
+and are run forever (restarted after they finish, forever).
+\item {\bf -f} : specifies the \vC files to be analyzed. 
+Multiple \vC files may be specified.  An object must be
+defined before it is used, so the \vC files must be 
+specified in the correct order.
+\item {\bf -O} : optimize the generated VHDL by compacting
+the control-path.  This does not change the resulting
+hardware, but makes the generated VHDL file smaller.
+\item {\bf -C} : the VHDL code has a system test bench which
+interfaces to foreign code using a VHPI/Modelsim-FLI interface.
+If this is not specified, the  generated test bench simply
+instantiates the system and starts all top-level modules
+off (you will need to fill in your own test bench here).
+The C testbench is usually easier to write (it probably
+already exists in the form of the original program).
+\item {\bf -a} : try to minimize the area of the resulting
+VHDL by sharing operators to the maximum extent possible
+(allowing potential contention for resources).  This will
+result in a slower (usually by 2X) system, but will
+also reduce the area (usually by 0.5X).  If not specified,
+two operations will be mapped to the same
+operator  only if it can be proved that they cannot be active simultaneously.
+\item {\bf -e top-entity-name} : The generated top-level VHDL entity
+corresponding to the AHIR system is named top-entity-name.  The default
+is test\_system.
+\item {\bf -w} :  If specified, the VHDL system and test-bench are
+generated as separate unformatted VHDL files.  You will need to
+format these using the vhdlFormat command.
+\item {\bf -s ghdl/modelsim} :  If {\bf ghdl} is specified
+with the -s option, then the generated testbench (if -C is specified)
+uses the VHPI interface to link with foreign code.  Otherwise,
+the generated testbench (if -C is specified) uses the Modelsim FLI
+interface to link with foreign code.
+\end{itemize}
+
+The tool performs concurrency analysis to determine operations which
+can be mapped to the same physical operator without the need for
+arbitration.  It also instantiates separate memory subsystems for
+the disjoint memory spaces (in practice many of the memory spaces
+are small and are converted to register banks).
+
+\subsection{Miscellaneous: {\bf vcFormat} and {\bf vhdlFormat}}
+
+The outputs produced by {\bf Aa2VC} and {\bf vc2vhdl} are
+not well formatted.  One can format \Aa and \vC files
+using  {\bf vcFormat} as follows
+\begin{verbatim}
+vcFormat < unformatted-vc/aa-file  > formatted-vc/aa-file
 \end{verbatim}
-The two functions can be compiled into separate processes
-or threads in a multi-threaded program (using pthreads, for
-example), and one gets a software pipeline, in which
-the inter-process communication is done using the
-read/write function calls, which use named pipes.
-
-One can also map this pipeline to a VHDL system, using the
-following flow.  One uses the standard flow that has already
-been described:
-\begin{verbatim}
-clang -std=gnu89 -I../../../iolib/ -emit-llvm -c prog.c
-llvm2aa prog.o | vcFormat >  prog.o.aa
-Aa2VC -O -I mempool -C prog.o.aa | vcFormat > prog.o.aa.vc
-vc2vhdl -C -s ghdl -T foo -T bar -f prog.o.aa.vc\
-           | vhdlFormat > system.vhdl
-\end{verbatim}
-Note that in this case, we use the -T option in vc2vhdl to
-specify that foo and bar are free-running top-level modules.
-The resulting AHIR system in system.vhdl implements the pipeline:
-the system has interfaces corresponding to the pipes inpipe and
-outpipe through which data is exchanged.
+and similarly use {\bf vhdlFormat} to format generated
+VHDL files.
 

### 2011-08-16 v2/CtestBench/docs/CtestBench.tex
Some documentation in order..

@@
 
+\section{An example}
+
+Consider the following program (lets say it is in file ``prog.c''):
+\begin{verbatim}
+#include <stdlib.h>
+#include <stdint.h>
+#include <stdio.h>
+#include <iolib.h>
+
+uint32_t sum;
+
+void set_sum(uint32_t x)
+{
+    sum = x;
+}
+
+uint32_t get_sum()
+{
+    return(sum);
+}
+
+void accumulate()
+{
+    while(1)
+    {
+        int nxt = read_uint32("in_data");
+#ifdef SW
+        printf("read %u\n", nxt);
+#endif
+        sum = (sum + nxt);
+        write_uint32("out_data",sum);
+#ifdef SW
+        printf("wrote %u\n", sum);
+#endif
+    }
+}
+
+\end{verbatim}
+
+This program describes a {\em system} which listens
+for data on a pipe ``in\_data'', and sends data
+out on a pipe ``out\_data''.  The incoming data
+is accumulated into the variable {\em sum}, and
+there are two methods to set and get the value
+of {\em sum}.
+
+Now to test this program, we can write a test-bench
+such as this one (lets call this file ``testbench.c'').
+\begin{verbatim}
+#include <pthread.h>
+#include <signal.h>
+#include <stdio.h>
+#include <stdlib.h>
+
+#ifdef SW 
+// for the read_*/write_* methods
+#include <iolib.h>
+#include "prog.h"
+#else
+// includes for the read_*/write_* methods
+// as well as stubs for functions moved
+// to hardware.
+#include "vhdlCStubs.h"
+#endif
+
+void Exit(int sig)
+{
+   fprintf(stderr, "## Break! ##\n");
+   exit(0);
+}
+
+void *accumulate_(void* fargs)
+{
+   accumulate();
+}
+
+void *write_pipe_(void* a)
+{
+   write_uint32_n("in_data",(uint32_t*)a, 10);
+}
+
+void *read_pipe_(void* a)
+{
+   read_uint32_n("out_data",(uint32_t*)a, 10);
+}
+
+int main(int argc, char* argv[])
+{
+        signal(SIGINT,  Exit);
+        signal(SIGTERM, Exit);
+
+        uint32_t data_in[10], data_out[10];
+        int i;
+
+        // initial value of sum.
+        set_sum(1);
+
+        for(i = 0; i < 10; i++)
+        {
+          data_in[i] = i;
+        }
+
+        pthread_t acc_t, wpipe_t, rpipe_t;
+
+#ifdef SW
+        pthread_create(&acc_t,NULL,&accumulate_,NULL);
+#endif
+        pthread_create(&wpipe_t,NULL,&write_pipe_,(void*)data_in);
+        pthread_create(&rpipe_t,NULL,&read_pipe_,(void*)data_out);
+
+        pthread_join(wpipe_t,NULL);
+        pthread_join(rpipe_t,NULL);
+
+        fprintf(stdout,"from out_data, we read ");
+        for(i=0; i < 10; i++)
+          fprintf(stdout," %u ", data_out[i]);
+        fprintf(stdout,"\n");
+        fprintf(stdout,"final sum is %u\n", get_sum());
+
+#ifdef SW
+        pthread_cancel(acc_t);
+#endif
+}
+
+\end{verbatim}
+The test-bench sets an initial value for sum, and starts
+three threads: one to write data to pipe in\_data, one
+to read data from pipe out\_data, and one to run the
+accumulate function.  After the last data is read back
+from out\_data, the test-bench gets the value of sum
+and prints it out.
+
+Obviously, we would prefer to use the same test-bench
+to verify that the VHDL system generated from ``prog.c'' 
+functions correctly.  
+The difference is that instead of using methods in iolib,
+the test-bench now uses methods in SocketLib.  Further, the
+VHDL is executed in a VHDL simulator; the simulator communicates
+with the testbench using sockets.
+The {\em ifdef's} in the test-bench and the system program
+indicate the difference between the pure software version
+of the system-test-bench combination and the hardware-software
+version.
+
+\section{An example}
+
+Look at the subdirectory ``example'' which contains the files ``prog.c'' and
+``testbench.c''.  A Makefile and README are also present.  Its all quite
+self-explanatory.
+
 \end{document}

### 2011-08-16 v2/CtestBench/docs/example/prog.c

@@ -1,0 +1,34 @@
+#include <stdlib.h>
+#include <stdint.h>
+#include <iolib.h>
+#include <stdio.h>
+
+
+uint32_t sum;
+
+void set_sum(uint32_t x)
+{
+    sum = x;
+}
+
+uint32_t get_sum()
+{
+    return(sum);
+}
+
+void accumulate()
+{
+    while(1)
+    {
+	int nxt = read_uint32("in_data");
+#ifdef SW
+	printf("read %u\n", nxt);
+#endif
+	sum = (sum + nxt);
+	write_uint32("out_data",sum);
+#ifdef SW
+	printf("wrote %u\n", sum);
+#endif
+    }
+}
+

### 2011-08-25 v2/CtestBench/docs/CtestBench.tex
Release area updated.

@@
 
-\section{An example}
+The following Makefile builds a software-only testbench
+executable, and also converts the system described in prog.c
+to VHDL.  The same testbench can be used to test the VHDL
+also.
+\begin{verbatim}
+# build software version of testbench 
+# (to check the "desired behaviour")
+IOLIB_INCLUDE=../../iolib/include
+IOLIB_LIB=../../iolib/lib
+SOCKETLIB_INCLUDE=../../CtestBench/include
+SOCKETLIB_LIB=../../CtestBench/lib
+VHDL_LIB=../../vhdl
+VHDL_VHPI_LIB=../../CtestBench/vhdl
+all: SW HW
 
-Look at the subdirectory ``example'' which contains the files ``prog.c'' and
-``testbench.c''.  A Makefile and README are also present.  Its all quite
-self-explanatory.
+# compile with SW defined.
+# note the use of IOLIB in building the testbench.
+SW: prog.c prog.h testbench.c 
+        gcc -c -DSW -I$(IOLIB_INCLUDE) prog.c
+        gcc -c -DSW -I$(IOLIB_INCLUDE) testbench.c
+        gcc -o testbench_sw prog.o testbench.o -L$(IOLIB_LIB) -lio -lpthread
 
+# five steps from C to vhdl simulator.
+HW: c2llvmbc llvmbc2aa aa2vc vc2vhdl vhdlsim
+
+# C to llvm byte-code.. use clang.
+c2llvmbc: prog.c prog.h testbench.c
+        clang -std=gnu89 -I$(IOLIB_INCLUDE) -emit-llvm -c prog.c
+        #llvm-dis prog.o
+        opt -O3	 prog.o > prog.opt.o
+        llvm-dis prog.opt.o
+
+# llvm byte-code to Aa..
+llvmbc2aa:  prog.opt.o
+        llvm2aa prog.opt.o | vcFormat >  prog.opt.o.aa
+
+# Aa to vC
+aa2vc: prog.opt.o.aa
+        Aa2VC -O -C prog.opt.o.aa | vcFormat > prog.opt.o.aa.vc
+
+# vC to VHDL
+vc2vhdl: prog.opt.o.aa.vc
+        vc2vhdl -C -e ahir_system -w -s ghdl\
+                -T accumulate -t set_sum -t get_sum\
+                -f prog.opt.o.aa.vc 
+        vhdlFormat < ahir_system.unformatted_vhdl > ahir_system.vhdl
+        vhdlFormat < ahir_system_test_bench.unformatted_vhdl\
+                    > ahir_system_test_bench.vhdl
+
+# build testbench and ghdl executable
+# note the use of SOCKETLIB in building the testbench.
+vhdlsim: ahir_system.vhdl ahir_system_test_bench.vhdl\
+           testbench.c vhdlCStubs.h vhdlCStubs.c
+        gcc -c vhdlCStubs.c -I./ -I$(SOCKETLIB_INCLUDE)
+        gcc -c testbench.c -I./ -I$(SOCKETLIB_INCLUDE)
+        gcc -o testbench_hw testbench.o vhdlCStubs.o\
+               -L$(SOCKETLIB_LIB) -lSocketLib -lpthread
+        ghdl --clean
+        ghdl --remove
+        ghdl -i --work=ahir  $(VHDL_LIB)/ahir.vhdl
+        ghdl -i --work=ieee_proposed  $(VHDL_LIB)/ieee_proposed.vhdl
+        ghdl -i --work=work $(VHDL_VHPI_LIB)/Utility_Package.vhdl
+        ghdl -i --work=work $(VHDL_VHPI_LIB)/Vhpi_Package.vhdl
+        ghdl -i --work=work ahir_system.vhdl
+        ghdl -i --work=work ahir_system_test_bench.vhdl
+        ghdl -m --work=work -Wc,-g -Wl,-L$(SOCKETLIB_LIB)\
+                 -Wl,-lVhpi ahir_system_test_bench
+
+clean:
+        rm -f *.o* *.cf *.*vhdl vhdlCStubs.*\
+             *.vcd in_data* out_data* testbench_sw\
+              testbench_hw ahir_system_test_bench\
+              vhpi.log
+
+PHONY: all clean	
+\end{verbatim}
+To test the software, run testbench\_sw.  To verify the hardware (using
+the VHDL simulator GHDL), start testbench\_sw in one shell, and then start ahir\_system\_test\_bench
+in a different shell.
 \end{document}

### 2011-08-25 v2/llvmbc2Aa/docs/llvm2aa.tex
updated

@@
 the \Aa code for this purpose.
-\item {\bf llvmpasses} : {\bf llvm2aa} uses the LLVM compiler
+\item {\bf llvm2aa} uses the LLVM compiler
 infrastructure to perform LLVM byte-code optimizations.  A large list
@@
 vector and void types are the only ones currently supported.
-\item The LLVM switch instruction is currently not supported,
-but support is planned in the near future.
 \item The LLVM indirect-branch, invoke, unwind and unreachable
@@
 \item The LLVM division and remainder instructions are currently
-not supported.  Support for these is planned.
+not supported. 
 \item LLVM vector instructions are currently not supported.

### 2011-08-29 v2/llvmbc2Aa/docs/llvm2aa.tex
added --pipedepths option to llvm2aa

@@
 \begin{itemize}
-\item {\bf -modules=listfile} : Specify the list of functions in the bytecode
+\item {\bf --modules=listfile} : Specify the list of functions in the bytecode
 which should be converted to \Aa.   The names of these functions should be
@@
 are converted.
-\item {\bf -storageinit} :  Storage objects in the llvm bytecode
+\item {\bf --storageinit} :  Storage objects in the llvm bytecode
 are explicitly initialized in the generated \Aa code.   An initializer
@@
 the \Aa code for this purpose.
+\item {\bf -pipedepths=depthfile} : Specify the maximum depths of
+pipes in the generated Aa program. The file ``depthfile'' is a list
+of pairs (one pair per line).  The first element of the pair specifies
+a pipe name, and the second, its maximum depth.   The default depth
+of any pipe is $1$.
 \item {\bf llvm2aa} uses the LLVM compiler

### 2011-10-15 Documents/AhirV2/AhirV2.tex
Tested a simple example to validate new initialization infrastructure.

@@
 
-This tool takes a list of \Aa files, elaborates the program,
+This linker tool takes a list of \Aa files, elaborates the program,
+creates a global storage initializer, and
 does memory space decomposition.  The externally visible memory space is
@@
 declared as an error!
+
+If the programs being linked contain memory initialization
+routines, the linker generates a global storage initialization
+function which is named {\bf global\_storage\_initializer\_}.
+This global initializer calls all the memory initializers in
+the programs being linked.
 

### 2011-10-15 v2/libAhirV2/src/vcSystem.cpp

@@ -520,6 +520,7 @@ void vcSystem::Print_VHDL_Vhpi_Test_Bench(ostream& ofile)
 	  ofile << arg_name << " <= Unpack_String(val_string," << w->Get_Size() << ");" << endl;
 	}
 
+      ofile << "wait for 0 ns;" << endl;
       ofile << "if " << start << " = '1' then -- {" << endl;
       ofile << "while true loop --{" << endl;
       ofile << "wait until clk = '1';" << endl;

### 2011-12-26 v2/Ahir-asm/devel/docs/AaLRM.tex
Fixed the handling of undefined memory accesses..  If Aa2VC sees a memory access which "points nowhere", it will treat the result as undefined...  that is, the access will just be ignored.
Continual reorganization.  Also changed vc2vhdl so that the number of banks in a memory space defaults to 1 (overrides through attribute settings in the vc file only).
Some other observations: performance differences
between -a and no -a are minimal..  The preferred
default case should be -q -a for min area, max
rate.

@@
 
-\section{Program Variables}
+\section{Program Variables} \label{sec:Objects}
 
@@
 
+\section{Inline and Macro Modules}
+
+By default, each distinct module in an \Aa program
+is eventually implemented as a separate VHDL entity
+by downstream AhirV2 tools (though downstream
+tools are free to instantiate multiple
+copies of this entity).
+
+However, in order to give the \Aa programmer
+more flexibility,  a module in \Aa may be marked
+to be inlined or substituted into the program
+at the point of the call.  The AhirV2 utility
+{\bf AaOpt} is responsible for this substitution.
+
+\subsection{Inlined Modules}
+
+An inlined module is substituted into the
+point of a call to it in a three-step manner.
+The input arguments are sampled at the point
+of the call, the body of
+the module is invoked and the output arguments
+are transferred.  The semantics of an inlined
+function call are defined by this three step
+procedure.  The input arguments are sampled
+at the point of the call, and the output
+arguments are updated at the termination of
+the call.
+
+For example, the following
+program 
+\begin{verbatim}
+$inline $module [add] $in (a: $uint<32> b: $uint<32>)
+    $out (c: $uint<32>) $is
+{
+    c := (a+b)
+}
+
+$module [add_check] $in (p: $uint<32> q: $uint<32>) 
+    $out (r : $uint<32>) $is
+{
+    $call add (p q) (r)
+}
+\end{verbatim}
+is translated to
+\begin{verbatim}
+$module [add_check]
+$in ( p : $uint<32>   q : $uint<32>  )
+$out ( r : $uint<32>  )
+$is
+{
+  $seriesblock[add_18] 
+  {
+    $parallelblock[InArgs] 
+    {
+      add_18_a := p
+      add_18_b := q
+    }
+    ( add_18_a => add_18_a add_18_b => add_18_b )
+    $seriesblock[body] 
+    {
+      add_18_c := (add_18_a + add_18_b)
+    }
+    (  add_18_c => add_18_c )
+    $parallelblock[OutArgs] 
+    {
+      r := add_18_c
+    }
+    ( r => r )
+  }
+  ( r => r )
+}
+\end{verbatim}
+The three step substitution of add into add\_check
+is clearly seen.
+
+\subsection{Macro Modules}
+
+A module defined as a macro (described below)
+is directly substituted at the point of a call
+to it.  That is, all references to input and
+output arguments
+are replaced by references to the corresponding
+expression passed to the called function.  
+The semantics are equivalent to a MACRO instantiation
+in programming languages such as C/C++.
+
+For example:
+\begin{verbatim}
+$macro $module [shift] $in (a: $uint<32>) 
+    $out (c: $uint<32>) $is
+{
+    c := a 
+}
+
+$pipe inpipe: $uint<32>
+$pipe midpipe: $uint<32>
+$pipe outpipe: $uint<32>
+
+$module [shift_check] $in () 
+    $out () $is
+{
+    $call shift (inpipe) (midpipe)
+    $call shift (midpipe) (outpipe)
+}
+
+\end{verbatim}
+is equivalent to 
+\begin{verbatim}
+$pipe  inpipe : $uint<32>  $depth 1 
+$pipe  midpipe : $uint<32>  $depth 1 
+$pipe  outpipe : $uint<32>  $depth 1 
+$module [shift_check]
+$in ()
+$out ()
+$is
+{
+  $seriesblock[shift_14] 
+  {
+    midpipe := inpipe
+  }
+  $seriesblock[shift_17] 
+  {
+    outpipe := midpipe
+  }
+}
+\end{verbatim}
+The two calls to shift are replaced by
+series blocks with input/output arguments
+replaced by the expressions passed to them
+at the point of the call.
+
+Since it is possible to pass a pipe as an 
+argument to a call, a macro-module is permitted
+to violate the rule that there should
+be at most one write to an interface object
+within a module.
+
 \section{Foreign Modules}

### 2011-12-26 vhdl/ahir-obsolete/docs/OperatorSpecs.tex
Continual reorganization.  Also changed vc2vhdl so that the number of banks in a memory space defaults to 1 (overrides through attribute settings in the vc file only).
Some other observations: performance differences
between -a and no -a are minimal..  The preferred
default case should be -q -a for min area, max
rate.

\documentclass{beamer}
\usepackage{epsfig}
\usepackage{alltt}
\begin{document}

\frame[containsverbatim]{\frametitle{Datatypes: type boolean}
We will use the  VHDL boolean type to model control
signals.  The following array type will also be
needed
\begin{verbatim}
type BooleanArray is array(integer range <>) of boolean;
\end{verbatim}
}

\frame[containsverbatim]{\frametitle{Datatypes: type natural, integer}
We will use the VHDL {\bf integer} and {\bf natural} types
to model generics.  The following array types will also
be needed
\begin{verbatim}
type IntegerArray is array(integer range <>) of integer;
type NaturalArray is array(integer range <>) of natural;
\end{verbatim}
}

\frame[containsverbatim]{\frametitle{Datatypes:  2D std\_logic array {\bf StdLogicArray2D}}
A two dimensional array of std\_logic, defined as
\begin{verbatim}
type StdLogicArray2D 
  is 
  array(integer range <>, integer range <>) 
  of std_logic;
\end{verbatim}
}

\frame[containsverbatim]{\frametitle{Datatypes: type {\bf ApInt}}
The LLVM arbitrary precision integer. 
The VHDL equivalent is
\begin{verbatim}
type ApInt is array(integer range <>) of std_logic;
\end{verbatim}
Object declarations will fix the width.  For example
\begin{verbatim}
signal a: ApInt(5 downto 0);
\end{verbatim}
}

\frame[containsverbatim]{\frametitle{Datatypes: type {\bf ApIntArray}}
An array of ApInt.  The VHDL equivalent is
\begin{verbatim}
type ApIntArray is 
  array(integer range <>, integer range <>) 
  of std_logic;
  -- 8 ApInts, each of width 32.
signal a: ApIntArray(0 to 7, 31 downto 0);
\end{verbatim}
The range for the second dimension will always be of the type
"high downto low".
}

\frame[containsverbatim]{\frametitle{Functions: extract/insert on {\bf ApIntArray}}

The following extract and insert functions will be used to
access elements of the array.
\begin{verbatim}
function 
  Extract (constant x: ApIntArray; 
           constant idx: integer)
  return ApInt;
procedure
  Insert (x: out ApIntArray; 
          y: in ApInt; 
          idx: in integer);
\end{verbatim}

The parameter idx must be an element of \begin{verbatim} x'range(1). \end{verbatim}

}

\frame[containsverbatim]{\frametitle{Operators: conversion operators on {\bf ApInt}}
\begin{verbatim}
component ApInt_S_1
             port (x: in ApIntArray; 
                   clk,reset: in std_logic; 
                   req: in BooleanArray;
                   ack: out booleanArray;
                   z: out ApIntArray);
end component;
\end{verbatim}
This is the component corresponding
to single input shared operations.
The component can be mapped using the following
configurations:
\begin{verbatim}
ApIntToApIntSigned, ApIntToApIntUnsigned, ApIntNot.
\end{verbatim}
}

\frame[containsverbatim]{\frametitle{Operators: {\bf ApInt} arithmetic/logical operators }
\begin{verbatim}
component ApInt_S_2
	     generic (colouring: NaturalArray);
             port (x,y: in ApIntArray; 
                   clk,reset: in std_logic; 
                   reqR,reqC: in BooleanArray;
                   ackR,ackC: out BooleanArray;
                   z: out ApIntArray);
end component;
\end{verbatim}
This is the component corresponding
to two input shared operations.
The following configurations can be applied to this component:
\begin{verbatim}
ApIntAdd, ApIntSub, ApIntAnd, ApIntOr, ApIntXor, ApIntMul,
ApIntSHL, ApIntLSHR, ApIntASHR (sign bit extension)
ApIntEq, ApIntNe, ApIntUgt, ApIntUge, ApIntUlt,
ApIntUle, ApIntSgt, ApIntSge, ApIntSlt, ApIntSle.
\end{verbatim}
For the shift operators, the argument y is interpreted as
the shift-amount (assumed to be unsigned!).
}

\frame[containsverbatim]{Operators: \frametitle{{\bf ApInt} binary operators with one constant operand }
If the second argument is statically known, then the following component should be
used.
\begin{verbatim}
component ApInt_S_2_C
                generic (y: ApInt; colouring : NaturalArray);
                port ( x: in ApIntArray; 
                   clk,reset: in std_logic; 
                   reqR,reqC: in BooleanArray;
                   ackR,ackC: out BooleanArray;
                   z: out ApIntArray);
end component;
\end{verbatim}
The following configurations are similarly defined for
this component:
\begin{verbatim}
ApIntAddC, ApIntSubC, ApIntAndC, ApIntOrC, ApIntXorC, 
ApIntMulC, ApIntSHLC, ApIntLSHRC, ApIntASHRC ApIntEqC, 
ApIntNeC, ApIntUgtC, ApIntUgeC, ApIntUltC,ApIntUleC, 
ApIntSgtC, ApIntSgeC, ApIntSltC,ApIntSleC.
\end{verbatim}
}

\frame[containsverbatim]{Operators: \frametitle{Operators: {\bf ApInt} multiplexor operators }
\begin{verbatim}
component ApIntPhi 
                   port(x,y: in  ApInt; 
                        reqx,reqy: in boolean;
                        z: out ApInt; 
                        ack: out boolean;
                        clk,reset: in bit);
end component;
component ApIntSelect 
                   port(x,y,sel: in  ApInt; 
                        req: in boolean;
                        z: out ApInt; 
                        ack: out boolean;
                        clk,reset: in bit);
end component;
\end{verbatim}
These operators are special, and are not shared.
}

\frame[containsverbatim]{\frametitle{Operators: {\bf ApInt} branch operator}
\begin{verbatim}
component ApIntBranch 
             port (condition: in ApInt; 
                   clk,reset: in std_logic; 
                   req: in Boolean;
                   ack0: out Boolean;
                   ack1: out Boolean);
end component;
\end{verbatim}
This operator is special and is not shared.
}

\frame[containsverbatim]{\frametitle{Operators: load components {\bf ApLoadReq} }
\begin{verbatim}
component ApLoadReq
          generic(width: NaturalArray; 
                  colouring : NaturalArray;
                  suppress_immediate_ack: BooleanArray);
          port(addr: in StdLogicArray2D;
               req: in BooleanArray;
               ack: out BooleanArray;
               mreq: out std_logic_vector;
               mack: in std_logic_vector;
               maddr: out std_logic_vector;
               mtag: out std_logic_vector;
               clk,reset: in std_logic);
end component;
\end{verbatim}
A shared Load-Request operator.  The generic colouring is used
to implement interlock (a tag output is generated to
identify requester if relevant).  
 If suppress\_immediate\_ack(I) 
is true, then requester I will not be acked immediately even if
it is possible to do so.
}

\frame[containsverbatim]{\frametitle{Operators: load components {\bf ApLoadComplete}}
\begin{verbatim}
component ApLoadComplete
          generic(width: NaturalArray; 
                  colouring : NaturalArray);
          port(req: in BooleanArray;
               data: out StdLogicArray2D;
               ack: out BooleanArray;
               mreq: out std_logic_vector;
               mack: in std_logic_vector;
               mdata: in std_logic_vector;
               mtag : in std_logic_vector;
               clk,reset: in std_logic);
end component;
\end{verbatim}
A shared Load-Complete operator.  The generic colouring is used
to implement interlock (together with the tag input). 
}

\frame[containsverbatim]{\frametitle{Operators: store components {\bf ApStoreReq} }
\begin{verbatim}
component ApStoreReq 
          generic(width: NaturalArray; 
                  colouring: NaturalArray;
                  suppress_immediate_ack: BooleanArray);
          port(addr,data: in StdLogicArray2D;
               req: in BooleanArray;
               ack: out BooleanArray;
               mreq: out std_logic_vector;
               mack: in std_logic_vector;
               maddr,
                mdata: out std_logic_vector;
               mtag: out std_logic_vector;
               clk,reset: in std_logic);
end component;
\end{verbatim}
A shared Store-Request operator.  The generic colouring is used
to implement interlock (a tag output is generated to
identify requester if relevant). 
 If suppress\_immediate\_ack(I) 
is true, then requester I will not be acked immediately even if
it is possible to do so.
}

\frame[containsverbatim]{\frametitle{Operators: store components {\bf ApStoreComplete} }
\begin{verbatim}
component ApStoreComplete 
          generic(width: NaturalArray; 
                  colouring : NaturalArray);
          port(req: in BooleanArray;
               ack: out BooleanArray;
               mreq: out std_logic_vector;
               mack: in std_logic_vector;
               mtag : in std_logic_vector;
               clk,reset: in std_logic);
end component;
\end{verbatim}
A shared Store-Complete operator.  The generic colouring is used
to implement interlock (together with the tag input). 
}

\frame[containsverbatim]{\frametitle{Datatypes:  type {\bf ApFloat}}
The LLVM float type has two primary parameters: the length of
the characteristic and the length of the mantissa.
The overall number is signbit-characteristic-mantissa.
The equivalent VHDL type is 
\begin{verbatim}
type ApFloat is array(integer range <>)
        of std_logic;
\end{verbatim}
Thus, a signal a with a characteristic of 5 bits
and a mantissa of 3 bits will be represented as
\begin{verbatim}
signal a: ApFloat(5 downto -3);
-- a(5) is the sign,
-- a(4 downto 0) is the characteristic
-- a(-1 downto -3) is the mantissa.
\end{verbatim}
Operations on the ApFloat follow the LLVM spec wherever applicable,
and the IEEE spec where the LLVM spec is silent.
}

\frame[containsverbatim]{\frametitle{Datatypes: type {\bf ApFloatArray}}
An array of floats.
\begin{verbatim}
type ApFloatArray is 
  array(integer range <>, integer range <>) 
  of std_logic;
signal a: ApFloatArray(0 to 7, 5 downto -3);
  -- 8 ApFloats, each with
  -- characteristic of 5 bits, 
  -- mantissa of 3 bits.
\end{verbatim}
}
\frame[containsverbatim]{\frametitle{Functions: {\bf ApFloatArray} access functions }
The following extract and insert functions will be used to access
elements of the array.
\begin{verbatim}
function
   Extract (constant x: ApFloatArray;
              constant idx: integer)
   return ApFloat;
procedure
   Insert (x: out ApFloatArray;
            y: in ApFloat;
            idx: in integer);

\end{verbatim}
The parameter idx must be an element of \begin{verbatim} x'range(1). \end{verbatim}
}

\frame[containsverbatim]{\frametitle{Operators: {\bf ApFloat $\leftrightarrow$ ApInt} conversion operators }
\begin{verbatim}
component ApIntUnsignedToApFloat
             generic(colouring: NaturalArray);
             port (x: in ApIntArray;
                   clk,reset: in std_logic; 
                   reqR,reqC: in BooleanArray;
                   ackR,ackC: out BooleanArray;
                   z: out ApFloatArray);
end component;
\end{verbatim}
The operator  ApIntSignedToApFloat is similar to ApIntUnsignedToApFloat
}

\frame[containsverbatim]{\frametitle{Operators: {\bf ApFloat $\leftrightarrow$ ApInt} conversion operators }
\begin{verbatim}
component ApFloatToApIntUnsigned 
             generic(colouring: NaturalArray);
             port (x: in ApFloatArray; 
                   clk,reset: in std_logic; 
                   reqR,reqC: in BooleanArray;
                   ackR,ackC: out BooleanArray;
                   z: out ApIntArray);
end component;
\end{verbatim}
The operator  ApFloatToApIntSigned is similar to ApFloatToApIntUnsigned.
}
\frame[containsverbatim]{\frametitle{Operators: {\bf ApFloat} conversion operators }

\begin{verbatim}
component ApFloat_S_1 
             generic(colouring: NaturalArray);
             port (x: in ApFloatArray;
                   clk,reset: in std_logic; 
                   reqR,reqC: in BooleanArray;
                   ackR,ackC: out BooleanArray;
                   z: out ApFloatArray);
end component;
\end{verbatim}
The following configuration can be applied to
ApFloat\_S\_1.
\begin{verbatim}
ApFloatResize
\end{verbatim}
}

\frame[containsverbatim]{\frametitle{Operators: {\bf ApFloat} arithmetic operators }
\begin{verbatim}
component ApFloat_S_2
             generic(colouring: NaturalArray);
             port (x,y: in ApFloatArray; 
                   clk,reset: in std_logic; 
                   reqR,reqC: in BooleanArray;
                   ackR,ackC: out BooleanArray;
                   z: out ApFloatArray);
end component;
\end{verbatim}
The following configurations can be applied:
\begin{verbatim}
ApFloatAdd, ApFloatSub, ApFloatMul.
\end{verbatim}
}
\frame[containsverbatim]{\frametitle{Operators: {\bf ApFloat} comparison operators}
\begin{verbatim}
component ApFloat_Cmp_2
             generic(colouring: NaturalArray);
             port (x,y: in ApFloatArray; 
                   clk,reset: in std_logic; 
                   reqR,reqC: in BooleanArray;
                   ackR,ackC: out BooleanArray;
                   z: out ApIntArray);
end component;
\end{verbatim}
The following configurations can be applied:
\begin{verbatim}
ApFloatOeq, ApFloatOne, ApFloatOrd 
ApFloatUeq, ApFloatUne, ApFloatUno
ApFloatOgt, ApFloatOge, ApFloatOlt,ApFloatOle
ApFloatUgt, ApFloatUge, ApFloatUlt,ApFloatUle.
\end{verbatim}
}

\frame[containsverbatim]{\frametitle{Operators: {\bf ApFloat} arithmetic operators with 
one constant operand}
\begin{verbatim}
component ApFloat_S_2_C
             generic (y: ApFloatArray; colouring: NaturalArray);
             port (x: in ApFloatArray; 
                   clk,reset: in std_logic; 
                   reqR,reqC: in BooleanArray;
                   ackR,ackC: out BooleanArray;
                   z: out ApFloatArray);
end component;
\end{verbatim}
The following configurations can be applied to ApFloat\_S\_2\_C:
\begin{verbatim}
ApFloatAddC, ApFloatSubC, ApFloatMulC.
\end{verbatim}
}

\frame[containsverbatim]{\frametitle{Operators: {\bf ApFloat} comparison operators with 
one constant operand}
\begin{verbatim}
component ApFloat_Cmp_2_C
          generic (y: ApFloatArray; 
                   colouring: NaturalArray);
          port    (x: in ApFloatArray; 
                   clk,reset: in std_logic; 
                   reqR,reqC: in BooleanArray;
                   ackR,ackC: out BooleanArray;
                   z: out ApIntArray);
end component;
\end{verbatim}
The following configurations can be used:
\begin{verbatim}
ApFloatOeqC, ApFloatOneC, ApFloatOrdC ApFloatUeqC, 
ApFloatUneC, ApFloatUnoC ApFloatOgtC, ApFloatOgeC, 
ApFloatOltC, ApFloatOleC ApFloatUgtC, ApFloatUgeC, 
ApFloatUltC, ApFloatUleC.
\end{verbatim}
}

\frame[containsverbatim]{\frametitle{Operators: {\bf ApFloat} multiplexor operators }
\begin{verbatim}
component ApFloatPhi 
                   port(x,y: in  ApFloat; 
                        reqx,reqy: in boolean;
                        z: out ApFloat; 
                        ack: out boolean;
                        clk,reset: in bit);
end component;
component ApFloatSelect 
                   port(x,y: in  ApFloat; 
                        sel: in ApInt;
                        req: in boolean;
                        z: out ApFloat; 
                        ack: out boolean;
                        clk,reset: in bit);
end component;
\end{verbatim}
These are special, unshared operators.
}

\begin{frame}[fragile]
  \frametitle{Input Port}
\begin{alltt}
component InputPort
  generic (colouring: NaturalArray);
  port (
    -- pulse interface with the data-path
    req       : in  BooleanArray;
    ack       : out BooleanArray;
    data      : out StdLogicArray2D;

    -- ready/ready interface with outside world
    oreq       : out std_logic;
    oack       : in  std_logic;
    odata      : in  std_logic_vector;

    clk, reset : in  std_logic);
end component;
\end{alltt}
If colouring indicates that mutual 
exclusion does not hold, then {\bf one} of the requesters
will succeed.
\end{frame}

\begin{frame}[fragile]
  \frametitle{Output Port}
\begin{alltt}
component OutputPort
  generic (colouring: NaturalArray);
  port (
    req       : in  BooleanArray;
    ack       : out BooleanArray;
    data      : in  StdLogicArray2D;
    oreq       : out std_logic;
    oack       : in  std_logic;
    odata      : out std_logic_vector;
    clk, reset : in  std_logic);
end component;
\end{alltt}
If colouring indicates that mutual 
exclusion does not hold, then {\bf one} of the requesters
will succeed.
\end{frame}

\begin{frame}
  \frametitle{Function Calls}
  \begin{itemize}
  \item {\bf Call} made to another module.
    \begin{itemize}
    \item Use OutputPort.
      \begin{itemize}
      \item Data port is  an std\_logic\_vector consisting of
        arguments and address for the called module.
      \end{itemize}
    \end{itemize}
  \item {\bf Response} received for the corresponding call.
    \begin{itemize}
      \item Use InputPort.
        \begin{itemize}
        \item Return-value packed in an std\_logic\_vector.
        \end{itemize}
    \end{itemize}
  \item {\bf Accept} an incoming call.
    \begin{itemize}
    \item Use InputPort.
      \begin{itemize}
      \item Arguments packed in an std\_logic\_vector.
      \end{itemize}
    \end{itemize}
  \item {\bf Return} an incoming call.
    \begin{itemize}
    \item Use OutputPort.
      \begin{itemize}
      \item Return value packed in an std\_logic\_vector.
      \end{itemize}
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Arbiter for function calls}
\small
\begin{alltt}
entity CallArbiter is
  port ( -- ready/ready handshake on all ports
    -- ports for the caller
    call_reqs   : in  std_logic_vector;
    call_acks   : out std_logic_vector;
    call_data   : in  StdLogicArray2D;
    -- call port connected to the called module
    call_mreq   : out std_logic;
    call_mack   : in  std_logic;
    call_mdata  : out std_logic_vector;
    call_mtag   : out std_logic_vector;
    -- similarly for return, initiated by the caller
    return_reqs : in  std_logic_vector;
    return_acks : out std_logic_vector;
    return_data : out StdLogicArray2D;
    -- return from function 
    return_mreq : in std_logic;
    return_mack : out std_logic;
    return_mdata : in  std_logic_vector;
    return_mtag : in  std_logic_vector);
end CallArbiter;
\end{alltt}
\end{frame}

\end{document}

% LocalWords:  StdLogicArray ApIntArray ApFloatArray ApInt ApFloat downto ireq
% LocalWords:  InputPort iack idata oreq oack odata OutputPort addr mreq maddr
% LocalWords:  mtag CallRequest CallComplete IncomingCallRequest
% LocalWords:  IncomingCallComplete
\end{document}

### 2011-12-26 v2/Ahir-asm/devel/include/AaObject.h

@@ -106,6 +106,7 @@ class AaInterfaceObject: public AaObject
 
   AaInterfaceObject(AaScope* scope_tpr, string oname, AaType* otype, string mode);
   ~AaInterfaceObject();
+  virtual string Get_Name();
   virtual string Kind() {return("AaInterfaceObject");}
   virtual bool Is_Interface_Object() {return(true);}
   // uses AaObject::Print method

### 2011-12-26 v2/Ahir-asm/devel/util/Aa2VC.cpp

@@ -74,6 +74,8 @@ int main(int argc, char* argv[])
     };
 
 
+  AaProgram::_tool_name = "Aa2VC";
+
   string fname;
   string mod_name;
   string opt_string;

### 2011-12-26 v2/libAhirV2/util/vcAnalyze.cpp

@@ -1,0 +1,148 @@
+#include <signal.h>
+#include <vcSystem.hpp>
+#include <vcParser.hpp>
+#include <vcLexer.hpp>
+
+using namespace std;
+
+void Handle_Segfault(int signal)
+{
+  vcSystem::Error("in vcAnalyze: segmentation fault! giving up!!");
+  exit(-1);
+}
+
+void Usage_VcAnalyze()
+{
+
+  cerr << "brief description: analyzes specified VC files,  " << endl;
+  cerr << "Usage: " << endl;
+  cerr << "vcAnalyze ... options ... " << endl;
+  cerr << " options " << endl;
+  cerr <<  " -t mod-name: will make mod-name a top-level module which must" << endl
+       <<  "              be started from outside the AHIR system.  Such a module can " << endl
+       <<  "              have input/output arguments from/to the outside world" << endl
+       <<  " Notes: " << endl
+       <<  "   1. alternate form: --top mod-name" << endl
+       <<  "   2. option can be repeated to specify multiple top-level modules" << endl;
+  cerr << endl;  
+  cerr <<  " -f vcfile-name: specifies a VC file to be parsed.  Repetitions of this option can be" << endl
+       <<  "                 used to specify more than one file.  Files are parsed in the order specified." << endl
+       <<  " alternate form:  --file vcfile-name" << endl;
+  cerr << endl;
+  cerr << "example: " << endl
+       << "    vcAnalyze  -t foo -f file1.vc -f file2.vc -t bar" << endl;
+  cerr << "file1.vc and file2.vc will be parsed in order, and the instantiated " << endl
+       << "system will have modules foo and bar as top-modules, which are controllable  " << endl
+       << "from outside the system. vcAnalyze will make certain checks on these files.." << endl;
+}
+
+
+int main(int argc, char* argv[])
+{
+  
+
+  signal(SIGSEGV, Handle_Segfault);
+
+  if(argc < 2)
+    {
+	Usage_VcAnalyze();
+      	exit(1);
+    }
+
+  // command-line parsing
+  struct option long_options[] =
+    {
+      /* These options set a flag. */
+      {"help", no_argument, 0, 'h'},
+      {"set_as_top",  required_argument, 0, 't'},
+      {"vcfile",    required_argument, 0, 'f'},
+      {0, 0, 0, 0}
+    };
+
+
[…]

### 2011-12-26 v2/libAhirV2/util/vcCPCheck.cpp

@@ -1,49 +1,0 @@
-#include <signal.h>
-#include <vcSystem.hpp>
-#include <vcParser.hpp>
-#include <vcLexer.hpp>
-
-using namespace std;
-
-void Handle_Segfault(int signal)
-{
-  vcSystem::Error("in vcAnalyze: segmentation fault! giving up!!");
-  exit(-1);
-}
-
-int main(int argc, char* argv[])
-{
-  
-  string sys_name = "test_system";
-  vcSystem test_system(sys_name);
-
-  signal(SIGSEGV, Handle_Segfault);
-
-  if(argc < 2)
-    {
-      cerr << "Usage: vcAnalyze <filename> <filename> ... " << endl;
-
-      exit(1);
-    }
-
-
-  for(int idx = 1; idx < argc; idx++)
-    {
-
-      string filename = argv[idx];
-      
-      test_system.Parse(filename);
-      
-      if(vcSystem::Get_Error_Flag())
-	{
-	  cerr << "Error: there were errors during parsing, check the log" << endl;
-	  return(1);
-	}
-    }
-
-  test_system.Elaborate();
-  test_system.Print_Control_Structure(cout);
-  return(0);
-}
-
-

### 2012-02-20 v2/libAhirV2/docs/vC_LRM.tex
several modifications.   - added a pipeHandler which uses sockets (as a replacement for     named pipes, which were a pain to debug).  pipeHandler is     kept in Ctestbench and can be used to write pthreads apps     to check your program before going to hardware     (plan:  move to gnu-pth to make things easier to debug and      used message pipes instead of sockets).
  - some additions to Aa.  allow constant references in
    switch alternatives, hex constants.

  - got rid of escaped identifiers from ieee trimmed vhdl
    library.  some simulators (note: IMAGE) dont like it.

\documentclass{article}

\title{vC Language Reference Manual}
\author{Madhav Desai \\ Department of Electrical Engineering \\ Indian Institute of Technology \\
	Mumbai 400076 India}

\newcommand{\Aa}{{\bf Aa}~}
\newcommand{\vC}{{\bf vC}~}

\begin{document}
\maketitle

\section{Introduction}

{\bf vC} is a language used for the description
of a virtual circuit which is decomposed into
control, data and storage aspects.

A \vC system consists of modules and memory spaces.
Each module may itself contain memory spaces.  Further,
each module has a {\em control-path} and a {\em data-path}.
The control-path is a Petri-net constructed using
a set of rules (as describe in \cite{ref:SameerPhD}).1
The data-path is essentially a directed graph of {\em operator } nodes
communicating with each other and the outside world using {\em wires}.

A memory-space stores data and provides access to this
data through load/store operators.
The \vC system also  contains {\em pipes} which
are first-in-first out queues. These pipes can
be accessed from a data-path using input/output
operators.

In the rest of this document, we outline the
structure and the syntax of the \vC language,
and also describe the semantics of a \vC
program, especially the execution model and
the behaviour of the equivalent system.

\section{Program structure}

A program (equivalently, a system) in \vC consists of a sequence of
declarations and module definitions.  
\begin{verbatim}
program := ( module-definition | 
              memory-space-definition | 
                pipe-declaration | 
                  attribute-specification )*
\end{verbatim}

A module in \vC is the basic unit of compilation, and
has the following structure
\begin{verbatim}
$module [module-name] 
{
   input-arguments (can be empty)
   output-arguments (can be empty)
   memory-space-definitions (optional)
   pipe-declarations (optional)
   control-path (required)
   data-path (required)
   control-data-links (optional)
   attribute-specifications (optional)
}
\end{verbatim}

Thus, a module has a name, has input and output arguments,
pipes, memory spaces, a control-path and a data path.
Pipes declared in the module are visible only in the
module body.

\section{Variables} \label{sec:Objects}

Variables in  a \vC system can belong to the following classes:
\begin{itemize}
\item {\bf Argument} variable declared at the module-level.
\item {\bf Wire} variables declared in the data-path.
\item {\bf Pipe} variables declared at the program level or at the module level.
\end{itemize}

Each variable has a type.  In \vC a fairly rich array of
types is supported, as described in Section \ref{sec:Types}.

{\bf Wire} variables are declared in the data-path, and
serve as connectors between operators.  A wire can have
at most one operator node which drives it, but can
be read by any number of operators.

{\bf Pipe} variables declared at the program scope
are globally visible first-in-first-out (FIFO) buffers that can 
be accessed from any module in the program.  A pipe
is a FIFO buffer with a specified number of
slots.  A write to a pipe succeeds
only if there is an empty slot in the FIFO (a successful
write uses up an empty slots.  A read
from the pipe succeeds only if there is at
least one occupied slot in the FIFO (a successful read
creates an empty slot in the FIFO).

\section{The control-path} \label{sec:Controlpath}

The control-path is a Petri-net which consists of
places and transitions.  In \vC this Petri-net
is described hierarchically in terms of 
control-path elements. 
\begin{itemize}
\item A single transition is a control-path element. It
fires when all of its predecessors have (released) a token.
After it fires, a token is released to each of
its successors.
\begin{verbatim}
$T [fire]
\end{verbatim}
This defines a transition with name $fire$. 
\item A single place is a control-path element.  When
any of its predecessors releases a token, the place
takes it and forwards it to one of its successors.
\begin{verbatim}
$P [loopback]
\end{verbatim}
\item A {\em series region} is a sequence of control
path elements.  The predecessor of each element is
the previous element and the successor is the next
element in the sequence.  
The predecessor of
the first element is an implicit entry transition
of the series region, and the successor of the last
element is an implicit exit transition of
the series region.
\begin{verbatim}
;;[chain] {
       $T [req]
       $T [ack]
}
\end{verbatim}
\item A {\em parallel region} is a sequence of control
path elements.  The implicit entry transition
is the predecessor of all the elements, while the
implicit exit transition is the successor of
all the elements.
\begin{verbatim}
||[pardo] {
       ;;[job_1] { ... }
       ;;[job_2] { ... }
}
\end{verbatim}
\item A {\em branch region} is a sequence of control
path elements with additional statements which describe
the connections between this sequence (branches, merges).
\begin{verbatim}
<>[branch_example] {
       $P [Enter]
       Enter <-| ($entry)
       Enter |-> (job_1)
       ;;[job_1] { ... }
       ;;[job_2] { ... }
       $P [Chk]
       $P [End]
       Chk <-| (job_1)
       Chk |-> (job_1 job_2)
       End <-| (job_2)
       End |-> ($exit)
}
\end{verbatim}
In this example, the keywords \$entry and \$exit correspond
to implicit entry and exit transitions associated with the branch block.
The $<-|$ specifies that a token leaving the right will enter
the left (place) and $|->$ signifies that a token entering the left
(place) will enter one of the elements in the list on the right.
\item A {\em fork region} is a sequence of control path
elements with additional statements which describe 
forking and joining.
\begin{verbatim}
::[fork_example] {
       $entry &-> (job_1 job_2)
       ;;[job_1] { ... }
       ;;[job_2] { ... }
       ;;[job_3] { ... }
       $T [f12]
       f12 <-& (job_1 job_2)
       f12 &-> (job_3)
       $exit <-& (job_3)
}
\end{verbatim}
In this example, the keywords \$entry and \$exit correspond
to implicit transitions associated with the branch block.
The <-\& specifies that after a token leaves from every
element in the right list, a token will enter
the left (transition) and \&-> signifies that a token leaving the left
(transition) will be replicated so that a
token enters each of the elements in the list on the right.
\end{itemize}

The control-path itself is a series region. It syntax is
illustrated below:
\begin{verbatim}
$CP {
   ... sequence of control-path elements --
}
\end{verbatim}

\section{Types} \label{sec:Types}

Types in \vC can be either scalar types or composite types.  They
are always attached to objects such as pipes, wires and input/output
arguments (types are not declared separately!).

Scalar types can be one of 
\begin{itemize}
\item Integers:  An integer type
has a width parameter and is specified as 
\begin{verbatim}
$int<width>
\end{verbatim}
The width parameter can be any positive number.
Values corresponding to this type are to be
viewed as a binary sequence of the specified width.
\item Pointers:  A pointer is an unsigned
integer with an unspecified width which
is associated with a memory space.
The width of the pointer is determined by the
implementation at the point of conversion from
\vC to VHDL.  For example:
\begin{verbatim}
$pointer< memory_space_1 > 
\end{verbatim}
\item Floats: A float is parametrized by
two integers, the width of the exponent,
and the width of the mantissa.  
The specification is
\begin{verbatim}
$float<exponent,mantissa>
\end{verbatim} %$
where the exponent and mantissa must be positive
integers.  The float is represented by 
a word with $exponent+mantissa+1$ bits
(with the additional bit needed for the sign).
The standard IEEE 754 float and double
precision representations correspond
to
\begin{verbatim}
$float<8,23> 
\end{verbatim}
and
\begin{verbatim}
$float<11,52> 
\end{verbatim}
respectively.  Currently, these are the
only float types that are supported.
\end{itemize}

Composite types in \vC can be either
array types or record types.
\begin{itemize}
\item 
Array types in \vC have the form
\begin{verbatim}
$array [d1] $of <element-type-spec>
\end{verbatim} %$
The value of $d1$ must
be a positive integers, and element-type-spec
must refer to a type.
For example, 
\begin{verbatim}
$array [10] of $array [10] $of $int<32>
\end{verbatim}
is an array type whose
elements are one dimensional arrays of 32-bit
unsigned integers.
\item
Record types in \vC have the form
\begin{verbatim}
$record <type-1> <type-2> ... <type-n>
\end{verbatim}
An element of such a record type is
an aggregate whose first element is of type
type-1, second element is of type type-2 and so on.
For example:
\begin{verbatim}
$record <$uint<32> > <$array [10] $of $uint<32> > 
\end{verbatim}
\end{itemize}

\section{The Data-path} \label{sec:Datapath}

The Data-path is a directed graph of operator nodes
interconnected with wires.
A wire is declared as
\begin{verbatim}
$W [wname] : <wire-type>
\end{verbatim}
For example:
\begin{verbatim}
$W [reg_out] : $int<33>
\end{verbatim}

One can also have constant wires which are declared
as
\begin{verbatim}
$constant $W [wname] : <wire-type> := <wire-value>
\end{verbatim}
where wire-value is a constant literal.  A constant wire can only have 
scalar type.

Operator nodes can be differentiated based on two criteria.
Based on the number of inputs, they can be classified
as unary, binary, variable argument and mutliple-output
operators.  Based on their protocol of interaction with
the control path, they can be classified as those
with a split-protocol (separate start-complete request/ack
handshakes) or a unitary protocol (single request/ack handshake).

Operator nodes are instantiated by the syntax
\begin{verbatim}
<op-code> [operator-name] ( <input-wires> )  ( <output-wires>)
\end{verbatim}
For example, an add operator node instantiation is 
\begin{verbatim}
+ [wadd] (a b)  (c)
\end{verbatim}
The add operation takes two inputs and produces a single output.

Other binary operators with two inputs and a single
output are (all binary operators have a split protocol): 
\begin{verbatim}
op-code operation             
-       subtract             
*       multiply            
/       divide             
<<      shift-left        
>>      shift-right-logical 
$S>>    shift-right-arithmetic
==      equal              
$S>$S   signed greater-than
$S>=$S  signed greater-than-or-equal
$S<$S   signed less-than
$S<=$S  signed less-than-or-equal
>       unsigned greater-than
>=      unsigned greater-than-or-equal
<       unsigned less-than
<=      unsigned less-than-or-equal
!=      not-equal
|       logical-bitwise-or
&       logical-bitwise-and
^       logical-bitwise-xor
~|      logical-bitwise-nor
~&      logical-bitwise-nand
~^      logical-bitwise-xnor
\end{verbatim}

Operators with a single input and a single
output are (all have a split-protocol):
\begin{verbatim}
~       logical-not
:=      assign-operation
$S:=$U  signed-to-unsigned-conversion
$U:=$S  unsigned-to-signed-conversion
$S:=$S  signed-to-signed-conversion
$S:=$F  signed-to-floating-point-conversion
$F:=$S  floating-point-to-signed-conversion
$U:=$F  unsigned-to-floating-point-conversion
$F:=$U  floating-point-to-unsigned-conversion
$F:=$F  float-to-float-conversion
\end{verbatim}

There are some special operators which
need some explanation.
\begin{itemize}
\item The {\em bit-select} operator has two inputs and
one output.  The syntax is
\begin{verbatim}
[] [selbit] (a i) (b)
\end{verbatim}
Wire b will get the value of the $i^{th}$ most significant bit
of a.  This operator has a split protocol. 
\item  The {\em PHI} operator has a variable number of inputs
and one output.  The syntax is
\begin{verbatim}
$phi [oname] (u1 u2 u3 .. un) (y)
\end{verbatim}
This operator has a unitary protocol, but there are as many
req/ack pairs as there are inputs.  The output y
is assigned to an input for which a request has been made.
\item The {\em concatenate} operator has two inputs and a single
output.  The output bit sequence is a conctatenation of the input
bit sequences.
\begin{verbatim}
&& [cname] (a b) (c)
\end{verbatim}
This operator also has a split protocol.
\item The {\em slice} operator has three inputs, the first
of which is a wire, and the next two must be constants.
It has a single output:
\begin{verbatim}
[:] [cname] (a 5 3) (c)
\end{verbatim}
Then c = a[5 downto 3].
\end{itemize}

Finally, there are some special operators, all
of which have split protocols.
\begin{itemize}
\item The {\em load} and {\em store} operators are used to access
a memory space.
\begin{verbatim}
$load [fetch1] $from memory_space_1 (addr_wire) (data_wire)
$store [write1] $to memory_space_1 (addr_wire data_wire) ()
\end{verbatim}
\item The {\em input} and {\em output} operators are used
to read from and write to pipes (these have a unitary protocol).
\begin{verbatim}
$ioport $in  [R1] (pipe_to_be_read) (wire_to_be_written_into)
$ioport $out [W1] (wire_to_be_read) (pipe_to_be_written_into)
\end{verbatim}
\item The {\em call} operator is used to send arguments to
another module and wait until results are returned (it has
a split protocol).
\begin{verbatim}
$call [call_id] $module foo (in1 in2 .. inN) (out1 out2 .. outM)
\end{verbatim}
The number of input and output arguments must match those
on module foo.
\end{itemize}

\section{Control-data links}

The control path and the data path work in
a coordinated fashion.  Transitions in the control
path are associated with request and acknowledge
signals on operators in the data path.  % TODO.

\section{Memory Spaces}

Each memory space provides a certain amount of 
storage with a well defined access width.  % TODO.

\section{Syntax} \label{sec:Syntax}

The syntax for \vC follows the following
principles
\begin{itemize}
\item Most keywords begin with the \$ sign.
\item The region between \{ and \} defines a new scope.
\item Elements are space separated (no semicolons at all).
\end{itemize}

The parser is implemented using 
an LL(k) parser (written as rules to be parsed by antlr2 \cite{ref:antlr2}).
The grammar for the parser is (using the EBNF notation) given in 
the html file {\bf vcParser.html} which is part of this distribution. 
The set of tokens recognized by the lexical analyzer (or lexer).
is available in the html file {\bf vcLexer.html}.

\begin{thebibliography}{99}
\bibitem{ref:SameerPhD}
Sameer D. Sahasrabuddhe,
``A competitive pathway from high-level programs to hardware,''
Ph.D. thesis, IIT Bombay, 2009.

\bibitem{ref:antlr2}
http://www.antlr2.org.
\end{thebibliography}

\end{document}

### 2012-02-20 v2/CtestBench/include/SocketLib.h

@@ -25,7 +25,7 @@
 
 // unlikely to have more than 16 active
 // threads talking to the TB?
-#define DEFAULT_MAX_CONNECTIONS 256
+#define DEFAULT_MAX_CONNECTIONS 65535
 
 void  ADD_SPACE__(char* buf);
 

### 2012-02-20 v2/CtestBench/pipeHandlerTest/test_program.c

@@ -1,0 +1,53 @@
+#include <pthread.h>
+#include <SocketLib.h>
+#include <Pipes.h>
+#include <pipeHandler.h>
+#include <stdint.h>
+
+void* pipeHandler__()
+{
+	pipeHandler();
+}
+
+void* Sender()
+{
+	uint32_t word;
+	for(word = 0; word < 10; word++)
+	{
+		write_uint32("test_pipe",word);
+		fprintf(stdout,"Sent: %d\n", word);
+		write_float32("test_pipe",(float)word);
+		fprintf(stdout,"Sent: %f\n", (float) word);
+	}
+}
+
+void* Receiver()
+{
+	uint32_t word;
+	float fword;
+	while(1)
+	{
+		word = read_uint32("test_pipe");
+		fprintf(stdout,"Received: %d\n", word);
+		fword = read_float32("test_pipe");
+		fprintf(stdout,"Received: %f\n", fword);
+	}
+}
+
+int main(int argc, char* argv[])
+{
+	pthread_t phandler_t, sender_t, receiver_t;
+	pthread_create(&phandler_t,NULL,&pipeHandler__,NULL);
+
+	register_pipe("test_pipe",32,32);
+
+	pthread_create(&sender_t,NULL,&Sender,NULL);
+	pthread_create(&receiver_t,NULL,&Receiver,NULL);
+
+	pthread_join(sender_t,NULL);
+	usleep(1000);
+	pthread_cancel(receiver_t);
+	
+	killPipeHandler();
+	pthread_join(phandler_t,NULL);
+}

### 2012-02-20 v2/CtestBench/src/Pipes.c

@@ -3,7 +3,9 @@
 #include <SocketLib.h>
 #include <string.h>
 #include <assert.h>
+#include <SocketLib.h>
 #include <Pipes.h>
+
 
 #define READ__(id,t,w)  sprintf(buffer, "piperead.single %s ", id);	\
   append_int(buffer,0);ADD_SPACE__(buffer);\
@@ -16,6 +18,8 @@
   append_int(buffer,1);ADD_SPACE__(buffer);\
   append_##t(buffer,data);ADD_SPACE__(buffer);\
   append_int(buffer,0);\
+  ADD_SPACE__(buffer);\
+  append_int(buffer,w);\
   send_packet_and_wait_for_response(buffer,strlen(buffer)+1,"localhost",9999);
 
 
@@ -28,10 +32,26 @@
 			 bcopy(buf,buffer + strlen(buffer),(len*w/8)); \
 			 send_packet_and_wait_for_response(buffer,send_len,"localhost",9999);
 
+
+uint8_t register_pipe(const char* id, int pipe_depth, int pipe_width)
+{
+  uint8_t ret_val;
+  char buffer[MAX_BUF_SIZE];
+  char* ss;
+  sprintf(buffer,"registerpipe %s %d %d", id, pipe_depth,pipe_width);
+  send_packet_and_wait_for_response(buffer,strlen(buffer),"localhost",9999);
+#ifdef DEBUG
+  fprintf(stderr,"Info: in register_pipe, received %s\n", buffer);
+#endif
+  ret_val = get_uint8_t(buffer,&ss);
+  return(ret_val);
+  
+}
+
 uint64_t read_uint64(const char *id)
 {
   uint64_t ret_val;
-  char buffer[4096];
+  char buffer[MAX_BUF_SIZE];
   char* ss;
   READ__(id,uint64_t,64);
 #ifdef DEBUG
@@ -42,7 +62,7 @@ uint64_t read_uint64(const char *id)
 }
 void write_uint64(const char *id, uint64_t data)
 {
-  char buffer[4096];
+  char buffer[MAX_BUF_SIZE];
   char* ss;
   WRITE__(id,uint64_t,64);
 }
@@ -51,9 +71,9 @@ void write_uint64(const char *id, uint64_t data)
 // from pipe id, receive buf_len uint64_t's..
 void read_uint64_n(const char *id, uint64_t* buf, int buf_len)
 {
-  char buffer[5000];
+  char buffer[MAX_BUF_SIZE+4];
 
[…]

### 2012-02-20 v2/CtestBench/test/vhdlCStubs.c

@@ -3,10 +3,10 @@
 uint32_t foo(uint32_t a)
 {
 char buffer[1024];  char* ss;  sprintf(buffer, "call foo ");
-append_int(buffer,1);
-append_uint32_t(buffer,a);
-append_int(buffer,1);
-append_int(buffer,32);
+append_int(buffer,1);ADD_SPACE__(buffer);
+append_uint32_t(buffer,a);ADD_SPACE__(buffer);
+append_int(buffer,1);ADD_SPACE__(buffer);
+append_int(buffer,32);ADD_SPACE__(buffer);
 
 send_packet_and_wait_for_response(buffer,strlen(buffer)+1,"localhost",9999);
 uint32_t ret_val__ = get_uint32_t(buffer,&ss);

### 2012-03-04 v2/Ahir-asm/devel/docs/AaLRM.tex
Added $lifo keyword so that pipes can be configured to be either FIFO or LIFO.

@@
 points in the program.
+A storage variable declaration has the form
+\begin{verbatim}
+$storage <var-name>: <var-type> 
+\end{verbatim}
+For example
+\begin{verbatim}
+$storage mempool: $array [1024] $of $uint<32>
+\end{verbatim}
 \item {\bf Pipe} variables declared at the program scope
-are globally visible first-in-first-out (FIFO) buffers that can 
+are globally visible buffers (which
+can be defined to be either first-in-first-out (FIFO) or
+last-in-first-out (LIFO)) that can 
 be read from anywhere in the program and can be
 written into from anywhere in th program.  A pipe
-is a FIFO buffer with a specified number of
+is a FIFO (or FIFO) buffer with a specified number of
 slots.  A write to a pipe succeeds
@@
 from the pipe succeeds only if there is at
-least one occupied slot in the FIFO (a successful read
-creates an empty slot in the FIFO). 
-Thus, pipes provide a mechanism for synchronization and
+least one occupied slot in the pipe (a successful read
+creates an empty slot in the pipe). 
+Thus, pipes can be used to model queues as
+well as stacks, and also provide a mechanism for synchronization and
 for message passing in an \Aa program.
@@
 \end{itemize}
+A pipe declaration has the form
+\begin{verbatim}
+($lifo)? $pipe <pipe-name>: <pipe-type>  ($depth <integer>)?
+\end{verbatim}
+For example
+\begin{verbatim}
+$pipe myfifo: $uint<32> $depth 16 // 16-slot queue
+$lifo $pipe mystack: $uint<32> $depth 4 // 4-slot stack
+\end{verbatim}
 \item {\bf Constant} variables declared at
@@
 Thus, an assignment to a constant is an error.
+A constant declaration has the form
+\begin{verbatim}
+$constant <constant-name>: <constant-type> := <initial-value>
+\end{verbatim}
+For example
+\begin{verbatim}
+$constant wordlength: $uint<32> := 16
+\end{verbatim}
 \item {\bf Implicit} or {\bf Static Single Assignment}
@@
 read from several points in the program.
+Implicit variables are not declared, but are defined
+by the statements which write to them.
 \end{itemize}

### 2012-03-04 v2/libAhirV2/include/vcRoot.hpp

@@ -74,6 +74,7 @@ enum vcLexerKeytags
     __INLINE,
     __IOPORT,
     __PIPE,
+    __LIFO,
     __FROM,
     __AT,
     __CONSTANT,
@@ -191,6 +192,7 @@ static string vcLexerKeywords[] =
       "$inline",
       "$ioport",
       "$pipe",
+      "$lifo",
       "$from",
       "$at",
       "$constant",

### 2012-03-04 v2/libAhirV2/include/vcSystem.hpp

@@ -67,7 +67,7 @@ class vcSystem: public vcRoot
     else
       return NULL;
   }
-  void Add_Pipe(string pipe_id, int width, int depth);
+  void Add_Pipe(string pipe_id, int width, int depth, bool lifo_mode);
   int Get_Pipe_Width(string pipe_id);
   int Get_Pipe_Depth(string pipe_id);
   void Print_Pipes(ostream& ofile);

### 2012-03-04 v2/libAhirV2/src/vcDataPath.cpp

@@ -83,6 +83,7 @@ void vcPipe::Print_VHDL_Instance(ostream& ofile)
       ofile << "num_reads => " << num_reads << "," << endl;
       ofile << "num_writes => " << num_writes << "," << endl;
       ofile << "data_width => " << pipe_width << "," << endl;
+      ofile << "lifo_mode => " << (this->Get_Lifo_Mode() ? "true" : "false") << "," << endl;
       ofile << "depth => " << pipe_depth << " --}\n)" << endl;
       ofile << "port map( -- { " << endl;
       ofile << "read_req => " << pipe_id << "_pipe_read_req," << endl 

### 2012-03-04 v2/CtestBench/docs/CtestBench.tex
Added support for stacks in llvm2aa (through pipedepth files).  Also modified the Queue/SynchFifo to try to merge pipes and stacks in the same entity.. but it is getting messy and it may be best to split the two.

--- 2011-07-15 v2/CtestBench/docs/CtestBench.tex
@@
 
+\section{An example}
+
+Consider the following program (lets say it is in file ``prog.c''):
+\begin{verbatim}
+#include <stdlib.h>
+#include <stdint.h>
+#include <stdio.h>
+// this file provides definitions
+// of pipe access functions read/write.
+#include <Pipes.h>
+
+// note: initialized value..
+uint32_t sum1 = 23;
+uint32_t sum2 = 39;
+
+// note: no problems with pointers :-)
+uint32_t* tgt[2] = {&sum1, &sum2};
+
+void accumulate()
+{
+    int i = 0;
+    while(1)
+    {
+	int nxt = read_uint32("in_data");
+#ifdef SW
+	printf("read %u\n", nxt);
+#endif
+	// ugly, but this is just a demo!
+	*(tgt[i])= ((*tgt[i]) + nxt);
+
+	write_uint32("out_data",*(tgt[i]));
+#ifdef SW
+	printf("wrote %u\n", *(tgt[i]));
+#endif
+	i = 1 - i;
+    }
+}
+
+uint32_t get_sum(uint32_t idx)
+{
+    return(*(tgt[idx]));
+}
+
+\end{verbatim}
+
+This program describes a {\em system} which listens
+for data on a pipe ``in\_data'', and sends data
+out on a pipe ``out\_data''.  The incoming data
+is accumulated into the variable {\em sum}, and
+there are two methods to set and get the value
+of {\em sum}.
+
+Now to test this program, we can write a test-bench
+such as this one (lets call this file ``testbench.c'').
+\begin{verbatim}
+#include <pthread.h>
+#include <signal.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <stdint.h>
+
+#ifdef SW
+// for the pipe read/write functions
+#include <Pipes.h>
+// for the pipeHandler daemon..
+#include <pipeHandler.h>
+#include "prog.h"
+#else
+#include "vhdlCStubs.h"
+#endif
+
+void Exit(int sig)
+{
+	fprintf(stderr, "## Break! ##\n");
+	exit(0);
+}
+
+void *accumulate_(void* fargs)
+{
+   accumulate();
+}
+
+void *write_pipe_(void* a)
+{
+   write_uint32_n("in_data",(uint32_t*)a, 10);
+}
+
+void *read_pipe_(void* a)
+{
+   read_uint32_n("out_data",(uint32_t*)a, 10);
+}
+
+void* pipeHandler__(void* a)
+{
+	pipeHandler();
+}
+
+int main(int argc, char* argv[])
+{
+	signal(SIGINT,  Exit);
+  	signal(SIGTERM, Exit);
+
+	uint32_t data_in[10], data_out[10];
+        int i;
+
+#ifdef SW
+
+	// in the "software" case, we start a pipe-handler
+	// to manage the pipes..  Earlier, we were using named
+	// pipes, which were very flaky and difficult to debug.
+	// the pipeHandler generates a log file (pipeHandler.log)
+	// which is very useful for figuring out what happened.
+	// one can also use gdb to trace activity.
+	pthread_t phandler_t;
+	pthread_create(&phandler_t,NULL,&pipeHandler__,NULL);
+
+	usleep(100);
+
+	// register two FIFOs..
+	register_pipe("in_data",10,32,0);
+	register_pipe("out_data",10,32,0);
+
+#endif
+
+#ifndef SW
+	// to set the initial value of sum.
+	// in the hardware version, storage
+	// variables are initialized by calling
+	// this function (auto-generated by
+	// the Aa linker AaLinkExtMem)
+	global_storage_initializer_();
+#endif
+
+        for(i = 0; i < 10; i++)
+	{
+		data_in[i] = i;
+	}
+
+	pthread_t acc_t, wpipe_t, rpipe_t;
+
+#ifdef SW
+
+	pthread_create(&acc_t,NULL,&accumulate_,NULL);
+#endif
+	pthread_create(&wpipe_t,NULL,&write_pipe_,(void*)data_in);
+	pthread_create(&rpipe_t,NULL,&read_pipe_,(void*)data_out);
+
+	pthread_join(wpipe_t,NULL);
+	pthread_join(rpipe_t,NULL);
+
+ 	fprintf(stdout,"from out_data, we read ");
+	for(i=0; i < 10; i++)
+ 	  fprintf(stdout," %u ", data_out[i]);
+	fprintf(stdout,"\n");
+
+	fprintf(stdout,"Sum 0 is %d\n",get_sum(0));
+	fprintf(stdout,"Sum 1 is %d\n",get_sum(1));
+
+#ifdef SW
+	pthread_cancel(acc_t);
+	killPipeHandler();
+#endif
+}
+\end{verbatim}
+The test-bench starts
+three threads: one to write data to pipe in\_data, one
+to read data from pipe out\_data, and one to run the
+accumulate function.  After the last data is read back
+from out\_data, the test-bench gets the value of sum
+and prints it out.
+
+Obviously, we would prefer to use the same test-bench
+to verify that the VHDL system generated from ``prog.c'' 
+functions correctly.  
+The difference is that instead of using the pipeHandler,
+the test-bench now uses methods in SocketLib.  Further, the
+VHDL is executed in a VHDL simulator; the simulator communicates
+with the testbench using sockets.
+The {\em ifdef's} in the test-bench and the system program
+indicate the difference between the pure software version
+of the system-test-bench combination and the hardware-software
+version.
+
+The following Makefile builds a software-only testbench
+executable, and also converts the system described in prog.c
+to VHDL.  The same testbench can be used to test the VHDL
+also.
+\begin{verbatim}
+SOCKETLIB_INCLUDE=../../CtestBench/include
+SOCKETLIB_LIB=../../CtestBench/lib
+VHDL_LIB=../../vhdl
+VHDL_VHPI_LIB=../../CtestBench/vhdl
+all: SW HW
+
+# compile with SW defined... this
+# tests the SW.
+SW: prog.c prog.h testbench.c 
+	gcc -c -DSW -I$(SOCKETLIB_INCLUDE) prog.c
+	gcc -c -DSW -I$(SOCKETLIB_INCLUDE) testbench.c
+	gcc -o testbench_sw prog.o testbench.o -L$(SOCKETLIB_LIB) -lSocketLib -lpthread
+
+# five steps from C to vhdl simulator.
+HW: c2llvmbc llvmbc2aa aa2vc vc2vhdl vhdlsim
+
+# C to llvm byte-code.. use clang.
+c2llvmbc: prog.c prog.h testbench.c
+	clang -std=gnu89 -I$(SOCKETLIB_INCLUDE) -emit-llvm -c prog.c
+	llvm-dis prog.o
+
+# llvm byte-code to Aa..
+llvmbc2aa:  prog.o
+	llvm2aa --storageinit=true prog.o | vcFormat >  prog.aa
+
+# Aa to vC
+aa2vc: prog.aa
+	AaLinkExtMem prog.aa | vcFormat > prog.linked.aa
+	Aa2VC -O -C prog.linked.aa | vcFormat > prog.vc
+
+# vC to VHDL
+vc2vhdl: prog.vc
+	vc2vhdl -C -e ahir_system -w -s ghdl -T accumulate\
+               -t get_sum -t global_storage_initializer_ -f prog.vc 
+	vhdlFormat < ahir_system.unformatted_vhdl > ahir_system.vhdl
+	vhdlFormat < ahir_system_test_bench.unformatted_vhdl\
+                       > ahir_system_test_bench.vhdl
+
+# build testbench and ghdl executable
+# note the use of SOCKETLIB in building the testbench.
+vhdlsim: ahir_system.vhdl ahir_system_test_bench.vhdl testbench.c vhdlCStubs.h vhdlCStubs.c
+	gcc -c vhdlCStubs.c -I./ -I$(SOCKETLIB_INCLUDE)
+	gcc -c testbench.c -I./ -I$(SOCKETLIB_INCLUDE)
+	gcc -o testbench_hw testbench.o vhdlCStubs.o  -L$(SOCKETLIB_LIB) -lSocketLib -lpthread
+	ghdl --clean
+	ghdl --remove
+	ghdl -i --work=ahir  $(VHDL_LIB)/ahir.vhdl
+	ghdl -i --work=ieee_proposed  $(VHDL_LIB)/ieee_proposed.vhdl
+	ghdl -i --work=work $(VHDL_VHPI_LIB)/Utility_Package.vhdl
+	ghdl -i --work=work $(VHDL_VHPI_LIB)/Vhpi_Package.vhdl
+	ghdl -i --work=work ahir_system.vhdl
+	ghdl -i --work=work ahir_system_test_bench.vhdl
+	ghdl -m --work=work -Wc,-g -Wl,-L$(SOCKETLIB_LIB) -Wl,-lVhpi ahir_system_test_bench
+
+clean:
+	rm -f *.o* *.cf *.*vhdl vhdlCStubs.* *.vcd in_data* out_data* testbench_sw testbench_hw ahir_system_test_bench vhpi.log *.aa *.vc
+
+PHONY: all clean	
+\end{verbatim}
+To test the software, run testbench\_sw.  To verify the hardware (using
+the VHDL simulator GHDL), start testbench\_sw in one shell, and then start ahir\_system\_test\_bench
+in a different shell.
 \end{document}

### 2012-03-04 v2/CtestBench/pipeHandlerTest/simple_check.c

@@ -15,7 +15,7 @@ int main(int argc, char* argv[])
 	pthread_t phandler_t;
 	pthread_create(&phandler_t,NULL,&pipeHandler__,NULL);
 
-	register_pipe("test_pipe",32,32);
+	register_pipe("test_pipe",32,32,0);
 	usleep(1000);
 
 	write_uint32("test_pipe",0);
@@ -26,5 +26,5 @@ int main(int argc, char* argv[])
 	v = read_uint32("test_pipe");
 	fprintf(stderr,"TEST: received %d\n", v);
 
-	pthread_cancel(phandler_t);
+        killPipeHandler();
 }

### 2012-03-04 v2/CtestBench/pipeHandlerTest/test_program.c

@@ -9,45 +9,42 @@ void* pipeHandler__()
 	pipeHandler();
 }
 
-void* Sender()
+void Sender()
 {
 	uint32_t word;
 	for(word = 0; word < 10; word++)
 	{
-		write_uint32("test_pipe",word);
+		write_uint32("test_fifo",word);
 		fprintf(stdout,"Sent: %d\n", word);
-		write_float32("test_pipe",(float)word);
+		write_float32("test_lifo",(float)word);
 		fprintf(stdout,"Sent: %f\n", (float) word);
 	}
 }
 
-void* Receiver()
+void Receiver()
 {
 	uint32_t word;
 	float fword;
-	while(1)
+        uint32_t idx;
+	for(idx = 0; idx < 10; idx++)
 	{
-		word = read_uint32("test_pipe");
-		fprintf(stdout,"Received: %d\n", word);
-		fword = read_float32("test_pipe");
+		fword = read_float32("test_lifo");
 		fprintf(stdout,"Received: %f\n", fword);
+		word = read_uint32("test_fifo");
+		fprintf(stdout,"Received: %d\n", word);
 	}
 }
 
 int main(int argc, char* argv[])
 {
-	pthread_t phandler_t, sender_t, receiver_t;
+	pthread_t phandler_t;
 	pthread_create(&phandler_t,NULL,&pipeHandler__,NULL);
 
-	register_pipe("test_pipe",32,32);
+	register_pipe("test_fifo",32,32, 0);// last arg = 1 => FIFO
+	register_pipe("test_lifo",32,32, 1);// last arg = 1 => LIFO
 
-	pthread_create(&sender_t,NULL,&Sender,NULL);
-	pthread_create(&receiver_t,NULL,&Receiver,NULL);
-
-	pthread_join(sender_t,NULL);
-	usleep(1000);
-	pthread_cancel(receiver_t);
+	Sender();
+	Receiver();
 	
 	killPipeHandler();
-	pthread_join(phandler_t,NULL);
 }

### 2013-01-08 v2/Ahir-asm/devel/include/AaStatement.h
further progress.. the end is neeeaaaarrrrr...

@@ -910,9 +910,6 @@ class AaMergeStatement: public AaSeriesBlockStatement
   set<string,StringCompare> _merge_label_set;
   vector<AaPlaceStatement*> _wait_on_statements;
 
-
-
-
   string _vc_source_link; // name of VC source link.
 
   bool _has_entry_place;
@@ -1025,8 +1022,6 @@ class AaPhiStatement: public AaStatement
   virtual void Write_VC_Datapath_Instances(ostream& ofile);
 
   virtual void Write_VC_Source_Control_Paths(ostream& ofile);
-  virtual void Write_VC_Source_Links(string hier_id, ostream& ofile);
-
 
   virtual void Propagate_Constants();
   virtual bool Is_Constant();
@@ -1035,6 +1030,27 @@ class AaPhiStatement: public AaStatement
 
   void Set_In_Do_While(bool v) { _in_do_while = v;}
   bool Get_In_Do_While() {return(_in_do_while);}
+
+  // this is the big one! added to support loop pipelining.
+  virtual void Write_VC_Control_Path_Optimized(bool pipeline_flag,
+					       set<AaRoot*>& visited_elements,
+					       map<string, vector<AaExpression*> >& ls_map,
+					       map<string,vector<AaExpression*> >& pipe_map,
+					       ostream& ofile);
+
+
+
+  // called only if it appears in a do while loop.
+  virtual void Write_VC_Links_Optimized(string hier_id, ostream& ofile);
+
+  virtual string Get_VC_Reenable_Update_Transition_Name(set<AaRoot*>& visited_elements)
+  {
+    return(this->Get_VC_Name() + "_reenable_");
+  }
+  virtual string Get_VC_Reenable_Sample_Transition_Name(set<AaRoot*>& visited_elements)
+  {
+    return(this->Get_VC_Name() + "_reenable_");
+  }
   friend class AaMergeStatement;
 };
 
@@ -1270,16 +1286,6 @@ class AaDoWhileStatement: public AaStatement
   virtual string Get_VC_Exit_Place_Name() {return(this->Get_VC_Name() + "__exit__");}
 
   virtual void Get_Target_Places(set<AaPlaceStatement*>& target_places); // do nothing.
-
-  // write out all the places associated with PHI statements
-  // in the merge.
-  void Write_VC_Phi_Places(ostream& ofile);
-
-  // write the Phi-sequencers.
[…]

### 2013-02-19 v2/Ahir-asm/devel/docs/AaLRM.tex
further progress.. the end is neeeaaaarrrrr...
Linearity checker example completed.  Also started updating docs (long overdue).

@@
 label-list for the merge must contain the identifier \$entry.
+The target of a phi-statement must be an implicit variable
+(declared by the statement itself, not a storage or a pipe
+object).
 \end{itemize}
@@
 
+\subsection{Pipelinable-loops: the {\bf do-while} construct}
+
+The \Aa language provides a single construct for describing loops.
+The form of the construct is
+\begin{verbatim}
+  $do
+     $merge $entry $loopback
+       <phi-statements>
+     $endmerge
+     atomic-statement-sequence
+  $while <condition-expression>
+\end{verbatim}
+The do-while statement can appear only in branch-blocks.  Control enters
+the do-while from a predecessor in the branch-block, and stays in the
+do-while as long as the condition-expression is false.  The \$loopback
+place is used to repeat the loop-body.  The leading merge statement
+in the do-while is used to merge the control-flow coming in either from
+the entry to the do-while or from the loopback place.
+
+The downstream tools will always attempt to pipeline the do-while loop by
+starting the next iteration while the current one is still in progress.
+Dependency analysis on values in the loop-body is used to determine the
+extent of loop pipelining.  Note that it is always possible to describe
+a do-while loop using a branch-block, and this is what should be done
+if loop-pipelining is not desired.
+
 Thus, an \Aa program is constructed as a collection of modules,

### 2013-02-19 release/examples/linearityChecker/src/pE.c

@@ -1,69 +1,0 @@
-#include <stdint.h>
-#include "pEMacro.h"
-
-#define __Row 0
-#define __Col 0
-#define __NRow 1
-#define __NCol 1
-#define __PRow -1
-void pE00() 
-{
-	float X[4096], Y[4096];
-	uint32_t vec_size, num_vecs,vec_index,I,J,K,XIDX,YIDX,NX,NY;
-	__pE;
-}
-
-#undef __Row 
-#undef __Col 
-#undef __NRow 
-#undef __NCol 
-#undef __PRow 
-
-#define __Row 0
-#define __Col 1
-#define __NRow 1
-#define __NCol -1
-#define __PRow -1
-
-void pE01()
-{
-	float X[4096], Y[4096];
-	uint32_t vec_size, num_vecs,vec_index,I,J,K,XIDX,YIDX,NX,NY;
-	__pE(0,1,1,-1);
-}
-
-#undef __Row 
-#undef __Col 
-#undef __NRow 
-#undef __NCol 
-#undef __PRow 
-
-#define __Row 1
-#define __Col 1
-#define __NRow -1
-#define __NCol -1
-#define __PRow 0
-void pE11()
-{
-	float X[4096], Y[4096];
-	uint32_t vec_size, num_vecs,vec_index,I,J,K,XIDX,YIDX,NX,NY;
-	__pE(1,1,-1,-1);
-}
-
-#undef __Row 
-#undef __Col 
-#undef __NRow 
-#undef __NCol 
-#undef __PRow 
-
-#define __Row 1
-#define __Col 0
-#define __NRow -1
-#define __NCol 1
-#define __PRow 0
-void pE10()
-{
-	float X[4096], Y[4096];
-	uint32_t vec_size, num_vecs,vec_index,I,J,K,XIDX,YIDX,NX,NY;
-	__pE(1,0,-1,1);
-}

### 2013-02-19 release/examples/linearityChecker/src/prog.c

@@ -3,12 +3,8 @@
 #include <Pipes.h>
 #include <stdio.h>
 #include "prog.h"
+#include "pEMacro.h"
 
-// PE_ARRAY_DIM must be a power of 2.
-#define PE_ARRAY_DIM   2
-
-// mask
-#define SEL_MASK ~((uint32_t) PE_ARRAY_DIM) 
 
 // There are N vectors v0, v1 ... vN-1.
 // The vectors x0,x1,...xN-Omega correspond to v0,v1,...,vN-Omega
@@ -16,178 +12,210 @@
 //
 // All differences || xI - yJ || for I<=J need to be computed.
 //
-// The processing elements are indexed as (I,J), 0 <= I,J <= PE_ARRAY_DIM.
-// The element (I,J) will maintain all xP, yQ such that
-// P mod PE_ARRAY_DIM = I and Q mod PE_ARRAY_DIM = J.
-// 
-// Each processing element (I,J) will then compute the differences for pairs
-// XU,YW such that U <= W.
-//
 //
 
-// 32 buffers each with VEC_SIZE entries.
-float vec_buffers[32*VEC_SIZE];
 
-// globals
+// storage for the vectors, divided into 4 blocks.
+float X0[CHUNK_SIZE], X1[CHUNK_SIZE], X2[CHUNK_SIZE], X3[CHUNK_SIZE];
+
 // number of vectors.
 uint32_t g_num_vecs;
 
 
-void init()
-{
-	int i;
-	for(i = 0; i < 1024; i=i+VEC_SIZE)
-	{
-		// load the buffer.
-		write_uintptr("free_buffer_pipe",(uint32_t*) &(vec_buffers[i*VEC_SIZE]));
-	}
-}
-
-// transmit control information to the left border of the PE array.
-void TxHControl(uint32_t val)
-{
-	__WHC(0,0,val);
-	__WHC(1,0,val);
-}
-
-// transmit control information to the top border of the PE array.
-void TxVControl(uint32_t val)
-{
-	__WVC(0,0,val);
-	__WVC(0,1,val);
-}
-// transmit data values to the the left and top borders
-// of the PE array.
-void TxValues()
-{
-	while(1)
-	{
-		// wait to hear from tx_vec_id pipe, to get
-		// a vector index.
-		uint32_t vec_id = read_uint32("tx_vec_id");
-
-		if(vec_id == 0xffffffff)
-		{
[…]

### 2013-02-22 Documents/Publications/dsd2013/paper.tex
Started writing a paper on loop-pipelining for EUROMICRO.

\documentclass[conference]{IEEEtran} 
\usepackage{graphicx}
\usepackage[tight,footnotesize]{subfigure}
\usepackage{amsmath}
\usepackage{alltt}
\usepackage{multirow}
\usepackage{times}
\usepackage{url}

\title{On the efficient implementation of loops in hardware}

\author{Madhav P. Desai\\
  Indian Institute of Technology -- Bombay, Powai, Mumbai -- 400076, INDIA\\
  Email: madhav@ee.iitb.ac.in }
\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]

\newcommand{\sym}[1]{$\operatorname{#1}$}

\newbox\subfigbox             % Create a box to hold the subfigure.
\makeatletter
  \newenvironment{subfloat}% % Create the new environment.
    {\def\caption##1{\gdef\subcapsave{\relax##1}}%
     \let\subcapsave=\@empty % Save the subcaption text.
     \let\sf@oldlabel=\label
     \def\label##1{\xdef\sublabsave{\noexpand\label{##1}}}%
     \let\sublabsave\relax    % Save the label key.
     \setbox\subfigbox\hbox
       \bgroup}%              % Open the box...
      {\egroup                % ... close the box and call \subfigure.
     \let\label=\sf@oldlabel
     \subfigure[\subcapsave]{\box\subfigbox\sublabsave}}%
\makeatother

\pagestyle{empty}

\begin{document}

\maketitle
\thispagestyle{empty}

\begin{abstract}

  In the context of mapping high-level algorithms to hardware,
  we consider the basic problem of generating an efficient hardware 
  implementation of a loop.  We describe a control-flow mechanism through
  which multiple iterations of a loop can be dynamically pipelined in
  the generated hardware,  and study the extent to which this 
  mechanism affects the properties of the generated hardware.
  First, we observe that the additional resources needed by this
  mechanism are small.  Further, we observe that this mechanism, when
  combined with static loop unrolling, offers an order-of-magnitude improvement 
  in performance for common linear-algebra kernels such as matrix multiplication 
  and inner product computations.  Finally, we observe that static
  loop unrolling together with dynamic loop pipelining can give us
  efficient hardware with excellent utilization of the available arithmetic
  resources in the circuit.

\end{abstract}

\section{Introduction}

%% The importance and problem of loops.

%% How compilers have handled this problem.

%% Algorithm-to-hardware:  can similar techniques work?

\section{Related work}

%% compiler techniques.

%% hls techniques to handle loops?

\section{Model of hardware generated by our compiler}

%%
%%  control path and data-flow
%%

\subsection{Control-path}

%% 
%% Petri-net, constructed according to
%% certain rules.  Req/Ack transitions
%%

\subsection{Data-path}

%%
%% Directed hyper-graph.  Nodes are operators
%% hyper-arcs are wires.  Each operator node
%% has a split protocol.
%%

\subsection{Memory-subsystem}

%%
%% time-stamping scheme, in order completion.
%%

\subsection{An example of a loop}

%%
%% for(i=0;i < 100; i++)
%% {
%%    c[i] = a[i] + b[i];
%% }
%%

\section{Control-flow mechanism for loop-pipelining}

%%
%% reenable rules.
%%
%% loop-terminator.
%%
%% advantages of dynamic pipelining: any loop
%% can be handled..
%%

\section{Experimental Results}

%%
%% examples: vector add, dot-product, matrix-multiply, linfinity-norm.
%%
%%
%%       show results without-unrolling-pipelining, with-unrolling-pipelining, etc.
%%

\section{Conclusion}

%%
%% loop-pipelining and loop-unrolling have high impact as
%% in pipelined processors.
%%
%% hardware mechanism for loop-pipelining is cheap and effective.
%%
%% can get 10X improvement and very high utilization of 
%% hardware resources.
%%

\bibliography{ref}
\bibliographystyle{IEEEtran}

\end{document}

% LocalWords:  Req Ack LL DP CP cmpgt br cdfg cp dp ir intra req init NCA SA mW
% LocalWords:  TPR STPR STPRs ccccc RTL ASIC petri ack versa liveness LRG LLVM
% LocalWords:  VHDL SRAM Synopsys AES LPK Linpack FFT RBT Synposys OSU TSMC nJ
% LocalWords:  nm

### 2013-03-01 Documents/Publications/dsd2013/paper.tex
Added examples for dsd2013 paper..

@@
 
-\title{On the efficient implementation of loops in hardware}
+\title{On the impact of loop optimizations in the mapping of high-level programs to hardware}
 
@@
   combined with static loop unrolling, offers an order-of-magnitude improvement 
-  in performance for common linear-algebra kernels such as matrix multiplication 
-  and inner product computations.  Finally, we observe that static
-  loop unrolling together with dynamic loop pipelining can give us
-  efficient hardware with excellent utilization of the available arithmetic
-  resources in the circuit.
+  in performance for common kernels such as the fast-fourier transform,
+  matrix multiplication and inner product computations.  
+  Finally, we observe that these loop-optimization techniques in
+  yield efficient mappings of single threads to hardware.  %% need more stuff here.
 
@@
 %% The importance and problem of loops.
+A high-level-synthesis (HLS) compiler takes an algorithm described in
+a high-level programming language (such as C, for example) and produces a circuit
+description (a register-transfer-level description in a hardware
+description language such as VHDL, for example).
+High-level synthesis systems have been available for some time
+now (for example, \cite{pegasus-cash}, \cite{spark-vlsi-paper} and
+commercial tools such as Catapult-C\footnote{\url{http://www.mentor.com/products/esl/high_level_synthesis/}} from Mentor Graphics
+C-to-Silicon\footnote{\url{http://www.cadence.com/products/sd/silicon_compiler/}} from Cadence,
+Synopsys HLS\footnote{\url{http://www.synopsys.com/Tools/SLD/HLS/}} from Synopsys).
+
+The main attraction of such a compiler is a substantial reduction in the design and verification effort.  The
+main concern is the quality of the hardware that is produced.
+For the same starting point, the quality of the hardware produced may be objectively
+measured in terms of factors such as area, energy-consumption and performance (in the
+sense of time required for completion, or in terms of the rate at which the hardware
+finishes the computational tasks).  An {\em internal measure} of the quality of the
+hardware produced is the utilization factor of the major resources in the hardware (that is,
+the utilization factor of expensive resources).
 
 %% How compilers have handled this problem.
+In this paper, we consider the problem of mapping loops to hardware.  It
+is well known that most compute intensive programs spend a large fraction
+of their time in inner loops.  Thus, the optimization of these loops is
+a primary problem for any optimizing compiler which is targetting
+either a deeply pipelined or parallel processor.  Several loop optimizations
+have been considered in literature, such as loop-unrolling, loop-peeling,
+software-loop-pipelining etc. \cite{wolfe}, \cite{muchnick}, with the intent of
+these optimizations being the extraction of as much
+parallelism as possible from the source program.
 
 %% Algorithm-to-hardware:  can similar techniques work?
+One expects that any technique which extracts parallelism from
+the source program would also be useful in improving the efficiency
+of hardware generated from that source program.  We consider two
+such loop optimizations:
+\begin{itemize}
+\item {\em Loop-unrolling}, a {\em static} compile-time optimization technique:  
+an inner loop is unrolled by instantiating multiple copies of the loop-body
+while simultaneously reducing the number of loop-iterations.  For example:
+\begin{verbatim}
+for(i=0; i < 8; i++) {
+   x += a[i]*b[i];
+}
+\end{verbatim}
+is transformed to
+\begin{verbatim}
+for(i=0; i < 4; i+=2) {
+   x += a[i]*b[i];
+   x += a[i+1]*b[i+1];
+}
+\end{verbatim}
+This unrolling increases the size of the basic block (that is, the maximal sequence of
+statements without any branches), and provides the possibility of extracting more
+parallelism in the loop.
+\item {Loop-pipelining}, a {dynamic} run-time optimization in hardware:  the hardware
+itself starts executing multiple iterations of the loop in parallel, while respecting
+all dependencies within and across loop iterations.   This is analogous to software
+loop-pipelining, except that software-loop-pipelining is static in nature, and is
+determined at compile time, whereas we are looking at a dynamic technique in which
+the generated hardware itself keeps multiple loop iterations in flight while maintaining
+dependencies within and across loop iterations.
+\end{itemize}
+
+In the remainder of this paper, we will first briefly describe the model of
+the hardware that is produced by our HLS compiler and illustrate how this model
+can incorporate dynamic run-time loop pipelining.  The chief issues here are the
+hardware overhead (area, energy, delay) incurred by the need to provide 
+this run-time support in the hardware generated by the compiler, and
+the corresponding improvement in performance that results from this optimization.
+
+In order to address this issue, we will present a set of observations from
+experiments performed on representative inner loops that occur in 
+some important applications such as the fast-fourier transform, the
+matrix product, vector dot-product, and a digital filtering algorithm.
+These observations report the hardware performance on four different
+loop-optimization choices: with no loop optimization, with static unrolling
+alone, with dynamic loop pipelining alone, and with static unrolling combined with
+dynamic loop pipelining.  Hardware resource utilization and delays are
+computed by synthesizing the generated hardware for an FPGA target.
+
+The observations indicate the following:
+\begin{itemize}
+\item The performance improvement with loop-pipelining alone is in the
+2X-3X range, with a small increase in the hardware resources needed.
+\item The performance improvement with loop-unrolling alone is in the 
+2X range, but substantial hardware overheads are incurred.
+\item The combination of loop-pipelining and loop-unrolling leads
+to a performance improvement between 8-10X, with the hardware overheads
+being similar to those incurred with unrolling alone.
+\end{itemize}
+
+%%
+%% TODO..  so what?
+%%
 
@@
 \section{Model of hardware generated by our compiler}
+
+The hardware generated by our compiler consists of three
+cooperating components: the control-path, the data-path and
+the storage system \cite{ahir}
+
+To illustrate the model, we consider a simple example.
+\begin{verbatim}
+float a[1024], b[1024];
+float dotp = 0.0;
+for(i=0; i < 1024; i++)
+{
+   dotp += a[i]*b[i];
+}
+\end{verbatim}
+To compile this code, we use the clang-2.8 \cite{clang} C compiler, 
+which is used to emit LLVM \cite{llvm} byte-code.  The LLVM byte-code
+is transformed through a series of steps by our compiler tools to
+produce the final hardware, which is depicted in Figure \ref{fig:dotp}.
 

### 2013-03-03 Documents/Presentations/eurasip/Talk.tex
Committing talk at Eurasip.

\documentclass{beamer}
\usepackage{epsfig}
\title{IV EURASIP Seminar on the Hardware Design of DSP Systems\\
Program-to-hardware transformation and its use in heterogenous high-performance DSP}
\author{Madhav P. Desai \\ Department of Electrical Engineering\\
  Indian Institute of Technology -- Bombay, Powai, Mumbai -- 400076, INDIA}
\date{March 6, 2013}
\begin{document}

\maketitle

\frame[containsverbatim]{\frametitle{Overview}

High-performance DSP can take advantage of several platforms:
\vspace{0.1in}

\begin{itemize}
\item Processors such as multi-cores, GPUs: ``easy'' to use, but power
	dissipation is an issue.
\item Programmable hardware:  lots of parallelism in principle, potential
	power benefits, difficult to map algorithms to  platform.  Can they
	beat power-performance characteristics of processors?
\item ASICs:  lots of parallelism, high efficiency, will surely beat power-performance
	characteristics of processors.  Difficult to design, and expensive to
	manufacture.
\end{itemize}

In terms of efficiency (performance/watt):
\begin{verbatim}
ASIC  > >  Processor, FPGA
Processor < ? >  FPGA 
\end{verbatim}

}

\frame[containsverbatim]{\frametitle{Our work}

\begin{itemize}
\item Make the path to ASIC/FPGA simpler for algorithm developers.
\item Develop (and use) a compiler flow for mapping algorithms to hardware.
\item Currently, we have a fully functional tool-chain which takes a C program to VHDL.  
	A multi-threaded C program (with statically defined threads) can
	be easily mapped to a hardware pipeline.
\item We are now working on optimizations (will talk more about that later)
and {\bf applications}.
\end{itemize}

}

\frame[containsverbatim]{\frametitle{How good is the compiler flow?}
We have some experimental data (from 2010)\footnote{S. Sahasrabuddhe, S. Subramanian,
K. Ghosh, K. Arya, M.P. Desai, ``A C-to-RTL flow as an energy efficient alternative
to embedded processors in digital systems'', EUROMICRO 2010, Lille France.} :
	\begin{itemize}
	\item comparison in terms of area, delay and {\bf energy}.
	\item for a set of programs, use standard commercial synthesis tools to obtain
		 an ASIC starting from the RTL generated by the C-to-RTL flow.
	\item compare with processor implementations of the same set of programs.
	\end{itemize}
}

\frame[containsverbatim]{\frametitle{How applicable is it?}
We have some data from a practical use-scenario (from 2012)\footnote{T. Rinta-Aho, M. Karlstedt,
M. P. Desai, ``The Click2Netfpga toolchain'', USENIX ATC 2012, Boston USA.}:
	\begin{itemize}
	\item The Click2Netfpga toolchain (joint work with T. Rinta-Aho, M. Karlstedt
		at Ericsson research) converts a C++ program (generated by the Click
		modular router) to VHDL.
	\item Verified on FPGA, performance up to 50\% of hand-coded Verilog.
	\end{itemize}
}

\frame[containsverbatim]{\frametitle{A system in AHIR}

\begin{figure}
\begin{centering}
\centerline{\psfig{figure=AhirSystem.eps,width=2.5in,height=2.0in}}
\end{centering}
\end{figure}

\begin{itemize}
\item A system description consists of a collection of modules (a module is
similar to a function in a C program).  The modules communicate
with the environment, with each other (through calls, and through
FIFO pipes), and with memory subsystems. 
\end{itemize}

}

\frame[containsverbatim]{\frametitle{An AHIR module}

An AHIR module description is of the form:

\begin{displaymath}
{\bf CP} \ \times \ {\bf DP} \ \times \ {\bf S}
\end{displaymath}

where ${\bf CP}$ represents control flow, ${\bf DP}$ represents 
data flow, and {\bf S} represents storage.
}

\frame[containsverbatim]{\frametitle{The control path ${\bf CP}$}

\begin{itemize}
\item The control path is modeled by a petri-net.  Each transition 
is associated with a symbol, which can either be an input symbol
or an output symbol.
\item When a transition labeled by an input symbol is enabled, 
it fires when the input symbol is received.
\item When a transition labeled by an output symbol is enabled,
it fires eventually and emits the output symbol.
\item The petri-net has a special structure, which guarantees that
{\bf CP} is live and safe.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{The data path ${\bf DP}$}

\begin{itemize}
\item A directed graph of operators.
\item Each operator has a set of request symbols and a set of acknowledge symbols.
The operator responds to the arrival of request symbols by eventually generating
appropriate acknowledge symbols.
\item Operator library: all the standard arithmetic and logical operators, as
well as a multiplexor and load/store operators.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{The storage subsystem ${\bf S}$}
\begin{itemize}
\item The storage subsystem connects to the load/store operators in the datapath,
and is required to eventually service the load/store requests.
\item Declared variables in the source program are grouped
into storage spaces.  
\item The storage spaces are determined by
static reference analysis (the system can have
several distinct memory subsystems).
\item The memory subsystems have a relaxed consistency model: all accesses are time-stamped,
and accesses to the same memory location are finished in first-come-first-served
order.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{An example: the dot-product} 
Consider the program fragment:
\begin{verbatim}
float A[1024], B[1024];
float dotP = 0.0;
for(I=0; I < 1024; I++) {
   dotP += A[I]*B[I];
}
\end{verbatim}
}

\frame[containsverbatim]{\frametitle{Dot-product circuit} 
\begin{figure}
\begin{centering}
\centerline{\psfig{figure=dotP.eps,width=4.0in,height=3.0in}}
\end{centering}
\end{figure}
}

\frame[containsverbatim]{\frametitle{Our C-to-VHDL flow} 
\begin{itemize}
\item The starting point is a program written in C
\begin{itemize}
\item restrictions: no cycles in call graph, and no {\bf function} pointers.
\end{itemize}
\item Use the CLANG front end to convert C program into LLVM byte code (CDFG).
\item Map the CDFG to an Ahir assembly program (using the
{\bf Aa} language, in which parallelism can be expressed in a native
way). 
\item Map the {\bf Aa} description to a virtual circuit consisting
of interacting AHIR modules.
\begin{itemize}
\item identify disjoint memory spaces.
\item dependency analysis to maximize parallelism in straight-line code.
\end{itemize}
\item Map the virtual circuit to VHDL.
\begin{itemize}
\item identify arbiter-less resource sharing opportunities to reduce hardware
cost.
\item instantiate the system: modules with their control and data-paths, the
inter-module link layer to handle calls, the memory subsystem.
\end{itemize}
\end{itemize}
}

\frame[containsverbatim]{\frametitle{Our C-to-VHDL flow} 
\begin{figure}
\begin{centering}
\centerline{\psfig{figure=AhirFlow.eps,width=3.0in,height=3.0in}}
\end{centering}
\end{figure}

}
\frame[containsverbatim]{\frametitle{Mapping to VHDL} 
\begin{itemize}
\item A synchronous, single clock, positive edge-triggered paradigm 
is used.
\item Transitions are coded as pulses which are sampled high by one
clock edge. 
\item Only join transitions need flip-flops for their implementation. 
Other elements are purely combinational in nature, and in many cases,
can be optimized away.
\item The datapath is mapped to an equivalent VHDL netlist constructed
using a library of operators.  When a set of operations shares
a single operator, input multiplexors and output demultiplexors are introduced.
\item The memory subsystem is implemented using multiple banks, and can
offer multiple ports (performance-cost tradeoff).
\end{itemize}
}

\frame[containsverbatim]{\frametitle{Experimental Evaluation 1}
\begin{itemize}
\item Select a range of programs
\begin{itemize}
\item A5 (stream cipher), AES encryption, Red-black Trees (data-structure),
      Linpack (LU factorization), Fast-Fourier Transform (FFT).
\end{itemize}
\item Using the C-to-VHDL flow, map each program to RTL.  The run-time for
mapping is neglible in all cases (less than a minute).
\item Use standard synthesis tools (Synopsys Design Compiler, Cadence SOC encounter)
to implement ASIC from RTL (we use the 180nm TSMC CMOS process, with OSU standard
cell libraries).  Extract area, delay and energy numbers for the implemented ASIC.
\begin{itemize}
\item Operators are not pipelined or optimized in any fashion.
\end{itemize}
\item Run each program on a  processor based platform (we use the Intel Atom N270 as
a reference).  Extract area, delay, energy numbers using the processor data sheet.
\item Compare the ASIC numbers with the processor numbers.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{Comparison of normalized (to 45nm) C-to-VHDL circuits with 45nm processor}
\begin{table}[htb]
  \centering
  \caption{Area/Delay/Power/Energy RATIOS (processor values relative to the scaled C-to-RTL circuit values)}
  \label{table:Ratios}
  \renewcommand\arraystretch{1.2}
  \setlength{\tabcolsep}{1ex}
  \begin{tabular}{c|c|c|c|c|c}
    \hline
    & Area & Freq & Delay & Power & Energy \\
    \hline
    \hline
    A5/1 & 275.8 & 5.6 & 1.7 & 68.5 & 116.6  \\
    \hline
    AES & 61.5 & 5.6 & 0.34 & 14.8 & 4.9 \\
    \hline
    FFT & 78.4 & 9.6 & 0.57 & 52.6 & 30.3  \\
    \hline
    LPK & 14.8 & 9.6 & 0.84 & 18.3 & 15.3  \\
    \hline
    RBT & 22.2 & 9.6 & 0.15 & 32.7 & 4.7 \\
    \hline
  \end{tabular}
\end{table}
}

\frame[containsverbatim]{\frametitle{Analysis}

\begin{itemize}
\item The C-to-RTL circuits are between 4.7X and 116X more energy efficient than the processor.
\item The processor delays are lower than those of the C-to-RTL circuit in most
cases (except for A5) but the difference is less than one order of magnitude in
all cases.
\begin{itemize}
\item pipelining of operators and retiming of the C-to-RTL circuits should reduce
this gap.
\end{itemize}
\end{itemize}
For single-threaded applications, the C-to-RTL circuits have much better 
energy efficiency than the processor, but have lower performance than the processor.  
\vspace{0.1in}

{\bf Update: }  Loop optimizations seem to be very useful for improving single
thread performance (we will talk about this later).

}

\frame[containsverbatim]{\frametitle{Experimental Evaluation 2: Click2Netfpga}

This work is a joint effort with Teemu Rinta Aho and 
Mika Karlstedt at Ericsson Research, and will be presented
at the USENIX ATC 2012 (Boston).

\begin{itemize}
\item Click is a modular framework for describing networking
devices (e.g. routers, classifiers etc.).  A Click description
is converted to C++ code, which can then be compiled and executed.
\item The Click2Netfpga tool-chain takes Click-generated C++ code
and produces VHDL which can be mapped to the Stanford NetFPGA card
and validated in real-time.
\begin{itemize}
\item The Click2llvm front-end generates LLVM
modules (which form a pipeline) from
the C++ code.
\item The AHIR tool flow maps the LLVM modules to a hardware pipeline.
\end{itemize}
\item A router generated in this manner gives up to 50\% of the
performance of a reference (hand-coded Verilog) design.
\end{itemize}

{\bf Update:} Loop optimizations should reduce the gap between hand-designed
and compiler generated hardware (will talk about this later).
}

\frame[containsverbatim]{\frametitle{Trends}

\begin{itemize}
\item On purely sequential code, the hardware that is generated
consumes substantially less energy than a processor running the
same code.  Performance of single-threaded hardware needs to
improve.
\item For processing pipelines inferred from application
specific C++ code, the performance is upto 50\% of hand-coded
Verilog.   The performance gap needs attention, but 50\% of
hand-coded is ``not bad''.
\end{itemize}

Offers a viable option to the system designer (in addition
to the use of embedded processors and custom hardware).
}

\frame[containsverbatim]{\frametitle{Optimizations: loop-pipelining}

Clearly, there is scope for improvement. 
The most critical problem in improving single-thread performance
is the inner loop.  
\begin{verbatim}
float A[1024], B[1024];
float dotP = 0.0;
for(I=0; I < 1024; I++) {
   dotP += A[I]*B[I];
}
\end{verbatim}
Execute multiple-iterations of a loop whenever possible.  
}

\frame[containsverbatim]{\frametitle{Dynamic loop-pipelining}

In an AHIR system, loop-pipelining can be managed if 
the control-path can maintain dependencies across loops:
\begin{itemize}
\item $F_k \rightarrow F_{k+1}$ for all operations $F$, for all $k$.
\item If $G$ depends on $F$, then $F_k \rightarrow G_k \rightarrow F_{k+1}$.
\item If $P, Q$ are memory operations with an explicit dependency (WAR, RAW, WAW),
then $P_k \rightarrow Q_k \rightarrow P_{k+1} \rightarrow Q_{k+1}$.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{Modified control-path for dynamic loop-pipelining}
\begin{figure}
\begin{centering}
\centerline{\psfig{figure=pipelinedCP.eps,width=3.0in,height=3.0in}}
\end{centering}
\end{figure}
}

\frame[containsverbatim]{\frametitle{Impact of dynamic loop-pipelining on single-threads}
We looked at some simple, but important examples:
\begin{itemize}
\item Dot-product.
\item FFT.
\item Matrix-multiply.
\end{itemize}

In each case, single-threaded performance improvements were measured.
Number of FP adders and multipliers used = 1 of each.
}

\frame[containsverbatim]{\frametitle{Impact of loop-optimizations on the 64-point dot-product}

For the 64-point dot-product (time/area numbers taken from FPGA logic
synthesis targeting a Xilinx Virtex-6):
\begin{verbatim}

             plain     pipelined   unrolled   unrolled
                                              and 
                                              pipelined

Cycles       4071       1874        1894        582
LUTs         4288       5195        4809       7344
FFs          3799       4345        4378       5911
Freq.(MHz)   199.7      199.7       199.7      199.7
\end{verbatim}

Best case FLOPS/cycle = 127/582 = 0.22.
}

\frame[containsverbatim]{\frametitle{Impact of loop-optimizations on the 64-point FFT}

For a single stage of the 64-point FFT (32 butterflies, 300 FP ops).
\begin{verbatim}

            plain     pipelined   unrolled   unrolled
                                             and 
                                             pipelined

Cycles       4151       2885       3110       1064
LUTs        12155      23139      16032       37831
FFs         11955      18632      15299       28698
Freq.(MHz)  186.9      164.1      186.9       164.1
\end{verbatim}

Best case FP-ops/cycle = 300/1064 = 0.3.
}

\frame[containsverbatim]{\frametitle{Impact of loop-optimizations on the 16x16 matrix-multiply}

Here, some aggressive loop-unrolling was used.
\begin{verbatim}

             plain    pipelined  unrolled   unrolled
                                            and 
                                            pipelined
Cycles       161K       77K         13K        7810
LUTs         6323      9408       16032       31744
FFs          6974      8818       15299       21060
Freq.(MHz)   199.7     199.7        186       164.1
\end{verbatim}

Best case FLOPS/cycle = 7936/7810 =1.01.
}

\frame[containsverbatim]{\frametitle{Observations}

\begin{itemize}
\item Excellent performance improvement with loop-pipelining and unrolling.
\item Single-threaded hardware performance is pretty OK.
\begin{itemize}
\item BUT: resource utilization levels need to be improved (anything above
50\% is reasonable).
\item Main cause is latency in non-parallel code (Amdahl's law).  This effect
can be reduced by reducing latency in the logic (in progress).
\item As the problem size is increased, utilization levels tend to increase.
\end{itemize}
\item In practice, we will need to use multiple threads to exploit advantages
of hardware relative to the processor.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{Future directions for our research}
\begin{itemize}
\item Applications: signal processing (image/multi-media), cryptology, 
CFD, Databases etc.
\item More optimizations:  latency reduction in non-parallel paths.
\item Algorithm-to-ASIC.
\item OpenMP, OpenCL support.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{Conclusion}

The AHIR flow is open-source.  If you wish to try
it out, please get in touch with me: madhav@ee.iitb.ac.in

Thank you!
}

\end{document}

### 2013-03-03 Documents/Publications/dsd2013/paper.tex
Committing talk at Eurasip.

@@
 \usepackage{graphicx}
-\usepackage[tight,footnotesize]{subfigure}
 \usepackage{amsmath}
 \usepackage{alltt}
-\usepackage{multirow}
 \usepackage{times}
@@
 
-\title{On the impact of loop optimizations in the mapping of high-level programs to hardware}
+%% revisit the title.
+\title{A dynamic loop pipelining mechanism for the efficient realization of algorithms as hardware}
 
@@
 \newcommand{\sym}[1]{$\operatorname{#1}$}
-
-\newbox\subfigbox             % Create a box to hold the subfigure.
-\makeatletter
-  \newenvironment{subfloat}% % Create the new environment.
-    {\def\caption##1{\gdef\subcapsave{\relax##1}}%
-     \let\subcapsave=\@empty % Save the subcaption text.
-     \let\sf@oldlabel=\label
-     \def\label##1{\xdef\sublabsave{\noexpand\label{##1}}}%
-     \let\sublabsave\relax    % Save the label key.
-     \setbox\subfigbox\hbox
-       \bgroup}%              % Open the box...
-      {\egroup                % ... close the box and call \subfigure.
-     \let\label=\sf@oldlabel
-     \subfigure[\subcapsave]{\box\subfigbox\sublabsave}}%
-\makeatother
 
@@
   implementation of a loop.  We describe a control-flow mechanism through
-  which multiple iterations of a loop can be dynamically pipelined in
+  which multiple iterations of an arbitrary loop can be dynamically pipelined in
   the generated hardware,  and study the extent to which this 
   mechanism affects the properties of the generated hardware.
-  First, we observe that the additional resources needed by this
-  mechanism are small.  Further, we observe that this mechanism, when
-  combined with static loop unrolling, offers an order-of-magnitude improvement 
-  in performance for common kernels such as the fast-fourier transform,
-  matrix multiplication and inner product computations.  
-  Finally, we observe that these loop-optimization techniques in
-  yield efficient mappings of single threads to hardware.  %% need more stuff here.
+  We consider some common loop kernels occurring in fast-fourier transform,
+  matrix multiplication and inner product computations, and
+  observe that this dynamic pipelining mechanism typically results in
+  a doubling of the performance of the hardware.  When 
+  combined with static loop unrolling,  the performance improvement
+  ranges from 6X to 20X across the loop kernels that we have studied.
+  These optimizations do have a hardware overhead, however.  We characterize
+  the performance/cost trade-off using an FPGA target, and find
+  that the loop optimizations not only improve performance, but
+  also the cost/performance ratio of the resulting hardware.
 
@@
 
-The main attraction of such a compiler is a substantial reduction in the design and verification effort.  The
+The main attraction of such a compiler is a substantial reduction in 
+the design and verification effort.  The
 main concern is the quality of the hardware that is produced.
@@
 %% Algorithm-to-hardware:  can similar techniques work?
-One expects that any technique which extracts parallelism from
-the source program would also be useful in improving the efficiency
-of hardware generated from that source program.  We consider two
-such loop optimizations:
-\begin{itemize}
-\item {\em Loop-unrolling}, a {\em static} compile-time optimization technique:  
-an inner loop is unrolled by instantiating multiple copies of the loop-body
+Several loop-optimization techniques have been explored and
+reported in the literature related to reconfigurable hardware (for
+example \cite{Weinhardt}, \cite{Cardoso}, \cite{Kastner}).  For instance,
+in the work reported by \cite{Weinhardt},
+the loop optimizations are done in a manner analogous to the static
+techniques used in software compilers, in which index 
+expressions which depend on the induction variable
+are analysed to identify dependencies and schedule operations
+in the loop body.  An explicitly timed controller is synthesized
+for the pipeline.  Another approach which works in a similar
+manner is described in \cite{Cardoso}.  These approaches rely on a static analysis of
+the loop, and the cases to which the approach can be applied
+are restricted (but still sufficiently general for most linear
+algebra and digital signal processing kernels).
+
+In our approach, the hardware model consists of a data-path (which is
+a graph of operators interconnected by wires) and an untimed control-path
+which is modeled as a Petri-net.  This representation allows the implementation
+of loop pipelining by a simple modification to the control Petri-net
+{\em without altering} the data-path.  Thus,
+it is possible to pipeline any loop, even those that do not have 
+explicit induction variables (such as while loops).
+
+The experimental results in this paper are based on the dynamic loop-pipelining
+optimization applied by itself and in conjunction with static loop-unrolling.
+By loop-unrolling, we mean a {\em static} compile-time optimization technique in
+which   an inner loop is unrolled by instantiating multiple copies of the loop-body
 while simultaneously reducing the number of loop-iterations.  For example:
@@
 parallelism in the loop.
-\item {Loop-pipelining}, a {dynamic} run-time optimization in hardware:  the hardware
-itself starts executing multiple iterations of the loop in parallel, while respecting
-all dependencies within and across loop iterations.   This is analogous to software
-loop-pipelining, except that software-loop-pipelining is static in nature, and is
-determined at compile time, whereas we are looking at a dynamic technique in which
-the generated hardware itself keeps multiple loop iterations in flight while maintaining
-dependencies within and across loop iterations.
-\end{itemize}
 
@@
 dynamic loop pipelining.  Hardware resource utilization and delays are
-computed by synthesizing the generated hardware for an FPGA target.
+computed by synthesizing and simulating the generated hardware for an FPGA target.
 
@@
 \begin{itemize}
-\item The performance improvement with loop-pipelining alone is in the
-2X-3X range, with a small increase in the hardware resources needed.
+\item The performance improvement with loop-pipelining applied alone is in the
+2X-3X range.
 \item The performance improvement with loop-unrolling alone is in the 
-2X range, but substantial hardware overheads are incurred.
+2X range, and with aggressive unrolling (as was tried in the matrix multiplication
+case), the improvement is as high as 10X.
 \item The combination of loop-pipelining and loop-unrolling leads
-to a performance improvement between 8-10X, with the hardware overheads
-being similar to those incurred with unrolling alone.
+to a performance improvement which is at least as high as the
+product of the improvements due to the individual optimizations.  
+For matrix multiplication, this improvement is as large as 20X.
+\item The hardware overheads for implementing these optimizations are substantial,
+but the cost-to-performance ratio improves substantially in all cases.
 \end{itemize}
@@
 
-\section{Related work}
+\section{A note on our compiler flow}
+
+We give a brief description of our compiler flow.  The
+details are not relevant to this paper, and the interested
+reader can find them in \cite{dsd2010}.
+
+Our compiler starts with a C program and produces VHDL.  For the
+C front-end, we use the clang-2.8 compiler\footnote{www.clang.org}.
+This compiler is used to emit LLVM byte-code\footnote{www.llvm.org}, 
+which is then transformed to VHDL using the following transformations:
+\begin{enumerate}
+\item The LLVM byte-code is translated to an internal intermediate
+format, which is itself a static-single assignment centric 
+control-flow language (named {\bf Aa}) which allows the description of parallelism
+using fork-join structures as well as arbitrary branching \cite{Aa}.
+\item The {\bf Aa} description is translated to a virtual circuit (the model
+is described in the next section).  During this translation, the
+following major optimizations
+are performed:  declared storage objects are partitioned into disjoint memory
+spaces using pointer reference analysis, and dependency analysis is used to
+generate appropriate sequencing of operations in order to maximize the 
+parallelism.
+\item The virtual circuit is then translated to VHDL.  At this point,
+decisions about operator sharing are taken.  Concurrency analysis is
+used if a shared hardware unit needs arbitration.  The generated
+VHDL uses a pre-designed library of useful operators ranging from
+multiplexors, arbiters to pipelined floating point arithmetic units.
+\end{enumerate}
+
+The compiler flow has been characterized over a wide variety
+of applications \cite{dsd2010}.   The Click-to-NetFPGA tool-chain \cite{usenix2012} 
+uses this compiler flow to map modular network appliances written in C++ to
+an FPGA platform, and achieved about half the
+performance possible using equivalent hand designed hardware.  
+Thus,  the flow can be applied to ``real'' software
+to produce competitive results.  The optimizations indicated in this
+paper, should in principle, make the flow even more competitive.
 
@@
 is transformed through a series of steps by our compiler tools to
-produce the final hardware, which is depicted in Figure \ref{fig:dotp}.
+produce the final hardware, which is depicted in Figure \ref{fig:dotP}.
+\begin{figure*}[ht]
+  \centering
+  \includegraphics[width=15cm]{dotP.eps}
+  \caption{Control-data-storage hardware model.}
+  \label{fig:dotP}
+\end{figure*}
+The circuit in Figure \ref{fig:dotP} has three components,
+described below.
 
+\subsection{Control-path}
 %%
@@
 %%
-
-\subsection{Control-path}
 
@@
 %%
+The hardware model consists of a control-path (shown on the left
+in Figure \ref{fig:dotP}) which is modeled as a Petri-net with
+a unique entry point and a unique exit point.  The Petri-net
+is constructed using a set of production rules which guarantee
+liveness and safeness \cite{ahir}.  Transitions in the Petri-net
+are associated with output symbols to the data-path (these can
+be described by the regular expressions *sr and *cr)
+and input symbols from the data-path (these are of the form *sa and
+*ca).  The *sr symbols instruct an element in the data-path to
+sample its inputs and the *cr symbols instruct an element in the
+data-path to update its outputs (all outputs of data-path elements
+are registered).  The *sa and *ca symbols are acknowledgements
+from the data-path which indicate that the corresponding requests
+have been served.
 
 \subsection{Data-path}
-
 %%
@@
 %%
+The data-path is a directed hyper-graph.  The operators
+are shown as boxes and have input and output nets.  Each
+net has at most one element which drives it.  Further, most
+operators have  a split protocol handshake with
+the control-path:  two pairs of request/acknowledge 
+associations (*sr/*sa for sampling the inputs  and *cr/*ca for
+updating the outputs).    The sequencing is required to be
+\begin{verbatim}
+sr -> sa -> cr -> ca
+\end{verbatim}
 
-\subsection{Memory-subsystem}
+Some data-path elements (such as the multiplexor
+shown on the top and the decision operator shown at the bottom
+left in Figure \ref{fig:dotP} follow a simpler protocol.  The multiplexor
+has a pair of requests and a single acknowledge, with the condition
+that at most one of the requests is received at any time instant.
+The input corresponding to the request is then sampled and updates
+the output of the multiplexor.
+acknowledge lines.
+The decision element has a single request and two acknowledes.  Upon
+receipt of the request symbol, the decision element checks its input
+and emits one of the two acknowledges depending on whether the input
+is zero/non-zero.
+
+In figure \ref{fig:dotP}, the following data-path elements
+are instantiated:
+\begin{verbatim}
+mI, mdotP  multiplexors for I, dotP.
+INCR       incrementor for I++
+LA         load-operator for a[I]
+LB         load-operator for b[I]
+FMUL       multiplier for p=a[I]*b[I]
+FADD       adder for dotP += a*b
+CMP EQ     comparator COND=I==1023
+D          decision  COND?
+\end{verbatim}
+
+\subsection{Storage subsystem}
 
@@
 %%
+The load (and in general, store) operators in the data-path
+are associated with memory subsystems.  In general, there
+can be multiple disjoint memory subsystems inferred by our
+compiler.  In this particular case, the array a[] and b[] 
+are mapped to disjoint memories, due to which the two
+loads are allowed to proceed in parallel (the relaxed consistency
+model is enforced).
+In order to maintain the relaxed consistency model, the
+memory subsystems are designed to use a time-stamping 
+scheme which guarantees first-come-first-served access to
+the same memory location.
 
-\subsection{An example of a loop}
+\subsection{Dependency analysis and extraction of parallelism}
 
-%%
-%% for(i=0;i < 100; i++)
-%% {
-%%    c[i] = a[i] + b[i];
-%% }
-%%
+Our compiler does a routine dependency analysis and enforces
+the following dependencies between operations in the
+data-path (these dependencies are enforced by the control-path):
+\begin{itemize}
+\item RAW: if the wire updated by operation A is used in an
+operation B, the the sample-request to B is generated only
+after the update-acknowledge from A.
+\item WAR: if B writes to a wire which was earlier being read
+by A (in the same loop iteration), then the update-request to B
+can only be generated after the sample-acknowledge from A.
+\item Load-Store ordering: A relaxed consistency model
+is used.  If a load L follows a store S
+and both are associated with the same memory subsystem, then
+sample-start of the load L must begin after sample-acknowledge
+of S (note that the memory subsystem maintains the service order).
+Store-after-load and store-after-store dependencies are
+handled in the same manner.  Successive loads without
+an intervening store are allowed to finish in arbitrary order.
+\end{itemize}
 
-\section{Control-flow mechanism for loop-pipelining}
+The control-path in \ref{fig:dotP} shows the sequencing
+generated by these rules.  Note that the data-path
+is not party to any sequencing decisions.  The data-path
+elements must adhere to the request/acknowledge protocol
+defined by the control-path.
+
+\section{A control-flow mechanism for loop-pipelining}
 
@@
 %%
+Suppose that we want to modify the control-path in order to
+permit the second (and maybe third etc.) iteration of a loop
+to begin while the first iteration is still in progress.  
+The
 

### 2013-03-25 Documents/Presentations/eurasip/Talk.tex
eurasip paper.

@@
 Cycles       161K       77K         13K        7810
-LUTs         6323      9408       16032       31744
-FFs          6974      8818       15299       21060
+LUTs         6323      9408       14891       31744
+FFs          6974      8818       12041       21060
 Freq.(MHz)   199.7     199.7        186       164.1

### 2013-03-25 Documents/Publications/dsd2013/paper.tex
eurasip paper.

@@
 %% revisit the title.
-\title{A dynamic loop pipelining mechanism for the efficient realization of algorithms as hardware}
+\title{Inner loop optimizations in mapping single-threaded programs to hardware}
 
@@
   we consider the basic problem of generating an efficient hardware 
-  implementation of a loop.  We describe a control-flow mechanism through
-  which multiple iterations of an arbitrary loop can be dynamically pipelined in
-  the generated hardware,  and study the extent to which this 
-  mechanism affects the properties of the generated hardware.
-  We consider some common loop kernels occurring in fast-fourier transform,
-  matrix multiplication and inner product computations, and
-  observe that this dynamic pipelining mechanism typically results in
-  a doubling of the performance of the hardware.  When 
-  combined with static loop unrolling,  the performance improvement
-  ranges from 6X to 20X across the loop kernels that we have studied.
-  These optimizations do have a hardware overhead, however.  We characterize
-  the performance/cost trade-off using an FPGA target, and find
-  that the loop optimizations not only improve performance, but
-  also the cost/performance ratio of the resulting hardware.
+  implementation of a single threaded program, in particular, that
+  of an inner loop.  
+  We describe a control-flow mechanism through
+  which multiple iterations of an arbitrary inner loop can be made simultaneously
+  active in the generated hardware, thus providing dynamic loop-pipelining
+  capability in hardware.   
+  We study the impact of this loop-pipelining scheme in conjunction with
+  source-level loop-unrolling.
+  In particular, we consider some common loop kernels occurring in fast-fourier transform,
+  matrix multiplication and inner product computations, with the resulting
+  hardware descriptions synthesized to an FPGA target.  We observe
+  that the use of dynamic loop-pipelining mechanism alone typically results in
+  a significant improvements in the performance of the hardware.  If the loop is
+  statically unrolled and if loop-pipelining is applied to the unrolled
+  program, then the performance improvement is still substantial.  When
+  both optimizations are used in conjunction, the improvement in performance
+  ranges from 6X to 20X (in terms of number of clock cycles needed for the computation)
+  across the loop kernels that we have studied.
+  These optimizations do have a hardware overhead, but,
+  in spite of this, we observe that the joint use of these 
+  loop optimizations not only improves performance, but also the 
+  performance/cost ratio of the resulting hardware.
 
@@
 
+In our work, we have developed a set of compiler tools which take a C-description of
+an algorithm (with the only important restrictions being the requirement that
+cycles are not present in the call graph of the source program, and the lack of 
+support for {\bf function} pointers, normal pointers are supported), and produces
+a VHDL description which is provably equivalent to the original program \cite{SameerPhD}.
+Results on the application of this compiler to single-threaded programs show
+that the circuits (when synthesized to VLSI layout) are substantially more
+energy efficient than a processor, but that the performance of the processor
+is usually better than that of the synthesized circuit \cite{dsd2010}.
+If multi-threaded source programs are used as a starting point, the circuits produced 
+are quite competitive with hand-designed RTL \cite{usenix2012}.
+Quantitatively, the results in \cite{dsd2010} and \cite{usenix2012} indicate
+that the gap between the synthesized hardware and modern
+pipelined processors (for single-threaded
+programs) and between the synthesized hardware and hand-crafted hardware (for
+multi-threaded programs) is of the order of 2-3X.  That is, the performance
+of the synthesized hardware needs to improve by 2-3X in order to close
+the respective performance gaps.
+
 %% How compilers have handled this problem.
-In this paper, we consider the problem of mapping loops to hardware.  It
+In this paper, we consider the problem of improving the performance
+of hardware generated from single threaded programs; in
+particular, the problem of mapping loops to hardware.  It
 is well known that most compute intensive programs spend a large fraction
-of their time in inner loops.  Thus, the optimization of these loops is
-a primary problem for any optimizing compiler which is targetting
-either a deeply pipelined or parallel processor.  Several loop optimizations
+of their time in inner loops.  Thus, the optimal implementation
+of such loops is of primary importance, whether the target is
+a processor or hardware.  Such an improvement is essential if
+synthesized hardware is to be performance competitive with
+high performance processors or with hand-crafted hardware.
+
+In the context of compilation to a pipelined processor, 
+several loop optimizations
 have been considered in literature, such as loop-unrolling, loop-peeling,
@@
 these optimizations being the extraction of as much
-parallelism as possible from the source program.
+parallelism as possible from the single-threaded source program.
 
 %% Algorithm-to-hardware:  can similar techniques work?
-Several loop-optimization techniques have been explored and
+Similar loop-optimization techniques have been explored and
 reported in the literature related to reconfigurable hardware (for
@@
 in the work reported by \cite{Weinhardt},
-the loop optimizations are done in a manner analogous to the static
+loop optimizations are done in a manner analogous to the static
 techniques used in software compilers, in which index 
@@
 are analysed to identify dependencies and schedule operations
-in the loop body.  An explicitly timed controller is synthesized
+across iterations of the loop body.  An explicitly timed controller is synthesized
 for the pipeline.  Another approach which works in a similar
@@
 the loop, and the cases to which the approach can be applied
-are restricted (but still sufficiently general for most linear
+are restricted (but are still sufficiently general for most linear
 algebra and digital signal processing kernels).
 
-In our approach, the hardware model consists of a data-path (which is
-a graph of operators interconnected by wires) and an untimed control-path
-which is modeled as a Petri-net.  This representation allows the implementation
+In our approach, the hardware model is abstracted as
+a virtual circuit which consists of a data-path (a graph of operations 
+interconnected by wires) and a control-path
+which is modeled as a Petri-net.   The operations in the data-path are
+not tightly scheduled, with dependencies being taken care of
+by the control Petri-net (for example: operation $X$ can start only
+after operations $Y,Z$ have finished etc.). This representation allows the implementation
 of loop pipelining by a simple modification to the control Petri-net
-{\em without altering} the data-path.  Thus,
-it is possible to pipeline any loop, even those that do not have 
-explicit induction variables (such as while loops).
+{\em without altering} the data-path.  It is possible to pipeline any loop, 
+even those that do not have explicit induction variables (such as while loops).  
+We will describe this loop-pipelining mechanism in a later section in
+this paper.
 
@@
 optimization applied by itself and in conjunction with static loop-unrolling.
-By loop-unrolling, we mean a {\em static} compile-time optimization technique in
+By loop-unrolling, we mean a {\em static} source-level or compile-time optimization technique in
 which  an inner loop is unrolled by instantiating multiple copies of the loop-body
@@
 for(i=0; i < 4; i+=2) {
+   int i1 = i+1; int i2 = i+2; int i3 = i+3;
    x += a[i]*b[i];
-   x += a[i+1]*b[i+1];
+   x += a[i1]*b[i1];
+   x += a[i2]*b[i2];
+   x += a[i3]*b[i3];
 }
@@
 statements without any branches), and provides the possibility of extracting more
-parallelism in the loop.
+parallelism in the loop.  Note that this unrolling can be done manually
+by the programmer, or automatically by an optimizing compiler.
 
@@
 the corresponding improvement in performance that results from this optimization.
+In the results presented in this paper, the unrolling has been done 
+manually at the source code level.
 
@@
 For matrix multiplication, this improvement is as large as 20X.
-\item The hardware overheads for implementing these optimizations are substantial,
+\item The hardware overheads for implementing these optimizations are considerable,
 but the cost-to-performance ratio improves substantially in all cases.
 \end{itemize}
+The results indicate that loop-unrolling combined with dynamic loop-pipelining
+can close the performance gap noted above.
+
+For the subsequent discussion, we assume that the inner
+loop which is being optimized consists of a single
+basic block, that is, there are no branching instructions
+in the loop body (no jumps, if constructs, switch constructs etc.).
+If such constructs are present, we assume that they have been
+eliminated using the mechanism of guarded (or predicated) execution.
 
@@
 decisions about operator sharing are taken.  Concurrency analysis is
-used if a shared hardware unit needs arbitration.  The generated
+used to determine if a shared hardware unit needs arbitration.  Optimizations
+related to clock-frequency maximization are also carried out here.
+The generated
 VHDL uses a pre-designed library of useful operators ranging from
@@
 
-\section{Model of hardware generated by our compiler}
+\section{Model of the virtual circuit generated by our compiler}
 
-The hardware generated by our compiler consists of three
+The virtual circuit generated by our compiler consists of three
 cooperating components: the control-path, the data-path and
@@
   \includegraphics[width=15cm]{dotP.eps}
-  \caption{Control-data-storage hardware model.}
+  \caption{Control-data-storage virtual circuit model.}
   \label{fig:dotP}
 \end{figure*}
-The circuit in Figure \ref{fig:dotP} has three components,
+The virtual circuit in Figure \ref{fig:dotP} has three components,
 described below.
+
+\subsection{Data-path}
+%%
+%% Directed hyper-graph.  Nodes are operators
+%% hyper-arcs are wires.  Each operator node
+%% has a split protocol.
+%%
+The data-path is a directed hyper-graph with nodes being
+operations and arcs being nets (shown as ovals).  Each
+net has at most one operation which drives it.  Further, most
+operations have  a split protocol handshake with
+the control-path:  two pairs of request/acknowledge 
+associations (*sr/*sa for sampling the inputs  and *cr/*ca for
+updating the outputs).    The operation samples its inputs
+on receiving the sr request symbol and acknowledges the completion
+of this action by emitting the sa acknowledge symbol.  After receiving
+the cr symbol, the operation will update its output net
+using the newly computed value. The sequencing is required to be
+\begin{verbatim}
+sr -> sa -> cr -> ca
+\end{verbatim}
+Note that an operation can be re-triggered while an earlier
+edition of the operation
+is in progress (this is important if the operation is implemented
+in a pipelined operator).
+
+Some data-path operations (such as the multiplexor
+shown on the top and the decision operation shown at the bottom
+left in Figure \ref{fig:dotP} follow a simpler protocol.  The multiplexor
+has a pair of requests and a single acknowledge, with the condition
+that at most one of the requests is received at any time instant.
+The input corresponding to the request is then sampled and stored
+in the output net of the multiplexor.
+The decision operation has a single request and two acknowledes.  Upon
+receipt of the request symbol, the decision operation checks its input net
+and emits one of the two acknowledges depending on whether the input
+is zero/non-zero.
+
+In Figure \ref{fig:dotP}, the following data-path operations
+are instantiated:
+\begin{verbatim}
+mI, mdotP  multiplexors for I, dotP.
+INCR       increment for I++
+LA         load for a[I]
+LB         load for b[I]
+FMUL       multiply for p=a[I]*b[I]
+FADD       add for dotP += a*b
+CMP EQ     compare for COND=(I==1023)
+D          decision  COND?
+\end{verbatim}
+
+Note that the data-path only shows the operations and their interconnection.
+When the data-path is implemented as hardware, multiple operations may
+be mapped to a single operator depending on cost/performance tradeoffs.  When
+this is done, multiplexing logic has to be introduced in the hardware.  These
+decisions and manipulations are performed in the compiler stage which is
+responsible for transforming the virtual circuit to VHDL.
+
+\subsection{Storage subsystem}
+
+%%
+%% time-stamping scheme, in order completion.
+%%
+The load and store operations in the data-path
+are associated with memory subsystems.  In general, there
+can be multiple disjoint memory subsystems inferred by our
+compiler.  In this particular case, the arrays a[] and b[] 
+are mapped to disjoint memories, due to which the two
+loads are allowed to proceed in parallel (the relaxed consistency
+model is enforced).
+In order to maintain the relaxed consistency model, the
+memory subsystems are designed to use a time-stamping 
+scheme which guarantees first-come-first-served access to
+the same memory location.
 
@@
 %%
-The hardware model consists of a control-path (shown on the left
-in Figure \ref{fig:dotP}) which is modeled as a Petri-net with
+The control-path in the virtual circuit encodes
+all the sequencing that is necessary for correct
+operation of the assembly.
+The control-path (shown on the left
+in Figure \ref{fig:dotP}) is modeled as a Petri-net with
 a unique entry point and a unique exit point.  The Petri-net
@@
 
-\subsection{Data-path}
-%%
-%% Directed hyper-graph.  Nodes are operators
-%% hyper-arcs are wires.  Each operator node
-%% has a split protocol.
-%%
-The data-path is a directed hyper-graph.  The operators
-are shown as boxes and have input and output nets.  Each
-net has at most one element which drives it.  Further, most
-operators have  a split protocol handshake with
-the control-path:  two pairs of request/acknowledge 
-associations (*sr/*sa for sampling the inputs  and *cr/*ca for
-updating the outputs).    The sequencing is required to be
+The following classes of dependencies are
+encoded in the control Petri-net:
+\begin{itemize}
+\item Read-after-write (RAW):  If the result of operator A is used
+as an input to operator B, the sr symbol to B can be emitted only after
+the ca symbol from A has been received.
+\item Write-after-read (WAR): If B writes to a net whose value needs
+to have been used by A earlier, for example as in
 \begin{verbatim}
-sr -> sa -> cr -> ca
+a = (b+c) -- operation A reads c
+c = (p*q) -- operation B writes to c
 \end{verbatim}
-
-Some data-path elements (such as the multiplexor
-shown on the top and the decision operator shown at the bottom
-left in Figure \ref{fig:dotP} follow a simpler protocol.  The multiplexor
-has a pair of requests and a single acknowledge, with the condition
-that at most one of the requests is received at any time instant.
-The input corresponding to the request is then sampled and updates
-the output of the multiplexor.
-acknowledge lines.
-The decision element has a single request and two acknowledes.  Upon
-receipt of the request symbol, the decision element checks its input
-and emits one of the two acknowledges depending on whether the input
-is zero/non-zero.
-
-In figure \ref{fig:dotP}, the following data-path elements
-are instantiated:
-\begin{verbatim}
-mI, mdotP  multiplexors for I, dotP.
-INCR       incrementor for I++
-LA         load-operator for a[I]
-LB         load-operator for b[I]
-FMUL       multiplier for p=a[I]*b[I]
-FADD       adder for dotP += a*b
-CMP EQ     comparator COND=I==1023
-D          decision  COND?
-\end{verbatim}
-
-\subsection{Storage subsystem}
-
-%%
-%% time-stamping scheme, in order completion.
-%%
-The load (and in general, store) operators in the data-path
-are associated with memory subsystems.  In general, there
-can be multiple disjoint memory subsystems inferred by our
-compiler.  In this particular case, the array a[] and b[] 
-are mapped to disjoint memories, due to which the two
-loads are allowed to proceed in parallel (the relaxed consistency
-model is enforced).
-In order to maintain the relaxed consistency model, the
-memory subsystems are designed to use a time-stamping 
-scheme which guarantees first-come-first-served access to
-the same memory location.
-
-\subsection{Dependency analysis and extraction of parallelism}
-
-Our compiler does a routine dependency analysis and enforces
-the following dependencies between operations in the
-data-path (these dependencies are enforced by the control-path):
-\begin{itemize}
-\item RAW: if the wire updated by operation A is used in an
-operation B, the the sample-request to B is generated only
-after the update-acknowledge from A.
-\item WAR: if B writes to a wire which was earlier being read
-by A (in the same loop iteration), then the update-request to B
-can only be generated after the sample-acknowledge from A.
-\item Load-Store ordering: A relaxed consistency model
-is used.  If a load L follows a store S
-and both are associated with the same memory subsystem, then
-sample-start of the load L must begin after sample-acknowledge
-of S (note that the memory subsystem maintains the service order).
-Store-after-load and store-after-store dependencies are
-handled in the same manner.  Successive loads without
-an intervening store are allowed to finish in arbitrary order.
+where there is a WAR dependency through c,
+ then the cr request to B can be
+issued only after the sa acknowledge from A has been received.
+\item Load-Store ordering:  If P,Q are load/store operations
+to the same memory subsystem, and if at least one of P,Q is a
+store, and if P is supposed to happen before Q,  then the sr request
+to Q must be emitted only after the sa acknowledge from Q
+has been received.  The memory subsystem itself guarantees
+that requests finish in the same order that they were
+initiated.  This takes care of WAR, RAW and WAW memory
+dependencies.
 \end{itemize}
@@
 generated by these rules.  Note that the data-path
-is not party to any sequencing decisions.  The data-path
-elements must adhere to the request/acknowledge protocol
-defined by the control-path.
+is not party to any sequencing decisions (other than
+responding to the request symbols).
 
@@
 to begin while the first iteration is still in progress.  
-The
+Consider the example of the dot product.  The original
+loop was
+\begin{verbatim}
+float a[1024], b[1024];
+float dotp = 0.0;
+for(i=0; i < 1024; i++)
+{
+   dotp += a[i]*b[i];
+}
+\end{verbatim}
+The {\em fully unrolled} version of this loop would be
+\begin{verbatim}
+float a[1024], b[1024];
+float dotp = 0.0;
+dotp += a[0]*b[0];
+dotp += a[1]*b[1];
+dotp += a[2]*b[2];
+...
+dotp += a[1023]*b[1023];
+\end{verbatim}
+Let $A$ denote the $*$ operation and let $B$ denote the $+$
+operation, $L_a,\ L_b$ denote the loads from $a$ and $b$ respectively.  
+In principle, all the loads can occur simultaneously, and all the multiplies
+can happen simultaneously once the loads complete.  The adds
+would need to be ordered because of the multiply-accumulate nature
+of the code as it is written.  Any ordering of these operations which
+satisfies these dependencies will be termed a {\em fully-unrolled consistent}
+ordering.
+
+In our proposed, loop pipelining scheme, we use the
+following ordering scheme.
+If A is an operation in the loop body,
+denote the
+$k^{th}$ execution of A by $A_k$.  Since each operation
+has  events sr, sa, cr, ca, we denote these by
+$A_k.sr$, $A_k.sa$, $A_k.cr$, $A_k.ca$ respectively. We impose 
+the following dependency rules on operations across
+loop iterations.
+\begin{itemize}
+\item $A_k.sa \rightarrow A_{k+1}.sr$ for all operations A: that is,
+the next execution of A cannot start until the current execution has
+finished sampling the inputs.
+\item $A_k.ca \rightarrow A_{k+1}.cr$ for all operations A: that is,
+the completion of the next execution of A can be initiated only
+after the current execution of A has completed.
+\item If $A \rightarrow B$ is a RAW dependency, then 
+$B_k.sa \rightarrow A_{k+1}.cr$.  That is, until B has
+sampled the result of the current A, the next completion
+of A cannot start.
+\item If $A \rightarrow B$ is a WAR dependency, then
+$B_k.ca \rightarrow A_{k+1}.sr$.  That is, the next
+A cannot start until the current B has completed.
+\item If P, Q are successive load/stores, with at least
+one of them being a store, then 
+$Q_k.sa \rightarrow P_{k+1}.sr$.  That is, the next P
+cannot start until the current Q has acknowledged that it
+has started.
+\end{itemize}
+The mechanism for incorporating RAW and WAR dependencies is
+illustrated in Figure \ref{fig:pipelineMechanism}
+for RAW and WAR dependencies within the loop body.
+The reverse dotted arc is a marked arc (it initially
+carries a single token).
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=7cm]{pipelineMechanism.eps}
+  \caption{Control-path mechanism for handling RAW, WAR dependencies for loop-pipelining.}
+  \label{fig:pipelineMechanism}
+\end{figure}
+
+It is easy to confirm that these additional dependencies ensure
+that the loop execution subject to these dependencies is a
+fully-unrolled consistent ordering.  
+The modified control path is shown in Figure \ref{fig:pipelinedCP}.
+The loop-terminator element has three inputs:  a loop-taken transition,
+a loop-not-taken transition and a loop-body-exit transition.  The loop-taken/not-taken
+pair indicates whether a new iteration is to be started or whether the 
+loop has terminated.  The loop-body-exit
+transition indicates that the body of the loop has finished executing an
+iteration.  The loop-terminator initiates a new iteration as long as the
+number of active iterations is within a specified limit $M$ (usually, we keep
+this limit to $M=8$).  Thus, all the places in the modified control path
+in Figure \ref{fig:pipelinedCP} now must have a capacity of $M$ and the
+cost of implementing each place in the control path goes up by a
+factor of $\log M$.  This is the major additional cost incurred by
+the pipelining mechanism.
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=7cm]{pipelinedCP.eps}
+  \caption{Modified control-path with loop-pipeline dependencies (dotted lines).}
+  \label{fig:pipelinedCP}
+\end{figure}
 
@@
 
-%%
-%% examples: vector add, dot-product, matrix-multiply, linfinity-norm.
-%%
+We considered three examples, the dot product, the fast-fourier-transform,
+and matrix multiplications, each with critical inner loops.  Four 
+configurations were tested in each case: the basic code, the unrolled code,
+the basic code with loop-pipelining and the unrolled code with loop-pipelining.
+In each case, the generated VHDL code was synthesized to a Xilinx Virtex-6
+FPGA and synthesis results were used to estimate the resource usage,
+the clock frequency and the number of cycles required by the inner
+loop.
+
+\section{The dot product}
+
+The basic code (a,b are arrays, dotP is the accumulated dot-product):
+\begin{verbatim}
+for(I=0; I < 64; I++)
+{
+  dotP += a[I]*b[I];
+}
+\end{verbatim}
+
+The unrolled version of the code used:
+\begin{verbatim}
+for(I=0; I < 64; I += 4)
+{
+  I1 = I+1; I2 = I+2; I3 =I+3;
+  dotP += ((a[I]*b[I]) + (a[I1]*b[I1]) 
+           + (a[I2]*b[I2]) + (a[I3]*b[I3]));
+}
+\end{verbatim}
+
 %%
@@
 %%
+The results are shown in Table \ref{table:dotP}.   The number of clock-cycles
+needed to complete the inner loop, the number of look-up tables needed, the
+ number of flip-flops needed and the post-synthesis clock frequency estimate
+are reported in the table.  The last two rows correspond to the normalized
+performance (time needed by the plain case relative to the optimized case, higher is better),
+and the normalized performance/cost ratio (time/(LUTs+FF)  ratio normalized
+with respect to the plain case, higher is better).
+\begin{table}[htb]
+  \centering
+  \caption{Dot-product results with and without loop-optimizations}
+  \label{table:dotP}
+  \renewcommand\arraystretch{1.2}
+  \setlength{\tabcolsep}{1ex}
+  \begin{tabular}{c|c|c|c|c}
+  \hline
+  & plain & pipelined & unrolled & pipelined  \\
+  &       &           &          & + unrolled\\
+  \hline
+
+Cycles   &    4071  &   1874  &     1894  &     582 \\
+LUTs     &   4288   &   5195  &     4809  &    7344 \\
+FFs      &    3799  &   4345  &     4378  &    5911 \\
+Freq.(MHz)  & 199.7  &    199.7 &      199.7 &    199.7 \\
+Norm. Perf. & 1      &  2.17  & 2.15 & 7 \\
+Norm. Perf/Cost & 1  &  1.83  & 1.89 & 4.27 \\
+\hline
+  \end{tabular}
+\end{table}
+
+From Table \ref{table:dotP}, we see 
+\begin{itemize}
+\item If loop-pipelining is 
+applied to the plain program, performance improves
+by about 2X relative to the non-pipelined, plain case.
+\item If loop-pipelining is applied to the unrolled program,
+performance improves by more than 3X relative to the non-pipelined,
+unrolled case.  
+\item In terms of the performance/cost
+ratio, the pipelined-unrolled version is more than 3X better than
+the plain version.   The normalized performance/cost ratio is 4X
+better when both loop optimizations are used.
+\end{itemize}
+
+\section{The fast-fourier-transform (FFT)}
+
+A 64 point FFT program (radix two, in-place, twiddle factors computed
+apriori) with the following loop-structure was used:
+\begin{verbatim}
+for(STAGE=0; STAGE < 6; STAGE++)
+{
+  for(BFLY=0; BFLY < 32; BFLY++)
+  {
+    Butterfly(STAGE,BFLY);
+  }
+}
+\end{verbatim} 
+where the function {\bf Butterfly} is inlined and implements
+the radix-two butterfly with index BFLY in stage STAGE using
+the precomputed twiddle factors.
+In the unrolled version, the inner-loop was rewritten as
+\begin{verbatim}
+  for(BFLY=0; BFLY < 32; BFLY += 4)
+  {
+    Butterfly(STAGE,BFLY);
+    Butterfly(STAGE,BFLY+1);
+    Butterfly(STAGE,BFLY+2);
+    Butterfly(STAGE,BFLY+3);
+  }
+\end{verbatim}
+
+The results are shown in Table \ref{table:fft} (the cycle count is for a single
+stage of the FFT).
+\begin{table}[htb]
+  \centering
+  \caption{FFT-results  with and without loop-optimizations}
+  \label{table:fft}
+  \renewcommand\arraystretch{1.2}
+  \setlength{\tabcolsep}{1ex}
+  \begin{tabular}{c|c|c|c|c}
+  \hline
+  & plain & pipelined & unrolled & pipelined  \\
+  &       &           &          & + unrolled\\
+  \hline
+
+Cycles   &    4151  &   2885   &     3110  &     1064 \\
+LUTs     &   12155   &  23139  &     16032  &    37831 \\
+FFs      &    11955  &  18632  &     15299  &    28698 \\
+Freq.(MHz)  & 186.9  &    164.1 &      186.9 &    164.7 \\
+Norm. Perf. & 1      &  1.26  & 1.33 & 3.42 \\
+Norm. Perf/Cost & 1  &  0.73  & 1.02 & 1.24 \\
+\hline
+  \end{tabular}
+\end{table}
+From Table \ref{table:fft}, we see 
+\begin{itemize}
+\item If loop-pipelining is 
+applied to the plain program, performance improves
+by about 1.26X relative to the non-pipelined, plain case.
+\item If loop-pipelining is applied to the unrolled program,
+performance improves by more than 2.5X relative to the non-pipelined,
+unrolled case.  
+\item In terms of the performance/cost
+ratio, the pipelined-unrolled version is only 1.24X better than
+the plain version.   
+\end{itemize}
+The reason for the poorer results in this case is the use of the in-place
+algorithm.  The bottleneck in this case becomes the access to the memory
+subsystem in which the array is stored.
+
+\section{Matrix multiplication}
+
+The plain triple loop matrix multiplication algorithm was
+used as a starting point.
+\begin{verbatim}
+float a[16]16], b[16][16], c[16][16];
+for(i = 0; i < 16; i++) {
+  for(j=0; j < 16; j++) {
+    float v = 0.0;
+    for(k = 0; k < 16; k++) {
+       v += a[i][k]*b[k][j];
+    }
+    c[i][j] = v;
+  }
+}
+\end{verbatim}
+The unrolled version was aggressively generated so that
+the inner loop simultaneously computes 16 entries of the
+product at a time.
+\begin{verbatim}
+float a[16]16], b[16][16], c[16][16];
+for(i = 0; i < 16; i += 4) {
+  for(j=0; j < 16; j += 4) {
+    float v00 = 0.0, v01 = 0.0,
+          v02 = 0.0, ... v33 = 0.0;
+    for(k = 0; k < 16; k += 4) {
+       v00 += (a[i][k]*b[k][j] 
+             + a[i][k+1]*b[k+1][j]
+             + a[i][k+2]*b[k+2][j]
+             + a[i][k+3]*b[k+3][j]);
+       ...
+       v33 += (a[i+3][k]*b[k][j+3] 
+             + a[i+3][k+1]*b[k+1][j+3]
+             + a[i+3][k+2]*b[k+2][j+3]
+             + a[i+3][k+3]*b[k+3][j+3]);
+    }
+    c[i][j]   = v00;
+    ..
+    c[i+3][j+3] = v33;
+  }
+}
+\end{verbatim}
+
+The observations are shown in Table \ref{table:mmultiply}.
+\begin{table}[htb]
+  \centering
+  \caption{Matrix-multiplication-results  with and without loop-optimizations}
+  \label{table:mmultiply}
+  \renewcommand\arraystretch{1.2}
+  \setlength{\tabcolsep}{1ex}
+  \begin{tabular}{c|c|c|c|c}
+  \hline
+  & plain & pipelined & unrolled & pipelined  \\
+  &       &           &          & + unrolled\\
+  \hline
+
+Cycles   &    161K  &   77K   &     13K  &     7810 \\
+LUTs     &   6323   &  9408  &     14891  &    31744 \\
+FFs      &    6974  &  8818  &     12041  &    21060 \\
+Freq.(MHz)  & 199.7  &   199.7 &   186.1 &    164.1 \\
+Norm. Perf. & 1      &  2.09  & 11.5 & 16.9 \\
+Norm. Perf/Cost & 1  &  1.52  & 5.67 & 4.25 \\
+\hline
+  \end{tabular}
+\end{table}
+From Table \ref{table:fft}, we see 
+\begin{itemize}
+\item If loop-pipelining is 
+applied to the plain program, performance improves
+by about 2X relative to the non-pipelined, plain case.
+\item If loop-pipelining is applied to the unrolled program,
+performance improves by 1.5X relative to the non-pipelined,
+unrolled case.  
+\item In terms of the performance/cost
+ratio, the pipelined-unrolled version is only 4.25X better than
+the plain version.   
+\end{itemize}
+In this instance, the aggressive loop-unrolling shows
+excellent performance.  Loop-pipelining when combined with loop-unrolling,
+gives a 20X improvement in the cycle count.  The normalized
+performance and performance/cost improvements are also substantial.
 
@@
 %%
+We have considered the problem of optimizing inner loop
+implementations in an algorithm-to-hardware compilation system.
+Two optimizations were considered: static source-level loop
+unrolling and dynamic hardware supported loop-pipelining.
+The loop-pipelining mechanism is implemented by modifying
+the control-flow in the generated hardware (without disturbing
+the data-path).
+
+The data obtained from three common and important inner loop
+kernels is encouraging.  In all three cases, both loop-pipelining
+and loop-unrolling lead to substantial performance gains.  Further,
+using both optimizations together results in multiplicative gains
+and in call cases, leads to hardware which is substantially
+faster and more efficient (in terms of the performance/cost ratio).
+The performance gain is lower if there is a bottleneck in the
+algorithm itself, such as in the FFT case, in which accesses to
+the in-place array reduce the performance gains seen due to
+the loop optimizations.
+
+Loop-unrolling is a common parallelization optimization
+in compilers that target pipelined processors.  Our data indicates
+that such unrolling offers significant benefits (up to an 
+order of magnitude) in an algorithm-to-hardware translation.
+Further, the use of hardware based dynamic loop-pipelining techniques 
+seems to offer a further boost in performance in the synthesis
+of hardware from single-threaded programs.  
+The performance boost provided by the dynamic loop-pipelining
+in hardware seems to indicate that its use,
+especially in conjunction with aggressive loop unrolling can offer 
+a substantial reduction in the gap between 
+the quality of automatically generated hardware and hand crafted 
+hardware implementations of the same algorithm.  This needs
+to be investigated further.
 

### 2013-03-27 Documents/Publications/dsd2013/paper.tex
Updated, ready to submit.

@@
 
-\author{Madhav P. Desai\\
-  Indian Institute of Technology -- Bombay, Powai, Mumbai -- 400076, INDIA\\
-  Email: madhav@ee.iitb.ac.in }
+\author{Author's name Suppressed for Blind Review}
 \newtheorem{definition}{Definition}[section]
@@
 energy efficient than a processor, but that the performance of the processor
-is usually better than that of the synthesized circuit \cite{dsd2010}.
+is usually better than that of the synthesized circuit \cite{ahirDsd2010}.
 If multi-threaded source programs are used as a starting point, the circuits produced 
-are quite competitive with hand-designed RTL \cite{usenix2012}.
-Quantitatively, the results in \cite{dsd2010} and \cite{usenix2012} indicate
+are quite competitive with hand-designed RTL \cite{ahirUsenix2012}.
+Quantitatively, the results in \cite{ahirDsd2010} and \cite{ahirUsenix2012} indicate
 that the gap between the synthesized hardware and modern
@@
 have been considered in literature, such as loop-unrolling, loop-peeling,
-software-loop-pipelining etc. \cite{wolfe}, \cite{muchnick}, with the intent of
+software-loop-pipelining etc. \cite{Wolfe}, \cite{Muchnick}, with the intent of
 these optimizations being the extraction of as much
@@
 reported in the literature related to reconfigurable hardware (for
-example \cite{Weinhardt}, \cite{Cardoso}, \cite{Kastner}).  For instance,
-in the work reported by \cite{Weinhardt},
+example \cite{WeinhardtLuk}, \cite{Cardoso}, \cite{KastnerPhd}).  For instance,
+in the work reported by \cite{WeinhardtLuk},
 loop optimizations are done in a manner analogous to the static
@@
 details are not relevant to this paper, and the interested
-reader can find them in \cite{dsd2010}.
+reader can find them in \cite{ahirDsd2010}.
 
@@
 The compiler flow has been characterized over a wide variety
-of applications \cite{dsd2010}.   The Click-to-NetFPGA tool-chain \cite{usenix2012} 
+of applications \cite{ahirDsd2010}.   The XXXX\footnote{name suppressed for blind review.} tool-chain \cite{ahirUsenix2012} 
 uses this compiler flow to map modular network appliances written in C++ to
@@
 cooperating components: the control-path, the data-path and
-the storage system \cite{ahir}
+the storage system \cite{ahirDsd2010}.
 
@@
 \end{verbatim}
-To compile this code, we use the clang-2.8 \cite{clang} C compiler, 
-which is used to emit LLVM \cite{llvm} byte-code.  The LLVM byte-code
+To compile this code, we use the clang-2.8  C compiler, 
+which is used to emit LLVM byte-code.  The LLVM byte-code
 is transformed through a series of steps by our compiler tools to
-produce the final hardware, which is depicted in Figure \ref{fig:dotP}.
+produce a virtual circuit, which is depicted in Figure \ref{fig:dotP}.
 \begin{figure*}[ht]
@@
 shown on the top and the decision operation shown at the bottom
-left in Figure \ref{fig:dotP} follow a simpler protocol.  The multiplexor
+left in Figure \ref{fig:dotP}) follow a simpler protocol.  The multiplexor
 has a pair of requests and a single acknowledge, with the condition
@@
 
+\noindent
+{\bf Remark}
 Note that the data-path only shows the operations and their interconnection.
@@
 be mapped to a single operator depending on cost/performance tradeoffs.  When
-this is done, multiplexing logic has to be introduced in the hardware.  These
+this is done, multiplexing logic is introduced in the hardware.  These
 decisions and manipulations are performed in the compiler stage which is
@@
 is constructed using a set of production rules which guarantee
-liveness and safeness \cite{ahir}.  Transitions in the Petri-net
+liveness and safeness \cite{ahirDsd2010}.  Transitions in the Petri-net
 are associated with output symbols to the data-path (these can
@@
 
-In our proposed, loop pipelining scheme, we use the
+In our dynamic loop pipelining scheme, we use the
 following ordering scheme.

### 2013-05-11 v2/llvmbc2Aa/docs/llvm2aa.tex
updated the pipe_dispatch example to use my_strcmp to do string matches..
Also, added comment line option in modules.txt
and updated llvm2aa documentation
updated documentation to reflect changes.
Committing changes to llvm2aa for suppression of do-while extraction when a special function is called in the loop.

@@
 which should be converted to \Aa.   The names of these functions should be
-listed in the text-file listfile. If this option is not specified, all functions
+listed in the text-file listfile.  One function name must be present
+on each line (the {\bf exact} function name should be provided on the line, without leading
+or trailing spaces).  Lines which start with the character '\#' are
+ignored as comments.  If this option is not specified, all functions
 are converted.  
@@
 of any pipe is $1$.
+\item {\bf -extract\_do\_while=true}: If specified, mark inner loops
+as do-while loops, whose implementation will be
+pipelined.  An inner loop is a basic block whose terminator
+statement has a branch back to the beginning of the basic block. 
+This optimization is suppressed if the inner loop body contains a
+call to the special function {\bf nooptimize}.  If not specified,
+inner loops are not marked as do-while loops.
 \item {\bf llvm2aa} uses the LLVM compiler

### 2013-08-26 v2/Ahir-asm/devel/src/AaExpressionWriteVCOptimized.cpp
Some more bugs, simple processor seems to run (for a while) at 3-cycle latency until dependencies catch up!

@@ -123,6 +123,7 @@ void AaExpression::Write_VC_Guard_Dependency(bool pipeline_flag, set<AaRoot*>& v
 {
 	if(this->Get_Guard_Expression() != NULL)
 	{
+		ofile << "// Guard dependency" << endl;
 		AaExpression* expr = this->Get_Guard_Expression();
 		if(expr->Is("AaSimpleObjectReference"))
 		{
@@ -131,6 +132,7 @@ void AaExpression::Write_VC_Guard_Dependency(bool pipeline_flag, set<AaRoot*>& v
 			if(visited_elements.find(root) != visited_elements.end())
 			{
 				__J(__SST(this), __UCT(root));
+				__J(__UST(this), __UCT(root));
 			}
 		}
 		else
@@ -143,6 +145,8 @@ void AaExpression::Write_VC_Guard_Dependency(bool pipeline_flag, set<AaRoot*>& v
 			// when this completes, the guard can be re-evaluated.
 			__MJ(expr->Get_VC_Reenable_Update_Transition_Name(visited_elements),
 				 __UCT(this));
+			__MJ(expr->Get_VC_Reenable_Update_Transition_Name(visited_elements),
+				 __SCT(this));
 		}
 	}
 }

### 2013-09-03 v2/Ahir-asm/devel/docs/AaLRM.tex
Some more bugs, simple processor seems to run (for a while) at 3-cycle latency until dependencies catch up!
Simple processor works ok.  Time to try the extreme pipelining here to figure out if any benefits accrue...

@@
 Here $p,q,r,s,t$ can be expressions.
+
+Complete atomic statements can have a guard predicate
+specified, so that they are executed only if the guard
+predicate is true.  This is specified either by
+\begin{verbatim}
+$guard ( <guard-variable> )  <statement-spec>
+\end{verbatim}
+or 
+\begin{verbatim}
+$guard ( ~ <guard-variable> )  <statement-spec>
+\end{verbatim}
+For example
+\begin{verbatim}
+$guard (sel_add) r := (a + b)
+\end{verbatim}
+or
+\begin{verbatim}
+$guard (~ sel_add) $call afun (a b) (r)
+\end{verbatim}
 
@@
 
-\subsection{Pipelinable-loops: the {\bf do-while} construct}
+\subsection{Pipelinable-loops: the {\bf dopipeline-while} construct}
 
@@
 \begin{verbatim}
-  $do
+  $dopipeline
      $merge $entry $loopback
@@
 Dependency analysis on values in the loop-body is used to determine the
-extent of loop pipelining.  Note that it is always possible to describe
+extent of loop pipelining.  While doing the pipelining, one can specify
+two parameters as follows
+\begin{verbatim}
+  $dopipeline $depth 8 $buffering 2
+    ...
+  $while ....
+\end{verbatim}
+This specifies that up to 8 iterations of the loops will be kept alive
+in hardware, and the amount of buffering per operation will be kept to
+2.
+
+Note that it is always possible to describe
 a do-while loop using a branch-block, and this is what should be done

### 2013-09-03 v2/Ahir-asm/devel/docs/AaLRM.tex
Added a comment to highlight the early variable feedback issue in the pipelining of loops.

@@
 a do-while loop using a branch-block, and this is what should be done
-if loop-pipelining is not desired.
+if loop-pipelining is not desired\footnote{There are still some issues
+with loop-pipelining which are still being sorted out.}.
+
+\subsubsection{Issues with loop-pipelining (in progress)}
+
+There is one outstanding issue: loop run-away.  This issue arises
+when an early value is fed back into the loop body which uses delayed
+versions of this early value.  The loop races ahead, priming the
+delayed versions with junk.  The solution is to always feed back the
+most delayed version of a variable.
 

### 2013-09-23 v2/Ahir-asm/devel/docs/AaLRM.tex
Document updated.

--- 2013-09-03 v2/Ahir-asm/devel/docs/AaLRM.tex
@@
 the do-while from a predecessor in the branch-block, and stays in the
-do-while as long as the condition-expression is false.  The \$loopback
+do-while as long as the condition-expression is false (the test is done
+every time the loop body has finished).  The \$loopback
 place is used to repeat the loop-body.  The leading merge statement
@@
 
+Additionally, it is possible to specify that the loop-pipelining
+be done to its extreme possibility, that is, to the maximum possible
+rate that can be achieved.
+\begin{verbatim}
+  $dopipeline $depth 8 $buffering 2
+  $fullrate // specify that extreme pipelining 
+	    // should be attempted.
+    ...
+  $while ....
+\end{verbatim}
+
 Note that it is always possible to describe
@@
 
-Thus, an \Aa program is constructed as a collection of modules,
+To summarize, an \Aa program is constructed as a collection of modules,
 each of which is a sequence of statements.  The use of series,

### 2014-01-25 Documents/Tutorials/Intro/Intro.tex
Added a tutorial for making the whole edifice a bit more accessible.

\documentclass{beamer}
\usepackage{epsfig}
\title{An introduction to the AHIR-v2 tool-set}
\author{Madhav P. Desai}
\begin{document}
\maketitle

\frame[containsverbatim]{\frametitle{The purpose of the toolset}

\begin{itemize}
\item Aid in the design and implementation of complex systems (in hardware).
\item Describe the algorithm (behaviour) of the system using high-level
languages (currently C), and migrate the algorithm to a hardware
implementation.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{For example}
Design a circuit that takes two inputs, computes the maximum of
the two and returns the maximum.
\begin{verbatim}
uint32_t maxOfTwo(uint32_t a, uint32_t b)
{
  uint32_t c = ((a > b) ? a : b);
}
\end{verbatim}
}

\frame[containsverbatim]{\frametitle{Lets write a test-program to check this code}
\begin{verbatim}
uint32_t maxOfTwo(uint32_t, uint32_t);
int main(int argc, char* argv[])
{
  if(argc < 3)
  {
     printf(stderr, "%s <uint32_t> <uint32_t>\n", argv[0]);
     return(1);
  }
  uint32_t c = maxOfTwo(atoi(argv[1]), atoi(argv[2]));
  printf(stdout,"Result = %d.\n", c);
  return(0);
}
\end{verbatim}
}

\frame[containsverbatim]{\frametitle{We compile and run it}

Go to ex1/ and check it out.

}

\frame[containsverbatim]{\frametitle{Conversion to hardware}

The AHIR-v2 flow translates the code to hardware in four steps:
\begin{enumerate}
\item The source C code is compiled using the open-source compiler
{\bf clang} (we version 2.8), which produces byte-code in {\bf llvm}
format.
\item We take the {\bf llvm} bytecode and translate it to our internal
representation called {\bf Aa}.  This is done by a utility called {\bf llvm2aa}.
\item Multiple {\bf Aa} files can be linked using an {\bf Aa} linker
tool called {\bf AaLinkExtMem}.
\item The linked {\bf Aa} code is then converted to a virtual circuit using
a {\bf vC} representation.  This is done using a utility called {\bf aa2vc}.
\item Finally, the {\bf vC} code is converted to VHDL which can then
be simulated or synthesized.  This conversion is done using a utility
called {\bf vc2vhdl}.
\end{enumerate}
At each stage, various optimizations are carried out.
}

\frame[containsverbatim]{\frametitle{Lets see what happens}

Go back to ex1/ and check it out.

}

\frame[containsverbatim]{\frametitle{Validating the VHDL}

\begin{itemize}
\item In order to test the VHDL, we will use the same program
that was used to test the original C code.
\item Two processes are started: one executing the test-program
and the other executing a VHDL simulator which simulates the
generated VHDL.  The two processes communicate through sockets.
\item There is no need to write a new test-bench!
\end{itemize}
}

\frame[containsverbatim]{\frametitle{Lets take a look}
Go back to examples/ex1 and check it out.
}

\frame[containsverbatim]{\frametitle{Lets make it a  little more interesting}
\begin{verbatim}
void maxDaemon()
{
  while(1)
  {
       uint32_t a = read_uint32("in_data");
       uint32_t b = read_uint32("in_data");
       c = ((a > b) ? a : b);
       write_uint32("out_data", c);
  }
}
\end{verbatim}
}

\frame[containsverbatim]{\frametitle{The testbench for the Daemon}
\begin{verbatim}
int main(int argc, char* argv[])
{
  uint32_t a, b, c;
  while(1)
  {
     scanf("%d", &a);
     scanf("%d", &b);
     write_uint32("in_data", a);
     write_uint32("in_data", b);
     c = read_uint32("out_data");
     printf(stdout,"Result = %d.\n", c);
  }
  return(0);
}
\end{verbatim}
}

\frame[containsverbatim]{\frametitle{The testbench for the Daemon}
Lets check out the software and the hardware generated from it. 
Go to ex2/.
}

\frame[containsverbatim]{\frametitle{Installing AHIR-v2}

\begin{itemize}
\item You will need to install llvm-2.8, clang-2.8, the BOOST graph library.
\item You can download the source code from gitHub.
\item Go to the v2/ folder in the git repository.  You will need to
source ``build\_bashrc'', and then type ``scons'' to build the executables.
You can then make a release by going back to the root of the git repository
and then typing ``make -f ReleaseMakeFile''.  This populates the Release
directory with binaries, examples etc. which should be enough to get you
going.
\end{itemize}
}

\end{document}

### 2014-02-22 Documents/Tutorials/Aa/Aa.tex
Added tutorial. Also fixed a bug in writing the VHDL test-bench.

\documentclass{beamer}
\usepackage{epsfig}
\title{An introduction to the {\bf Aa} language}
\author{Madhav P. Desai}
\begin{document}
\maketitle

\frame[containsverbatim]{\frametitle{The {\bf Aa} language}

\begin{itemize}
\item Serves as an intermediate representation in the AHIR-V2 flow.
\item Control-flow (imperative) language with support for parallelism
and branching.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{A simple program in {\bf Aa}}
\begin{verbatim}
$module [maxOfTwo] $in (a: $uint<32> b:$uint<32>)
   $out (c: $uint<32>) $is
{
   c := ( $mux (a > b) a b )
}
\end{verbatim}
}

\frame[containsverbatim]{\frametitle{Program-structure in {\bf Aa} }
\begin{verbatim}
aA_Program :=  ( aA_Module | 
                 aA_Object_Declaration | 
                 aA_Named_Type_Declaration)*
\end{verbatim}
}

\frame[containsverbatim]{\frametitle{A module}
\begin{verbatim}
aA_Module := $module [identifier]
                $in ( aA_Module_Args) 
                $out (aA_Module_Args) 
{
        aA_Object_Declaration*
        aA_Atomic_Statement_Sequence
}
\end{verbatim}
}

\frame[containsverbatim]{\frametitle{{\bf Aa} Types}

{\bf Aa} provides a comprehensive set of types.
\begin{itemize}
\item {\em Unsigned integers} 
\begin{verbatim}
$uint<1>, $uint<32> etc.
\end{verbatim}
\item {\em Signed integers}   
\begin{verbatim}
$int<1>, $int<32> etc.
\end{verbatim}
\item {\em Sized floats}      
\begin{verbatim} 
$float<8,32>, $float<11,52>
\end{verbatim}
\item {\em Pointers} : 
\begin{verbatim}
$pointer<$uint<32>> etc..
\end{verbatim}
\item {\em Arrays}:  
\begin{verbatim}
$array[32][4] $of $uint<32> etc.
\end{verbatim}
\item {\em Records}: 
\begin{verbatim}
$record $uint<32> $uint<1>
\end{verbatim}
\item {\em Named Records}: 
\begin{verbatim}
$record [MyRec] $pointer<MyRec>) $uint<32>
\end{verbatim}
\end{itemize}
}

\frame[containsverbatim]{\frametitle{{\bf Aa} Objects}
\begin{itemize}
\item Storage objects:
\begin{verbatim}
$storage A: $array [1024] $of $uint<32>
\end{verbatim}
\item Constant objects:
\begin{verbatim}
$constant A: $uint<4> := _b0011
$constant B: $uint<32> := _hf0f0f0f0
$constant C: $float<8,23> := _f2.3465e+0
\end{verbatim}
\item Pipe objects:
\begin{verbatim}
$pipe A: $uint<32> $depth 4
$lifo $pipe B: $uint<8> $depth 8
\end{verbatim}
\item Implicit Variables:  these are defined by
statements:
\begin{verbatim}
a := (A + B)
\end{verbatim}
They are also called {\em static-single-assignment}
or SSA variables.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{{\bf Aa} Expressions}
\begin{itemize}
\item Constants: 
\begin{verbatim}
_b00011
_habf1
\end{verbatim}
\item Simple references:
\begin{verbatim}
a
\end{verbatim}
\item Array references:
\begin{verbatim}
a[0][1] 
a[(I+1)][J][K] 
\end{verbatim}
\item Unary expressions:
\begin{verbatim}
(<op> expr)
e.g.  (~ a)
\end{verbatim}
\item Binary expressions:
\begin{verbatim}
(a <op> B)
<op> can be +,-,*,/,<<,>>,|,&,~|,~&,^,~^
            ==,!=,>,>=,<,<=
\end{verbatim}
\end{itemize}
}

\frame[containsverbatim]{\frametitle{{\bf Aa} More Expressions}
\begin{itemize}
\item Ternary expressions:
\begin{verbatim}
($mux <test-expr> <if-expr> <else-expr> )
e.g.  ($mux (a > 0) (b+1) (b-1))
\end{verbatim}
\item Concatenation expression:
\begin{verbatim}
(a && b)
\end{verbatim}
\item Bit-select expression:
\begin{verbatim}
(a [] I)
\end{verbatim}
\item Address-of expression:
\begin{verbatim}
@(a)
@(a[I])
\end{verbatim}
\item Pointer-dereference expression:
\begin{verbatim}
->(ptr)
\end{verbatim}
If it appears on the left-hand-side, it is
a store, else it is a load.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{{\bf Aa} Statements}
\begin{itemize}
\item {\em Atomic} statements.
\item {\em non-Atomic} statements.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{Atomic {\bf Aa} Statements}
\begin{itemize}
\item {\em Simple} statements.
\item {\em Block} statements.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{{\bf Aa} Atomic Simple Statements}
\begin{itemize}
\item Assignment statements:
\begin{verbatim}
target-expression := source_expression
e.g.
a := (b + c)
\end{verbatim}
\item Call statements:
\begin{verbatim}
$call fpadd32 (A B) (C)
\end{verbatim}
\end{itemize}
}

\frame[containsverbatim]{\frametitle{{\bf Aa} Atomic Block Statements}
\begin{itemize}
\item Series-block statements.
\item Parallel-block statements.
\item Branch-block statements.
\item Fork-block statements.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{{\bf Aa} Series Block Statements}
\begin{verbatim}
$seriesblock [SB] {
   $storage a: $uint<32>
   a := (b + c)
   d := (a + e)
} (d => D)
\end{verbatim}

Control-flow is sequential: statements are executed in order, token
leaves statement after last statement finishes.  
\vspace{0.5cm}
A module body is also a series-block.
}

\frame[containsverbatim]{\frametitle{{\bf Aa} Parallel Block Statements}
\begin{verbatim}
$parallelblock [PB] {
   a := (b + c)
   p := (q + r)
}
\end{verbatim}
Control-flow: both statements started in parallel,  token leaves
statement after both statements have finished.
}

\frame[containsverbatim]{\frametitle{{\bf Aa} Branch Block Statements}
\begin{verbatim}
$branchblock [BB] {
   $merge $entry loopback
       $phi I := ($bitcast($uint<32>) 0) $on $entry
                    NI $on loopback
   $endmerge
   a[I] := (b[I] + c[I])
   NI := (I+1)
   $if (NI < 16) $then 
      $place [loopback] 
   $endif
}
\end{verbatim}
Control-flow:  sequential, but control flow is altered by
merge, place, if and switch statements.
}

\frame[containsverbatim]{\frametitle{{\bf Aa} Fork Block Statements}
\begin{verbatim}
$forkblock [FB] {
   $seriesblock [S1] { ... }
   $seriesblock [S2] { ... }

   $join S1 S2 $fork
      $seriesblock [S3] { ... }
      $seriesblock [S4] { ... }
   $endjoin

   $join S3 S4 $endjoin
}
\end{verbatim}
Control-flow:  all statements will start in parallel, join-forks
will trigger new statements etc.  Token exits block when all statements
finish.
}

\frame[containsverbatim]{\frametitle{{\bf Aa} Do-While Statements}
These are not atomic, and can occur only inside a branch-block.
\begin{verbatim}
$do
   $merge $entry $loopback
        $phi I := ($bitcast ($uint<32>) 0) 
                      $on $entry
                   NI $on $loopback
   $endmerge
   a[I] := (b[I] + c[I])
   NI   := (I+1)
$while (I < 16)
\end{verbatim}
Control-flow:  sequential, controlled by the condition check.  The
places \$entry and \$loopback are implicitly defined.  When control
enters the do-while, the token gets placed in \$entry and when
control loops-back from the condition check, the token gets placed
in \$loopback.
}

\frame[containsverbatim]{\frametitle{Getting to hardware starting from an {\bf Aa} Program}
}

\end{document}

### 2014-02-22 v2/Ahir-asm/devel/docs/AaLRM.tex
Added tutorial. Also fixed a bug in writing the VHDL test-bench.

@@
 in an \Aa program corresponds to a petri net
-of a specific class (the type-2 petri-net as
-introduced in \cite{ref:SameerPhD}).  
+of a specific class.
 The data-flow is constructed using static
@@
 \end{verbatim}
+Pipe variable declarations can also be appended by
+the following keywords:
+\begin{verbatim}
+   ($in | $out)? ($port)? ($signal)? 
+\end{verbatim}
+These additional keywords mean the following things:
+\begin{itemize}
+\item Inclusion of ``\$in'' (respectively ``\$out'') means
+that the pipe is a read-only (respectively write-only) object.
+\item The ``\$port'' keyword means that the pipe is meant to
+be connected to the outside world {\em without} any handshake,
+such as for example in an asynchronous interface to the outside
+world.
+\item The ``\$signal'' keyword means that the interface to
+the external world is a single bit.
+\end{itemize}
 \item {\bf Constant} variables declared at

### 2014-04-05 Documents/Tutorials/Pipelines/Pipelines.tex
Fixed a problem in which flow-through was not percolating into the InterlockBuffer.

\documentclass{beamer}
\usepackage{epsfig}
\title{Describing Pipelines}
\author{Madhav P. Desai}
\begin{document}
\maketitle

\frame[containsverbatim]{\frametitle{Two ways of describing pipelines}

\begin{itemize}
\item with distinct threads.
\item by pipelining a loop.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{An example}
\begin{itemize}
\item We want to compute $d(k) = (a(k)+b(k)) \times c(k)$ for a sequence of 
numbers. 
\item The system should accept a sequence of numbers $\{a(k),b(k),c(k)\}$
and produce the sequence $\{ d(k) \}$.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{As separate threads}
\begin{verbatim}
void Stage_0()
{
  while(1)
  {
    float a = read_float32("a_pipe");
    float b = read_float32("b_pipe");
    float c = read_float32("c_pipe");
    write_float32("c_forward_pipe", c);
    write_float32("s0_result", a+b);
  }
}
void Stage_1()
{
  while(1)
  {
    float c = read_float32("c_forward_pipe");
    float aSb = read_float32("s0_result");
    write_float32("d_pipe", aSb * c);
  }
}
\end{verbatim}
}

\frame[containsverbatim]{\frametitle{In a single loop}
\begin{verbatim}
void Daemon()
{
  while(1)
  {
    float a = read_float32("a_pipe");
    float b = read_float32("b_pipe");
    float c = read_float32("c_pipe");
    write_float32("d_pipe", (a+b)*c);
  }
}
\end{verbatim}
}

\frame[containsverbatim]{\frametitle{Loop Pipelining}
\begin{itemize}
\item Loop pipelining involves executing multiple iterations of a
loop at the same time.
\item While doing this, all dependencies must be taken care of.
\begin{itemize}
\item Operation order.
\item Data dependency.
\item Memory accesses.
\item Pipe accesses.
\end{itemize}
\end{itemize}
}

\frame[containsverbatim]{ \frametitle{Example}
}
\end{document}

### 2014-04-14 Documents/Tutorials/MemorySpaces/MemorySpaces.tex
Fixed a bug in AaWriter.cpp.  The operand for a call was getting incorrectly generated when it was to be the pointer to a global variable. Added new function prepare_operand(Value*, Type*).

\documentclass{beamer}
\usepackage{epsfig}
\title{Memory Spaces in AHIRV2}
\author{Madhav P. Desai}
\begin{document}
\maketitle

\frame[containsverbatim]{\frametitle{Storage Variables and Memory Spaces}

\begin{itemize}
\item Storage variables are grouped into memory spaces.
\item Each memory space is implemented in hardware by a memory sub-system. 
\item Memory spaces are identified by doing a static aliasing analysis.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{Aliasing analysis}
\begin{verbatim}
int A, B, C, D; // storage objects.
int foo()
{
   int ret_val = 0;
   int* p1 = &A;
   ret_val += *p1;
   int *p2 = &B;
   ret_val += *p2;
   int *p3 = &C;
   ret_val += *p3;
   p3 = &D;
   ret_val += *p3;
   return(ret_val);
}
\end{verbatim}
}

\frame[containsverbatim]{\frametitle{Groupings}

\begin{itemize}
\item The memory spaces are $\{A\}$, $\{ B\}$, $\{ C, D \}$.
\item $C$ and $D$ are grouped together because pointer $p3$ can
point to either $C$ or $D$.
\end{itemize}

}

\frame[containsverbatim]{\frametitle{Tool-flow for identifying memory-spaces.}

\begin{verbatim}
AaLinkExtMem -I 1 -E mempool prog1.aa prog2.aa | vcFormat > linked.aa
Aa2VC -E mempool  linked.aa ....
\end{verbatim}

}

\frame[containsverbatim]{\frametitle{AaLinkExtMem}
\begin{verbatim}
AaLinkExtMem -I 1 -E mempool prog1.aa prog2.aa | vcFormat > linked.aa
\end{verbatim}
\begin{itemize}
\item -I 1 : introduce a dummy storage variable of size 1 byte as target for unclassified
pointer dereferences.
\item -E mempool:  this dummy variable is called mempool.
\item Use the -E mempool option with Aa2VC to pass this information to vC and further to VHDL.
\end{itemize}
}

\frame[containsverbatim]{ \frametitle{Example}
}
\end{document}

### 2014-04-14 Documents/Tutorials/MemorySpaces/ex1/src/prog.h

@@ -1,0 +1,2 @@
+void vectorSum();
+uint32_t doSum(uint32_t*, uint32_t);

### 2014-08-10 v2/Ahir-asm/devel/docs/AaLRM.tex
Get github into synch.

@@
 \item Unary expressions.
+\item Slice expressions.
+\item Bitmap (bit permutation) expressions.
 \item Binary expressions.
@@
 
+\subsection{Slice expressions}
+
+Slice expressions can be used to extract a contiguous
+bit-field from a value.  They have the
+following syntax:
+\begin{verbatim}
+LPAREN $slice <expression> <high-index> <low-index> RPAREN
+\end{verbatim}
+where high-index and low-index are constant non-negative
+integers.  The result of executing the expression is
+an unsigned integer whose width is $(high-index - low-index)+1$.
+The value produced by expression may be of any type, but
+for the slicing, it is flattened in the natural manner
+(see the notes about aggregate types).  All bit-vectors
+are assumed to indexed as $HIGH$ dowto $LOW$.
+
+For example
+\begin{verbatim}
+( $slice ($bitcast ($uint<16>) _h1234) 8 4) = _h3
+\end{verbatim}
+
+\subsection{Bitmap expressions}
+
+These expressions can be used to express bit permutations
+in a word.  The syntax is as follows
+\begin{verbatim}
+LPAREN $bitmap <expression> ( <from-index> <to-index> )+ RPAREN
+\end{verbatim}
+The flattened bit-vector value of expression is permuted according
+to the information provided by the $(from,to)$ pairs (result(to) = source(from)).
+
+For example
+\begin{verbatim}
+( $slice ($bitcast ($uint<4>) _h3) 0 3 1 2 2 1 3 0 ) = _he
+\end{verbatim}
+
 \subsection{Binary expressions}
@@
 SHR               >> 
+ROL               <o<
+ROR               >o>
 

### 2014-08-10 v2/Ahir-asm/devel/src/AaUtil.cpp

@@ -103,6 +103,12 @@ string C_Name(AaOperation op)
     case __SHR:
       ret_string = "__SHR";
       break;
+    case __ROL:
+      ret_string = "__ROL";
+      break;
+    case __ROR:
+      ret_string = "__ROR";
+      break;
     case __PLUS:
       ret_string = "__PLUS";
       break;
@@ -176,6 +182,12 @@ string Aa_Name(AaOperation op)
       break;
     case __SHR:
       ret_string = ">>";
+      break;
+    case __ROL:
+      ret_string = "<o<";
+      break;
+    case __ROR:
+      ret_string = ">o>";
       break;
     case __PLUS:
       ret_string = "+";

### 2014-08-10 v2/CtestBench/examples/clickwrapper/ToVHDL.sh

@@ -1,23 +1,0 @@
-# functions in prog.c to be mapped to vhdl
-# generate llvm-byte-code
-clang -std=gnu89 -I../../../iolib/ -emit-llvm -c prog.c
-# disassemble byte-code for human readibility.
-opt --instnamer prog.o > prog.opt.o
-# llvm-bytecode to Aa
-llvm-dis prog.opt.o
-llvm2aa prog.opt.o | vcFormat >  prog.o.aa
-# map external memory references from system to
-# internal object mempool (of size 1 byte).
-AaLinkExtMem -I 1 -E mempool prog.o.aa | vcFormat > prog.o.linked.aa
-# take linked Aa and convert to virtual circuit.
-Aa2VC -O -I mempool -C prog.o.linked.aa | vcFormat > prog.o.linked.aa.vc
-#
-# take virtual circuit and convert to VHDL
-# NOTE: the -T option tells vc2vhdl that io_module is to be an ever-running
-# top-level module, which will not need to be started from the outside, but
-# will be free-running in the system (such modules cannot have input/output
-# arguments, but can only listen/send on pipes).
-#
- vc2vhdl -C -s modelsim -e ahir_system -w -T wrapper_input -T wrapper_output -T free_queue_manager -f prog.o.linked.aa.vc 
-vhdlFormat < ahir_system.unformatted_vhdl > ahir_system.vhdl
-vhdlFormat < ahir_system_test_bench.unformatted_vhdl > ahir_system_test_bench.vhdl

### 2014-08-10 v2/libAhirV2/include/vcDataPath.hpp

@@ -12,6 +12,7 @@ class vcLoad;
 class vcStore;
 class vcSelect;
 class vcSlice;
+class vcPermutation;
 class vcBranch;
 class vcInport;
 class vcOutport;
@@ -351,6 +352,12 @@ protected:
 	}
   }
 
+  virtual void Print_Flow_Through(ostream& ofile) 
+  {
+	if(this->Get_Flow_Through())
+		ofile << " $flowthrough ";
+  }
+
   void Set_Input_Buffering(vcWire* w, int buffering)
   {
 	_input_wire_buffering_map[w] = buffering;
@@ -410,6 +417,7 @@ class vcDataPath: public vcRoot
   map<string, vcWire*> _wire_map;
   map<string, vcSelect*> _select_map;
   map<string, vcSlice*> _slice_map;
+  map<string, vcPermutation*> _permutation_map;
   map<string, vcBranch*> _branch_map;
   map<string, vcRegister*> _register_map;
   map<string, vcEquivalence*> _equivalence_map;
@@ -467,6 +475,9 @@ class vcDataPath: public vcRoot
   void Add_Slice(vcSlice* p);
   vcSlice* Find_Slice(string id);
 
+  void Add_Permutation(vcPermutation* p);
+  vcPermutation* Find_Permutation(string id);
+
   void Add_Branch(vcBranch* p);
   vcBranch* Find_Branch(string id);
 
@@ -522,6 +533,7 @@ class vcDataPath: public vcRoot
   void Print_VHDL_Phi_Instances(ostream& ofile);
   void Print_VHDL_Select_Instances(ostream& ofile);
   void Print_VHDL_Slice_Instances(ostream& ofile);
+  void Print_VHDL_Permutation_Instances(ostream& ofile);
   void Print_VHDL_Register_Instances(ostream& ofile);
   void Print_VHDL_Interlock_Buffer_Instances(ostream& ofile);
   void Print_VHDL_Binary_Logical_Operator_Instances(ostream& ofile);

### 2014-08-10 v2/libAhirV2/include/vcRoot.hpp

@@ -97,6 +97,8 @@ enum vcLexerKeytags
     __DIV_OP           , 
     __SHL_OP           , 
     __SHR_OP           , 
+    __ROL_OP           , 
+    __ROR_OP           , 
     __SGT_OP            , 
     __SGE_OP            , 
     __EQ_OP            , 
@@ -122,6 +124,7 @@ enum vcLexerKeytags
     __NAND_OP          ,
     __XNOR_OP          ,
     __EQUIVALENCE_OP,
+    __PERMUTE_OP,
     __OPEN,
     __DELAY,
     __SHRA_OP,
@@ -231,6 +234,8 @@ static string vcLexerKeywords[] =
       "/",
       "<<",
       ">>",
+      "<o<",
+      ">o>",
       "$S>$S",
       "$S>=$S",
       "==",
@@ -256,6 +261,7 @@ static string vcLexerKeywords[] =
       "~&",
       "~^",
       "&/",
+      ":X=",
       "$open",
       "$delay",
       // signed shr

### 2014-08-10 v2/llvmbc2Aa/include/AaWriter.hpp

@@ -36,12 +36,25 @@ namespace Aa {
     bool _guard_complement;
     std::string _guard_variable;
 
+
+    // flags for controlling how a
+    // basic block is to be printed.
     bool _suppress_leading_merge;
     bool _suppress_terminator;
-    llvm::BasicBlock* _chain_trailer;
+    BasicBlockChainDAG* _active_dag;
 
+
+    // kept as strings.. a hack.
     std::map<std::string,std::set<std::string> > bb_predecessor_map;
     std::map<std::string, llvm::BasicBlock*> bb_name_map;
+
+    // kept as pointers..  
+    std::map<llvm::BasicBlock*,std::vector<llvm::BasicBlock*> > basic_block_predecessor_map;
+
+
+    // get predecessors.
+    void get_predecessors(llvm::BasicBlock* bb, std::vector<llvm::BasicBlock*>& preds);
+    void get_predecessors(std::string& bb, std::vector<std::string>& preds);
 
     std::map<llvm::Value*, std::string> value_name_map;
     std::map<std::string,std::string> pipe_map;
@@ -52,6 +65,7 @@ namespace Aa {
     std::map<llvm::BasicBlock*, llvm::BasicBlock*> chain_representative_map;
     std::vector<llvm::BasicBlock*> ordered_chain_rep_vector;
 
+    std::vector<BasicBlockChainDAG*> _bb_chain_dag_vector;
 
     int _pointer_width;
     void Set_Pointer_Width(int w) { _pointer_width = w; }
@@ -84,17 +98,17 @@ namespace Aa {
     void Set_Guard_Variable(std::string gs) {_guard_variable  = gs;}
     std::string Get_Guard_Variable() {return(_guard_variable);}
     std::string Get_Guard_String() {
-	if(this->_guard_complement)
-		return( "( ~ " + _guard_variable + " )");
-	else
-		return("(" + _guard_variable + ")");
+	    if(this->_guard_complement)
+		    return( "( ~ " + _guard_variable + " )");
+	    else
+		    return("(" + _guard_variable + ")");
     }
     std::string Get_Guard_RHS_Expression() {
-	if(this->_guard_complement)
-		return( "( ~ " + _guard_variable + " )");
-	else
-		return(_guard_variable);
-   }
+	    if(this->_guard_complement)
+		    return( "( ~ " + _guard_variable + " )");
+	    else
+		    return(_guard_variable);
+    }
 
[…]

### 2014-08-10 v2/llvmbc2Aa/src/ModuleGen.cpp

@@ -13,6 +13,7 @@
 #include <set>
 #include <iostream>
 #include <fstream>
+#include "BasicBlockChain.hpp"
 #include "AaWriter.hpp"
 #include "Utils.hpp"
 
@@ -426,14 +427,14 @@ namespace {
 	
 	std::cout << "$branchblock [body] {"  << std::endl;
 
-	// build basic-block chains.
-	aa_writer->Build_Basic_Block_Chains(F);
+	// build basic-block chain dags.
+	aa_writer->Build_Basic_Block_Chain_DAGs(F);
 
 
 	// generate AA code for all the basic-block chains.
-	for(int J = 0, fJ = aa_writer->ordered_chain_rep_vector.size(); J < fJ; J++)
+	for(int J = 0, fJ = aa_writer->_bb_chain_dag_vector.size(); J < fJ; J++)
 	{
-		aa_writer->Write_Aa_Code(aa_writer->ordered_chain_rep_vector[J],_extract_do_while);
+		aa_writer->Write_Aa_Code(aa_writer->_bb_chain_dag_vector[J],_extract_do_while);
 	}
 
 	/*

### 2014-08-10 v2/llvmbc2Aa/src/Utils.cpp

@@ -12,7 +12,9 @@
 #include "Utils.hpp"
 #include <stdio.h>
 #include <limits.h>
+#include <math.h>
 
+using namespace std;
 using namespace llvm;
 using namespace Aa;
 
@@ -121,23 +123,36 @@ std::string Aa::get_aa_constant_string(llvm::Constant *konst)
   else if(isa<ConstantFP>(konst))
     {
       const ConstantFP *fkonst = dyn_cast<ConstantFP>(konst);
+      bool is_normal_fp = true;
 
       // fix this.  this should be a binary string..
       char buffer[1024];
       if(type->isFloatTy())
 	{
-	  sprintf(buffer,"%e",fkonst->getValueAPF().convertToFloat());
+	  float fval = fkonst->getValueAPF().convertToFloat();
+          
+          is_normal_fp = (fpclassify(fval) == FP_NORMAL);
+          if(is_normal_fp) // if it is not a nan/inf, then keep it as a float.
+	  	sprintf(buffer,"_f%e",fval);
+	  else // else, print it as a hex number and prepend _h.
+		sprintf(buffer,"_h%x", ((uint32_t) *((uint32_t*)&fval)));
+		
 	}
       else if(type->isDoubleTy())
 	{
-	  sprintf(buffer,"%e",fkonst->getValueAPF().convertToDouble());
+	  double dval = fkonst->getValueAPF().convertToDouble();
+          is_normal_fp = (fpclassify(dval) == FP_NORMAL);
+	  if(is_normal_fp)// if it is not a nan/inf, then keep it as a float.
+	  	sprintf(buffer,"_f%e",dval);
+          else // else, print it as a hex number and prepend _h.
+		sprintf(buffer,"_h%llx", ((uint64_t) *((uint64_t*)&dval)));
 	}
       else
 	{
 	  std::cerr << "Error: unsupported floating point type (only float and double are allowed)"
 		    << std::endl;
 	}
-      ret_val = "_f" + std::string(buffer);
+      ret_val = std::string(buffer);
     }
   else if(isa<ConstantArray>(konst) || isa<ConstantStruct>(konst) || isa<ConstantVector>(konst))
     {
@@ -1055,9 +1070,7 @@ bool Aa::is_zero(llvm::Constant* konst)
 }
 
 
-// if the terminator contains a branch back to this
[…]

### 2014-09-02 Documents/AhirV2/AhirV2.tex
Get the documentation into better shape.

@@
 
-AhirV2 is a set of tools 
-\begin{itemize}
-\item which can convert a C/C++ progam to  an equivalent hardware
-implementation (described in VHDL).
-\item which provide an algorithmic approach to the design of 
-hardware.
-\end{itemize}
+AhirV2 is a set of tools which can convert a C 
+description of a system to  an equivalent hardware
+implementation (described in VHDL).  Using these tools,
+it is possible to take an algorithmic approach to 
+the design of  hardware.
 
@@
 \begin{itemize}
-\item Given a high-level C/C++ program, we rely on
+\item Given a high-level C, we rely on
 an LLVM (www.llvm.org) compatible compiler such as clang (www.clang.org)
-or dragon-egg (dragonegg.llvm.org) to produce LLVM byte code.
+to produce LLVM byte code.
 Currently, the AhirV2 flow uses LLVM byte code as a starting point.
@@
 \Aa  which serves as a target for sequential
-programming languages (such as C/C++) as
+programming languages (such as C) as
 well as for parallel programming languages.
 An \Aa program consists of modules (analogous
-to sub-programs in C/C++) which can call each 
+to sub-programs in C) which can call each 
 other, and can communicate through storage objects
@@
 description language \vC) is generated.
-The chief optimization carried out at this
-step is the decomposition of the system memory
-into disjoint spaces based on usage
+The chief optimizations carried out at this
+step are dependency based operation
+ordering, dynamic loop-pipelining and decomposition of the system memory
+into disjoint spaces based on static pointer
 analysis (this considerably improves the
@@
 Thus, to generate hardware using the AhirV2 flow,
-it is possible to start at C/C++, at the \Aa level
+it is possible to start at the C-level, at the \Aa level
 at the \vC level or at the VHDL level (or a combination
@@
 easier for the programmer, but using lower level representations
-will probably lead to more efficient hardware.
-Typically, if we start from C/C++, circuits
+will usually lead to more efficient hardware.
+Typically, if we start from C, circuits
 produced by the AhirV2 flow are upto two orders of
@@
 Currently, there are only two restrictions in mapping
-a C/C++ program to VHDL using the AhirV2 flow:
+a C program to VHDL using the AhirV2 flow:
 \begin{itemize}
@@
 \end{verbatim}
-In this example, the read/write functions are assumed
-to use named pipes for inter-process communication.
+In this example, the read/write functions are provided
+as part of a {\em pipeHandler} library which is bundled with
+the AhirV2 distribution.
 
@@
 \begin{verbatim}
-llvm2aa [-modules=<listfile>] [-storageinit] bytecode.o > bytecode.aa
+llvm2aa options bytecode.o > bytecode.aa
 \end{verbatim}
@@
 are explicitly initialized in the generated \Aa code.   An initializer
-routine {\bf global\_storage\_initializer\_} is instantiated in
+routine named
+\begin{verbatim}
+    global_storage_initializer
+\end{verbatim}
+is instantiated in
 the \Aa code for this purpose.
+\item {\bf -pipedepths=filename} : Specifies a file which contains
+the depths of pipes which are part of the generated \Aa code.
+\item {\bf -extract\_do\_while} : Innermost loops which are marked
+using a call to the special function 
+\begin{verbatim}
+    _loop_pipelining_on_
+\end{verbatim}
+are extracted as pipelined do-while loops.  This is necessary
+for automatic parallelization of inner loops in the 
+generated hardware (substantial performance benefits can be
+realized).
 \end{itemize}
@@
 \begin{verbatim}
-AaLinkExtMem [-I n] [-E obj-name] bytecode.o > bytecode.aa
+AaLinkExtMem options file1.aa file2.aa ...  > linked.aa
 \end{verbatim}
@@
 
+\subsection{{\bf AaOpt}}
+
+The optimization utility {\bf AaOpt} takes an \Aa program (list of
+\Aa files) and produces an optimized version of the source program.
+\begin{verbatim}
+AaOpt options file1.aa file2.aa ... > optimized.aa
+\end{verbatim}
+The optimized \Aa code is printed to {\bf stdout}.  On success, 
+the tool returns a 0 (else a non-zero).  Macro and inlined
+function calls in the source code are substituted in place
+in the optimized code.
+
+The options:
+\begin{itemize}
+\item {\bf -r module-name} (optional) : specifies a root module in the
+system.  Multiple root modules can be specified.  All dead code (which
+is not reachable from a root module) is eliminated.
+\item {\bf -I extmem-object} (optional) : similar to AaLinkExtMem,
+this option specifies the name of the extmem-object in the
+source \Aa files.
+\item {-B} (optional) : if specified, add buffering to balance pipelined
+loops so that loop performance is not bottlenecked by inadequate
+buffering.
+\end{itemize}
+
 \subsection{{\bf Aa2VC}}
@@
 \begin{verbatim}
-Aa2VC [-O] [-C] [-I obj-name] file1.aa file2.aa ...  > result.vc
+Aa2VC options file1.aa file2.aa ...  > result.vc
 \end{verbatim}
@@
 the VHDL code generated by downstream tools.
+\item {\bf -U} : memory subsystems will be unordered (that is,
+will not guarantee in-order completion of accesses).  This
+leads to a simpler memory subsystem, but more conservative
+control flow.  The default is that all memory subsystems
+are ordered (will complete read/write requests in the order that
+they are accepted).
+\item {\bf -r root-module} (optional): specifies a root module.
+Code which is not accessible from a root-module is considered
+as dead code and is ignored.
 \item {\bf -I obj-name} : if specified, all external memory references
@@
 \begin{verbatim}
-vc2vhdl [-O] [-C] [-a] [-e <entity-name] [-w]\
+vc2vhdl [-O] [-C] [-q] [-a] [-e <entity-name] [-w]\
          -t/-T foo [-t/-T bar -t/-T bar2 ...]\
@@
 and are run forever (restarted after they finish, forever).
-\item {\bf -f} : specifies the \vC files to be analyzed. 
+\item {\bf -f file-name} : specifies the \vC files to be analyzed. 
 Multiple \vC files may be specified.  An object must be
@@
 operator  only if it can be proved that they cannot be active simultaneously.
+\item {\bf -q } : if specified, do aggressive register insertion
+to minimize the clock period.
+\item {\bf -S bypass-stride } : by specifying the bypass stride
+(an integer $\geq 1$),
+the user can trade-off clock cycles versus clock period.  
+The lowest clock period will be obtained for -S 1.   
 \item {\bf -e top-entity-name} : The generated top-level VHDL entity
@@
 is test\_system.
+\item {-L function-library} : AhirV2 provides some built in operator
+functions which can be called from your code.  These are organized
+as function libraries and this option specifies a function library
+to look into when generating VHDL.  For example {\em -L fpu} gives
+access to the floating point library which provides some useful
+built in functions (e.g. fpalu32, fpalu64 etc.).
 \item {\bf -w} :  If specified, the VHDL system and test-bench are

### 2014-09-02 v2/CtestBench/docs/CtestBench.tex
Get the documentation into better shape.

--- 2011-08-16 v2/CtestBench/docs/CtestBench.tex
@@
 #include <stdint.h>
+#include <Pipes.h>
 #include <stdio.h>
-#include <iolib.h>
 
-uint32_t sum;
+// note: initialized value..
+uint32_t sum1 = 23;
+uint32_t sum2 = 39;
 
-void set_sum(uint32_t x)
+// note: no problems with pointers :-)
+uint32_t* tgt[2] = {&sum1, &sum2};
+
+uint32_t get_sum(uint32_t idx)
 {
-    sum = x;
-}
-
-uint32_t get_sum()
-{
-    return(sum);
+  return(*(tgt[idx]));
 }
@@
 {
+  int i = 0;
   while(1)
@@
 #endif
-        sum = (sum + nxt);
-        write_uint32("out_data",sum);
+     // ugly, but this is just a demo,
+     // we are showing off.
+     *(tgt[i])= ((*tgt[i]) + nxt);
+
+     write_uint32("out_data",*(tgt[i]));
 #ifdef SW
-        printf("wrote %u\n", sum);
+     printf("wrote %u\n", *(tgt[i]));
 #endif
+     i = 1 - i;
   }
@@
 #include <stdlib.h>
+#include <stdint.h>
 
+#include <Pipes.h>
 #ifdef SW
-// for the read_*/write_* methods
-#include <iolib.h>
+#include <pipeHandler.h>
 #include "prog.h"
 #else
-// includes for the read_*/write_* methods
-// as well as stubs for functions moved
-// to hardware.
 #include "vhdlCStubs.h"
@@
 
-void *accumulate_(void* fargs)
-{
-   accumulate();
-}
-
-void *write_pipe_(void* a)
-{
-   write_uint32_n("in_data",(uint32_t*)a, 10);
-}
-
-void *read_pipe_(void* a)
-{
-   read_uint32_n("out_data",(uint32_t*)a, 10);
-}
+// The model of the hardware accumulate thread
+// is necessary when building the SW testbench.
+#ifdef SW
+DEFINE_THREAD(accumulate)
+#endif
 
@@
 
-        // initial value of sum.
-        set_sum(1);
+#ifdef SW
+  init_pipe_handler();
+
+  // register FIFO
+  register_pipe("in_data",10,32,0);
+  // register FIFO..
+  register_pipe("out_data",10,32,0);
+
+#endif
+
+#ifndef SW
+  // to set the initial value of sum.
+  // in the hardware version, storage
+  // variables are initialized by calling
+  // this function (auto-generated by
+  // the Aa linker AaLinkExtMem)
+  global_storage_initializer_();
+#endif
+
+#ifdef SW
+  // start the accumulate thread.
+  // (this is necessary only in the SW model)
+  PTHREAD_DECL(accumulate)
+  PTHREAD_CREATE(accumulate)
+#endif
 
@@
 
-        pthread_t acc_t, wpipe_t, rpipe_t;
+  // write 10 things to in_data, it has enough room..
+  write_uint32_n("in_data",(uint32_t*)data_in, 10);
 
-#ifdef SW
-        pthread_create(&acc_t,NULL,&accumulate_,NULL);
-#endif
-        pthread_create(&wpipe_t,NULL,&write_pipe_,(void*)data_in);
-        pthread_create(&rpipe_t,NULL,&read_pipe_,(void*)data_out);
-
-        pthread_join(wpipe_t,NULL);
-        pthread_join(rpipe_t,NULL);
+  // read back 10 things from out_data..
+  read_uint32_n("out_data",(uint32_t*)data_out, 10);
 
@@
   fprintf(stdout,"\n");
-        fprintf(stdout,"final sum is %u\n", get_sum());
+
+  fprintf(stdout,"Sum 0 is %d\n",get_sum(0));
+  fprintf(stdout,"Sum 1 is %d\n",get_sum(1));
 
 #ifdef SW
-        pthread_cancel(acc_t);
+  close_pipe_handler();
 #endif
 }
-
 \end{verbatim}
-The test-bench sets an initial value for sum, and starts
-three threads: one to write data to pipe in\_data, one
-to read data from pipe out\_data, and one to run the
-accumulate function.  After the last data is read back
-from out\_data, the test-bench gets the value of sum
-and prints it out.
+In the SW case, the test-bench starts the accumulate thread,
+writes data to the hardware, and reads back stuff from the
+hardware.  In the non-SW case, there is no need to 
+start the accumulate thread, since it exists in the 
+hardware model.
 
@@
 functions correctly.  
-The difference is that instead of using methods in iolib,
+The difference is that instead of using the pipeHandler,
 the test-bench now uses methods in SocketLib.  Further, the
@@
 
-\section{An example}
+The following Makefile builds a software-only testbench
+executable, and also converts the system described in prog.c
+to VHDL.  The same testbench can be used to test the VHDL
+also.
+\begin{verbatim}
+# build software version of testbench (to check the "desired behaviour")
+SOCKETLIB_INCLUDE=$(AHIR_RELEASE)/CtestBench/include
+SOCKETLIB_LIB=$(AHIR_RELEASE)/CtestBench/lib
+PIPEHANDLER_INCLUDE=$(AHIR_RELEASE)/pipeHandler/include
+PIPEHANDLER_LIB=$(AHIR_RELEASE)/pipeHandler/lib
+PTHREADUTILS_INCLUDE=$(AHIR_RELEASE)/pthreadUtils/include
+VHDL_LIB=$(AHIR_RELEASE)/vhdl
+VHDL_VHPI_LIB=$(AHIR_RELEASE)/CtestBench/vhdl
+FUNCTIONLIB=$(AHIR_RELEASE)/functionLibrary/
+SRC=./src
+all: SW HW 
+TOAA:c2llvmbc llvmbc2aa  aalink
+TOVC:c2llvmbc llvmbc2aa  aalink aa2vc 
+VC2VHDL: vc2vhdl  vhdlsim
+AA2VHDLSIM: aa2vc  vc2vhdl  vhdlsim
+TOVHDL:TOVC vc2vhdl
 
-Look at the subdirectory ``example'' which contains the files ``prog.c'' and
-``testbench.c''.  A Makefile and README are also present.  Its all quite
-self-explanatory.
+# llvm2aa opts: pipelined case, extract-do-while.
+#LLVM2AAOPTS=--storageinit=true
+LLVM2AAOPTS=-extract_do_while=true --storageinit=true -pipedepths=pipedepths.txt
 
+PROGDEFS=
+
+# the top-level modules
+# -T specifies top-level daemon module
+# -t specifies top-level slave module.
+#
+TOPMODULES=-T accumulate -t get_sum -t global_storage_initializer_
+
+# compile with SW defined.
+# note the use of IOLIB in building the testbench.
+SW: $(SRC)/prog.c $(SRC)/prog.h $(SRC)/testbench.c 
+	gcc -g -c -DSW $(PROGDEFS) -I$(PIPEHANDLER_INCLUDE)\
+              -I$(FUNCTIONLIB)/include -I$(SRC) $(SRC)/prog.c
+	gcc -g -c -DSW $(PROGDEFS) -I$(PIPEHANDLER_INCLUDE)\
+              -I$(PTHREADUTILS_INCLUDE) -I$(SRC) $(SRC)/testbench.c
+	gcc -g -o testbench_sw prog.o testbench.o\
+              -L$(PIPEHANDLER_LIB) -lPipeHandler -lpthread
+
+# five steps from C to vhdl simulator.
+HW: c2llvmbc llvmbc2aa  aalink aa2vc  vc2vhdl  vhdlsim
+
+AA2VHDL: aa2vc vc2vhdl vhdlsim
+
+# C to llvm byte-code.. use clang.
+c2llvmbc: $(SRC)/prog.c $(SRC)/prog.h
+	clang -O3 -std=gnu89 $(PROGDEFS)  -I$(SOCKETLIB_INCLUDE)\
+              -I$(FUNCTIONLIB)/include -emit-llvm -c $(SRC)/prog.c
+	opt --indvars --loopsimplify prog.o -o prog.opt.o
+	llvm-dis prog.opt.o
+
+# llvm byte-code to Aa..
+llvmbc2aa:  prog.opt.o 
+	llvm2aa $(LLVM2AAOPTS)  prog.opt.o | vcFormat >  prog.aa
+
+# Aa to vC
+aalink: prog.aa 
+	AaLinkExtMem prog.aa | vcFormat > prog.linked.aa
+	AaOpt -B prog.linked.aa | vcFormat > prog.linked.opt.aa
+
+aa2vc: prog.linked.opt.aa
+	Aa2VC -O -C prog.linked.opt.aa | vcFormat > prog.vc
+
+# vC to VHDL
+vc2vhdl: prog.vc
+	vc2vhdl -O -S 4 -I 2 -v -a -C -e ahir_system\
+                 -w -s ghdl $(TOPMODULES) -f prog.vc 
+	vhdlFormat < ahir_system_global_package.unformatted_vhdl\
+                 > ahir_system_global_package.vhdl
+	vhdlFormat < ahir_system.unformatted_vhdl\
+                 > ahir_system.vhdl
+	vhdlFormat < ahir_system_test_bench.unformatted_vhdl\
+                 > ahir_system_test_bench.vhdl
+
+# build testbench and ghdl executable
+# note the use of SOCKETLIB in building the testbench.
+vhdlsim: vhdlTb ghdlModel
+
+vhdlTb: $(SRC)/testbench.c vhdlCStubs.h vhdlCStubs.c
+	gcc -c vhdlCStubs.c  -I$(SRC)\
+                -I./ -I$(SOCKETLIB_INCLUDE)
+	gcc -c $(SRC)/testbench.c\
+                -I$(PTHREADUTILS_INCLUDE) -I$(SRC)\
+                -I./ -I$(SOCKETLIB_INCLUDE)
+	gcc -o testbench_hw testbench.o vhdlCStubs.o\
+                -L$(SOCKETLIB_LIB) -lSocketLib -lpthread
+
+ghdlModel: ahir_system.vhdl ahir_system_test_bench.vhdl ahir_system_global_package.vhdl
+	ghdl --clean
+	ghdl --remove
+	ghdl -i --work=GhdlLink  $(VHDL_LIB)/GhdlLink.vhdl
+	ghdl -i --work=ahir  $(VHDL_LIB)/ahir.vhdl
+	ghdl -i --work=aHiR_ieee_proposed\
+                $(VHDL_LIB)/aHiR_ieee_proposed.vhdl
+	ghdl -i --work=work ahir_system_global_package.vhdl
+	ghdl -i --work=work ahir_system.vhdl
+	ghdl -i --work=work ahir_system_test_bench.vhdl
+	ghdl -m --work=work -Wl,-L$(SOCKETLIB_LIB) -Wl,-lVhpi ahir_system_test_bench
+
+PHONY: all clean	
+\end{verbatim}
+To test the software, run testbench\_sw.  To verify the hardware (using
+the VHDL simulator GHDL), start testbench\_sw in one shell, and then start ahir\_system\_test\_bench
+in a different shell.
 \end{document}

### 2014-11-05 Documents/Publications/loopPipelining/paper.tex
Revived loop-pipelining paper for arxiv.

--- 2013-03-27 Documents/Publications/dsd2013/paper.tex
@@
 
-\author{Author's name Suppressed for Blind Review}
+\author{Madhav P. Desai\\ Department of Electrical Engineering\\ IIT Bombay, Mumbai, India\\ madhav@ee.iitb.ac.in}
 \newtheorem{definition}{Definition}[section]
@@
   of an inner loop.  
-  We describe a control-flow mechanism through
-  which multiple iterations of an arbitrary inner loop can be made simultaneously
-  active in the generated hardware, thus providing dynamic loop-pipelining
-  capability in hardware.   
+  We describe a control-flow mechanism which provides dynamic loop-pipelining
+  capability in hardware, so that multiple iterations of an arbitrary inner loop 
+  can be made simultaneously active in the generated hardware, 
   We study the impact of this loop-pipelining scheme in conjunction with
   source-level loop-unrolling.  
-  In particular, we consider some common loop kernels occurring in fast-fourier transform,
-  matrix multiplication and inner product computations, with the resulting
-  hardware descriptions synthesized to an FPGA target.  We observe
+  In particular, we apply this technique to some common loop kernels: regular kernels
+  such as the  fast-fourier transform and matrix multiplication, as well
+  as an example of an inner loop whose body has branching.
+  The resulting hardware descriptions are synthesized to an FPGA target,
+  and then characterized for performance and resource utilization.  
+  We observe
   that the use of dynamic loop-pipelining mechanism alone typically results in
-  a significant improvements in the performance of the hardware.  If the loop is
+  a significant improvements in the performance of the hardware.  
+  If the loop is
   statically unrolled and if loop-pipelining is applied to the unrolled
-  program, then the performance improvement is still substantial.  When
-  both optimizations are used in conjunction, the improvement in performance
+  program, then the performance improvement is still substantial.  
+  When dynamic loop pipelining is used in conjunction with static loop unrolling, the improvement in performance
   ranges from 6X to 20X (in terms of number of clock cycles needed for the computation)
@@
 
-%% The importance and problem of loops.
-A high-level-synthesis (HLS) compiler takes an algorithm described in
-a high-level programming language (such as C, for example) and produces a circuit
-description (a register-transfer-level description in a hardware
-description language such as VHDL, for example).
-High-level synthesis systems have been available for some time
-now (for example, \cite{pegasus-cash}, \cite{spark-vlsi-paper} and
-commercial tools such as Catapult-C\footnote{\url{http://www.mentor.com/products/esl/high_level_synthesis/}} from Mentor Graphics
-C-to-Silicon\footnote{\url{http://www.cadence.com/products/sd/silicon_compiler/}} from Cadence,
-Synopsys HLS\footnote{\url{http://www.synopsys.com/Tools/SLD/HLS/}} from Synopsys).
-
-The main attraction of such a compiler is a substantial reduction in 
-the design and verification effort.  The
-main concern is the quality of the hardware that is produced.
-For the same starting point, the quality of the hardware produced may be objectively
-measured in terms of factors such as area, energy-consumption and performance (in the
-sense of time required for completion, or in terms of the rate at which the hardware
-finishes the computational tasks).  An {\em internal measure} of the quality of the
-hardware produced is the utilization factor of the major resources in the hardware (that is,
-the utilization factor of expensive resources).
-
-In our work, we have developed a set of compiler tools which take a C-description of
-an algorithm (with the only important restrictions being the requirement that
-cycles are not present in the call graph of the source program, and the lack of 
-support for {\bf function} pointers, normal pointers are supported), and produces
-a VHDL description which is provably equivalent to the original program \cite{SameerPhD}.
-Results on the application of this compiler to single-threaded programs show
-that the circuits (when synthesized to VLSI layout) are substantially more
-energy efficient than a processor, but that the performance of the processor
-is usually better than that of the synthesized circuit \cite{ahirDsd2010}.
-If multi-threaded source programs are used as a starting point, the circuits produced 
-are quite competitive with hand-designed RTL \cite{ahirUsenix2012}.
-Quantitatively, the results in \cite{ahirDsd2010} and \cite{ahirUsenix2012} indicate
-that the gap between the synthesized hardware and modern
-pipelined processors (for single-threaded
-programs) and between the synthesized hardware and hand-crafted hardware (for
-multi-threaded programs) is of the order of 2-3X.  That is, the performance
-of the synthesized hardware needs to improve by 2-3X in order to close
-the respective performance gaps.
-
 %% How compilers have handled this problem.
-In this paper, we consider the problem of improving the performance
+We consider the problem of improving the performance
 of hardware generated from single threaded programs; in
-particular, the problem of mapping loops to hardware.  It
+particular, the important problem of mapping loops to hardware.  It
 is well known that most compute intensive programs spend a large fraction
@@
 high performance processors or with hand-crafted hardware.
+
+%% The importance and problem of loops.
+% A high-level-synthesis (HLS) compiler takes an algorithm described in
+% a high-level programming language (such as C, for example) and produces a circuit
+% description (a register-transfer-level description in a hardware
+% description language such as VHDL, for example).
+% High-level synthesis systems have been available for some time
+% now (for example, \cite{pegasus-cash}, \cite{spark-vlsi-paper} and
+% commercial tools such as Catapult-C\footnote{\url{http://www.mentor.com/products/esl/high_level_synthesis/}} from Mentor Graphics
+% C-to-Silicon\footnote{\url{http://www.cadence.com/products/sd/silicon_compiler/}} from Cadence,
+% Synopsys HLS\footnote{\url{http://www.synopsys.com/Tools/SLD/HLS/}} from Synopsys).
+
+% The main attraction of such a compiler is a substantial reduction in 
+% the design and verification effort.  The
+% main concern is the quality of the hardware that is produced.
+% For the same starting point, the quality of the hardware produced may be objectively
+% measured in terms of factors such as area, energy-consumption and performance (in the
+% sense of time required for completion, or in terms of the rate at which the hardware
+% finishes the computational tasks).  An {\em internal measure} of the quality of the
+% hardware produced is the utilization factor of the major resources in the hardware (that is,
+% the utilization factor of expensive resources).
+
+% In our work, we have developed a set of compiler tools which take a C-description of
+% an algorithm (with the only important restrictions being the requirement that
+% cycles are not present in the call graph of the source program, and the lack of 
+% support for {\bf function} pointers, normal pointers are supported), and produces
+% a VHDL description which is provably equivalent to the original program \cite{SameerPhD}.
+% Results on the application of this compiler to single-threaded programs show
+% that the circuits (when synthesized to VLSI layout) are substantially more
+% energy efficient than a processor, but that the performance of the processor
+% is usually better than that of the synthesized circuit \cite{ahirDsd2010}.
+% If multi-threaded source programs are used as a starting point, the circuits produced 
+% are quite competitive with hand-designed RTL \cite{ahirUsenix2012}.
+% Quantitatively, the results in \cite{ahirDsd2010} and \cite{ahirUsenix2012} indicate
+% that the gap between the synthesized hardware and modern
+% pipelined processors (for single-threaded
+% programs) and between the synthesized hardware and hand-crafted hardware (for
+% multi-threaded programs) is of the order of 2-3X.  That is, the performance
+% of the synthesized hardware needs to improve by 2-3X in order to close
+% the respective performance gaps.
 
@@
 \item The performance improvement with loop-pipelining applied alone is in the
-2X-3X range. 
+2X-8X range.   This improvement is observed both in the case of an inner
+loop whose body is a single basic block, as well as in the case when the
+inner loop body has branches.
 \item The performance improvement with loop-unrolling alone is in the 
@@
 
-For the subsequent discussion, we assume that the inner
-loop which is being optimized consists of a single
-basic block, that is, there are no branching instructions
-in the loop body (no jumps, if constructs, switch constructs etc.).
-If such constructs are present, we assume that they have been
-eliminated using the mechanism of guarded (or predicated) execution.
-
 %%
@@
 
-We give a brief description of our compiler flow.  The
+The dynamic pipelining mechanism described in this paper
+is implemented in a compiler flow which takes 
+a C program and produces an equivalent VHDL description.
+We give a brief description of this compiler flow.  The
 details are not relevant to this paper, and the interested
@@
 The compiler flow has been characterized over a wide variety
-of applications \cite{ahirDsd2010}.   The XXXX\footnote{name suppressed for blind review.} tool-chain \cite{ahirUsenix2012} 
-uses this compiler flow to map modular network appliances written in C++ to
-an FPGA platform, and achieved about half the
-performance possible using equivalent hand designed hardware.  
-Thus,  the flow can be applied to ``real'' software
-to produce competitive results.  The optimizations indicated in this
-paper, should in principle, make the flow even more competitive.
-
-%% compiler techniques.
-
-%% hls techniques to handle loops?
+of applications \cite{ahirDsd2010}, \cite{ahirUsenix2012}.
 
@@
 \subsection{Control-path}
+
 %%
@@
 
-\section{A control-flow mechanism for loop-pipelining}
+\section{A control-flow mechanism for dynamic loop-pipelining}
+
+For the subsequent discussion, we assume that the inner
+loop which is being optimized consists of a single
+basic block, that is, there are no branching instructions
+in the loop body (no jumps, if constructs, switch constructs etc.).
+If such constructs are present, these are first eliminated
+using the mechanism of guarded (that is, predicated) execution.
 
@@
 of the code as it is written.  Any ordering of these operations which
-satisfies these dependencies will be termed a {\em fully-unrolled consistent}
+satisfies these dependencies will be termed a {\em loop-consistent}
 ordering.
@@
 that the loop execution subject to these dependencies is a
-fully-unrolled consistent ordering.  
+loop-consistent ordering.  
 The modified control path is shown in Figure \ref{fig:pipelinedCP}.
@@
 
-We considered three examples, the dot product, the fast-fourier-transform,
+We considered four examples:
+\begin{itemize}
+\item Three examples where the loop body was a single basic block: the dot product, the fast-fourier-transform,
 and matrix multiplications, each with critical inner loops.  Four 
@@
 loop.
+\item One example where the loop body exhibits branching: a stream processor kernel which
+operates on a stream of numbers and performs operations on the stream depending on
+an op-code stream.  This example illustrates that the dynamic loop-pipelining mechanism
+is effective in complex loop bodies as well.
+\end{itemize}
 
@@
 
+\section{A stream processor}
+
+The following loop was used to test a situation in
+which the loop body has branching.  
+\begin{verbatim}
+while(1)
+{
+  float x = read_float32("x_pipe");
+  float y = read_float32("y_pipe");
+  uint8_t op_code = read_uint8("op_pipe");
+
+  float result = 0;
+  if(op_code == 0)
+    result = x*y;
+  else if(op_code == 1)
+    result = x+y;
+  else if(op_code == 2)
+    result = (x*x) - (y*y);
+  else if(op_code == 3)
+    result = (x + y) * (x + y);
+  else
+    result = 0;
+  write_float32("z_pipe",result);
+}
+\end{verbatim}
+In this loop, $x,y$ and $op\_code$ are read from
+three input streams.  Depending on the value of
+$op\_code$, a result is calculated and written out
+to an output stream.
+
+In order to pipeline this loop, the conditional
+statements are first eliminated using guards.   This
+is done by calculating the predicates 
+\begin{verbatim}
+(op_code == 0)
+(op_code == 1)
+(op_code == 2)
+(op_code == 3)
+\end{verbatim}
+and using these predicates to guard the execution of
+the statements which depend on these conditions.  This
+is done automatically in our compiler.
+
+The observations are shown in Table \ref{table:streamProcessor}.
+The time reported is that needed to process 16 elements from
+the streams (that is, to complete 16 iterations of the loop).
+\begin{table}[htb]
+  \centering
+  \caption{Stream processor inner loop observations: pipelined versus non-pipelined}
+  \label{table:streamProcessor}
+  \renewcommand\arraystretch{1.2}
+  \setlength{\tabcolsep}{1ex}
+  \begin{tabular}{c|c|c}
+  \hline
+  & plain & pipelined \\
+  \hline
+
+Cycles   &    2913  &   334  \\
+LUTs     &    4603  &  12501  \\
+FFs      &    4248  &  8808  \\
+Freq.(MHz)  & 184.9  &   103.9 \\
+Norm. Perf. & 1      &  4.86 \\
+Norm. Perf/Cost & 1  &  2.01  \\
+\hline
+  \end{tabular}
+\end{table}
+
+We observe a 4X improvement in performance and a 2X improvement
+in the performance/cost ratio.
+
 \section{Conclusion}
@@
 
-The data obtained from three common and important inner loop
-kernels is encouraging.  In all three cases, both loop-pipelining
+The data obtained from four inner loop
+kernels is encouraging. The first three  were examples in
+which the inner loop body consisted of a single basic 
+block. In these cases, both loop-pipelining
 and loop-unrolling lead to substantial performance gains.  Further,
@@
 the loop optimizations.
+In the fourth inner loop kernel, we considered a loop body
+which had branching.  In this case, considerable performance
+and performance/cost gains were observed when loop pipelining
+was enabled.
 
-Loop-unrolling is a common parallelization optimization
-in compilers that target pipelined processors.  Our data indicates
-that such unrolling offers significant benefits (up to an 
-order of magnitude) in an algorithm-to-hardware translation.
-Further, the use of hardware based dynamic loop-pipelining techniques 
-seems to offer a further boost in performance in the synthesis
-of hardware from single-threaded programs.  
+Thus, the use of hardware based dynamic loop-pipelining techniques 
+offers a significant boost in performance in 
+hardware synthesized from single-threaded programs.  
 The performance boost provided by the dynamic loop-pipelining

### 2015-02-07 Documents/AhirV2/AhirV2.tex
Update the Ahir docs with Aa2C related stuff

@@
 are extracted as pipelined do-while loops.  This is necessary
-for automatic parallelization of inner loops in the 
+for automatic cross-iteration parallelization of inner loops in the 
 generated hardware (substantial performance benefits can be
@@
 
+\subsection{{\bf Aa2C}: convert an {\bf Aa} description into a {\bf C} program}
+
+The AhirV2 flow offers considerable flexibility to a system designer.
+For example, it is possible to write code directly in {\bf Aa} in order
+to get more optimal implementations (relative to those obtained 
+starting from {\bf C}).  In such cases, if we wish to simulate
+the {\bf Aa} description, we would use the {\bf Aa2C} utility to
+convert the {\bf Aa} code to ANSI {\bf C}, and then compile it
+in the usual way.
+
+The {\bf Aa2C} program can be summarized as
+\begin{verbatim}
+Aa2C [-I <ext-mem-object>] <aa-file> (<aa-file>)*
+\end{verbatim}
+The only option is:
+\begin{itemize}
+\item {\bf -I <ext-mem-object>}:  the same behaviour as in {\bf Aa2VC}.
+\end{itemize}
+The remaining arguments are {\bf Aa} files which will be linked and
+converted to {\bf C} code.  Two outputs files are created:
+\begin{itemize}
+\item aa\_c\_model.h :  a header file declaring functions in the generated
+source code.
+\item aa\_c\_model.c : a source file containing function definitions
+corresponding to the {\bf Aa} modules.
+\end{itemize}
+
+External calls into the generated {\bf C} code must have the
+form:
+\begin{verbatim}
+void foo ( Ctype_1 in_1, Ctype_2 in_2, Ctype_3* out_1, Ctype_4* out_2);
+\end{verbatim}
+where Ctype is either a float or double or (int/uint)(64/32/16/8)\_t type.
+You can then link your external code with the generated {\bf C} code
+in the usual way.
+
+\subsubsection{Restrictions in using {\bf Aa2C}}
+
+The current implementation of {\bf Aa2C} produces un-threaded code.
+Thus, if you have a parallel block in your {\bf Aa} code, the
+statements in the parallel block are serialized in the resulting
+{\bf C} program.  This can result in the generated {\bf C} program
+potentially hanging (if one of the statements in the parallel
+block runs for-ever).   Another situation is when two
+concurrent blocks in the {\bf Aa} program are writing and reading
+from the same pipe.  In such a case, the serialized code may
+get dead-locked.  You need to be careful that your {\bf Aa} code
+does not have such situations (the simplest option is
+to not use parallel blocks in the {\bf Aa} code!).
+
+This issue will be fixed in a future release of {\bf Aa2C}.
+
 \subsection{Miscellaneous: {\bf vcFormat} and {\bf vhdlFormat}}

### 2015-02-11 Documents/Presentations/eeColloquium/Talk.tex
Added ee colloquium talk.

--- 2013-03-25 Documents/Presentations/eurasip/Talk.tex
@@
 \usepackage{epsfig}
-\title{IV EURASIP Seminar on the Hardware Design of DSP Systems\\
-Program-to-hardware transformation and its use in heterogenous high-performance DSP}
+\title{Mapping algorithms to hardware\\ the AHIR story}
 \author{Madhav P. Desai \\ Department of Electrical Engineering\\
   Indian Institute of Technology -- Bombay, Powai, Mumbai -- 400076, INDIA}
-\date{March 6, 2013}
+\date{January 21, 2015}
 \begin{document}
@@
 
-\frame[containsverbatim]{\frametitle{Overview}
-
-High-performance DSP can take advantage of several platforms:
-\vspace{0.1in}
+\frame[containsverbatim]{\frametitle{The premise}
 
 \begin{itemize}
-\item Processors such as multi-cores, GPUs: ``easy'' to use, but power
-	dissipation is an issue.
-\item Programmable hardware:  lots of parallelism in principle, potential
-	power benefits, difficult to map algorithms to  platform.  Can they
-	beat power-performance characteristics of processors?
-\item ASICs:  lots of parallelism, high efficiency, will surely beat power-performance
-	characteristics of processors.  Difficult to design, and expensive to
-	manufacture.
+\item Hardware capacity has scaled, but design capability has remained stagnant.
+\begin{itemize}
+\item evidence: anecdotal, most chips today are cut-and-paste/repetitive.
 \end{itemize}
-
-In terms of efficiency (performance/watt):
-\begin{verbatim}
-ASIC  > >  Processor, FPGA
-Processor < ? >  FPGA 
-\end{verbatim}
+\item Hardware capability/efficiency is not fully exploited in implementing complex algorithms.
+\end{itemize}
 
@@
 
+}
+
+\frame[containsverbatim]{\frametitle{Acknowledgements}
+
+Sameer Sahasrabudhe worked out many of the basic ideas as part of
+his Ph.D. thesis  and also contributed most of the initial code for the first version of the tools. 
+Thanks to Kavi Arya for many interesting discussions and inputs.  Maybe some day we will have
+a front end for functional programming languages.
+Kartik Lakhotia contributed to the FPGA implementation effort.
+Thanks to Teemu Rinta-Aho and G. Caffarena for providing valuable user feedback.
 }
@@
 
-\frame[containsverbatim]{\frametitle{A system in AHIR}
+\frame[containsverbatim]{\frametitle{The starting point: the algorithm}
+We view the algorithm as a program:
+\begin{itemize}
+\item which describes the behaviour of a collection of static threads (no spawning on the fly).
+\item the threads interact using shared memory and FIFOs.
+\item can be expressed in C.
+\end{itemize}
+}
+
+\frame[containsverbatim]{\frametitle{The end point: a logic circuit}
 
@@
 \begin{itemize}
-\item A system description consists of a collection of modules (a module is
+\item A collection of modules (a module is
 similar to a function in a C program).  The modules communicate
@@
 
+\frame[containsverbatim]{\frametitle{Type-2 petri-nets} 
+Type-2 petri-nets\footnote{Introduced in Sameer Sahasrabudhe's Ph.D. thesis} 
+are live and safe petri-nets constructed using a
+set of simple rules.  The simplest type-2 petri-net has the form
+shown in the figure.
+\begin{figure}
+\begin{centering}
+\centerline{\psfig{figure=Type2Base.eps,width=2.5in,height=2.0in}}
+\end{centering}
+\end{figure}
+}
+
+\frame[containsverbatim]{\frametitle{Production rules} 
+Replace a transition by a transition-place-transition (any region produced
+from a transition is called a t-region):
+\begin{figure}
+\begin{centering}
+\centerline{\psfig{figure=Type2T1.eps,width=2.5in,height=1.5in}}
+\end{centering}
+\end{figure}
+}
+
+\frame[containsverbatim]{\frametitle{Production rules} 
+Replace a transition by an acyclic graph constructed using
+transitions and p-regions: the acyclic graph must have a
+single source and a single sink.
+\begin{figure}
+\begin{centering}
+\centerline{\psfig{figure=Type2T2.eps,width=2.5in,height=2.0in}}
+\end{centering}
+\end{figure}
+}
+
+\frame[containsverbatim]{\frametitle{Production rules} 
+Replace a place by a possibly cyclic graph consisting of
+places and t-regions.
+\begin{figure}
+\begin{centering}
+\centerline{\psfig{figure=Type2P1.eps,width=2.3in,height=2.0in}}
+\end{centering}
+\end{figure}
+}
+
+\frame[containsverbatim]{\frametitle{Production rules: pipelines} 
+Using token-preserving regions, we can construct a well-behaved
+pipeline.
+\begin{figure}
+\begin{centering}
+\centerline{\psfig{figure=PipelineTPR.eps,width=4.5in,height=2.0in}}
+\end{centering}
+\end{figure}
+
+Note: this can be generalized to DAG.
+}
+
+\frame[containsverbatim]{\frametitle{Liveness and Safety}
+
+\begin{itemize}
+\item   Assume that an enabled input transition eventually fires.  Then the type-2
+petri-net is live and safe.
+\item The type-2 petri-net is also powerful enough to describe the control
+flow in imperative languages such as C/C++ as well as in synchronous
+languages such as Esterel.
+\end{itemize}
+
+}
+
 \frame[containsverbatim]{\frametitle{The data path ${\bf DP}$}
@@
 
+\frame[containsverbatim]{\frametitle{CP-DP interactions}
+\begin{figure}
+\begin{centering}
+\centerline{\psfig{figure=CPxDPxInteraction.eps,width=4.0in,height=3.0in}}
+\end{centering}
+\end{figure}
+}
+
 \frame[containsverbatim]{\frametitle{The storage subsystem ${\bf S}$}
@@
 several distinct memory subsystems).
-\item The memory subsystems have a relaxed consistency model: all accesses are time-stamped,
+\item The memory subsystems impose a strong consistency model: all accesses are time-stamped,
 and accesses to the same memory location are finished in first-come-first-served
@@
 
-\frame[containsverbatim]{\frametitle{Dot-product circuit} 
+\frame[containsverbatim]{\frametitle{Dot-product virtual circuit} 
 \begin{figure}
@@
 \end{figure}
+}
+
+\frame[containsverbatim]{\frametitle{Optimizations} 
+\begin{itemize}
+\item Local dependency analysis: extract local parallelism.
+\item Operator sharing: e.g. non-concurrent operations can be mapped to the same
+physical operator.
+\item Memory decomposition: different data structures can be mapped to different
+memories.
+\item Loop pipelining: the hardware can keep multiple iterations of a loop in
+progress so that throughput is improved.
+\end{itemize}
+}
+
+\frame[containsverbatim]{\frametitle{Memory decomposition: static aliasing analysis}
+\begin{verbatim}
+int A, B, C, D; // storage objects.
+int foo()
+{
+   int ret_val = 0;
+   int* p1 = &A;
+   ret_val += *p1;
+   int *p2 = &B;
+   ret_val += *p2;
+   int *p3 = &C;
+   ret_val += *p3;
+   p3 = &D;
+   ret_val += *p3;
+   return(ret_val);
+}
+\end{verbatim}
+}
+
+\frame[containsverbatim]{\frametitle{Memory decomposition: grouping}
+
+\begin{itemize}
+\item The memory spaces are $\{A\}$, $\{ B\}$, $\{ C, D \}$.
+\item $C$ and $D$ are grouped together because pointer $p3$ can
+point to either $C$ or $D$.
+\end{itemize}
+
+}
+
+\frame[containsverbatim]{\frametitle{Loop pipelining}
+\begin{verbatim}
+void Daemon()
+{
+  while(1)
+  {
+    float a = read_float32("a_pipe");
+    float b = read_float32("b_pipe");
+    float c = read_float32("c_pipe");
+    write_float32("d_pipe", (a+b)*c);
+  }
+}
+\end{verbatim}
+}
+
+\frame[containsverbatim]{\frametitle{Loop Pipelining: dependencies across iterations}
+\begin{itemize}
+\item Loop pipelining involves executing multiple iterations of a
+loop at the same time.
+\item While doing this, all dependencies must be taken care of.
+\begin{itemize}
+\item Operation order.
+\item Data dependency.
+\item Memory accesses.
+\item Pipe accesses.
+\end{itemize}
+\end{itemize}
+Currently, we are a bit conservative.  High performance compiler
+community has made substantial progress here.
 }
@@
 CFD, Databases etc.
-\item More optimizations:  latency reduction in non-parallel paths.
-\item Algorithm-to-ASIC.
-\item OpenMP, OpenCL support.
+\item Optimizations:  latency reductions, alternative DP structures.
+\item Algorithm-to-ASIC, potential asynchronous implementations.
+\item Front-end stuff:
+\begin{itemize}
+\item Other front-ends (going beyond C), OpenMP, OpenCL support.
+\item More aggressive compiler optimizations.
+\end{itemize}
 \end{itemize}
@@
 it out, please get in touch with me: madhav@ee.iitb.ac.in
+or search on github (standard disclaimers apply).
 

### 2015-02-22 v2/Ahir-asm/devel/docs/AaLRM.tex
Get the documentation in better shape.

@@
                     $endif
-               } 
+               } (s => sb1_s)
                $branchblock[sb2]{
@@
                     $endif
-               } 
-           }
+               } (s => sb2_s)
+           }(sb1_s => sb1_s sb2_s => sb2_s)
 
            // combine results from parallel statement above
-           snontriv := (%parsum%sb1:s  + %parsum%sb2:s)
+           snontriv := (sb1_s  + sb2_s)
            $place [nontrivsum]

### 2015-02-22 v2/Ahir-asm/devel/docs/AaLRM.tex
Get the documentation into better shape.

@@
 \begin{verbatim}
-phi-spec target-expr := expr1 on-spec place-spec-1  [expr2 on-spec place-spec-2 ... ]
+phi-spec target-expr := expr1 on-spec place-spec-1  
+                         [expr2 on-spec place-spec-2 ... ]
 \end{verbatim}
@@
 token in its default entry place, one of the labels in the
-label-list for the merge must contain the identifier \$entry.
+label-list for the merge must contain the identifier \$entry,
+which corresponds to the merge statement being triggered
+by the immediately preceding statement.
 The target of a phi-statement must be an implicit variable

### 2015-10-02 v2/Ahir-asm/devel/docs/AaLRM.tex
Fixed bug in generating volatile vhdl model.

@@
 
-\section{Inline and Macro Modules}
+\section{Module classes: pipeline, volatile, operator, inline, macro, foreign }
 
@@
 more flexibility,  a module in \Aa may be marked
-to be inlined or substituted into the program
-at the point of the call.  The AhirV2 utility
-{\bf AaOpt} is responsible for this substitution.
+as belonging to a particular class.  These
+are described below.
+
+\subsection{Pipelined Modules}
+
+A module is marked as a pipeline-module if its
+hardware implementation is to be converted into 
+a pipeline.  This allows efficient use of the module
+resources.
+
+\begin{verbatim}
+$pipeline $module [foo] $in (...) $out (...) $is
+{
+...
+}
+\end{verbatim}
+
+A pipeline module {\bf must} have {\bf at least} one input
+and one output argument.  Further, at least one input argument
+must actually be sampled inside the module, and one output argument
+must actually be updated inside the module.
+
+\subsection{Operator Modules}
+
+A module is marked as an operator-module if it
+is to be substituted at every point where it
+is called (thus, multiple instances of the module
+will be generated).
+This allows low latency access to the module
+functionality in hardware.
+
+\begin{verbatim}
+$operator $module [foo] $in (...) $out (...) $is
+{
+...
+}
+\end{verbatim}
+
+An operator module {\bf must} have {\bf at least} one input
+and one output argument.  Further, at least one input argument
+must actually be sampled inside the module, and one output argument
+must actually be updated inside the module.
+
+An operator module may also be marked as pipelined.
+\begin{verbatim}
+$pipeline $operator $module [foo] $in (..) $out (..)
+\end{verbatim}
+
+\subsection{Volatile Modules}
+
+A module is marked as a volatile-module if 
+\begin{itemize}
+\item Its hardware implementation is to be 
+a combinational (memory-less) circuit.
+\item It is to be substituted at every point where it
+is called (thus, multiple instances of the module
+will be generated).
+\end{itemize}
+This allows low level implementation control
+to the programmer.
+
+\begin{verbatim}
+$volatile $module [foo] $in (...) $out (...) $is
+{
+...
+}
+\end{verbatim}
+
+A volatile module cannot make any memory accesses or 
+reads/writes to pipes.  Inside a volatile module, 
+a function call must be to volatile modules only.
 
@@
 is clearly seen.  
+There are no restrictions on the
+module that is being inlined.
 
@@
 within a module.
+
+There are no restrictions on the
+module that is being inlined.
 

### 2015-11-27 v2/Ahir-asm/devel/docs/AaLRM.tex
Update LRM

@@
 \begin{verbatim}
-($lifo)? $pipe <pipe-name>: <pipe-type>  ($depth <integer>)?
+($lifo | $noblock)? $pipe <pipe-name>: <pipe-type>  ($depth <integer>)?
 \end{verbatim}
@@
 $lifo $pipe mystack: $uint<32> $depth 4 // 4-slot stack
-\end{verbatim}
+$noblock $pipe cmd_pipe: $uint<8> // non-blocking reads.
+\end{verbatim}
+A non-blocking pipe when read, will deliver $0$ if
+it does not contain data (thus, you will have to reserve
+the $0$ for invalid data).
+
 Pipe variable declarations can also be appended by
@@
 \begin{verbatim}
-   ($in | $out)? ($port)? ($signal)? 
+   ($in | $out)? ($signal)? 
 \end{verbatim}
@@
 that the pipe is a read-only (respectively write-only) object.
-\item The ``\$port'' keyword means that the pipe is meant to
-be connected to the outside world {\em without} any handshake,
-such as for example in an asynchronous interface to the outside
-world.
-\item The ``\$signal'' keyword means that the interface to
-the external world is a single bit.
+\item The ``\$signal'' keyword means that the pipe behaves
+like a variable.  That is, it remembers the last value
+that was written to it.    
 \end{itemize}

### 2016-03-18 v2/Ahir-asm/devel/docs/AaLRM.tex
Lots of clean up.  Moved critical examples to regression-tests folder. Can we automate the regression testing?
Also lots of clean up around the PHI handling.  We will treat the
PHI block as a set of concurrent statements.  This allows for
cleaner handling, and why would you want to have dependencies
across the PHI's anyway.

@@
 
+\subsubsection{Careful! The phi-statements in a merge block are concurrent}
+
+Note that the phi-statements in a merge are concurrent, and
+may execute in any order.
+An example:
+\begin{verbatim}
+  $merge loopback $entry // $entry is the entry place
+     $phi p := 0 $on $entry np $on loopback
+     $phi q := 0 $on $entry p  $on loopback
+  $endmerge
+\end{verbatim}
+In this case, the two $phi$ statements may execute
+in any order, and thus the value on $q$ after the execution
+is indeterminate.  To prevent such race conditions, do not
+use the result of a $phi$ as an input to another $phi$ in
+the same merge block.
+
 \subsection{Pipelinable-loops: the {\bf dopipeline-while} construct}
@@
 \end{verbatim}
+
+\section{Specifying buffering for expressions}
+
+By default, the result of each expression may be
+buffered by the AHIR compilers.  The default amount of
+buffering is $1$.  If you wish to specify more buffering, 
+say $2$ units, then we add 
+\begin{verbatim}
+$buffering 2
+\end{verbatim}
+after the expression.
+
+For example
+\begin{verbatim}
+( A $buffering 2  + (B + C) $buffering 3) $buffering 2
+\end{verbatim}
+means that the expression $A$ has buffering $2$,
+the expression $(B+C)$ has buffering $3$ and the
+overall expression $(A + (B+C))$ has buffering $2$.
 

### 2016-03-18 v2/Ahir-asm/devel/src/AaObject.cpp

@@ -154,7 +154,6 @@ void AaObject::PrintC_Declaration(ofstream& ofile)
 {
 	AaType* t = this->Get_Type();
 	AaModule* m = this->Get_Module();
-	
 	bool static_flag = ((m != NULL) && m->Static_Flag_In_C());
 	Print_C_Declaration(this->C_Reference_String(), static_flag, t, ofile);
 
@@ -178,6 +177,7 @@ void AaObject::PrintC_Global_Initialization(ofstream& ofile)
 {
 	AaType* t = this->Get_Type();
 	Print_C_Global_Initialization(this->C_Reference_String(), t,  ofile);
+
 	if(_value != NULL)
 	{
 		this->_value->Evaluate();

### 2016-03-23 v2/HierSysBuild/include/hierSystem.h
Started writing system partitioner to produce hw/sw  groups from common hsys description.

@@ -145,6 +145,7 @@ class hierSystemInstance: public hierRoot
 	public:
 
 	hierSystem* _parent;
+	hierSystemInstance* _parent_instance;
 	hierSystem* _base_system;
 	map<string, string> _port_map;
 	map<string, string> _reverse_port_map;
@@ -184,10 +185,14 @@ class hierSystemInstance: public hierRoot
 	void Print_Vhdl(ostream& ofile);
 };
 
+class hierInstanceGraph;
 class hierSystem: public hierRoot
 {
         string _library;
 	hierSystem* _parent;
+
+
+	set<hierPipe*> _pipe_set;
 
 	map<string, hierPipe* > _pipe_map;
 
@@ -241,7 +246,17 @@ public:
 	void List_Out_Pipe_Names(vector<string>& pvec);
 	void List_Internal_Pipe_Names(vector<string>& pvec);
 	
-	
+	void List_In_Pipes(vector<hierPipe*>& pvec);
+	void List_Out_Pipes(vector<hierPipe*>& pvec);
+	void List_Internal_Pipes(vector<hierPipe*>& pvec);
+
+	hierPipe* Find_Pipe(string pname)
+	{
+		if(_pipe_map.find(pname) != _pipe_map.end())
+			return(_pipe_map[pname]);
+		else	
+			return(NULL);
+	}	
 
 	void Add_Noblock_Pipe(string pname)
 	{
@@ -363,10 +378,12 @@ public:
 		{
 			p = new hierPipe(pid, pipe_width, pipe_depth);
 			_pipe_map[pid] = p;
+			_pipe_set.insert(p);
 		}
 		return(p);
 	}
 
+	bool Has_Pipe(hierPipe* p) {return(_pipe_set.find(p) != _pipe_set.end());}
 	void Add_Pipe_To_Map(map<string, hierPipe*>& pipe_map, string pipe_name, int width, int depth,
 					string pipe_type)
 	{
@@ -607,9 +624,159 @@ public:
 	string Get_Pipe_Vhdl_Read_Ack_Name(string pipe);
 
 	void Print_Vhdl_Rtl_Type_Package(ostream& ofile);
+
+	void Build_Instance_Hierarchy(hierInstanceGraph** instance_graph);
+	void Build_Instance_Graph_Arcs (hierInstanceGraph* instance_graph);
+
+};
+
+class hierInstanceGraph;
+class hierPipeInstance: public hierRoot
+{
[…]

### 2016-04-11 v2/Ahir-asm/devel/docs/AaLRM.tex
Started writing system partitioner to produce hw/sw  groups from common hsys description.
Updated a bit

@@
 in hardware, and the amount of buffering per operation will be kept to
-2.
+at least 2 (or any specified integer).  Further the buffering specification implies that at most
+2 versions of the same operations will be kept in flight across multiple
+loop iterations.
 
@@
 \end{verbatim}
+When the full-rate is specified, the generated circuit will
+be aggressively implemented to obtain a 1-cycle loop initiation
+latency.  This may lead to long combinational paths, which can
+be controlled by using double buffering.
 
@@
 
-\section{Scoping Rules} \label{sec:Scopes}
+\section{Scoping Rules (Deprecated!) } \label{sec:Scopes}
 
@@
 
+Pipeline modules can have several parameters as follows:
+\begin{verbatim}
+$pipeline ($depth <integer>)? ($buffering <integer>)? ($fullrate)?
+        $module [foo] $in (...) $out (...) $is
+\end{verbatim}
+The pipeline parameters are similar to the {\bf \$dopipeline} 
+construct.
+
 \subsection{Operator Modules}
@@
 \end{verbatim}
+
+An operator module is not allowed to access pipes from within.
+Inside an operator module, function calls can be made only
+to operator/volatile modules.
 

### 2016-06-19 v2/Ahir-asm/devel/docs/AaLRM.tex
Added decode, encode, bitreduce operators to Aa.

@@
 \item Unary expressions.
+\item Useful Unary logical expressions.
 \item Slice expressions.
@@
 {\bf required}.
+
+\subsection{Useful Unary Logical expressions}
+
+Some very useful logical building blocks are decoders,
+encoders and bit-reduction operators.
+
+\begin{itemize}
+\item The unary expression
+\begin{verbatim}
+($encode x)
+\end{verbatim}
+expresses the encoded version of $x$ (produces a \$uint 
+whose length is $\log{length(x)}$).  It operates only on \$uint types,
+and produces an encoding of $x$ under the assumption that only
+one of the bits in $x$ is $1$.
+\item The unary expression 
+\begin{verbatim}
+($p_encode x)
+\end{verbatim}
+expresses the priority-encoded version of $x$ (produces a \$uint which
+whose length is $length(x)$).  It operates only on \$uint types,
+and identifies the highest bit position of $x$ which is $1$. For example
+if $x$ is $01010$, then the expression will evaluate to $01000$.
+\item The unary expressions
+\begin{verbatim}
+($bitreduce | x)
+($bitreduce & x)
+($bitreduce ^ x)
+\end{verbatim}
+produce the reduced single-bit version of $x$ with the specified operation.
+For example, the bit-reduced OR is the OR of all the bits of $x$.
+\end{itemize}
 

### 2016-06-19 v2/libAhirV2/src/vcOperator.cpp

@@ -765,48 +765,76 @@ void vcUnarySplitOperator::Print_Flow_Through_VHDL(ostream& ofile)
 {
 	string X = this->Get_X()->Get_VHDL_Signal_Id();
 	string Z = this->Get_Z()->Get_VHDL_Signal_Id();
-	bool ip_is_float = (this->Get_X()->Get_Type()->Is("vcFloatType")); 
-	bool op_is_float = (this->Get_Z()->Get_Type()->Is("vcFloatType")); 
 
-	int ip_exponent_width, ip_fraction_width;
-	if(ip_is_float)
+	if(this->_op_id == "$decode")
 	{
-		ip_exponent_width = ((vcFloatType*) (this->Get_X()->Get_Type()))->Get_Characteristic_Width();
-		ip_fraction_width = ((vcFloatType*) (this->Get_X()->Get_Type()))->Get_Mantissa_Width();
+		ofile << Z << " <= GenericDecode("   << X << ");" << endl;
 	}
-	int op_exponent_width, op_fraction_width;
-	if(op_is_float)
+	else if(this->_op_id == "$encode")
 	{
-		op_exponent_width = ((vcFloatType*) (this->Get_Z()->Get_Type()))->Get_Characteristic_Width();
-		op_fraction_width = ((vcFloatType*) (this->Get_Z()->Get_Type()))->Get_Mantissa_Width();
+		ofile << Z << " <= GenericEncode("   << X << ");" << endl;
 	}
-
-	ofile << "-- unary operator " << this->Get_VHDL_Id() << endl;
-	ofile << "process(" << this->Get_X()->Get_VHDL_Signal_Id() << ") -- {" << endl;
-	ofile << "variable tmp_var : " << this->Get_Output_Type()->Get_VHDL_Type_Name() << "; -- }" << endl;
-	ofile << "begin -- { " << endl;
-	string vhdl_op_id  = Get_VHDL_Op_Id(this->_op_id, this->Get_Input_Type(), this->Get_Output_Type(), false);
-
-	if(ip_is_float & op_is_float)
+	else if(this->_op_id == "$priority_encode")
 	{
-		X = StdLogicToFloatConversion(X, this->Get_X()->Get_Type());
-		ofile << "ApFloatResize_proc(" << X << ", " << op_exponent_width << ", " 
[…]

### 2016-10-23 v2/Ahir-asm/devel/docs/AaLRM.tex
Added concept of parameters to allow abstraction of magic numbers to meaningful symbols/expresions.

@@
 \begin{itemize}
+\item {\bf Parameters} can be used to configure the Aa program to
+a certain extent.  These are 32-bit integers declared as
+\begin{verbatim}
+$parameter PIPEDEPTH 16
+$parameter PIPEDEPTH2  (2 * PIPEDEPTH) 
+\end{verbatim}
+for example.  
+Parameters can be used to specify some 
+constants inside
+the program (this will be indicated whenever relevant).
+Parameters
+should be known at the point at which they are used.  Also,
+parameters can be declared only at global program scope (that is
+not inside a module).
 \item {\bf Storage} variables declared in the program
@@
 \end{itemize}
+Pipe depths can be specified by parameters or parameter
+expressions.
 \item {\bf Constant} variables declared at
@@
 
+Note that the pipeline depth and buffering can be
+specified as parameters or parameter expressions.
+
 To summarize, an \Aa program is constructed as a collection of modules,
@@
 \end{verbatim}
-The width parameter can be any positive number.
+The width can be any positive integer (or a parameter
+expression that evaluated to a positive integer).
 Values corresponding to this type are to be
@@
 where the exponent and mantissa must be positive
-integers.  The float is represented by 
+integers (or parameter expressions that evaluate
+to positive integers).  The float is represented by 
 a word with $exponent+mantissa+1$ bits
@@
 where high-index and low-index are constant non-negative
-integers.  The result of executing the expression is
+integers (or parameter expressions that evaluate to
+non-negative integers).  The result of executing the expression is
 an unsigned integer whose width is $(high-index - low-index)+1$.
@@
 \begin{verbatim}
-( $slice ($bitcast ($uint<4>) _h3) 0 3 1 2 2 1 3 0 ) = _he
-\end{verbatim}
+( $bitmap ($bitcast ($uint<4>) _h3) 0 3 1 2 2 1 3 0 ) = _he
+\end{verbatim}
+
+Note: this expression is not currently supported in hardware,
+but will eventually be, maybe.
 

### 2016-10-23 v2/Ahir-asm/devel/docs/AaLRM.tex
Added parameter support to be used to parametrize an Aa program to a certain extent.

@@
 $parameter PIPEDEPTH 16
-$parameter PIPEDEPTH2  (2 * PIPEDEPTH) 
-\end{verbatim}
-for example.  
+$parameter PIPEDEPTH2  {2 * PIPEDEPTH} 
+\end{verbatim}
+for example.    Note that the parameter
+PIPEDEPTH2 is obtained by evaluating the
+parameter expression 
+\begin{verbatim}
+{2 * PIPEDEPTH} 
+\end{verbatim}
+Parameter expressions  are described in more detail Section \ref{sec:ParamExprs}.
 Parameters can be used to specify some 
@@
 \end{verbatim}
+\item Parameters can be used as simple identifiers.
+These are converted to integers (whose width as a uint
+is unknown) provided they are defined at the point
+of reference.
 \end{itemize}
@@
 overall expression $(A + (B+C))$ has buffering $2$.
+
+\section{Parameter Expressions} \label{sec:ParamExprs}
+
+Parameters are like constants, but have an implicit
+integer types.  They are to be used in lieu of
+constants in order to provide a measure of configurability
+to the Aa program, as well as to improve readability of
+the code.
+
+Pararmeters can be combined in expressions in the
+much the same way as other objects.  The only difference
+is that the delimitations is done using curly braces
+instead of parenthesis, for example
+\begin{verbatim}
+{ a + {b + c} }
+\end{verbatim}
+Here, $a,b,c$ must be parameters, and the expression
+is evaluated as soon as it is encountered and replaced
+by its equivalent integer value.
+
+All operations which are relevant for integers
+are supported (including the ternary operation).
 

### 2017-02-12 v2/Ahir-asm/devel/docs/AaLRM.tex
A lacuna in WAR dependency handling.  closed.
Update LRM to mention $mark $synch $delay constructs.

@@
 like a variable.  That is, it remembers the last value
-that was written to it.    
+that was written to it\footnote{    
+Note that a read from a signal will block until the
+signal has been written into at least once!}.
 \end{itemize}
@@
 \end{verbatim}
+
+Complete atomic statements can be explicitly sequenced using
+the \$mark and \$synch constructs.  For example
+\begin{verbatim}
+a := (b + c) $mark FIRST
+e := (f + g) $mark SECOND $synch (FIRST)
+h := (i + j) $synch (FIRST SECOND)
+\end{verbatim}
+means that the second statement will be {\em started}
+only after the first statement has {\em started}.  The
+third statement will be started only after the first
+two have started. This
+is necessary in some cases where dependency analysis
+says that two statements are independent, but you wish
+to force explicit ordering.
+
+A related construct is \$delay which explicitly sets
+the delay between the completion of two statements
+(which otherwise seem unrelated)
+\begin{verbatim}
+// from_pipe, to_pipe are pipes.
+to_pipe := (a + b) $mark WPIPE
+e := from_pipe  $delay (WPIPE 10)
+\end{verbatim}
+This says that there is delay of 10 units (clock cycles)
+between the write to to\_pipe and the read from from\_pipe.
+This delay information is used by the compiler to 
+balance pipelines.
 

### 2017-04-05 v2/libAhirV2/docs/vC_LRM.tex
Throw an error if a volatile statement involves a non-trivial expression.

@@
 token enters each of the elements in the list on the right.
+\item A pipelined-fork block defines a fork block which can support
+several active instantiations at the same time.
+\begin{verbatim}
+:|: [<name>] {
+   same-structure-as-fork
+   but-can-have-marked-joins
+}
+\end{verbatim}
+This uses a special join called a marked-join
+\begin{verbatim}
+jT o<- (s1 1 s2 1)
+(s1 can be $entry)
+\end{verbatim}
+Which says that jT is a join of s1 and s2, with a pre-condition
+that s1 is assumed to have fired twice and s2 is assumed to have fired
+twice before time began.  As a result it is possible for jT to fire
+twice before s1 and s2 fire in normal operation. 
 \end{itemize}
@@
 \begin{verbatim}
-<op-code> [operator-name] ( <input-wires> )  ( <output-wires>)
+($guard (~)? guard-wire)? <op-code> [operator-name] ( <input-wires> )  ( <output-wires>)
 \end{verbatim}
@@
 The add operation takes two inputs and produces a single output.
+Some other variants:
+\begin{verbatim}
+$guard (p) + [wadd1] (a b)  (c)
+\end{verbatim}
+signifies that the wadd1 operation is actually executed only if
+the wire $p$ is 1.
+\begin{verbatim}
+$guard (~p) + [wadd2] (a b)  (c)
+\end{verbatim}
+signifies that the wadd2 operation is actually executed only if
+the wire $p$ is 0.
 
@@
 ~^      logical-bitwise-xnor
+&&      concatenate
 \end{verbatim}
@@
 \end{verbatim}
-This operator has a unitary protocol, but there are as many
-req/ack pairs as there are inputs.  The output y
+This operator has a split protocol, but there are as many
+sample-req/sample-ack pairs as there are inputs.  There is a
+single update-req/update-ack pair.  The output y
 is assigned to an input for which a request has been made.
-\item The {\em concatenate} operator has two inputs and a single
-output.  The output bit sequence is a conctatenation of the input
-bit sequences.
-\begin{verbatim}
-&& [cname] (a b) (c)
-\end{verbatim}
-This operator also has a split protocol.
 \item The {\em slice} operator has three inputs, the first
@@
 \end{itemize}
+
+{\bf Note: All operators can be guarded!}
 

### 2017-04-17 v2/Ahir-asm/devel/docs/AaLRM.tex
LRM updated with  bitmap (permutation) stuff.

@@
 
-For example
-\begin{verbatim}
-( $bitmap ($bitcast ($uint<4>) _h3) 0 3 1 2 2 1 3 0 ) = _he
-\end{verbatim}
-
-Note: this expression is not currently supported in hardware,
-but will eventually be, maybe.
+For example, 
+\begin{verbatim}
+// X has type $uint<4>
+( $bitmap ($bitcast X) 0 3 1 2 2 1 3 0 ) = rX
+\end{verbatim}
+implements
+\begin{verbatim}
+rX[3] <- X[0], rX[1] <- X[2], 
+   rX[2] <- X[1], rX[0] <- X[3]
+\end{verbatim}
+and
+\begin{verbatim}
+// X has type $uint<4>
+( $bitmap ($bitcast X) 0 3 3 0) = rX
+\end{verbatim}
+implements (not that unmapped indices
+are left undisturbed)
+\begin{verbatim}
+rX[3] <- X[0], rX[1] <- X[1], 
+   rX[2] <- X[2], rX[3] <- X[0]
+\end{verbatim}
 

### 2018-03-05 Documents/Tutorials/Aa/Aa.tex
Updated the tutorial

@@
 \author{Madhav P. Desai}
+\date{March 8, 2018}
 \begin{document}
@@
 $lifo $pipe B: $uint<8> $depth 8
+$noblock $pipe C: $uint<39> $depth 2
 \end{verbatim}
-\item Implicit Variables:  these are defined by
+\item Implicit objects:  these are defined by
 statements:
@@
 or SSA variables.
+\item Interface objects:  Inputs and outputs of modules are
+treated like SSA variables.  Inputs cannot be written into.
+\end{itemize}
+}
+
+\frame[containsverbatim]{\frametitle{{\bf Aa} Storage Objects}
+\begin{itemize}
+\item  Will be implemented in memory spaces.
+\item  Access to a storage object takes 3+ cycles.
+\item  Should use sparingly.
+\end{itemize}
+}
+
+\frame[containsverbatim]{\frametitle{{\bf Aa} Pipe Objects}
+\begin{itemize}
+\item Blocking in nature:  provide synchronization mechanism.
+\item FIFO or LIFO data transfer within and across modules.
+\item Non-blocking pipes: return a 0 if read is attempted on an empty pipe.
+\end{itemize}
+}
+\frame[containsverbatim]{\frametitle{{\bf Aa} SSA Objects}
+\begin{itemize}
+\item Implemented as registers.
+\item Fast access (immediate read, single cycle write).
+\item Should use as often as possible.
+\item Note that the type of these objects is inferred from the
+context in which they appear.
+\begin{verbatim}
+a := (A + B)
+\end{verbatim}
+$A$ and $B$ must have the same type, and the type of $a$ is
+inferred to be the type of $A$.
 \end{itemize}
@@
 
+\frame[containsverbatim]{\frametitle{{\bf Aa} Phi Statements}
+\begin{itemize}
+\item Can appear only in branch blocks, and within merge statements.
+\item Indicate the action to be taken depending on how control got
+to this statement.
+\item  Example:
+\begin{verbatim}
+       $phi I := ($bitcast($uint<32>) 0) $on $entry
+                    NI $on loopback
+\end{verbatim}
+
+\end{itemize}
+}
+
+\frame[containsverbatim]{\frametitle{{\bf Aa} Do-Pipeline-While Statements}
+These are not atomic, and can occur only inside a branch-block.
+\begin{verbatim}
+$dopipeline $depth 7 $buffering 1 $fullrate
+   $merge $entry $loopback
+        $phi I := ($bitcast ($uint<32>) 0) 
+                      $on $entry
+                   NI $on $loopback
+   $endmerge
+   a[I] := (b[I] + c[I])
+   NI   := (I+1)
+$while (I < 16)
+\end{verbatim}
+Control-flow:  sequential, controlled by the condition check.  The
+places \$entry and \$loopback are implicitly defined.  When control
+enters the do-while, the token gets placed in \$entry and when
+control loops-back from the condition check, the token gets placed
+in \$loopback.  The compiler will pipeline the loop by keeping
+\$depth iterations alive.
+}
+
 \frame[containsverbatim]{\frametitle{{\bf Aa} Fork Block Statements}
@@
 
-\frame[containsverbatim]{\frametitle{{\bf Aa} Do-While Statements}
-These are not atomic, and can occur only inside a branch-block.
+\frame[containsverbatim]{\frametitle{{\bf Aa} Volatile Statements}
+\begin{itemize}
+\item If you do not want to use a register:
 \begin{verbatim}
-$do
-   $merge $entry $loopback
-        $phi I := ($bitcast ($uint<32>) 0) 
-                      $on $entry
-                   NI $on $loopback
-   $endmerge
-   a[I] := (b[I] + c[I])
-   NI   := (I+1)
-$while (I < 16)
+$volatile p := (q + r)
 \end{verbatim}
-Control-flow:  sequential, controlled by the condition check.  The
-places \$entry and \$loopback are implicitly defined.  When control
-enters the do-while, the token gets placed in \$entry and when
-control loops-back from the condition check, the token gets placed
-in \$loopback.
+\item $p$ will not necessarily be implemented as a register.   You
+can only assume that $p$ is equivalent to $(q  +  r)$.
+\item Useful in describing combinational logic.
+\end{itemize}
 }
 
-\frame[containsverbatim]{\frametitle{Getting to hardware starting from an {\bf Aa} Program}
+\frame[containsverbatim]{\frametitle{{\bf Aa} Summary}
+
+\begin{itemize}
+\item An {\bf Aa} program consists of a collection of module
+descriptions and object declarations.
+\item Each module is a sequence of statements.
+\item Statements can be simple or block-structured.
+\item Specific blocks allow branching, while others describe
+parallelism.
+\item Objects in an {\bf Aa} program can be FIFOs (pipes), storage
+or implicitly inferred.
+\item Almost everything you can express in C can be expressed, but
+we are closer to hardware.
+\end{itemize}
+}
+
+\frame[containsverbatim]{\frametitle{Look at the FIR again}
 }

### 2018-03-05 Documents/Tutorials/DesignProblem/DesignProblem.tex
Updated the tutorial

\documentclass{beamer}
\usepackage{graphics}
\title{An Exercise}
\author{Madhav P. Desai\\ Department of Electrical Engg.\\ IIT Bombay, Mumbai India}
\date{March 8, 2018}
\begin{document}
\maketitle

\frame[containsverbatim]{\frametitle{Specification}

\begin{itemize}
\item Design an 8-tap programmable FIR filter (PFIRF).
\item The PFIRF has a single 32-bit FIFO input and a single
32-bit FIFO output.
\item After reset, the PFIRF first reads the 8 tap values from the FIFO input,
and on subsequent inputs, starts functioning like an FIR.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{Possible solution}
\begin{verbatim}
void pfirf ()
{
    int i;
    float taps[8];
    float xvals[8];
    for (i = 0; i < 8; i++)
    {
       taps[i] = read_float32("in_data");
       xvals [i] = 0;
    }
\end{verbatim}
}

\frame[containsverbatim]{\frametitle{Possible solution: continued}
\begin{verbatim}
    int HEAD = 7;
    while (1) // outer loop.
    {
      // read input sample x and store it.
      float x = read_float32("in_data");
      xvals [HEAD] =  x;  
      // calculate the FIR output
      float output_val = 0.0;
      // inner loop..
      for (i = 0; i < 8; i++)
        output_val += xvals [(i + HEAD) & 0x7]*taps[7 -i]; 
      // Send FIR output on pipe out_data.
      write_float32 ("out_data", output_val);
      // Adjust head of circular queue.
      HEAD = (HEAD - 1) & 0x7;
    }
}
\end{verbatim}
}

\frame[containsverbatim]{\frametitle{Try to write a test-bench}
\begin{itemize}
\item Testbench should first send in the taps.
\item Then it should send in the filter inputs and
read out the filter outputs.
\item Try to do this in a burst so that you can check filter 
throughput.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{Makefile}
See the model in vanillaC/
}

\frame[containsverbatim]{\frametitle{Can we do better?}
\begin{itemize}
\item Can you optimize the performance?  
\begin{itemize}
\item Use loop pipelining.
\item Use loop unrolling.
\end{itemize}
\end{itemize}
}

\frame[containsverbatim]{\frametitle{How about writing the PFIRF in {\bf Aa}}
\begin{itemize}
\item Almost exactly like C, but use a branch block and do-while.
\item Inner loop is unrolled 4-times and pipelined.
\item 1.5X better performance with marginally more hardware.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{Makefile for Aa}

See the model in optAa/.   We will convert the Aa to C to 
verify the functionality before going to hardware.

}

\frame[containsverbatim]{\frametitle{Other interesting problems}
\begin{itemize}
\item Sorter:  read N integers, sort them and output the sorted list.
\item Correlator: Given an input sequence and a reference sequence, find
the amount by which the input sequence needs to be rotated to maximize
its correlation with the reference sequence. 
\item Etc..
\end{itemize}
}

\end{document}

### 2018-03-05 Documents/Tutorials/DpCpEtc/DpCpEtc.tex
Updated the tutorial

\documentclass{beamer}
\usepackage{graphics}
\title{Structure and organization of generated hardware}
\author{Madhav P. Desai\\Department of Electrical Engineering\\IIT Bombay, Mumbai, India}
\date{March 8, 2018}
\begin{document}
\maketitle

\frame[containsverbatim]{\frametitle{Top-level view}
\begin{figure}
  \centering
  \includegraphics[width=9cm]{figs/TopLevel.pdf}
  \caption{Structure of AHIR system}
\end{figure}
}

\frame[containsverbatim]{\frametitle{The Data-path: a graph of operators and wires}
\begin{figure}
  \centering
  \includegraphics[width=6cm]{figs/DataPath.pdf}
  \caption{Data-path in module of AHIR system}
\end{figure}
}

\frame[containsverbatim]{\frametitle{A generic operator in the data-path}
\begin{figure}
  \centering
  \includegraphics[width=8cm]{figs/GenericOperator.pdf}
  \caption{Generic operator in the data-path}
\end{figure}
}

\frame[containsverbatim]{\frametitle{Operator classes in the data-path}
\begin{itemize}
\item Pure operators: arithmetic, logic.
\item I/O port operators:  input and output ports.
\item Load/Store operators.
\item Shared Call/Return operators:  call-start and call-complete operators.
\item Function operators.
\item Branch operators.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{Input/Output operators in the data-path}
\begin{figure}
  \centering
  \includegraphics[width=8cm]{figs/IOPorts.pdf}
  \caption{Input/output port operators in the data-path}
\end{figure}
}

\frame[containsverbatim]{\frametitle{Load/Store operators in the data-path}
\begin{figure}
  \centering
  \includegraphics[width=8cm]{figs/LoadStore.pdf}
  \caption{Load/Store operators in the data-path}
\end{figure}
}

\frame[containsverbatim]{\frametitle{Call/Return operators in the data-path}
\begin{figure}
  \centering
  \includegraphics[width=10cm]{figs/CallReturn.pdf}
  \caption{Call/Return operators in the data-path}
\end{figure}
}

\frame[containsverbatim]{\frametitle{Function operators in the data-path}
\begin{figure}
  \centering
  \includegraphics[width=8cm]{figs/Operator.pdf}
  \caption{Function operators in the data-path}
\end{figure}
}

\frame[containsverbatim]{\frametitle{Branch operators in the data-path}
\begin{figure}
  \centering
  \includegraphics[width=6cm]{figs/BranchOperator.pdf}
  \caption{Branch operators in the data-path}
\end{figure}
}

\frame[containsverbatim]{\frametitle{Data-path: summary}
\begin{itemize}
\item  A  network of wires and data-path operators.
\item  The outputs of data-path operators are registered (buffered).
\item  The amount of buffering at the inputs and outputs can be controlled.
\item  Operators are triggered by control pulses and indicate completion by
status pulses.
\item  A split protocol is followed in all operators.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{The Control-path: a safe Petri-net}
\begin{itemize}
\item A Petri-net is a directed graph with two kinds of vertices
\begin{itemize}
\item Places, which can be thought of as containers which can hold a token.
\item Transitions, which can be thought of as events.
\end{itemize}
\item There are arcs from places to transitions and from transitions to places.
\item A transition is enabled when all its input places have a token.
\item An enabled transition can fire, and firing of a token removes a token
from all input places and puts a token into all output places.
\item The net is safe if it is impossible for a place to ever have more than
one token.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{Examples of Petri-nets}
\begin{figure}
  \centering
  \includegraphics[width=9cm]{figs/ExamplePetriNets.pdf}
  \caption{Examples of Petri-nets}
\end{figure}
}

\frame[containsverbatim]{\frametitle{Construction of control net}
\begin{figure}
  \centering
  \includegraphics[width=8cm]{figs/BaseCP.pdf}
  \caption{Starting point: control Petri-net}
\end{figure}
}

\frame[containsverbatim]{\frametitle{Series augmentation}
\begin{figure}
  \centering
  \includegraphics[width=10cm]{figs/SeriesCP.pdf}
  \caption{Series augmentation}
\end{figure}
}

\frame[containsverbatim]{\frametitle{Parallel augmentation}
\begin{figure}
  \centering
  \includegraphics[width=10cm]{figs/ParallelCP.pdf}
  \caption{Series augmentation}
\end{figure}
}

\frame[containsverbatim]{\frametitle{Branch augmentation}
\begin{figure}
  \centering
  \includegraphics[width=10cm]{figs/BranchCP.pdf}
  \caption{Branch augmentation}
\end{figure}
}

\frame[containsverbatim]{\frametitle{Fork augmentation}
\begin{figure}
  \centering
  \includegraphics[width=10cm]{figs/ForkCP.pdf}
  \caption{Fork augmentation}
\end{figure}
}

\frame[containsverbatim]{\frametitle{Implementation of control net}
\begin{figure}
  \centering
  \includegraphics[width=10cm]{figs/Join.pdf}
  \caption{Two-input join logic implementation}
\end{figure}
}

\frame[containsverbatim]{\frametitle{The control path:  a summary}
\begin{itemize}
\item Constructed as a Petri-net by following certain rules.
\begin{itemize}
\item The resultant Petri-net is safe (and hence can be implemented
as a logic circuit).
\end{itemize}
\item Implementation of the control path using logic gates is easy.
\begin{itemize}
\item Only join transitions need actual logic gates.
\end{itemize}
\item  Sufficient to model control-flow in most high level programming
languages.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{Other constituents of the AHIR hardware}
\begin{itemize}
\item Pipes:  FIFO with multiple writers and readers.
\item Memory sub-systems: first initiated first served guarantee.
\item Multiplexors: with tagged and untagged functionality.
\item Arbiters: fair arbitration (round-robin).
\end{itemize}
}

\frame[containsverbatim]{\frametitle{Pipes}
\begin{figure}
  \centering
  \includegraphics[width=10cm]{figs/Pipe.pdf}
  \caption{Pipes}
\end{figure}
}

\frame[containsverbatim]{\frametitle{Memory sub-system}
\begin{figure}
  \centering
  \includegraphics[width=10cm]{figs/MemorySubsystem.pdf}
  \caption{Memory sub-system}
\end{figure}
}

\frame[containsverbatim]{\frametitle{Lets take a look}
Go back to earlier example of FIR filter and check it out.
}
\end{document}

### 2018-03-05 Documents/Tutorials/Intro/Intro.tex
Updated the tutorial

@@
 \documentclass{beamer}
-\usepackage{epsfig}
+\usepackage{graphics}
 \title{An introduction to the AHIR-v2 tool-set}
-\author{Madhav P. Desai}
+\author{Madhav P. Desai\\ Department of Electrical Engg.\\ IIT Bombay, Mumbai India}
+\date{March 8, 2018}
 \begin{document}
@@
 \item Describe the algorithm (behaviour) of the system using high-level
-languages (currently C), and migrate the algorithm to a hardware
+languages, and migrate the algorithm to a hardware
 implementation.
+\begin{itemize}
+\item C.
+\item Aa.
+\end{itemize}
+\end{itemize}
+}
+
+\frame[containsverbatim]{\frametitle{Plan}
+\begin{itemize}
+\item Introduction: going from C to VHDL and simulating the VHDL.
+\begin{itemize}
+\item Exercise: go through a couple of examples.
+\end{itemize}
+\item Understand the hardware target model used by the AHIR-V2 tools.
+\begin{itemize}
+\item Exercise: re-visit the examples to take a closer look.
+\end{itemize}
+\item Optimizations.
+\item Loop-pipelining.
+\item {\bf Aa} language.
+\item Design problem.
+\begin{itemize}
+\item Design a programmable FIR filter.
+\end{itemize}
 \end{itemize}
@@
 \frame[containsverbatim]{\frametitle{For example}
+
 Design a circuit that takes two inputs, computes the maximum of
@@
 
+\frame[containsverbatim]{\frametitle{AHIR flow}
+\begin{figure}
+  \centering
+  \includegraphics[width=10cm]{figs/AhirFlow.pdf}
+  \caption{AHIR flow}
+\end{figure}
+}
+
 \frame[containsverbatim]{\frametitle{Conversion to hardware}
@@
 \item The source C code is compiled using the open-source compiler
-{\bf clang} (we version 2.8), which produces byte-code in {\bf llvm}
+{\bf clang} (we use version 2.8), which produces byte-code in {\bf llvm}
 format.
@@
 
+\frame[containsverbatim]{\frametitle{VHDL Structure}
+\begin{figure}
+  \centering
+  \includegraphics[width=10cm]{ex1/figs/VhdlStructure.eps}
+  \caption{VHDL structure of generated example}
+\end{figure}
+}
+
 \frame[containsverbatim]{\frametitle{Validating the VHDL}
@@
 \frame[containsverbatim]{\frametitle{Lets take a look}
-Go back to examples/ex1 and check it out.
+Go back to ex1/ and check it out.
 }
@@
 \begin{verbatim}
-void maxDaemon()
+float x_vals[2];
+void firDaemon()
 {
+  x_vals[0] = 0.0; x_vals[1] = 0.0;
+  uint8_t head_pointer = 0;
+  // fir loop.
   while(1)
   {
-       uint32_t a = read_uint32("in_data");
-       uint32_t b = read_uint32("in_data");
-       c = ((a > b) ? a : b);
-       write_uint32("out_data", c);
+    float x = read_float32("in_data");
+    float y = 0.8*x + (0.15*x_vals[head_pointer]) + 
+                (0.05*x_vals[(head_pointer+1)&0x1]);
+    write_float32("out_data",y);
+    head_pointer = (head_pointer + 1)&0x1;
+    x_vals[head_pointer] = x;
   }
@@
 {
-  uint32_t a, b, c;
   while(1)
   {
-     scanf("%d", &a);
-     scanf("%d", &b);
-     write_uint32("in_data", a);
-     write_uint32("in_data", b);
-     c = read_uint32("out_data");
-     printf(stdout,"Result = %d.\n", c);
+    float a;
+    scanf("%f", &a);
+    write_float32("in_data",a);
+    float c = read_float32("out_data");
+    fprintf(stdout,"Result = %f.\n", c);
   }
@@
 
-\frame[containsverbatim]{\frametitle{Installing AHIR-v2}
-
+\frame[containsverbatim]{\frametitle{Download and Install}
 \begin{itemize}
-\item You will need to install llvm-2.8, clang-2.8, the BOOST graph library.
-\item You can download the source code from gitHub.
-\item Go to the v2/ folder in the git repository.  You will need to
-source ``build\_bashrc'', and then type ``scons'' to build the executables.
-You can then make a release by going back to the root of the git repository
-and then typing ``make -f ReleaseMakeFile''.  This populates the Release
-directory with binaries, examples etc. which should be enough to get you
-going.
+\item Clone git repository
+\begin{verbatim}
+git clone https://github.com/madhavPdesai/ahir.git .
+\end{verbatim}
+\item Install llvm-2.8, clang-2.8 from www.llvm.org (from source, builds on ubuntu 12, centos 6.5, may need  to do minor edits in includes in ubuntu 14).
+\item Install boost development packages.
+\item Install antlr-2.7.7 (from source, builds on ubuntu 12, centos 6.5, some includes need to be added in ubuntu 14).
+\item Install java run-time (openjdk).
+\item Install ghdl (or modelsim).
+\item Go to release directory. 
+\begin{verbatim}
+source ahir_bashrc
+mkdir bin
+./build.sh
+\end{verbatim}
+\item Should work..
 \end{itemize}

### 2018-03-05 Documents/Tutorials/Optimizations/Optimizations.tex
Updated the tutorial

\documentclass{beamer}
\usepackage{graphics}
\title{Optimizations carried out by the AHIR tool chain}
\author{Madhav P. Desai\\Department of Electrical Engineering\\ IIT Bombay, Mumbai India}
\date{March 8, 2018}
\begin{document}
\maketitle

\frame[containsverbatim]{\frametitle{Optimizations}
\begin{itemize}
\item Dependency analysis.
\begin{itemize}
\item Reduce latency through a group of statements.
\end{itemize}
\item Loop pipelining.
\begin{itemize}
\item Increase throughput of a loop.
\end{itemize}
\item Operator sharing.
\begin{itemize}
\item Reduce cost of hardware.
\end{itemize}
\item Identification of disjoint memory spaces.
\begin{itemize}
\item Increase memory bandwidth.
\end{itemize}
\end{itemize}
}

\frame[containsverbatim]{\frametitle{Dependency analysis}
\begin{verbatim}
   a := (b + c)
   d := (b + c)
   e := (a + d)
\end{verbatim}
The first two statements can be executed concurrently, but the third depends
on the first two (RAW dependency).  Other dependencies that are analyzed and
enforced are WAR and WAW.  For pipe accesses, RAR dependencies are also imposed.
}

\frame[containsverbatim]{\frametitle{Loop pipelining}
A loop may be implemented as a pipeline in hardware.  We will discuss this
separately.
}

\frame[containsverbatim]{\frametitle{Operator sharing}

\begin{verbatim}
   a := (b + c)
   d := (b + c)
   e := (a + d)
\end{verbatim}
If $+$ can be replicated, we observe that only two copies of $+$ are required
to implement the three additions.

}

\frame[containsverbatim]{\frametitle{Storage Variables and Memory Spaces}
\begin{itemize}
\item Storage variables are grouped into memory spaces.
\item Each memory space is implemented in hardware by a memory sub-system. 
\item Memory spaces are identified by doing a static aliasing analysis.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{Aliasing analysis}
\begin{verbatim}
int A, B, C, D; // storage objects.
int foo()
{
   int ret_val = 0;
   int* p1 = &A;
   ret_val += *p1;
   int *p2 = &B;
   ret_val += *p2;
   int *p3 = &C;
   ret_val += *p3;
   p3 = &D;
   ret_val += *p3;
   return(ret_val);
}
\end{verbatim}
}

\frame[containsverbatim]{\frametitle{Groupings}

\begin{itemize}
\item The memory spaces are $\{A\}$, $\{ B\}$, $\{ C, D \}$.
\item $C$ and $D$ are grouped together because pointer $p3$ can
point to either $C$ or $D$.
\end{itemize}

}

\frame[containsverbatim]{\frametitle{Note: Orphan references}

An orphan reference is pointer which points to an unknown storage object.
In such cases, the orphan pointer is directed to a default storage object
(sort of like /dev/null).

\begin{verbatim}
AaLinkExtMem -I 1 -E mempool prog1.aa prog2.aa 
       | vcFormat > linked.aa
Aa2VC -E mempool  linked.aa ....
\end{verbatim}

All orphan pointers in linked.aa now point to mempool which is a dummy storage object whose
size is $1$ byte.  This mempool should never actually be accessed...

}

\end{document}

### 2018-03-05 Documents/Tutorials/Pipelines/Pipelines.tex
Updated the tutorial

@@
 \usepackage{epsfig}
-\title{Describing Pipelines}
-\author{Madhav P. Desai}
+\title{Pipelines:  a fundamental building block for efficient hardware}
+\author{Madhav P. Desai\\Department of Electrical Engineering\\IIT-Bombay, Mumbai India}
+\date{March 8, 2018}
 \begin{document}
@@
 
-\frame[containsverbatim]{\frametitle{Two ways of describing pipelines}
-
+\frame[containsverbatim]{\frametitle{An example}
 \begin{itemize}
-\item with distinct threads.
-\item by pipelining a loop.
+\item We want to compute $result(k) = a(k)*(a(k) + 1.0)$ for a sequence of 
+numbers. 
+\item The system should accept a sequence of numbers $\{a(k))\}$
+and produce the sequence $\{ result(k) \}$.
 \end{itemize}
@@
 
-\frame[containsverbatim]{\frametitle{An example}
-\begin{itemize}
-\item We want to compute $d(k) = (a(k)+b(k)) \times c(k)$ for a sequence of 
-numbers. 
-\item The system should accept a sequence of numbers $\{a(k),b(k),c(k)\}$
-and produce the sequence $\{ d(k) \}$.
-\end{itemize}
-}
-
-\frame[containsverbatim]{\frametitle{As separate threads}
-\begin{verbatim}
-void Stage_0()
-{
-  while(1)
-  {
-    float a = read_float32("a_pipe");
-    float b = read_float32("b_pipe");
-    float c = read_float32("c_pipe");
-    write_float32("c_forward_pipe", c);
-    write_float32("s0_result", a+b);
-  }
-}
-void Stage_1()
-{
-  while(1)
-  {
-    float c = read_float32("c_forward_pipe");
-    float aSb = read_float32("s0_result");
-    write_float32("d_pipe", aSb * c);
-  }
-}
-\end{verbatim}
-}
-
-\frame[containsverbatim]{\frametitle{In a single loop}
+\frame[containsverbatim]{\frametitle{The program}
 \begin{verbatim}
@@
     float a = read_float32("a_pipe");
-    float b = read_float32("b_pipe");
-    float c = read_float32("c_pipe");
-    write_float32("d_pipe", (a+b)*c);
+    float ap1 = (a + 1.0);
+    float result = a*ap1;
+    write_float32("result_pipe", result);
   }
@@
 
+\frame[containsverbatim]{\frametitle{The hardware we would like to obtain}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=9cm]{figs/ExpectedHardware.pdf}
+  \caption{The hardware we would like to see.}
+\end{figure}
+
+}
+
+\frame[containsverbatim]{\frametitle{The AHIR  tools produce the expected hardware}
+\begin{verbatim}
+void __loop_pipelining_on__(int,int,int);
+void Daemon()
+{
+  while(1)
+  {
+    // tell the AHIR compiler to pipeline this loop..
+    // and keep 32 iterations alive at any time.
+    __loop_pipelining_on__(32,2,1);
+
+    float a = read_float32("a_pipe");
+    float ap1 = (a + 1.0);
+    float result = a*ap1;
+    write_float32("result_pipe", result);
+  }
+}
+\end{verbatim}
+
+}
+
 \frame[containsverbatim]{\frametitle{Loop Pipelining}
 \begin{itemize}
-\item Loop pipelining involves executing multiple iterations of a
-loop at the same time.
+\item The  AHIR compiler modifies the control path to
+allow multiple iterations of the loop to be active simultaneously.
 \item While doing this, all dependencies must be taken care of.
@@
 \item Pipe accesses.
+\item Producer-consumer dependencies are the key.
 \end{itemize}
+\item The data-path is modified by inserting buffers to balance
+the pipeline.
 \end{itemize}
@@
 
-\frame[containsverbatim]{ \frametitle{Example}
+\frame[containsverbatim]{\frametitle{Control path modifications}
+\begin{figure}
+  \centering
+  \includegraphics[width=9cm]{figs/ProducerConsumer.pdf}
+  \caption{Producer consumer dependencies in the pipeline}
+\end{figure}
 }
+
+\frame[containsverbatim]{\frametitle{Cost of loop pipelining}
+\begin{itemize}
+\item In the data-path: additional buffering.
+\item In the control-path, the cost of each join increases by
+a factor of $\log N$, where $N$ is the number of iterations that
+need to be kept alive.
+\item Need to choose $N$ carefully and avoid unbalanced loop bodies
+in order to manage cost of loop pipelining.
+\item Performance benefits can be substantial.
+\end{itemize}
+}
+
+\frame[containsverbatim]{\frametitle{An example with and without loop-pipelining}
+
+\begin{itemize}
+\item Look at ex1/ and ex1\_no\_loop\_pipeling to see the difference.
+\item More than 20X speedup by pipelining the loop.
+\end{itemize}
+}
+
+\frame[containsverbatim]{\frametitle{Loop-pipelining is critical!}
+\begin{itemize}
+\item  Think of the hardware as interacting threads, each of which
+is a pipelined loop.
+\end{itemize}
+}
+
 \end{document}

### 2018-03-05 Documents/Tutorials/Optimizations/ex1/src/prog.h

@@ -1,0 +1,1 @@
+void vectorSum();

### 2018-03-05 Documents/Tutorials/Optimizations/ex1/src/prog0.c

@@ -1,0 +1,29 @@
+#include <stdlib.h>
+#include <stdint.h>
+#include <Pipes.h>
+#include <stdio.h>
+#include "prog.h"
+
+uint32_t doSum(uint32_t*, uint32_t);
+uint32_t A, B, C[10];
+void vectorSum()
+{
+	uint32_t D;
+	A = read_uint32("in_data");
+	B = read_uint32("in_data");
+
+	uint32_t ret_val = (A+B);
+
+	C[0] = read_uint32("in_data");
+	D = read_uint32("in_data");
+
+	uint32_t* ptr = &(C[0]);
+	ret_val = doSum(ptr, ret_val);
+
+	ptr = &D;
+	ret_val = doSum(ptr, ret_val);
+
+	write_uint32("out_data", ret_val);
+}
+
+

### 2018-03-05 release/examples/regression_tests/arrayAccessesViaPtr/src/testbench.c

@@ -17,28 +17,13 @@ void Exit(int sig)
 	exit(0);
 }
 
-	
-float expected_result[ORDER];
-void Sender()
-{
-	int idx;
-	float val[ORDER];
-	for(idx = 0; idx < ORDER; idx++)
-	{
-		val[idx] = drand48();
-		expected_result[idx] = (val[idx])*(val[idx] + 1.0);
-	}
-	write_float32_n("a_pipe",val,ORDER);
-}
-
 #ifdef SW
 DEFINE_THREAD(vectorSum)
 #endif
-DEFINE_THREAD(Sender)
-
 int main(int argc, char* argv[])
 {
-	float result[ORDER];
+	float result;
+	float expected_result[ORDER];
 	signal(SIGINT,  Exit);
   	signal(SIGTERM, Exit);
 
@@ -47,21 +32,24 @@ int main(int argc, char* argv[])
 	PTHREAD_DECL(vectorSum);
 	PTHREAD_CREATE(vectorSum);
 #endif
-	PTHREAD_DECL(Sender);
-	PTHREAD_CREATE(Sender);
 
 	uint8_t idx;
-	
-	read_float32_n("result_pipe",result,ORDER);
+	for(idx = 0; idx < ORDER; idx++)
+	{
+		float val = drand48();
+		expected_result[idx] = 2.0*val;
+		write_float32("in_data_pipe",val);
+	}
+
 
 	for(idx = 0; idx < ORDER; idx++)
 	{
-		fprintf(stdout,"Result = %f, expected = %f.\n", result[idx],expected_result[idx]);
+		result = read_float32("out_data_pipe");
+		fprintf(stdout,"Result = %f, expected = %f.\n", result,expected_result[idx]);
 	}
-	PTHREAD_CANCEL(Sender);
 #ifdef SW
 	close_pipe_handler();
 	PTHREAD_CANCEL(vectorSum);
-#endif
 	return(0);
+#endif
 }

### 2018-03-07 Documents/AhirV2/AhirV2.tex
Updated the tutorial etc.

@@
 \usepackage{epsfig}
+\usepackage{graphics}
 
@@
 \caption{AhirV2 flow}
+\end{centering}
  \label{fig:AhirFlow}
-\end{centering}
 \end{figure}
@@
 \begin{verbatim}
-entity test_system is  -- system
+entity ahir_system is  -- system
   port (--
@@
 \begin{verbatim}
-entity test_system is  -- system 
+entity ahir_system is  -- system 
   port (-- 
@@
 routines, the linker generates a global storage initialization
-function which is named {\bf global\_storage\_initializer\_}.
+function which is named 
+\begin{verbatim}
+global_storage_initializer
+\end{verbatim}
 This global initializer calls all the memory initializers in
@@
 corresponding to the AHIR system is named top-entity-name.  The default
-is test\_system.
+is ahir\_system.
 \item {-L function-library} : AhirV2 provides some built in operator

### 2018-03-07 Documents/Tutorials/Intro/Intro.tex
Updated.
Updated the tutorial etc.

@@
 
-\frame[containsverbatim]{\frametitle{We compile and run it}
-
-Go to ex1/ and check it out.
-
-}
-
 \frame[containsverbatim]{\frametitle{AHIR flow}
@@
 \end{figure}
+}
+
+\frame[containsverbatim]{\frametitle{We compile and run it}
+
+Go to ex1/ and check it out.
+\begin{itemize}
+\item In the Makefile, we define rules to produce a ``software'' executable (testbench\_sw)
+and two ``hardware'' executables, testbench\_hw  and ahir\_system\_test\_bench.
+\item You can execute testbench\_sw to verify that your C code is OK.
+\item You can concurrently execute testbench\_hw and ahir\_system\_test\_bench to
+verify that the the VHDL is OK.
+\item Note: the same testbench.c file is used to verify both the software and hardware.
+\end{itemize}
+
 }
@@
   \centering
-  \includegraphics[width=10cm]{ex1/figs/VhdlStructure.eps}
+  \includegraphics[width=10cm]{ex1/figs/VhdlStructure.pdf}
   \caption{VHDL structure of generated example}
@@
 \item Install llvm-2.8, clang-2.8 from www.llvm.org (from source, builds on ubuntu 12, centos 6.5, may need  to do minor edits in includes in ubuntu 14).
-\item Install boost development packages.
+\item Install boost development packages, GNU pth, pdflatex (packages exist for ubuntu/centos).
 \item Install antlr-2.7.7 (from source, builds on ubuntu 12, centos 6.5, some includes need to be added in ubuntu 14).

### 2018-03-07 v2/CtestBench/docs/CtestBench.tex
Updated the tutorial etc.

@@
 The difference is that instead of using the pipeHandler,
-the test-bench now uses methods in SocketLib.  Further, the
+the test-bench now uses methods in SocketLibPipeHandler.  Further, the
 VHDL is executed in a VHDL simulator; the simulator communicates
@@
 # build software version of testbench (to check the "desired behaviour")
-SOCKETLIB_INCLUDE=$(AHIR_RELEASE)/CtestBench/include
-SOCKETLIB_LIB=$(AHIR_RELEASE)/CtestBench/lib
-PIPEHANDLER_INCLUDE=$(AHIR_RELEASE)/pipeHandler/include
-PIPEHANDLER_LIB=$(AHIR_RELEASE)/pipeHandler/lib
-PTHREADUTILS_INCLUDE=$(AHIR_RELEASE)/pthreadUtils/include
+AHIR_INCLUDE=$(AHIR_RELEASE)/include
+AHIR_LIB=$(AHIR_RELEASE)/lib
 VHDL_LIB=$(AHIR_RELEASE)/vhdl
-VHDL_VHPI_LIB=$(AHIR_RELEASE)/CtestBench/vhdl
+VHDL_VHPI_LIB=$(AHIR_RELEASE)/vhdl
 FUNCTIONLIB=$(AHIR_RELEASE)/functionLibrary/
@@
 SW: $(SRC)/prog.c $(SRC)/prog.h $(SRC)/testbench.c 
-	gcc -g -c -DSW $(PROGDEFS) -I$(PIPEHANDLER_INCLUDE)\
+	gcc -g -c -DSW $(PROGDEFS) -I$(AHIR_INCLUDE)\
               -I$(FUNCTIONLIB)/include -I$(SRC) $(SRC)/prog.c
-	gcc -g -c -DSW $(PROGDEFS) -I$(PIPEHANDLER_INCLUDE)\
-              -I$(PTHREADUTILS_INCLUDE) -I$(SRC) $(SRC)/testbench.c
+	gcc -g -c -DSW $(PROGDEFS) -I$(AHIR_INCLUDE)\
+               -I$(SRC) $(SRC)/testbench.c
 	gcc -g -o testbench_sw prog.o testbench.o\
-              -L$(PIPEHANDLER_LIB) -lPipeHandler -lpthread
+              -L$(AHIR_LIB) -lPipeHandler -lpthread
 
@@
 c2llvmbc: $(SRC)/prog.c $(SRC)/prog.h
-	clang -O3 -std=gnu89 $(PROGDEFS)  -I$(SOCKETLIB_INCLUDE)\
+	clang -O3 -std=gnu89 $(PROGDEFS)  -I$(AHIR_INCLUDE)\
               -I$(FUNCTIONLIB)/include -emit-llvm -c $(SRC)/prog.c
@@
 	gcc -c vhdlCStubs.c  -I$(SRC)\
-                -I./ -I$(SOCKETLIB_INCLUDE)
+                -I./ -I$(AHIR_INCLUDE)
 	gcc -c $(SRC)/testbench.c\
-                -I$(PTHREADUTILS_INCLUDE) -I$(SRC)\
-                -I./ -I$(SOCKETLIB_INCLUDE)
+                -I$(AHIR_INCLUDE) -I$(SRC)\
+                -I./ 
 	gcc -o testbench_hw testbench.o vhdlCStubs.o\
-                -L$(SOCKETLIB_LIB) -lSocketLib -lpthread
+                -L$(AHIR_LIB) -lSocketLibPipeHandler -lpthread
 
@@
 	ghdl -i --work=work ahir_system_test_bench.vhdl
-	ghdl -m --work=work -Wl,-L$(SOCKETLIB_LIB) -Wl,-lVhpi ahir_system_test_bench
+	ghdl -m --work=work -Wl,-L$(AHIR_LIB) -Wl,-lVhpi ahir_system_test_bench
 

### 2018-04-20 v2/Ahir-asm/devel/docs/AaLRM.tex
Added some dope on volatiles to the AaLRM.  Also signal read ack is always true.
Added $opaque module class to prevent spurious dependency chasing.

@@
 Section \ref{sec:Syntax}.
+
+An assignment statement can be marked as a volatile
+\begin{verbatim}
+$volatile a := (b+c)
+\end{verbatim}
+which means that $a$ will be generated from $b,c$, but
+the logic which updates $a$ can be combinational.
+
 \item A {\bf call} statement is of the form
@@
 
+A call statement must be marked as a volatile
+if the corresponding called module is volatile
+\begin{verbatim}
+$volatile $call foo (b c) (a)
+\end{verbatim}
+which means that $a$ will be generated from $b,c$, but
+the logic which updates $a$ can be combinational.
+
+\item The {\bf null} statement does nothing, and just passes the
+token onwards.  This is how it looks like
+\begin{verbatim}
+$null
+\end{verbatim}
+\end{itemize}
+
 Complete atomic statements can have a guard predicate
@@
 Section \ref{sec:Syntax}.
-
-\item The {\bf null} statement does nothing, and just passes the
-token onwards.  This is how it looks like
-\begin{verbatim}
-$null
-\end{verbatim}
-\end{itemize}
 
@@
 
+\subsection{Opaque Modules}
+
+The {\bf Aa} compiler normally looks inside module
+descriptions to identify load/store or pipe access
+dependencies.  If you mark a module as opaque, then
+the calling module will ignore dependencies inside
+this module while generating the control path.
+
+An example of an opaque module:
+\begin{verbatim}
+$opaque $module [foo] $in (a: $uint<9>)
+   $out (b: $uint<8>) $is
+{
+    // out_pipe is a pipe.
+    out_pipe := a
+
+    b := in_pipe
+}
+\end{verbatim}
+
 \subsection{Inlined Modules}

### 2018-04-20 v2/Ahir-asm/devel/include/AaRoot.h

@@ -190,6 +190,8 @@ class AaRoot
   virtual bool Is_Null_Like_Statement() {return(false); }
   virtual bool Is_Assignment_Statement() {return(false); }
   virtual bool Is_Call_Statement() {return(false); }
+  virtual bool Is_Opaque_Call_Statement() {return(false);}
+
 
   // do we really need this? keep it for now
   virtual bool Is(string class_name);

### 2018-10-31 v2/Ahir-asm/devel/docs/AaLRM.tex
AaLrm modified to reflect change in macro-module handling.

@@
 
-\section{Module classes: pipeline, volatile, operator, inline, macro, foreign }
+\section{Module classes: pipeline, volatile, operator,  macro, foreign }
 
@@
 
-\subsection{Inlined Modules}
-
-An inlined module is substituted into the
-point of a call to it in a three-step manner.
-The input arguments are sampled at the point
-of the call, the body of
-the module is invoked and the output arguments
-are transferred.  The semantics of an inlined
-function call are defined by this three step
-procedure.  The input arguments are sampled
-at the point of the call, and the output
-arguments are updated at the termination of
-the call.
-
-For example, the following
-program 
-\begin{verbatim}
-$inline $module [add] $in (a: $uint<32> b: $uint<32>)
-    $out (c: $uint<32>) $is
-{
-    c := (a+b)
-}
-
-$module [add_check] $in (p: $uint<32> q: $uint<32>) 
-    $out (r : $uint<32>) $is
-{
-    $call add (p q) (r)
-}
-\end{verbatim}
-is translated to
-\begin{verbatim}
-$module [add_check]
-$in ( p : $uint<32>   q : $uint<32>  )
-$out ( r : $uint<32>  )
-$is
-{
-  $seriesblock[add_18] 
-  {
-    $parallelblock[InArgs] 
-    {
-      add_18_a := p
-      add_18_b := q
-    }
-    ( add_18_a => add_18_a add_18_b => add_18_b )
-    $seriesblock[body] 
-    {
-      add_18_c := (add_18_a + add_18_b)
-    }
-    (  add_18_c => add_18_c )
-    $parallelblock[OutArgs] 
-    {
-      r := add_18_c
-    }
-    ( r => r )
-  }
-  ( r => r )
-}
-\end{verbatim}
-The three step substitution of add into add\_check
-is clearly seen.  
-There are no restrictions on the
-module that is being inlined.
-
 \subsection{Macro Modules}
@@
 {
-  $seriesblock[shift_14] 
-  {
     midpipe := inpipe
-  }
-  $seriesblock[shift_17] 
-  {
     outpipe := midpipe
 }
-}
 \end{verbatim}
 The two calls to shift are replaced by
-series blocks with input/output arguments
+the body of the shift module with input/output arguments
 replaced by the expressions passed to them
@@
 
-There are no restrictions on the
-module that is being inlined.
+Macro modules should consist of only assignment
+statements and call statements.  There should
+be no object declarations inside a macro module.
 

### 2018-11-25 v2/Ahir-asm/devel/docs/AaLRM.tex
Uncovered a beauty of a bug due to rollover of time stamps in the memory subsystem.  For now, increase the time stamp range to reduce the likelihood of any screw ups.  Keep it at least 10 for now..

@@
 {\bf Aa} is a programming language for the 
-description of algorithms.   The control-flow
+description of algorithms, with the ultimate aim
+being the conversion of these algorithms into 
+a hardware implementation.   The control-flow
 in an \Aa program corresponds to a petri net
@@
 
-\subsection{Operator Modules}
+\subsection{Deterministic Pipelined Modules}
+
+In certain situations, we are sure that all operations inside
+a module are deterministic, that is, we know that they take
+a fixed amount of time.  Examples of such operations are 
+addition, subtraction etc. which are mapped to unshared
+operators.  Examples of operations which are not deterministic
+are pipe accesses and accesses to shared memory.
+
+If we are sure that all operations are deterministic, then
+we can add a qualifier to the pipelined module.  For instance
+\begin{verbatim}
+$pipeline $depth 3 $deterministic $operator
+        $module [foo] $in (A B: $uint<8>) $out (C: $uint<8>) $is
+{
+   U := A
+   V := (U + B)
+   W := (V + A)
+   C  := (W + B)
+}
+\end{verbatim}
+This module will then eventually be implemented in VHDL
+using much simpler control logic.
+
+Pipelined modules have the following restrictions:
+\begin{itemize}
+\item Their internal structure is that of an
+operator module (Section \ref{sec:OperatorModules}).
+\item They can not access memory.
+\item They can not access pipes.
+\item The only write-after-read sequences permitted
+are of an accumulation type, for example
+\begin{verbatim}
+a := (a + 1)
+\end{verbatim}
+\item They can call other modules only if the called
+modules are either volatile (Section \ref{sec:VolatileModules}) or deterministic pipelines.
+\end{itemize}
+
+In many cases, the core computation is completely
+deterministic, and we can save considerable hardware
+by collecting it into a deterministic pipeline.
+
+\subsection{Operator Modules} \label{sec:OperatorModules}
 
@@
 
-\subsection{Volatile Modules}
+\subsection{Volatile Modules} \label{sec:VolatileModules}
 

### 2018-11-25 v2/libAhirV2/src/vcModule.cpp

@@ -56,6 +56,7 @@ vcModule::vcModule(vcSystem* sys, string module_name):vcRoot(module_name)
 	this->_pipeline_depth = 1;
 	this->_is_function_library_module = false;
 	this->_delay = 2;
+	this->_deterministic_longest_path = -1;
 }
 
 int vcModule::Get_Delay()
@@ -345,6 +346,9 @@ void vcModule::Print_VHDL(ostream& ofile)
 
 		if(this->Get_Pipeline_Deterministic_Flag())
 		{
+			if(this->Get_Deterministic_Longest_Path() < 0)
+				this->Calculate_And_Set_Deterministic_Longest_Path();
+
 			vcSystem::Print_VHDL_Inclusions(ofile);
 			this->Print_VHDL_Deterministic_Pipeline_Operator_Entity(ofile);
 			this->Print_VHDL_Deterministic_Pipeline_Operator_Architecture(ofile);

### 2018-12-12 v2/Ahir-asm/devel/docs/AaLRM.tex
Added $update note in synch.

@@
 e := (f + g) $mark SECOND $synch (FIRST)
-h := (i + j) $synch (FIRST SECOND)
+h := (i + j) $synch ($update FIRST SECOND)
 \end{verbatim}
@@
 third statement will be started only after the first
-two have started. This
+has finished and the second has started. This
 is necessary in some cases where dependency analysis

### 2019-03-05 v2/Ahir-asm/devel/include/AaExpression.h
Added function-call-expression.  Gives us a better way to plug in custom vhdl implementations.

@@ -315,6 +315,7 @@ class AaExpression: public AaRoot
 
 	virtual bool Is_Implicit_Variable_Reference() {return(false);}
 	virtual bool Is_Interface_Object_Reference() {return(false);}
+	virtual bool Is_Volatile_Function_Call() {return(false);}
 
 	virtual bool Is_Implicit_Object() {return(false);}
 
@@ -1759,6 +1760,119 @@ class AaTernaryExpression: public AaExpression
 	virtual bool Is_Trivial() {return(this->Get_Is_Intermediate());}
 	//virtual bool Is_Trivial() {return(false);}
 };
+
+class AaFunctionCallExpression: public AaExpression
+{
+	string _module_identifier;
+	AaModule* _called_module;
+
+	vector<AaExpression*> _arguments;
+	public:
+
+	AaFunctionCallExpression(AaScope* scope, string module_id,
+					vector<AaExpression*>& arguments);
+	~AaFunctionCallExpression();	
+
+	virtual void Print(ostream& ofile); 
+
+	AaExpression*      Get_Argument(int id) 
+	{
+		if((id >= 0) && (id < _arguments.size()))
+			return(_arguments[id]);
+		else
+			return(NULL);
+	}
+	virtual string Kind() {return("AaFunctionCallExpression");}
+	virtual void Map_Source_References(set<AaRoot*>& source_objects);
+	virtual void Mark_As_Visited(set<AaRoot*>& visited_elements)
+	{
+		visited_elements.insert(this);
+		for(int I = 0, fI = _arguments.size(); I < fI; I++)
+		{
+			this->_arguments[I]->Mark_As_Visited(visited_elements);
+		}
+	}
+
+	virtual void Set_Associated_Statement(AaStatement* stmt)
+	{
+		_associated_statement = stmt;
+		for(int I = 0, fI = _arguments.size(); I < fI; I++)
+		{
+			this->_arguments[I]->Set_Associated_Statement(stmt);
+		}
+	}
+	virtual void Set_Phi_Source_Index(int v)
+	{
+		_phi_source_index = v;
+		for(int I = 0, fI = _arguments.size(); I < fI; I++)
+		{
+			this->_arguments[I]->Set_Phi_Source_Index(v);
+		}
+	}
+	virtual void Set_Pipeline_Parent(AaStatement* dws)
+	{
+		_pipeline_parent = dws;
[…]

### 2019-03-05 v2/Ahir-asm/devel/src/AaProgram.cpp

@@ -683,11 +683,11 @@ bool AaProgram::Propagate_Types()
 	  AaType* itype = NULL;
 	  
 	  // two passes: first find the type and then set it
-	  for(set<AaRoot*>::iterator siter = type_eq_class_map[i].begin();
-	      siter != type_eq_class_map[i].end();
-	      siter++)
+	  for(set<AaRoot*>::iterator ssiter = type_eq_class_map[i].begin();
+	      ssiter != type_eq_class_map[i].end();
+	      ssiter++)
 	    {
-	      AaRoot* item = *siter;
+	      AaRoot* item = *ssiter;
 	      AaType* ntype = NULL;
 	      if(item->Is_Expression())
 		{

### 2019-03-05 v2/Ahir-asm/devel/src/AaStatementWriteVCOptimized.cpp

@@ -212,7 +212,11 @@ void AaAssignmentStatement::Write_VC_Control_Path_Optimized(bool pipeline_flag,
 		if(this->Get_Is_Volatile())
 			ofile << "// volatile! " << endl;
 
-		bool source_is_implicit = (_source->Is_Signal_Read() || _source->Is_Implicit_Variable_Reference());
+		bool source_is_implicit = 
+			(_source->Is_Signal_Read() ||
+			 _source->Is_Volatile_Function_Call() ||
+			 _source->Is_Implicit_Variable_Reference());
+
 		bool target_is_implicit = _target->Is_Implicit_Variable_Reference();
 
 		if(!this->Get_Is_Volatile() && source_is_implicit && target_is_implicit)
@@ -355,7 +359,11 @@ void AaAssignmentStatement::Write_VC_Links_Optimized(string hier_id, ostream& of
 				ofile);
 		this->_target->Write_VC_Links_As_Target_Optimized(hier_id,
 				ofile);
-		bool source_is_implicit = (_source->Is_Signal_Read() || _source->Is_Implicit_Variable_Reference());
+		bool source_is_implicit = 
+			(_source->Is_Signal_Read() ||
+			 _source->Is_Volatile_Function_Call() ||
+			 _source->Is_Implicit_Variable_Reference());
+
 		bool target_is_implicit = _target->Is_Implicit_Variable_Reference();
 
 		if(source_is_implicit && target_is_implicit)
@@ -1713,6 +1721,7 @@ void AaPhiStatement::Write_VC_Control_Path_Optimized(bool pipeline_flag,
 		{
 			AaExpression* sge = source_expr->Get_Guard_Expression();
 			if((sge != NULL) && !sge->Is_Constant() && (sge != source_expr) && 
+					!sge->Is_Volatile_Function_Call() && 
 					!sge->Is_Implicit_Variable_Reference() && 
 					!sge->Is_Signal_Read() && !sge->Is_Flow_Through())
 			{
@@ -1733,7 +1742,8 @@ void AaPhiStatement::Write_VC_Control_Path_Optimized(bool pipeline_flag,
 				sge->Mark_As_Visited(visited_elements);
 
 			bool src_has_no_dpe  = (source_expr->Is_Implicit_Variable_Reference() ||  source_expr->Is_Signal_Read());
-			if(src_has_no_dpe)
[…]

### 2019-03-05 v2/libAhirV2/src/vcModule.cpp

@@ -1690,3 +1690,13 @@ void vcModule::Print_Reduced_CP_As_Dot_File()
 	this->_control_path->Print_Reduced_Control_Path_As_Dot_File(cp_file);
 	cp_file.close();	
 }
+
+void vcModule::Print_DP_As_Dot_File()
+{
+	string dp_file_name = this->Get_VHDL_Id() + "_DP.dot";
+	ofstream dp_file;
+	dp_file.open(dp_file_name.c_str());
+	this->_data_path->Print_Data_Path_As_Dot_File(dp_file);
+	dp_file.close();	
+}
+

### 2019-04-13 v2/Ahir-asm/devel/util/AaOpt.cpp
Some major cleanups in handling WAR.  There seemed to be an unnecessary delay in the SCT -> UST dependency for WARs.
Also cleaned up the elaborate algorithm for Aa.  First parse
and then map is the approach to be used.  This was not
being followed completely.

Finally, do not delay WAR object refs.  Also, delete orphan
statements (unless qualified with $keep).  Checked on
some examples including AJIT CPU.

Issue remaining:  Combinational cycle delay introduced? Lets
check.

@@ -82,6 +82,7 @@ int main(int argc, char* argv[])
 
   // inline in outfile
   AaProgram::_print_inlined_functions_in_caller = true;
+  AaProgram::_do_not_print_orphans = true;
 
   AaProgram::_tool_name = "AaOpt";
   AaProgram::_buffering_bits_added_during_path_balancing = 0;

### 2019-04-20 v2/Ahir-asm/devel/docs/AaLRM.tex
Some major cleanups in handling WAR.  There seemed to be an unnecessary delay in the SCT -> UST dependency for WARs.
Also cleaned up the elaborate algorithm for Aa.  First parse
and then map is the approach to be used.  This was not
being followed completely.

Finally, do not delay WAR object refs.  Also, delete orphan
statements (unless qualified with $keep).  Checked on
some examples including AJIT CPU.

Issue remaining:  Combinational cycle delay introduced? Lets
check.
Added fifo mem with asynch read in queue implementation.

@@
 balance pipelines.
+
+The {\bf Aa} compiler will optimize away dangling
+assignment statements (whose targets are not
+used any where in the program). In order to prevent
+this from happening, you can use the \$keep keyword.
+\begin{verbatim}
+sacrifical_value := signal_from_somwhere $keep
+\end{verbatim}
+The \$keep keyword must be placed after the \$delay
+keyword.
 
@@
 \end{verbatim}
+
+\section{Synthetic instructions}
+
+For convenience, there are some synthetic operations which
+you can use to simplify coding an Aa program.  These
+are listed below
+\begin{itemize}
+\item The \$reduce operation:
+\begin{verbatim}
+($reduce |  a  b  c)   
+    (note: this is the same as  ((a+b)+c))
+\end{verbatim}
+Any associative operation may be used instead
+of the or (add, mul, xor, and).
+\item The \$split operation:  Suppose you want
+to split a 32 bit variable into two 8-bit variables
+and four 4-bit variables.  This can be done using
+\begin{verbatim}
+($split X 8 8 4 4 4 4) (a8 b8 c4 d4 e4 f4)
+   (note: a8 := ($slice X 31 24) etc..)
+\end{verbatim}
+\item The \$concat operation:  This is the
+reverse of a \$split.  
+\begin{verbatim}
+Y := ($concat a8 b8 c4 d4 e4 f4)
+  (note: Y := (a8 && (b8 && (c4 && d4 ....))
+\end{verbatim}
+\item The \$excmux operations:  A multiplexor
+whose selects are expected to be mutually exclusive.
+\begin{verbatim}
+Y := ($excmux s0 X0 s1 X1 s2 X2)
+\end{verbatim}
+This is equivalent to 
+\begin{verbatim}
+Y := ($reduce |
+           ($mux s0 X0 0)
+           ($mux s1 X1 0)
+           ($mux s2 X2 0))
+\end{verbatim}
+Note that if all of s0,s1,s2 are 0, then
+Y is assigned to 0.
+\item The \$prioritymux operations:  A multiplexor
+whose selects may not be mutually exclusive.
+\begin{verbatim}
+Y := ($prioritymux s0 X0 s1 X1 s2 X2
+                $default 1)
+\end{verbatim}
+This is equivalent to 
+\begin{verbatim}
+Y := ($mux s0 X0
+           ($mux s1 X1
+                $mux s2 X2 1))
+\end{verbatim}
+\end{itemize}
 

### 2019-04-29 v2/Ahir-asm/devel/docs/AaLRM.tex
Guards for function call expressions were not being printed.

@@
 Section \ref{sec:Syntax}.
+
+\subsection{A synthetic statement: the \$split statement}
+
+Often, we need to split a large vector into several pieces.
+Suppose you want
+to split a 32 bit variable into two 8-bit variables
+and four 4-bit variables.  This can be done using
+\begin{verbatim}
+$split (X 8 8 4 4 4 4) (a8 b8 c4 d4 e4 f4)
+\end{verbatim}
+This is actually translated to the following
+sequence of assignment statements.
+\begin{verbatim}
+a8 := ($slice X 31 24)
+a7 := ($slice X 23 16)
+c4 := ($slice X 15 12)
+...
+f4 := ($slice X 3 0)
+\end{verbatim}
 
@@
 of the or (add, mul, xor, and).
-\item The \$split operation:  Suppose you want
-to split a 32 bit variable into two 8-bit variables
-and four 4-bit variables.  This can be done using
-\begin{verbatim}
-($split X 8 8 4 4 4 4) (a8 b8 c4 d4 e4 f4)
-   (note: a8 := ($slice X 31 24) etc..)
-\end{verbatim}
 \item The \$concat operation:  This is the

### 2019-09-27 Documents/AhirV2/AhirV2ToolsReference.tex
The following case is not handled correctly:     a := (b_sig + c_sig) where b_sig and c_sig are signals.  The only thing that is allowed with signals is     p := b_sig The fix is easy, but am going to go for a fence. This is now flagged as an error.

--- 2018-03-07 Documents/AhirV2/AhirV2.tex
@@
 
-\title{AhirV2: from algorithms to hardware \\ An overview }
+\title{AhirV2: Tools which take algorithms to hardware \\ A reference manual }
 \author{Madhav Desai \\ Department of Electrical Engineering \\ Indian Institute of Technology \\
@@
 will usually lead to more efficient hardware.
-Typically, if we start from C, circuits
-produced by the AhirV2 flow are upto two orders of
-magnitude more energy-efficient than
-a processor \cite{ref:dsd2010}.
 
@@
 \end{itemize}
-
-\section{Use Scenario 1:  a transformational system}
-
-Many programs are transformational in nature.  There
-is some input data $X$, and the program can be modeled as
-a function $f$ which acts on $X$ to produce output data $Y$.
-
-\subsection{A trivial example} \label{sec:Example}
-
-Consider the following trivial example:
-\begin{verbatim}
-int add(int a, int b)
-{
-        int c = (a+b);
-        return(c);
-}
-\end{verbatim}
-We wish to generate a circuit which {\em implements}
-the specification implied by this program.
-
-We convert the program to LLVM byte code using
-the {\bf clang} compiler (www.llvm.org)
-\begin{verbatim}
-      clang -std=gnu89 -emit-llvm -c add.c
-\end{verbatim}
-This produces a binary file {\bf add.o} which is
-the LLVM byte-code.  To make the byte-code human
-readable, we dis-assemble it using an LLVM utility
-\begin{verbatim}
-     llvm-dis add.o
-\end{verbatim}
-This is what the LLVM assembly code looks like
-\begin{verbatim}
-; ModuleID = 'add.o'
-target datalayout = "e-p ...... "
-target triple = "i386-pc-linux-gnu"
-
-define i32 @add(i32 %a, i32 %b) nounwind {
-  %1 = alloca i32, align 4
-  %2 = alloca i32, align 4
-  %c = alloca i32, align 4
-  store i32 %a, i32* %1, align 4
-  store i32 %b, i32* %2, align 4
-  %3 = load i32* %1, align 4
-  %4 = load i32* %2, align 4
-  %5 = add nsw i32 %3, %4
-  store i32 %5, i32* %c, align 4
-  %6 = load i32* %c, align 4
-  ret i32 %6
-}
-\end{verbatim}
-To get to this point, we could have used several
-optimizations which are available in the LLVM frame-work.
-But we work with the unoptimized version to illustrate
-the storage decomposition which is carried out by
-the AhirV2 tools.
-
-The LLVM byte-code is our starting point.  We first convert it
-to \Aa.
-\begin{verbatim}
-llvm2aa add.o | vcFormat > add.o.aa
-\end{verbatim}
-This produces an \Aa program
-\begin{verbatim}
-// Aa code produced by llvm2aa (version 1.0)
-$module [add]
-// arguments
-$in (a : $uint<32> b : $uint<32> )
-$out (ret_val__ : $uint<32>)
-$is
-{
-  $storage stored_ret_val__ : $uint<32>
-  $branchblock [add]
-  {
-    //begin: basic-block bb_0
-    $storage iNsTr_0 : $uint<32>
-    $storage iNsTr_1 : $uint<32>
-    $storage c : $uint<32>
-    iNsTr_0 := a
-    iNsTr_1 := b
-    // load
-    iNsTr_4 := iNsTr_0
-    // load
-    iNsTr_5 := iNsTr_1
-    iNsTr_6 := (iNsTr_4 + iNsTr_5)
-    c := iNsTr_6
-    // load
-    iNsTr_8 := c
-    stored_ret_val__ := iNsTr_8
-    $place [return__]
-    $merge return__ $endmerge
-    ret_val__ := stored_ret_val__
-  }
-}
-\end{verbatim}
-
-Now, this \Aa code is converted to a virtual circuit \vC representation.
-\begin{verbatim}
-     Aa2VC -O add.o.aa | vcFormat > add.o.aa.vc
-\end{verbatim}
-The virtual circuit representation is a bit too verbose to reproduce entirely
-here, but we show some critical fragments
-\begin{verbatim}
-$module [add] 
-{
-  $in a:$int<32> b:$int<32>
-  $out ret_val__:$int<32>
-  $memoryspace [memory_space_0] 
-  {
-    $capacity 1
-    $datawidth 32
-    $addrwidth 1
-    // ret-val is kept here
-    $object [xxaddxxstored_ret_val__] : $int<32>
-  }
-  $memoryspace [memory_space_1] 
-  {
-    $capacity 1
-    $datawidth 32
-    $addrwidth 1
-    // a is kept here.
-    $object [xxaddxxaddxxiNsTr_0] : $int<32>
-  }
-  $memoryspace [memory_space_2] 
-  {
-    $capacity 1
-    $datawidth 32
-    $addrwidth 1
-    // b is kept her
-    $object [xxaddxxaddxxiNsTr_1] : $int<32>
-  }
-  $memoryspace [memory_space_3] 
-  {
-    $capacity 1
-    $datawidth 32
-    $addrwidth 1
-    // c is kept here.
-    $object [xxaddxxaddxxc] : $int<32>
-  }
-  $CP 
-  {
-     // a control-flow petri-net..  verbose..
-  }
-  // end control-path
-  $DP 
-  {
-     // wires and operators.
-  }
-
-   // links between CP and DP
-}
-\end{verbatim}
-
-Note that the stored objects a,b,c and
-ret\_val\_\_ are mapped to different memory spaces.  Thus, the
-chief difference between a \vC description and a processor is 
-that the \vC program partitions storage into small units which
-are accessed only by operators that need them.
-
-Finally, we take the \vC description and convert it to
-VHDL
-\begin{verbatim}
-vc2vhdl -t add -f add.o.aa.vc | vhdlFormat > system.vhdl
-\end{verbatim}
-This produces a VHDL implementation of the system with
-{\bf add} marked as a top-level module.  The VHDL that
-is produced is too voluminous to reproduce here, but
-the top-level system entity is
-\begin{verbatim}
-entity ahir_system is  -- system
-  port (--
-    add_a : in  std_logic_vector(31 downto 0);
-    add_b : in  std_logic_vector(31 downto 0);
-    add_ret_val_x_x : out  std_logic_vector(31 downto 0);
-    add_tag_in: in std_logic_vector(0 downto 0);
-    add_tag_out: out std_logic_vector(0 downto 0);
-    add_start_req : in std_logic;
-    add_start_ack : out std_logic;
-    add_fin_req   : in std_logic;
-    add_fin_ack   : out std_logic;
-    clk : in std_logic;
-    reset : in std_logic); --
-  --
-end entity;
-\end{verbatim}
-
-The AHIR system is a purely synchronous implementation
-which uses only the rising edge of the {\bf clk} input.
-The {\bf reset} is synchronous, and is active high.
-
-The VHDL implementation of a module in the AHIR system
-corresponds to that of a pipeline stage (see Figure \ref{fig:ModulePipeline}).
-\begin{figure}
-\begin{centering}
-\centerline{\psfig{figure=ModulePipeline.eps,width=3.0in,height=3.0in}}
-\caption{System and Module Interfaces}
- \label{fig:ModulePipeline}
-\end{centering}
-\end{figure}
-The ports 
-\begin{verbatim}
-add_a add_b
-\end{verbatim}
-correspond to the input arguments of the top-level 
-function {\bf add}.  The port 
-\begin{verbatim}
-ret_val_x_x
-\end{verbatim}
-corresponds to the value returned by {\bf add}.
-The pair of ports
-\begin{verbatim}
-add_start_req  add_start_ack
-\end{verbatim}
-implement a {\bf start} protocol; The environment
-asserts {add\_start\_req} when it wants to start
-{\bf add}, and the AHIR system asserts {\bf add\_start\_ack}
-whenever it is ready to start.  The environment is required
-to hold the input arguments steady until it observes
-the acknowledge from the AHIR system.
-The pair of ports
-\begin{verbatim}
-add_fin_req  add_fin_ack
-\end{verbatim}
-implement a {\bf finish} protocol; The environment
-asserts {add\_fin\_req} when it is in a position to accept
-the returned value from a previously started 
-{\bf add}, and the AHIR system asserts {\bf add\_fin\_ack}
-whenever it has a return-value avaiable.  The returned value
-is valid only when the acknowledge from the AHIR system is
-asserted. 
-The ports
-\begin{verbatim}
-    add_tag_in  add_tag_out
-\end{verbatim}
-provide a mechanism by which a tag can be presented
-by the environment to the AHIR system to identify
-a particular invocation of the {\bf add} function.  The
-width of the tag is chosen (by the AhirV2 tool-chain,
-specifically, the vc2vhdl tool) to be large enough
-that each active call to {\bf add} can be identified
-uniquely.
-
-\section{Use Scenario 2: pipelines}
-
-A pipeline is a collection of parallel processes which
-work together to accomplish a certain function (or to finish
-a job).   Pipelines are commonly used in software and
-in hardware systems.
-
-The concept of pipes or sockets provide a natural mechanism
-for communication between parallel processes in a software
-pipeline.  In \Aa (and \vC), pipes are first-in-first-out
-buffers which provide a corresponding communication construct
-in the generated hardware.
-
-\subsection{An example of a trivial pipeline}
-
-Consider the following pipeline which has two stages
-{\bf foo} and {\bf bar}:  
-\begin{verbatim}
-Environment --> foo --> bar --> Environment
-\end{verbatim}
-The stage {\bf foo} takes a 32-bit
-integer from the external world, complements it and passes it on to stage
-bar.  The stage {\bf bar} takes the 32-bit number from {\bf foo}, complements
-it and sends it to the external world (in effect, nothing useful
-is done, this is only an illustration).
-
-If we are writing this pipe-line as a program, we could implement
-two independent processes (or threads) and use named pipes to
-perform the communication between these processes/threads (and the
-``outside´´ world).  For example:
-\begin{verbatim}
-#include <iolib.h>
-void foo()
-{
-   while(1)
-   {
-     uint32_t data = read_uint32("inpipe");
-     write_uint32("midpipe", ~data);
-   }
-}
-
-void bar()
-{
-   while(1)
-   {
-     uint32_t data = read_uint32("midpipe");
-     write_uint32("outpipe", ~data);
-   }
-}
-\end{verbatim}
-In this example, the read/write functions are provided
-as part of a {\em pipeHandler} library which is bundled with
-the AhirV2 distribution.
-
-The two functions can be compiled into separate processes
-or can be used in threads in a multi-threaded program (using pthreads, for
-example), and one gets a software pipeline, in which
-the inter-process communication is done using the
-read/write function calls, which use named pipes.
-
-One can also map this pipeline to a VHDL system, using the
-following flow.  One uses the standard flow that has already
-been described:
-\begin{verbatim}
-clang -std=gnu89 -I../../../iolib/ -emit-llvm -c prog.c
-llvm2aa prog.o | vcFormat >  prog.o.aa
-Aa2VC -O -I mempool -C prog.o.aa | vcFormat > prog.o.aa.vc
-vc2vhdl -C -s ghdl -T foo -T bar -f prog.o.aa.vc\
-           | vhdlFormat > system.vhdl
-\end{verbatim}
-Note that in this case, we use the -T option in vc2vhdl to
-specify that foo and bar are free-running top-level modules.
-\begin{figure}
-\begin{centering}
-\centerline{\psfig{figure=FooBarPipeline.eps,width=3.0in,height=3.0in}}
-\caption{Hardware implementation of Foo-Bar pipeline}
- \label{fig:FooBarPipeline}
-\end{centering}
-\end{figure}
-The resulting AHIR system in system.vhdl implements the pipeline
-(the block diagram of the system is shown in Figure \ref{fig:FooBarPipeline})
-with the following interface:
-\begin{verbatim}
-entity ahir_system is  -- system 
-  port (-- 
-    clk : in std_logic;
-    reset : in std_logic;
-    inpipe_pipe_write_data: in std_logic_vector(31 downto 0);
-    inpipe_pipe_write_req : in std_logic_vector(0 downto 0);
-    inpipe_pipe_write_ack : out std_logic_vector(0 downto 0);
-    outpipe_pipe_read_data: out std_logic_vector(31 downto 0);
-    outpipe_pipe_read_req : in std_logic_vector(0 downto 0);
-    outpipe_pipe_read_ack : out std_logic_vector(0 downto 0)); -- 
-  -- 
-end entity;
-\end{verbatim}
-The system has interfaces corresponding to the pipes inpipe and
-outpipe through which data is exchanged.
-
-In practice, one can have any number of pipes and 
-interacting processes in a pipeline implementation.
-
-\section{Storage variables and memory spaces}
-
-In an \Aa program, variables can be of three kinds:
-storage variables, pipe variables, or single-static-assignment
-variables.  Storage variables are implemented in memory,
-pipe variables as FIFO buffers, and single-static-assignment
-variables as registers.
-
-While transforming an \Aa description down to a \vC description,
-the storage variables in the \Aa program need to be grouped
-into memory spaces.  Two storage variables are put in the 
-same memory space only if we
-determine that a pointer de-reference in the \Aa
-program can point into either of the two storage variables
-(this is determined by a conservative static analysis).
-
-A memory space in a \vC description is characterized by
-a word-length (the greatest common divisor of the widths
-of accesses to this memory space), an address-width (wide enough
-to allow access to all words in the memory space), and
-a capacity (the number of words in the memory space).  Typically,
-a program will have many small memory spaces corresponding to
-scratch storage and some large memory spaces which correspond
-to arrays etc.
-
-There is one small issue, however.
-Consider the following C program:
-\begin{verbatim}
-int main(int* b)
-{
-   int q[2];
-   q[0] = *b;
-   q[1] = q[0];
-   return(q[1]);
-}
-\end{verbatim}
-When this program is mapped to a circuit, we identify 
-two distinct memory spaces, one which contains the
-array $q$ and the other corresponding to the 
-external world (the one referred to by the pointer $b$).
-Where is the external memory physically located?  In 
-the AhirV2 flow, we can either locate it outside
-the system or inside the system which is being 
-described by this program.
-
-If the external memory is to be placed outside, then
-accesses to it from within the system must be routed
-outside the system.  On the other hand if it is
-to be placed inside, a storage object corresponding 
-to it must be created and all accesses to the external
-memory must be directed at this storage object.
-Further, the external world must have a mechanism for
-accessing this storage object.
-
-Both options are supported in the AhirV2 flow through
-{\bf AaLinkExtMem}.
-
-\subsection{Keeping the external memory outside the system}
-
-In this scenario, all memory accesses which are resolved
-to be to a storage object which is not declared in the
-\Aa program are redirected outside the system by using
-pipes.
-
-If you want to keep the external memory
-outside, you will have to go through the following sequence
-\begin{verbatim}
-# first use clang (or llvm-gcc) to generate llvm-byte-code
-clang -std=gnu89 -emit-llvm -c foo.c
-#
-# disassemble so that you can make sense of the llvm bc.
-llvm-dis foo.o
-#
-# OK, now take the llvm byte code
-# and generate an Aa description.
-# use the storageinit option to initialize
-# global storage.
-# (the pipe to vcFormat is to prettify the output)
- llvm2aa -storageinit foo.o | vcFormat > foo.o.aa
-#
-#
-# Do an Aa -> Aa transformation: map external
-# memory outside..
-AaLinkExtMem foo.o.aa | vcFormat > foo.o.memlinked.ExternalOutside.aa
-#
-# Now take the Aa code and generate a virtual
-# circuit..
-# the -O flag does dependency analysis in straight-line
-# code and parallelizes it.
-#
-Aa2VC -O foo.o.memlinked.ExternalOutside.aa | vcFormat\
-                 > foo.o.memlinked.ExternalOutside.aa.vc
-#
-# finally, generate vhdl from the vc description.  Note that
-# you will have to mark the module foo as well as the
-# extmem_store_32/load_32 modules as top-level modules
-# so that it is possible for the outside world to serve
-# requests made from inside.
-#
-vc2vhdl -O -t foo -t extmem_store_32 -t extmem_load_32\ 
-         -f foo.o.memlinked.ExternalOutside.aa.vc | vhdlFormat\
-              > foo_o_aa_memlinked_external_outside_vc.vhdl
-
-\end{verbatim}
-
-If you look at the generated top-level VHDL entity, its ports
-will be
-\begin{verbatim}
-entity ahir_system is  -- system
-  port (--
-    clk : in std_logic;
-    reset : in std_logic;
-
-    --   some-lines-omitted --
-    --   foo-related ports --
-    --   some-lines-omitted --
-
-    extmem_read_address_32_pipe_read_data: out std_logic_vector(31 downto 0);
-    extmem_read_address_32_pipe_read_req : in std_logic_vector(0 downto 0);
-    extmem_read_address_32_pipe_read_ack : out std_logic_vector(0 downto 0);
-    extmem_read_data_32_pipe_write_data: in std_logic_vector(31 downto 0);
-    extmem_read_data_32_pipe_write_req : in std_logic_vector(0 downto 0);
-    extmem_read_data_32_pipe_write_ack : out std_logic_vector(0 downto 0);
-    extmem_write_address_32_pipe_read_data: out std_logic_vector(31 downto 0);
-    extmem_write_address_32_pipe_read_req : in std_logic_vector(0 downto 0);
-    extmem_write_address_32_pipe_read_ack : out std_logic_vector(0 downto 0);
-    extmem_write_data_32_pipe_read_data: out std_logic_vector(31 downto 0);
-    extmem_write_data_32_pipe_read_req : in std_logic_vector(0 downto 0);
-    extmem_write_data_32_pipe_read_ack : out std_logic_vector(0 downto 0)); --
-  --
-end entity;
-\end{verbatim}
-The external memory read and write address and data are clearly
-visible.   The outside world is responsible for serving the
-read/write requests made from the inside.
-
-\subsection{Keeping the external memory inside the system}
-
-In this scenario, we will assume that all accesses to 
-storage variables not defined in the \Aa program are 
-to be directed to a storage variable which is to be
-a viewed as a shared memory pool that is visible to the
-\Aa program as well as to the outside world.  The visibility
-to the outside world is provided by access functions
-that the outside world can use to read/write from this
-shared memory pool.
-
-You will have to go through the following sequence:
-\begin{verbatim}
-# use clang (or llvm-gcc) to generate llvm-byte-code
-clang -std=gnu89 -emit-llvm -c foo.c
-#
-# disassemble so that you can make sense of the llvm bc.
-llvm-dis foo.o
-#
-# OK, now take the llvm byte code
-# and generate an Aa description.
-# use the storageinit option to initialize
-# global storage.
-# (the pipe to vcFormat is to prettify the output)
- llvm2aa -storageinit foo.o | vcFormat > foo.o.aa
-#
-#
-# Do an Aa -> Aa transformation: map external
-# memory to a storage area inside the system...
-# -I 1024 says that the amount of memory that will be
-# referred to is 1024 bytes.
-# -E mempool says that the storage object corresponding
-# to external memory is named mempool.
-AaLinkExtMem  -I 1024 -E mempool foo.o.aa | vcFormat\
-               > foo.o.memlinked.ExternalInside.aa
-#
-# Now take the Aa code and generate a virtual
-# circuit..
-# the -O flag does dependency analysis in straight-line
-# code and parallelizes it.
-# the -I mempool option says that external memory is
-# to be mapped inside the system to object mempool..
-#
-Aa2VC -O -I mempool foo.o.memlinked.ExternalInside.aa\
-           | vcFormat > foo.o.memlinked.ExternalInside.aa.vc
-#
-# finally, generate vhdl from the vc description.
-# note that you will have to mark mem_load__ and mem_store__
-# as top-level modules, so that the external world can
-# access its memory pool inside the system.
-#
-vc2vhdl -O -t foo -t mem_load__ -t mem_store__ \ 
-     -f foo.o.memlinked.ExternalInside.aa.vc\
-       | vhdlFormat > foo_o_aa_memlinked_external_inside_vc.vhdl
-\end{verbatim}
-
-In this example, we are saying that the shared memory pool
-variable is named mempool, and it is an array of 1024 {\bf bytes}.
-The generated top-level VHDL entity has the following
-ports:
-\begin{verbatim}
-entity ahir_system is  -- system
-  port (--
-    foo_b : in  std_logic_vector(31 downto 0);
-    foo_ret_val_x_x : out  std_logic_vector(31 downto 0);
-    foo_tag_in: in std_logic_vector(0 downto 0);
-    foo_tag_out: out std_logic_vector(0 downto 0);
-    foo_start : in std_logic;
-    foo_fin   : out std_logic;
-    mem_load_x_x_address : in  std_logic_vector(31 downto 0);
-    mem_load_x_x_data : out  std_logic_vector(7 downto 0);
-    mem_load_x_x_tag_in: in std_logic_vector(0 downto 0);
-    mem_load_x_x_tag_out: out std_logic_vector(0 downto 0);
-    mem_load_x_x_start_req : in std_logic;
-    mem_load_x_x_start_ack : out std_logic;
-    mem_load_x_x_fin_req   : in std_logic;
-    mem_load_x_x_fin_ack   : out std_logic;
-    mem_store_x_x_address : in  std_logic_vector(31 downto 0);
-    mem_store_x_x_data : in  std_logic_vector(7 downto 0);
-    mem_store_x_x_tag_in: in std_logic_vector(0 downto 0);
-    mem_store_x_x_tag_out: out std_logic_vector(0 downto 0);
-    mem_store_x_x_start_req : in std_logic;
-    mem_store_x_x_start_ack : out std_logic;
-    mem_store_x_x_fin_req   : in std_logic;
-    mem_store_x_x_fin_ack   : out std_logic;
-    clk : in std_logic;
-    reset : in std_logic); --
-  --
-end entity;
-\end{verbatim}
-
-The system provides  memory load and memory store
-function interfaces to the external world 
-(through  mem\_load.. and mem\_store..).
-The shared memory variable is guaranteed to have a base address of $0$.  Thus,
-byte mempool[I] will be present at address $I$.
 
@@
 realized).
+\item There are zillions of other LLVM optimizations that are
+available in llvm2aa (type llvm2aa --help to see these).  These
+involve program level optimizations such as loop unrolling,
+garbage collection optimizations etc. In general anything that
+increases instruction level parallelism will result in faster
+hardware, but with a cost.
 \end{itemize}
@@
 source \Aa files.
+\item {-C} (optional) : if specified, try to eliminate registers
+to the maximum extent possible by marking statements which 
+have a fanout of one as volatile.
 \item {-B} (optional) : if specified, add buffering to balance pipelined
@@
 \end{itemize}
+Note that the -B and -C options cannot be used together.
 
@@
 \begin{itemize}
+\item {\bf -h}: print help message and quit.
 \item {\bf -O} : if used, sequential statement blocks are parallelized
@@
 storage objects declared inside the \Aa program.
+\item {\bf -P} (optional):  Normally, the {\bf Aa} compiler resolves
+accesses to pipes and storage objects by analysing modules.  But if
+the -P flag is specified, all modules are treated as opaque and
+are assumed to not have any side-effects.  This can be dangerous.
 \end{itemize}
@@
 \begin{verbatim}
-vc2vhdl [-O] [-C] [-q] [-a] [-e <entity-name] [-w]\
-         -t/-T foo [-t/-T bar -t/-T bar2 ...]\
+vc2vhdl [-O] [-C] [-q] [-a] [-e <entity-name] [-w]
+         -t/-T foo [-t/-T bar -t/-T bar2 ...]
+         [-s ghdl/modelsim] [-L <file-name>]
+         [-v] [-D] [-W <work-lib>]
+         [-U] [-H]
          -f file1.vc -f file2.vc ...  > system.vhdl
@@
 interface to link with foreign code.
+\item {\bf -U}: input/output pipes will be printed with depth 0. 
+\item {\bf -H}: system interface in .hsys format will be printed.   This
+is used by another AHIR-V2 tool called hierSysBuild which can put together
+multiple AHIR systems using this hsys information.
 \end{itemize}
@@
 \begin{verbatim}
-Aa2C [-I <ext-mem-object>] <aa-file> (<aa-file>)*
+Aa2C [-I <ext-mem-object>] [-P <prefix>]  <aa-file> (<aa-file>)*
 \end{verbatim}
@@
 \begin{itemize}
-\item {\bf -I <ext-mem-object>}:  the same behaviour as in {\bf Aa2VC}.
+\item {\bf -I ext-mem-object}:  the same behaviour as in {\bf Aa2VC}.
+\item {\bf -P prefix}:  Specify a prefix string (default is the empty string)
+which will be used to name generated files (see below).
 \end{itemize}
@@
 \begin{itemize}
-\item aa\_c\_model.h :  a header file declaring functions in the generated
+\item [prefix]aa\_c\_model.h :  a header file declaring functions in the generated
 source code.
-\item aa\_c\_model.c : a source file containing function definitions
+\item [prefix]aa\_c\_model\_internal.h :  an internal header file with lots of macros.
+\item [prefix]aa\_c\_model.c : a source file containing function definitions
 corresponding to the {\bf Aa} modules.
@@
 
-\begin{thebibliography}{99}
-\bibitem{ref:dsd2010}
-Sameer D. Sahasrabuddhe, Sreenivas Subramanian, Kunal P. Ghosh, Kavi Arya, Madhav P. Desai, 
-"A C-to-RTL Flow as an Energy Efficient Alternative to Embedded Processors in Digital Systems," 
-DSD, pp.147-154, 2010 13th Euromicro Conference on Digital 
-System Design: Architectures, Methods and Tools, 2010
-\end{thebibliography}
 \end{document}

### 2020-07-26 v2/Ahir-asm/devel/docs/AaLRM.tex

@@
 Section \ref{sec:Syntax}.
+
+\subsection{Reporting and trace statements}
+
+It is possible to convert an {\bf Aa} description
+to C and then execute this C description.  During
+the execution, we can generate log messages using the
+following statement:
+\begin{verbatim}
+$rreport (Descriptor  Comment  var1 var2 .. varn)
+\end{verbatim}
+This generates a message during execution with
+the following format.
+\begin{verbatim}
+Descriptor    Comment
+         var1 <var1-value>
+         var2 <var1-value>
+         ....
+         varn <varn-value>
+\end{verbatim}
+
+Another useful statement is the assert statement.
+\begin{verbatim}
+$assert (expr)
+   $rreport (Descriptor  Comment  var1 var2 .. varn)
+\end{verbatim}
+This prints a report if expr is false, and then
+exits the execution.
+
+Finally, we have
+\begin{verbatim}
+$trace simple-identifier [ ( integer-level) ]
+\end{verbatim}
+This produces a message
+\begin{verbatim}
+TRACE simple-identifier
+\end{verbatim}
+it the trace-level (set in the execution of the AA2C
+model) is set to integer-level or lower.
+
+Note that these statements are not converted to hardware.
 

### 2020-10-28 Documents/AhirV2/AhirV2ToolsReference.tex
Fixed butt stupid bug in register_file_1w_1r_port.vhd. Also some issues related to constant phi statements. Updated docs a bit.

@@
 \begin{figure}
-\begin{centering}
-\centerline{\psfig{figure=AHIRv2Flow.eps,width=3.0in,height=3.0in}}
+\centering
+\includegraphics[width=8cm]{AHIRv2Flow.pdf}
 \caption{AhirV2 flow}
-\end{centering}
  \label{fig:AhirFlow}
@@
 
+\subsection{{\bf AaPreprocess}: Preprocessor for {\bf Aa} description}
+
+The {\bf AaPreprocess} utility takes an input \Aa file and applies pre-processing
+directives in the file  to produce an output \Aa file.   The invocation of the
+{\bf AaPreprocess} utility is as follows:
+\begin{verbatim}
+AaPreprocess [-I <search-directory>]* -o <output-aa-file> <list-of-input-aa-files>
+\end{verbatim}
+
+The utility then applies the pre-processing directives in the input \Aa files
+to produce an output \Aa file.
+
+The pre-process directives that are supported are
+\begin{description}
+\item[\#include]  Specify a file to be included.  This is typically specified
+as 
+\begin{verbatim}
+#include foo.aa
+\end{verbatim}
+The directories specified by the -I option are searched and the first
+matching file with name ``foo.aa'' is substituted in place at the point
+of the include.
+\item[\#define]  Specifies a key-word definition.  For example
+\begin{verbatim}
+#define P  const_56
+\end{verbatim}
+This defines the  string "P" to be an alias of the string "const\_56".
+A \#define can override another \#define.  The last definition is used.
+\item[\#undefine]  This removes the specified define from the list
+of preprocessor defines.
+\begin{verbatim}
+#undefine P
+\end{verbatim}
+Any further references to ``P'' in the source will be
+flagged as errors by the pre-preocessor.
+\item[\#\#]  This the paste directive.   For example, if we have
+defined
+\begin{verbatim}
+#define P  const_56
+\end{verbatim}
+and we have a line in the input file of the form
+\begin{verbatim}
+a := ( ##P + 1)
+\end{verbatim}
+then, this will be expanded to
+\begin{verbatim}
+a := (const_56 + 1)
+\end{verbatim}
+\item [\#if - \#endif] 
+Code can be conditionally included using this construct.
+\begin{verbatim}
+#define TFLAG 1
+#define SFLAG 0
+#if TFLAG
+a := b
+#endif 
+#if SFLAG
+b := a
+#endif
+\end{verbatim}
+will be expanded to
+\begin{verbatim}
+a := b
+\end{verbatim}
+\end{description}
+
+To summarize: the preprocessor can be used to 
+\begin{itemize}
+\item include a file into the code.
+\item paste strings into the code.
+\item conditionally include code.
+\end{itemize}
+This adds flexibility to the programmer and
+can reduce the amount of repetitive code.
+
 \subsubsection{Restrictions in using {\bf Aa2C}}

### 2020-11-21 v2/libAhirV2/docs/vC_LRM.tex
Bug fix in single-source phi (removed special casing) and in llvm2aa block chain guard handling.

@@
 The control-path is a Petri-net constructed using
-a set of rules (as describe in \cite{ref:SameerPhD}).1
+a set of rules (as described in \cite{ref:SameerPhD}).
 The data-path is essentially a directed graph of {\em operator } nodes
@@
 to implicit entry and exit transitions associated with the branch block.
-The $<-|$ specifies that a token leaving the right will enter
-the left (place) and $|->$ signifies that a token entering the left
+The 
+\begin{verbatim}
+<-|
+\end{verbatim}
+ specifies that a token leaving the right will enter
+the left (place) and 
+\begin{verbatim}
+|->
+\end{verbatim}
+signifies that a token entering the left
 (place) will enter one of the elements in the list on the right.
@@
 \end{verbatim}
+
 In this example, the keywords \$entry and \$exit correspond
 to implicit transitions associated with the branch block.
-The <-\& specifies that after a token leaves from every
+The
+\begin{verbatim}
+ <-&
+\end{verbatim}
+ specifies that after a token leaves from every
 element in the right list, a token will enter
-the left (transition) and \&-> signifies that a token leaving the left
+the left (transition) and the
+\begin{verbatim}
+&->
+\end{verbatim}
+signifies that a token leaving the left
 (transition) will be replicated so that a
@@
 Which says that jT is a join of s1 and s2, with a pre-condition
-that s1 is assumed to have fired twice and s2 is assumed to have fired
-twice before time began.  As a result it is possible for jT to fire
-twice before s1 and s2 fire in normal operation. 
+that s1 is assumed to have fired and s2 is assumed to have fired
+before time began.  As a result it is possible for jT to fire
+before s1 and s2 fire in normal operation. 
 \end{itemize}
@@
 \end{verbatim}
+
+\subsection{Special constructs in the control path}
+
+Some special constructs are used in the control path.
+\begin{itemize}
+\item Transition merge:  This is specified as
+\begin{verbatim}
+$transition_merge [merge_name] (p q r) (s)
+\end{verbatim}
+Where p, q, r are transition, and s fires whenever any 
+of p,q,r fire.
+\item  Phi-sequencer:   This is specified
+as
+\begin{verbatim}
+$phisequencer [phiseq_name] :  
+    trigger_1
+       src_1_sample_req src_1_sample_ack
+       src_1_update_req src_1_update_ack
+    trigger_2
+       src_2_sample_req src_2_sample_ack
+       src_2_update_req src_2_update_ack:
+    phi_sample_start phi_sample_completed
+       phi_update_start phi_update_completed :
+    trigger_1_sample_req  trigger_2_sample_req:
+       phi_mux_ack
+\end{verbatim}
+and it describes the control connections between the Phi operator and
+its sources.
+\item Loop terminator: This is specified as
+\begin{verbatim}
+$terminate (loop_exit loop_taken loop_body_done) 
+              (loop_back loop_terminate)
+\end{verbatim}
+The loop terminator looks at the result of the do-while condition
+check via the input transitions loop\_exit and loop\_taken, 
+and uses the loop\_body\_done transition to determine the
+number of transitions in flight and hence to trigger the
+next do-while iteration (firing of loop\_back) or to
+decide that the loop has terminated.
+\end{itemize}
 

### 2020-11-21 v2/libAhirV2/src/vcOperator.cpp

@@ -143,7 +143,7 @@ void vcEquivalence::Print_Flow_Through_VHDL(bool level_flag, ostream& ofile)
 		int L = (H- w->Get_Size())+1;
 		ofile << Z
 			<< " <= ov(" << H << " downto " << L << ");" << endl;
-		H = H-L;
+		H = H - w->Get_Size();
     	}
 	ofile << "--}" << endl;
 	ofile << "end process;" << endl;

### 2021-10-01 v2/Ahir-asm/devel/docs/AaExamples.tex
AA Examples.

\documentclass[12pt]{article}

\title{Some examples of the use of Aa relative to C}
\author{Madhav Desai \\ Department of Electrical Engineering \\ Indian Institute of Technology \\
	Mumbai 400076 India}

\newcommand{\Aa}{{\bf Aa}~}
\newcommand{\vC}{{\bf vC}~}

\begin{document}
\maketitle

We describe some examples of functional descriptions in {\bf C} and their equivalents in \Aa.

\section{Finding the maximum of $32$ numbers}

Consider a system which reads in 32 numbers from an input FIFO and stores them into an array.
Then the system finds the smallest item in the array, as well as the index
of the smallest item.  These two values are written out to an output FIFO.

In C the program could be written as follows.
\begin{verbatim}
void minItem() {
    uint32_t items[32];
    uint32_t I;

    // read loop
    for(I = 0;  I < 32; I++)
       items[I] = read_uint32("input_pipe");

    // min item loop
    uint32_t smallest = 0xffffffff;
    uint32_t smallest_index = 33;
    for(I = 0;  I < 32; I++)
    {
       if(items[I] <= smallest)
       {
           smallest = items[I];
           smallest_index = I;
       }
    }

    // write out.
    write_uint32("out_pipe", smallest);
    write_uint32("out_pipe", smallest_index);
}
\end{verbatim}

The almost exact equivalent of this C program in \Aa could
be written as follows.

\begin{verbatim}
$pipe input_pipe output_pipe: $uint<32>
$module [minItem] $in () $out () $is
{
   $storage items: $array[32] $of $uint<32>

   // read loop
   $branchblock[readLoop] {
       $merge $entry loopback
          $phi I := $zero<8> $on $entry nI $on loopback
       $endmerge

       $volatile nI := (I + 1)
       items[I] := input_pipe

       $if (I < 31) $then
           $place[loopback]
       $endif

       // min item loop
       $merge $entry min_item_loopback
          $phi J := $zero<8> $on $entry 
                                    nJ $on min_loopback
          $phi min_val := (~$zero<32>) $on $entry 
                                    n_min_val $on min_loopback
          $phi min_index := ($bitcast ($uint<32>) 33) $on $entry 
                                    n_min_index $on min_loopback
       $endmerge
       $volatile nJ := (J + 1)
       X := items[J]
       $if (X <= min_val) $then
           n_min_val := X
           n_min_index := J
       $endif

       $if (J < 31) $then
           $place[loopback]
       $endif 
   }  (n_min_val => MINVAL n_min_index => MININDEX)

   out_pipe := MINVAL
   out_pipe := MININDEX
}
\end{verbatim}

\section{Using double loops: matrix multiplication}

We read in a square matrix $A$ and calculate the 
matrix $A^2$.  The C program for this could be as follows.
\begin{verbatim}
void matrixSquarer() {
   uint32_t I, J, K;
   uint32_t A[32][32];
   uint32_t ASQ[32][32];

   // read matrix
   for (I = 0; I < 32; I++) {
      for (J = 0; J < 32; J++) {
        A[I][J] = read_uint32("input_pipe");
      }
   }
   // matrix multiply
   for (I = 0; I < 32; I++) {
      for (J = 0; J < 32; J++) {
         uint32_t tmp = 0;
         for (K = 0; J < 32; J++) {
            tmp += A[I][K]*A[K][J];
         }
         ASQ[I][J] = tmp;
      }
   }
   // send out.
   for (I = 0; I < 32; I++) {
      for (J = 0; J < 32; J++) {
        write_uint32("output_pipe", A[I][J]);
      }
   }
}
\end{verbatim}

The equivalent \Aa program could be as follows.
\begin{verbatim}
$module [matrixSquarer] $in () $out () $is {
   $storage A B: $array[32][32] $of $uint<32>

   // read loop
   $branchblock[readLoop] {

       ////////////////////   read loop /////////////////////
       $merge $entry outer_loopback
          $phi I := $zero<8> $on $entry nI $on outer_loopback
       $endmerge

       $merge $entry inner_loopback
          $phi J := $zero<8> $on $entry nJ $on inner_loopback
       $endmerge

       $volatile nI := (I + 1)
       $volatile nJ := (J + 1)
       A[I][J] := input_pipe

       $if (J < 31) $then
           $place[inner_loopback]
       $endif 
       $if (I < 31) $then
           $place [outer_loopback]
       $endif

       ////////////////////   matrix-multiply loop /////////////////////
       $merge $entry mm_outer_loopback
          $phi II := $zero<8> $on $entry nII $on mm_outer_loopback
       $endmerge

       $merge $entry mm_inner_loopback
          $phi JJ := $zero<8> $on $entry nJJ $on mm_inner_loopback
       $endmerge

       $volatile nII := (II + 1)
       $volatile nJJ := (JJ + 1)

       $merge $entry mm_dotp_loopback
          $phi K := $zero<8> $on $enry nK $on mm_dotp_loopback
          $phi TMP := $zero<32> $on $entry nTMP $on mm_dotp_loopback
       $endmerge

       nTMP := (TMP + (A[II][K] * A[K][JJ])

       $if (K < 31) $then
           $place [mm_dotp_loopback]
       $else
           B[II][JJ] := nTMP
       $endif

       $if (JJ < 31) $then
          $place [mm_inner_loopback]
       $endif

       $if (II < 31) $then
          $place [mm_outer_loopback]
       $endif_

       ////////////////////   send loop /////////////////////
       $merge $entry w_outer_loopback
          $phi III := $zero<8> $on $entry nIII $on w_outer_loopback
       $endmerge

       $merge $entry w_inner_loopback
          $phi JJJ := $zero<8> $on $entry nJJJ $on w_inner_loopback
       $endmerge

       $volatile nIII := (III + 1)
       $volatile nJJJ := (JJJ + 1)
       output_pipe := B[III][JJJ]

       $if (JJJ < 31) $then
           $place[w_inner_loopback]
       $endif 
       $if (III < 31) $then
           $place [w_outer_loopback]
       $endif

   } 
}
\end{verbatim}

\section{Using wide SSA variables to implement a shift regiser}

In this system, we read in 32-bit values from an input pipe and pass them through
an 8 stage shift register.   Outgoing values are written into
an output pipe.  Instead of using storage variables, we use wide
SSA variables to maintain the state of the shift register.

\begin{verbatim}
$pipe input_pipe output_pipe: $uint<32>
$module [shiftReg8] $in () $out () $is
{
   $branchblock[loop] {
        $merge $entry loopback
           $phi SR := $zero<256> $on $entry nSR $on loopback
           $phi V  := $zero<8>   $on $entry nV $on loopback
        $endmerge
        X := input_pipe

        // shift in from the right.
        nSR := ($concat ($slice SR 223 0) X)
        nV  := ($concat ($slice V 7 1) $one<1>)

        $if (V [] 7) $then
            output_pipe := ($slice SR 255 224)
        $endif
   }
}
\end{verbatim}

\end{document}

### 2021-10-24 v2/Ahir-asm/devel/docs/AaExamples.tex

@@
        $if (J < 31) $then
-           $place[loopback]
+           $place[min_loopback]
        $endif 
@@
       for (J = 0; J < 32; J++) {
-        write_uint32("output_pipe", A[I][J]);
+        write_uint32("output_pipe", ASQ[I][J]);
       }

### 2022-07-10 v2/Ahir-asm/devel/docs/AaLRM.tex
RtlThread generated VHDL sometimes infers a spurious latch. Removed.

@@
 like a variable.  That is, it remembers the last value
-that was written to it\footnote{    
-Note that a read from a signal will block until the
-signal has been written into at least once!}.
+that was written to it\footnote{Note that the signal is initialized to 0.}.
 \end{itemize}
@@
 \item Ternary expressions.
+\item Zero, One expressions.
 \end{itemize}
@@
 ($mux a (b+1) (c+d))
+\end{verbatim}
+
+\subsection{Zero, One expressions}
+
+For convenience, the following expressions represent
+a zero and one value with the specified uint type:
+\begin{verbatim}
+$zero<width>
+$one<width>
+\end{verbatim}
+
+The following expression 
+\begin{verbatim}
+$zero<16>
+\end{verbatim}
+expands to
+\begin{verbatim}
+($bitcast ($uint<16>) 0)
+\end{verbatim}
+and the  following expression
+\begin{verbatim}
+$one<16>
+\end{verbatim}
+expands to
+\begin{verbatim}
+($bitcast ($uint<16>) 1)
 \end{verbatim}

### 2023-05-29 v2/HierSysBuild/docs/tutorial/HierSysTools.tex
Added a HierSysBuild tutorial, long overdue.

\documentclass{beamer}
\usepackage{graphics}
\title{Hierarchical System Builder Toolset}
\author{Madhav P. Desai\\ Department of Electrical Engg.\\ IIT Bombay, Mumbai India}
\begin{document}
\maketitle

\frame[containsverbatim]{\frametitle{Overview}

\begin{itemize}
\item The hierarchical system builder toolset is intended for 
assembly of a system in terms of sub-systems.
\item The interfaces of the system and sub-systems are either
{\em pipes} or {\em signals}.   Pipes can be blocking or non-blocking.
\item It is possible to embed register transfer level fragments
in the system description in order to model glue logic.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{Hierachical system description format}

A system description takes the following form:

\begin{verbatim}
$system system_name $library system_library
   $in 
       ... input pipes and signals ...
   $out 
       ... output pipes and signals ...
{
   ... (optional) internals of the system ....
}
\end{verbatim}

The description is kept in a file with extension ``.hsys''.   For example
\begin{verbatim}
system_name.hsys
\end{verbatim}

}

\frame[containsverbatim]{\frametitle{A simple system}

\begin{verbatim}
// comments start with //
$system sysmem $library sysmem_library
   $in 
       // input pipe bus_request, width 32, depth 2
       $pipe bus_request 32 $depth 2
   $out 
       // output pipe bus_response, width 32, depth 2
       $pipe bus_response 32 $depth 2
{
   // empty..
}
\end{verbatim}
This describes a leaf level system which could
have been implemented by translation from Aa to vhdl
by the AHIR V2 toolset.  It's VHDL description  
is supposed to be compiled into the library sysmem\_library.

}

\frame[containsverbatim]{\frametitle{Hierarchy}

A 4-stage shift register constructed out of two simpler
2-stage shift registers.
\begin{verbatim}
$system shift_reg_4 $library shiftreg_library
   $in 
       $pipe in_data 32 $depth 2
   $out 
       $pipe out_data 32 $depth 2
{
   // internal pipe description..
   $pipe mid_data 32 depth 2
   $instance sr_left shiftreg_library :  shift_reg_2
                in_data => in_data
                out_data => mid_data
   $instance sr_right shiftreg_library :  shift_reg_2
                in_data => mid_data
                out_data => out_data
}
\end{verbatim}
}

\frame[containsverbatim]{\frametitle{Hierarchy: continued}
shift\_register\_2 may itself be build out of 1-stage
shift registers.
\begin{verbatim}
$system shift_reg_2 $library shiftreg_library
   $in 
       $pipe in_data 32 $depth 2
   $out 
       $pipe out_data 32 $depth 2
{
   // internal pipe description..
   $pipe mid_data 32 depth 2
   $instance sr_left shiftreg_library :  shift_reg_1
                in_data => in_data
                out_data => mid_data
   $instance sr_right shiftreg_library :  shift_reg_1
                in_data => mid_data
                out_data => out_data
}
\end{verbatim}
}

\frame[containsverbatim]{\frametitle{Hierarchy: reached the leaf}
\begin{verbatim}
$pipe in_data out_data: $uint<32> $depth 2

$module [shift_reg_1_daemon] $in () $out () $is
{
  $branchblock[loop] {
    $merge $entry loopback $endmerge
    X := in_data
    out_data := X

    $place [loopback]
  }
}
\end{verbatim}

}

\frame[containsverbatim]{\frametitle{Hierarchy: build organization and methodology. See examples/shift\_reg\_4.}
Organize the system into directories which reflect the 
hierarchy.
\begin{verbatim}
Top directory 
    shift_reg_4/

Contains
        shift_reg_2/

which contains
           shift_reg_1/
\end{verbatim}
}

\frame[containsverbatim]{\frametitle{Hierarchy: build organization and methodology continued}
At the leaf directory shift\_reg\_1, we may have an {\bf Aa} project which we
used to generate VHDL  and an aa2c model etc.  At non-leaf directories, we describe
the hierarchy at that level in an hsys file.
\begin{verbatim}
shift_reg_4/
    contains hsys file
    shift_reg_2/
      contains hsys file
      shift_reg_1/
         contains Aa project  
\end{verbatim}
}

\frame[containsverbatim]{\frametitle{Hierarchy: build scripts and make-files at each level}

\begin{verbatim}

shift_reg_4/
    build.sh script generates aa2c of entire hierarchy
    and VHDL at this level.
    shift_reg_2/
       Makefile generates VHDL model at this level
      shift_reg_1/
          Makefile generates aa2c and VHDL of leaf
          shift_reg_1

\end{verbatim}
}

\frame[containsverbatim]{\frametitle{Testbench: See examples/shift\_reg\_4/tb}

\begin{itemize}
\item See tb/tb.c, and compile.sh
\item Use testbench tb\_aa2c to validate the aa2c model.
\item Use testbench tb\_ghdl to validate the VHDL model.
\end{itemize}

}

\frame[containsverbatim]{\frametitle{Glue logic insertion using RTL}

\begin{itemize}
\item In addition to structural descriptions, we can use the
hsys file to describe glue logic using an RTL description.
\item For example, suppose we wish to process the result coming out
of the left shift\_register\_2 by complementing it.
\item We will insert a state machine between the left and
right instances of shift\_register\_2 in the top level
shift\_register\_4.
\item Note: this mechanism cannot be verified by C simulations,
but only by VHDL simulations.
\end{itemize}

}

\frame[containsverbatim]{\frametitle{RTL State Machine}
\begin{verbatim}
$thread complementerFsm
   $in $pipe input_data: $unsigned<32>
   $out $pipe ouput_data: $unsigned<32>
   $variable  data_reg: $unsigned<32>
   $constant z1  : $unsigned<1> := ($unsigned<1>) 0
   $constant o1  : $unsigned<1> := ($unsigned<1>) 1
$default
   //... continued... 
\end{verbatim}
}
\frame[containsverbatim]{\frametitle{RTL State Machine: continued}
\begin{verbatim}
   $now input_data$req  := o1 // default value of pipe read req
   $now output_data$req := z1 // default valu of pipe write req
   <st_empty> {
          $if input_data$ack { // input pipe has data
              data_reg := (~ input_data) // register!
              $goto st_full
          }  
   }
   <st_full> {
          $now output_data$req := o1 // request write to pipe
          $now output_data := data_reg // data to pipe
          $if output_data$ack { // pipe accepts..
                $goto st_empty
          }
          $else { $goto st_full }
   }
   $now $tick
\end{verbatim}
}

\frame[containsverbatim]{\frametitle{RTL State Machine: instantiating}
\begin{verbatim}
$system shift_reg_4 $library shiftreg_library
   $in 
       $pipe in_data 32 $depth 2
   $out 
       $pipe out_data 32 $depth 2
{
   $pipe mid_data_left mid_data_right 32 depth 2
   $instance sr_left shiftreg_library :  shift_reg_2
                in_data => in_data
                out_data => mid_data_left
   $instance sr_right shiftreg_library :  shift_reg_2
                in_data => mid_data_right
                out_data => out_data

   // thread description of complementFsm not shown...

   $string complement_inst: complementFsm 
      input_data => mid_data_left output_data => mid_data_right
}
\end{verbatim}

}

\frame[containsverbatim]{\frametitle{Using RTL State Machine: example shift\_reg\_4\_rtl}
\begin{itemize}
\item Essentially the same as shift\_reg\_4 example, but without aa2c simulation setup.
\item Examine the build and make files to see the differences.
\item Note: only VHDL sim verificationis supported for now.
\end{itemize}
}

\frame[containsverbatim]{\frametitle{Summary}

\begin{itemize}
\item Convenient way to assemble large systems which use pipes and signals
to communicate.
\item Build script: buildHierarchicalModel.py can generate aa2c and vhdl models.
\item Lots of error checking: dangling connections, width mismatch, direction mismatch.
\item Can embed glue RTL in description files 
\begin{itemize}
\item If you use embedded RTL, then only vhdl sims are possible for verification.
\end{itemize}
\item Tried and proven tools in AJIT, NAVIC project!
\end{itemize}

}
\end{document}

### 2023-05-29 v2/HierSysBuild/docs/tutorial/examples/shift_reg_4/tb/compile.sh

@@ -1,0 +1,6 @@
+gcc -g  -c -DAA2C -I ./ -I ../aa2c -I $AHIR_RELEASE/include tb.c
+gcc -g tb.o -o tb_aa2c -L $AHIR_RELEASE/lib -L ../lib -lshift_register_lib -lBitVectors -lSockPipes -lPipeHandlerDebugPthreads    -lpthread
+rm *.o 
+gcc -g -c  -I ./ -I ../aa2c -I $AHIR_RELEASE/include tb.c
+gcc -g  tb.o -o tb_ghdl -L $AHIR_RELEASE/lib -L ../lib -lBitVectors -lSockPipes   -lpthread
+rm *.o 

### 2023-05-29 v2/HierSysBuild/docs/tutorial/examples/shift_reg_4_rtl/tb/compile.sh

@@ -1,0 +1,3 @@
+gcc -g -c  -I ./ -I ../aa2c -I $AHIR_RELEASE/include tb.c
+gcc -g  tb.o -o tb_ghdl -L $AHIR_RELEASE/lib -L ../lib -lBitVectors -lSockPipes   -lpthread
+rm *.o 

### 2023-09-16 v2/Ahir-asm/devel/docs/AaLRM.tex
Small fix:    when  a guarded statement is present in AA      $guard (op) p := (q + r)    the latency of the operation (q+r) at the point    of the guard increases by 1.  This needs to    be accounted for in pipeline balancing.
   Further, the rate behaviour of this statement
   in a pipeline will be 1/2 capacity unless the
   statement is double buffered..

   This has been fixed by making the guarded (in this case, add)
   operator have a double buffered output wire.  The fix
   is applied in printing the vC.

@@
 by checking in the scope that contains the scope X, and so on.
-\item A scope can read from variables that are defined in descendant
-scopes.  
-\item A scope can export a variable defined within its scope to
-its parent scope.   The name  of the exported variable must not
-clash with the name of any other variable visible in the
-parent scope.
 \item A scope can read from variables that are defined in ancestor scopes.
@@
 
-A variable reference in a statement may be specified 
-as follows
-\begin{verbatim}
-a            look for variable a in current scope; 
-             if not found look in the parent.
-
-:a           same as the previous case
-
-../:a        look for variable a starting from the parent of
-             the current scope.
-
-%p:a         look for variable a starting from the child scope 
-             with label p (child scope of the current scope).
-
-../../:a     look for variable starting from the parent of
-             the parent of the current scope.
-
-%p%q:a       look for variable starting from the child scope 
-             with label q of the child scope with label p of 
-             the current scope
-
-a[10]        look for a storage variable a defined in the 
-             current scope.  If not found, look for it in
-             the parent scope, and so on.  If eventually found, 
-             access the corresponding element of the composite
-             object.
-\end{verbatim}
-Thus, a generic variable reference has the form 
-\begin{verbatim} 
-    scope-reference : variable-reference 
-\end{verbatim} 
-and the scoping rules forbid a scope from accessing variables
-which are not defined in either an ancestor or a descendant
-of the scope (the export mechanism should be used if this
-is needed).
+If, from within a scope, we wish to read the value of
+a variable defined in a child scope, then the child scope
+must export the value of the variable so that the parent
+scope can access it.
 
@@
 use the results of the \Aa compilation process.
+
+\section{Gated clocks}
+
+	Gated clocks can be declared and controlled within the Aa
+	language framework.   There are two control scenarios which
+	are supported.
+\begin{itemize}
+\item  Automatic mode:
+	   the gated clock can be attached to a module, and is
+	  enabled whenever the module is activated (by the
+	  start\_req/start\_ack pair), and disabled when the
+	  module is no longer active, i.e., it has no further
+	  jobs in progress inside it.
+\item Explicit mode:
+	   the gated clock enable signal is generated from within
+	   some logic described in Aa.
+\end{itemize}
+
+\subsection{Creating and using an explicitly controlled gated clock}
+
+	To create an explicit gated clock, we need to declare an enable signal
+	and a gated clock to which it is matched.  For example
+\begin{verbatim}
+// signal declaration of an enable.
+$pipe  data_1_clock_enable: $uint<1> $signal
+// gated clock attached to enable.
+$gated_clock  data_1_clock data_1_clock_enable
+\end{verbatim}
+
+	Now the declared gated clock can be applied to
+	a module foo by the statement
+\begin{verbatim}
+	$use_gated_clock foo  data_1_clock
+\end{verbatim}
+ 	It is the responsibility of the programmer
+	to control the enable signal (data\_1\_clock\_enable above).
+	This enable signal will control the gated clock data\_1\_clock
+	which is used by the module foo.
+
+\subsection{Declaring and using automatic gated clocks}
+
+	In this case, all we need to declare is that 
+	a module ``bar'' will run on a gated clock which
+	is enabled when the module is working on an active job
+	and disabled otherwise.  This is done by the
+	statement
+\begin{verbatim}
+$use_gated_clock bar
+\end{verbatim}
 

### 2023-10-15 v2/Ahir-asm/devel/src/AaStatement.cpp
Finished reimplementation of Guard interface    - bypass bit queue    - rewrite update side state-machine.
Confirmed on VHDL sims of AJIT processor.

@@ -2045,17 +2045,6 @@ void AaAssignmentStatement::Write_VC_Datapath_Instances(ostream& ofile)
 				{
 
 					int bufval = this->Get_Buffering();
-				
-					// corner case: guarded, unit-delay, full-rate.	
-					if((bufval < 2) && 
-						(this->Get_Guard_Expression() != NULL) &&
-						this->Is_Part_Of_Fullrate_Pipeline())
-					{
-						ofile << "// Buffering of fullrate unit-delay guarded simple-reference is set to 2." << endl;
-
-						bufval = 2;
-					}
-
 					if(bufval > 1)
 					{
 						ofile << "$buffering  $out " << dpe_name << " " << tgt_name << " " << bufval << endl;
@@ -6558,8 +6547,6 @@ void AaAssignmentStatement::Update_Adjacency_Map(map<AaRoot*, vector< pair<AaRoo
 
 
 	AaExpression* tgt_expression = this->Get_Target();
-	bool has_guard = (tgt_expression->Get_Guard_Expression() != NULL);
-
 	tgt_expression->Update_Adjacency_Map(adjacency_map,visited_elements);
 
 	if(this->_guard_expression != NULL)
@@ -6586,10 +6573,6 @@ void AaAssignmentStatement::Update_Adjacency_Map(map<AaRoot*, vector< pair<AaRoo
 	{
 		delay = src_expression->Get_Delay();
 	}
-
-	// increment delay if guard is not null.
-	if(has_guard)
-		delay++;
 
 	// arc from root to tgt_expression.
 	__InsMap(adjacency_map,this,tgt_expression,delay);
@@ -6637,8 +6620,7 @@ void AaCallStatement::Update_Adjacency_Map(map<AaRoot*, vector< pair<AaRoot*, in
 	{
 		AaExpression* tgt = _output_args[idx];
 		tgt->Update_Adjacency_Map(adjacency_map, visited_elements);
-		bool has_guard = (tgt->Get_Guard_Expression() != NULL);
-		__InsMap(adjacency_map,this,tgt,(has_guard ? delay : (delay + 1)));
+		__InsMap(adjacency_map,this,tgt,delay);
 	}
 
 

### 2023-12-26 v2/Ahir-asm/devel/docs/AaLRM.tex
Updated handling of records....   and checked out.

@@
 
+\subsection{The bit-level layout of a Record-type}
+
+Consider a record type:
+\begin{verbatim}
+$record [myrec] <$uint<8> >  < $uint<16> >
+\end{verbatim}
+which has an 8-bit field and a 16-bit field.  The
+record type when implemented in hardware will
+use 24 bits organized as follows
+\begin{verbatim}
+[23:8]                [7:0]
+second-field          first-field
+   16bit                8bit
+\end{verbatim}
+
+A cast from the record type to an integer type
+will be consistent with this bit layout.
+For example
+\begin{verbatim}
+// concatenate fields last-field to first-field.
+r := ($cast (myrec) ($concat u16 u8))
+
+//
+// The cast from the record gives the 
+// last field as the left-most.
+// w[23:8] will be field 2, 
+// w[7:0]  will be field 1.
+w := ($bitcast ($uint<24>) r) 
+\end{verbatim}
+
+Note that an assignment to an SSA variable
+with record type which writes only to
+a part of a record is not permitted.
+That is
+\begin{verbatim}
+r [0] := ....
+\end{verbatim}
+is not permitted if $r$ is an SSA variable
+with a record type.
+
 \section{Expressions} \label{sec:Expressions}
@@
 \item Zero, One expressions.
+\item Funcion call expressions.
 \end{itemize}
@@
 \end{verbatim}
+
+\subsection{Function call expressions}
+
+When a module is declared with only a single
+output argument, then it can be ``called'' in
+another way, using a function call expression.
+
+For example
+\begin{verbatim}
+$module [add8] $in (X Y: $uint<8>) 
+    $out(Z: $uint<8>)
+$is
+{
+  Y := (X + 1)
+}
+\end{verbatim}
+
+can be invoked in an expression as
+\begin{verbatim}
+P := ($call add8 (U V))
+\end{verbatim}
+for convenience.
 
@@
 
-\subsection{Opaque Modules}
+\subsection{Opaque Modules} \label{sec:Opaque}
 

### 2023-12-26 examples/regression_tests/record_types/1/src/testbench.c

@@ -1,0 +1,44 @@
+#include <signal.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <stdint.h>
+#include <pthread.h>
+#include <pthreadUtils.h>
+#include <Pipes.h>
+#include <pipeHandler.h>
+#ifndef SW
+#include "vhdlCStubs.h"
+#endif
+
+#define ORDER 16
+uint32_t expected_result[ORDER];
+
+void Exit(int sig)
+{
+	fprintf(stderr, "## Break! ##\n");
+	exit(0);
+}
+
+	
+
+int main(int argc, char* argv[])
+{
+	uint32_t result[ORDER];
+	signal(SIGINT,  Exit);
+  	signal(SIGTERM, Exit);
+
+	uint8_t I;
+	for(I = 0; I < 255; I++)
+	{
+		write_uint8("input_data", I);
+		write_uint8("input_data", I + 1);
+
+		uint8_t J = read_uint8("output_data");
+		if(J != (uint8_t) (I + I + 1))
+		{
+			fprintf(stderr,"Error: saw %d, expected %d\n", J, (uint8_t) (I + I + 1));
+		}
+	}
+	fprintf(stderr,"Done!\n");
+	return(0);
+}

### 2023-12-26 v2/Ahir-asm/devel/util/AaLinkExtMem.cpp

@@ -182,6 +182,7 @@ int main(int argc, char* argv[])
   string gsi_file_name = ".gsi.aa";
   ofstream gsi_file(gsi_file_name.c_str(),ios::out);
   AaProgram::Print_Global_Storage_Initializer(gsi_file);
+  AaProgram::Print_Record_Type_Helpers (gsi_file);
   gsi_file.close();
 
   // now parse the temporary file.

### 2023-12-26 v2/libAhirV2/src/vcRoot.cpp

@@ -50,7 +50,7 @@ string To_VHDL(string x)
   bool replace_underscore = false;;
   for (int i = 0; i < x.size(); i++)
     {
-      if(x[i] == '/' || x[i] == '$')
+      if(x[i] == '/' || x[i] == '$' || x[i] == '.')
 	{
 	  ret_string += 'X';
 	}

### 2025-01-10 v2/HierSysBuild/docs/tutorial/HierSysTools.tex
Updated Hier Sys Builder to allow the use of internally generated clocks (good luck!)

@@
 
-A 4-stage shift register constructed out of two simpler
+A 4-stage shift register is constructed out of two simpler
 2-stage shift registers.
@@
 
-\frame[containsverbatim]{\frametitle{Using RTL State Machine: example shift\_reg\_4\_rtl}
+\frame[containsverbatim]{\frametitle{Using RTL State Machines: example shift\_reg\_4\_rtl}
 \begin{itemize}
@@
 \end{itemize}
+}
+
+\frame[containsverbatim]{\frametitle{Clocks and resets}
+\begin{itemize}
+\item By default, each block in the system operates on 
+clock {\em clk} and reset {\em reset}.  The reset is
+active high and synchronized to {\em clk}.
+\item It is possible to define alternative clock and
+reset signals used in a system.
+\item It is possible to map one of these alternative clocks
+and resets to the default clock/reset on a subsystem instance.
+\end{itemize}
+}
+
+\frame[containsverbatim]{\frametitle{Declaring clock-like and reset-like signals}
+\begin{verbatim}
+$system Top $library TopLib
+   $in
+     $pipe A
+     $signal $clk EXTCLK 1 $depth 1
+     $signal $reset EXTRESET 1 $depth 1
+   $out
+     $pipe B
+{
+}
+\end{verbatim}
+Here, we are stating that the input signal EXTCLK is clock-like
+and the input signal RESET is reset-like.
+}
+
+\frame[containsverbatim]{\frametitle{Using clock-like and reset-like signals}
+\begin{verbatim}
+// ... as above ... 
+{
+ $pipe TMP $clk => EXTCLK $reset =>  EXTRESET
+ $instance i0  S1:Stage1
+         ..
+         $clk => EXTCLK
+         $reset => EXTRESET
+
+ $instance i1  S2:Stage2
+         ...
+         $clk => EXTCLK
+         $reset => EXTRESET
+
+\end{verbatim}
+}
+
+\frame[containsverbatim]{\frametitle{Using clock-like and reset-like signals}
+\begin{itemize}
+\item We have declared EXTCLK as clock-like and EXTRESET as reset-like.
+\item Pipe TMP and instance i0 will operate with EXTCLK connected
+to their default clock input and EXTRESET connected to their default
+reset input.
+\item We have generated a local clock EXTCLK\_REPEATED using an
+RTL machine.
+\item Instance i0 will operate with EXTCLK\_REPEATED connected
+to its default clock input and EXTRESET connected to its default
+reset input.
+\item Only a clock-like signal can be mapped to the default clock.
+\item Only a reset-like signal can be mapped to the default reset.
+\end{itemize}
+}
+
+\frame[containsverbatim]{\frametitle{Example: using clock-like and reset-like signals}
+See the example clocks\_and\_resets.
+
 }

### 2025-01-10 v2/HierSysBuild/src/hierSystem.cpp

@@ -375,7 +375,8 @@ void hierSystemInstance::Set_Default_Clock(string clk_id)
 	bool actual_is_input = (parent->Get_Input_Pipe_Width(clk_id) > 0);
 	bool actual_is_null  = (clk_id == "$null");
 
-	if(actual_is_null || (actual_is_clock && actual_is_input))
+	//if(actual_is_null || (actual_is_clock && actual_is_input))
+	if(actual_is_null || actual_is_clock)
 	{
 		this->_default_clock = clk_id;
 		if(!actual_is_null)
@@ -397,7 +398,8 @@ void hierSystemInstance::Set_Default_Reset(string reset_id)
 	bool actual_is_input = (parent->Get_Input_Pipe_Width(reset_id) > 0);
 	bool actual_is_null  = (reset_id == "$null");
 
-	if(actual_is_null || (actual_is_reset && actual_is_input))
+	//if(actual_is_null || (actual_is_reset && actual_is_input))
+	if(actual_is_null || actual_is_reset)
 	{
 		this->_default_reset = reset_id;
 		if(!actual_is_null)
